### 2013-08-27 writeup_and_notes/male_dependent_drive.tex

%%!TEX TS-program = latex
\documentclass[12pt,letterpaper]{article}
\usepackage{epsfig}                 
\usepackage[authoryear]{natbib}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{psfrag}
\usepackage{mathabx}
\renewcommand{\baselinestretch}{1.6}
\large
\pagenumbering{arabic}
\usepackage[usenames,dvipsnames]{color}
\usepackage{fullpage}
\bibliographystyle{genetics}
\usepackage{multirow}
\newcommand{\ye}{\hat{y}}
\newcommand{\xe}{\hat{x}}
\usepackage{color}
\usepackage[normalem]{ulem}

\title{Sperm modifiers of female drive.}
\author{Yaniv Brandvain \\ email: ybrandvain@gmail.com  \and Graham Coop \\ email: gmcoop@ucdavis.edu }

\usepackage{fancyhdr}
\pagestyle{fancy}
\lhead{}
\rhead{}
\renewcommand{\headrulewidth}{0.0pt}
\rfoot{}
\cfoot{\thepage}
\date{}
\begin{document}
\maketitle
\begin{center}
Center for Population Biology \& Section in Evolution and Ecology \\ University of California - Davis \\ Davis, CA, 95616
\end{center}

%FIgures 
{\bf Abstract:}
\newpage

\section*{Introduction}

Since all alleles within an individual rely on individual survival and reproduction for evolutionary success, most of life in a diploid eukaryotic genome is harmonious.
However, this harmony is incomplete, as in numerous cases an allele can benefit (in the short or long term) from taking advantage of its host individual.
A clear opportunity for this conflict arises during gametogenesis where alternative alleles are in direct competition for representation in the next generation.
Female meiosis is particularly unfair, only one of the four products of meiosis becomes an egg, while the other threes
are disgarded into polar bodies. 
Alleles that can segregate into the egg more than half the time in heterozygotes, by exploiting asymmetries, 
can potentially increase in frequency in the population (i.e. experiencing true meiotic drive).
If such alleles have fitness consequences they are a source of genetic conflict, 
and can become balanced in the population if their host in homozygotes 
out weighs their ability to spread to spread through heterozygotes. 
Such balanced drive systems can cause the subsequent evolution of linked enhancers of drive
and of supressors of drive throughout the genome. 
Indeed, the known polymorphic female meiotic drive systems (OM, maize knobs, IN, mimulus, others?) habor
a great diversity of enhancers and supressors.

Another unusual aspect of animal female meiosis is the fact that it is paused at XXX.
In fact in many species of animal the final stages of female meiosis are only completed upon fertlization of the egg by sperm.There's considerable variation in 
which stage of meiosis requires fertilization.
This raises the possibility that an allele in sperm could influence the outcome of meiosis in its favor. 
For example imagine an allele that when a sperm bearing this allele were to encounter an egg, whose meiotic product was currently heterozygote for 
another copy of itself, the sperm allele biased meiosis in favor of the other copy of itself.
At first sight, it seems as although female meiosis is primed to be exploited by selfish sperm systems.

Why then is the requirement of fertilization to complete female meiosis so unbiquitous? 
It is certainly not the case that mechanistically animals are incapable of evolving past this requirement.
A number of animal clades (should we try and lower bound how many transitions?) have evolved
to allow the completion of female meiosis upon ovulation. 
%(Although even in these organisms female meiosis 
%may also be subjection to influence by sperm alleles, if the sperm arrives at the egg soon enough after 
%ovulation to influence meiosis.)

This is not just idle speculation, as there is also direct evidence that sperm components and indeed different sperm genotypes can affect female meiosis.
%are mechanistic reasons to think that this is more than idle speculation.

In C elegans supressing XXX leads to premature deployment of the aster (a vitial component of mitotic mechinary) provided by the sperm, distrupting MII meiotic segregation in the egg, leading to a triploid zygote. 
This demonstrates that physical components derived from the sperm are are capable of affecting the outcome of female meiosis.
Furthermore, there is also more direct genetic evidence that the allele of a sperm can influence the outcome of female meiotic drive in their favor.
There are two female meiotic drive systems in mouse, In and Om, both of which are thought to operate
through drive at MII. In these systems the outcome of female meiosis seems to depend on 
the genotype of the fertilizing sperm.

In this article we explore through simple models the consequences of alleles that influence the outcome 
of female meiosis. 
We use these models to argue that it is actually surprisingly hard for the influence of
sperm on the outcome of meiosis to drive sustained conflict. In fact we find that sperm and egg 
genomes' interests are often aligned as they are both invested in the fate of the zygote they will form.
This suggests that females are unlikely to evolve to prevent the influence of sperm on meiosis,
and indeed feature of meiosis may evolve that facilitate the interaction of sperm with female meiosis.

\section*{Results}
Rough order of rest of text
\subsection*{Lay out Model} 
We imagine that drive acts in such a way that an allele promote itself into a fraction of d
 eggs ($d> 1/2$, when it is present in a heterozygote, forcing the other allele to be present in $1-d$ eggs. 
In mammals, fertilization takes place at MII, so we imagine this drive must be taking place at MII in order
for a sperm to have any influence. For drive to take place at MII there has to be an uneven number of
 crossovers between the centromere and the drive locus, such that realistically $d$ is bounded to be $<XX$. 
In other systems where fertilization occurs during MI, sperm could influence either drive at MI or MII, 
and drivers at MI can have a $d=1$ if they occur in tight linkage with the centromere.

Cartoon figure of sperms meeting eggs with different genotypes?

\begin{figure}
	\rotatebox{270}{\includegraphics[width = 0.8
          \textwidth]{Figures/sperm_egg_cartoon1.ps}}
\caption{transmission probabilities for alleles through female
  meiosis depend on sperm genotype. 2 allele models}  
\label{Eggsperm_2_allele_cartoon}
\end{figure}

Cost of alleles

We'll setup a model where the sperm allele affects outcome of female meiosis. This may not be biologically
realistic as there is very little sperm expression. More realistically the male genotype that produced the sperm
may be more relevant, as males could place products in their sperm that influenced the outcome of female meoisis.
However, in practice models that allow the influence of fertilizing male genotype on female meiosis 
seem to behave very similiarly to those based on the allele of the fertilizing sperm.
For simplicity we'll focus on models based on the allele of the sperm, as it's bloodly complicated enough already.

\subsection*{ A) Invasion by single self promoting allele}
show that such alleles:
\begin{enumerate}
\item can't be balanced, \\
\item and homozygous problems are tested out at low freq.  \\
\item Any heterozygous problems, leads to a bistable allele\\
\item If these alleles take off they speed through to fixation\\
\item If allele has any drive ability in absence of sperm effect that is what allows it to enter the population
and sperm effect isn't a further cause of conflict. What if anything do we mean by this?\\
\end{enumerate}

Conclusion, such alleles are unlikely to cause evolution of female supressors, they test hemselves in a homozygous
state when they enter the population, and sweep quickly (all the way to fixation) if they enter the pop at all.\\

\begin{figure}
	\rotatebox{270}{\includegraphics[width = 0.8
          \textwidth]{Figures/sperm_egg_cartoon2.ps}}
\caption{transmission probabilities righthand allele through female
  meiosis}  \label{Eggsperm_3_allele_cartoon}
\end{figure}

\subsection*{B) A more biologically realistic selfish sperm system}
Three allele systems. See Figure \ref{Eggsperm_3_allele_cartoon} \\ 
Perhaps these self promoting alleles are biologically unrealistic as new alleles needs 
both ability to drive in female meiosis and to influence that drive in sperm.
Perhaps more realistic to think that a female drive system evolves, 
and then a sperm modifier appears. This will only occur if original system is trapped 
at drive-selection balance. So driver must have homozygous cost.

\begin{enumerate}
\item Setup a drive-selection balanced polymorphism in std. drive model. 
Do this by imagining the sperm-influence allele arising on the background of the driver, 
so the allele has drive capabilities, and can have had time to evolve new biology. 
Evolving on the new background means that the allele suffers the fitness consequences of the 
driver.  See A in Figure \ref{Eggsperm_3_allele_cartoon} \\
\item Sperm-based enhancers of drive can't invade (can they in some situations?). \\
\item Intuition is that the driver has already driven to a frequency 
where it is held in check by its cost in homozygotes. The sperm allele 
thus can't really help as it creates zygotes which suffer the homozygous fitness consequences.
\end{enumerate}

\subsection{So what can evolve?}
So what can happen?
\begin{enumerate}
\item Alleles that arise in linkage with drive systems, which when in sperm switch off drive, 
can spread. They benefit from drive, but avoid some of the
consequences ( (Bi) in Figure \ref{Eggsperm_3_allele_cartoon} ). Overall as a side product they are benefiting all in pop.\\
\item Presumably alleles that actually switch the allele that drives may do even better? As they'd end up in 
hets. Although they'd not drive, so hard to say.\\
\item Alleles that cause sperm to switch off drive that arise on other background or unlinked to the drive system
are selected, and spread as fast as female supressors of drive ( (Bii) in Figure \ref{Eggsperm_3_allele_cartoon} )..\\
\item Alleles that in females facilitate the action of sperm supressors of drive (or vis versa) can spread. Haven't actually checked this.\\
\end{enumerate}

\section*{Conclusions.}
Discussion of general conclusion that females have little reason to evolve supression mechanisms to prevent sperm influence on meiosis. 
General logic that sperm genome has to live in a zygote with consequences of its effect on female meiosis, so
it can not generate too dire a consequence.

This logic may not hold for sex chromosomes. In ZW systems 
 Male modification of recombination rates
POssibility that this could happen in plants if pollen emit signals to ``egg''
Discussion of OM and IN.

\end{document}

### 2013-08-27 writeup_and_notes/old_Yaniv_male_dep_drive.tex

%%!TEX TS-program = latex
\documentclass[12pt,letterpaper]{article}
\usepackage{epsfig}                 
\usepackage[authoryear]{natbib}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{psfrag}
\usepackage{mathabx}
\renewcommand{\baselinestretch}{1.6}
\large
\pagenumbering{arabic}
\usepackage[usenames,dvipsnames]{color}
\usepackage{fullpage}
\bibliographystyle{genetics}
\usepackage{multirow}
\newcommand{\ye}{\hat{y}}
\newcommand{\xe}{\hat{x}}
\usepackage{color}
\usepackage[normalem]{ulem}  
\title{Sperm modifiers of female drive.}
\usepackage{fancyhdr}
\pagestyle{fancy}
\lhead{}
\rhead{}
\renewcommand{\headrulewidth}{0.0pt}
\rfoot{}
\cfoot{\thepage}
%\author{Yaniv Brandvain \\ email: ybrandvain@gmail.com  \and Graham Coop \\ email: gmcoop@ucdavis.edu }
\date{}
\begin{document}
\maketitle
\begin{center}
Center for Population Biology \& Section in Evolution and Ecology \\ University of California - Davis \\ Davis, CA, 95616
\end{center}

%FIgures 
% 2 cartoon5.eps
% 3 Unlinked.eps
% 4 Linked.eps
% 5 3locus.eps
\newpage

{\bf Abstract:}
Gametic drive, the selfish distortion of gametogenesis by an allele to increase its representation in the next generation is potentially common evolutionary force, which has received much attention. 
What has received less attention is the possibility that an allele in one individual can modify drive in his mate, the subject of this note.
Critically, the biological opportunity for this model occurs because in many organisms, the final division of female meiosis does not occur until after fertilization.
%We show that a green bearded driver which enhances female drive examine both the case of `green beard' drive, where an allele in sperm enhances drive during female meiosis II, and `red beard' drive drive, where an allele in sperm hinders drive of that same allele in females.
%We show that green bearded drivers are difficult to maintain at equilibrium because positive frequency dependent selection on these alleles overwhelms the negative frequency dependance of recessive fitness cost.
%By contrast, female drivers which act to suppress their own drive in females can spread when rare and be maintained at a polymorphic equilibrium. 
These unexpected results highlight the diverse consequences of intragenomic conflict.
\newpage

\section*{Introduction}
Since all alleles within an individual rely on individual survival and reproduction for evolutionary success, most of life in a diploid eukaryotic genome is harmonious.
However, this harmony is incomplete, as in numerous cases an allele can benefit (in the short or long term) from taking advantage of its host individual.
A clear opportunity for this conflict arises during gametogenesis where alternative alleles are in direct competition for representation in the next generation.
The evolution of gametic drive in both spermatogenesis and oogenesis, and the consequences of drive have been well explored, and convincing cases of meiotic drive have been documented.
Additionally, because these selfish drivers potentially increase in frequency at the cost of individual fitness numerous models have explored mechanisms by which unlinked loci act to suppress the effect of drive.

What has been less appreciated is the opportunity for loci in sperm to modify the effect of a driving allele, perhaps because this possibility seems somewhat far-fetched.
However, since in many species, the final division female female meiotic division does not occur until after fertilization, there is an opportunity for an allele in sperm to influence the outcome of meiotic drive.
This opportunity appears to have been realized in mice where drive at both the XXX and YYYY loci depend on sperm genotype.

Here, we present a first theoretical treatment of this problem, examining the evolutionary dynamics of loci  in sperm that act to enhance or suppress female meiotic drive.
We begin with an allele that drives in female meiosis and has reached an equilibrium between drive and selection.
We then explore the dynamics of an allele that acts in sperm to enhance or suppress drive in female meiosis II, and investigate how this locus evolves when linked or unlinked to the female drive locus.

\newpage
\section*{The Model}
In Table 1, we present a standard model of female drive. 
Here, the female drive locus is balletic with D and d alleles at frequency $f_D$ and $f_d = 1 -f_D$, respectively.
The D  allele biases the $2^{nd}$ meiotic division in its favor by $\alpha$ when there has been a recombination event between itself and the centromere, which occurs with probability $r$. 
This dependence on recombination follows from the biology of female meiotic drive -- for an allele to drive during meiosis II, diads must be heteromorphic such that there is an opportunity for selection, and we focus on the on an MII driver because it this form of female drive which could be plausibly influenced by sperm genotype.
Thus, for example, when a driver is complete (i.e. $\alpha = 1$) the D allele is transmitted to all eggs of drive heterozygotes in which there was a recombination event between driver and centromere.

After the female drive locus reaches an drive-viability equilibrium, (which equals XXX in the case of a recessive lethal driver), we introduce a third allele at the drive locus.
This allele acts to modify female drive such that the efficacy of drive depends on sperm genotype. 
 equilibrium frequency deter  Under this model, the equilibrium frequency of the female driver locus equals MATH, where the fitness of ecotypes dd, Dd and DD equals $1$, $1-s_1$ and $1-s_2$, respectively.

In Table 1B we present a model where female drive is modified by a sperm's haplotype. 
In this model, the efficacy of drive is linearly enhanced by $\delta$ when fertilized by an M bearing sperm. 
We ask how the m locus evolves as a function of $\delta$ and the linkage relationship between male modifier (m) and female drive (d) loci, and how this influences the frequency of the female drive allele.

Throughout our analyses we begin at an equilibrium frequency of the driving allele and introduce a rare modifier.
We assume random mating among genotypes; however, the potential recessive cost of female drive and the influence of sperm on the function of drive prevent us from assuming linkage equilibrium or Hardy-Weinberg frequencies.

\begin{table}[ht]
\begin{center}
\begin{tabular}{ccc}
  \hline
1A  & Egg \\   \hline
mat 	& 		d 			& 		D 	\\  \hline
dd		& 		1		&		0	\\
Dd		& $(1-r \alpha)/2$	&	 $(1+r \alpha)/2$ 	\\
DD		& 		0			&		1	\\
   \hline
\end{tabular}
\end{center}
\end{table}

\begin{table}[ht]
\begin{center}
\begin{tabular}{cccc}
  \hline
1B  		& 			& Egg \\   \hline
mat 	& Sperm	&		d 					& 		D 	\\  \hline
dd		& M	 or m	&	1						&		0	\\
Dd		& m			&$(1-r \alpha)/2$			&	 $(1+r \alpha)/2$ 	\\
Dd		& M			&$(1-r (\alpha +\delta))/2$	&	 $(1+r( \alpha+\delta))/2$ 	\\
DD		& M	 or m	&0							&		1	\\
   \hline
\end{tabular}
\end{center}
\end{table}

\subsection*{Complete linkage, gametic phase }

\subsection*{Complete linkage, repulsion phase }

\subsection*{Unlinked loci}

\end{document}

\label{predictions}
\begin {tabular} { | p{1.5cm} | p{1.5cm} |  p{2.75cm} | p{2.75cm} | p{2.75cm} |p{2.75cm}  |}
\hline
&&  Offspring  \\  \hline
Mom & Dad &DD & Dd &dD &  dd  \\  \hline
DD & DD & 1 & 0  & 0 & 0 \\  \hline
DD & Dd & 1/2 & 1/2 &0&0\\  \hline
DD & dd & 0 & 1&0&0 \\  \hline
Dd & DD & $\frac{1 - r \alpha_D}{2}$ & 0&$\frac{1 + r \alpha_D}{2}$  &  0 \\  \hline
Dd & Dd & $\frac{1 - r \alpha_D}{4}$ & $\frac{1 - r \alpha_d}{4}$ &$\frac{1 + r \alpha_D}{4}$ & $\frac{1 + r \alpha_d}{4}$ \\  \hline
Dd & dd & 0 & $\frac{1 - r \alpha_d}{2}$&0&$\frac{1+ r \alpha_d}{2}$ \\  \hline
dd & DD & 0 & 0 & 1 & 0 \\  \hline
dd & Dd & 0 & 0 &1/2&1/2\\  \hline
dd & dd & 0 & 0 &0&1 \\  \hline
\end {tabular}
\caption{Relative male ($r_\Mars$) to female ($r_\Venus$) recombination rates as predicted by the meiotic drive theory.} 
\label{}
\end{table}

We then find the frequency of the $ij^{th}$ genotype in the next generation, $f_{ij}'$, by taking the weighted sum of the frequency of $ij$ genotypes produced in the $k^{th}$ family

\begin{equation}
	f_{ij}' =  \sum \limits_{k} X_i \phi_{ij|k}
\end{equation}
Where $X_{k}$ is the frequency of the $k^{th}$ family, and $\phi_{ij|k}$ is the frequency of genotype $ij$ in offspring of the $k^{th}$ family before selection:
\\
\begin{equation}
  f_{ij}'=\begin{cases}
    f_{i\Venus} f_{j\Mars} - f_{Dd\Venus}f_{j\Mars}r\alpha_j/2& \text{  \hspace{1cm} if $j=D$}.\\
    f_{i\Venus} f_{j\Mars} + f_{Dd\Venus}f_{j\Mars}r\alpha_j/2& \text{  \hspace{1cm} if $j=d$}.\\
  \end{cases}
  \label{hapgenofreq}
\end{equation}
Where $i$ and $j$ are maternally and paternally derived alleles, respectively, and are calculated after selection in the previous generation. 
Note that in equation, \ref{hapgenofreq}, Dd and dD heterozygotes are calculated separately, so these must be summed to obtain the frequency of heterozygotes.

\end{document}

### 2013-09-02 writeup_and_notes/male_dependent_drive.tex
YB modified intro and  added comments

@@
 \date{}
+\bibliographystyle{plain}
 \begin{document}
@@
 
-Since all alleles within an individual rely on individual survival and reproduction for evolutionary success, most of life in a diploid eukaryotic genome is harmonious.
-However, this harmony is incomplete, as in numerous cases an allele can benefit (in the short or long term) from taking advantage of its host individual.
-A clear opportunity for this conflict arises during gametogenesis where alternative alleles are in direct competition for representation in the next generation.
-Female meiosis is particularly unfair, only one of the four products of meiosis becomes an egg, while the other threes
-are disgarded into polar bodies. 
-Alleles that can segregate into the egg more than half the time in heterozygotes, by exploiting asymmetries, 
-can potentially increase in frequency in the population (i.e. experiencing true meiotic drive).
-If such alleles have fitness consequences they are a source of genetic conflict, 
-and can become balanced in the population if their host in homozygotes 
-out weighs their ability to spread to spread through heterozygotes. 
-Such balanced drive systems can cause the subsequent evolution of linked enhancers of drive
-and of supressors of drive throughout the genome. 
-Indeed, the known polymorphic female meiotic drive systems (OM, maize knobs, IN, mimulus, others?) habor
-a great diversity of enhancers and supressors.
+%Since all alleles within an individual rely on individual survival and reproduction for evolutionary success, most of life in a diploid eukaryotic genome is harmonious.
+%However, this harmony is incomplete, as in numerous cases an allele can benefit (in the short or long term) from taking advantage of its host.
+Despite the apparent unity of the organism, occasionally alleles can gain an evolutionary advantage at a cost to  individual fitness \citep{Burt2006}.
+%A clear opportunity for this conflict arises during gametogenesis where alternative alleles are in direct competition for representation in the next generation.
+Female meiosis, an event in which only one of two alternate alleles enters the  egg, is one such occasion \citep{Sandler1957,Pardo-ManuelDeVillena2001a}. 
+%Female meiosis is particularly unfair, only one of the four products of meiosis becomes an egg, while the other threes are disgarded into polar bodies. 
+An allele that biases female meiosis in its favor (i.e.
+a meiotic driver), may increase in frequency even if this driver entails a pleiotropic fitness cost \citep{Prout1973}, generating a genetic conflict between the success of the driver and organismal fitness.
+%Alleles that can segregate into the egg more than half the time in heterozygotes, by exploiting asymmetries, 
+%can potentially increase in frequency in the population (i.e. experiencing true meiotic drive).
+%If such alleles have fitness consequences they are a source of genetic conflict, 
+%and can become balanced in the population if their host in homozygotes 
+%out weighs their ability to spread to spread through heterozygotes. 
+Meiotic drivers observed in nature (in both plants \citep{Buckler1999,Fishman2005,Fishman2008}, and animals \citep{Agulnik1990,Wu2005,Pardo-ManuelDeVillena2001c}) highlight this conflict -- the selfish benefits of drive and the associated pleiotropic fitness costs sustain a balanced polymorphism, and often generate on ongoing evolutionary escalation of drive suppressors and enhancers \citep{Dawe1996,Fishman2008}.
+%Such balanced drive systems can cause the subsequent evolution of linked enhancers of drive
+%and of supressors of drive throughout the genome. 
+%Indeed, the known polymorphic female meiotic drive systems (OM, maize knobs, IN, mimulus, others?) habor
+%a great diversity of enhancers and supressors. 
+The threat of meiotic drive to organismal fitness is potentially so severe that that many basic properties of meiosis and o\"{o}genesis, including the initial genome doubling in meiosis I \citep{Haig1991}, the commonness of arrested female meiosis \citep{Mira1998}, and sex differences in the recombination rate \citep{Haig2010,Brandvain2012} have putatively evolved to discourage the evolution of meiotic drive an enforce fairness.
 
-Another unusual aspect of animal female meiosis is the fact that it is paused at XXX.
-In fact in many species of animal the final stages of female meiosis are only completed upon fertlization of the egg by sperm.There's considerable variation in 
-which stage of meiosis requires fertilization.
-This raises the possibility that an allele in sperm could influence the outcome of meiosis in its favor. 
-For example imagine an allele that when a sperm bearing this allele were to encounter an egg, whose meiotic product was currently heterozygote for 
-another copy of itself, the sperm allele biased meiosis in favor of the other copy of itself.
+It is therefore somewhat surprising that despite the intense evolutionary pressure on female meiosis to prevent meiotic drive, 
+it is potentially open to sabotage by a virtual stranger -- a haploid sperm genome.
+That is, in many animal species, the completion of female meiosis requires fertilization of the egg, 
+and there is ample opportunity for interaction between the sperm and female meiotic machinery.
+If, for example, an allele in sperm could facilitate meiotic drive by a genetically equivalent allele in a 
+heteromorphic dyad, such an allele could presumably bias meiosis in its favor and rapidly spread through the population.
 At first sight, it seems as although female meiosis is primed to be exploited by selfish sperm systems.
 
-Why then is the requirement of fertilization to complete female meiosis so unbiquitous? 
-It is certainly not the case that mechanistically animals are incapable of evolving past this requirement.
-A number of animal clades (should we try and lower bound how many transitions?) have evolved
+%Another unusual aspect of animal female meiosis is the fact that it is paused at XXX.
+%In fact in many species of animal the final stages of female meiosis are only completed upon fertlization of the egg by sperm.There's considerable variation in 
+%which stage of meiosis requires fertilization.
+%This raises the possibility that an allele in sperm could influence the outcome of meiosis in its favor. 
+%For example imagine an allele that when a sperm bearing this allele were to encounter an egg, whose meiotic product was currently heterozygote for 
+%another copy of itself, the sperm allele biased meiosis in favor of the other copy of itself.
+%At first sight, it seems as although female meiosis is primed to be exploited by selfish sperm systems.
+
+Why then is the requirement of fertilization to complete female meiosis so ubiquitous? 
+It is certainly not the case that animals are mechanistically incapable of evolving past this requirement.
+There is considerable variation in which stage of meiosis requires fertilization, and 
+a number of animal clades (should we try and lower bound how many transitions?) have evolved
 to allow the completion of female meiosis upon ovulation. 
@@
 
-This is not just idle speculation, as there is also direct evidence that sperm components and indeed different sperm genotypes can affect female meiosis.
+It is also not the case that sperm is mechanistically incapable of influencing the outcome of female meiosis.
+%This is not just idle speculation, as there is also direct evidence that sperm components and indeed different sperm genotypes can affect female meiosis.
 %are mechanistic reasons to think that this is more than idle speculation.
-
-In C elegans supressing XXX leads to premature deployment of the aster (a vitial component of mitotic mechinary) provided by the sperm, distrupting MII meiotic segregation in the egg, leading to a triploid zygote.
-This demonstrates that physical components derived from the sperm are are capable of affecting the outcome of female meiosis.
-Furthermore, there is also more direct genetic evidence that the allele of a sperm can influence the outcome of female meiotic drive in their favor.
-There are two female meiotic drive systems in mouse, In and Om, both of which are thought to operate
-through drive at MII. In these systems the outcome of female meiosis seems to depend on 
-the genotype of the fertilizing sperm.
+Mechanistic evidence for this possibility comes from \emph{C.
+elegans}, where experimentally suppressing XXX leads to premature deployment of the aster (a vital component of mitotic machinery) provided by the sperm, disrupting MII meiotic segregation in the egg, leading to a triploid zygote.
+%This demonstrates that physical components derived from the sperm are are capable of affecting the outcome of female meiosis.
+Additionally, genetic evidence suggests that the transmission patterns in heterozygous females can depend on sperm haplotype. 
+Specifically, the two best characterized female meiotic drive systems in mouse (In and Om), both operate by distorting the second meiotic division, 
+and in both systems the outcome of female meiosis depends the genotype of the fertilizing sperm \citep{Agulnik1993,Wu2005}.
 
@@
 sperm on the outcome of meiosis to drive sustained conflict. In fact we find that sperm and egg 
-genomes' interests are often aligned as they are both invested in the fate of the zygote they will form.
+genomes' interests are often aligned as they are both invested in the fate of the zygote they will form (as was suggested for the In locus \citep{Pomiankowski1993}).
 This suggests that females are unlikely to evolve to prevent the influence of sperm on meiosis,
@@
 \section*{Results}
+
+YB: I think we should do our best to be very short and sweet here. There are many opportunities to move off to cool math and cool biological scenarios. 
+We should fight all of them and focus on mission critical results.
+All others can go in supp or appease reviewers.
+Lets talk about why the critical results are.
+To me its (1) male enhancement of female drive cannot maintain a stable polymorphisms, and (2) An allele in sperm can evolve to suppress its drive in females.
+
 Rough order of rest of text
@@
 and sperm effect isn't a further cause of conflict. What if anything do we mean by this?\\
+\item PERHAPS HERE WE INTRODUCE A SELF-RESTRAINING ALLELE
 \end{enumerate}
@@
 \item Presumably alleles that actually switch the allele that drives may do even better? As they'd end up in 
-hets. Although they'd not drive, so hard to say.\\
+hets. Although they'd not drive, so hard to say. YB: There is evidence that In distorts meiosis in the other direction (they still drive when rare i.e. when not fused with drive supp sperm) \\
 \item Alleles that cause sperm to switch off drive that arise on other background or unlinked to the drive system
@@
 it can not generate too dire a consequence.
+I THINK THERE IS A MORE SUBTLE POINT . The sperm has special knowledge that if it allows drive it will end up in the low fitness homozygote.
 
@@
 
+\bibliography{refs}
 \end{document}

### 2013-09-02 scripts/invasion_analysis.R
added acknowledgement

@@ -1,6 +1,6 @@
 
 
-test.invasion<-function(d,s2,s3,x=.001){
+test.invasion<-function(d,s2,s3,x=.001,steps=10){
 	D["1 or 2","12"]<-0.5
 	D["1 or 2","13"]<-0.5
 	D["3","12"]<-.5
@@ -9,11 +9,14 @@ test.invasion<-function(d,s2,s3,x=.001){
 	D["3","23"]<-0.5
 	female.transmission.probs<-make.sperm.dep.female.transmission.prob(D)
 	par(mar=c(4,4,1,1))
-	old.geno.freqs<-iterate.1.locus.drive(s2=s2,s3=s3,num.iterations=10,female.transmission.probs=female.transmission.probs,initialize.allele.freqs =c(1-x, 0,x))
+	old.geno.freqs<-iterate.1.locus.drive(s2=s2,s3=s3,num.iterations=steps,female.transmission.probs=female.transmission.probs,initialize.allele.freqs =c(1-x, 0,x))
 
-	invading<-old.geno.freqs[["allele.freqs"]][10,3]>old.geno.freqs[["allele.freqs"]][1,3]
+	invading<-old.geno.freqs[["allele.freqs"]][steps,3]>old.geno.freqs[["allele.freqs"]][1,3]
 	return(invading)
 }
+
+
+###invasion of self-driver against homozy. cost.
 
 d.range<-seq(0.5,1,length=100)
 s3.range<-seq(0,1,length=200)
@@ -26,7 +29,7 @@ invasion.grid<-sapply(s3.range,function(s3){
 
 invasion.lines.s3.cutoff<-s3.range[apply(invasion.grid,1,function(x){max(which(x))})]
 
-image(d.range, s3.range,invasion.grid)
+image(d.range, s3.range,invasion.grid,xlab="Drive coeff. D",ylab="Homozygous cost, s")
 theory.s3.cutoff<-d.range-0.5
 lines(d.range,theory.s3.cutoff)
 
@@ -38,4 +41,54 @@ invasion.grid.HWE<-sapply(s3.range,function(s3){
 })
 
 invasion.lines.s3.HWE.cutoff<-s3.range[apply(invasion.grid.HWE,1,function(x){max(which(x))})]
+lines(d.range,invasion.lines.s3.HWE.cutoff)
+
+
+drive.coeffs<-c(0.6,.7,.8)
+selection.coeffs<-rep(0.1,3)
[…]

### 2013-09-02 scripts/iterate_drive_model.R

@@ -16,7 +16,7 @@ iterate.1.locus.drive<-function(s2,s3,num.iterations,female.transmission.probs,m
 		geno.freqs["12"]<-2*allele.freqs["1"]*allele.freqs["2"]
 		geno.freqs["13"]<-2*allele.freqs["1"]*allele.freqs["3"]
 		geno.freqs["23"]<-2*allele.freqs["3"]*allele.freqs["2"]
-		stopifnot(sum(geno.freqs)==1)
+		stopifnot(sum(geno.freqs)>0.9999999999 & sum(geno.freqs)< 1.0000000001)
 
 	}else{
 		print("initiating from my.geno.freqs")
@@ -328,13 +328,14 @@ examples.for.talk <-function(){
 		 D["1 or 2","12"]<-0.5
 	 D["1 or 2","13"]<-0.5
 	 D["3","12"]<-.5
-	 D["3","13"]<-1
+	 D["3","13"]<-0.6
 	 D["1 or 2","23"]<- 1-0.5
 	 D["3","23"]<-0.5
 		 female.transmission.probs<-make.male.geno.dep.female.transmission.prob(D)
 	par(mar=c(4,4,1,1))
-	old.geno.freqs<-iterate.1.locus.drive(s2=0,s3=1,num.iterations=10000,female.transmission.probs=female.transmission.probs,initialize.allele.freqs =c(0.99, 0,0.01 ))
+	old.geno.freqs<-iterate.1.locus.drive(s2=0,s3=.1,num.iterations=3000,female.transmission.probs=female.transmission.probs,initialize.allele.freqs =c(0.99, 0,0.01 ))
 	plot.freqs(old.geno.freqs)
+
 	
 }
 

### 2013-09-08 scripts/invasion_analysis.R
updated iterations to allow het. fitness, included fix to bistability code, new invasion figs

@@ -1,6 +1,12 @@
+library("RColorBrewer")
 
-
-test.invasion<-function(d,s2,s3,x=.001,steps=10){
+test.invasion.self.prop<-function(d,s.het,s.hom,x=.001,steps=10){
+	s.array<-rep(0,6)
+	names(s.array)<-c("11","12","13","23","22","33")
+	s.array["13"]<- s.het
+	s.array["33"]<- s.hom
+	
+	D<-matrix(1/2,nrow=2,ncol=3,dimnames=list(c("1 or 2","3"),c("12","13","23"))) ##entries are drive coeff of the 2nd 
 	D["1 or 2","12"]<-0.5
 	D["1 or 2","13"]<-0.5
 	D["3","12"]<-.5
@@ -8,22 +14,42 @@ test.invasion<-function(d,s2,s3,x=.001,steps=10){
 	D["1 or 2","23"]<- 1-0.5
 	D["3","23"]<-0.5
 	female.transmission.probs<-make.sperm.dep.female.transmission.prob(D)
-	par(mar=c(4,4,1,1))
-	old.geno.freqs<-iterate.1.locus.drive(s2=s2,s3=s3,num.iterations=steps,female.transmission.probs=female.transmission.probs,initialize.allele.freqs =c(1-x, 0,x))
+	old.geno.freqs<-iterate.1.locus.drive(s.array,num.iterations=steps,female.transmission.probs=female.transmission.probs,initialize.allele.freqs =c(1-x, 0,x))
 
 	invading<-old.geno.freqs[["allele.freqs"]][steps,3]>old.geno.freqs[["allele.freqs"]][1,3]
 	return(invading)
 }
 
+test.fixation.of.simple.driver<-function(d,s.het,s.hom,x=.999,steps=10){
+	s.array<-rep(0,6)
+	names(s.array)<-c("11","12","13","23","22","33")
+	s.array["12"]<- s.het
+	s.array["22"]<- s.hom
+	
+	D<-matrix(1/2,nrow=2,ncol=3,dimnames=list(c("1 or 2","3"),c("12","13","23"))) ##entries are drive coeff of the 2nd 
+	D["1 or 2","12"]<- d
+	D["1 or 2","13"]<-0.5
+	D["3","12"]<- 0.5
+	D["3","13"]<- 0.5
+	D["1 or 2","23"]<- 1-0.5
+	D["3","23"]<-0.5
+	female.transmission.probs<-make.sperm.dep.female.transmission.prob(D)
[…]

### 2013-09-08 scripts/iterate_drive_model.R

@@ -1,4 +1,4 @@
-iterate.1.locus.drive<-function(s2,s3,num.iterations,female.transmission.probs,my.geno.freqs,initialize.allele.freqs=c(0.999,0.001,0)){
+iterate.1.locus.drive<-function(s.array,num.iterations,female.transmission.probs,my.geno.freqs,initialize.allele.freqs=c(0.999,0.001,0)){
 	geno.freqs<-rep(NA,6)
 	names(geno.freqs)<-geno.names
 
@@ -28,9 +28,15 @@ iterate.1.locus.drive<-function(s2,s3,num.iterations,female.transmission.probs,m
 
 	for(i in 1:num.iterations){
 	#selection
-		geno.freqs["22"]<-geno.freqs["22"] * (1-s2)
-		geno.freqs[c("33","23")]<-geno.freqs[c("33","23")] * (1-s3)
+		geno.freqs["11"]<-geno.freqs["11"] * (1-s.array["11"])
+		geno.freqs["12"]<-geno.freqs["12"] * (1-s.array["12"])
+		geno.freqs["13"]<-geno.freqs["13"] * (1-s.array["13"])
 
+		geno.freqs["22"]<-geno.freqs["22"] * (1-s.array["22"])
+		geno.freqs["23"]<-geno.freqs["23"] * (1-s.array["23"])
+
+		geno.freqs["33"]<-geno.freqs["33"] * (1-s.array["33"])
+	
 		geno.freqs<-geno.freqs/sum(geno.freqs)	
 
 		geno.freqs.array<-outer(rep(geno.freqs,each=2),rep(geno.freqs,each=2),"*")

### 2013-09-10 writeup_and_notes/male_dependent_drive.tex
setting invasion figure for paper

@@
 \begin{center}
-Center for Population Biology \& Section in Evolution and Ecology \\ University of California - Davis \\ Davis, CA, 95616
+Center for Population Biology \& Department of Evolution and Ecology \\ University of California - Davis \\ Davis, CA, 95616
 \end{center}
@@
 %However, this harmony is incomplete, as in numerous cases an allele can benefit (in the short or long term) from taking advantage of its host.
-Despite the apparent unity of the organism, occasionally alleles can gain an evolutionary advantage at a cost to  individual fitness \citep{Burt2006}.
+Despite the apparent unity of the organism, occasionally alleles can
+gain an evolutionary advantage at a cost to  individual fitness
+\citep{Burt2006}, often by exploiting meiosis and gametogensis.
 %A clear opportunity for this conflict arises during gametogenesis where alternative alleles are in direct competition for representation in the next generation.
-Female meiosis, an event in which only one of two alternate alleles enters the  egg, is one such occasion \citep{Sandler1957,Pardo-ManuelDeVillena2001a}. 
+Female meiosis, an asymmetric event in which only one of two alternate alleles
+enters the  egg, with the others consigned to the polar body, is one such occasion \citep{Sandler1957,Pardo-ManuelDeVillena2001a}. 
 %Female meiosis is particularly unfair, only one of the four products of meiosis becomes an egg, while the other threes are disgarded into polar bodies. 
@@
 %out weighs their ability to spread to spread through heterozygotes. 
-Meiotic drivers observed in nature (in both plants \citep{Buckler1999,Fishman2005,Fishman2008}, and animals \citep{Agulnik1990,Wu2005,Pardo-ManuelDeVillena2001c}) highlight this conflict -- the selfish benefits of drive and the associated pleiotropic fitness costs sustain a balanced polymorphism, and often generate on ongoing evolutionary escalation of drive suppressors and enhancers \citep{Dawe1996,Fishman2008}.
+Meiotic drivers observed in nature (in both plants
+\citep{Buckler1999,Fishman2005,Fishman2008}, and animals
+\citep{Agulnik1990,Wu2005,Pardo-ManuelDeVillena2001c}) highlight this
+conflict -- the selfish benefits of drive and the associated
+pleiotropic fitness costs sustain a balanced polymorphism
+\citep{Prout1973}, 
+and often generate on ongoing evolutionary escalation of drive suppressors and enhancers \citep{Dawe1996,Fishman2008}. 
 %Such balanced drive systems can cause the subsequent evolution of linked enhancers of drive
@@
 %a great diversity of enhancers and supressors. 
-The threat of meiotic drive to organismal fitness is potentially so severe that that many basic properties of meiosis and o\"{o}genesis, including the initial genome doubling in meiosis I \citep{Haig1991}, the commonness of arrested female meiosis \citep{Mira1998}, and sex differences in the recombination rate \citep{Haig2010,Brandvain2012} have putatively evolved to discourage the evolution of meiotic drive an enforce fairness.
+The threat of meiotic drive to organismal fitness is potentially so
+severe that it has been hypothesized that many basic properties of meiosis and o\"{o}genesis, including the initial genome doubling in meiosis I \citep{Haig1991}, the commonness of arrested female meiosis \citep{Mira1998}, and sex differences in the recombination rate \citep{Haig2010,Brandvain2012} have evolved to disrupt meiotic drive and enforce fairness.
 
@@
 
-In this article we explore through simple models the consequences of alleles that influence the outcome 
+In this article we explore through simple population genetic models the consequences of alleles that influence the outcome 
 of female meiosis. 
@@
 This suggests that females are unlikely to evolve to prevent the influence of sperm on meiosis,
-and indeed feature of meiosis may evolve that facilitate the interaction of sperm with female meiosis.
+and indeed features of meiosis may evolve that facilitate the interaction of sperm with female meiosis.
 
@@
 Cartoon figure of sperms meeting eggs with different genotypes?
+
+\begin{figure}
+	\rotatebox{270}{\includegraphics[width = 0.8
+          \textwidth]{Figures/sperm_egg_cartoon1.ps}}
+\caption{transmission probabilities for alleles through female
+  meiosis depend on sperm genotype. 2 allele models}  
+\label{Eggsperm_2_allele_cartoon}
+\end{figure}
 

### 2013-09-10 scripts/invasion_analysis.R

@@ -35,7 +35,6 @@ test.fixation.of.simple.driver<-function(d,s.het,s.hom,x=.999,steps=10){
 	D["3","23"]<-0.5
 	female.transmission.probs<-make.sperm.dep.female.transmission.prob(D)
 	old.geno.freqs<-iterate.1.locus.drive(s.array=s.array,num.iterations=steps,female.transmission.probs=female.transmission.probs,initialize.allele.freqs =c(1-x, x,0))
-
 	fixing<-old.geno.freqs[["allele.freqs"]][steps,2]>old.geno.freqs[["allele.freqs"]][1,2]
 	return(fixing)
 }
@@ -177,12 +176,46 @@ fixation.grid<-sapply(s3.range,function(s3){
 })
 fixation.lines.s3<-s3.range[apply(fixation.grid,1,function(x){max(which(x))})]
 
+
+my.cols<-brewer.pal(2,name="Dark2")
 pdf(file=paste(directory,"invasion_space_recessive_driver.pdf",sep=""))
-plot(c(0.5,1),c(0,1),xlab="Drive coefficient",ylab="selection coeff against homozygotes")
+plot(c(0.5,1),c(0,1),xlab="Drive coefficient",ylab="selection coeff against homozygotes",type="n")
 polygon(x=c(d.range,rev(d.range)),y=c(fixation.lines.s3,rep(1,length(d.range))),col="white")
-polygon(x=c(d.range,rev(d.range)),y=c(fixation.lines.s3,rev(invasion.lines.s3.cutoff)),col="red")
-polygon(x=c(d.range,rev(d.range)),y=c(invasion.lines.s3.cutoff,rep(0,length(d.range))),col="blue")
-text(0.7,0.6,"Simple driver invades but can't fix")
-text(0.9,0.3,"Simple driver can fix")
-text(0.8,0.05,"Simple driver & Self promoter driver can fix",col="white")
+polygon(x=c(d.range,rev(d.range)),y=c(fixation.lines.s3,rev(invasion.lines.s3.cutoff)),col=my.cols[1])
+polygon(x=c(d.range,rev(d.range)),y=c(invasion.lines.s3.cutoff,rep(0,length(d.range))),col=my.cols[2])
+text(0.8,0.35,"Simple driver invades but can't fix",srt=25)
[…]

### 2013-09-10 scripts/iterate_drive_model.R

@@ -26,6 +26,8 @@ iterate.1.locus.drive<-function(s.array,num.iterations,female.transmission.probs
 		
 		}
 
+	my.freqs<-rbind(my.freqs,geno.freqs)
+
 	for(i in 1:num.iterations){
 	#selection
 		geno.freqs["11"]<-geno.freqs["11"] * (1-s.array["11"])

### 2013-09-11 writeup_and_notes/male_dependent_drive.tex
added invasion figure file
added discussion of homozyg. cost invasion to paper

@@
 
-YB: I think we should do our best to be very short and sweet here. There are many opportunities to move off to cool math and cool biological scenarios. 
-We should fight all of them and focus on mission critical results.
-All others can go in supp or appease reviewers.
-Lets talk about why the critical results are.
-To me its (1) male enhancement of female drive cannot maintain a stable polymorphisms, and (2) An allele in sperm can evolve to suppress its drive in females.
+\yb{YB: To me its (1) male enhancement of female drive cannot maintain a stable polymorphisms, and (2) An allele in sperm can evolve to suppress its drive in females.}
 
 Rough order of rest of text
-\subsection*{Lay out Model} 
-We imagine that drive acts in such a way that an allele promote itself into a fraction of d
- eggs ($d> 1/2$, when it is present in a heterozygote, forcing the other allele to be present in $1-d$ eggs. 
+\subsection*{A single locus model} 
+The standard model of female meiotic drive is two allele model, where the
+allele $1$ is transmitted $d$ fraction of the time through female
+meiosis ($d > \half$), regardless of the genotype of the sperm. 
 In mammals, fertilization takes place at MII, so we imagine this drive must be taking place at MII in order
 for a sperm to have any influence. For drive to take place at MII there has to be an uneven number of
- crossovers between the centromere and the drive locus, such that realistically $d$ is bounded to be $<XX$. 
+ crossovers between the centromere and the drive locus, such that realistically $d$ is bounded to be $<0.75$?. 
 In other systems where fertilization occurs during MI, sperm could influence either drive at MI or MII, 
@@
 
-Cartoon figure of sperms meeting eggs with different genotypes?
+We can model a simple sperm dependent meiotic drive system, by
+modifying this model so that the allele only drives during female
+meiosis if it fertilized by a sperm carrying that allele
+ (see
+Figure \ref{Eggsperm_2_allele_cartoon}). Under what conditions can
+this self propograting driver allele spread, and could it drive the
+evolution of supressors?
+
+We first consider the situation where the driving allele has no
+fitness consequences in heterozygotes, only in homozygotes (whose
+relative fitness is $1-s$). A standard meiotic drive allele can always
+invade the population in this situation, as it is initially only
+present mostly in heterozygotes and rarely suffers the consequences of
+the problems in homozygotes. However, if the cost in homozygotes is
+strong enough ($s>XXd$), then the allele is prevented from reaching
+fixation and instead is maintained as a balanced polymorphism in the
+population \citep{Prout1973}. Such balanced polymorphisms can then
+drive the evolution of supressors of drive \citep{XX}, and most of our
+examples of drive systems are of this form.
+
+Under a recessive fitness cost model our sperm dependent driver
+actually has far more restrictive conditions to invade the population,
+than a standard meiotic driver (see Figure \ref{Invasion_space}).
+That is because our allele which drives only when it is present in the
+fertilizing sperm is guarenteed to form a homzygote. Thus, 
+to even spread in the population when it is rare, our sperm-dependent
+driver has to overcome the cost it suffers in the homozygote state.
+
+When conditions are suitable for such alleles to spread into the
+population, they spread slowly through low frequencies as they rarely
+drive as they are unlikely to be fertilized by a sperm with the right
+allele. Once they become established they spread rapidly due to their
+positive frequency dependent behavior (Inset in Figure \ref{Invasion_space}). 
+Unlike standard meiotic drive systems 
+these alleles can not become balanced in the population  by homozygote
+cost, as they have already over come this cost just to enter the population.
+
+HETERO
 
@@
 \caption{transmission probabilities for alleles through female
-  meiosis depend on sperm genotype. 2 allele models}  
+  meiosis depend on sperm genotype. 2 allele model}  
 \label{Eggsperm_2_allele_cartoon}
@@
 \begin{figure}
-	\rotatebox{270}{\includegraphics[width = 0.8
-          \textwidth]{Figures/sperm_egg_cartoon1.ps}}
-\caption{transmission probabilities for alleles through female
-  meiosis depend on sperm genotype. 2 allele models}  
-\label{Eggsperm_2_allele_cartoon}
+\includegraphics[width = 0.8 \textwidth]{Figures/invasion_space_recessive_driver.eps}
+\caption{Invasion analysis. Likely merge with Figure 1.} \label{Invasion_space}
 \end{figure}
@@
 \subsection*{ A) Invasion by single self promoting allele}
+
 show that such alleles:

### 2013-09-11 writeup_and_notes/male_dependent_drive.tex
finished a very rough draft of results, and added figs

@@
 \usepackage[normalem]{ulem}  
+	\newcommand{\gc}[1]{{ \color{red} #1}}
+	\newcommand{\yb}[1]{{ \color{blue} #1}}
 
-\title{Sperm modifiers of female drive.}
+%Why do animals allow male control of female meiosis?
+
+%Why does female meiosis allow the opportunity for exploitation by selfish sperm?
+
+%sperm and egg genomes both favour the peaceful resolution of female meiotic drive
+%strange bed fellows
+
+\title{Why does female meiosis allow the opportunity for exploitation
+  by selfish sperm? BETTER TITLE?}
 \author{Yaniv Brandvain \\ email: ybrandvain@gmail.com  \and Graham Coop \\ email: gmcoop@ucdavis.edu }
@@
 The threat of meiotic drive to organismal fitness is potentially so
-severe that it has been hypothesized that many basic properties of meiosis and o\"{o}genesis, including the initial genome doubling in meiosis I \citep{Haig1991}, the commonness of arrested female meiosis \citep{Mira1998}, and sex differences in the recombination rate \citep{Haig2010,Brandvain2012} have evolved to disrupt meiotic drive and enforce fairness.
+severe that it has been hypothesized that many basic properties of
+meiosis and o\"{o}genesis, including the initial genome doubling in
+meiosis I \citep{Haig1991}, arrested female meiosis
+\citep{Mira1998}, centromere mechinery \citep{MalikHenikoff}, and sex differences in the recombination rate \citep{Haig2010,Brandvain2012} have evolved to disrupt meiotic drive and enforce fairness.
 
@@
 %are mechanistic reasons to think that this is more than idle speculation.
-Mechanistic evidence for this possibility comes from \emph{C.
-elegans}, where experimentally suppressing XXX leads to premature deployment of the aster (a vital component of mitotic machinery) provided by the sperm, disrupting MII meiotic segregation in the egg, leading to a triploid zygote.
+Mechanistic evidence for this possibility comes from
+\emph{C. elegans}, where experimentally suppressing XXX leads to
+premature deployment of the aster (a vital component of mitotic
+machinery) provided by the sperm, disrupting MII meiotic segregation in the egg, leading to a triploid zygote. 
 %This demonstrates that physical components derived from the sperm are are capable of affecting the outcome of female meiosis.
-Additionally, genetic evidence suggests that the transmission patterns in heterozygous females can depend on sperm haplotype. 
+Additionally, genetic evidence suggests that the transmission patterns
+in heterozygous females many potentially depend on sperm haplotype. 
 Specifically, the two best characterized female meiotic drive systems in mouse (In and Om), both operate by distorting the second meiotic division, 
@@
 
-\yb{YB: To me its (1) male enhancement of female drive cannot maintain a stable polymorphisms, and (2) An allele in sperm can evolve to suppress its drive in females.}
+\subsection*{ A) Invasion of the population by a driving allele that promotes
+itself.}
 
-Rough order of rest of text
-\subsection*{A single locus model} 
 The standard model of female meiotic drive is two allele model, where the
 allele $1$ is transmitted $d$ fraction of the time through female
-meiosis ($d > \half$), regardless of the genotype of the sperm. 
+meiosis ($d > \frac{1}{2}$), regardless of the genotype of the sperm. 
 In mammals, fertilization takes place at MII, so we imagine this drive must be taking place at MII in order
@@
 meiosis if it fertilized by a sperm carrying that allele
- (see
-Figure \ref{Eggsperm_2_allele_cartoon}). Under what conditions can
+ (see Figure \ref{Eggsperm_2_allele_cartoon}). Under what conditions can
 this self propograting driver allele spread, and could it drive the
@@
 drive as they are unlikely to be fertilized by a sperm with the right
-allele. Once they become established they spread rapidly due to their
+allele. Once they become established they spread rapidly to fixation due to their
 positive frequency dependent behavior (Inset in Figure \ref{Invasion_space}). 
@@
 
-HETERO
+Things are even worse for the sperm-dependent driver if it suffers any
+fitness cost in the heterozygote. Our standard driver can invade the
+population from low frequency if its drive in heterozygotes out weighs
+the fitness cost they suffer (XXX). However, when at low frequency our
+sperm dependent driver is often in heterozygotes but will only drive
+if fertilized by the same allele. If our allele incurs even very weak costs
+in heterozygotes it can be prevented from spreading from very low
+frequencies, unless its drive is very strong. The dynamics of 
+our sperm-dependent drivers with an additive cost resemble an underdominant polymorphism
+(in that the system is bistable). Below a cutoff frequency the allele is
+unable to enter the population (in a deterministic model) as it is
+paying the cost of the drive but drive is ineffective (see Figure
+S\ref{bistable_additive}). Above this
+frequency the allele increases deterministically as its drive is sufficiently effective 
+becayse heterozygote eggs are fertilized by sperm with our allele
+sufficiently often to beat the costs. 
+This bistability suggests that sperm-dependent drivers will often be 
+unable to enter into reasonably sized populations, and if they do they
+should rapidly be spread fixation.
+
+A model where the sperm allele affects outcome of female meiosis may not be biologically
+realistic as there is very little sperm expression. More realistically the male genotype that produced the sperm
+may be more relevant, as males could place products in their sperm that influenced the outcome of female meoisis.
+However, in practice models that allow the influence of fertilizing male genotype on female meiosis 
+seem to behave qualitatively to those based on the allele of the
+fertilizing sperm (see Supplementary Figures XXX and XXX).
+
+Given the difficulty that meotic drive alleles that promote themselves
+through sperm have even entering the population, and the speed at
+which they fix if they do, it seems very unlikely that such alleles
+could drive the evolution of female
 
@@
 
-Cost of alleles
+\begin{figure}
+	\rotatebox{270}{\includegraphics[width = 0.8
+          \textwidth]{Figures/sperm_egg_cartoon2.ps}}
+\caption{transmission probabilities righthand allele through female
+  meiosis}  \label{Eggsperm_3_allele_cartoon}
+\end{figure}
 
-We'll setup a model where the sperm allele affects outcome of female meiosis. This may not be biologically
-realistic as there is very little sperm expression. More realistically the male genotype that produced the sperm
-may be more relevant, as males could place products in their sperm that influenced the outcome of female meoisis.
-However, in practice models that allow the influence of fertilizing male genotype on female meiosis 
-seem to behave very similiarly to those based on the allele of the fertilizing sperm.
-For simplicity we'll focus on models based on the allele of the sperm, as it's bloodly complicated enough already.
+\begin{figure}
+\includegraphics[width = 0.8 \textwidth]{Figures/trajectories_of_sperm_based_supressors.eps} 
+\caption{Trajectories of two sperm-based supressors of drive. Possibly
+merge this figure the 3 allele cartoon}  \label{Trajectories_of_supressors}
+\end{figure}
 
-\subsection*{ A) Invasion by single self promoting allele}
+\subsection*{B) A more biologically realistic selfish sperm-egg haplotype system}
+These self promoting alleles are somewhat biologically unrealistic as
+we are hypothesizing a newly evolved allele  
+that both drives in female meiosis and has the ability to influence that drive in sperm.
+Perhaps it is more realistic to think that a female drive system first
+evolves and is maintained as a polymorphism in the population. 
+This can then allow another allele to appears that further facilitates
+female meiotic drive through its action in the sperm.
 
+Our sperm-acting allele will be best able to spread if it arises in very close
+linkage with the initial meiotic drive system as then it could potentially
+hitchhiking due to the additional drive it causes. Polymorphic transmission distortion systems
+have often evolved complex sets of inversions to prevent their breakup
+\citep{XXX}, so it is reasonable to assume that for some female drive
+systems there may be a reasonably large mutational target allowing our
+allele to arise in tight linkage. Assuming very tight linkage we can
+model this system using a model with three alleles at the same
+locus. The first two alleles ($A$ and $B$) consist of our ancestral non-driver and
+our original driver allele. The third allele ($C$) is the 
+combined haplotype of our driver allele and the fully linked
+sperm-acting promotor of drive.
+ See Figure
+\ref{Eggsperm_3_allele_cartoon} for a simple schematic of this drive model.
+
+The problem for our sperm-acting promoter of drive (the novel C
+allele), is that in order to have had time to evolve our original
+drive system ($B$) has to be trapped at drive-homozygote disadvantage selection balance. 
+Our new allele has arisen on the drive background so it
+suffers from the same disadvantage. 
+Worse yet, when our new allele it is rare it is often not helping
+itself drive in females but instead $B$ ($C$ sperm meeting $AB$ eggs).
+As $C$ facilitiates $B$ these interactions will frequently form 
+$BC$ heterozygote which suffers from the full
+effect fitness costs as $BB$.
+
+Sperm-acting selfing alleles are therefore at a profound disadvantage
+in this scenerio, even more so than under the previous two allele model.
+We have found no parameter range of this
+three allele system that allows the selfish sperm-acting allele $C$ to
+invade the population.
+
+There are other types of sperm acting alleles that influence
+female meiosis that can invade the population. 
+In particular sperm-acting alleles that act to restore 
+the fairness of meiosis in females can spread. 
+We could imagine that these act by 
+disrupting whatever mechanistic asymmetry in female meiotic segregation the
+drive initially evolved to exploit \citep{Pardo-ManuelDeVillena2001a}.\\
+
+If these sperm-acting disrupting alleles arise at unlinked loci or on
+the $A$ background they can readily invade a population segregating
+for the drive system ($AB$, see Figure ). The invasion of such allele
+will lower the frequency of the original drive system (perhaps to zero)
+and will spread to fixation if they do not carry strong fitness costs
+(Figure \ref{Eggsperm_3_allele_cartoon} and \ref{Trajectories_of_supressors}A). \\
+%These alleles are the sperm-acting equivalent of female 
+%supressor of drive.
+
+Intriguingly these sperm-acting distrupting alleles can also spread
+when they arise tightly linked to the original drive system (B), forming
+an novel third allele/haplotype C (Figures \ref{Eggsperm_3_allele_cartoon} and \ref{Trajectories_of_supressors}B). This new allele spreads
+to displace the original drive system (B) as it benefits from drive
+but by disrupting drive when it is in the sperm it forms
+ the deleterious $BC$ and $CC$ combinations less often (than the
+ analogous combinations for the $B$ allele). 
+\gc{Do we mention the In and Om locus here or in the discussion as a
+  potential e.g. Do we mention the fact that the frequency of these
+  newly invade sperm dodging alleles can be lower than the original system?}\\
+
+Finally given a system of a balanced meiotic drive system which avoid
+its own ill effects through a sperm-based mechanism
+, such as that depicted in Figure XX and XX, female-based modifiers 
+can evolve that act to further facilitate the sperms action in
+disrupting driver.
+
+\section*{Discussion}
+
+This logic may not hold for sex chromosomes. In ZW systems 
+ Male modification of recombination rates
+POssibility that this could happen in plants if pollen emit signals to ``egg''
+Discussion of OM and IN.
+
+\section*{OUTLINE}
+\yb{YB: To me its (1) male enhancement of female drive cannot maintain a stable polymorphisms, and (2) An allele in sperm can evolve to suppress its drive in females.}
 show that such alleles:
@@
 state when they enter the population, and sweep quickly (all the way to fixation) if they enter the pop at all.\\
-
-\begin{figure}
-	\rotatebox{270}{\includegraphics[width = 0.8
-          \textwidth]{Figures/sperm_egg_cartoon2.ps}}
-\caption{transmission probabilities righthand allele through female
-  meiosis}  \label{Eggsperm_3_allele_cartoon}
-\end{figure}
-
-\subsection*{B) A more biologically realistic selfish sperm system}
-Three allele systems. See Figure \ref{Eggsperm_3_allele_cartoon} \\ 
-Perhaps these self promoting alleles are biologically unrealistic as new alleles needs 
-both ability to drive in female meiosis and to influence that drive in sperm.
-Perhaps more realistic to think that a female drive system evolves, 
-and then a sperm modifier appears. This will only occur if original system is trapped 
-at drive-selection balance. So driver must have homozygous cost.
 
@@
 it can not generate too dire a consequence.
-I THINK THERE IS A MORE SUBTLE POINT . The sperm has special knowledge that if it allows drive it will end up in the low fitness homozygote.
-
-This logic may not hold for sex chromosomes. In ZW systems 
- Male modification of recombination rates
-POssibility that this could happen in plants if pollen emit signals to ``egg''
-Discussion of OM and IN.
+I THINK THERE IS A MORE SUBTLE POINT . The sperm has special knowledge
+that if it allows drive it will end up in the low fitness homozygote.
 

### 2013-09-11 scripts/invasion_analysis.R

@@ -129,56 +129,69 @@ lines(d.range,fixation.lines.s3)
 
 
 ####bistable invasion of driver against heterozyg. cost
-pdf(file=paste(directory,"bistable_x_vs_d_additive_s.pdf",sep=""))
+#pdf(file=paste(directory,"bistable_x_vs_d_additive_s.pdf",sep=""))
 d.range<-seq(0.5,1,length=200)
 x.range<- 10^seq(-7,0,length=200)  #seq(0.001,.99,length=50)
-plot(range(d.range),c(10^(-6),1),type="n",log="y",xlab="D",ylab="x bistable cutoff")
 
 my.s<-c(0.01,0.001,0.0001,0.00001)
-my.cols<-brewer.pal(length(my.s),name="Dark2")
-my.x.bistable<-numeric()
 
-for(i in 1:length(my.s)){
+if(FALSE){
+	my.x.bistable<-numeric()
 	
-	invasion.grid.bistable<-sapply(d.range,function(d){
-		sapply(x.range,function(x){
-			test.invasion.self.prop(d=d,s.het=my.s[i],s.hom=my.s[i]*2.0,x=x,step=10)
+	for(i in 1:length(my.s)){
+		
+		invasion.grid.bistable<-sapply(d.range,function(d){
+			sapply(x.range,function(x){
+				test.invasion.self.prop(d=d,s.het=my.s[i],s.hom=my.s[i]*2.0,x=x,step=10)
+			})
 		})
-	})
-	#image(d.range,x.range,invasion.grid.bistable,log="y")
-	invasion.x.lines<-x.range[apply(invasion.grid.bistable,1,function(x){min(which(x))})]
-	lines(d.range[-length(d.range)],invasion.x.lines[-length(d.range)],col=my.cols[i],lwd=2)
-#	text(d.range[round(length(d.range)/2)],invasion.x.lines[round(length(d.range)/2)],s,col="red")
-	my.x.bistable<-rbind(my.x.bistable,invasion.x.lines)
+		#image(d.range,x.range,invasion.grid.bistable,log="y")
+		invasion.x.lines<-x.range[apply(invasion.grid.bistable,1,function(x){min(which(x))})]
+	#	text(d.range[round(length(d.range)/2)],invasion.x.lines[round(length(d.range)/2)],s,col="red")
+		my.x.bistable<-rbind(my.x.bistable,invasion.x.lines)
+	}
+	
[…]

### 2013-09-11 scripts/iterate_drive_model.R

@@ -161,12 +161,12 @@ plot.freqs<-function(new.geno.freqs,plot.HWE.dev=FALSE){
 }
 
 
-run.iterations<-function(D,s2,s3,sperm.or.geno.dependent,num.iter=4000,plot.initial.rise=FALSE){
+run.iterations<-function(D,s.array,sperm.or.geno.dependent,num.iter=4000,plot.initial.rise=FALSE){
 	if(sperm.or.geno.dependent=="sperm"){ female.transmission.probs<-make.sperm.dep.female.transmission.prob(D)}
 	if(sperm.or.geno.dependent=="geno"){ female.transmission.probs<-make.male.geno.dep.female.transmission.prob(D)}
 	
-	old.geno.freqs<-iterate.1.locus.drive(s2=s2,s3=s3,num.iterations=num.iter,female.transmission.probs=female.transmission.probs)
-	new.geno.freqs<-iterate.1.locus.drive(s2=s2,s3=s3,num.iterations=num.iter,female.transmission.probs=female.transmission.probs,my.geno.freqs=old.geno.freqs)
+	old.geno.freqs<-iterate.1.locus.drive(s.array,num.iterations=num.iter,female.transmission.probs=female.transmission.probs)
+	new.geno.freqs<-iterate.1.locus.drive(s.array=s.array,num.iterations=num.iter,female.transmission.probs=female.transmission.probs,my.geno.freqs=old.geno.freqs)
 	
 	if(plot.initial.rise) new.geno.freqs[["allele.freqs"]]<-rbind(old.geno.freqs[["allele.freqs"]],new.geno.freqs[["allele.freqs"]])
 	plot.freqs(new.geno.freqs)
@@ -179,6 +179,49 @@ directory<-"~/Dropbox/Ideas/Om/scripts/"
 geno.names<-c("11", "12","13", "22", "23", "33")
 transmitted.allele<-c(1,1,1,2,1,3,2,2,2,3,3,3)
 
+
+figures.of.traj.for.paper(){
+	layout(t(1:2)); par(mar=c(3,3,1,1))
+	##female driver followed by sperm supressor
+	
+	s.array<-rep(0,6);names(s.array)<-c("11","12","13","23","22","33")
+	s.array["22"]<- 1
+
+	D1 = 1
+	D2 = 0.5
[…]

### 2013-09-12 writeup_and_notes/male_dependent_drive.tex
Yaniv's changes on Thurs

--- 2013-09-10 writeup_and_notes/male_dependent_drive.tex
@@
 \begin{center}
-Center for Population Biology \& Department of Evolution and Ecology \\ University of California - Davis \\ Davis, CA, 95616
+Center for Population Biology \& Section in Evolution and Ecology \\ University of California - Davis \\ Davis, CA, 95616
 \end{center}
@@
 %However, this harmony is incomplete, as in numerous cases an allele can benefit (in the short or long term) from taking advantage of its host.
-Despite the apparent unity of the organism, occasionally alleles can
-gain an evolutionary advantage at a cost to  individual fitness
-\citep{Burt2006}, often by exploiting meiosis and gametogensis.
+Despite the apparent unity of the organism, occasionally alleles can gain an evolutionary advantage at a cost to  individual fitness \citep{Burt2006}.
 %A clear opportunity for this conflict arises during gametogenesis where alternative alleles are in direct competition for representation in the next generation.
-Female meiosis, an asymmetric event in which only one of two alternate alleles
-enters the  egg, with the others consigned to the polar body, is one such occasion \citep{Sandler1957,Pardo-ManuelDeVillena2001a}. 
+Female meiosis, an event in which only one of two alternate alleles enters the  egg, is one such occasion \citep{Sandler1957,Pardo-ManuelDeVillena2001a}. 
 %Female meiosis is particularly unfair, only one of the four products of meiosis becomes an egg, while the other threes are disgarded into polar bodies. 
@@
 %out weighs their ability to spread to spread through heterozygotes. 
-Meiotic drivers observed in nature (in both plants
-\citep{Buckler1999,Fishman2005,Fishman2008}, and animals
-\citep{Agulnik1990,Wu2005,Pardo-ManuelDeVillena2001c}) highlight this
-conflict -- the selfish benefits of drive and the associated
-pleiotropic fitness costs sustain a balanced polymorphism
-\citep{Prout1973}, 
-and often generate on ongoing evolutionary escalation of drive suppressors and enhancers \citep{Dawe1996,Fishman2008}. 
+Meiotic drivers observed in nature (in both plants \citep{Buckler1999,Fishman2005,Fishman2008}, and animals \citep{Agulnik1990,Wu2005,Pardo-ManuelDeVillena2001c}) highlight this conflict -- the selfish benefits of drive and the associated pleiotropic fitness costs sustain a balanced polymorphism, and often generate on ongoing evolutionary escalation of drive suppressors and enhancers \citep{Dawe1996,Fishman2008}.
 %Such balanced drive systems can cause the subsequent evolution of linked enhancers of drive
@@
 %a great diversity of enhancers and supressors. 
-The threat of meiotic drive to organismal fitness is potentially so
-severe that it has been hypothesized that many basic properties of meiosis and o\"{o}genesis, including the initial genome doubling in meiosis I \citep{Haig1991}, the commonness of arrested female meiosis \citep{Mira1998}, and sex differences in the recombination rate \citep{Haig2010,Brandvain2012} have evolved to disrupt meiotic drive and enforce fairness.
+The threat of meiotic drive to organismal fitness is potentially so severe that that many basic properties of meiosis and o\"{o}genesis, including the initial genome doubling in meiosis I \citep{Haig1991}, the commonness of arrested female meiosis \citep{Mira1998}, and sex differences in the recombination rate \citep{Haig2010,Brandvain2012} have putatively evolved to discourage the evolution of meiotic drive an enforce fairness.
 
@@
 
-In this article we explore through simple population genetic models the consequences of alleles that influence the outcome 
+In this article we explore through simple models the consequences of alleles that influence the outcome 
 of female meiosis. 
@@
 This suggests that females are unlikely to evolve to prevent the influence of sperm on meiosis,
-and indeed features of meiosis may evolve that facilitate the interaction of sperm with female meiosis.
+and indeed feature of meiosis may evolve that facilitate the interaction of sperm with female meiosis.
 
@@
 Cartoon figure of sperms meeting eggs with different genotypes?
-
-\begin{figure}
-	\rotatebox{270}{\includegraphics[width = 0.8
-          \textwidth]{Figures/sperm_egg_cartoon1.ps}}
-\caption{transmission probabilities for alleles through female
-  meiosis depend on sperm genotype. 2 allele models}  
-\label{Eggsperm_2_allele_cartoon}
-\end{figure}
 
@@
 so the allele has drive capabilities, and can have had time to evolve new biology.
+
+Multilocus drivers held together with inversions, may offer a relatively large mutational/functional target.
+
 Evolving on the new background means that the allele suffers the fitness consequences of the 

### 2013-09-14 writeup_and_notes/male_dependent_drive.tex
Yaniv's changes upto page 5 on Friday

--- 2013-09-11 writeup_and_notes/male_dependent_drive.tex
@@
 Female meiosis, an asymmetric event in which only one of two alternate alleles
-enters the  egg, with the others consigned to the polar body, is one such occasion \citep{Sandler1957,Pardo-ManuelDeVillena2001a}. 
+enters the  egg, with the others consigned to the polar body, \yb{particularly vulnerable to such exploitation}  \citep{Sandler1957,Pardo-ManuelDeVillena2001a}. 
 %Female meiosis is particularly unfair, only one of the four products of meiosis becomes an egg, while the other threes are disgarded into polar bodies. 
@@
 meiosis I \citep{Haig1991}, arrested female meiosis
-\citep{Mira1998}, centromere mechinery \citep{MalikHenikoff}, and sex differences in the recombination rate \citep{Haig2010,Brandvain2012} have evolved to disrupt meiotic drive and enforce fairness.
+\citep{Mira1998}, centromere machinery \citep{MalikHenikoff}, and sex differences in the recombination rate \citep{Haig2010,Brandvain2012} have evolved to disrupt meiotic drive and enforce fairness.
 
@@
 At first sight, it seems as although female meiosis is primed to be exploited by selfish sperm systems.  
+\yb{Why then is the requirement of fertilization to complete female meiosis so ubiquitous? }
 
@@
 
-Why then is the requirement of fertilization to complete female meiosis so ubiquitous? 
+%Why then is the requirement of fertilization to complete female meiosis so ubiquitous? 
 It is certainly not the case that animals are mechanistically incapable of evolving past this requirement.
@@
 Specifically, the two best characterized female meiotic drive systems in mouse (In and Om), both operate by distorting the second meiotic division, 
-and in both systems the outcome of female meiosis depends the genotype of the fertilizing sperm \citep{Agulnik1993,Wu2005}.
+and in both systems the outcome of female meiosis depends the genotype of the fertilizing sperm \citep{Agulnik1993,Wu2005}; 
+	\yb{however,  we acknowledge the difficulty in reeling out early mortality as an alternative explanation of these putative drivers}.
 
-In this article we explore through simple population genetic models the consequences of alleles that influence the outcome 
-of female meiosis. 
-We use these models to argue that it is actually surprisingly hard for the influence of
-sperm on the outcome of meiosis to drive sustained conflict. In fact we find that sperm and egg 
-genomes' interests are often aligned as they are both invested in the fate of the zygote they will form (as was suggested for the In locus \citep{Pomiankowski1993}).
-This suggests that females are unlikely to evolve to prevent the influence of sperm on meiosis,
-and indeed features of meiosis may evolve that facilitate the interaction of sperm with female meiosis.
+Here, we develop population genetic models exploring the evolution of alleles in sperm that influence the outcome of female meiosis 
+%In this article we explore through simple population genetic models the consequences of alleles that influence the outcome of female meiosis.
+These models argue that it is actually surprisingly hard for the influence of sperm on the outcome of meiosis to drive sustained conflict. 
+In fact, the interests of sperm and egg  genomes' are often aligned, as are both invested in the fate of the resultant zygote 
+	(as was speculated for the In locus \citep{Pomiankowski1993}).
+Therefore, there is little selective benefit to females in preventing sperm to influence female meioses,
+	indeed features of meiosis may evolve that facilitate the interaction of sperm with female meiosis.
 
@@
 
-The standard model of female meiotic drive is two allele model, where the
-allele $1$ is transmitted $d$ fraction of the time through female
-meiosis ($d > \frac{1}{2}$), regardless of the genotype of the sperm. 
-In mammals, fertilization takes place at MII, so we imagine this drive must be taking place at MII in order
-for a sperm to have any influence. For drive to take place at MII there has to be an uneven number of
- crossovers between the centromere and the drive locus, such that realistically $d$ is bounded to be $<0.75$?. 
-In other systems where fertilization occurs during MI, sperm could influence either drive at MI or MII, 
-and drivers at MI can have a $d=1$ if they occur in tight linkage with the centromere.
+In the standard model of female meiotic drive, the driving allele at a balletic site is transmitted to the egg with probability  $d > \frac{1}{2}$, regardless of sperm genotype. 
+%The standard model of female meiotic drive is two allele model, where the allele $1$ is transmitted $d$ fraction of the time through female meiosis ($d > \frac{1}{2}$), regardless of the genotype of the sperm.
+\yb{ To depict a case of a self-promoting (or `green-bearded') meiotic driver,  we modify this standard model such that the female meiotic driver is only effective when fertilized by a sperm carrying that allele (see Figure \ref{Eggsperm_2_allele_cartoon}).}
+We then identify the conditions under which this self propagating driver can spread, and if it drive the evolution of suppressors? 
+\yb{We note that the timing of fertilization relative to female meiosis places another constrain ton $d$, 
+	for example, if fertilization (and therefore, sperm dependent drive) takes place at MII (as in mammals),
+	female drive requires an uneven number of crossovers between the centromere and the drive locus, 
+	so $d$ is bounded to be $<0.75$?.}  
+\yb{I THINK THIS ORDER IS A BIT BETTER, BUT STILL ``OFF'', perhaps this is a footnote, or in methods or something...}
+%In mammals, fertilization takes place at MII, so we imagine this drive must be taking place at MII in order
+%for a sperm to have any influence. For drive to take place at MII there has to be an uneven number of
+% crossovers between the centromere and the drive locus, such that realistically $d$ is bounded to be $<0.75$?. 
+%In other systems where fertilization occurs during MI, sperm could influence either drive at MI or MII, 
+%and drivers at MI can have a $d=1$ if they occur in tight linkage with the centromere.
 
-We can model a simple sperm dependent meiotic drive system, by
-modifying this model so that the allele only drives during female
-meiosis if it fertilized by a sperm carrying that allele
- (see Figure \ref{Eggsperm_2_allele_cartoon}). Under what conditions can
-this self propograting driver allele spread, and could it drive the
-evolution of supressors?
+%We can model a simple sperm dependent meiotic drive system, by
+%modifying this model so that the allele only drives during female
+%meiosis if it fertilized by a sperm carrying that allele
+ %(see Figure \ref{Eggsperm_2_allele_cartoon}). Under what conditions can
+%this self propagating driver allele spread, and could it drive the
+%evolution of supressors?
 
-We first consider the situation where the driving allele has no
-fitness consequences in heterozygotes, only in homozygotes (whose
-relative fitness is $1-s$). A standard meiotic drive allele can always
-invade the population in this situation, as it is initially only
-present mostly in heterozygotes and rarely suffers the consequences of
-the problems in homozygotes. However, if the cost in homozygotes is
-strong enough ($s>XXd$), then the allele is prevented from reaching
-fixation and instead is maintained as a balanced polymorphism in the
-population \citep{Prout1973}. Such balanced polymorphisms can then
-drive the evolution of supressors of drive \citep{XX}, and most of our
-examples of drive systems are of this form.
+\yb{We first consider the situation where the driving allele is neutral in heterozygotes, 
+	but deleterious in homozygotes (whose relative fitness is $1-s$). } 
+%We first consider the situation where the driving allele is has no
+%fitness consequences in heterozygotes, only in homozygotes (whose
+%relative fitness is $1-s$). 
+\yb{In this model, a standard meiotic driver can always invade because 
+	when rare it occurs predominantly in heterozygotes and therefore it initially pays no fitness cost.} 
+%A standard meiotic drive allele can always
+%invade the population in this situation, as it is initially only
+%present mostly in heterozygotes and rarely suffers the consequences of
+%the problems in homozygotes. 
+\yb{However, a severe homozygotes fitness cost ($s>XXd$) prevents  
+	drivers from fixing  \citep{Prout1973}, 
+	and provides an opportunity for the
+	evolution of drive suppressors  \citep{XX}, corresponding well to 
+	empirical examples of female meiotic drive. }
 
-Under a recessive fitness cost model our sperm dependent driver
-actually has far more restrictive conditions to invade the population,
-than a standard meiotic driver (see Figure \ref{Invasion_space}).
-That is because our allele which drives only when it is present in the
-fertilizing sperm is guarenteed to form a homzygote. Thus, 
-to even spread in the population when it is rare, our sperm-dependent
-driver has to overcome the cost it suffers in the homozygote state.
+%However, if the cost in homozygotes is
+%strong enough ($s>XXd$), then the allele is prevented from reaching
+%fixation and instead is maintained as a balanced polymorphism in the
+%population \citep{Prout1973}. Such balanced polymorphisms can then
+%drive the evolution of supressors of drive \citep{XX}, and most of our
+%examples of drive systems are of this form.
 
-When conditions are suitable for such alleles to spread into the
-population, they spread slowly through low frequencies as they rarely
-drive as they are unlikely to be fertilized by a sperm with the right
-allele. Once they become established they spread rapidly to fixation due to their
-positive frequency dependent behavior (Inset in Figure \ref{Invasion_space}). 
-Unlike standard meiotic drive systems 
-these alleles can not become balanced in the population  by homozygote
-cost, as they have already over come this cost just to enter the population.
+\yb{In contrast to a traditional driver, which can drive but pay no fitness cost when rare, 
+	our green-bearded driver drives by creating homozygotes, and therefore pays a fitness cost even when rare. 
+To simply increase when rare, our sperm-dependent driver must overcome a homozygous fitness cost.
+Therefore with a recessive fitness cost,
+	the conditions allowing the invasion of our green-bearded driver 
+ 	are far more restrictive than those for a standard meiotic driver (see Figure \ref{Invasion_space}).}
 
-Things are even worse for the sperm-dependent driver if it suffers any
-fitness cost in the heterozygote. Our standard driver can invade the
-population from low frequency if its drive in heterozygotes out weighs
-the fitness cost they suffer (XXX). However, when at low frequency our
-sperm dependent driver is often in heterozygotes but will only drive
-if fertilized by the same allele. If our allele incurs even very weak costs
-in heterozygotes it can be prevented from spreading from very low
-frequencies, unless its drive is very strong. The dynamics of 
-our sperm-dependent drivers with an additive cost resemble an underdominant polymorphism
-(in that the system is bistable). Below a cutoff frequency the allele is
-unable to enter the population (in a deterministic model) as it is
-paying the cost of the drive but drive is ineffective (see Figure
-S\ref{bistable_additive}). Above this
-frequency the allele increases deterministically as its drive is sufficiently effective 
-becayse heterozygote eggs are fertilized by sperm with our allele
-sufficiently often to beat the costs. 
-This bistability suggests that sperm-dependent drivers will often be 
-unable to enter into reasonably sized populations, and if they do they
-should rapidly be spread fixation.
+%Under a recessive fitness cost model our sperm dependent driver
+%actually has far more restrictive conditions to invade the population,
+%than a standard meiotic driver (see Figure \ref{Invasion_space}).
+%That is because our allele which drives only when it is present in the
+%fertilizing sperm is guarenteed to form a homzygote. Thus, 
+%to even spread in the population when it is rare, our sperm-dependent
+%driver has to overcome the cost it suffers in the homozygote state.
+
+\yb{When a deleterious recessive green-bearded driver does spread, 
+	it spends much time at low frequency, because when rare it is unlikely to meet a complimentary sperm. 
+However, once common, a green bearded driver spreads rapidly to fixation due to their
+	positive frequency dependent behavior (Inset in Figure \ref{Invasion_space}).  PERHAPS COMPARE TO TRADITIONAL DRIVER IN THIS FIGURE
+Unlike standard meiotic drive systems, polymorphism of a green-bearded driver cannot be sustained by a homozygous cost, 
+	because it already overcame this cost just to enter the population.
+The lack of a balanced polymorphism precludes the evolution of an allele that suppresses this form of meiotic drive.}
+
+%When conditions are suitable for such alleles to spread into the
+%population, they spread slowly through low frequencies as they rarely
+%drive as they are unlikely to be fertilized by a sperm with the right
+%allele. Once they become established they spread rapidly to fixation due to their
+%positive frequency dependent behavior (Inset in Figure \ref{Invasion_space}). 
+%Unlike standard meiotic drive systems 
+%these alleles can not become balanced in the population  by homozygote
+%cost, as they have already over come this cost just to enter the population.
+
+\yb{Inclusion of a heterozygous fitness cost (e.g. additivity) further constrains the evolution of a sperm-dependent driver. 
+A  standard female  driver can invade from low frequency so long as drive in heterozygotes out weighs the fitness cost they suffer (XXX); 
+	however, when at low frequency our sperm dependent driver pays the heterozygous fitness cost but rarely drives because it is rare in sperm.
+Therefore, even very weak heterozygous fitness costs
+	prevent the spread of a very rare green-bearded driver, unless its drive is very strong. 
+With an additive cost, the dynamics of a green-bearded driver resembles that of an underdominant mutation (in that the system is bistable). 
+Below a cutoff frequency the allele deterministically decreases in frequency 
+	because it pays the  heterozygous fitness cost  but drives is ineffectively (see FigureS\ref{bistable_additive}). 
+Above this frequency, the allele increases deterministically because the commonness of sperm carrying this allele 
+	make drive effective enough to overcome the associated fitness cost.
+%becayse heterozygote eggs are fertilized by sperm with our allele %sufficiently often to beat the costs. 
+This bistability will  prevent sperm-dependent drivers from invading 	
+	reasonably sized populations, and assures that if they do invade they will rapidly fix.}
+
+%Things are even worse for the sperm-dependent driver if it suffers any
+%fitness cost in the heterozygote. Our standard driver can invade the
+%population from low frequency if its drive in heterozygotes out weighs
+%the fitness cost they suffer (XXX). However, when at low frequency our
+%sperm dependent driver is often in heterozygotes but will only drive
+%if fertilized by the same allele. If our allele incurs even very weak costs
+%in heterozygotes it can be prevented from spreading from very low
+%frequencies, unless its drive is very strong. The dynamics of 
+%our sperm-dependent drivers with an additive cost resemble an underdominant polymorphism
+%(in that the system is bistable). Below a cutoff frequency the allele is
+%unable to enter the population (in a deterministic model) as it is
+%paying the cost of the drive but drive is ineffective (see Figure
+%S\ref{bistable_additive}). Above this
+%frequency the allele increases deterministically as its drive is sufficiently effective 
+%becayse heterozygote eggs are fertilized by sperm with our allele
+%sufficiently often to beat the costs. 
+%This bistability suggests that sperm-dependent drivers will often be 
+%unable to enter into reasonably sized populations, and if they do they
+%should rapidly be spread fixation.
+
+XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX
+EDITED TO HERE
+XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX
 

### 2013-09-14 scripts/invasion_analysis.R
adding to male determined drive code

@@ -1,6 +1,6 @@
 library("RColorBrewer")
 
-test.invasion.self.prop<-function(d,s.het,s.hom,x=.001,steps=10){
+test.invasion.self.prop<-function(d,s.het,s.hom,x=.001,steps=10,sperm.or.male="sperm"){
 	s.array<-rep(0,6)
 	names(s.array)<-c("11","12","13","23","22","33")
 	s.array["13"]<- s.het
@@ -13,9 +13,11 @@ test.invasion.self.prop<-function(d,s.het,s.hom,x=.001,steps=10){
 	D["3","13"]<-d
 	D["1 or 2","23"]<- 1-0.5
 	D["3","23"]<-0.5
-	female.transmission.probs<-make.sperm.dep.female.transmission.prob(D)
+	stopifnot(sperm.or.male=="sperm" | sperm.or.male=="male")
+	if(sperm.or.male=="sperm"){ female.transmission.probs<-make.sperm.dep.female.transmission.prob(D)}
+	if(sperm.or.male=="male"){ female.transmission.probs<-make.male.geno.dep.female.transmission.prob(D) }
 	old.geno.freqs<-iterate.1.locus.drive(s.array,num.iterations=steps,female.transmission.probs=female.transmission.probs,initialize.allele.freqs =c(1-x, 0,x))
-
+recover()
 	invading<-old.geno.freqs[["allele.freqs"]][steps,3]>old.geno.freqs[["allele.freqs"]][1,3]
 	return(invading)
 }
@@ -225,10 +227,49 @@ for(i in 1:3){
 	 D["3","13"]<-drive.coeffs[i]
 	 D["1 or 2","23"]<- 1-0.5
 	 D["3","23"]<-0.5
-		 female.transmission.probs<-make.male.geno.dep.female.transmission.prob(D)
+		 female.transmission.probs<-make.sperm.dep.female.transmission.prob(D)
 
 	old.geno.freqs<-iterate.1.locus.drive(s.array=s.array,num.iterations=2000,female.transmission.probs=female.transmission.probs,initialize.allele.freqs =c(0.99, 0,0.01 ))
 	lines(old.geno.freqs[["allele.freqs"]][,3],col=my.point.cols[i],lwd=2)
 	}
 
 dev.copy2eps(file=paste(directory,"invasion_space_recessive_driver.eps",sep=""))
+
+
+#####invasion space for male genotype controlled allele
+
[…]

### 2013-09-14 scripts/iterate_drive_model.R

@@ -137,9 +137,45 @@ make.male.geno.dep.female.transmission.prob<-function(D){
 
 	transmission.probs[male.geno.3,colnames(transmission.probs) == "23" & transmitted.allele == 2] <- 1-D["3","23"]
 	transmission.probs[male.geno.3,colnames(transmission.probs) =="23" & transmitted.allele == 3] <- D["3","23"]
-	
+	recover()
 transmission.probs
 }
+
+
+make.male.geno.dep.female.transmission.prob.w.dom<-function(D){
+ 	geno.names<-c("11", "12","13", "22", "23", "33")
+	transmission.probs<-matrix(0.5,nrow=12,ncol=12)
+	rownames(transmission.probs)<-rep(geno.names,each=2)
+	colnames(transmission.probs)<-rep(geno.names,each=2)
+	transmitted.allele<-c(1,1,1,2,1,3,2,2,2,3,3,3)
+ 	
+ 	
+	transmission.probs[,colnames(transmission.probs) =="12" & transmitted.allele == 1] <- 1-D["1 or 2","12"]
+	transmission.probs[,colnames(transmission.probs) =="12" & transmitted.allele == 2] <- D["1 or 2","12"]
+
+	transmission.probs[,colnames(transmission.probs) =="13" & transmitted.allele == 1] <- 1-D["1 or 2","13"]
+	transmission.probs[,colnames(transmission.probs) =="13" & transmitted.allele == 3] <- D["1 or 2","13"]
+
+	transmission.probs[,colnames(transmission.probs) == "23" & transmitted.allele == 2] <- 1-D["1 or 2","23"]
+	transmission.probs[,colnames(transmission.probs) =="23" & transmitted.allele == 3] <- D["1 or 2","23"]
+
+
+	##males carrying allele 3 alter rate of female drive in dominant fashion. 
+	
+	male.geno.3<-grep("3",rownames(transmission.probs))
+	
+	transmission.probs[male.geno.3,colnames(transmission.probs) == "13" & transmitted.allele == 1] <- 1-D["3","13"]
+	transmission.probs[male.geno.3,colnames(transmission.probs) =="13" & transmitted.allele == 3] <- D["3","13"]
+
+	transmission.probs[male.geno.3,colnames(transmission.probs) == "12" & transmitted.allele == 1] <- 1-D["3","12"]
[…]

### 2013-09-14 writeup_and_notes/male_dependent_drive.tex

@@
 
+%yb version
 In the standard model of female meiotic drive, the driving allele at a balletic site is transmitted to the egg with probability  $d > \frac{1}{2}$, regardless of sperm genotype. 
-%The standard model of female meiotic drive is two allele model, where the allele $1$ is transmitted $d$ fraction of the time through female meiosis ($d > \frac{1}{2}$), regardless of the genotype of the sperm.
-\yb{ To depict a case of a self-promoting (or `green-bearded') meiotic driver,  we modify this standard model such that the female meiotic driver is only effective when fertilized by a sperm carrying that allele (see Figure \ref{Eggsperm_2_allele_cartoon}).}
+To depict a case of a self-promoting (or `green-bearded') meiotic driver,  we modify this standard model such that the female meiotic driver is only effective when fertilized by a sperm carrying that allele (see Figure \ref{Eggsperm_2_allele_cartoon}).
 We then identify the conditions under which this self propagating driver can spread, and if it drive the evolution of suppressors? 
-\yb{We note that the timing of fertilization relative to female meiosis places another constrain ton $d$, 
+We note that the timing of fertilization relative to female meiosis places another constrain ton $d$, 
 	for example, if fertilization (and therefore, sperm dependent drive) takes place at MII (as in mammals),
 	female drive requires an uneven number of crossovers between the centromere and the drive locus, 
-	so $d$ is bounded to be $<0.75$?.}  
+	so $d$ is bounded to be $<0.75$?.
 \yb{I THINK THIS ORDER IS A BIT BETTER, BUT STILL ``OFF'', perhaps this is a footnote, or in methods or something...}
+
+%coop version
+%The standard model of female meiotic drive is two allele model, where the allele $1$ is transmitted $d$ fraction of the time through female meiosis ($d > \frac{1}{2}$), regardless of the genotype of the sperm.
 %In mammals, fertilization takes place at MII, so we imagine this drive must be taking place at MII in order
@@
 
-\yb{We first consider the situation where the driving allele is neutral in heterozygotes, 
-	but deleterious in homozygotes (whose relative fitness is $1-s$). } 
+% YB VERSION
+We first consider the situation where the driving allele is neutral in heterozygotes, 
+	but deleterious in homozygotes (whose relative fitness is $1-s$).  
+In this model, a standard meiotic driver can always invade because 
+	when rare it occurs predominantly in heterozygotes and therefore it initially pays no fitness cost. 
+However, a severe homozygotes fitness cost ($s>XXd$) prevents  
+	drivers from fixing  \citep{Prout1973}, 
+	and provides an opportunity for the
+	evolution of drive suppressors  \citep{XX}, corresponding well to 
+	empirical examples of female meiotic drive.
+
+%COOP VERSION
+
 %We first consider the situation where the driving allele is has no
@@
 %relative fitness is $1-s$). 
-\yb{In this model, a standard meiotic driver can always invade because 
-	when rare it occurs predominantly in heterozygotes and therefore it initially pays no fitness cost.} 
 %A standard meiotic drive allele can always
@@
 %the problems in homozygotes. 
-\yb{However, a severe homozygotes fitness cost ($s>XXd$) prevents  
-	drivers from fixing  \citep{Prout1973}, 
-	and provides an opportunity for the
-	evolution of drive suppressors  \citep{XX}, corresponding well to 
-	empirical examples of female meiotic drive. }
-
 %However, if the cost in homozygotes is
@@
 
-\yb{In contrast to a traditional driver, which can drive but pay no fitness cost when rare, 
+%YB VERSION
+In contrast to a traditional driver, which can drive but pay no fitness cost when rare, 
 	our green-bearded driver drives by creating homozygotes, and therefore pays a fitness cost even when rare. 
@@
 	the conditions allowing the invasion of our green-bearded driver 
- 	are far more restrictive than those for a standard meiotic driver (see Figure \ref{Invasion_space}).}
+ 	are far more restrictive than those for a standard meiotic driver (see Figure \ref{Invasion_space}).
 
+%COOP VERSION
 %Under a recessive fitness cost model our sperm dependent driver
@@
 
-\yb{When a deleterious recessive green-bearded driver does spread, 
+%YB VERSION
+When a deleterious recessive green-bearded driver does spread, 
 	it spends much time at low frequency, because when rare it is unlikely to meet a complimentary sperm. 
@@
 	because it already overcame this cost just to enter the population.
-The lack of a balanced polymorphism precludes the evolution of an allele that suppresses this form of meiotic drive.}
+The lack of a balanced polymorphism precludes the evolution of an allele that suppresses this form of meiotic drive.
 
+%COOP VERSION
 %When conditions are suitable for such alleles to spread into the
@@
 
-\yb{Inclusion of a heterozygous fitness cost (e.g. additivity) further constrains the evolution of a sperm-dependent driver. 
+%YB VERSION
+Inclusion of a heterozygous fitness cost (e.g. additivity) further constrains the evolution of a sperm-dependent driver. 
 A  standard female  driver can invade from low frequency so long as drive in heterozygotes out weighs the fitness cost they suffer (XXX); 
@@
 This bistability will  prevent sperm-dependent drivers from invading 	
-	reasonably sized populations, and assures that if they do invade they will rapidly fix.}
+	reasonably sized populations, and assures that if they do invade they will rapidly fix.
 
+%COOP VERSION
 %Things are even worse for the sperm-dependent driver if it suffers any
@@
 
-XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX
-EDITED TO HERE
-XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX
-
-A model where the sperm allele affects outcome of female meiosis may not be biologically
-realistic as there is very little sperm expression. More realistically the male genotype that produced the sperm
-may be more relevant, as males could place products in their sperm that influenced the outcome of female meoisis.
-However, in practice models that allow the influence of fertilizing male genotype on female meiosis 
+%yb version
+%trying to make the models above sound less silly
+While the sperm allele  could plausible influence the outcome of female meiosis, the limited levels of gene expression in sperm 
+	suggest that sperm may influence the outcome of female meiosis as a function a the diploid genotype that produced the sperm perhaps due to proteins and RNAs packed in the sperm.
+Models where male genotype, rather than sperm haplotype, determines the outcome of female meiosis 
 	seem to behave qualitatively to those based on the allele of the
 	fertilizing sperm (see Supplementary Figures XXX and XXX).
+	\yb{THIS IS NOW WRONG WE NEED TO THINK ABOUT HOW TO WRITE THIS}
 
-Given the difficulty that meotic drive alleles that promote themselves
-through sperm have even entering the population, and the speed at
-which they fix if they do, it seems very unlikely that such alleles
-could drive the evolution of female
+%coop version
+%A model where the sperm allele affects outcome of female meiosis may not be biologically realistic as there is very little sperm expression. 
+%More realistically the male genotype that produced the sperm
+%may be more relevant, as males could place products in their sperm that influenced the outcome of female meoisis.
+%However, in practice models that allow the influence of fertilizing male genotype on female meiosis 
+%seem to behave qualitatively to those based on the allele of the
+%fertilizing sperm (see Supplementary Figures XXX and XXX).
+
+Given the difficulty that meiotic drive alleles that promote themselves
+through sperm have entering the population, the speed at
+which they fix if they do, \yb{and the absence of balanced polymorphisms at such loci,} it seems very unlikely that such alleles
+could drive the evolution of female \yb{suppressors}.
 
@@
 \end{figure}
+
+XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX\newline
+EDITED TO HERE\newline
+XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX\newline
 

### 2013-09-15 writeup_and_notes/male_dependent_drive.tex
Yaniv comments through discussion on sat

@@
 %A clear opportunity for this conflict arises during gametogenesis where alternative alleles are in direct competition for representation in the next generation.
-Female meiosis, an asymmetric event in which only one of two alternate alleles
-enters the  egg, with the others consigned to the polar body, \yb{particularly vulnerable to such exploitation}  \citep{Sandler1957,Pardo-ManuelDeVillena2001a}. 
+Because only one of the four products of female meiosis is transmitted to the egg, female meiosis \yb{is particularly vulnerable to such exploitation}  \citep{Sandler1957,Pardo-ManuelDeVillena2001a}. 
 %Female meiosis is particularly unfair, only one of the four products of meiosis becomes an egg, while the other threes are disgarded into polar bodies. 
@@
 The threat of meiotic drive to organismal fitness is potentially so
-severe that it has been hypothesized that many basic properties of
+severe that many basic properties of
 meiosis and o\"{o}genesis, including the initial genome doubling in
 meiosis I \citep{Haig1991}, arrested female meiosis
-\citep{Mira1998}, centromere machinery \citep{MalikHenikoff}, and sex differences in the recombination rate \citep{Haig2010,Brandvain2012} have evolved to disrupt meiotic drive and enforce fairness.
+\citep{Mira1998}, centromere machinery \citep{MalikHenikoff}, and sex differences in the recombination rate \citep{Haig2010,Brandvain2012} 
+have perhaps evolved to disrupt meiotic drive and enforce fairness.
 
@@
 it is potentially open to sabotage by a virtual stranger -- a haploid sperm genome.
-That is, in many animal species, the completion of female meiosis requires fertilization of the egg, 
-and there is ample opportunity for interaction between the sperm and female meiotic machinery.
+That is, in many animal species, the completion of female meiosis requires fertilization of the egg \yb{CITE}, 
+and there is ample opportunity for interaction between the sperm and female meiotic machinery \yb{CITE}.
 If, for example, an allele in sperm could facilitate meiotic drive by a genetically equivalent allele in a 
@@
 %Why then is the requirement of fertilization to complete female meiosis so ubiquitous? 
-It is certainly not the case that animals are mechanistically incapable of evolving past this requirement.
+It is certainly not the case that animals are mechanistically incapable of \yb{circumventing} this requirement.
 There is considerable variation in which stage of meiosis requires fertilization, and 
@@
 
-Here, we develop population genetic models exploring the evolution of alleles in sperm that influence the outcome of female meiosis 
+Here, we develop population genetic models exploring the evolution of \yb{sperm influence on female meiotic drive}. 
 %In this article we explore through simple population genetic models the consequences of alleles that influence the outcome of female meiosis.
-These models argue that it is actually surprisingly hard for the influence of sperm on the outcome of meiosis to drive sustained conflict. 
+These models show that \yb{sperm modifiers of female meiotic drive are unlikely to create a sustained conflict.} 
+%These models argue that it is surprisingly hard for the influence of sperm on the outcome of meiosis to drive sustained conflict. 
 In fact, the interests of sperm and egg  genomes' are often aligned, as are both invested in the fate of the resultant zygote 
@@
 %yb version
-In the standard model of female meiotic drive, the driving allele at a balletic site is transmitted to the egg with probability  $d > \frac{1}{2}$, regardless of sperm genotype. 
-To depict a case of a self-promoting (or `green-bearded') meiotic driver,  we modify this standard model such that the female meiotic driver is only effective when fertilized by a sperm carrying that allele (see Figure \ref{Eggsperm_2_allele_cartoon}).
-We then identify the conditions under which this self propagating driver can spread, and if it drive the evolution of suppressors? 
-We note that the timing of fertilization relative to female meiosis places another constrain ton $d$, 
+In the standard model of female meiotic drive, \yb{the driver is one of two alleles, and in heterozygotes is transmitted to the egg} with probability  $d > \frac{1}{2}$, regardless of sperm genotype. 
+To depict a case of a self-promoting (or `green-bearded') meiotic driver,  we modify this standard model such that the driver is only effective when fertilized by a sperm carrying that allele (see Figure \ref{Eggsperm_2_allele_cartoon}).
+We then identify the conditions under which this self propagating driver can spread, and evaluate whether a driver of this form could 
+	\yb{generate a sustained conflict favoring the evolution of suppressors}. 
+We note that the timing of fertilization relative to female meiosis places another constraint on $d$, 
 	for example, if fertilization (and therefore, sperm dependent drive) takes place at MII (as in mammals),
@@
 % YB VERSION
-We first consider the situation where the driving allele is neutral in heterozygotes, 
+We first consider a driving allele neutral in heterozygotes, 
 	but deleterious in homozygotes (whose relative fitness is $1-s$).  
 In this model, a standard meiotic driver can always invade because 
-	when rare it occurs predominantly in heterozygotes and therefore it initially pays no fitness cost. 
-However, a severe homozygotes fitness cost ($s>XXd$) prevents  
+	when rare it occurs predominantly in heterozygotes and it therefore drives without a fitness cost. 
+However, a severe homozygous fitness cost ($s>XXd$) prevents  
 	drivers from fixing  \citep{Prout1973}, 
-	and provides an opportunity for the
-	evolution of drive suppressors  \citep{XX}, corresponding well to 
+	providing an opportunity for the evolution of
+	drive suppressors \citep{XX}, corresponding well to 
 	empirical examples of female meiotic drive.
@@
 %YB VERSION
-In contrast to a traditional driver, which can drive but pay no fitness cost when rare, 
-	our green-bearded driver drives by creating homozygotes, and therefore pays a fitness cost even when rare. 
-To simply increase when rare, our sperm-dependent driver must overcome a homozygous fitness cost.
+In contrast to a traditional driver, which drives but pays no fitness cost as a rare deleterious recissive, 
+	our green-bearded driver specifically creates homozygotes when driving, and therefore there is a drive-associated fitness cost even when rare. 
+Our sperm-dependent driver must overcome this homozygous fitness cost simply to spread when rare.
 Therefore with a recessive fitness cost,
@@
 	it spends much time at low frequency, because when rare it is unlikely to meet a complimentary sperm. 
-However, once common, a green bearded driver spreads rapidly to fixation due to their
-	positive frequency dependent behavior (Inset in Figure \ref{Invasion_space}).  PERHAPS COMPARE TO TRADITIONAL DRIVER IN THIS FIGURE
-Unlike standard meiotic drive systems, polymorphism of a green-bearded driver cannot be sustained by a homozygous cost, 
+However, once common, a green-bearded driver spreads rapidly to fixation due to its
+	positive frequency dependent behavior (Inset in Figure \ref{Invasion_space}).  \yb{PERHAPS COMPARE TO TRADITIONAL DRIVER IN THIS FIGURE} 
+Unlike standard meiotic drive systems, a homozygous cost, cannot sustain polymorphism of a green-bearded driver, 
 	because it already overcame this cost just to enter the population.
@@
 Inclusion of a heterozygous fitness cost (e.g. additivity) further constrains the evolution of a sperm-dependent driver. 
-A  standard female  driver can invade from low frequency so long as drive in heterozygotes out weighs the fitness cost they suffer (XXX); 
-	however, when at low frequency our sperm dependent driver pays the heterozygous fitness cost but rarely drives because it is rare in sperm.
-Therefore, even very weak heterozygous fitness costs
-	prevent the spread of a very rare green-bearded driver, unless its drive is very strong. 
-With an additive cost, the dynamics of a green-bearded driver resembles that of an underdominant mutation (in that the system is bistable). 
-Below a cutoff frequency the allele deterministically decreases in frequency 
-	because it pays the  heterozygous fitness cost  but drives is ineffectively (see FigureS\ref{bistable_additive}). 
+In contrast to a  standard female  driver which can invade from low frequency so long as drive in heterozygotes outweighs the fitness cost they suffer (XXX); 
+	when rare our sperm dependent driver pays the heterozygous fitness cost but only weakly drives because it is rare in sperm. 
+This frequency independent fitness cost, and positive frequency dependent drive generates a bistable system. 
+%Therefore, even very weak heterozygous fitness costs
+%	prevent the spread of a very rare green-bearded driver, unless its drive is very strong. 
+%With an additive cost, the dynamics of a green-bearded driver resembles that of an underdominant mutation (in that the system is bistable). 
+Below a cutoff frequency the green-bearded driver deterministically decreases in frequency 
+	because it pays the  heterozygous fitness cost  but drives ineffectively (see FigureS\ref{bistable_additive}). 
 Above this frequency, the allele increases deterministically because the commonness of sperm carrying this allele 
@@
 %becayse heterozygote eggs are fertilized by sperm with our allele %sufficiently often to beat the costs. 
-This bistability will  prevent sperm-dependent drivers from invading 	
+This bistability prevents sperm-dependent drivers from invading 	
 	reasonably sized populations, and assures that if they do invade they will rapidly fix.
@@
 %trying to make the models above sound less silly
-While the sperm allele  could plausible influence the outcome of female meiosis, the limited levels of gene expression in sperm 
-	suggest that sperm may influence the outcome of female meiosis as a function a the diploid genotype that produced the sperm perhaps due to proteins and RNAs packed in the sperm.
+While an allele specifically active in sperm allele  could plausibly influence the outcome of female meiosis, the limited levels of gene expression in sperm 
+	suggest that sperm may influence female meiosis as a function a the diploid genotype that produced the sperm (perhaps due to proteins and RNAs packed in the sperm).
 Models where male genotype, rather than sperm haplotype, determines the outcome of female meiosis 
@@
 
-Given the difficulty that meiotic drive alleles that promote themselves
-through sperm have entering the population, the speed at
+Given the difficulty that green-bearded meiotic drivers have entering the population, the speed at
 which they fix if they do, \yb{and the absence of balanced polymorphisms at such loci,} it seems very unlikely that such alleles
@@
 \subsection*{B) A more biologically realistic selfish sperm-egg haplotype system}
-These self promoting alleles are somewhat biologically unrealistic as
-we are hypothesizing a newly evolved allele  
-that both drives in female meiosis and has the ability to influence that drive in sperm.
-Perhaps it is more realistic to think that a female drive system first
-evolves and is maintained as a polymorphism in the population. 
-This can then allow another allele to appears that further facilitates
-female meiotic drive through its action in the sperm.
+Above, we explored a model were a driver that drove in females when signaled by a complementary allele in sperm.  
+This complicated pleiotropic effect, while plausible is somewhat unlikely. 
+Here, we examine an alternative scenario (and more realistic) in which first a driver is at drive-viability equilibrium,
+	and subsequently a sperm dependent modifier of drive tightly linked to this allele invades.
+	\yb{COMMENT XXX This is unclear. I think invades is wrong, bc it doesn't invade.  I think also we need to say if further increases Drive from D to D+ XXX }
+\yb{Assuming tight linkage between female driver and sperm modifier is consistent with the nature of well characterized meiotic drive systems 
+	which are often maintained as polymorphic inversions with numerous linked modifiers (CITE).}
+\yb{??In the SUPP we explore the dynamics of a sperm modifier of female drive unlinked to the female driver??} 
+%somewhat surprising combination of pleiotropic effect
+%These self promoting alleles are somewhat biologically unrealistic as
+%we are hypothesizing a newly evolved allele  
+%that both drives in female meiosis and has the ability to influence that drive in sperm.
+%Perhaps it is more realistic to think that a female drive system first
+%evolves and is maintained as a polymorphism in the population. 
+%This can then allow another allele to appears that further facilitates
+%	female meiotic drive through its action in the sperm. 
+%Our sperm-acting allele will be best able to spread if it arises in very close
+%linkage with the initial meiotic drive system as then it could potentially
+%hitchhiking due to the additional drive it causes. Polymorphic transmission distortion systems
+%have often evolved complex sets of inversions to prevent their breakup
+%\citep{XXX}, so it is reasonable to assume that for some female drive
+%systems there may be a reasonably large mutational target allowing our
+%allele to arise in tight linkage. 
+We represent this as a three allele system where $A$ and $B$ are the ancestral non-driver and driver alleles, 
+	and allele $B'$ is the non-recombining haplotype of a driver allele and sperm-acting drive modifier. 
+	We note that in this model, $B'$ modifies drive of $B$ and $B'$ alleles equally, see Figure \ref{Eggsperm_3_allele_cartoon} for a simple schematic of this drive model.
 
-Our sperm-acting allele will be best able to spread if it arises in very close
-linkage with the initial meiotic drive system as then it could potentially
-hitchhiking due to the additional drive it causes. Polymorphic transmission distortion systems
-have often evolved complex sets of inversions to prevent their breakup
-\citep{XXX}, so it is reasonable to assume that for some female drive
-systems there may be a reasonably large mutational target allowing our
-allele to arise in tight linkage. Assuming very tight linkage we can
-model this system using a model with three alleles at the same
-locus. The first two alleles ($A$ and $B$) consist of our ancestral non-driver and
-our original driver allele. The third allele ($C$) is the 
-combined haplotype of our driver allele and the fully linked
-sperm-acting promotor of drive.
- See Figure
-\ref{Eggsperm_3_allele_cartoon} for a simple schematic of this drive model.
+Because the sperm-acting drive promoter (the novel $B'$ haplotype) is introduced when the ancestral driver ($B$) is at drive-selection balance, 
+	$B'$ will immediately suffer a homozygous fitness cost.  
+%Our new allele has arisen on the drive background so it suffers from the same disadvantage. 
+Worse yet, a novel $B'$ haplotype most often helps 
+	the $B$  allele drive ($B'$ sperm meeting $AB$ eggs), simply because $B$ is initially more common than $B'$.
+As $B'$ facilitates $B$ these interactions will frequently form 
+	$BB'$ heterozygote which suffers from the full
+	effect fitness costs as $BB$. 
+%The problem for our sperm-acting promoter of drive (the novel $B'$ 
+%allele), is that in order to have had time to evolve our original
+%drive system ($B$) has to be trapped at drive-homozygote disadvantage selection balance. 
+%Our new allele has arisen on the drive background so it
+%suffers from the same disadvantage. 
+%Worse yet, when our new allele it is rare it is often not helping
+%itself drive in females but instead $B$ ($B'$ sperm meeting $AB$ eggs).
+%As $B'$ facilitates $B$ these interactions will frequently form 
+%$BB'$ heterozygote which suffers from the full
+%effect fitness costs as $BB$. 
+Sperm-acting self promoting alleles are therefore at a profound disadvantage
+in this scenario, even more so than under the previous two allele model.
+We have found no parameter range of this
+three allele system that allows the selfish sperm-acting allele $B'$ to
+invade the population. \yb{MATHS?}
 
-The problem for our sperm-acting promoter of drive (the novel C
-allele), is that in order to have had time to evolve our original
-drive system ($B$) has to be trapped at drive-homozygote disadvantage selection balance. 
-Our new allele has arisen on the drive background so it
-suffers from the same disadvantage. 
-Worse yet, when our new allele it is rare it is often not helping
-itself drive in females but instead $B$ ($C$ sperm meeting $AB$ eggs).
-As $C$ facilitiates $B$ these interactions will frequently form 
-$BC$ heterozygote which suffers from the full
-effect fitness costs as $BB$.
-
-Sperm-acting selfing alleles are therefore at a profound disadvantage
-in this scenerio, even more so than under the previous two allele model.
-We have found no parameter range of this
-three allele system that allows the selfish sperm-acting allele $C$ to
-invade the population.
-
-There are other types of sperm acting alleles that influence
-female meiosis that can invade the population. 
-In particular sperm-acting alleles that act to restore 
-the fairness of meiosis in females can spread. 
-We could imagine that these act by 
-disrupting whatever mechanistic asymmetry in female meiotic segregation the
-drive initially evolved to exploit \citep{Pardo-ManuelDeVillena2001a}.\\
-
-If these sperm-acting disrupting alleles arise at unlinked loci or on
-the $A$ background they can readily invade a population segregating
+While  sperm enhancement of a female drive cannot displace a polymorphic female driver, sperm based drive suppression - that is sperm-acting alleles that act to restore 
+fairness of female meiosis can. 
+%In particular sperm-acting alleles that act to restore 
+%the fairness of meiosis in females can spread. 
+%We could imagine that these act by 
+%disrupting whatever mechanistic asymmetry in female meiotic segregation the
+%drive initially evolved to exploit \citep{Pardo-ManuelDeVillena2001a}.\\
+%There are other types of sperm acting alleles that influence
+%female meiosis that can invade the population. 
+%In particular sperm-acting alleles that act to restore 
+%the fairness of meiosis in females can spread. 
+%We could imagine that these act by 
+%disrupting whatever mechanistic asymmetry in female meiotic segregation the
+%drive initially evolved to exploit \citep{Pardo-ManuelDeVillena2001a}.\\
+ If these sperm-acting drive suppressors arise on
+	the $A$ background or are unlinked to the drive locus they readily invade a population segregating
 for the drive system ($AB$, see Figure ). The invasion of such allele
@@
 %supressor of drive.
+%If these sperm-acting disrupting alleles arise at unlinked loci or on
+%the $A$ background they can readily invade a population segregating
+%for the drive system ($AB$, see Figure ). The invasion of such allele
+%will lower the frequency of the original drive system (perhaps to zero)
+%and will spread to fixation if they do not carry strong fitness costs
+%(Figure \ref{Eggsperm_3_allele_cartoon} and \ref{Trajectories_of_supressors}A). \\
+%These alleles are the sperm-acting equivalent of female 
+%supressor of drive.
 
-Intriguingly these sperm-acting distrupting alleles can also spread
-when they arise tightly linked to the original drive system (B), forming
-an novel third allele/haplotype C (Figures \ref{Eggsperm_3_allele_cartoon} and \ref{Trajectories_of_supressors}B). This new allele spreads
-to displace the original drive system (B) as it benefits from drive
-but by disrupting drive when it is in the sperm it forms
- the deleterious $BC$ and $CC$ combinations less often (than the
- analogous combinations for the $B$ allele). 
+Intriguingly these sperm disruptors of female meiotic drive can spread
+	even when they arise tightly linked to the original drive system (B), forming
+	an novel third allele $B^*$ as above (Figures \ref{Eggsperm_3_allele_cartoon} and \ref{Trajectories_of_supressors}B). 
+$B^*$ displaces the original drive system ($B$) because it drives when not costly, 
+	but by disrupting drive when transmitted to sperm it avoids low-fitness drive homozygotes ($BB^*$ and $B^*B^*$. 
+Surprising despite displacing the ancestral driver ($B$), self avoiding driver, $B^*$ it's equilibrium frequency is not always higher than that of $B$. 
+	% as it benefits from drive but by disrupting drive when it is in the sperm it forms the deleterious $BC$ and $CC$ combinations less often (than the  analogous combinations for the $B$ allele). 
 \gc{Do we mention the In and Om locus here or in the discussion as a
@@
 can evolve that act to further facilitate the sperms action in
-disrupting driver.
+disrupting driver. \yb{THIS SOUNDS MORE DISCUSSIVE}
 

### 2013-09-19 writeup_and_notes/male_dependent_drive.tex
added 2 figs to the supp, and worked through yaniv's edits to results

@@
 it is potentially open to sabotage by a virtual stranger -- a haploid sperm genome.
-That is, in many animal species, the completion of female meiosis requires fertilization of the egg \yb{CITE}, 
-and there is ample opportunity for interaction between the sperm and female meiotic machinery \yb{CITE}.
+That is, in many animal species, the completion of female meiosis requires fertilization of the egg \citep{Masui_book}, 
+and there is ample opportunity for interaction between the sperm and female meiotic machinery.
 If, for example, an allele in sperm could facilitate meiotic drive by a genetically equivalent allele in a 
@@
 At first sight, it seems as although female meiosis is primed to be exploited by selfish sperm systems.  
-\yb{Why then is the requirement of fertilization to complete female meiosis so ubiquitous? }
+Why then is the requirement of fertilization to complete female meiosis so ubiquitous?
 
-%Another unusual aspect of animal female meiosis is the fact that it is paused at XXX.
-%In fact in many species of animal the final stages of female meiosis are only completed upon fertlization of the egg by sperm.There's considerable variation in 
-%which stage of meiosis requires fertilization.
-%This raises the possibility that an allele in sperm could influence the outcome of meiosis in its favor. 
-%For example imagine an allele that when a sperm bearing this allele were to encounter an egg, whose meiotic product was currently heterozygote for 
-%another copy of itself, the sperm allele biased meiosis in favor of the other copy of itself.
-%At first sight, it seems as although female meiosis is primed to be exploited by selfish sperm systems.
-
-%Why then is the requirement of fertilization to complete female meiosis so ubiquitous? 
-It is certainly not the case that animals are mechanistically incapable of \yb{circumventing} this requirement.
+It is certainly not the case that animals are mechanistically incapable of circumventing this requirement.
 There is considerable variation in which stage of meiosis requires fertilization, and 
 a number of animal clades (should we try and lower bound how many transitions?) have evolved
-to allow the completion of female meiosis upon ovulation. 
-%(Although even in these organisms female meiosis 
-%may also be subjection to influence by sperm alleles, if the sperm arrives at the egg soon enough after 
-%ovulation to influence meiosis.)
+to allow the completion of female meiosis upon ovulation \citep[see
+Table 1 of ][]{Masui_book}.
 
 It is also not the case that sperm is mechanistically incapable of influencing the outcome of female meiosis.
-%This is not just idle speculation, as there is also direct evidence that sperm components and indeed different sperm genotypes can affect female meiosis.
-%are mechanistic reasons to think that this is more than idle speculation.
 Mechanistic evidence for this possibility comes from
-\emph{C. elegans}, where experimentally suppressing XXX leads to
-premature deployment of the aster (a vital component of mitotic
-machinery) provided by the sperm, disrupting MII meiotic segregation in the egg, leading to a triploid zygote. 
-%This demonstrates that physical components derived from the sperm are are capable of affecting the outcome of female meiosis.
+\emph{C. elegans} the fact that premature deployment of the aster (a vital component of mitotic
+machinery) provided by the sperm can disrupting MII meiotic segregation
+in the egg, leading to a triploid zygote \citep{McNally2012}. 
 Additionally, genetic evidence suggests that the transmission patterns
@@
 and in both systems the outcome of female meiosis depends the genotype of the fertilizing sperm \citep{Agulnik1993,Wu2005}; 
-	\yb{however,  we acknowledge the difficulty in reeling out early mortality as an alternative explanation of these putative drivers}.
+However,  we acknowledge the difficulty in ruling out early mortality
+as an alternative explanation of the 
+specifics of these putative drivers \citep[Give page numbers to
+discussion of this in][]{Burt2006}. \gc{should be careful here as the
+  Om locus is pretty convincingly a meiotic driver, doubts have just
+  been raised about whether IN is. Whether the sperm dep nature of
+  either is robust to this is I think more at question.}
 
-Here, we develop population genetic models exploring the evolution of \yb{sperm influence on female meiotic drive}. 
+Here, we develop population genetic models exploring the evolution of sperm influence on female meiotic drive. 
 %In this article we explore through simple population genetic models the consequences of alleles that influence the outcome of female meiosis.
-These models show that \yb{sperm modifiers of female meiotic drive are unlikely to create a sustained conflict.} 
-%These models argue that it is surprisingly hard for the influence of sperm on the outcome of meiosis to drive sustained conflict. 
-In fact, the interests of sperm and egg  genomes' are often aligned, as are both invested in the fate of the resultant zygote 
-	(as was speculated for the In locus \citep{Pomiankowski1993}).
+These models show that sperm modifiers of female meiotic drive are
+unlikely to create a sustained conflict, making it unlikely that
+female meiosis will evolve to resist them.
+In fact, the interests of sperm and egg  genomes' are often aligned, as are both invested in the fate of the resultant zygote (as was speculated for the In locus \citep{Pomiankowski1993}).
 Therefore, there is little selective benefit to females in preventing sperm to influence female meioses,
@@
 
-\subsection*{ A) Invasion of the population by a driving allele that promotes
+\subsection*{ Invasion of the population by a driving allele that promotes
 itself.}
@@
 %yb version
-In the standard model of female meiotic drive, \yb{the driver is one of two alleles, and in heterozygotes is transmitted to the egg} with probability  $d > \frac{1}{2}$, regardless of sperm genotype. 
-To depict a case of a self-promoting (or `green-bearded') meiotic driver,  we modify this standard model such that the driver is only effective when fertilized by a sperm carrying that allele (see Figure \ref{Eggsperm_2_allele_cartoon}).
-We then identify the conditions under which this self propagating driver can spread, and evaluate whether a driver of this form could 
-	\yb{generate a sustained conflict favoring the evolution of suppressors}. 
-We note that the timing of fertilization relative to female meiosis places another constraint on $d$, 
-	for example, if fertilization (and therefore, sperm dependent drive) takes place at MII (as in mammals),
+In the standard model of female meiotic drive, the driver is one of two alleles, and in heterozygotes is transmitted to the egg with probability  $d > \frac{1}{2}$, regardless of sperm genotype. 
+To depict a case of a self-promoting (or `green-bearded') meiotic
+driver,  we modify this standard model such that the driver is only effective when fertilized by a sperm carrying that allele (see Figure
+\ref{Eggsperm_2_allele_cartoon}) \footnote{the timing of fertilization relative to female meiosis places another constraint on $d$, for example, if fertilization (and therefore, sperm dependent drive) takes place at MII (as in mammals),
 	female drive requires an uneven number of crossovers between the centromere and the drive locus, 
-	so $d$ is bounded to be $<0.75$?.
-\yb{I THINK THIS ORDER IS A BIT BETTER, BUT STILL ``OFF'', perhaps this is a footnote, or in methods or something...}
+	so $d$ is bounded to be $<0.75$? \citep[see ][ for discussion]{Buckler1999}. Could also put this in
+        caption of fig.}.
+We then identify the conditions under which this self propagating driver can spread, and evaluate whether a driver of this form could generate a sustained conflict favoring the evolution of suppressors.
 
-%coop version
-%The standard model of female meiotic drive is two allele model, where the allele $1$ is transmitted $d$ fraction of the time through female meiosis ($d > \frac{1}{2}$), regardless of the genotype of the sperm.
-%In mammals, fertilization takes place at MII, so we imagine this drive must be taking place at MII in order
-%for a sperm to have any influence. For drive to take place at MII there has to be an uneven number of
-% crossovers between the centromere and the drive locus, such that realistically $d$ is bounded to be $<0.75$?. 
-%In other systems where fertilization occurs during MI, sperm could influence either drive at MI or MII, 
-%and drivers at MI can have a $d=1$ if they occur in tight linkage with the centromere.
-
-%We can model a simple sperm dependent meiotic drive system, by
-%modifying this model so that the allele only drives during female
-%meiosis if it fertilized by a sperm carrying that allele
- %(see Figure \ref{Eggsperm_2_allele_cartoon}). Under what conditions can
-%this self propagating driver allele spread, and could it drive the
-%evolution of supressors?
-
-% YB VERSION
 We first consider a driving allele neutral in heterozygotes, 
@@
 
-%COOP VERSION
-
-%We first consider the situation where the driving allele is has no
-%fitness consequences in heterozygotes, only in homozygotes (whose
-%relative fitness is $1-s$). 
-%A standard meiotic drive allele can always
-%invade the population in this situation, as it is initially only
-%present mostly in heterozygotes and rarely suffers the consequences of
-%the problems in homozygotes. 
-%However, if the cost in homozygotes is
-%strong enough ($s>XXd$), then the allele is prevented from reaching
-%fixation and instead is maintained as a balanced polymorphism in the
-%population \citep{Prout1973}. Such balanced polymorphisms can then
-%drive the evolution of supressors of drive \citep{XX}, and most of our
-%examples of drive systems are of this form.
-
-%YB VERSION
-In contrast to a traditional driver, which drives but pays no fitness cost as a rare deleterious recissive, 
-	our green-bearded driver specifically creates homozygotes when driving, and therefore there is a drive-associated fitness cost even when rare. 
+In contrast to a traditional driver, which drives but pays effectively
+no fitness cost as a rare deleterious recessive, 
+	our green-bearded driver creates homozygotes
+        when it drives, and therefore there is a drive-associated fitness cost even when rare. 
 Our sperm-dependent driver must overcome this homozygous fitness cost simply to spread when rare.
@@
 However, once common, a green-bearded driver spreads rapidly to fixation due to its
-	positive frequency dependent behavior (Inset in Figure \ref{Invasion_space}).  \yb{PERHAPS COMPARE TO TRADITIONAL DRIVER IN THIS FIGURE} 
-Unlike standard meiotic drive systems, a homozygous cost, cannot sustain polymorphism of a green-bearded driver, 
+	positive frequency dependent behavior (Inset in Figure \ref{Invasion_space}).  
+Unlike standard meiotic drive systems, a homozygous cost, cannot sustain polymorphism of a green-bearded driver 
 	because it already overcame this cost just to enter the population.
@@
 %YB VERSION
-Inclusion of a heterozygous fitness cost (e.g. additivity) further constrains the evolution of a sperm-dependent driver. 
-In contrast to a  standard female  driver which can invade from low frequency so long as drive in heterozygotes outweighs the fitness cost they suffer (XXX); 
-	when rare our sperm dependent driver pays the heterozygous fitness cost but only weakly drives because it is rare in sperm. 
+Inclusion of a heterozygous fitness cost (e.g.
+additive fitness cost) further constrains the evolution of a sperm-dependent driver.
+In contrast to a  standard female  driver which can invade from low frequency so long as drive in heterozygotes outweighs the fitness cost they suffer (XXX); when rare our sperm dependent driver pays the heterozygous
+        fitness cost but only weakly drives because the right sperm
+        are rare. 
 This frequency independent fitness cost, and positive frequency dependent drive generates a bistable system. 
@@
 	make drive effective enough to overcome the associated fitness cost.
-%becayse heterozygote eggs are fertilized by sperm with our allele %sufficiently often to beat the costs. 
 This bistability prevents sperm-dependent drivers from invading 	
@@
 
-%COOP VERSION
-%Things are even worse for the sperm-dependent driver if it suffers any
-%fitness cost in the heterozygote. Our standard driver can invade the
-%population from low frequency if its drive in heterozygotes out weighs
-%the fitness cost they suffer (XXX). However, when at low frequency our
-%sperm dependent driver is often in heterozygotes but will only drive
-%if fertilized by the same allele. If our allele incurs even very weak costs
-%in heterozygotes it can be prevented from spreading from very low
-%frequencies, unless its drive is very strong. The dynamics of 
-%our sperm-dependent drivers with an additive cost resemble an underdominant polymorphism
-%(in that the system is bistable). Below a cutoff frequency the allele is
-%unable to enter the population (in a deterministic model) as it is
-%paying the cost of the drive but drive is ineffective (see Figure
-%S\ref{bistable_additive}). Above this
-%frequency the allele increases deterministically as its drive is sufficiently effective 
-%becayse heterozygote eggs are fertilized by sperm with our allele
-%sufficiently often to beat the costs. 
-%This bistability suggests that sperm-dependent drivers will often be 
-%unable to enter into reasonably sized populations, and if they do they
-%should rapidly be spread fixation.
-
-%yb version
-%trying to make the models above sound less silly
-While an allele specifically active in sperm allele  could plausibly influence the outcome of female meiosis, the limited levels of gene expression in sperm 
-	suggest that sperm may influence female meiosis as a function a the diploid genotype that produced the sperm (perhaps due to proteins and RNAs packed in the sperm).
-Models where male genotype, rather than sperm haplotype, determines the outcome of female meiosis 
-	seem to behave qualitatively to those based on the allele of the
-	fertilizing sperm (see Supplementary Figures XXX and XXX).
-	\yb{THIS IS NOW WRONG WE NEED TO THINK ABOUT HOW TO WRITE THIS}
-
-%coop version
-%A model where the sperm allele affects outcome of female meiosis may not be biologically realistic as there is very little sperm expression. 
-%More realistically the male genotype that produced the sperm
-%may be more relevant, as males could place products in their sperm that influenced the outcome of female meoisis.
-%However, in practice models that allow the influence of fertilizing male genotype on female meiosis 
-%seem to behave qualitatively to those based on the allele of the
-%fertilizing sperm (see Supplementary Figures XXX and XXX).
+While an allele specifically active in sperm could plausibly influence the outcome of female meiosis, the limited levels of gene expression in sperm \citep[e.g.][]{Joseph2004}
+	suggest it may also be plausible to consider 
+sperm may influence female meiosis via expression of the fertilizing male's
+diploid genotype (perhaps due to proteins and RNAs packaged into the sperm).
+This complicates the dynamics of the system, as the dominance of
+  the genotype of the male effect on female drive must be considered as a parameter. 
+While under some circumstances the dominance of the drive effect in
+males can allow the 
+maintainance of the allele a polymorphism, the parameter range that
+permits polymorphism is very narrow (and quite sensitive to the
+dominance of the male effect, see Figure S\ref{Effect_of_dominance}).
 
 Given the difficulty that green-bearded meiotic drivers have entering the population, the speed at
-which they fix if they do, \yb{and the absence of balanced polymorphisms at such loci,} it seems very unlikely that such alleles
-could drive the evolution of female \yb{suppressors}.
+which they fix if they do,
+and the narrow parameter range permitting balanced polymorphisms at such loci, 
+it seems very unlikely that such alleles
+could drive the evolution of female suppressors of sperm-enabled
+female meiotic drive.
 
@@
 \includegraphics[width = 0.8 \textwidth]{Figures/invasion_space_recessive_driver.eps}
-\caption{Invasion analysis. Likely merge with Figure 1.} \label{Invasion_space}
+\caption{Invasion analysis. Likely merge with Figure 1. \yb{PERHAPS COMPARE TO TRADITIONAL DRIVER IN THIS FIGURE} } \label{Invasion_space}
 \end{figure}
+
+\subsection*{ sperm-egg haplotype balanced drive systems}
 
@@
 
-XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX\newline
-EDITED TO HERE\newline
-XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX\newline
-
 \begin{figure}
@@
 
-\subsection*{B) A more biologically realistic selfish sperm-egg haplotype system}
 Above, we explored a model were a driver that drove in females when signaled by a complementary allele in sperm.  
 This complicated pleiotropic effect, while plausible is somewhat unlikely. 
-Here, we examine an alternative scenario (and more realistic) in which first a driver is at drive-viability equilibrium,
-	and subsequently a sperm dependent modifier of drive tightly linked to this allele invades.
-	\yb{COMMENT XXX This is unclear. I think invades is wrong, bc it doesn't invade.  I think also we need to say if further increases Drive from D to D+ XXX }
-\yb{Assuming tight linkage between female driver and sperm modifier is consistent with the nature of well characterized meiotic drive systems 
-	which are often maintained as polymorphic inversions with numerous linked modifiers (CITE).}
+Here, we examine an alternative scenario (and more realistic) in which
+initially a driver, with no sperm-depedency, is at drive-viability equilibrium (with two alleles
+$A$ and $B$ are the ancestral non-driver and driver alleles).
+Subsequently a sperm-acting modifier of female meotic drive arises tightly
+        linked to this balanced locus (effectively creating a third
+        allele/haplotype at this locus).
+Assuming tight linkage between female driver and sperm modifier is consistent with the nature of well characterized meiotic drive systems 
+	which are often maintained as polymorphic inversions with numerous linked modifiers (CITE).
+
 \yb{??In the SUPP we explore the dynamics of a sperm modifier of female drive unlinked to the female driver??}
-%somewhat surprising combination of pleiotropic effect
-%These self promoting alleles are somewhat biologically unrealistic as
-%we are hypothesizing a newly evolved allele  
-%that both drives in female meiosis and has the ability to influence that drive in sperm.
-%Perhaps it is more realistic to think that a female drive system first
-%evolves and is maintained as a polymorphism in the population. 
-%This can then allow another allele to appears that further facilitates
-%	female meiotic drive through its action in the sperm. 
-%Our sperm-acting allele will be best able to spread if it arises in very close
-%linkage with the initial meiotic drive system as then it could potentially
-%hitchhiking due to the additional drive it causes. Polymorphic transmission distortion systems
-%have often evolved complex sets of inversions to prevent their breakup
-%\citep{XXX}, so it is reasonable to assume that for some female drive
-%systems there may be a reasonably large mutational target allowing our
-%allele to arise in tight linkage. 
-We represent this as a three allele system where $A$ and $B$ are the ancestral non-driver and driver alleles, 
-	and allele $B'$ is the non-recombining haplotype of a driver allele and sperm-acting drive modifier. 
-	We note that in this model, $B'$ modifies drive of $B$ and $B'$ alleles equally, see Figure \ref{Eggsperm_3_allele_cartoon} for a simple schematic of this drive model.
 
-Because the sperm-acting drive promoter (the novel $B'$ haplotype) is introduced when the ancestral driver ($B$) is at drive-selection balance, 
+We will first consider the case where a sperm-acting facilitator of
+female meiotic drive arises very tightly linked to the $B$ background 
+creating the third allele $B^{+}$ (i.e. the non-recombining haplotype of a
+        driver allele and sperm-acting drive modifier). 
+This new allele which acts in sperm to increase the effectiveness of both
+the $B$ and  $B^{+}$ drive in $AB$ and $AB^{+}$ heterozygotes, see Figure \ref{Eggsperm_3_allele_cartoon} for a simple schematic of this drive model.  
+Naively this $B^{+}$ allele might be able to increase in frequency in the
+population as then it could be allowed to spread due to the additional
+drive it causes.
+That turns out to not be the case, 
+because the sperm-acting drive promoter (the novel $B^{+}$ haplotype) 
+is introduced when the ancestral driver ($B$) is at drive-selection balance, 
 	$B'$ will immediately suffer a homozygous fitness cost.  
-%Our new allele has arisen on the drive background so it suffers from the same disadvantage. 
-Worse yet, a novel $B'$ haplotype most often helps 
-	the $B$  allele drive ($B'$ sperm meeting $AB$ eggs), simply because $B$ is initially more common than $B'$.
-As $B'$ facilitates $B$ these interactions will frequently form 
-	$BB'$ heterozygote which suffers from the full
+Worse yet, a novel $B^{+}$ haplotype most often helps 
+the $B$  allele drive ($B'$ sperm meeting $AB$ eggs), simply because $B$ is initially more common than $B^{+}$.
+As $B^{+}$ facilitates $B$ these interactions will frequently form 
+	$BB^{+}$ heterozygote that suffers from the full
 	effect fitness costs as $BB$.
-%The problem for our sperm-acting promoter of drive (the novel $B'$ 
-%allele), is that in order to have had time to evolve our original
-%drive system ($B$) has to be trapped at drive-homozygote disadvantage selection balance. 
-%Our new allele has arisen on the drive background so it
-%suffers from the same disadvantage. 
-%Worse yet, when our new allele it is rare it is often not helping
-%itself drive in females but instead $B$ ($B'$ sperm meeting $AB$ eggs).
-%As $B'$ facilitates $B$ these interactions will frequently form 
-%$BB'$ heterozygote which suffers from the full
-%effect fitness costs as $BB$. 
+
 Sperm-acting self promoting alleles are therefore at a profound disadvantage
@@
 We have found no parameter range of this
-three allele system that allows the selfish sperm-acting allele $B'$ to
+three allele system that allows the selfish sperm-acting allele $B^{+}$ to
 invade the population. \yb{MATHS?}
@@
 fairness of female meiosis can. 
-%In particular sperm-acting alleles that act to restore 
-%the fairness of meiosis in females can spread. 
-%We could imagine that these act by 
-%disrupting whatever mechanistic asymmetry in female meiotic segregation the
-%drive initially evolved to exploit \citep{Pardo-ManuelDeVillena2001a}.\\
-%There are other types of sperm acting alleles that influence
-%female meiosis that can invade the population. 
-%In particular sperm-acting alleles that act to restore 
-%the fairness of meiosis in females can spread. 
-%We could imagine that these act by 
-%disrupting whatever mechanistic asymmetry in female meiotic segregation the
-%drive initially evolved to exploit \citep{Pardo-ManuelDeVillena2001a}.\\
  If these sperm-acting drive suppressors arise on
-	the $A$ background or are unlinked to the drive locus they readily invade a population segregating
-for the drive system ($AB$, see Figure ). The invasion of such allele
+	the $A$ background (creating a third allele $A^{-}$) 
+or are unlinked to the drive locus they readily invade a population segregating
+for the drive system ($AB$, see Figure). The invasion of such allele
 will lower the frequency of the original drive system (perhaps to zero)
@@
 (Figure \ref{Eggsperm_3_allele_cartoon} and \ref{Trajectories_of_supressors}A). \\
-%These alleles are the sperm-acting equivalent of female 
-%supressor of drive.
-%If these sperm-acting disrupting alleles arise at unlinked loci or on
-%the $A$ background they can readily invade a population segregating
-%for the drive system ($AB$, see Figure ). The invasion of such allele
-%will lower the frequency of the original drive system (perhaps to zero)
-%and will spread to fixation if they do not carry strong fitness costs
-%(Figure \ref{Eggsperm_3_allele_cartoon} and \ref{Trajectories_of_supressors}A). \\
-%These alleles are the sperm-acting equivalent of female 
-%supressor of drive.
 
@@
 	even when they arise tightly linked to the original drive system (B), forming
-	an novel third allele $B^*$ as above (Figures \ref{Eggsperm_3_allele_cartoon} and \ref{Trajectories_of_supressors}B). 
-$B^*$ displaces the original drive system ($B$) because it drives when not costly, 
-	but by disrupting drive when transmitted to sperm it avoids low-fitness drive homozygotes ($BB^*$ and $B^*B^*$. 
+	an novel third allele $B^{-}$ as above (Figures \ref{Eggsperm_3_allele_cartoon} and \ref{Trajectories_of_supressors}B). 
+$B^{-}$ displaces the original drive system ($B$) because it drives when not costly, 
+	but by disrupting drive when transmitted to sperm it avoids low-fitness drive homozygotes ($BB^-$ and $B^-B^-$. 
 Surprising despite displacing the ancestral driver ($B$), self avoiding driver, $B^*$ it's equilibrium frequency is not always higher than that of $B$.
-	% as it benefits from drive but by disrupting drive when it is in the sperm it forms the deleterious $BC$ and $CC$ combinations less often (than the  analogous combinations for the $B$ allele). 
+
 \gc{Do we mention the In and Om locus here or in the discussion as a
@@
 	 newly invade sperm dodging alleles can be lower than the original system?}\\
+
+\section*{Discussion}
 
@@
 can evolve that act to further facilitate the sperms action in
-disrupting driver. \yb{THIS SOUNDS MORE DISCUSSIVE}
-
-\section*{Discussion}
+disrupting driver.
 
@@
 Discussion of OM and IN.
+
+Discussion of general conclusion that females have little reason to evolve supression mechanisms to prevent sperm influence on meiosis. 
+General logic that sperm genome has to live in a zygote with consequences of its effect on female meiosis, so
+it can not generate too dire a consequence.
+I THINK THERE IS A MORE SUBTLE POINT . The sperm has special knowledge
+that if it allows drive it will end up in the low fitness homozygote.
+
+\bibliography{refs}
+
+\section*{Supplementary Stuff}
+
+\begin{figure}
+\includegraphics[width = 0.8 \textwidth]{Figures/bistable_x_vs_d_additive_s.eps} 
+\caption{Bistable green-bearded allele when selection cost (s) acts
+  additively, for an allele whose effect
+ on female meoisis is mediated by sperm allele. }  \label{bistable_additive}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width = 0.8 \textwidth]{Figures/effect_of_dominance_on_invasion_space_one_graph.eps} 
+\caption{Invasion analysis of green-bearded sperm allele whose effect
+ on female meoisis is mediated by the genotype of the fertilizing
+ male. The area below the solid/dotted line is the parameter region
+ where the allele can invade/fix (from $1/1000$ and $999/1000$
+ \gc{maybe should lower this}). Note
+ that all of the fixation lines are on top of each other as dominance
+ parameter makes little difference here. The allele will be
+ polymorphic in the parameter space where the solid (invasion) line
+ falls above the dotted (fixation) line. Note that this is a very
+ narrow parameter range, and only exists if the allele has a
+ reasonable degree on dominance}  \label{Effect_of_dominance}
+\end{figure}
+
+\end{document}
 
@@
 \section*{Conclusions.}
-Discussion of general conclusion that females have little reason to evolve supression mechanisms to prevent sperm influence on meiosis. 
-General logic that sperm genome has to live in a zygote with consequences of its effect on female meiosis, so
-it can not generate too dire a consequence.
-I THINK THERE IS A MORE SUBTLE POINT . The sperm has special knowledge
-that if it allows drive it will end up in the low fitness homozygote.
-
-\bibliography{refs}
-\end{document}

### 2013-09-19 scripts/invasion_analysis.R

@@ -22,6 +22,29 @@ recover()
 	return(invading)
 }
 
+####invasion test for male dependent drive w. dominance controled by d.dom
+test.invasion.self.prop.w.dom<-function(d,d.dom,s.het,s.hom,x=.001,steps=10){
+	s.array<-rep(0,6)
+	names(s.array)<-c("11","12","13","23","22","33")
+	s.array["13"]<- s.het
+	s.array["33"]<- s.hom
+	
+		D<-matrix(1/2,nrow=3,ncol=3,dimnames=list(c("1 or 2","3.het","3"),c("12","13","23"))) ##entries are drive coeff of the 2nd 
+ ##entries are drive coeff of the 2nd 
+	D["3","13"]<-d
+	D["3.het","13"]<-(d-0.5)*d.dom + 0.5  ##d.dom modifies the effect of drive 
+	D["1 or 2","23"]<- 1-0.5
+	D["3","23"]<-0.5
+	
+	female.transmission.probs<-make.male.geno.dep.female.transmission.prob.w.dom(D) 
+	old.geno.freqs<-iterate.1.locus.drive(s.array,num.iterations=steps,female.transmission.probs=female.transmission.probs,initialize.allele.freqs =c(1-x, 0,x))
+#recover()
+	invading<-old.geno.freqs[["allele.freqs"]][steps,3]>old.geno.freqs[["allele.freqs"]][1,3]
+	return(invading)
+}
+
+
+
 test.fixation.of.simple.driver<-function(d,s.het,s.hom,x=.999,steps=10){
 	s.array<-rep(0,6)
 	names(s.array)<-c("11","12","13","23","22","33")
@@ -235,9 +258,10 @@ for(i in 1:3){
 
 dev.copy2eps(file=paste(directory,"invasion_space_recessive_driver.eps",sep=""))
 
-
+#######################################################
 #####invasion space for male genotype controlled allele
 
+##fully dominant allele
 d.range<-seq(0.5,1,length=200)
 s3.range<-seq(0,1,length=200)
 
@@ -259,10 +283,54 @@ fixation.lines.s3.cutoff.male<-s3.range[apply(fixation.grid.male.control,1,funct
 plot(d.range,invasion.lines.s3.cutoff.male,col="red")
 lines(d.range,fixation.lines.s3.cutoff.male,col="green")
 
+if(FALSE){
[…]

### 2013-09-19 scripts/iterate_drive_model.R

@@ -4,7 +4,7 @@ iterate.1.locus.drive<-function(s.array,num.iterations,female.transmission.probs
 
 	my.freqs<-numeric()
 	if(missing(my.geno.freqs)){
-		print("running 2 allele system to stablity")
+#ŧprint("running 2 allele system to stablity")
 		allele.freqs<-initialize.allele.freqs	
 		names(allele.freqs)<-c("1","2","3")	
 
@@ -162,17 +162,30 @@ make.male.geno.dep.female.transmission.prob.w.dom<-function(D){
 
 	##males carrying allele 3 alter rate of female drive in dominant fashion. 
 	
-	male.geno.3<-grep("3",rownames(transmission.probs))
+	male.heterozy<-rownames(transmission.probs) %in%  c("13","23")
+	male.homozy<-rownames(transmission.probs) =="33"
 	
-	transmission.probs[male.geno.3,colnames(transmission.probs) == "13" & transmitted.allele == 1] <- 1-D["3","13"]
-	transmission.probs[male.geno.3,colnames(transmission.probs) =="13" & transmitted.allele == 3] <- D["3","13"]
+	
+	transmission.probs[male.heterozy,colnames(transmission.probs) == "13" & transmitted.allele == 1] <- 1-D["3.het","13"]
+	transmission.probs[male.heterozy,colnames(transmission.probs) =="13" & transmitted.allele == 3] <- D["3.het","13"]
 
-	transmission.probs[male.geno.3,colnames(transmission.probs) == "12" & transmitted.allele == 1] <- 1-D["3","12"]
-	transmission.probs[male.geno.3,colnames(transmission.probs) =="12" & transmitted.allele == 2] <- D["3","12"]
+	transmission.probs[male.heterozy,colnames(transmission.probs) == "12" & transmitted.allele == 1] <- 1-D["3.het","12"]
+	transmission.probs[male.heterozy,colnames(transmission.probs) =="12" & transmitted.allele == 2] <- D["3.het","12"]
 
-	transmission.probs[male.geno.3,colnames(transmission.probs) == "23" & transmitted.allele == 2] <- 1-D["3","23"]
-	transmission.probs[male.geno.3,colnames(transmission.probs) =="23" & transmitted.allele == 3] <- D["3","23"]
-	recover()
+	transmission.probs[male.heterozy,colnames(transmission.probs) == "23" & transmitted.allele == 2] <- 1-D["3.het","23"]
[…]

### 2013-09-20 writeup_and_notes/male_dependent_drive.tex
Yaniv comments through discussion on friday sept 20

@@
 These models show that sperm modifiers of female meiotic drive are
-unlikely to create a sustained conflict, making it unlikely that
-female meiosis will evolve to resist them.
-In fact, the interests of sperm and egg  genomes' are often aligned, as are both invested in the fate of the resultant zygote (as was speculated for the In locus \citep{Pomiankowski1993}).
+unlikely to create a sustained conflict, and therefore 
+female meiosis will have little incentive to evolve resistance to them.
+In fact, the interests of sperm and egg  genomes' are often aligned, as both are invested in the fate of the resultant zygote (as was speculated for the In locus \citep{Pomiankowski1993}).
 Therefore, there is little selective benefit to females in preventing sperm to influence female meioses,
@@
 	positive frequency dependent behavior (Inset in Figure \ref{Invasion_space}).  
-Unlike standard meiotic drive systems, a homozygous cost, cannot sustain polymorphism of a green-bearded driver 
-	because it already overcame this cost just to enter the population.
-The lack of a balanced polymorphism precludes the evolution of an allele that suppresses this form of meiotic drive.
+%Unlike standard meiotic drive systems, a homozygous cost cannot sustain polymorphism of a green-bearded driver 
+%	because it already overcame this cost just to enter the population.
+This lack of a balanced polymorphism precludes the evolution of an allele that suppresses this form of meiotic drive.
 
@@
 %YB VERSION
-Inclusion of a heterozygous fitness cost (e.g.
-additive fitness cost) further constrains the evolution of a sperm-dependent driver.
-In contrast to a  standard female  driver which can invade from low frequency so long as drive in heterozygotes outweighs the fitness cost they suffer (XXX); when rare our sperm dependent driver pays the heterozygous
-        fitness cost but only weakly drives because the right sperm
-        are rare. 
+Inclusion of a heterozygous fitness cost (e.g. additive fitness cost) further constrains the evolution of a sperm-dependent driver. 
+While a  standard female  driver invades from low frequency when the transmission advantage outweighs the fitness cost (XXX),
+	a rare sperm dependent driver pays a fitness cost, but drives only weakly because complementary sperm are uncommon. 
 This frequency independent fitness cost, and positive frequency dependent drive generates a bistable system. 
@@
 	because it pays the  heterozygous fitness cost  but drives ineffectively (see FigureS\ref{bistable_additive}). 
-Above this frequency, the allele increases deterministically because the commonness of sperm carrying this allele 
+Above this frequency, the allele increases deterministically because the commonness of complementary sperm 
 	make drive effective enough to overcome the associated fitness cost.
@@
 
-While an allele specifically active in sperm could plausibly influence the outcome of female meiosis, the limited levels of gene expression in sperm \citep[e.g.][]{Joseph2004}
-	suggest it may also be plausible to consider 
-sperm may influence female meiosis via expression of the fertilizing male's
+While an the allelic identity of sperm could plausibly influence the outcome of female meiosis, 
+	limited gene expression in sperm \citep[e.g.][]{Joseph2004}
+	suggest a model where sperm influence female meiosis via expression of the fertilizing male's
 	diploid genotype (perhaps due to proteins and RNAs packaged into the sperm).
-This complicates the dynamics of the system, as the dominance of
-  the genotype of the male effect on female drive must be considered as a parameter. 
-While under some circumstances the dominance of the drive effect in
-males can allow the 
-maintainance of the allele a polymorphism, the parameter range that
-permits polymorphism is very narrow (and quite sensitive to the
-dominance of the male effect, see Figure S\ref{Effect_of_dominance}).
+This adds an additional parameter to the system -- the dominance of the drive enhancer in heterozygous males, 
+	when this dominance is high, a narrow sliver of parameter space sustaining a polymorphism emerges (see Figure S\ref{Effect_of_dominance})
+ 	\yb{presumably because positive frequency dependence diminishes with the dominance of the male effect driver.}
+%complicates the dynamics of the system, as the dominance of
+ % the genotype of the male effect on female drive must be considered as a parameter. 
+%Some circumstances the dominance of the drive effect in
+%males can allow the  maintenance of the allele a polymorphism, the parameter range that
+%permits polymorphism is very narrow (and quite sensitive to the
+%dominance of the male effect, see Figure S\ref{Effect_of_dominance}).
 
@@
 
+\yb{********* DO WE NEED TO DO THE MALE GENOTYPE THING HERE TOO? *********}
+
 Above, we explored a model were a driver that drove in females when signaled by a complementary allele in sperm.  
@@
 Here, we examine an alternative scenario (and more realistic) in which
-initially a driver, with no sperm-depedency, is at drive-viability equilibrium (with two alleles
+initially a driver, with no sperm-dependency, is at drive-viability equilibrium (with two alleles
 $A$ and $B$ are the ancestral non-driver and driver alleles).
-Subsequently a sperm-acting modifier of female meotic drive arises tightly
+Subsequently a sperm-acting modifier of female meiotic drive arises tightly
         linked to this balanced locus (effectively creating a third
@@
 
-We will first consider the case where a sperm-acting facilitator of
-female meiotic drive arises very tightly linked to the $B$ background 
-creating the third allele $B^{+}$ (i.e. the non-recombining haplotype of a
+We consider a sperm-acting facilitator of
+	female meiotic drive tightly linked to the $B$ background, 
+	generating a third allele $B^{+}$ (i.e. the non-recombining haplotype of a
         driver allele and sperm-acting drive modifier). 
-This new allele which acts in sperm to increase the effectiveness of both
-the $B$ and  $B^{+}$ drive in $AB$ and $AB^{+}$ heterozygotes, see Figure \ref{Eggsperm_3_allele_cartoon} for a simple schematic of this drive model.  
-Naively this $B^{+}$ allele might be able to increase in frequency in the
-population as then it could be allowed to spread due to the additional
-drive it causes.
-That turns out to not be the case, 
-because the sperm-acting drive promoter (the novel $B^{+}$ haplotype) 
-is introduced when the ancestral driver ($B$) is at drive-selection balance, 
-	$B'$ will immediately suffer a homozygous fitness cost.  
+This new allele acts in sperm to increase the effectiveness of both
+	the $B$ and  $B^{+}$ drive in $AB$ and $AB^{+}$ heterozygotes, see Figure \ref{Eggsperm_3_allele_cartoon} 
+	for a simple schematic of this drive model.  
+Naively, the $B^{+}$ may spread by capitalizing on the additional drive it causes; hover,  
+%Naively the $B^{+}$ allele might be able to increase in frequency in the
+%	population as then it could be allowed to spread due to the additional
+%	drive it causes. 
+	this is not the case for a few simple reasons. 
+First, the sperm-acting drive promoter (the novel $B^{+}$ haplotype) 
+	arises when the ancestral driver ($B$) is at drive-selection balance, 
+	and therefore immediately suffers a homozygous fitness cost.  
 Worse yet, a novel $B^{+}$ haplotype most often helps 
-the $B$  allele drive ($B'$ sperm meeting $AB$ eggs), simply because $B$ is initially more common than $B^{+}$.
-As $B^{+}$ facilitates $B$ these interactions will frequently form 
-	$BB^{+}$ heterozygote that suffers from the full
-	effect fitness costs as $BB$.
-
-Sperm-acting self promoting alleles are therefore at a profound disadvantage
+	the $B$  allele drive ($B^+$ sperm meeting $AB$ eggs), because $B$ is initially more common than $B^{+}$. 
+As $B^{+}$ facilitates $B$ these interactions frequently form 
+	$BB^{+}$ heterozygotes that suffer the same fitness costs as $BB$. 
+Therefore, sperm-acting self promoting alleles experience a profound disadvantage
 	in this scenario, even more so than under the previous two allele model. 
@@
 
-While  sperm enhancement of a female drive cannot displace a polymorphic female driver, sperm based drive suppression - that is sperm-acting alleles that act to restore 
+While  sperm enhancement of a female drive cannot displace a polymorphic female driver, sperm based drive suppression - 
+	that is sperm-acting alleles that act to restore 
 	fairness of female meiosis can. 
 If these sperm-acting drive suppressors arise on
-	the $A$ background (creating a third allele $A^{-}$) 
+	the $A$ background (creating a third allele $A^{-}$) ;
 	or are unlinked to the drive locus they readily invade a population segregating
-for the drive system ($AB$, see Figure). The invasion of such allele
-will lower the frequency of the original drive system (perhaps to zero)
-and will spread to fixation if they do not carry strong fitness costs
+	for the drive system ($AB$, see Figure). 
+This allele lowers the frequency of the original drive system (perhaps to zero),
+	and spreads to fixation if it does not carry strong fitness costs
 	(Figure \ref{Eggsperm_3_allele_cartoon} and \ref{Trajectories_of_supressors}A). \\
+\yb{This result is unsurprising, as drive suppressors unlinked to drivers or in repulsion phase with them usually invade polymorphic drive systems (CITE).}
 
-Intriguingly these sperm disruptors of female meiotic drive can spread
-	even when they arise tightly linked to the original drive system (B), forming
-	an novel third allele $B^{-}$ as above (Figures \ref{Eggsperm_3_allele_cartoon} and \ref{Trajectories_of_supressors}B). 
-$B^{-}$ displaces the original drive system ($B$) because it drives when not costly, 
-	but by disrupting drive when transmitted to sperm it avoids low-fitness drive homozygotes ($BB^-$ and $B^-B^-$. 
-Surprising despite displacing the ancestral driver ($B$), self avoiding driver, $B^*$ it's equilibrium frequency is not always higher than that of $B$.
+More surprisingly, sperm disruptors of female meiotic drive can spread
+	even when tightly linked to the original drive system (B), forming
+	an novel third allele $B^{-}$ as above 
+	(Figures \ref{Eggsperm_3_allele_cartoon} and \ref{Trajectories_of_supressors}B). 
+$B^{-}$ displaces the original driver ($B$) by driving when not costly, 
+	but by disrupting drive when transmitted to sperm it avoids low-fitness drive homozygotes ($BB^-$ and $B^-B^-$). 
+Despite displacing the ancestral driver ($B$), self avoiding driver, $B^*$ it's equilibrium frequency is not always higher than that of $B$.
 

### 2013-09-20 writeup_and_notes/male_dependent_drive.tex
small changes to discussion

--- 2013-09-19 writeup_and_notes/male_dependent_drive.tex
@@
 
+XY systems don't care. ZW systems (males ZZ, females ZW). If there is
+sex ratio distortion how could this play out?
+
 Finally given a system of a balanced meiotic drive system which avoid
@@
 , such as that depicted in Figure XX and XX, female-based modifiers 
-can evolve that act to further facilitate the sperms action in
+can evolve that act to further facilitate the sperm's action in
 disrupting driver.
 
-This logic may not hold for sex chromosomes. In ZW systems 
  Male modification of recombination rates

### 2013-09-22 writeup_and_notes/male_dependent_drive.tex
draft of discussion

@@
 
-%Since all alleles within an individual rely on individual survival and reproduction for evolutionary success, most of life in a diploid eukaryotic genome is harmonious.
-%However, this harmony is incomplete, as in numerous cases an allele can benefit (in the short or long term) from taking advantage of its host.
 Despite the apparent unity of the organism, occasionally alleles can
@@
 \citep{Burt2006}, often by exploiting meiosis and gametogensis.
-%A clear opportunity for this conflict arises during gametogenesis where alternative alleles are in direct competition for representation in the next generation.
-Because only one of the four products of female meiosis is transmitted to the egg, female meiosis \yb{is particularly vulnerable to such exploitation}  \citep{Sandler1957,Pardo-ManuelDeVillena2001a}. 
-%Female meiosis is particularly unfair, only one of the four products of meiosis becomes an egg, while the other threes are disgarded into polar bodies. 
+Because only one of the four products of female meiosis is transmitted to the egg, female meiosis is particularly vulnerable to such exploitation \citep{Sandler1957,Pardo-ManuelDeVillena2001a}. 
 An allele that biases female meiosis in its favor (i.e.
 a meiotic driver), may increase in frequency even if this driver entails a pleiotropic fitness cost \citep{Prout1973}, generating a genetic conflict between the success of the driver and organismal fitness.
-%Alleles that can segregate into the egg more than half the time in heterozygotes, by exploiting asymmetries, 
-%can potentially increase in frequency in the population (i.e. experiencing true meiotic drive).
-%If such alleles have fitness consequences they are a source of genetic conflict, 
-%and can become balanced in the population if their host in homozygotes 
-%out weighs their ability to spread to spread through heterozygotes. 
 Meiotic drivers observed in nature (in both plants
@@
 and often generate on ongoing evolutionary escalation of drive suppressors and enhancers \citep{Dawe1996,Fishman2008}. 
-%Such balanced drive systems can cause the subsequent evolution of linked enhancers of drive
-%and of supressors of drive throughout the genome. 
-%Indeed, the known polymorphic female meiotic drive systems (OM, maize knobs, IN, mimulus, others?) habor
-%a great diversity of enhancers and supressors. 
 The threat of meiotic drive to organismal fitness is potentially so
@@
 
-%COOP VERSION
-%Under a recessive fitness cost model our sperm dependent driver
-%actually has far more restrictive conditions to invade the population,
-%than a standard meiotic driver (see Figure \ref{Invasion_space}).
-%That is because our allele which drives only when it is present in the
-%fertilizing sperm is guarenteed to form a homzygote. Thus, 
-%to even spread in the population when it is rare, our sperm-dependent
-%driver has to overcome the cost it suffers in the homozygote state.
-
-%YB VERSION
 When a deleterious recessive green-bearded driver does spread, 
 	it spends much time at low frequency, because when rare it is unlikely to meet a complimentary sperm. 
-However, once common, a green-bearded driver spreads rapidly to fixation due to its
+However, once relatively common, a green-bearded driver spreads rapidly to fixation due to its
 	positive frequency dependent behavior (Inset in Figure \ref{Invasion_space}).  
-%Unlike standard meiotic drive systems, a homozygous cost cannot sustain polymorphism of a green-bearded driver 
-%	because it already overcame this cost just to enter the population.
 This lack of a balanced polymorphism precludes the evolution of an allele that suppresses this form of meiotic drive.
-
-%COOP VERSION
-%When conditions are suitable for such alleles to spread into the
-%population, they spread slowly through low frequencies as they rarely
-%drive as they are unlikely to be fertilized by a sperm with the right
-%allele. Once they become established they spread rapidly to fixation due to their
-%positive frequency dependent behavior (Inset in Figure \ref{Invasion_space}). 
-%Unlike standard meiotic drive systems 
-%these alleles can not become balanced in the population  by homozygote
-%cost, as they have already over come this cost just to enter the population.
 
@@
 This frequency independent fitness cost, and positive frequency dependent drive generates a bistable system. 
-%Therefore, even very weak heterozygous fitness costs
-%	prevent the spread of a very rare green-bearded driver, unless its drive is very strong. 
-%With an additive cost, the dynamics of a green-bearded driver resembles that of an underdominant mutation (in that the system is bistable). 
 Below a cutoff frequency the green-bearded driver deterministically decreases in frequency 
@@
  	\yb{presumably because positive frequency dependence diminishes with the dominance of the male effect driver.}
-%complicates the dynamics of the system, as the dominance of
- % the genotype of the male effect on female drive must be considered as a parameter. 
-%Some circumstances the dominance of the drive effect in
-%males can allow the  maintenance of the allele a polymorphism, the parameter range that
-%permits polymorphism is very narrow (and quite sensitive to the
-%dominance of the male effect, see Figure S\ref{Effect_of_dominance}).
 
-Given the difficulty that green-bearded meiotic drivers have entering the population, the speed at
+\gc{Add short paragraph about sex chromosomes here? XY systems don't care. ZW systems (males ZZ, females ZW). If there is
+sex ratio distortion how could this play out?}
+
+Given the difficulty that self promiting meiotic drivers have entering the population, the speed at
 which they fix if they do,
@@
 
-XY systems don't care. ZW systems (males ZZ, females ZW). If there is
-sex ratio distortion how could this play out?
+Our results provide a plausible answer to why female oogenesis and meiosis has
+not evolved to prevent sperm alleles from hijacking . 
+Such self-promoting sperm alleles are usually prevented from 
+entering the population if they have fitness costs, and are not
+maintained as polymorphisms if they do. As such, these type of
+exploitation can not drive sustained genetic conflict and so it is unlikely
+that female oogenesis and meiosis will evolve to prevent their effect.
 
-Finally given a system of a balanced meiotic drive system which avoid
-its own ill effects through a sperm-based mechanism
-, such as that depicted in Figure XX and XX, female-based modifiers 
-can evolve that act to further facilitate the sperm's action in
-disrupting driver.
+Gametogenesis and meiosis are common times for genetic conflict. This
+is a consequence of the fact that as the set of alleles that consitute
+an individual are in the process of parting ways, 
+with only a subset of them being assured transmission.
+In constrast, the perspective of the fertilizing sperm's genome is different, it has to live with the consequences of any distortion it causes 
+to female meiosis through the entire next life cycle of the organism. 
+It is therefore in the interest of the fertilizing sperm's genome to
+make a high fitness zygote even if that means alleles in sperm
+disrupting other copies of themselves from gaining a 
+transmission advantage through female meiosis.
+
+The interests of fertilizing sperm and female are quite well aligned during
+syngamy. As noted above there are even opportunities for coevolution
+between sperm-based supressors of female meiotic drive and the female genome.
+ Thus it seems plausible that aspects of oogenesis may actively 
+evolve to facilitate the influence of sperm on the outcome of female meiosis.
+In light of that it is interesting to consider whether the phylogenetic variation
+in the stage of oogenesis when fertilization takes place reflects not
+only species' different reproductive ecologies, but also different
+selection pressures shaping the interaction of sperm alleles with
+female meioses.
+
+\gc{Other pointsf}\\
 
@@
 
-Discussion of general conclusion that females have little reason to evolve supression mechanisms to prevent sperm influence on meiosis. 
-General logic that sperm genome has to live in a zygote with consequences of its effect on female meiosis, so
-it can not generate too dire a consequence.
 I THINK THERE IS A MORE SUBTLE POINT . The sperm has special knowledge

### 2013-09-22 scripts/invasion_analysis.R

@@ -178,7 +178,7 @@ if(FALSE){
 	
 	save(file=paste(directory,"bistable_invasion_grid.Robj",sep=""),my.x.bistable,my.s)
 }
-
+load(file=paste(directory,"bistable_invasion_grid.Robj",sep=""))
 plot(range(d.range),c(10^(-6),1),type="n",log="y",xlab="d",ylab="x bistable cutoff freq.")
 for(i in 1:length(my.s)){
 	lines(d.range,my.x.bistable[i,],col=my.cols[i],lwd=2)
@@ -187,7 +187,7 @@ for(i in 1:length(my.s)){
 
 legend(x="bottomleft",legend=paste("s=",my.s),col=my.cols,lty=1,lwd=2)
 
-dev.copy2eps(file=paste(directory,"bistable_x_vs_d_additive_s.eps",sep="")
+dev.copy2eps(file=paste(directory,"bistable_x_vs_d_additive_s.eps",sep=""))
 #dev.off()
 
 ##############Phase diagram figure

### 2013-09-23 writeup_and_notes/male_dependent_drive.tex
Yaniv comments on discussion on sat sept 22

@@
 
-\gc{Add short paragraph about sex chromosomes here? XY systems don't care. ZW systems (males ZZ, females ZW). If there is
-sex ratio distortion how could this play out?}
+%\gc{Add short paragraph about sex chromosomes here? XY systems don't care. ZW systems (males ZZ, females ZW). If there is
+%sex ratio distortion how could this play out?}
 
-Given the difficulty that self promiting meiotic drivers have entering the population, the speed at
-which they fix if they do,
-and the narrow parameter range permitting balanced polymorphisms at such loci, 
-it seems very unlikely that such alleles
-could drive the evolution of female suppressors of sperm-enabled
+Given the difficulty that self-promoting meiotic drivers have entering the population, the speed at
+which they fix if they do, and the narrow parameter range permitting balanced polymorphisms at such loci,  
+it seems very unlikely that such alleles could drive the evolution of female suppressors of sperm-enabled
 female meiotic drive.
@@
 	and spreads to fixation if it does not carry strong fitness costs
-	(Figure \ref{Eggsperm_3_allele_cartoon} and \ref{Trajectories_of_supressors}A). \\
+	(Figure \ref{Eggsperm_3_allele_cartoon} and \ref{Trajectories_of_supressors}A). 
 \yb{This result is unsurprising, as drive suppressors unlinked to drivers or in repulsion phase with them usually invade polymorphic drive systems (CITE).}
@@
 	but by disrupting drive when transmitted to sperm it avoids low-fitness drive homozygotes ($BB^-$ and $B^-B^-$). 
-Despite displacing the ancestral driver ($B$), self avoiding driver, $B^*$ it's equilibrium frequency is not always higher than that of $B$.
+Despite displacing the ancestral driver ($B$), self avoiding driver, $B^*$ it's equilibrium frequency is not always higher than that of $B$ \yb{(point to supp?)}.
 
@@
 
-Our results provide a plausible answer to why female oogenesis and meiosis has
-not evolved to prevent sperm alleles from hijacking . 
-Such self-promoting sperm alleles are usually prevented from 
-entering the population if they have fitness costs, and are not
-maintained as polymorphisms if they do. As such, these type of
-exploitation can not drive sustained genetic conflict and so it is unlikely
+%Our results provide a plausible answer to why female oogenesis and meiosis has
+%not evolved to prevent sperm alleles from hijacking . 
+%Such self-promoting sperm alleles are usually prevented from 
+%entering the population if they have fitness costs, and are not
+%maintained as polymorphisms if they do. As such, these type of
+%exploitation can not drive sustained genetic conflict and so it is unlikely
+%that female oogenesis and meiosis will evolve to prevent their effect.
+
+Sexual reproduction is a high-stakes event that determins which chromosomes are transmitted to the next generation.  
+As a consequence of this intense competition for inclusion in the next generation, alleles that  gain a transmission advantage during reproduction 
+	can succeed evolutionarily, even if they have bad effects on the population, 
+	the mother whose offspring they will inhabbit, 
+	and even the organism in which they reside. 
+This potential generates numerous conflicts  \cite{genesInConflict} including sexual conflicts between mates \cite{sexualConflict}, 
+	as well as conflicts between alleles that are overtransmitted in meiosis and the organisms they injure while doing so.
+
+It seems that allowing sperm to influence the outcome of meiosis would generate a confluence of these potential conflicts -- 
+	sperm could actually assist an allele that distorts female meiosis.
+However, this is not the case.
+We find that an allele which when in sperm help distort female meiosis in its favor, %YB [too awkward / unclear?] 
+	can rarely spread through a population if it bears any cost. 
+Additionally, when this self-promoting driver can spread, it cannot be maintained as a protected polymorphism, and due to its positive frequency dependence 
+	it spends very little time at intermediate frequency.
+As such, these type of exploitation cannot drive sustained genetic conflict and so it is unlikely
 	that female oogenesis and meiosis will evolve to prevent their effect.  
+This likely explains why females can delay the completion of meiosis until after fertilization 
+	without exploitation by collaborations by female drivers and collaborative sperm.
+%Male modification of recombination rates. \yb{trying to think through this \dots{}}
 
-Gametogenesis and meiosis are common times for genetic conflict. This
-is a consequence of the fact that as the set of alleles that consitute
-an individual are in the process of parting ways, 
-with only a subset of them being assured transmission.
-In constrast, the perspective of the fertilizing sperm's genome is different, it has to live with the consequences of any distortion it causes 
-to female meiosis through the entire next life cycle of the organism. 
-It is therefore in the interest of the fertilizing sperm's genome to
-make a high fitness zygote even if that means alleles in sperm
-disrupting other copies of themselves from gaining a 
-transmission advantage through female meiosis.
+Why is it that an allele that biases female meiosis in its favor can generate a genetic conflict, but an allele in sperm that assists this female driver cannot? 
+So long as the transmission advantage of female meiotic drive out way the organismal fitness cost to heterozygotes the female driver can spread when rare, and it increases in 		
+	frequency until the fitness cost experienced when homozygous balances the transmission advantage.
+By contrast a sperm promotor of female drive is only effective when matched with a heterozgyote female -- meaning that when rare this allele rarely enhances female drive. 
+Even worse, when it does so it will find itself in a low fitness drive homozygote. 
+Not only are drive-promoting sperm alleles unable to create a sustained genetic conflict, 
+	but alleles in sperm with the opposite effect - that is those that prevent their own drive through female meiosis do.
+This is because such drive suppressing alleles reduced their chances of forming  low fitness homozygotes.
+In fact, an allele in sperm which causes the alternate allele to drive can be favored (FIG), a counterintuitive result 
+	consistent with the tenuous evidence that the non-driving $ln^-$ allele is over transmitted when fertilized by $ln^+$ sperm \cite{agulnick}.
 
-The interests of fertilizing sperm and female are quite well aligned during
-syngamy. As noted above there are even opportunities for coevolution
-between sperm-based supressors of female meiotic drive and the female genome.
- Thus it seems plausible that aspects of oogenesis may actively 
-evolve to facilitate the influence of sperm on the outcome of female meiosis.
-In light of that it is interesting to consider whether the phylogenetic variation
-in the stage of oogenesis when fertilization takes place reflects not
-only species' different reproductive ecologies, but also different
-selection pressures shaping the interaction of sperm alleles with
-female meioses.
+\yb{DOES THIS GO ANYWHERE? }
+ We speculate that other effects of sperm on female meiosis may also be tolerated by females as a way of ensuring meiotic fairness, 
+ 	for example in Drosophila and other taxa, sperm accessory gland proteins (ACPs) increase the female recombination rate \cite{SOMEthING}.
+Presumably this is a consequence of the resolution of double strand breaks initiated by the stressful/toxic composition of ACPs. 
+However, by increasing the recombination rate sperm may decrease the effectiveness of female meiotic drivers active during MI, 
+	pleiotropically enhancing female fitness in a way (while decreasing it in many others, \cite{Partridge and Hurst})
 
+%Gametogenesis and meiosis are common times for genetic conflict. 
+%This is a consequence of the fact that as the set of alleles that consitute
+%	an individual are in the process of parting ways, 
+%	with only a subset of them being assured transmission.
+%In constrast, the perspective of the fertilizing sperm's genome is different, it has to live with the consequences of any distortion it causes 
+%to female meiosis through the entire next life cycle of the organism. 
+%It is therefore in the interest of the fertilizing sperm's genome to
+%make a high fitness zygote even if that means alleles in sperm
+%disrupting other copies of themselves from gaining a 
+%transmission advantage through female meiosis.
+
+Therefore, the interests of fertilizing sperm and female are quite well aligned during syngamy -- producing a viable and potentially fit offspring. 
+%As noted above there are even opportunities for coevolution between sperm-based suppressors of female meiotic drive and the female genome.
+Thus it seems plausible that aspects of oogenesis may actively evolve to facilitate the influence of sperm on ensuring a fair  outcome of female meiosis which generates fit offspring.
+In light of that it is interesting to consider whether the phylogenetic variation in the stage of oogenesis when fertilization takes place reflects not
+	only species' different reproductive ecologies, but also different selection pressures shaping the interaction of sperm alleles with female meioses. 
+\\
+\\
+\\
+\\
 \gc{Other pointsf}\\
+\\
+\\
+\gc{Add short paragraph about sex chromosomes here? XY systems don't care. ZW systems (males ZZ, females ZW). If there is
+sex ratio distortion how could this play out?}
 
- Male modification of recombination rates
 POssibility that this could happen in plants if pollen emit signals to ``egg''
 Discussion of OM and IN.
-
-I THINK THERE IS A MORE SUBTLE POINT . The sperm has special knowledge
-that if it allows drive it will end up in the low fitness homozygote.
 

### 2013-12-07 writeup_and_notes/male_dependent_drive.tex
committing files

@@
 
-Sexual reproduction is a high-stakes event that determins which chromosomes are transmitted to the next generation.  
-As a consequence of this intense competition for inclusion in the next generation, alleles that  gain a transmission advantage during reproduction 
-	can succeed evolutionarily, even if they have bad effects on the population, 
-	the mother whose offspring they will inhabbit, 
+Sexual reproduction is a high-stakes event that determines which
+determines what gets transmitted to the next generation.  
+As a consequence of this intense competition, alleles that gain a transmission advantage during reproduction 
+	can succeed evolutionarily, even if they lower the 
+        fitness of the population, 
+	the mother whose offspring they will inhabit, 
 	and even the organism in which they reside. 
@@
 However, this is not the case.
-We find that an allele which when in sperm help distort female meiosis in its favor, %YB [too awkward / unclear?] 
+We find that an allele which acts through sperm to distort female meiosis in its favor, %YB [too awkward / unclear?] 
 	can rarely spread through a population if it bears any cost. 
-Additionally, when this self-promoting driver can spread, it cannot be maintained as a protected polymorphism, and due to its positive frequency dependence 
+Additionally, when this self-promoting driver can spread, it can only
+rarely 
+be maintained as a protected polymorphism, and due to its positive frequency dependence 
 	it spends very little time at intermediate frequency.
@@
 This likely explains why females can delay the completion of meiosis until after fertilization 
-	without exploitation by collaborations by female drivers and collaborative sperm.
+	without risking exploitation by collaborations by female
+        drivers and sperm alleles.
 %Male modification of recombination rates. \yb{trying to think through this \dots{}}
@@
 In fact, an allele in sperm which causes the alternate allele to drive can be favored (FIG), a counterintuitive result 
-	consistent with the tenuous evidence that the non-driving $ln^-$ allele is over transmitted when fertilized by $ln^+$ sperm \cite{agulnick}.
+	consistent with the tenuous evidence that the non-driving
+        $ln^-$ allele is over transmitted when fertilized by $ln^+$
+        sperm \cite{agulnick, Pomiankowski1993}.
 
 \yb{DOES THIS GO ANYWHERE? }
+
+Males can potentially influence the outcome of meiosis even before
+fertilization. 
+For example in Drosophila and other taxa, sperm accessory
+        gland proteins (ACPs) increase the female recombination rate
+        \cite{SOMEthING}, and there is some evidence that there is
+        variance across male genotypes in there effect on female
+        recombination rate \cite{Stevison2012}.
+Presumably this is a pleiotropic consequence of the resolution of double strand breaks initiated by the stressful/toxic composition of ACPs. 
+However, a pleiotropic effect of this is to increase the recombination
+rate which can decrease the effectiveness of female meiotic drivers
+active during MI.
+
+	pleiotropically enhancing female fitness in a way (while
+        decreasing it in many others, \cite{Partridge and Hurst}).
+
  We speculate that other effects of sperm on female meiosis may also be tolerated by females as a way of ensuring meiotic fairness,
- 	for example in Drosophila and other taxa, sperm accessory gland proteins (ACPs) increase the female recombination rate \cite{SOMEthING}.
-Presumably this is a consequence of the resolution of double strand breaks initiated by the stressful/toxic composition of ACPs. 
-However, by increasing the recombination rate sperm may decrease the effectiveness of female meiotic drivers active during MI, 
-	pleiotropically enhancing female fitness in a way (while decreasing it in many others, \cite{Partridge and Hurst})
 

### 2013-12-08 writeup_and_notes/male_dependent_drive.tex
modified text

@@
 %Why do animals allow male control of female meiosis?
-
 %Why does female meiosis allow the opportunity for exploitation by selfish sperm?
-
 %sperm and egg genomes both favour the peaceful resolution of female meiotic drive
@@
 
-\title{Why does female meiosis allow the opportunity for exploitation
-  by selfish sperm? BETTER TITLE?}
+%\title{Why does female meiosis allow the opportunity for exploitation by selfish sperm? BETTER TITLE?}
+\title{Sperm do not evolve to collaborate in  female meiotic drive}
+
 \author{Yaniv Brandvain \\ email: ybrandvain@gmail.com  \and Graham Coop \\ email: gmcoop@ucdavis.edu }
@@
 
-Despite the apparent unity of the organism, occasionally alleles can
+Despite the apparent unity of the organism, alleles can
 gain an evolutionary advantage at a cost to  individual fitness
@@
 \citep{Prout1973}, 
-and often generate on ongoing evolutionary escalation of drive suppressors and enhancers \citep{Dawe1996,Fishman2008}. 
+and often generate an ongoing evolutionary escalation of drive suppressors and enhancers \citep{Dawe1996,Fishman2008}. 
 The threat of meiotic drive to organismal fitness is potentially so
@@
 it is potentially open to sabotage by a virtual stranger -- a haploid sperm genome.
-That is, in many animal species, the completion of female meiosis requires fertilization of the egg \citep{Masui_book}, 
-and there is ample opportunity for interaction between the sperm and female meiotic machinery.
-If, for example, an allele in sperm could facilitate meiotic drive by a genetically equivalent allele in a 
-heteromorphic dyad, such an allele could presumably bias meiosis in its favor and rapidly spread through the population.
+That is, in many animal species, female meiosis is competed only after fertilization \citep{Masui_book}, 
+creating ample opportunity for interaction between the sperm and female meiotic machinery. 
+For example, an allele in sperm that facilitates meiotic drive by a genetically equivalent allele in female meiosis %(a  heteromorphic dyad), 
+     could presumably rapidly spread through the population.
 At first sight, it seems as although female meiosis is primed to be exploited by selfish sperm systems.  
@@
 It is certainly not the case that animals are mechanistically incapable of circumventing this requirement.
-There is considerable variation in which stage of meiosis requires fertilization, and 
-a number of animal clades (should we try and lower bound how many transitions?) have evolved
-to allow the completion of female meiosis upon ovulation \citep[see
-Table 1 of ][]{Masui_book}.
+The stage at which female meiosis requires fertilization varies widely across taxa, and in a few animal taxa, 
+	female meiosis is complete upon ovulation (that is fertilization is not required for female meiosis \citep[see
+	Table 1 of ][]{Masui_book}).
 
@@
 Specifically, the two best characterized female meiotic drive systems in mouse (In and Om), both operate by distorting the second meiotic division, 
-and in both systems the outcome of female meiosis depends the genotype of the fertilizing sperm \citep{Agulnik1993,Wu2005}; 
-However,  we acknowledge the difficulty in ruling out early mortality
-as an alternative explanation of the 
-specifics of these putative drivers \citep[Give page numbers to
-discussion of this in][]{Burt2006}. \gc{should be careful here as the
-  Om locus is pretty convincingly a meiotic driver, doubts have just
-  been raised about whether IN is. Whether the sperm dep nature of
-  either is robust to this is I think more at question.}
+and in both systems the outcome of female meiosis depends the genotype of the fertilizing sperm \citep{Agulnik1993,Wu2005}. 
+While such genetic evidence is suggestive, we note the difficulty in ruling out the alternative hypothesis of early selection    \citep[Give page numbers to discussion of this in][]{Burt2006}.
+
+%However,  we acknowledge the difficulty in ruling out early mortality
+%as an alternative explanation of the 
+%specifics of these putative drivers \citep[Give page numbers to
+%discussion of this in][]{Burt2006}. \gc{should be careful here as the
+ % Om locus is pretty convincingly a meiotic driver, doubts have just
+ % been raised about whether IN is. Whether the sperm dep nature of
+ % either is robust to this is I think more at question.}
 
@@
 itself.}
+%%%%%%%%%%%%%%%%%%%%%%%%%
+% PLEIOTROPY MODEL [self-promoting]
+%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%yb version
-In the standard model of female meiotic drive, the driver is one of two alleles, and in heterozygotes is transmitted to the egg with probability  $d > \frac{1}{2}$, regardless of sperm genotype. 
+In the standard single-locus, biallelic model of female meiotic drive, the driving allele is transmitted to the egg  in heterozygotes with probability  $d > \frac{1}{2}$, regardless of sperm genotype.
 To depict a case of a self-promoting (or `green-bearded') meiotic
@@
 We then identify the conditions under which this self propagating driver can spread, and evaluate whether a driver of this form could generate a sustained conflict favoring the evolution of suppressors.
+\yb{EQUATIONS A1-A4}
 
 We first consider a driving allele neutral in heterozygotes, 
-	but deleterious in homozygotes (whose relative fitness is $1-s$).  
+	but deleterious in homozygotes (whose relative fitness is $1-s$) \yb{(EQUATIONS A3-A4)}.  
 In this model, a standard meiotic driver can always invade because 
@@
 no fitness cost as a rare deleterious recessive, 
-	our green-bearded driver creates homozygotes
+	a self-promoting driver creates homozygotes
         when it drives, and therefore there is a drive-associated fitness cost even when rare. 
-Our sperm-dependent driver must overcome this homozygous fitness cost simply to spread when rare.
+Our self-promoting driver must overcome this homozygous fitness cost simply to spread when rare.
 Therefore with a recessive fitness cost,
-	the conditions allowing the invasion of our green-bearded driver 
+	the conditions allowing the invasion of a self-promoting driver
  	are far more restrictive than those for a standard meiotic driver (see Figure \ref{Invasion_space}).
 
-When a deleterious recessive green-bearded driver does spread, 
+When a deleterious recessive self-promoting driver does spread, 
 	it spends much time at low frequency, because when rare it is unlikely to meet a complimentary sperm. 
-However, once relatively common, a green-bearded driver spreads rapidly to fixation due to its
+However, once relatively common, a self-promoting driver spreads rapidly to fixation due to its
 	positive frequency dependent behavior (Inset in Figure \ref{Invasion_space}).  
 This lack of a balanced polymorphism precludes the evolution of an allele that suppresses this form of meiotic drive.
+\yb{POINT TO SOME MATHS OR FIGURE}
 
-%YB VERSION
-Inclusion of a heterozygous fitness cost (e.g. additive fitness cost) further constrains the evolution of a sperm-dependent driver. 
+Inclusion of a heterozygous fitness cost (e.g. additive fitness cost) further constrains the evolution of a self-promoting driver. 
 While a  standard female  driver invades from low frequency when the transmission advantage outweighs the fitness cost (XXX),
-	a rare sperm dependent driver pays a fitness cost, but drives only weakly because complementary sperm are uncommon. 
+	a rare self-promoter pays a fitness cost, but drives only weakly because complementary sperm are uncommon. 
 This frequency independent fitness cost, and positive frequency dependent drive generates a bistable system. 
-Below a cutoff frequency the green-bearded driver deterministically decreases in frequency 
+Below a cutoff frequency the self-promoter driver deterministically decreases in frequency 
 	because it pays the  heterozygous fitness cost  but drives ineffectively (see FigureS\ref{bistable_additive}). 
@@
 	make drive effective enough to overcome the associated fitness cost.
-This bistability prevents sperm-dependent drivers from invading 	
+This bistability prevents self-promoting drivers from invading 	
 	reasonably sized populations, and assures that if they do invade they will rapidly fix.
+\yb{Our model therefore predicts that self-promoting drivers will not be observed as polymorphisms in natural populations. }
 
-While an the allelic identity of sperm could plausibly influence the outcome of female meiosis, 
+While the allelic identity of sperm could plausibly influence the outcome of female meiosis, 
 	limited gene expression in sperm \citep[e.g.][]{Joseph2004}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%
+% TWO LOCUS MODEL [modify (general) / facilitate (helps ) / prevent (stops)] 
+%%%%%%%%%%%%%%%%%%%%%%%%%
 \yb{********* DO WE NEED TO DO THE MALE GENOTYPE THING HERE TOO? *********}
@@
 Above, we explored a model were a driver that drove in females when signaled by a complementary allele in sperm.  
-This complicated pleiotropic effect, while plausible is somewhat unlikely. 
-Here, we examine an alternative scenario (and more realistic) in which
-initially a driver, with no sperm-dependency, is at drive-viability equilibrium (with two alleles
+We complement this single locus model  of pleiotropy with an alternative model of two linked loci - one female driver and the other a sperm allele which modifies the effect of drive upon fertilization.
+%Here, we examine an alternative scenario (and more realistic) in which
+In this model initially a female meiotic driver, with no sperm-dependency, reaches drive-viability equilibrium (with two alleles
 $A$ and $B$ are the ancestral non-driver and driver alleles).
-Subsequently a sperm-acting modifier of female meiotic drive arises tightly
+Subsequently, a sperm-acting modifier of female meiotic drive arises tightly
         linked to this balanced locus (effectively creating a third
@@
 
-We consider a sperm-acting facilitator of
+%%%sperm-acting facilitator
+We first consider a sperm-acting facilitator of
 	female meiotic drive tightly linked to the $B$ background, 
@@
 Naively, the $B^{+}$ may spread by capitalizing on the additional drive it causes; hover,  
-%Naively the $B^{+}$ allele might be able to increase in frequency in the
-%	population as then it could be allowed to spread due to the additional
-%	drive it causes. 
 	this is not the case for a few simple reasons. 
-First, the sperm-acting drive promoter (the novel $B^{+}$ haplotype) 
+First, the sperm-acting drive facilitator (the novel $B^{+}$ haplotype) 
 	arises when the ancestral driver ($B$) is at drive-selection balance, 
@@
 	$BB^{+}$ heterozygotes that suffer the same fitness costs as $BB$. 
-Therefore, sperm-acting self promoting alleles experience a profound disadvantage
+Therefore, sperm-acting drive facilitator alleles experience a profound disadvantage 
 	in this scenario, even more so than under the previous two allele model. 
 We have found no parameter range of this
-	three allele system that allows the selfish sperm-acting allele $B^{+}$ to
+	three allele system that allows the sperm-acting drive facilitator $B^{+}$ to
 	invade the population. \yb{MATHS?}
 
+%%%sperm-acting drive suppressors
 While  sperm enhancement of a female drive cannot displace a polymorphic female driver, sperm based drive suppression - 
-	that is sperm-acting alleles that act to restore 
-	fairness of female meiosis can. 
+	that is sperm-acting alleles that act to restore  fairness of female meiosis can. 
 If these sperm-acting drive suppressors arise on
-	the $A$ background (creating a third allele $A^{-}$) ;
+	the non-driving $A$ background (creating a third allele $A^{-}$) ;
 	or are unlinked to the drive locus they readily invade a population segregating
@@
 	(Figure \ref{Eggsperm_3_allele_cartoon} and \ref{Trajectories_of_supressors}A). 
-\yb{This result is unsurprising, as drive suppressors unlinked to drivers or in repulsion phase with them usually invade polymorphic drive systems (CITE).}
+\yb{This result is consistent with previous work showing that drive suppressors unlinked to, or in repulsion phase with drivers usually invade polymorphic drive systems \citep[e.g.
+][]{Brandvain2012}.}
 
-More surprisingly, sperm disruptors of female meiotic drive can spread
+More surprisingly, sperm-acting suppressors of female drive can spread
 	even when tightly linked to the original drive system (B), forming
@@
 
-Sexual reproduction is a high-stakes event that determines which
-determines what gets transmitted to the next generation.  
+Sexual reproduction is a high-stakes event that determines what gets transmitted to the next generation.  
 As a consequence of this intense competition, alleles that gain a transmission advantage during reproduction 
@@
 	and even the organism in which they reside. 
-This potential generates numerous conflicts  \cite{genesInConflict} including sexual conflicts between mates \cite{sexualConflict}, 
+This potential generates numerous conflicts  \cite{Burt2006} including sexual conflicts between mates \cite{Arnqvist2005}, 
 	as well as conflicts between alleles that are overtransmitted in meiosis and the organisms they injure while doing so. 
+Such conflicts and their resolution likely play a major role in the structure and evolution of many basic biological processes \citep{Rice2013}.
 
@@
 However, this is not the case.
-We find that an allele which acts through sperm to distort female meiosis in its favor, %YB [too awkward / unclear?] 
+We find that an allele which acts through sperm to distort female meiosis in its favor %YB [too awkward / unclear?] 
 	can rarely spread through a population if it bears any cost. 
-Additionally, when this self-promoting driver can spread, it can only
-rarely 
+Additionally, when this self-promoting driver can spread, it can only rarely 
 	be maintained as a protected polymorphism, and due to its positive frequency dependence 
 	it spends very little time at intermediate frequency.
-As such, these type of exploitation cannot drive sustained genetic conflict and so it is unlikely
+As such, this type of exploitation cannot generate a sustained genetic conflict and so it is unlikely
 	that female oogenesis and meiosis will evolve to prevent their effect.  
@@
 Why is it that an allele that biases female meiosis in its favor can generate a genetic conflict, but an allele in sperm that assists this female driver cannot? 
-So long as the transmission advantage of female meiotic drive out way the organismal fitness cost to heterozygotes the female driver can spread when rare, and it increases in 		
+So long as the transmission advantage of female meiotic drive outweigh the organismal fitness cost to heterozygotes the female driver can spread when rare, and it increases in 		
 	frequency until the fitness cost experienced when homozygous balances the transmission advantage.
 By contrast a sperm promotor of female drive is only effective when matched with a heterozgyote female -- meaning that when rare this allele rarely enhances female drive. 
-Even worse, when it does so it will find itself in a low fitness drive homozygote. 
+Even worse, when it does so it will preferentially find itself in a low fitness drive homozygote. 
 Not only are drive-promoting sperm alleles unable to create a sustained genetic conflict, 
-	but alleles in sperm with the opposite effect - that is those that prevent their own drive through female meiosis do.
+	but alleles in sperm with the opposite effect - that is those that prevent their own drive through female meiosis do maintain a polymorphisms and 
+	provide evolution  with time and opportunity to further minimize drive.
 This is because such drive suppressing alleles reduced their chances of forming  low fitness homozygotes.
@@
 
-Males can potentially influence the outcome of meiosis even before
-fertilization. 
+Males can potentially influence the outcome of meiosis even before fertilization. 
 For example in Drosophila and other taxa, sperm accessory
@@
 Presumably this is a pleiotropic consequence of the resolution of double strand breaks initiated by the stressful/toxic composition of ACPs. 
-However, a pleiotropic effect of this is to increase the recombination
-rate which can decrease the effectiveness of female meiotic drivers
-active during MI.
-
-	pleiotropically enhancing female fitness in a way (while
-        decreasing it in many others, \cite{Partridge and Hurst}).
-
- We speculate that other effects of sperm on female meiosis may also be tolerated by females as a way of ensuring meiotic fairness,
+However, a pleiotropic effect of this is to increase the recombination rate which can decrease the effectiveness of female meiotic drivers
+active during MI \citep{Brandvain}. 
+We speculate that other effects of sperm on female meiosis may also be tolerated by females as a way of ensuring meiotic fairness -- 
+	that is females evolve to allow male influence on female meiosis in cases in which the male influence is  egalitarian.
 

### 2013-12-10 writeup_and_notes/male_dependent_drive.tex
added eqn. to describe geno. freq evolution for simplest sperm dep. drive model

@@
 %\title{Why does female meiosis allow the opportunity for exploitation by selfish sperm? BETTER TITLE?}
-\title{Sperm do not evolve to collaborate in  female meiotic drive}
+\title{Co-operation between the sexes trumps selfish control of female
+  meiotic drive. Or 
+Sperm do not evolve to collaborate in female meiotic drive. Or Why does female meiosis allow the opportunity for exploitation by selfish sperm? }
 
@@
 in heterozygous females many potentially depend on sperm haplotype. 
-Specifically, the two best characterized female meiotic drive systems in mouse (In and Om), both operate by distorting the second meiotic division, 
+Specifically, two female meiotic drive systems in mouse (In and Om), both operate by distorting the second meiotic division, 
 and in both systems the outcome of female meiosis depends the genotype of the fertilizing sperm \citep{Agulnik1993,Wu2005}. 
-While such genetic evidence is suggestive, we note the difficulty in ruling out the alternative hypothesis of early selection    \citep[Give page numbers to discussion of this in][]{Burt2006}.
+While such genetic evidence is suggestive, we note the difficulty in
+ruling out the alternative hypothesis of early selection  \citep[pages
+52-54][]{Burt2006}.
 
@@
 
+\section*{Appendix Stuff}
+$\Venus$
+Genotype frequencies at birth $x_{AA},~x_{AB},~x_{BB}$ cost in
+homozygotes 
+\begin{equation}
+x'_{BB} \propto w_{BB} x_{BB}^{\Venus} (w_{BB} x_{BB}^{\Mars} +\frac{1}{2}
+  w_{AB} x_{AB}^{\Mars} ) + w_{AB} x_{AB}^{\Venus} (d w_{BB}
+  x_{BB}^{\Mars} +\frac{1}{2} d
+w_{AB} x_{AB}^{\Mars} )  
+\end{equation}
+
+\begin{equation}
+x'_{AA} \propto w_{AA} x_{AA}^{\Venus} (w_{AA} x_{AA}^{\Mars} + \frac{1}{2}
+   w_{AB} x_{AB}^{\Mars}) + w_{AB} x_{AB}^{\Venus}(\frac{1}{2} w_{AA} x_{AA}^{\Mars}+ \frac{1}{2} 
+  w_{AB} x_{AB}^{\Mars}) 
+\end{equation}
+
+\begin{eqnarray}
+x'_{AB} \propto  & w_{AA} x_{AA}^{\Venus}  ( w_{BB}  x_{BB}^{\Mars} + \frac{1}{2}
+  w_{AB}  x_{AB}^{\Mars} ) +w_{BB} x_{BB}^{\Venus} ( w_{AA}  x_{AA}^{\Mars} + \frac{1}{2}
+  w_{AB}  x_{AB}^{\Mars} )  \nonumber \\
+& + x_{AB}^{\Venus}(\frac{1}{2}  w_{AA}  x_{AA}^{\Mars}+\frac{1}{2} (1-d)
+ w_{AB}  x_{AB}^{\Mars} + \frac{1}{4}  w_{AB}
+ x_{AB}^{\Mars} + (1-d)  w_{BB}  x_{BB}^{\Mars} ) 
+\end{eqnarray}
+
 \bibliography{refs}

### 2014-02-04 writeup_and_notes/male_dependent_drive.tex
redid figures for paper

--- 2013-12-08 writeup_and_notes/male_dependent_drive.tex
@@
 gain an evolutionary advantage at a cost to  individual fitness
-\citep{Burt2006}, often by exploiting meiosis and gametogensis.
+\citep{Burt2006}, often by exploiting meiosis and gametogenesis.
 Because only one of the four products of female meiosis is transmitted to the egg, female meiosis is particularly vulnerable to such exploitation \citep{Sandler1957,Pardo-ManuelDeVillena2001a}. 
@@
 in heterozygous females many potentially depend on sperm haplotype. 
-Specifically, the two best characterized female meiotic drive systems in mouse (In and Om), both operate by distorting the second meiotic division, 
+Specifically, two female meiotic drive systems in mouse (In and Om), both operate by distorting the second meiotic division, 
 and in both systems the outcome of female meiosis depends the genotype of the fertilizing sperm \citep{Agulnik1993,Wu2005}. 
-While such genetic evidence is suggestive, we note the difficulty in ruling out the alternative hypothesis of early selection    \citep[Give page numbers to discussion of this in][]{Burt2006}.
+While such genetic evidence is suggestive, we note the difficulty in
+ruling out the alternative hypothesis of early selection on zygotes \citep[pages
+52-54][]{Burt2006}.
 
@@
 We first consider a driving allele neutral in heterozygotes, 
-	but deleterious in homozygotes (whose relative fitness is $1-s$) \yb{(EQUATIONS A3-A4)}.  
+	but deleterious in homozygotes (whose relative fitness is
+        $1-s$). This model is laid out in eqn. \ref{eqn.A1}-\ref{eqn.A3}.
 In this model, a standard meiotic driver can always invade because 
@@
 	positive frequency dependent behavior (Inset in Figure \ref{Invasion_space}).  
-This lack of a balanced polymorphism precludes the evolution of an allele that suppresses this form of meiotic drive.
+This lack of a balanced polymorphism ( Figure \ref{Invasion_space}) precludes the evolution of an
+allele that suppresses this form of meiotic drive in females.
 \yb{POINT TO SOME MATHS OR FIGURE}
@@
 	a rare self-promoter pays a fitness cost, but drives only weakly because complementary sperm are uncommon. 
-This frequency independent fitness cost, and positive frequency dependent drive generates a bistable system. 
+This frequency-independent fitness cost, and positive frequency dependent drive generates a bistable system. 
 Below a cutoff frequency the self-promoter driver deterministically decreases in frequency 
-	because it pays the  heterozygous fitness cost  but drives ineffectively (see FigureS\ref{bistable_additive}). 
+	because it pays the  heterozygous fitness cost  but drives
+        ineffectively (see Figure S\ref{bistable_additive}). 
 Above this frequency, the allele increases deterministically because the commonness of complementary sperm 
@@
 	reasonably sized populations, and assures that if they do invade they will rapidly fix.
-\yb{Our model therefore predicts that self-promoting drivers will not be observed as polymorphisms in natural populations. }
+Our model therefore predicts that self-promoting drivers will not be observed as polymorphisms in natural populations.
 
@@
 	when this dominance is high, a narrow sliver of parameter space sustaining a polymorphism emerges (see Figure S\ref{Effect_of_dominance})
- 	\yb{presumably because positive frequency dependence diminishes with the dominance of the male effect driver.}
+ presumably because positive frequency dependence diminishes with the dominance of the male effect driver.
 
@@
 \begin{figure}
-	\rotatebox{270}{\includegraphics[width = 0.8
-          \textwidth]{Figures/sperm_egg_cartoon1.ps}}
+	\rotatebox{90}{
+\includegraphics[width = 0.5
+          \textwidth]{Figures/Sperm_egg_cartoon_paper_Fig1.eps}}
 \caption{transmission probabilities for alleles through female
@@
 \includegraphics[width = 0.8 \textwidth]{Figures/invasion_space_recessive_driver.eps}
-\caption{Invasion analysis. Likely merge with Figure 1. \yb{PERHAPS COMPARE TO TRADITIONAL DRIVER IN THIS FIGURE} } \label{Invasion_space}
+\caption{Invasion analysis for a female meiotic drive allele with
+  recessive costs. First consider a simple female meiotic
+  drive system, where the allele is transmitted a fraction $d>1/2$ of
+  meioses in female heterozygotes, inspective
+  of the allele carried by the fertilizing sperm. Such an allele can
+  invade over all of the parameter space, and will be maintained as a
+  protected polymorphism in the white region, and can fix in the
+  population in all of the parameter space colored green,
+  and will fix  or orange. Now consider an allele that is fairly transmitted through
+  female meiosis when fertilized by a sperm carrying a different
+  allele, but a fraction $d>1/2$ when a heterozygote egg is fertilized
+  by a sperm carrying the allele (as depicted in Figure
+  \ref{Eggsperm_2_allele_cartoon}). Such an allele can only invade in
+  the region of parameter space colored orange, and it fixes over all
+  of this parameter region. 
+ In the inset we show trajectories of sperm-dependent female drive
+ alleles that can invade the population, each allele has $s =0.1$ against the
+ homozygotes and $d =0.6$, $0.7$, $0.8$ shown in yellow, blue, and red. 
+ \gc{Likely make this Figure 1B.} } \label{Invasion_space}
 \end{figure}
@@
 \begin{figure}
-	\rotatebox{270}{\includegraphics[width = 0.8
-          \textwidth]{Figures/sperm_egg_cartoon2.ps}}
-\caption{transmission probabilities righthand allele through female
+	\rotatebox{90}{\includegraphics[width = 0.8
+          \textwidth]{Figures/Sperm_egg_cartoon_paper_Fig2.eps}}
+\caption{Transmission probabilities righthand allele through female
   meiosis}  \label{Eggsperm_3_allele_cartoon}
@@
 
-\begin{figure}
-\includegraphics[width = 0.8 \textwidth]{Figures/trajectories_of_sperm_based_supressors.eps} 
-\caption{Trajectories of two sperm-based supressors of drive. Possibly
-merge this figure the 3 allele cartoon}  \label{Trajectories_of_supressors}
-\end{figure}
+% \begin{figure}
+% \includegraphics[width = 0.8 \textwidth]{Figures/trajectories_of_sperm_based_supressors.eps} 
+% \caption{Trajectories of two sperm-based supressors of drive. Possibly
+% merge this figure the 3 allele cartoon}  \label{Trajectories_of_supressors}
+% \end{figure}
 
@@
 Above, we explored a model were a driver that drove in females when signaled by a complementary allele in sperm.  
-We complement this single locus model  of pleiotropy with an alternative model of two linked loci - one female driver and the other a sperm allele which modifies the effect of drive upon fertilization.
+We complement this single locus model of an allele with effects in
+both male and with femalr, with  an alternative model of two linked loci - one female driver and the other a sperm allele which modifies the effect of drive upon fertilization. 
 %Here, we examine an alternative scenario (and more realistic) in which
@@
 	(Figure \ref{Eggsperm_3_allele_cartoon} and \ref{Trajectories_of_supressors}A). 
-\yb{This result is consistent with previous work showing that drive suppressors unlinked to, or in repulsion phase with drivers usually invade polymorphic drive systems \citep[e.g.
-][]{Brandvain2012}.}
+This result is consistent with previous work showing that drive suppressors unlinked to, or in repulsion phase with drivers usually invade polymorphic drive systems \citep[e.g. ][]{Brandvain2012}.
 
@@
 
-\gc{Do we mention the In and Om locus here or in the discussion as a
-	 potential e.g. Do we mention the fact that the frequency of these
-	 newly invade sperm dodging alleles can be lower than the original system?}\\
+%\gc{Do we mention the In and Om locus here or in the discussion as a
+%	 potential e.g. Do we mention the fact that the frequency of these
+%	 newly invade sperm dodging alleles can be lower than the original system?}\\
 
 \section*{Discussion}
-
-%Our results provide a plausible answer to why female oogenesis and meiosis has
-%not evolved to prevent sperm alleles from hijacking . 
-%Such self-promoting sperm alleles are usually prevented from 
-%entering the population if they have fitness costs, and are not
-%maintained as polymorphisms if they do. As such, these type of
-%exploitation can not drive sustained genetic conflict and so it is unlikely
-%that female oogenesis and meiosis will evolve to prevent their effect.
 
@@
 
-It seems that allowing sperm to influence the outcome of meiosis would generate a confluence of these potential conflicts -- 
+It seems that allowing sperm to influence the outcome of female meiosis would generate a confluence of these potential conflicts -- 
 	sperm could actually assist an allele that distorts female meiosis.
@@
 	provide evolution  with time and opportunity to further minimize drive.
-This is because such drive suppressing alleles reduced their chances of forming  low fitness homozygotes.
+This is because such drive suppressing alleles reduced their chances
+of forming low fitness homozygotes. 
 In fact, an allele in sperm which causes the alternate allele to drive can be favored (FIG), a counterintuitive result 
 	consistent with the tenuous evidence that the non-driving
-        $ln^-$ allele is over transmitted when fertilized by $ln^+$
+        $ln^-$ allele is over transmitted through female meiosis in $ln^-/ln^+$ heterozygote when fertilized by $ln^+$
         sperm \cite{agulnick, Pomiankowski1993}.   
+More generally both linked and unlinked alleles may often spread that act through sperm to
+reduce the opportunity of female meiotic drive in order to disable
+segregating drive systems, by for example reducing the mechanistic
+asymmetries that female drivers exploit.
 
-\yb{DOES THIS GO ANYWHERE? }
+\gc{Is there some nice way to work in green beard ideas}
 
-Males can potentially influence the outcome of meiosis even before fertilization. 
-For example in Drosophila and other taxa, sperm accessory
-        gland proteins (ACPs) increase the female recombination rate
-        \cite{SOMEthING}, and there is some evidence that there is
-        variance across male genotypes in there effect on female
-        recombination rate \cite{Stevison2012}.
-Presumably this is a pleiotropic consequence of the resolution of double strand breaks initiated by the stressful/toxic composition of ACPs. 
-However, a pleiotropic effect of this is to increase the recombination rate which can decrease the effectiveness of female meiotic drivers
-active during MI \citep{Brandvain}. 
-We speculate that other effects of sperm on female meiosis may also be tolerated by females as a way of ensuring meiotic fairness -- 
-	that is females evolve to allow male influence on female meiosis in cases in which the male influence is  egalitarian.
+The interests of fertilizing sperm and female are quite well aligned during syngamy -- producing a viable and potentially fit offspring. 
+%As noted above there are even opportunities for coevolution between sperm-based suppressors of female meiotic drive and the female genome.
+Thus it seems plausible that aspects of oogenesis may actively evolve to facilitate the influence of sperm on ensuring a fair  outcome of female meiosis which generates fit offspring.
+In light of that it is interesting to consider whether the phylogenetic variation in the stage of oogenesis when fertilization takes place reflects not
+	only species' different reproductive ecologies, but also different selection pressures shaping the interaction of sperm alleles with female meioses. 
+%\gc{Add short paragraph about sex chromosomes here? XY systems don't care. ZW systems (males ZZ, females ZW). If there is
+%sex ratio distortion how could this play out?}
+
+%\yb{DOES THIS GO ANYWHERE? }
+
+% Males can potentially influence the outcome of meiosis even before fertilization. 
+% For example in Drosophila and other taxa, sperm accessory
+%         gland proteins (ACPs) increase the female recombination rate
+%         \cite{SOMEthING}, and there is some evidence that there is
+%         variance across male genotypes in there effect on female
+%         recombination rate \cite{Stevison2012}.
+% Presumably this is a pleiotropic consequence of the resolution of double strand breaks initiated by the stressful/toxic composition of ACPs. 
+% However, a pleiotropic effect of this is to increase the recombination rate which can decrease the effectiveness of female meiotic drivers
+% active during MI \citep{Brandvain}. 
+% We speculate that other effects of sperm on female meiosis may also be tolerated by females as a way of ensuring meiotic fairness -- 
+% 	that is females evolve to allow male influence on female meiosis in cases in which the male influence is  egalitarian.
 
@@
 
-Therefore, the interests of fertilizing sperm and female are quite well aligned during syngamy -- producing a viable and potentially fit offspring. 
-%As noted above there are even opportunities for coevolution between sperm-based suppressors of female meiotic drive and the female genome.
-Thus it seems plausible that aspects of oogenesis may actively evolve to facilitate the influence of sperm on ensuring a fair  outcome of female meiosis which generates fit offspring.
-In light of that it is interesting to consider whether the phylogenetic variation in the stage of oogenesis when fertilization takes place reflects not
-	only species' different reproductive ecologies, but also different selection pressures shaping the interaction of sperm alleles with female meioses. 
-\\
-\\
-\\
-\\
-\gc{Other pointsf}\\
-\\
-\\
-\gc{Add short paragraph about sex chromosomes here? XY systems don't care. ZW systems (males ZZ, females ZW). If there is
-sex ratio distortion how could this play out?}
-
-POssibility that this could happen in plants if pollen emit signals to ``egg''
-Discussion of OM and IN.
+\section*{Appendix}
+In our first scenerio allele $A$ is the ancestral allele and allele
+$B$ is our sperm dependent drive allele.
+The sperm dependence of the drive is a form of assortative
+transmission (mating), and so we have to keep track of the genotype frequencies.
+Our genotype frequencies at birth are $x_{AA},~x_{AB},$ and
+$x_{BB}$. These three genotypes have fitnesses $w_{AA},~w_{AB},$ and
+$w_{BB}$, which are identical between males and females. As the
+fitnesses are the same in males and females the frequencies of the
+genotypes are identical in males and females, but in the following
+equations they are marked with $\Mars$ and $\Venus$ respectively to
+allow the reader to more easily follow the logic.
+The frequency of the three genotypes at birth in the next generation is:
+\begin{equation}
+x'_{BB} =\frac{1}{\bar{w} } \Big( w_{BB} x_{BB}^{\Venus} (w_{BB} x_{BB}^{\Mars} +\frac{1}{2}
+  w_{AB} x_{AB}^{\Mars} ) + w_{AB} x_{AB}^{\Venus} (d w_{BB}
+  x_{BB}^{\Mars} +\frac{1}{2} d
+w_{AB} x_{AB}^{\Mars} ) \Big),  \label{eqn.A1}
+\end{equation}
+\begin{equation}
+x'_{AA} =\frac{1}{\bar{w} } \Big(  w_{AA} x_{AA}^{\Venus} (w_{AA} x_{AA}^{\Mars} + \frac{1}{2}
+   w_{AB} x_{AB}^{\Mars}) + w_{AB} x_{AB}^{\Venus}(\frac{1}{2} w_{AA} x_{AA}^{\Mars}+ \frac{1}{2} 
+  w_{AB} x_{AB}^{\Mars}) \Big),  \label{eqn.A2}
+\end{equation}
+and
+\begin{eqnarray}
+x'_{AB} = & \frac{1}{\bar{w} } \Big( w_{AA} x_{AA}^{\Venus}  ( w_{BB}  x_{BB}^{\Mars} + \frac{1}{2}
+  w_{AB}  x_{AB}^{\Mars} ) +w_{BB} x_{BB}^{\Venus} ( w_{AA}  x_{AA}^{\Mars} + \frac{1}{2}
+  w_{AB}  x_{AB}^{\Mars} )  \nonumber \\
+& + x_{AB}^{\Venus}(\frac{1}{2}  w_{AA}  x_{AA}^{\Mars}+\frac{1}{2} (1-d)
+ w_{AB}  x_{AB}^{\Mars} + \frac{1}{4}  w_{AB}
+ x_{AB}^{\Mars} + (1-d)  w_{BB}  x_{BB}^{\Mars} ) \Big) \label{eqn.A3}
+\end{eqnarray}
+where the mean fitness of the population is
+\begin{equation}
+\bar{w} = w_{AA} x_{AA}+w_{AB}  x_{AB} +w_{BB}  x_{BB}.
+\end{equation}
 

### 2014-02-07 writeup_and_notes/male_dependent_drive.tex
wrote figure captions

@@
           \textwidth]{Figures/Sperm_egg_cartoon_paper_Fig2.eps}}
-\caption{Transmission probabilities righthand allele through female
-  meiosis}  \label{Eggsperm_3_allele_cartoon}
+\caption{Dynamics in 3 allele system
+\bf{A)} The sperm-dependent transmission probabilities of alleles through
+female meiosis (female germ cells depicted as circles, sperm indicated by wiggly
+tails). In each case the top and bottom entry in each cell give the
+transmisson probability of the left and right allele in the female
+being transmitted to the egg. In the left panel we show the transmission probabilities for
+  A, B, and B'. Where B' is a driver allele which is either promotes
+  itself through through female meiosis when sperm (a B$^+$ allele,
+  $d'>d$) or self-avoiding through the interaction of sperm with female meiosis (a B$^-$ allele,
+$d'<d$). In the right panel we depict the transmission probabilities for
+  A, B, and A$^-$ allele a non-driven allele that acts to reduce the
+  effects of female drive through sperm.
+\bf{B) Frequency trajectories in the three allele system. }
+}  \label{Eggsperm_3_allele_cartoon}
 \end{figure}

### 2014-02-14 writeup_and_notes/male_dependent_drive.tex
adding to appendix

@@
 \caption{Dynamics in 3 allele system
-\bf{A)} The sperm-dependent transmission probabilities of alleles through
+{\bf (A) } The sperm-dependent transmission probabilities of alleles through
 female meiosis (female germ cells depicted as circles, sperm indicated by wiggly
@@
   effects of female drive through sperm.
-\bf{B) Frequency trajectories in the three allele system. }
-}  \label{Eggsperm_3_allele_cartoon}
+{\bf (B)} Frequency trajectories in the three allele system. }
+ \label{Eggsperm_3_allele_cartoon}
 \end{figure}
@@
 
-\yb{??In the SUPP we explore the dynamics of a sperm modifier of female drive unlinked to the female driver??}
+%\yb{??In the SUPP we explore the dynamics of a sperm modifier of female drive unlinked to the female driver??}
 
@@
 %%%sperm-acting drive suppressors
-While  sperm enhancement of a female drive cannot displace a polymorphic female driver, sperm based drive suppression - 
-	that is sperm-acting alleles that act to restore  fairness of female meiosis can. 
+While  sperm enhancement of a female drive cannot displace a polymorphic female driver, sperm based drive suppression - that is sperm-acting alleles that act to restore  fairness of female meiosis can.
 If these sperm-acting drive suppressors arise on
@@
 \end{equation}
+We iteration can then iterate this recursion over generations, for
+example to produce the frequency trajectories shown in the inset
+Figure \ref{Invasion_space} by plotting $x_{BB}+ \frac{1}{2}x_{AB}$
+over time. We can also use this to assess the
+invasion criteria for the allele by starting allele B at very low
+frequency, and seeing if it increases across the generations.
 
+\paragraph{Dominance} In the supplement we also explore a model where
+the genotype of the male who produced the fertilizing sperm affects female drive,
+instead of just the allelic type of the sperm. We assume that if a
+heterozygous egg ($AB$) is fertilized by a spermwith the following
+male ($\Mars$) genotypes the $B$ allele will be transmitted through
+female meiosis with the follow probabilites
+\begin{tabular}{cccc}
+$\Mars$ geno. & AA & AB  & BB\\
+B transmission prob. & $\frac{1}{2}$ &  $\frac{1}{2} + h(d-\frac{1}{2})$  & $d$\\ 
+\end{tabular}
+where $d > 1/2$ and $h$ ($0<h<1$) acts to specify the dominance of the $B$
+allele's effect on female meiotic drive through the males genotype. We
+set up a set of recursions, similar to those in eqn.
+\ref{eqn.A1}-\ref{eqn.A1}, but now featuring these male genotype
+transmission probabilites. In Supplementary Figure \ref{Effect_of_dominance} we show the invasion
+criteria for both the A and the B allele when they are rare in the
+population. We find that there is a narrow parameter range the B allele
+can invade but not fix if it is dominant (i.e. $h>0$) and the cost in homozygotes
+is not too great. This reflects the fact that when the allele is rare
+it can increase in frequency due to fertilization of heterozygous eggs
+by sperm from heterozygote males, where the sperm do not themselves
+carry the allele. But as the B allele becomes common it will frequently be
+present in the homozygous state in males, and so will be unable to
+avoid the production of unfit homozygote zygotes leaving it trapped in
+the population.
+
+\paragraph{three allele system}
+%$x_{AA}$, $x_{AB}$,$x_{AB'}$, $w_{BB'} x_{BB'}$, $w_{B'B'} x_{BB}$, $w_{B'B'} x_{B'B'}$
+We denote the frequency of the three genotypes in the population at birth are
+$x_{AA}$, $x_{AB}$, $x_{AB'}$, $x_{BB'}$, $x_{BB}$, $x_{B'B'}$. 
+The mean relative fitness of the population is 
+\begin{equation}
+x_{AA}+x_{AB}+ x_{AB'}+ w_{BB'} +x_{BB'}+w_{B'B'}x_{BB} + w_{B'B'} x_{B'B'} 
+\end{equation}
+The frequency of the genotypes in meiotic-cells of female adults is 
+\begin{equation}
+\frac{1}{\bar{w}}\Big( x_{AA}, ~x_{AB},~ x_{AB'},~ w_{BB'} x_{BB'},~ w_{B'B'}x_{BB},
+~w_{B'B'} x_{B'B'} \Big)
+\end{equation}
+The frequency of fertilization events can be found by combining these
+female frequencies with the frequency of the three alleles in sperm is
+\begin{eqnarray}
+x_A = & x_{AA} + \frac{1}{2} x_{AB}+ \frac{1}{2} x_{AB'} \\
+x_B = & w_{BB}  x_{BB} +   w_{BB'} \frac{1}{2} x_{BB'}+ \frac{1}{2} x_{AB} \\
+x_{B'} = & w_{B'B'}  x_{B'B'} +   w_{BB'} \frac{1}{2} + x_{BB'}+ \frac{1}{2} x_{AB'}  
+\end{eqnarray}
+in combination with the transmission probabilities given in Figure \ref{Eggsperm_3_allele_cartoon}A.
+
+\begin{table}
+\begin{tabular}{lllll}
+\hline 
+  &  \multicolumn{4}{c}{Sperm genotype}   \\
+  &  & A & B & B'  \\
+$\Venus$ geno. & Freq.  & $x_{AA} + \frac{1}{2} x_{AB}+ \frac{1}{2} x_{AB'} $ & $ w_{BB}  x_{BB} +   w_{BB'} \frac{1}{2} x_{BB'}+
+\frac{1}{2} x_{AB}$ & $ w_{B'B'}  x_{B'B'} +   w_{BB'} \frac{1}{2}
+x_{BB'}+ \frac{1}{2} x_{AB'}  $\\
+\hline\\
+AA & $x_{AA}$ &  & & \\
+AB & $x_{AB}$ &  & & \\
+AB' & $x_{AB'}$ &  & & \\
+BB' & $w_{BB'} x_{BB'}$ &  & & \\
+BB & $w_{B'B'} x_{BB}$ &  & & \\
+B'B' & $w_{B'B'} x_{B'B'}$ &  & & \\
+\end{tabular}
+\caption{ALTERNATIVELY I COULD FILL IN THIS MATRIX  with genotype }
+\end{table}
 \bibliography{refs}
@@
 \includegraphics[width = 0.8 \textwidth]{Figures/bistable_x_vs_d_additive_s.eps} 
-\caption{Bistable green-bearded allele when selection cost (s) acts
+\caption{Bistable of a drive allele when selection cost (s) acts
   additively, for an allele whose effect
@@
 \includegraphics[width = 0.8 \textwidth]{Figures/effect_of_dominance_on_invasion_space_one_graph.eps} 
-\caption{Invasion analysis of green-bearded sperm allele whose effect
+\caption{Invasion analysis of an allele whose effect
  on female meoisis is mediated by the genotype of the fertilizing

### 2014-02-16 writeup_and_notes/male_dependent_drive.tex
adding to appendix

@@
 %transmission advantage through female meiosis.
-
+  \appendix
 \section*{Appendix}
@@
 
+\paragraph{cost of driver in heterozygotes}
+We also explored a range of other costs to the drive alleles other than a
+simple recessive cost, for example an additive cost. We modified the recursions given in eqn
+\ref{eqn.A1}-\ref{eqn.A3} to feature an additive cost $s$ for each
+copy of the $B$ allele. A none recessive cost to the drive allele can
+cause bistability of the drive allele. The B allele is unable to enter
+the population when rare, this results from the fact that
+when the allele is rare ($x_B \ll 1$) sperm carrying it rarely
+encounter heterozygous eggs ($\sim x^2$) but the allele still pays a
+cost in heterozygous individuals.  While when the B allele is common ($x_B
+\gg 0$) encounters between heterozygous eggs and $B$ allele sperm
+become more common allowing drive to occur this can allow the $B$
+allele to spread when common. In Supplementary Figure
+\ref{bistable_additive} we show the unstable equilibrium frequency
+above which the $B$ allele can invade. This equilibrium frequency was found by a simple
+parameter search.  \gc{wonder if a bit of work could find this}
+
 \paragraph{Dominance} In the supplement we also explore a model where
@@
 instead of just the allelic type of the sperm. We assume that if a
-heterozygous egg ($AB$) is fertilized by a spermwith the following
+heterozygous egg ($AB$) is fertilized by a sperm with the following
 male ($\Mars$) genotypes the $B$ allele will be transmitted through
-female meiosis with the follow probabilites
+female meiosis with the follow probabilites\\
 \begin{tabular}{cccc}
 $\Mars$ geno. & AA & AB  & BB\\
-B transmission prob. & $\frac{1}{2}$ &  $\frac{1}{2} + h(d-\frac{1}{2})$  & $d$\\ 
-\end{tabular}
-where $d > 1/2$ and $h$ ($0<h<1$) acts to specify the dominance of the $B$
+P(B). & $\frac{1}{2}$ &  $\frac{1}{2} + h(d-\frac{1}{2}) $ & $d$ 
+\end{tabular}\\
+where $d > 1/2$ and $h$ ($0 \geq h \geq 1$) acts to specify the dominance of the $B$
 allele's effect on female meiotic drive through the males genotype. We
-set up a set of recursions, similar to those in eqn.
+set up a set of recursions using a recessive sex independent cost to
+the driver, similar to those in eqn.
 \ref{eqn.A1}-\ref{eqn.A1}, but now featuring these male genotype
@@
 avoid the production of unfit homozygote zygotes leaving it trapped in
-the population.
+the population. \gc{Perhaps we should show the equil. freq. in the
+  fig, as I'm guessing it might be rare}
 
@@
 \end{eqnarray}
-in combination with the transmission probabilities given in Figure \ref{Eggsperm_3_allele_cartoon}A.
+in combination with the transmission probabilities given in Figure
+\ref{Eggsperm_3_allele_cartoon}A.
+
+\gc{Alternatively we could give the table below:}
 

### 2014-03-04 writeup_and_notes/male_dependent_drive.tex
minor typo fixing and wordsmithing, also ref. updating

@@
 %\title{Why does female meiosis allow the opportunity for exploitation by selfish sperm? BETTER TITLE?}
-\title{Co-operation between the sexes trumps selfish control of female
-  meiotic drive. Or 
-Sperm do not evolve to collaborate in female meiotic drive. Or Why does female meiosis allow the opportunity for exploitation by selfish sperm? }
+\title{
+%Co-operation between the sexes trumps selfish control of female  meiotic drive. 
+%Or 
+Sperm do not evolve to collaborate in female meiotic drive. 
+%Or Why does female meiosis allow the opportunity for exploitation by selfish sperm? 
+}
 
@@
 \citep{Prout1973}, 
-and often generate an ongoing evolutionary escalation of drive suppressors and enhancers \citep{Dawe1996,Fishman2008}. 
+and often generate ongoing evolutionary escalations of drive suppressors and enhancers \citep{Dawe1996,Fishman2008}. 
 The threat of meiotic drive to organismal fitness is potentially so
@@
 meiosis I \citep{Haig1991}, arrested female meiosis
-\citep{Mira1998}, centromere machinery \citep{MalikHenikoff}, and sex differences in the recombination rate \citep{Haig2010,Brandvain2012} 
-have perhaps evolved to disrupt meiotic drive and enforce fairness.
+\citep{Mira1998}, centromere machinery \citep{Malik2002a,Malik2009,Malik2009}, and sex differences in the recombination rate \citep{Haig2010,Brandvain2012} 
+have perhaps evolved to enforce fairness by disrupting meiotic drive \citep{Rice2013}.
 
@@
 The stage at which female meiosis requires fertilization varies widely across taxa, and in a few animal taxa, 
-	female meiosis is complete upon ovulation (that is fertilization is not required for female meiosis \citep[see
-	Table 1 of ][]{Masui_book}).
+	female meiosis is complete upon ovulation (that is fertilization is not required for female meiosis \citep[see table 1 of ref 
+][]{Masui_book}).
 
-It is also not the case that sperm is mechanistically incapable of influencing the outcome of female meiosis.
-Mechanistic evidence for this possibility comes from
-\emph{C. elegans} the fact that premature deployment of the aster (a vital component of mitotic
-machinery) provided by the sperm can disrupting MII meiotic segregation
+It is also not the case that sperm is mechanistically incapable of influencing female meiosis.
+In \emph{C. elegans} premature deployment of the aster (a vital component of mitotic
+machinery) provided by the sperm can disrupts MII meiotic segregation
 in the egg, leading to a triploid zygote \citep{McNally2012}. 
-Additionally, genetic evidence suggests that the transmission patterns
-in heterozygous females many potentially depend on sperm haplotype. 
-Specifically, two female meiotic drive systems in mouse (In and Om), both operate by distorting the second meiotic division, 
+Thus, there is direct mechanistic evidence for that sperm can influence female meiosis.
+ Additionally, genetic evidence from the In and Om  drive system sin mice suggests that the 
+ transmission of alternate alleles in heterozygous females can depend on sperm haplotype.  
+ Both of these female drive systems operate by distorting the second meiotic division, 
  and in both systems the outcome of female meiosis depends the genotype of the fertilizing sperm \citep{Agulnik1993,Wu2005}.  
@@
 In the standard single-locus, biallelic model of female meiotic drive, the driving allele is transmitted to the egg  in heterozygotes with probability  $d > \frac{1}{2}$, regardless of sperm genotype.
-To depict a case of a self-promoting (or `green-bearded') meiotic
+To depict a case of a self-promoting meiotic
 driver,  we modify this standard model such that the driver is only effective when fertilized by a sperm carrying that allele (see Figure
@@
 	providing an opportunity for the evolution of
-	drive suppressors \citep{XX}, corresponding well to 
+	drive suppressors (CITE - citep{XX}), corresponding well to 
 	empirical examples of female meiotic drive.
@@
         when it drives, and therefore there is a drive-associated fitness cost even when rare. 
-Our self-promoting driver must overcome this homozygous fitness cost simply to spread when rare.
+The self-promoting driver must overcome this homozygous fitness cost simply to spread when rare.
 Therefore with a recessive fitness cost,
@@
 
-While the allelic identity of sperm could plausibly influence the outcome of female meiosis, 
+Although the allelic identity of sperm could plausibly influence the outcome of female meiosis, 
 	limited gene expression in sperm \citep[e.g.][]{Joseph2004}
@@
 Above, we explored a model were a driver that drove in females when signaled by a complementary allele in sperm.  
-We complement this single locus model of an allele with effects in
-both male and with femalr, with  an alternative model of two linked loci - one female driver and the other a sperm allele which modifies the effect of drive upon fertilization. 
+We complement this single locus model of an allele with pleiotropic effects in both sexes, 
+ with  an alternative model of two linked loci - one female driver and the other a sperm allele which modifies the effect of drive upon fertilization. 
 %Here, we examine an alternative scenario (and more realistic) in which
-In this model initially a female meiotic driver, with no sperm-dependency, reaches drive-viability equilibrium (with two alleles
+In this model, a female meiotic driver with no sperm-dependency, initially reaches drive-viability equilibrium (with two alleles
 $A$ and $B$ are the ancestral non-driver and driver alleles).
@@
 	for a simple schematic of this drive model.  
-Naively, the $B^{+}$ may spread by capitalizing on the additional drive it causes; hover,  
+Naively, the $B^{+}$ may spread by capitalizing on the additional drive it causes; however,  
 	this is not the case for a few simple reasons. 
@@
 	the mother whose offspring they will inhabit, 
-	and even the organism in which they reside. 
-This potential generates numerous conflicts  \cite{Burt2006} including sexual conflicts between mates \cite{Arnqvist2005}, 
-	as well as conflicts between alleles that are overtransmitted in meiosis and the organisms they injure while doing so. 
+	or the organism in which they reside. 
+This generates numerous conflicts  \cite{Burt2006} including sexual conflicts between mates \cite{Arnqvist2005}, 
+	and conflicts between alleles that are overtransmitted in meiosis and the organisms they injure while doing so. 
 Such conflicts and their resolution likely play a major role in the structure and evolution of many basic biological processes \citep{Rice2013}.
@@
 	it spends very little time at intermediate frequency.
-As such, this type of exploitation cannot generate a sustained genetic conflict and so it is unlikely
+As such, this type of exploitation cannot generate a sustained genetic conflict.
+It is therefore unlikely
 	that female oogenesis and meiosis will evolve to prevent their effect.  
@@
         $ln^-$ allele is over transmitted through female meiosis in $ln^-/ln^+$ heterozygote when fertilized by $ln^+$
-        sperm \cite{agulnick, Pomiankowski1993}.   
-More generally both linked and unlinked alleles may often spread that act through sperm to
-reduce the opportunity of female meiotic drive in order to disable
-segregating drive systems, by for example reducing the mechanistic
-asymmetries that female drivers exploit.
+        sperm (CITE cite{agulnick, Pomiankowski1993}).   
+More generally, alleles that act through sperm to
+ reduce the opportunity of female meiotic drive regardless of linkage or phase.
+
+%More generally both linked and unlinked alleles may often spread that act through sperm to
+%reduce the opportunity of female meiotic drive in order to disable
+%segregating drive systems, by for example reducing the mechanistic
+%asymmetries that female drivers exploit.
 
@@
 $x_{BB}$. These three genotypes have fitnesses $w_{AA},~w_{AB},$ and
-$w_{BB}$, which are identical between males and females. As the
-fitnesses are the same in males and females the frequencies of the
-genotypes are identical in males and females, but in the following
+$w_{BB}$, which are identical between males and females. Because 
+fitnesses are equivalent in both sexes, genotypic frequencies
+are identical in males and females, but in the following
 equations they are marked with $\Mars$ and $\Venus$ respectively to
@@
 \end{equation}
-We iteration can then iterate this recursion over generations, for
+We can then iterate this recursion over generations, for
 example to produce the frequency trajectories shown in the inset
@@
 encounter heterozygous eggs ($\sim x^2$) but the allele still pays a
-cost in heterozygous individuals.  While when the B allele is common ($x_B
+cost in heterozygous individuals.  By contrast, when the B allele is common ($x_B
 \gg 0$) encounters between heterozygous eggs and $B$ allele sperm
@@
 the genotype of the male who produced the fertilizing sperm affects female drive,
-instead of just the allelic type of the sperm. We assume that if a
-heterozygous egg ($AB$) is fertilized by a sperm with the following
-male ($\Mars$) genotypes the $B$ allele will be transmitted through
-female meiosis with the follow probabilites\\
-\begin{tabular}{cccc}
-$\Mars$ geno. & AA & AB  & BB\\
-P(B). & $\frac{1}{2}$ &  $\frac{1}{2} + h(d-\frac{1}{2}) $ & $d$ 
-\end{tabular}\\
-where $d > 1/2$ and $h$ ($0 \geq h \geq 1$) acts to specify the dominance of the $B$
-allele's effect on female meiotic drive through the males genotype. We
-set up a set of recursions using a recessive sex independent cost to
+instead of just the allelic type of the sperm.
+ We assume a heterozygous female will transmit the $B$ allele 
+  with probabilities  $\frac{1}{2}$,  $\frac{1}{2} + h(d-\frac{1}{2}) $ or $d$, 
+ if she mated with a n $AA$, $AB$, or $BB$ male,  respectively.
+ Above, $d > 1/2$ and $h$ ($0 \geq h \geq 1$) acts to specify the dominance of the $B$
+%allele's effect on female meiotic drive through the males genotype.
+%YB TRY_1 We assume that in a mating between a heterozygous female and an $AA$, $AB$, or $BB$ male, the female will transmit the $B$ allele 
+% with probabilities  $\frac{1}{2}$,  $\frac{1}{2} + h(d-\frac{1}{2}) $, and $d$, respectively.
+  %Here $d > 1/2$ and $h$ ($0 \geq h \geq 1$) acts to specify the dominance of the $B$
+%allele's effect on female meiotic drive through the males genotype.
+% GC ORIGINAL  We assume that if a
+%heterozygous egg ($AB$) is fertilized by a sperm with the following
+%male ($\Mars$) genotypes the $B$ allele will be transmitted through
+%female meiosis with the following probabilities \\
+%\begin{tabular}{cccc}
+%$\Mars$ geno. & AA & AB  & BB\\
+%P(B). & $\frac{1}{2}$ &  $\frac{1}{2} + h(d-\frac{1}{2}) $ & $d$ 
+%\end{tabular}\\
+%where $d > 1/2$ and $h$ ($0 \geq h \geq 1$) acts to specify the dominance of the $B$
+%allele's effect on female meiotic drive through the males genotype. 
+We set up a set of recursions using a recessive sex independent cost to
 the driver, similar to those in eqn.
@@
 transmission probabilites. In Supplementary Figure \ref{Effect_of_dominance} we show the invasion
-criteria for both the A and the B allele when they are rare in the
-population. We find that there is a narrow parameter range the B allele
-can invade but not fix if it is dominant (i.e. $h>0$) and the cost in homozygotes
-is not too great. This reflects the fact that when the allele is rare
+criteria for both the A and B alleles when they are rare. 
+We find a narrow parameter range where B 
+can invade but not fix.
+This corresponds to B being less than fully recessive (i.e. $h>0$) and to a modest cost in homozygotes. 
+This combination of parameters reflects the fact that when the allele is rare, 
 it can increase in frequency due to fertilization of heterozygous eggs

### 2014-04-23 writeup_and_notes/male_dependent_drive.tex
Added some analytical results and fixed a few typos

@@
 a meiotic driver), may increase in frequency even if this driver entails a pleiotropic fitness cost \citep{Prout1973}, generating a genetic conflict between the success of the driver and organismal fitness.
-Meiotic drivers observed in nature (in both plants
+Meiotic drivers observed in both plants
 \citep{Buckler1999,Fishman2005,Fishman2008}, and animals
-\citep{Agulnik1990,Wu2005,Pardo-ManuelDeVillena2001c}) highlight this
+\citep{Agulnik1990,Wu2005,Pardo-ManuelDeVillena2001c} highlight this
 conflict -- the selfish benefits of drive and the associated
@@
 creating ample opportunity for interaction between the sperm and female meiotic machinery. 
-For example, an allele in sperm that facilitates meiotic drive by a genetically equivalent allele in female meiosis %(a  heteromorphic dyad), 
+Therefore, an allele in sperm that facilitates meiotic drive by a genetically equivalent allele in female meiosis %(a  heteromorphic dyad), 
      could presumably rapidly spread through the population. 
-At first sight, it seems as although female meiosis is primed to be exploited by selfish sperm systems.  
+At first sight, female meiosis appears primed for exploitation by selfish sperm systems.  
 Why then is the requirement of fertilization to complete female meiosis so ubiquitous?
@@
 It is also not the case that sperm is mechanistically incapable of influencing female meiosis.
-In \emph{C. elegans} premature deployment of the aster (a vital component of mitotic
-machinery) provided by the sperm can disrupts MII meiotic segregation
+In \emph{C. elegans} premature deployment of the aster, a vital component of mitotic
+machinery, provided by the sperm disrupts MII meiotic segregation
 in the egg, leading to a triploid zygote \citep{McNally2012}. 
-Thus, there is direct mechanistic evidence for that sperm can influence female meiosis.
- Additionally, genetic evidence from the In and Om  drive system sin mice suggests that the 
+Thus, there is direct mechanistic evidence that sperm can influence female meiosis.
+ Additionally, genetic evidence from the \emph{In} and \emph{Om}  drive systems in mice suggests that the 
  transmission of alternate alleles in heterozygous females can depend on sperm haplotype.  
- Both of these female drive systems operate by distorting the second meiotic division, 
- and in both systems the outcome of female meiosis depends the genotype of the fertilizing sperm \citep{Agulnik1993,Wu2005}.  
-While such genetic evidence is suggestive, we note the difficulty in
-ruling out the alternative hypothesis of early selection on zygotes \citep[pages
-52-54][]{Burt2006}.
-
-%However,  we acknowledge the difficulty in ruling out early mortality
-%as an alternative explanation of the 
-%specifics of these putative drivers \citep[Give page numbers to
-%discussion of this in][]{Burt2006}. \gc{should be careful here as the
- % Om locus is pretty convincingly a meiotic driver, doubts have just
- % been raised about whether IN is. Whether the sperm dep nature of
- % either is robust to this is I think more at question.}
+ While ruling out the alternative hypothesis of early selection on zygotes \citep[pages
+52-54][]{Burt2006} in these cases is challenging, it appears that the extent to which \emph{In} and \emph{Om} distort the second female meiotic division partially depends on the genotype of the fertilizing sperm \citep{Agulnik1993,Wu2005}.
+% Both of these female drive systems operate by distorting the second meiotic division, 
+% and in both systems the outcome of female meiosis depends the genotype of the fertilizing sperm \citep{Agulnik1993,Wu2005}.  
+%While such genetic evidence is suggestive, we note the difficulty in
+%ruling out the alternative hypothesis of early selection on zygotes \citep[pages
+%52-54][]{Burt2006}.
 
@@
 	when rare it occurs predominantly in heterozygotes and it therefore drives without a fitness cost. 
-However, a severe homozygous fitness cost ($s>XXd$) prevents  
+However, a severe homozygous fitness cost prevents  
 	drivers from fixing  \citep{Prout1973}, 
@@
  	are far more restrictive than those for a standard meiotic driver (see Figure \ref{Invasion_space}).
+\yb{When fully recessive, a self-promoting driver can only invade when $s$, the selection against the drive homozygote, is less than $\frac{2 d - 1}{4 d}$ (The tan region in Figure  \ref{Invasion_space}).}
 
@@
 allele that suppresses this form of meiotic drive in females.
-\yb{POINT TO SOME MATHS OR FIGURE}
+\yb{In fact, while a rare self promoting driving can only invade if $s<\frac{2 d - 1}{4 d}$, a common self-promoting driver can fix over a broader range of parameter space (i.e.
+if $s<\frac{2d-1}{2}$).
+This system therefore displays bistability, when $\frac{2 d - 1}{4 d}<s<\frac{2d-1}{2}$, the driver will deterministically fix when introduced at a frequency greater than $\frac{1-2d+4ds}{2s(2d-1)}$, but will be lost otherwise.}
 
@@
 
+\yb{\paragraph{Recessive fitness cost of self-promoting driver}
+When the self-promoting driver has a recessive fitness cost, $s$ genotype frequencies after selection and drive equal
+\begin{equation}	x'_{BB} =\frac{x_{B}^2(1-s)}{\bar{w} }  \Big( F+(1-F)(x_{B}+2x_{A} d) \Big) 	\end{equation}
+\begin{equation}	x'_{AA} =\frac{1}{\bar{w} }  \Big(x_{A}^2\Big) 	\end{equation}
+and
+\begin{equation}	x'_{AB} =\frac{2 x_{A}x_{B} }{\bar{w} }  \Big(1-x_{B}(d-1/2)(1-F) \Big)	\end{equation}
+
+its change in frequency across generations $\Delta f_B$ is
+\begin{equation}
+\frac{f_A f_B^2}{\bar{w}} \left( [1-F][ d(1-2 f_A s) - f_B s-1/2] - s  F \right)
+\end{equation}
+where $F$ is the deviation from Hardy-Weinberg expectations, and $f_A$ is the frequency of the non-driving allele. 
+For analytical solutions presented above, we assume that $F$ 
+}
+
 \paragraph{cost of driver in heterozygotes}

### 2014-04-27 writeup_and_notes/male_dependent_drive.tex
added heterozygous bistabiltiy analysis and editeed through single locus sperm effect

@@
 \section*{Introduction}
-
 Despite the apparent unity of the organism, alleles can
@@
 creating ample opportunity for interaction between the sperm and female meiotic machinery. 
-Therefore, an allele in sperm that facilitates meiotic drive by a genetically equivalent allele in female meiosis %(a  heteromorphic dyad), 
-     could presumably rapidly spread through the population. 
+Therefore, a `green-bearded' \citep{Gardner2010} sperm-acting allele that recognizes and facilitates the meiotic drive of a genetically equivalent allele in female meiosis %(a  heteromorphic dyad), 
+     could presumably rapidly spread through a population. 
 At first sight, female meiosis appears primed for exploitation by selfish sperm systems.  
@@
 It is certainly not the case that animals are mechanistically incapable of circumventing this requirement.
-The stage at which female meiosis requires fertilization varies widely across taxa, and in a few animal taxa, 
+The stage at which female meiosis requires fertilization varies widely across taxa, and in a few animals, 
 	female meiosis is complete upon ovulation (that is fertilization is not required for female meiosis \citep[see table 1 of ref 
@@
 It is also not the case that sperm is mechanistically incapable of influencing female meiosis.
-In \emph{C. elegans} premature deployment of the aster, a vital component of mitotic
-machinery, provided by the sperm disrupts MII meiotic segregation
+In \emph{C. elegans}, premature deployment of the aster, a vital component of mitotic
+machinery provided by the sperm, disrupts MII meiotic segregation
 in the egg, leading to a triploid zygote \citep{McNally2012}. 
@@
 
-Here, we develop population genetic models exploring the evolution of sperm influence on female meiotic drive. 
+We develop population genetic models exploring the evolution of sperm influence on female meiotic drive, focussing on models in which `self-promoting' alleles in sperm facilitate drive of like genotypes during gametogenesis in heterozygous females.
 %In this article we explore through simple population genetic models the consequences of alleles that influence the outcome of female meiosis.
@@
 female meiosis will have little incentive to evolve resistance to them.
-In fact, the interests of sperm and egg  genomes' are often aligned, as both are invested in the fate of the resultant zygote (as was speculated for the In locus \citep{Pomiankowski1993}).
+In fact, the interests of sperm and egg  genomes' are often aligned, as both are invested in the fate of the resultant zygote (as was speculated for the \emph{In} locus \citep{Pomiankowski1993}).
 Therefore, there is little selective benefit to females in preventing sperm to influence female meioses,
@@
 driver,  we modify this standard model such that the driver is only effective when fertilized by a sperm carrying that allele (see Figure
-\ref{Eggsperm_2_allele_cartoon}) \footnote{the timing of fertilization relative to female meiosis places another constraint on $d$, for example, if fertilization (and therefore, sperm dependent drive) takes place at MII (as in mammals),
-	female drive requires an uneven number of crossovers between the centromere and the drive locus, 
-	so $d$ is bounded to be $<0.75$? \citep[see ][ for discussion]{Buckler1999}. Could also put this in
-        caption of fig.}.
-We then identify the conditions under which this self propagating driver can spread, and evaluate whether a driver of this form could generate a sustained conflict favoring the evolution of suppressors.
-\yb{EQUATIONS A1-A4}
+\ref{Eggsperm_2_allele_cartoon}). 
+%\footnote{the timing of fertilization relative to female meiosis places another constraint on $d$, for example, if fertilization (and therefore, sperm dependent drive) takes place at MII (as in mammals),
+%	female drive requires an uneven number of crossovers between the centromere and the drive locus, 
+%	so $d$ is bounded to be $<0.75$? \citep[see ][ for discussion]{Buckler1999}. Could also put this in
+  %      caption of fig.}.
+We then identify the conditions allowing the spread of this self-promoting driver, 
+	and evaluate whether a driver of this form could generate a sustained conflict favoring the evolution of suppressors. 
+%\yb{ We present a general framework for this problem in equations A1-A4, and evaluate specialized cases below.}
 
-We first consider a driving allele neutral in heterozygotes, 
-	but deleterious in homozygotes (whose relative fitness is
-        $1-s$). This model is laid out in eqn. \ref{eqn.A1}-\ref{eqn.A3}.
-In this model, a standard meiotic driver can always invade because 
-	when rare it occurs predominantly in heterozygotes and it therefore drives without a fitness cost. 
-However, a severe homozygous fitness cost prevents  
-	drivers from fixing  \citep{Prout1973}, 
+For comparative purposes, we present the standard drive model in equations \ref{eqn.A1}-\ref{eqn.A3}, 
+	and refer readers to the previous literature on this model (\citep[e.g. ][]{Prout1973,Ubeda2004} for additional results). 
+In this case, assuming the driving allele is deleterious but fully recessive 
+	(i.e. homozygous fitness equals $w_{BB}=1-s$ and other genotypic fitnesses equal $w_{AA}=w_{AB}=1$), 
+	it always invades because,  when rare, it occurs predominantly in heterozygotes and therefore drives without a fitness cost. 
+However, when $s$ is large drivers cannot fix \citep{Prout1973}, 
 	providing an opportunity for the evolution of
-	drive suppressors (CITE - citep{XX}), corresponding well to 
-	empirical examples of female meiotic drive.
+	drive suppressors, corresponding well to 
+	empirical examples of female meiotic drive \citep[reviewed in ][]{Burt2006}.
 
 In contrast to a traditional driver, which drives but pays effectively
-no fitness cost as a rare deleterious recessive, 
-	a self-promoting driver creates homozygotes
-        when it drives, and therefore there is a drive-associated fitness cost even when rare. 
-The self-promoting driver must overcome this homozygous fitness cost simply to spread when rare.
-Therefore with a recessive fitness cost,
-	the conditions allowing the invasion of a self-promoting driver
- 	are far more restrictive than those for a standard meiotic driver (see Figure \ref{Invasion_space}).
-\yb{When fully recessive, a self-promoting driver can only invade when $s$, the selection against the drive homozygote, is less than $\frac{2 d - 1}{4 d}$ (The tan region in Figure  \ref{Invasion_space}).}
+	no fitness cost when rare, 
+	a self-promoting driver specifically creates homozygotes by driving. 
+%, and therefore there is a drive-associated fitness cost even when rare. 
+Thus it must overcome a drive-associated homozygous fitness cost simply to spread when rare,
+	and the conditions allowing the invasion of a self-promoting driver
+ 	are far more restrictive than those for a standard meiotic driver.
+When rare, a fully recessive, self-promoting driver can only invade when $s$ 
+	%, the selection against the drive homozygote, 
+	is less than $\frac{2 d - 1}{4 d}$. 
+This analytical result, derived  from Equation \eqref{deltadriver} assuming Hardy-Weinberg, 
+	closely matches results obtained by exact numerical iteration (Figure \ref{Invasion_space}).
 
-When a deleterious recessive self-promoting driver does spread, 
-	it spends much time at low frequency, because when rare it is unlikely to meet a complimentary sperm. 
-However, once relatively common, a self-promoting driver spreads rapidly to fixation due to its
+When a self-promoting driver does spread, 
+	it spends much time at low frequency, 
+	because the paucity of complimentary sperm compromises its ability to drive. 
+However, once relatively common, it  rapidly achieves fixation due to its
 	positive frequency dependent behavior (Inset in Figure \ref{Invasion_space}).  
-This lack of a balanced polymorphism ( Figure \ref{Invasion_space}) precludes the evolution of an
+This lack of a balanced polymorphism (Figure \ref{Invasion_space}) precludes the evolution of an
 	allele that suppresses this form of meiotic drive in females.
-\yb{In fact, while a rare self promoting driving can only invade if $s<\frac{2 d - 1}{4 d}$, a common self-promoting driver can fix over a broader range of parameter space (i.e.
-if $s<\frac{2d-1}{2}$).
-This system therefore displays bistability, when $\frac{2 d - 1}{4 d}<s<\frac{2d-1}{2}$, the driver will deterministically fix when introduced at a frequency greater than $\frac{1-2d+4ds}{2s(2d-1)}$, but will be lost otherwise.}
 
-Inclusion of a heterozygous fitness cost (e.g. additive fitness cost) further constrains the evolution of a self-promoting driver. 
-While a standard female driver invades from low frequency when the transmission advantage outweighs the fitness cost (XXX),
-	a rare self-promoter pays a fitness cost, but drives only weakly because complementary sperm are uncommon. 
-This frequency-independent fitness cost, and positive frequency dependent drive generates a bistable system. 
-Below a cutoff frequency the self-promoter driver deterministically decreases in frequency 
-	because it pays the  heterozygous fitness cost but drives
-        ineffectively (see Figure S\ref{bistable_additive}). 
-Above this frequency, the allele increases deterministically because the commonness of complementary sperm 
-	make drive effective enough to overcome the associated fitness cost.
+The positive frequency dependence of a self-promoting driver can induce bistability in its spread. 
+When rare, a  recessive self-promoting driver can spread only if $s < \frac{2 d - 1}{4 d}$;
+	however, once common such a driver can fix over a broader range of parameter space (i.e. if $s<\frac{2d-1}{2}$). 
+Therefore, when $\frac{2 d - 1}{4 d}<s<\frac{2d-1}{2}$, 
+	a recessive self-promoting driver will deterministically fix if drift, mutation, 
+	or migration pressure bring its frequency above $\frac{1-2d+4ds}{2s(2d-1)}$, 
+	but it will be lost otherwise.
+Inclusion of a heterozygous fitness cost (i.e. $w_{AB}=1-h_s$)
+	%(e.g. additive fitness cost) 
+	further constrains the evolution of a self-promoting driver. 
+This case also displays bistability -- a rare self-promoting driver cannot spread; 
+	however, if $s < \frac{2 d - 1+s_h(3-2d)}{2}$ it will fix deterministically if its 
+	frequency is greater than some threshold, $f_B^*$ [Equation \eqref{thresh2}, and \yb{FIGURES SXX}].
+%	because it pays the  heterozygous fitness cost but drives
+ %      ineffectively (see Figure S\ref{bistable_additive})
+%While a standard female driver invades from low frequency when the transmission advantage outweighs the fitness cost (XXX),
+%	a rare self-promoter pays a fitness cost, but drives only weakly because complementary sperm are uncommon. 
+%This frequency-independent fitness cost, and positive frequency dependent drive generates a bistable system. 
+%Below a cutoff frequency the self-promoter driver deterministically decreases in frequency 
+%	because it pays the  heterozygous fitness cost but drives
+ %      ineffectively (see Figure S\ref{bistable_additive}). 
+%Above this frequency, the allele increases deterministically because the commonness of complementary sperm 
+%	make drive effective enough to overcome the associated fitness cost.
 This bistability prevents self-promoting drivers from invading 	
-	reasonably sized populations, and assures that if they do invade they will rapidly fix.
+	reasonably sized populations, and assures that, if they do invade, they will rapidly fix.
 Our model therefore predicts that self-promoting drivers will not be observed as polymorphisms in natural populations.
@@
   meioses in female heterozygotes, inspective
-  of the allele carried by the fertilizing sperm. Such an allele can
-  invade over all of the parameter space, and will be maintained as a
+  of the allele carried by the fertilizing sperm (we note that the timing of fertilization relative to female meiosis places another constraint on $d$, for example, when drive takes place druing the second meitoic division
+	it requires an uneven number of crossovers between the centromere and the drive locus \citep[see ][ for discussion]{Buckler1999}). 
+Such an allele can invade over all of the parameter space, and will be maintained as a
   protected polymorphism in the white region, and can fix in the
@@
 This generates numerous conflicts  \cite{Burt2006} including sexual conflicts between mates \cite{Arnqvist2005}, 
-	and conflicts between alleles that are overtransmitted in meiosis and the organisms they injure while doing so. 
+	and conflicts between alleles that are over-transmitted in meiosis and the organisms they injure while doing so. 
 Such conflicts and their resolution likely play a major role in the structure and evolution of many basic biological processes \citep{Rice2013}.
@@
 	frequency until the fitness cost experienced when homozygous balances the transmission advantage.
-By contrast a sperm promotor of female drive is only effective when matched with a heterozgyote female -- meaning that when rare this allele rarely enhances female drive. 
+By contrast a sperm promotor of female drive is only effective when matched with a heterozygote female -- meaning that when rare this allele rarely enhances female drive. 
 Even worse, when it does so it will preferentially find itself in a low fitness drive homozygote. 
@@
 \frac{f_A f_B^2}{\bar{w}} \left( [1-F][ d(1-2 f_A s) - f_B s-1/2] - s  F \right)
+\label{deltadriver}
 \end{equation}
 where $F$ is the deviation from Hardy-Weinberg expectations, and $f_A$ is the frequency of the non-driving allele. 
-For analytical solutions presented above, we assume that $F$ 
+For analytical solutions presented above, we assume that $F$ equals zero. 
 }
@@
 \paragraph{cost of driver in heterozygotes}
-We also explored a range of other costs to the drive alleles other than a
+We explored a more complex model where the self-promoting driver has a cost, $s_h$ in heterozygotes (so $w_{AB} = 1-h_s$) in addition to the cost $s$, in homozygotes.
+Model construction is similar to that above.
+So long as there is any heterozygous fitness cost (i.e. $s_h > 0$) the self-promoting driver cannot invade when rare;
+	however, as is the case without a heterozygous cost, this system is bistable.
+Assuming no deviation from Hardy-Weinberg a self-promoting driver with a heterozygous fitness cost deterministically fixes if it's frequency exceeds some threshold, $f_B^*$.
+
+\begin{equation}
+	f_B^* = \frac{a_1+\sqrt{a_1^2+8 s_h a_2}}{2 a_2}
+	\label{thresh2}
+\end{equation}
+where, $a_1=1-2 ds_h+4 d s-2 d-3 s_h$ and $a_2=2(s-s_h)(2d-1)$.
+
+ range of other costs to the drive alleles other than a
 simple recessive cost, for example an additive cost. We modified the recursions given in eqn

### 2014-04-28 writeup_and_notes/male_dependent_drive.tex
added male-genotype dependent results

@@
 	diploid genotype (perhaps due to proteins and RNAs packaged into the sperm).
-This adds an additional parameter to the system -- the dominance of the drive enhancer in heterozygous males, 
-	when this dominance is high, a narrow sliver of parameter space sustaining a polymorphism emerges (see Figure S\ref{Effect_of_dominance})
- presumably because positive frequency dependence diminishes with the dominance of the male effect driver.
+%GRAHAM I HAVE REWRITTEN THIS SECTION TO REFLECT MATHS AND MY PARAMETERIZATION. WE CAN WORK ON SYNCHING OUR NOTATION
+%This adds an additional parameter to the system -- the dominance of the drive enhancer in heterozygous males, 
+%	when this dominance is high, a narrow sliver of parameter space sustaining a polymorphism emerges (see Figure S\ref{Effect_of_dominance})
+ %presumably because positive frequency dependence diminishes with the dominance of the male effect driver.
+This male-genotype dependent model requires one additional parameter, as we exchange $d$ in the sperm dependent case for $d_{het}$ and $d_{hom}$ which describe the transmission of the drive allele in a heterozygous female mating with males heterozygous and homozygous for the self-promoting drive allele, respectively.
+When $d_{het}$ is large relative to $d_{hom}$, a narrow sliver of parameter space sustains a polymorphism (see Figure S\ref{Effect_of_dominance}), presumably because positive frequency dependence diminishes with the dominance of the male effect driver.
+As an analytical example, consider the case of male-genotype dependent deleterious recessive female driver 
+	(assuming, as above, $w_{BB}=1-s$, and $w_{AA}=w_{AB}=1$). 
+Here, a rare driver invades when $s<\frac{d_{het}-1/2}{d_{het}}$, and fixes when $s < d_{hom} - \frac{1}{2}$. 
+Intermediate values of $s$, which represent a small proportion of parameter space can maintain a polymorphism.
 

### 2014-04-30 writeup_and_notes/male_dependent_drive.tex
fixed minor typos, changes in text inequalities to single-line rather than fract formatting

@@
 
-In the standard single-locus, biallelic model of female meiotic drive, the driving allele is transmitted to the egg  in heterozygotes with probability  $d > \frac{1}{2}$, regardless of sperm genotype.
+In the standard single-locus, biallelic model of female meiotic drive, the driving allele is transmitted to the egg  in heterozygotes with probability  $d > 1/2$, regardless of sperm genotype. 
 To depict a case of a self-promoting meiotic
@@
 For comparative purposes, we present the standard drive model in equations \ref{eqn.A1}-\ref{eqn.A3}, 
-	and refer readers to the previous literature on this model (\citep[e.g. ][]{Prout1973,Ubeda2004} for additional results). 
+	and refer readers to the previous literature on this model \citep[e.g. ][]{Prout1973,Ubeda2004} for additional results. 
 In this case, assuming the driving allele is deleterious but fully recessive 
@@
 %, and therefore there is a drive-associated fitness cost even when rare. 
-Thus it must overcome a drive-associated homozygous fitness cost simply to spread when rare,
-	and the conditions allowing the invasion of a self-promoting driver
- 	are far more restrictive than those for a standard meiotic driver.
+It must therefore overcome a drive-associated homozygous fitness cost simply to spread when rare. 
+The conditions allowing the invasion of a self-promoting driver
+ 	are consequently far more restrictive than those for a standard meiotic driver.
 When rare, a fully recessive, self-promoting driver can only invade when $s$ 
 	%, the selection against the drive homozygote, 
-	is less than $\frac{2 d - 1}{4 d}$. 
-This analytical result, derived  from Equation \eqref{deltadriver} assuming Hardy-Weinberg, 
+	is less than $(2 d - 1)/(4 d)$. 
+This analytical approximation, derived  from Equation \eqref{deltadriver} assuming Hardy-Weinberg, 
 	closely matches results obtained by exact numerical iteration (Figure \ref{Invasion_space}).
@@
 The positive frequency dependence of a self-promoting driver can induce bistability in its spread. 
-When rare, a  recessive self-promoting driver can spread only if $s < \frac{2 d - 1}{4 d}$;
-	however, once common such a driver can fix over a broader range of parameter space (i.e. if $s<\frac{2d-1}{2}$). 
-Therefore, when $\frac{2 d - 1}{4 d}<s<\frac{2d-1}{2}$, 
-	a recessive self-promoting driver will deterministically fix if drift, mutation, 
-	or migration pressure bring its frequency above $\frac{1-2d+4ds}{2s(2d-1)}$, 
+Compared to the invasion criteria, parameter space allowing fixations is much broader -- 
+	a common, recessive, self-promoting driver fixes if  $s$ is less than $(2d-1)/2$. 
+Therefore, when $(2 d - 1)/(4 d)<s<(2d-1)/(2)$, 
+	a recessive, self-promoting driver will deterministically fix if drift, mutation, 
+	or migration pressure bring its frequency above $(1-2d+4ds)/(2s(2d-1))$, 
 	but it will be lost otherwise.
@@
 	further constrains the evolution of a self-promoting driver. 
-This case also displays bistability -- a rare self-promoting driver cannot spread; 
-	however, if $s < \frac{2 d - 1+s_h(3-2d)}{2}$ it will fix deterministically if its 
-	frequency is greater than some threshold, $f_B^*$ [Equation \eqref{thresh2}, and \yb{FIGURES SXX}].
+This case also displays bistability -- with any heterozygous fitness cost, a rare self-promoting driver cannot spread. 
+However, if $s < d - 1/2+s_h(3-2d)/2$, 
+	it will fix deterministically if its 
+	frequency is greater than a threshold, $f_B^*$ [Equation \eqref{thresh2}, and \yb{FIGURES SXX}].
 %	because it pays the  heterozygous fitness cost but drives
@@
 	suggest a model where sperm influence female meiosis via expression of the fertilizing male's
-	diploid genotype (perhaps due to proteins and RNAs packaged into the sperm).
+	diploid genotype (perhaps due to proteins and RNAs packaged into the sperm), rather than sperm haplotype.
 %GRAHAM I HAVE REWRITTEN THIS SECTION TO REFLECT MATHS AND MY PARAMETERIZATION. WE CAN WORK ON SYNCHING OUR NOTATION
@@
 	(assuming, as above, $w_{BB}=1-s$, and $w_{AA}=w_{AB}=1$). 
-Here, a rare driver invades when $s<\frac{d_{het}-1/2}{d_{het}}$, and fixes when $s < d_{hom} - \frac{1}{2}$. 
+Here, a rare driver invades when $s$ is less than $(d_{het}-1/2)/d_{het}$, and fixes when $s$ is less than $d_{hom} - 1/2$. 
 Intermediate values of $s$, which represent a small proportion of parameter space can maintain a polymorphism.

### 2014-05-04 scripts/invasion_analysis.R
added Yaniv's theory lines to invasion space plot

@@ -217,13 +217,20 @@ save(file=paste(directory,"invasion_grid_homozy_cost.Robj",sep=""),fixation.line
 show(load(paste(directory,"invasion_grid_homozy_cost.Robj",sep="")))
 
 my.cols<-brewer.pal(2,name="Dark2")
-plot(c(0.5,1),c(0,1),xlab="Drive coefficient",ylab="selection coeff against homozygotes",type="n")
+plot(c(0.5,1),c(0,1),xlab="",ylab="",type="n")
+mtext("Drive coefficient, d",side=1,line=2)
+mtext("selection against homozygotes, s",side=2,line=2)
+
 polygon(x=c(d.range,rev(d.range)),y=c(fixation.lines.s3,rep(1,length(d.range))),col="white")
-polygon(x=c(d.range,rev(d.range)),y=c(fixation.lines.s3,rev(invasion.lines.s3.cutoff)),col=my.cols[1])
-polygon(x=c(d.range,rev(d.range)),y=c(invasion.lines.s3.cutoff,rep(0,length(d.range))),col=my.cols[2])
+polygon(x=c(d.range,rev(d.range)),y=c(fixation.lines.s3,rev(invasion.lines.s3.cutoff)),col=my.cols[1],border=NA)
+polygon(x=c(d.range,rev(d.range)),y=c(invasion.lines.s3.cutoff,rep(0,length(d.range))),col=my.cols[2],border=NA)
 text(0.8,0.35,"Simple driver invades but can't fix",srt=25)
 text(0.85,0.27,"Simple driver can invade & fix",col="white",srt=15)
 text(0.8,0.05,"Simple driver & Self promoter driver can invade & fix",col="white")
+lines(d.range,(2*d.range- 1)/2,col="black",lwd=3,lty=1) ##simple driver fixes below this line
+lines(d.range,(2*d.range- 1)/(4*d.range),col="black",lwd=3,lty=2) ##self prom. fixes below this line
+
+text(x=0.95,y=0.95,"A.",cex=2)
 
 par(fig=c(grconvertX(0.51, from = "user", to = "ndc"), grconvertX(0.75, from = "user", to = "ndc"), grconvertY(0.4, from = "user", to = "ndc"), grconvertY(1, from = "user", to = "ndc")), new = TRUE); 
 par(mar=c(2,2,1,0))
@@ -255,7 +262,7 @@ for(i in 1:3){
[…]

### 2014-05-05 writeup_and_notes/male_dependent_drive.tex
updated caption

@@
 
+% (we note that the timing of fertilization relative to female meiosis places another constraint on $d$, for example, when drive takes place druing the second meitoic division
+%	it requires an uneven number of crossovers between the centromere and the drive locus \citep[see ][ for discussion]{Buckler1999}).
+
 \begin{figure}
 \includegraphics[width = 0.8 \textwidth]{Figures/invasion_space_recessive_driver.eps}
-\caption{Invasion analysis for a female meiotic drive allele with
-  recessive costs. First consider a simple female meiotic
+\caption{{\bf A.} Invasion analysis for a female meiotic drive allele with
+  recessive costs (selection coefficient $s$). First consider a simple female meiotic
   drive system, where the allele is transmitted a fraction $d>1/2$ of
   meioses in female heterozygotes, inspective
-  of the allele carried by the fertilizing sperm (we note that the timing of fertilization relative to female meiosis places another constraint on $d$, for example, when drive takes place druing the second meitoic division
-	it requires an uneven number of crossovers between the centromere and the drive locus \citep[see ][ for discussion]{Buckler1999}). 
-Such an allele can invade over all of the parameter space, and will be maintained as a
+  of the allele carried by the fertilizing sperm. 
+Such an allele can invade the population over all of the parameter space, and will be maintained as a
   protected polymorphism in the white region, and can fix in the
-  population in all of the parameter space colored green,
-  and will fix  or orange. Now consider an allele that is fairly transmitted through
+  population in all of the parameter space colored green or orange. 
+The solid back line marks the criteria for the allele to fix $s< (2d-1)/2$.
+Now consider an allele that is fairly transmitted through
   female meiosis when fertilized by a sperm carrying a different
-  allele, but a fraction $d>1/2$ when a heterozygote egg is fertilized
+  allele, but at a fraction $d>1/2$ when a heterozygote egg is fertilized
   by a sperm carrying the allele (as depicted in Figure
@@
   the region of parameter space colored orange, and it fixes over all
-  of this parameter region. 
- In the inset we show trajectories of sperm-dependent female drive
+  of this parameter region. The approximate criteria for this allele
+  to fix is given by $s < (2d- 1)/(4d)$. 
+ {\bf B.} In the inset we show trajectories of sperm-dependent female drive
  alleles that can invade the population, each allele has $s =0.1$ against the

### 2014-05-05 writeup_and_notes/male_dependent_drive.tex
updated text of 2 allele system

@@
 
-In the standard single-locus, biallelic model of female meiotic drive, the driving allele is transmitted to the egg  in heterozygotes with probability  $d > 1/2$, regardless of sperm genotype. 
-To depict a case of a self-promoting meiotic
-driver,  we modify this standard model such that the driver is only effective when fertilized by a sperm carrying that allele (see Figure
+In the standard single-locus, biallelic model of female meiotic drive,
+the driving allele is transmitted to the egg  in heterozygotes with
+probability  $d > 1/2$, regardless of sperm genotype. To depict a case
+of a self-promoting meiotic driver,  we modify this standard model such that the driver is only effective when fertilized by a sperm carrying that allele (see Figure
 \ref{Eggsperm_2_allele_cartoon}). 
-%\footnote{the timing of fertilization relative to female meiosis places another constraint on $d$, for example, if fertilization (and therefore, sperm dependent drive) takes place at MII (as in mammals),
-%	female drive requires an uneven number of crossovers between the centromere and the drive locus, 
-%	so $d$ is bounded to be $<0.75$? \citep[see ][ for discussion]{Buckler1999}. Could also put this in
-  %      caption of fig.}.
+\footnote{the timing of fertilization relative to female meiosis places another constraint on $d$, for example, if fertilization (and therefore, sperm dependent drive) takes place at MII (as in mammals),
+	female drive requires an uneven number of crossovers between the centromere and the drive locus, 
+	so $d$ is bounded to be $<0.75$? \citep[see ][ for
+        discussion]{Buckler1999}. }. %PUT THIS BACK IN
 We then identify the conditions allowing the spread of this self-promoting driver, 
 	and evaluate whether a driver of this form could generate a sustained conflict favoring the evolution of suppressors.
-%\yb{ We present a general framework for this problem in equations A1-A4, and evaluate specialized cases below.}
 
@@
 	and refer readers to the previous literature on this model \citep[e.g. ][]{Prout1973,Ubeda2004} for additional results. 
-In this case, assuming the driving allele is deleterious but fully recessive 
+In this case, assuming the driving allele is deleterious in both sexes but fully recessive 
 	(i.e. homozygous fitness equals $w_{BB}=1-s$ and other genotypic fitnesses equal $w_{AA}=w_{AB}=1$), 
 	it always invades because,  when rare, it occurs predominantly in heterozygotes and therefore drives without a fitness cost. 
-However, when $s$ is large drivers cannot fix \citep{Prout1973}, 
-	providing an opportunity for the evolution of
-	drive suppressors, corresponding well to 
-	empirical examples of female meiotic drive \citep[reviewed in ][]{Burt2006}.
+However, when $s$ is large ($s>(2d-1)/(2)$) a driver cannot fix, and
+will be maintained as a protected polymorphism \citep{Prout1973}. The
+        parameter space where the allele can invade but not fix is shown in white
+        in Figure \ref{Invasion_space}A., and where it can fix in the green
+        and orange areas \ref{Invasion_space}B. When the allele is
+        maintained as a polymorphism it providing an opportunity for the evolution of
+	drive suppressors, corresponding well to empirical examples of
+        female meiotic drive \citep[reviewed in ][]{Burt2006}.
 
@@
 	no fitness cost when rare, 
-	a self-promoting driver specifically creates homozygotes by driving. 
+	a self-promoting driver specifically creates homozygote
+        zygotes when it drives in heterozygotes. 
 %, and therefore there is a drive-associated fitness cost even when rare. 
@@
 
-The positive frequency dependence of a self-promoting driver can induce bistability in its spread. 
-Compared to the invasion criteria, parameter space allowing fixations is much broader -- 
-	a common, recessive, self-promoting driver fixes if  $s$ is less than $(2d-1)/2$. 
+The positive frequency dependence of a self-promoting driver can
+induce bistability in its spread, i.e. there is a critical frequency
+above which it can spread and below which it moves towards loss. 
+Compared to the invasion criteria, the bistable parameter space allowing fixations is much broader -- 
+	a common, recessive, self-promoting driver fixes if
+        $s<(2d-1)/2$ approximately. 
 Therefore, when $(2 d - 1)/(4 d)<s<(2d-1)/(2)$, 
@@
 Inclusion of a heterozygous fitness cost (i.e. $w_{AB}=1-h_s$)
-	%(e.g. additive fitness cost) 
 	further constrains the evolution of a self-promoting driver. 
@@
 	reasonably sized populations, and assures that, if they do invade, they will rapidly fix.
-Our model therefore predicts that self-promoting drivers will not be observed as polymorphisms in natural populations.
+Our model therefore predicts that self-promoting drivers will not be
+observed as stable polymorphisms in natural populations.
 
@@
   of this parameter region. The approximate criteria for this allele
-  to fix is given by $s < (2d- 1)/(4d)$. 
+  to fix is given by the dashed line showing $s < (2d- 1)/(4d)$. 
  {\bf B.} In the inset we show trajectories of sperm-dependent female drive

### 2014-05-07 writeup_and_notes/male_dependent_drive.tex
fixed bug in iterate_drive_model.R

@@
 above which it can spread and below which it moves towards loss. 
-Compared to the invasion criteria, the bistable parameter space allowing fixations is much broader -- 
+Compared to the invasion criteria, the parameter space
+allowing the fixation of the allele if it is above a cutoff frequency is much broader -- 
 	a common, recessive, self-promoting driver fixes if

### 2014-05-07 scripts/invasion_analysis.R

@@ -17,7 +17,7 @@ test.invasion.self.prop<-function(d,s.het,s.hom,x=.001,steps=10,sperm.or.male="s
 	if(sperm.or.male=="sperm"){ female.transmission.probs<-make.sperm.dep.female.transmission.prob(D)}
 	if(sperm.or.male=="male"){ female.transmission.probs<-make.male.geno.dep.female.transmission.prob(D) }
 	old.geno.freqs<-iterate.1.locus.drive(s.array,num.iterations=steps,female.transmission.probs=female.transmission.probs,initialize.allele.freqs =c(1-x, 0,x))
-recover()
+#recover()
 	invading<-old.geno.freqs[["allele.freqs"]][steps,3]>old.geno.freqs[["allele.freqs"]][1,3]
 	return(invading)
 }
@@ -190,7 +190,7 @@ legend(x="bottomleft",legend=paste("s=",my.s),col=my.cols,lty=1,lwd=2)
 dev.copy2eps(file=paste(directory,"bistable_x_vs_d_additive_s.eps",sep=""))
 #dev.off()
 
-##############Phase diagram figure
+##############Phase diagram figure for the paper!!!!
 if(FALSE){
 	D<-matrix(1/2,nrow=2,ncol=3,dimnames=list(c("1 or 2","3"),c("12","13","23"))) ##entries are drive coeff of the 2nd allele listed against the 1st.
 	
@@ -230,7 +230,7 @@ text(0.8,0.05,"Simple driver & Self promoter driver can invade & fix",col="white
 lines(d.range,(2*d.range- 1)/2,col="black",lwd=3,lty=1) ##simple driver fixes below this line
 lines(d.range,(2*d.range- 1)/(4*d.range),col="black",lwd=3,lty=2) ##self prom. fixes below this line
 
-text(x=0.95,y=0.95,"A.",cex=2)
+text(x=0.95,y=0.95,"B.",cex=2)
 
 par(fig=c(grconvertX(0.51, from = "user", to = "ndc"), grconvertX(0.75, from = "user", to = "ndc"), grconvertY(0.4, from = "user", to = "ndc"), grconvertY(1, from = "user", to = "ndc")), new = TRUE); 
 par(mar=c(2,2,1,0))
@@ -262,8 +262,33 @@ for(i in 1:3){
[…]

### 2014-05-07 scripts/iterate_drive_model.R

@@ -229,7 +229,7 @@ geno.names<-c("11", "12","13", "22", "23", "33")
 transmitted.allele<-c(1,1,1,2,1,3,2,2,2,3,3,3)
 
 
-figures.of.traj.for.paper(){
+figures.of.traj.for.paper<-function(){
 	layout(t(1:2)); par(mar=c(3,3,1,1))
 	##female driver followed by sperm supressor
 	

### 2014-05-07 scripts/iterate_drive_model.R
fixed bug in iterate_drive_model.R

@@ -441,6 +441,7 @@ examples.for.talk <-function(){
 	
 }
 
+if(FALSE){
 ##
 x=.01;my.x2<-numeric()
 for(i in 1:4000){
@@ -494,4 +495,4 @@ s3=s
  HWEx2<-c(HWEx2,x2)
  HWEx3<-c(HWEx3,x3)
  }
- 
+ }

### 2014-05-07 scripts/homozyg_bistability.R
added homozyg_bistability.R

@@ -1,0 +1,48 @@
+ source("iterate_drive_model.R")
+ 
+
+test.invasion.self.prop<-function(d,s.het,s.hom,x=.001,steps=10,sperm.or.male="sperm"){
+	s.array<-rep(0,6)
+	names(s.array)<-c("11","12","13","23","22","33")
+	s.array["13"]<- s.het
+	s.array["33"]<- s.hom
+	
+	D<-matrix(1/2,nrow=2,ncol=3,dimnames=list(c("1 or 2","3"),c("12","13","23"))) ##entries are drive coeff of the 2nd 
+	D["1 or 2","12"]<-0.5
+	D["1 or 2","13"]<-0.5
+	D["3","12"]<-.5
+	D["3","13"]<-d
+	D["1 or 2","23"]<- 1-0.5
+	D["3","23"]<-0.5
+	stopifnot(sperm.or.male=="sperm" | sperm.or.male=="male")
+	if(sperm.or.male=="sperm"){ female.transmission.probs<-make.sperm.dep.female.transmission.prob(D)}
+	if(sperm.or.male=="male"){ female.transmission.probs<-make.male.geno.dep.female.transmission.prob(D) }
+	old.geno.freqs<-iterate.1.locus.drive(s.array,num.iterations=steps,female.transmission.probs=female.transmission.probs,initialize.allele.freqs =c(1-x, 0,x))
+#recover()
+	invading<-old.geno.freqs[["allele.freqs"]][steps,3]>old.geno.freqs[["allele.freqs"]][1,3]
+	return(invading)
+}
+
+
+#approx.bistable.freq<-(1-2*d+4*d*s)/(2*s*(2*d-1))
+d.range<-seq(0.5,1,length=200)
+s3.range<-seq(0,1,length=200)
+x.range<-seq(0.00001,1,length=100);
+	my.x.bistable<-numeric()
+
+	for(i in 1:length(s3.range)){
+		
+		invasion.grid.bistable<-sapply(d.range,function(d){
+			sapply(x.range,function(x){
+				test.invasion.self.prop(d=d,s.het=0,s.hom=s3.range[i],x=x,step=10)
+			})
+		})
+		cat(i," ")
+		#image(d.range,x.range,invasion.grid.bistable,log="y")
+		invasion.x.lines<-x.range[apply(invasion.grid.bistable,2,function(x){min(which(x))})]
[…]

### 2014-05-07 scripts/invasion_analysis.R

@@ -271,21 +271,22 @@ d.range<-seq(0.5,1,length=200)
 s3.range<-seq(0,1,length=200)
 x.range<-seq(0.00001,1,length=100);
 	my.x.bistable<-numeric()
-	my.s=c(0.01,0.02)
-	x.range<-seq(0,1,length=10)
 
-	for(i in 1:length(my.s)){
+
+
+	for(i in 1:length(s3.range)){
 		
 		invasion.grid.bistable<-sapply(d.range,function(d){
 			sapply(x.range,function(x){
-				test.invasion.self.prop(d=d,s.het=0,s.hom=my.s[i],x=x,step=10)
+				test.invasion.self.prop(d=d,s.het=0,s.hom=s3.range[i],x=x,step=10)
 			})
 		})
-		recover()
+		cat(i," ")
 		#image(d.range,x.range,invasion.grid.bistable,log="y")
 		invasion.x.lines<-x.range[apply(invasion.grid.bistable,2,function(x){min(which(x))})]
 	#	text(d.range[round(length(d.range)/2)],invasion.x.lines[round(length(d.range)/2)],s,col="red")
 		my.x.bistable<-rbind(my.x.bistable,invasion.x.lines)
+		save(file="homozyg_bistability.Robj",my.x.bistable)
 	}
 
 

### 2014-05-07 scripts/invasion_analysis.R
added files associated with homozyg_bistability supp fig

@@ -265,14 +265,14 @@ for(i in 1:3){
 text(x=50,y=0.95,"C.",cex=1)
 dev.copy2eps(file=paste(directory,"invasion_space_recessive_driver.eps",sep=""))
 ##############Phase diagram figure showing bistability 
-
-approx.bistable.freq<-(1-2*d+4*d*s)/(2*s*(2*d-1))
+##done BATCH in homozyg_bistability.R
+#
 d.range<-seq(0.5,1,length=200)
 s3.range<-seq(0,1,length=200)
+if(FALSE){
 x.range<-seq(0.00001,1,length=100);
+
 	my.x.bistable<-numeric()
-
-
 
 	for(i in 1:length(s3.range)){
 		
@@ -288,8 +288,34 @@ x.range<-seq(0.00001,1,length=100);
 		my.x.bistable<-rbind(my.x.bistable,invasion.x.lines)
 		save(file="homozyg_bistability.Robj",my.x.bistable)
 	}
+}
 
+load("homozyg_bistability.Robj")
+layout(t(1:2))
+par(mar=c(3,3,1,1))
+my.x.bistable<-t(my.x.bistable)
+my.x.bistable<-cbind(my.x.bistable,matrix(nrow=200,ncol=67)) ##fill out rest of array 
+image(d.range,s3.range,my.x.bistable,xlab="",ylab="")
+lines(d.range,(2*d.range- 1)/2,col="black",lwd=3,lty=1) ##simple driver fixes below this line
+lines(d.range,(2*d.range- 1)/(4*d.range),col="black",lwd=3,lty=2) ##self prom. fixes below this line
+mtext("Drive coefficient, d",side=1,line=2)
+mtext("selection against homozygotes, s",side=2,line=2)
+contour(d.range,s3.range,my.x.bistable,add=TRUE,lty=2)
+text(0.8,0.6,"Self promoter cannot invade\n no matter what freq.")
+text(0.85,0.1,"Self promoter driver can\n invade & fix when rare",col="white",cex=0.8)
+##HWE
+approx.bistable.freq<-outer(d.range,s3.range,function(d,s){(1-2*d+4*d*s)/(2*s*(2*d-1))})
+approx.bistable.freq<-apply(approx.bistable.freq,1,function(x){x[x>=1]<-NA;x[x<=0]<-0;x})
+image(d.range,s3.range,t(approx.bistable.freq),xlab="",ylab="")
[…]

### 2014-05-07 writeup_and_notes/male_dependent_drive.tex
updated text on bistability, fixed bug in additive bistability

--- 2014-05-05 writeup_and_notes/male_dependent_drive.tex
@@
 induce bistability in its spread, i.e. there is a critical frequency
-above which it can spread and below which it moves towards loss. 
-Compared to the invasion criteria, the bistable parameter space allowing fixations is much broader -- 
-	a common, recessive, self-promoting driver fixes if
-        $s<(2d-1)/2$ approximately. 
-Therefore, when $(2 d - 1)/(4 d)<s<(2d-1)/(2)$, 
-	a recessive, self-promoting driver will deterministically fix if drift, mutation, 
-	or migration pressure bring its frequency above $(1-2d+4ds)/(2s(2d-1))$, 
-	but it will be lost otherwise.
+above which it can spread quickly to fixation and below which it moves towards loss (an
+unstable equilibrium frequency). In 
+the appendix we devlop a simple
+approximation to this bistability 
+This bistability occurs in the region of parameter space in Figure
+\ref{Invasion_space}. In Supplementary Figure
+\ref{Bistab_homozyg_cost_fig} we further explore this bistability and
+show that the unstable equilibrium frequency is high across much of
+this parameter space, and so in practice the self promoting allele
+only invade in the orange space. (In the appendix we develop a simple
+approximation to the unstable equilibrium frequency
+\eqref{bistab_homozy_approx} that is also shown in  Supplementary Figure
+\ref{Bistab_homozyg_cost_fig}).
+% Compared to the invasion criteria, the parameter space
+% allowing the fixation of the allele if it is above a cutoff frequency is much broader -- 
+% 	a common, recessive, self-promoting driver fixes if
+%         $s<(2d-1)/2$ approximately. 
+% Therefore, when $(2 d - 1)/(4 d)<s<(2d-1)/(2)$, 
+% 	a recessive, self-promoting driver will deterministically fix if drift, mutation, 
+% 	or migration pressure bring its frequency above $(1-2d+4ds)/(2s(2d-1))$, 
+% 	but it will be lost otherwise.
 Inclusion of a heterozygous fitness cost (i.e. $w_{AB}=1-h_s$)
-	further constrains the evolution of a self-promoting driver. 
-This case also displays bistability -- with any heterozygous fitness cost, a rare self-promoting driver cannot spread. 
-However, if $s < d - 1/2+s_h(3-2d)/2$, 
-	it will fix deterministically if its 
+strongly constrains the evolution of a self-promoting driver. 
+In fact with any heterozygous fitness cost, a rare self-promoting
+driver cannot spread deterministically. However, this case also displays bistability -- 
+if $s < d - 1/2+s_h(3-2d)/2$ approximately, it will fix deterministically if its 
 	frequency is greater than a threshold, $f_B^*$ [Equation \eqref{thresh2}, and \yb{FIGURES SXX}].
@@
 
-\yb{\paragraph{Recessive fitness cost of self-promoting driver}
+\paragraph{Recessive fitness cost of self-promoting driver}
 When the self-promoting driver has a recessive fitness cost, $s$ genotype frequencies after selection and drive equal
@@
 \end{equation}
-where $F$ is the deviation from Hardy-Weinberg expectations, and $f_A$ is the frequency of the non-driving allele. 
+where $F$ is the deviation from Hardy-Weinberg expectations (HWE), and $f_A$ is the frequency of the non-driving allele. 
 For analytical solutions presented above, we assume that $F$ equals zero.
-}
+
+We used this HWE approximation to investifgate the bistability of the
+invasion criteria. Under this approximation we find that a common, recessive, self-promoting driver fixes if
+ $s<(2d-1)/2$. Furthermore, when $(2 d - 1)/(4 d)<s<(2d-1)/(2)$, a
+ recessive, self-promoting driver will deterministically fix if drift,
+ mutation, or migration pressure bring its frequency above approximately
+\begin{equation} 
+(1-2d+4ds)/(2s(2d-1)) \label{bistab_homozy_approx}
+\end{equation}
+ but it will be lost if it is introduced below this frequency
+ (i.e. this is an unstable equilibrium point). In Supplementary Figure \ref{Bistab_homozyg_cost_fig} we show
+ this approximation condition for bistability andd how it compared to
+ that obtained using the recursion. This equation
+ \eqref{bistab_homozy_approx} offers a reasonable, but imperfect
+ approximation 
+and seems to serve as a lower bound to the
+true unstable equilibrium frequency.
 
@@
 Assuming no deviation from Hardy-Weinberg a self-promoting driver with a heterozygous fitness cost deterministically fixes if it's frequency exceeds some threshold, $f_B^*$.
-
 \begin{equation}
@@
 
-\paragraph{Dominance} In the supplement we also explore a model where
+\paragraph{Dominance} we also explored a model where
 the genotype of the male who produced the fertilizing sperm affects female drive,
@@
 
-\gc{Alternatively we could give the table below:}
+% \gc{Alternatively we could give the table below:}
 
-\begin{table}
-\begin{tabular}{lllll}
-\hline 
-  &  \multicolumn{4}{c}{Sperm genotype}   \\
-  &  & A & B & B'  \\
-$\Venus$ geno. & Freq.  & $x_{AA} + \frac{1}{2} x_{AB}+ \frac{1}{2} x_{AB'} $ & $ w_{BB}  x_{BB} +   w_{BB'} \frac{1}{2} x_{BB'}+
-\frac{1}{2} x_{AB}$ & $ w_{B'B'}  x_{B'B'} +   w_{BB'} \frac{1}{2}
-x_{BB'}+ \frac{1}{2} x_{AB'}  $\\
-\hline\\
-AA & $x_{AA}$ &  & & \\
-AB & $x_{AB}$ &  & & \\
-AB' & $x_{AB'}$ &  & & \\
-BB' & $w_{BB'} x_{BB'}$ &  & & \\
-BB & $w_{B'B'} x_{BB}$ &  & & \\
-B'B' & $w_{B'B'} x_{B'B'}$ &  & & \\
-\end{tabular}
-\caption{ALTERNATIVELY I COULD FILL IN THIS MATRIX  with genotype }
-\end{table}
+% \begin{table}
+% \begin{tabular}{lllll}
+% \hline 
+%   &  \multicolumn{4}{c}{Sperm genotype}   \\
+%   &  & A & B & B'  \\
+% $\Venus$ geno. & Freq.  & $x_{AA} + \frac{1}{2} x_{AB}+ \frac{1}{2} x_{AB'} $ & $ w_{BB}  x_{BB} +   w_{BB'} \frac{1}{2} x_{BB'}+
+% \frac{1}{2} x_{AB}$ & $ w_{B'B'}  x_{B'B'} +   w_{BB'} \frac{1}{2}
+% x_{BB'}+ \frac{1}{2} x_{AB'}  $\\
+% \hline\\
+% AA & $x_{AA}$ &  & & \\
+% AB & $x_{AB}$ &  & & \\
+% AB' & $x_{AB'}$ &  & & \\
+% BB' & $w_{BB'} x_{BB'}$ &  & & \\
+% BB & $w_{B'B'} x_{BB}$ &  & & \\
+% B'B' & $w_{B'B'} x_{B'B'}$ &  & & \\
+% \end{tabular}
+% \caption{ALTERNATIVELY I COULD FILL IN THIS MATRIX  with genotype }
+% \end{table}
 \bibliography{refs}
@@
 \section*{Supplementary Stuff}
+
+\begin{figure}
+	\rotatebox{90}{
+\includegraphics[width = 0.5
+          \textwidth]{Scripts/homozyg_bistability.eps}}
+\caption{Invasion analysis for a self promoting female meiotic drive allele with
+  recessive costs (selection coefficient $s$), showing the region of
+  bistability. The colors, and the thin dashed contours, indicate the
+  frequency the allele must be above to invade the population (note
+  that these alleles reach fixation conditional on invading). In the
+  white area the allele can not invade ($x \geq1$), in the solid red
+  area the allele can invade and fix when rare. In the left panel we
+  show the results obtained by a grid search using the recursion, on
+  the right we show the approximation obtained assuming that HWE
+  holds. }  
+\label{Bistab_homozyg_cost_fig}
+\end{figure}
 
@@
 \end{document}
-
-\section*{OUTLINE}
-\yb{YB: To me its (1) male enhancement of female drive cannot maintain a stable polymorphisms, and (2) An allele in sperm can evolve to suppress its drive in females.}
-show that such alleles:
-\begin{enumerate}
-\item can't be balanced, \\
-\item and homozygous problems are tested out at low freq.  \\
-\item Any heterozygous problems, leads to a bistable allele\\
-\item If these alleles take off they speed through to fixation\\
-\item If allele has any drive ability in absence of sperm effect that is what allows it to enter the population
-and sperm effect isn't a further cause of conflict. What if anything do we mean by this?\\
-\item PERHAPS HERE WE INTRODUCE A SELF-RESTRAINING ALLELE
-\end{enumerate}
-
-Conclusion, such alleles are unlikely to cause evolution of female supressors, they test hemselves in a homozygous
-state when they enter the population, and sweep quickly (all the way to fixation) if they enter the pop at all.\\
-
-\begin{enumerate}
-\item Setup a drive-selection balanced polymorphism in std. drive model. 
-Do this by imagining the sperm-influence allele arising on the background of the driver, 
-so the allele has drive capabilities, and can have had time to evolve new biology.
-
-Multilocus drivers held together with inversions, may offer a relatively large mutational/functional target.
-
-Evolving on the new background means that the allele suffers the fitness consequences of the 
-driver.  See A in Figure \ref{Eggsperm_3_allele_cartoon} \\
-\item Sperm-based enhancers of drive can't invade (can they in some situations?). \\
-\item Intuition is that the driver has already driven to a frequency 
-where it is held in check by its cost in homozygotes. The sperm allele 
-thus can't really help as it creates zygotes which suffer the homozygous fitness consequences.
-\end{enumerate}
-
-\subsection{So what can evolve?}
-So what can happen?
-\begin{enumerate}
-\item Alleles that arise in linkage with drive systems, which when in sperm switch off drive, 
-can spread. They benefit from drive, but avoid some of the
-consequences ( (Bi) in Figure \ref{Eggsperm_3_allele_cartoon} ). Overall as a side product they are benefiting all in pop.\\
-\item Presumably alleles that actually switch the allele that drives may do even better? As they'd end up in 
-hets. Although they'd not drive, so hard to say. YB: There is evidence that In distorts meiosis in the other direction (they still drive when rare i.e. when not fused with drive supp sperm) \\
-\item Alleles that cause sperm to switch off drive that arise on other background or unlinked to the drive system
-are selected, and spread as fast as female supressors of drive ( (Bii) in Figure \ref{Eggsperm_3_allele_cartoon} )..\\
-\item Alleles that in females facilitate the action of sperm supressors of drive (or vis versa) can spread. Haven't actually checked this.\\
-\end{enumerate}
-
-\section*{Conclusions.}

### 2014-05-07 scripts/invasion_analysis.R

@@ -178,12 +178,25 @@ if(FALSE){
 	
 	save(file=paste(directory,"bistable_invasion_grid.Robj",sep=""),my.x.bistable,my.s)
 }
-load(file=paste(directory,"bistable_invasion_grid.Robj",sep=""))
-plot(range(d.range),c(10^(-6),1),type="n",log="y",xlab="d",ylab="x bistable cutoff freq.")
+show(load(file=paste(directory,"bistable_invasion_grid.Robj",sep="")))
+par(mar=c(3,3,1,1))
+plot(range(d.range),c(10^(-6),1),type="n",log="y",xlab="",ylab="")
+mtext(side=1,line=2,"Drive coefficient, d")
+mtext(side=2,line=2,"Unstable equilibrium frequency")
+	my.cols<-brewer.pal(length(my.s),name="Dark2")
+
 for(i in 1:length(my.s)){
 	lines(d.range,my.x.bistable[i,],col=my.cols[i],lwd=2)
-	my.cols<-brewer.pal(length(my.s),name="Dark2")
+	
+	sh<-my.s[i]
+	s<-my.s[i]*2
+	a1<-1-2*d.range*sh+4*d.range*s-2*d.range-3*sh;
+	a2<-2*(s-sh)*(2*d.range-1)
+
+	f.HWE.bistab<- a1+sqrt(a1^2+8*sh*a2)/(2*a2)
+	lines(d.range,f.HWE.bistab,col=my.cols[i],lwd=2,lty=2)
 }
+
 
 legend(x="bottomleft",legend=paste("s=",my.s),col=my.cols,lty=1,lwd=2)
 
@@ -298,19 +311,21 @@ my.x.bistable<-cbind(my.x.bistable,matrix(nrow=200,ncol=67)) ##fill out rest of 
 image(d.range,s3.range,my.x.bistable,xlab="",ylab="")
 lines(d.range,(2*d.range- 1)/2,col="black",lwd=3,lty=1) ##simple driver fixes below this line
 lines(d.range,(2*d.range- 1)/(4*d.range),col="black",lwd=3,lty=2) ##self prom. fixes below this line
+mtext("Bistability of invasion w. self promoter w. homozyg. cost",side=3,line=0)
 mtext("Drive coefficient, d",side=1,line=2)
-mtext("selection against homozygotes, s",side=2,line=2)
+mtext("Selection against homozygotes, s",side=2,line=2)
 contour(d.range,s3.range,my.x.bistable,add=TRUE,lty=2)
[…]

### 2014-05-08 writeup_and_notes/male_dependent_drive.tex
added bistable additive figure and code
added male geno. control figure effect_of_dominance_on_invasion_space_four_graph.eps

@@
 
-\section*{Supplementary Stuff}
+\section*{Supplementary Material}
 
 \begin{figure}
-	\rotatebox{90}{
+
 \includegraphics[width = 0.5
-          \textwidth]{Scripts/homozyg_bistability.eps}}
+          \textwidth]{../Scripts/homozyg_bistability.eps}
 \caption{Invasion analysis for a self promoting female meiotic drive allele with
@@
 \includegraphics[width = 0.8 \textwidth]{Figures/bistable_x_vs_d_additive_s.eps} 
-\caption{Bistable of a drive allele when selection cost (s) acts
-  additively, for an allele whose effect
- on female meoisis is mediated by sperm allele. }  \label{bistable_additive}
+\caption{The unstable equilibrium frequency for a self promoting
+  female meiotic drive allele with an additive cost ($s_h=s/2$) as a
+  function of the drive parameter. The solid line shows results
+  obtained using the recursion, the dots our approximation given by \eqref{thresh2}/}  \label{bistable_additive}
 \end{figure}

### 2014-05-08 scripts/invasion_analysis.R

@@ -193,12 +193,12 @@ for(i in 1:length(my.s)){
 	a1<-1-2*d.range*sh+4*d.range*s-2*d.range-3*sh;
 	a2<-2*(s-sh)*(2*d.range-1)
 
-	f.HWE.bistab<- a1+sqrt(a1^2+8*sh*a2)/(2*a2)
-	lines(d.range,f.HWE.bistab,col=my.cols[i],lwd=2,lty=2)
+	f.HWE.bistab<- (a1+sqrt(a1^2+8*sh*a2))/(2*a2)
+	points(d.range[c(1:4,seq(5,200,by=5))],f.HWE.bistab[c(1:4,seq(5,200,by=5))],col=my.cols[i],lwd=2,lty=2,pch=19,cex=1)
 }
 
 
-legend(x="bottomleft",legend=paste("s=",my.s),col=my.cols,lty=1,lwd=2)
+legend(x="bottomleft",legend=paste("s=",2*my.s),col=my.cols,lty=1,lwd=2)
 
 dev.copy2eps(file=paste(directory,"bistable_x_vs_d_additive_s.eps",sep=""))
 #dev.off()
@@ -393,13 +393,28 @@ dev.off()
 save(file=paste(directory,"effect_of_dominance_male_control.Robj",sep=""),invasion.w.dom,fixation.w.dom)
 }
 load(file=paste(directory,"effect_of_dominance_male_control.Robj",sep=""))
-
 my.cols<-brewer.pal(nrow(invasion.w.dom),name="Dark2")
 plot(c(0.5,1),c(0,0.5),type="n",,xlab="Drive coefficient",ylab="selection coeff against homozygotes")
 sapply(1:nrow(invasion.w.dom),function(i){lines(d.range,invasion.w.dom[i,],lty=1,col=my.cols[i])})
 sapply(1:nrow(invasion.w.dom),function(i){lines(d.range,fixation.w.dom[i,],lty=2,col=my.cols[i])})
 legend("topleft",lty=c(1,2,rep(NA,nrow(invasion.w.dom))),pch=c(NA,NA,rep(19,nrow(invasion.w.dom))),col=c(rep("black",2),my.cols),legend=c("invasion line","fixation line",paste("dominance = ",seq(0,1,length=5))))
 dev.copy2eps(file=paste(directory,"effect_of_dominance_on_invasion_space_one_graph.eps",sep=""))
+
+layout(t(1:5))
+par(mar=c(3,3,1,0.1))
[…]

### 2014-05-08 writeup_and_notes/male_dependent_drive.tex
fixed tupos, modified the bistabilty section

@@
 it is potentially open to sabotage by a virtual stranger -- a haploid sperm genome.
-That is, in many animal species, female meiosis is competed only after fertilization \citep{Masui_book}, 
+That is, in many animal species, female meiosis is completed only after fertilization \citep{Masui_book}, 
 creating ample opportunity for interaction between the sperm and female meiotic machinery. 
@@
      could presumably rapidly spread through a population. 
-At first sight, female meiosis appears primed for exploitation by selfish sperm systems.  
+At first sight, female meiosis appears primed for exploitation by such selfish systems.  
 Why then is the requirement of fertilization to complete female meiosis so ubiquitous?
@@
 The stage at which female meiosis requires fertilization varies widely across taxa, and in a few animals, 
-	female meiosis is complete upon ovulation (that is fertilization is not required for female meiosis \citep[see table 1 of ref 
+	female meiosis is complete upon ovulation (that is, fertilization is not required for female meiosis \citep[see table 1 of ref 
 ][]{Masui_book}).
@@
 of a self-promoting meiotic driver,  we modify this standard model such that the driver is only effective when fertilized by a sperm carrying that allele (see Figure
-\ref{Eggsperm_2_allele_cartoon}). 
+\ref{Eggsperm_2_allele_cartoon})
 \footnote{the timing of fertilization relative to female meiosis places another constraint on $d$, for example, if fertilization (and therefore, sperm dependent drive) takes place at MII (as in mammals),
@@
         and orange areas \ref{Invasion_space}B. When the allele is
-        maintained as a polymorphism it providing an opportunity for the evolution of
+        maintained as a polymorphism, it provides an opportunity for the evolution of
 	drive suppressors, corresponding well to empirical examples of
@@
 	a self-promoting driver specifically creates homozygote
-        zygotes when it drives in heterozygotes. 
+        zygotes when it drives. %in heterozygotes. 
 %, and therefore there is a drive-associated fitness cost even when rare. 
@@
 
+%I SUBSTANTIALLY REWROTE THIS
+\yb{
 The positive frequency dependence of a self-promoting driver can
-induce bistability in its spread, i.e. there is a critical frequency
-above which it can spread quickly to fixation and below which it moves towards loss (an
-unstable equilibrium frequency). In 
-the appendix we devlop a simple
-approximation to this bistability 
-This bistability occurs in the region of parameter space in Figure
-\ref{Invasion_space}. In Supplementary Figure
-\ref{Bistab_homozyg_cost_fig} we further explore this bistability and
-show that the unstable equilibrium frequency is high across much of
-this parameter space, and so in practice the self promoting allele
-only invade in the orange space. (In the appendix we develop a simple
-approximation to the unstable equilibrium frequency
-\eqref{bistab_homozy_approx} that is also shown in  Supplementary Figure
-\ref{Bistab_homozyg_cost_fig}).
+induce bistability in its spread.
+%OLD GRAHAM WORDS
+ %, i.e. there is a critical frequency
+%above which it can spread quickly to fixation and below which it moves towards loss (an
+%unstable equilibrium frequency). In 
+%the appendix, we develop a simple
+%approximation to this bistability. 
+When $s$ is too large to allow a rare driver to increase in frequency, but small enough to allow a common driver to fix (the inequality in EQUATION XXX),
+there is  an unstable equilibrium frequency (approximated in EQUATION XXX)  
+	above which it can spread quickly to fixation and below which it moves towards loss 
+	(see Figure \ref{Bistab_homozyg_cost_fig} for exact results).	
+%OLD GRAHAM WORDS
+%This bistability occurs in the region of parameter space in Figure
+%\ref{Invasion_space}. In Supplementary Figure
+%\ref{Bistab_homozyg_cost_fig} we further explore this bistability and
+%show that the unstable equilibrium frequency is high across much of
+%this parameter space, and so in practice the self promoting allele
+%only invade in the orange space. (In the appendix we develop a simple
+%approximation to the unstable equilibrium frequency
+%\eqref{bistab_homozy_approx} that is also shown in  Supplementary Figure
+%\ref{Bistab_homozyg_cost_fig}).
+This unstable equilibrium frequency is high across much of
+	parameter space, likely too high to be reached by drift, 
+	and therefore the more restrictive invasion criteria rather than the fixation criteria, will  likely determine the fate of self promoting drivers. 
+}
+
+%OLD YANIV WORDS
 % Compared to the invasion criteria, the parameter space
@@
 Inclusion of a heterozygous fitness cost (i.e. $w_{AB}=1-h_s$)
-strongly constrains the evolution of a self-promoting driver. 
-In fact with any heterozygous fitness cost, a rare self-promoting
-driver cannot spread deterministically. However, this case also displays bistability -- 
+further constrains the evolution of a self-promoting driver. 
+In fact, with any heterozygous fitness cost, a rare self-promoting
+driver cannot spread deterministically. 
+However, this case also displays bistability -- 
 if $s < d - 1/2+s_h(3-2d)/2$ approximately, it will fix deterministically if its 

### 2014-05-09 writeup_and_notes/male_dependent_drive.tex
edited polymorphic figure

--- 2014-05-08 writeup_and_notes/male_dependent_drive.tex
@@
 \begin{figure}
-\includegraphics[width = 0.8 \textwidth]{Figures/effect_of_dominance_on_invasion_space_one_graph.eps} 
+\includegraphics[width = 0.8 \textwidth]{../Scripts/effect_of_dominance_on_invasion_space_four_graph.eps} 
 \caption{Invasion analysis of an allele whose effect
  on female meoisis is mediated by the genotype of the fertilizing
- male. The area below the solid/dotted line is the parameter region
- where the allele can invade/fix (from $1/1000$ and $999/1000$
- \gc{maybe should lower this}). Note
- that all of the fixation lines are on top of each other as dominance
- parameter makes little difference here. The allele will be
- polymorphic in the parameter space where the solid (invasion) line
- falls above the dotted (fixation) line. Note that this is a very
- narrow parameter range, and only exists if the allele has a
- reasonable degree on dominance}  \label{Effect_of_dominance}
+ male.  As above wee assume a heterozygous female will transmit the $B$ allele 
+  with probabilities  $\frac{1}{2}$,  $\frac{1}{2} + h(d-\frac{1}{2}) $ or $d$, 
+ if she mated with a n $AA$, $AB$, or $BB$ male,  respectively.  The allele suffers a recessive fitness cost $s$.  
+ The four panels correspond to different dominance relationships.
+In the parameter space below the invasion (solid) line the self promotor
+ driver can invade. In the parameter space below the fixation (long
+ dash) line the self promotor can fix in the population. In the last
+ two panels the invasion line is above the fixation line and so the
+ allele can be maintained as a polymorphism in that thin slice of
+ parameter space between the two lines. These results were 
+  obtained using the recursion. In the final panel we show the
+  fixation line (small dashes) as predicted by our HWE approximation.
+}  \label{Effect_of_dominance}
 \end{figure}

### 2014-05-09 writeup_and_notes/male_dependent_drive.tex
edited male geno. text

@@
 %I SUBSTANTIALLY REWROTE THIS
-\yb{
 The positive frequency dependence of a self-promoting driver can
@@
 	and therefore the more restrictive invasion criteria rather than the fixation criteria, will  likely determine the fate of self promoting drivers.
-}
 
@@
 This male-genotype dependent model requires one additional parameter, as we exchange $d$ in the sperm dependent case for $d_{het}$ and $d_{hom}$ which describe the transmission of the drive allele in a heterozygous female mating with males heterozygous and homozygous for the self-promoting drive allele, respectively.
-When $d_{het}$ is large relative to $d_{hom}$, a narrow sliver of parameter space sustains a polymorphism (see Figure S\ref{Effect_of_dominance}), presumably because positive frequency dependence diminishes with the dominance of the male effect driver.
+However, when
+the distorting effect of genotype displays strong dominance in its
+effect on female meiosis ($d_{het}$ is close to $d_{hom}$), 
+there is a a narrow sliver of parameter space sustains a polymorphism
+when the cost of the drive is recessive
+(see Figure S\ref{Effect_of_dominance}). 
+Presumably this zone of polymorphism comes about because positive frequency dependence diminishes with the dominance of the male effect driver. 
 As an analytical example, consider the case of male-genotype dependent deleterious recessive female driver 
@@
 Intermediate values of $s$, which represent a small proportion of parameter space can maintain a polymorphism. 
+However, it does not seem particularly realistic to think that the
+effect of the drive allele would be dominant in its action through the
+male genotype, while the cost would be recessive. Thus we regard this
+case as interesting but not very general.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%
-\yb{********* DO WE NEED TO DO THE MALE GENOTYPE THING HERE TOO? *********}
 
@@
 avoid the production of unfit homozygote zygotes leaving it trapped in
-the population. \gc{Perhaps we should show the equil. freq. in the
-  fig, as I'm guessing it might be rare}
+the population.
 
@@
   obtained using the recursion. In the final panel we show the
-  fixation line (small dashes) as predicted by our HWE approximation.
+  fixation line (small dashes) as predicted by our HWE approximation
+  ($(d_{het}-1/2)/d_{het}$), see the `dominance' section of the
+  Appendix for more details.
 }  \label{Effect_of_dominance}

### 2014-05-10 writeup_and_notes/male_dependent_drive.tex
shortened and modified male-geno case, also hyphenated all self-promo..

@@
 	parameter space, likely too high to be reached by drift, 
-	and therefore the more restrictive invasion criteria rather than the fixation criteria, will  likely determine the fate of self promoting drivers.
+	and therefore the more restrictive invasion criteria rather than the fixation criteria, will  likely determine the fate of self-promoting drivers.
 
@@
 when the cost of the drive is recessive
-(see Figure S\ref{Effect_of_dominance}). 
+(see Figure S\ref{Effect_of_dominance}, \yb{and Equations XXX in the Appendix}).  
 Presumably this zone of polymorphism comes about because positive frequency dependence diminishes with the dominance of the male effect driver. 
-As an analytical example, consider the case of male-genotype dependent deleterious recessive female driver 
-	(assuming, as above, $w_{BB}=1-s$, and $w_{AA}=w_{AB}=1$). 
-Here, a rare driver invades when $s$ is less than $(d_{het}-1/2)/d_{het}$, and fixes when $s$ is less than $d_{hom} - 1/2$.
-Intermediate values of $s$, which represent a small proportion of parameter space can maintain a polymorphism. 
-However, it does not seem particularly realistic to think that the
+%As an analytical example, consider the case of male-genotype dependent deleterious recessive female driver (assuming, as above, $w_{BB}=1-s$, and $w_{AA}=w_{AB}=1$). 
+%Here, a rare driver invades when $s$ is less than $(d_{het}-1/2)/d_{het}$, and fixes when $s$ is less than $d_{hom} - 1/2$.
+%Intermediate values of $s$, which represent a small proportion of parameter space can maintain a polymorphism. 
+While mathematically interesting, it does not seem particularly realistic to think that the
 effect of the drive allele would be dominant in its action through the
-male genotype, while the cost would be recessive. Thus we regard this
-case as interesting but not very general.
+male genotype, while the cost would be recessive. 
+Therefore, although this model can sustain a polymorphism, the lack of biological reality underlying the small portion of parameter values required for this polymorphism make us doubt it general applicability.
+%Thus we regard this case as interesting but not very general.
 
@@
           \textwidth]{../Scripts/homozyg_bistability.eps}
-\caption{Invasion analysis for a self promoting female meiotic drive allele with
+\caption{Invasion analysis for a self-promoting female meiotic drive allele with
   recessive costs (selection coefficient $s$), showing the region of
@@
 \includegraphics[width = 0.8 \textwidth]{Figures/bistable_x_vs_d_additive_s.eps} 
-\caption{The unstable equilibrium frequency for a self promoting
+\caption{The unstable equilibrium frequency for a self-promoting
   female meiotic drive allele with an additive cost ($s_h=s/2$) as a
@@
  The four panels correspond to different dominance relationships.
-In the parameter space below the invasion (solid) line the self promotor
+In the parameter space below the invasion (solid) line the self-promotor
  driver can invade. In the parameter space below the fixation (long
- dash) line the self promotor can fix in the population. In the last
+ dash) line the self-promotor can fix in the population. In the last
  two panels the invasion line is above the fixation line and so the

### 2014-05-12 writeup_and_notes/male_dependent_drive.tex
edited text

--- 2014-05-09 writeup_and_notes/male_dependent_drive.tex
@@
 	diploid genotype (perhaps due to proteins and RNAs packaged into the sperm), rather than sperm haplotype.
-%GRAHAM I HAVE REWRITTEN THIS SECTION TO REFLECT MATHS AND MY PARAMETERIZATION. WE CAN WORK ON SYNCHING OUR NOTATION
-%This adds an additional parameter to the system -- the dominance of the drive enhancer in heterozygous males, 
-%	when this dominance is high, a narrow sliver of parameter space sustaining a polymorphism emerges (see Figure S\ref{Effect_of_dominance})
- %presumably because positive frequency dependence diminishes with the dominance of the male effect driver.
 This male-genotype dependent model requires one additional parameter, as we exchange $d$ in the sperm dependent case for $d_{het}$ and $d_{hom}$ which describe the transmission of the drive allele in a heterozygous female mating with males heterozygous and homozygous for the self-promoting drive allele, respectively.
@@
 case as interesting but not very general.
-
-%\gc{Add short paragraph about sex chromosomes here? XY systems don't care. ZW systems (males ZZ, females ZW). If there is
-%sex ratio distortion how could this play out?}
 
@@
  with  an alternative model of two linked loci - one female driver and the other a sperm allele which modifies the effect of drive upon fertilization. 
-%Here, we examine an alternative scenario (and more realistic) in which
 In this model, a female meiotic driver with no sperm-dependency, initially reaches drive-viability equilibrium (with two alleles
@@
         linked to this balanced locus (effectively creating a third
-        allele/haplotype at this locus).
-Assuming tight linkage between female driver and sperm modifier is consistent with the nature of well characterized meiotic drive systems 
-	which are often maintained as polymorphic inversions with numerous linked modifiers (CITE).
+        allele/haplotype at this locus). Tight linkage offers the best
+        chance for a collaboration to evolve between a driver and
+       an enhancer of the drive, as recombination breaks up drive haplotypes \citep{Thomson1974,Charlesworth1978,Haig1991}. 
+In addition tight linkage between female driver and sperm modifier is consistent with the nature of well characterized drive systems which are often maintained as polymorphic inversions with numerous linked modifiers (CITE).
 
@@
 	three allele system that allows the sperm-acting drive facilitator $B^{+}$ to
-	invade the population. \yb{MATHS?}
+	invade the population.
 
@@
         $ln^-$ allele is over transmitted through female meiosis in $ln^-/ln^+$ heterozygote when fertilized by $ln^+$
-        sperm (CITE cite{agulnick, Pomiankowski1993}).   
+        sperm \cite{agulnick, Pomiankowski1993}.   
 More generally, alleles that act through sperm to
@@
 %asymmetries that female drivers exploit.
-
-\gc{Is there some nice way to work in green beard ideas}
 

### 2014-05-12 writeup_and_notes/male_dependent_drive.tex

--- 2014-05-10 writeup_and_notes/male_dependent_drive.tex
@@
 %Thus we regard this case as interesting but not very general.
-
-%\gc{Add short paragraph about sex chromosomes here? XY systems don't care. ZW systems (males ZZ, females ZW). If there is
-%sex ratio distortion how could this play out?}
 

### 2014-05-12 writeup_and_notes/male_dependent_drive.tex
edits [with emphasis on the two-locus models], Added abstract draft.

@@
 {\bf Abstract:}
+Because only one of four products of female meiosis is transmitted to offspring, oogenesis is particularly
+	 susceptible to genomic conflicts arising when alleles gains in frequency at a cost to organismal fitness. 
+In fact, many facets of oogenesis have been interpreted as mechanisms of protection against genomic outlaws. 
+Why then is female meiosis often left uncompleted until after fertilization - 
+	potentially providing an opportunity for sperm to meddle with its outcome and facilitate meiotic drive? 
+The population genetic theory presented herein suggests that
+	sperm nearly always evolve to increase the fairness of female meiosis when possible. 
+These results are constant with current knowledge of sperm-dependent meiotic drivers, 
+	and suggest that the requirement of fertilization for the completion of meiosis may represent 
+	yet another mechanism of female meiosis designed to prevent meiotic drive.
 \newpage
@@
 probability  $d > 1/2$, regardless of sperm genotype. To depict a case
-of a self-promoting meiotic driver,  we modify this standard model such that the driver is only effective when fertilized by a sperm carrying that allele (see Figure
-\ref{Eggsperm_2_allele_cartoon})
+of a self-promoting meiotic driver,  we modify this standard model such 
+	that the driver is only effective when fertilized by a sperm carrying that allele (see Figure \ref{Eggsperm_2_allele_cartoon})
 \footnote{the timing of fertilization relative to female meiosis places another constraint on $d$, for example, if fertilization (and therefore, sperm dependent drive) takes place at MII (as in mammals),
@@
         discussion]{Buckler1999}. }. %PUT THIS BACK IN
-We then identify the conditions allowing the spread of this self-promoting driver, 
+We then identify the conditions allowing for the spread of this self-promoting driver, 
 	and evaluate whether a driver of this form could generate a sustained conflict favoring the evolution of suppressors.
@@
 In this case, assuming the driving allele is deleterious in both sexes but fully recessive 
-	(i.e. homozygous fitness equals $w_{BB}=1-s$ and other genotypic fitnesses equal $w_{AA}=w_{AB}=1$), 
+	(i.e. the fitness of drive homozygotes equals $w_{BB}=1-s$ and other genotypic fitnesses equal $w_{AA}=w_{AB}=1$), 
 	it always invades because,  when rare, it occurs predominantly in heterozygotes and therefore drives without a fitness cost. 
@@
         parameter space where the allele can invade but not fix is shown in white
-        in Figure \ref{Invasion_space}A., and where it can fix in the green
-        and orange areas \ref{Invasion_space}B. When the allele is
+        in Figure \ref{Invasion_space}A, and where it can fix in the green
+        and orange areasin Figure \ref{Invasion_space}B. When the allele is
         maintained as a polymorphism, it provides an opportunity for the evolution of
@@
 	parameter space, likely too high to be reached by drift, 
-	and therefore the more restrictive invasion criteria rather than the fixation criteria, will  likely determine the fate of self-promoting drivers.
+	and therefore  the fate of self-promoting drivers  will  likely be determined
+	by the more restrictive invasion criteria rather than the fixation criteria.
 
@@
 However, this case also displays bistability -- 
-if $s < d - 1/2+s_h(3-2d)/2$ approximately, it will fix deterministically if its 
-	frequency is greater than a threshold, $f_B^*$ [Equation \eqref{thresh2}, and \yb{FIGURES SXX}].
+	mathematical approximation assuming Hardy-Weinberg reveal that 
+	if $s < d - 1/2+s_h(3-2d)/2$, it will fix deterministically if its 
+	frequency is greater than a threshold, $f_B^*$ [Equation \eqref{thresh2}. 
+	See \yb{FIGURES SXX} for exact results].
 %	because it pays the  heterozygous fitness cost but drives
@@
 This male-genotype dependent model requires one additional parameter, as we exchange $d$ in the sperm dependent case for $d_{het}$ and $d_{hom}$ which describe the transmission of the drive allele in a heterozygous female mating with males heterozygous and homozygous for the self-promoting drive allele, respectively.
-However, when
-the distorting effect of genotype displays strong dominance in its
+Here, a rare driver invades when $s$ is less than $(d_{het}-1/2)/d_{het}$, and usually fixes when it invades.
+	% fixes when $s$ is less than $d_{hom} - 1/2$.
+However, when the distorting effect of genotype displays strong dominance in its
 	effect on female meiosis ($d_{het}$ is close to $d_{hom}$), 
@@
 	(see Figure S\ref{Effect_of_dominance}, \yb{and Equations XXX in the Appendix}).  
-Presumably this zone of polymorphism comes about because positive frequency dependence diminishes with the dominance of the male effect driver. 
+%Presumably this zone of polymorphism comes about because positive frequency 
+%	dependence diminishes with the dominance of the male effect driver. 
 %As an analytical example, consider the case of male-genotype dependent deleterious recessive female driver (assuming, as above, $w_{BB}=1-s$, and $w_{AA}=w_{AB}=1$). 
@@
 	male genotype, while the cost would be recessive. 
-Therefore, although this model can sustain a polymorphism, the lack of biological reality underlying the small portion of parameter values required for this polymorphism make us doubt it general applicability.
+Therefore, although this model can sustain a polymorphism, 
+	the lack of biological reality underlying the small portion of parameter values 
+	required for this polymorphism make us doubt it general applicability. 
 %Thus we regard this case as interesting but not very general.
@@
 
-Above, we explored a model were a driver that drove in females when signaled by a complementary allele in sperm.  
-We complement this single locus model of an allele with pleiotropic effects in both sexes, 
- with  an alternative model of two linked loci - one female driver and the other a sperm allele which modifies the effect of drive upon fertilization. 
-In this model, a female meiotic driver with no sperm-dependency, initially reaches drive-viability equilibrium (with two alleles
+Above, we explored a model of an allele drove in females when signaled by a complementary allele in sperm.  
+We complement this single locus model %of an allele with pleiotropic effects in both sexes, 
+	with  an alternative model of two linked loci - one a female driver, 
+	and the other a sperm allele which modifies the effect of drive upon fertilization. 
+In this model, a female meiotic driver with no sperm-dependency 
+	initially reaches drive-viability equilibrium (with two alleles
 	$A$ and $B$ are the ancestral non-driver and driver alleles). 
@@
 %%%sperm-acting facilitator
-We first consider a sperm-acting facilitator of
+We first consider a sperm-acting enhancer of
 	female meiotic drive tightly linked to the $B$ background, 
-	generating a third allele $B^{+}$ (i.e. the non-recombining haplotype of a
+	generating a third `$B^{+}$' allele (i.e. the non-recombining haplotype of a
         driver allele and sperm-acting drive modifier). 
 This new allele acts in sperm to increase the effectiveness of both
-	the $B$ and  $B^{+}$ drive in $AB$ and $AB^{+}$ heterozygotes, see Figure \ref{Eggsperm_3_allele_cartoon} 
-	for a simple schematic of this drive model.  
+	the $B$ and  $B^{+}$ drive in $AB$ and $AB^{+}$ heterozygotes (see Figure \ref{Eggsperm_3_allele_cartoon})
+%	for a simple schematic of this drive model.  
 Naively, the $B^{+}$ may spread by capitalizing on the additional drive it causes; however,  
@@
 %%%sperm-acting drive suppressors
-While  sperm enhancement of a female drive cannot displace a polymorphic female driver, sperm based drive suppression - that is sperm-acting alleles that act to restore  fairness of female meiosis can.
-If these sperm-acting drive suppressors arise on
-	the non-driving $A$ background (creating a third allele $A^{-}$) ;
-	or are unlinked to the drive locus they readily invade a population segregating
-	for the drive system ($AB$, see Figure). 
+While  sperm enhancement of a female drive cannot displace a polymorphic female driver, sperm based drive suppression can. 
+%	that is sperm-acting alleles that act to restore  fairness of female meiosis, can. %If these sperm-acting drive suppressors arise on
+If a sperm-acting allele that restores fairness to female meiosis arises on
+	the non-driving $A$ background (creating a third allele $A^{-}$), 
+	or is unlinked to the drive locus, it readily invades a population segregating
+	for the drive system ($AB$, see Figure and Equations XXX). 
 This allele lowers the frequency of the original drive system (perhaps to zero),
@@
 More surprisingly, sperm-acting suppressors of female drive can spread
-	even when tightly linked to the original drive system (B), forming
-	an novel third allele $B^{-}$ as above 
-	(Figures \ref{Eggsperm_3_allele_cartoon} and \ref{Trajectories_of_supressors}B). 
-$B^{-}$ displaces the original driver ($B$) by driving when not costly, 
-	but by disrupting drive when transmitted to sperm it avoids low-fitness drive homozygotes ($BB^-$ and $B^-B^-$). 
-Despite displacing the ancestral driver ($B$), self avoiding driver, $B^*$ it's equilibrium frequency is not always higher than that of $B$ \yb{(point to supp?)}.
+	even when tightly linked to the original drive system (Figures \ref{Eggsperm_3_allele_cartoon} and \ref{Trajectories_of_supressors}B). 
+The $B^{-}$ haplotype (representing the tightly linked drive/drive suppressor complex) displaces the original driver 
+	by driving when not costly, but avoiding low fitness homozygotes by disrupting drive when transmitted to sperm. %  ($BB^-$ and $B^-B^-$). 
+Despite displacing the ancestral driver (the $B$ haplotype), the equilibrium frequency of the $B^-$ haplotype is often lower than that of $B$ \yb{(point to supp?)}.
 

### 2014-05-13 writeup_and_notes/male_dependent_drive.tex
In the process of overhauling the in text appendix. Finished the single-locus, sperm dep model

@@
 
-For comparative purposes, we present the standard drive model in equations \ref{eqn.A1}-\ref{eqn.A3}, 
+For comparative purposes, we present the standard drive model as Model 1 in the Appendix, 
+%equations \ref{eqn.A1}-\ref{eqn.A3}, 
 	and refer readers to the previous literature on this model \citep[e.g. ][]{Prout1973,Ubeda2004} for additional results. 
@@
 	positive frequency dependent behavior (Inset in Figure \ref{Invasion_space}).  
-This lack of a balanced polymorphism (Figure \ref{Invasion_space}) precludes the evolution of an
+This lack of a balanced polymorphism %(Figure \ref{Invasion_space}) 
+	precludes the evolution of an
 	allele that suppresses this form of meiotic drive in females.
@@
 %approximation to this bistability. 
-When $s$ is too large to allow a rare driver to increase in frequency, but small enough to allow a common driver to fix (the inequality in EQUATION XXX),
-there is  an unstable equilibrium frequency (approximated in EQUATION XXX)  
+When $s$ is too large to allow a rare driver to increase in frequency, but small enough to allow a common driver to fix (the inequality in Equation \ref{ineqA}),
+there is  an unstable equilibrium frequency, $f^*_{B\text{ recessive}}$  
 	above which it can spread quickly to fixation and below which it moves towards loss 
-	(see Figure \ref{Bistab_homozyg_cost_fig} for exact results).	
+	($f^*_{B\text{ recessive}}$ is approximated in Equation \ref{bistab_homozy_approx}.  
+	We compare this approximation to the exact result in Figure \ref{Bistab_homozyg_cost_fig}).	
 %OLD GRAHAM WORDS
@@
 % 	but it will be lost otherwise.
-Inclusion of a heterozygous fitness cost (i.e. $w_{AB}=1-h_s$)
+Inclusion of a heterozygous fitness cost (i.e. $w_{AB}=1-s_h$)
 further constrains the evolution of a self-promoting driver. 
@@
 tails). In each case the top and bottom entry in each cell give the
-transmisson probability of the left and right allele in the female
+transmission probability of the left and right allele in the female
 being transmitted to the egg. In the left panel we show the transmission probabilities for
@@
   \appendix
+
 \section*{Appendix}
-In our first scenerio allele $A$ is the ancestral allele and allele
-$B$ is our sperm dependent drive allele.
-The sperm dependence of the drive is a form of assortative
-transmission (mating), and so we have to keep track of the genotype frequencies.
-Our genotype frequencies at birth are $x_{AA},~x_{AB},$ and
-$x_{BB}$. These three genotypes have fitnesses $w_{AA},~w_{AB},$ and
-$w_{BB}$, which are identical between males and females. Because 
-fitnesses are equivalent in both sexes, genotypic frequencies
-are identical in males and females, but in the following
-equations they are marked with $\Mars$ and $\Venus$ respectively to
-allow the reader to more easily follow the logic. 
-The frequency of the three genotypes at birth in the next generation is:
+Below, we highlight some of our derivations and analytical results. 
+The fitness of genotype, $i$ is $w_g$ and its frequency before drive, after drive, and after drive and selection are $f_g$, $f_g'$ and $f_g''$, respectively, and population mean fitness,  $\bar{w}$ equals  $\Sigma f_g'w_g$.
+Genotypic fitnesses are $w_{AA} = 1$, $w_{AB} = 1-s_h$, and $w_{BB}=1-s$, and are sex-independent.
+All models include a balletic locus with non-driving and driving alleles in frequencies $f_A$ and $f_B$, respectively, while more complex models include more loci and/or alleles.  
+Because many of our analyses assume Hardy Weinberg Equilibrium (HWE), we note which analytical results are approximations. 
+We verify these approximations with exact numerical iterations in supplementary figures and scripts, and detail the derivation and details of these results in our mathematical supplement.
+
+\subsection*{Model 1. Traditional driver}
+In the standard drive model meiosis in males is fair such that $A/B$ heterozygotes contribute $A$ and $B$ alleles with equal probabilities; however, $A/B$ females transmit the $B$ allele with probability $d>1/2$.
+After drive and random mating genotype frequencies are 
+	$f_{AA}'=f_A (f_{AA }+ f_{AB} (1 - d))$, 
+	$f_{AB}'=f_A (f_{BB} + f_{ABd}) + f_B (f_{AA} + f_{AB} (1 - d))$,
+	and $f_{BB}= f_B (f_{AB} d + f_{BB})$. 
+%Assuming fitnesses $w_{AA}=1$, $w_{AB}-1 s_h$ and $w_{BB}=1-s$, 
+Frequencies of genotype, $i$, after drive and selection are 
+	\begin{equation} f_i''= f_i'w_i/\bar{w}\end{equation}
+	where $w_i$ is genotypic fitness, and mean fitness, $\bar{w} = \Sigma f_i'w_i$. 
+	Mathematical approximations assuming HWE show that a rare driver will spread when $s_h <(2d-1)/(1+2d)$, and will fix when $s< d -1/2 + 3 s_h /2- d s_h$.
+This later inequality reduces to $s<(2d-1)/2$ when the cost of drive is fully recessive.
+
+\subsection*{Model 2. Single locus, sperm-dependent drive.}
+Our single-locus model of sperm dependent drive resembles the traditional driver with the caveat that the $B$ allele drives in heterozygous females only when fertilized by $B$ bearing sperm. 
+Therefore, genotype frequencies after drive are 
+	$f_{AA}'=f_A^2$, 
+	$f_{AB}'=f_A f_B+f_B (f_{AA} + f_{AB}(1 - d)) $,
+	and $f_{BB}'= f_B (f_{AB} d + f_{BB} )$. 
+As above, exact frequencies after drive, random mating and selection are $f_i''= f_i'w_i/\bar{w}$. 
+We iterate this recursion over generations, to produce the frequency trajectories shown in the inset
+Figure \ref{Invasion_space} by plotting $x_{BB}+ \frac{1}{2}x_{AB}$
+over time, and to assess the analytical approximations, below.
+
+% the $B$ allele In our first scenerio allele $A$ is the ancestral allele and allele
+%$B$ is our sperm dependent drive allele.
+%The sperm dependence of the drive is a form of assortative
+%transmission (mating), and so we have to keep track of the genotype frequencies.
+%Our genotype frequencies at birth are $x_{AA},~x_{AB},$ and
+%$x_{BB}$. These three genotypes have fitnesses $w_{AA},~w_{AB},$ and
+%$w_{BB}$, which are identical between males and females. Because 
+%fitnesses are equivalent in both sexes, genotypic frequencies
+%are identical in males and females, but in the following
+%equations they are marked with $\Mars$ and $\Venus$ respectively to
+%allow the reader to more easily follow the logic. 
+%The frequency of the three genotypes at birth in the next generation is:
+%\begin{equation}
+%x'_{BB} =\frac{1}{\bar{w} } \Big( w_{BB} x_{BB}^{\Venus} (w_{BB} x_{BB}^{\Mars} +\frac{1}{2}
+%  w_{AB} x_{AB}^{\Mars} ) + w_{AB} x_{AB}^{\Venus} (d w_{BB}
+%  x_{BB}^{\Mars} +\frac{1}{2} d
+%w_{AB} x_{AB}^{\Mars} ) \Big),  \label{eqn.A1}
+%\end{equation}
+%\begin{equation}
+%x'_{AA} =\frac{1}{\bar{w} } \Big(  w_{AA} x_{AA}^{\Venus} (w_{AA} x_{AA}^{\Mars} + \frac{1}{2}
+%   w_{AB} x_{AB}^{\Mars}) + w_{AB} x_{AB}^{\Venus}(\frac{1}{2} w_{AA} x_{AA}^{\Mars}+ \frac{1}{2} 
+%  w_{AB} x_{AB}^{\Mars}) \Big),  \label{eqn.A2}
+%\end{equation}
+%and
+%\begin{eqnarray}
+%x'_{AB} = & \frac{1}{\bar{w} } \Big( w_{AA} x_{AA}^{\Venus}  ( w_{BB}  x_{BB}^{\Mars} + \frac{1}{2}
+%  w_{AB}  x_{AB}^{\Mars} ) +w_{BB} x_{BB}^{\Venus} ( w_{AA}  x_{AA}^{\Mars} + \frac{1}{2}
+%  w_{AB}  x_{AB}^{\Mars} )  \nonumber \\
+%& + x_{AB}^{\Venus}(\frac{1}{2}  w_{AA}  x_{AA}^{\Mars}+\frac{1}{2} (1-d)
+% w_{AB}  x_{AB}^{\Mars} + \frac{1}{4}  w_{AB}
+ %x_{AB}^{\Mars} + (1-d)  w_{BB}  x_{BB}^{\Mars} ) \Big) \label{eqn.A3}
+%\end{eqnarray}
+%where the mean fitness of the population is
+%\begin{equation}
+%\bar{w} = w_{AA} x_{AA}+w_{AB}  x_{AB} +w_{BB}  x_{BB}.
+%\end{equation}
+%We can then iterate this recursion over generations, for
+%example to produce the frequency trajectories shown in the inset
+%Figure \ref{Invasion_space} by plotting $x_{BB}+ \frac{1}{2}x_{AB}$
+%over time. We can also use this to assess the
+%invasion criteria for the allele by starting allele B at very low
+%frequency, and seeing if it increases across the generations.
+
+\paragraph{Recessive fitness cost of self-promoting driver:} %When the self-promoting driver has a recessive fitness cost, $s$ genotype frequencies after selection and drive equal
+%\begin{equation}	x'_{BB} =\frac{x_{B}^2(1-s)}{\bar{w} }  \Big( F+(1-F)(x_{B}+2x_{A} d) \Big) 	\end{equation}
+%\begin{equation}	x'_{AA} =\frac{1}{\bar{w} }  \Big(x_{A}^2\Big) 	\end{equation}
+%and
+%\begin{equation}	x'_{AB} =\frac{2 x_{A}x_{B} }{\bar{w} }  \Big(1-x_{B}(d-1/2)(1-F) \Big)	%\end{equation}
+When fully recessive, the change in frequency of the self-promoting driver across generations $\Delta f_B$ is exactly 
 \begin{equation}
-x'_{BB} =\frac{1}{\bar{w} } \Big( w_{BB} x_{BB}^{\Venus} (w_{BB} x_{BB}^{\Mars} +\frac{1}{2}
-  w_{AB} x_{AB}^{\Mars} ) + w_{AB} x_{AB}^{\Venus} (d w_{BB}
-  x_{BB}^{\Mars} +\frac{1}{2} d
-w_{AB} x_{AB}^{\Mars} ) \Big),  \label{eqn.A1}
-\end{equation}
-\begin{equation}
-x'_{AA} =\frac{1}{\bar{w} } \Big(  w_{AA} x_{AA}^{\Venus} (w_{AA} x_{AA}^{\Mars} + \frac{1}{2}
-   w_{AB} x_{AB}^{\Mars}) + w_{AB} x_{AB}^{\Venus}(\frac{1}{2} w_{AA} x_{AA}^{\Mars}+ \frac{1}{2} 
-  w_{AB} x_{AB}^{\Mars}) \Big),  \label{eqn.A2}
-\end{equation}
-and
-\begin{eqnarray}
-x'_{AB} = & \frac{1}{\bar{w} } \Big( w_{AA} x_{AA}^{\Venus}  ( w_{BB}  x_{BB}^{\Mars} + \frac{1}{2}
-  w_{AB}  x_{AB}^{\Mars} ) +w_{BB} x_{BB}^{\Venus} ( w_{AA}  x_{AA}^{\Mars} + \frac{1}{2}
-  w_{AB}  x_{AB}^{\Mars} )  \nonumber \\
-& + x_{AB}^{\Venus}(\frac{1}{2}  w_{AA}  x_{AA}^{\Mars}+\frac{1}{2} (1-d)
- w_{AB}  x_{AB}^{\Mars} + \frac{1}{4}  w_{AB}
- x_{AB}^{\Mars} + (1-d)  w_{BB}  x_{BB}^{\Mars} ) \Big) \label{eqn.A3}
-\end{eqnarray}
-where the mean fitness of the population is
-\begin{equation}
-\bar{w} = w_{AA} x_{AA}+w_{AB}  x_{AB} +w_{BB}  x_{BB}.
-\end{equation}
-We can then iterate this recursion over generations, for
-example to produce the frequency trajectories shown in the inset
-Figure \ref{Invasion_space} by plotting $x_{BB}+ \frac{1}{2}x_{AB}$
-over time. We can also use this to assess the
-invasion criteria for the allele by starting allele B at very low
-frequency, and seeing if it increases across the generations.
-
-\paragraph{Recessive fitness cost of self-promoting driver}
-When the self-promoting driver has a recessive fitness cost, $s$ genotype frequencies after selection and drive equal
-\begin{equation}	x'_{BB} =\frac{x_{B}^2(1-s)}{\bar{w} }  \Big( F+(1-F)(x_{B}+2x_{A} d) \Big) 	\end{equation}
-\begin{equation}	x'_{AA} =\frac{1}{\bar{w} }  \Big(x_{A}^2\Big) 	\end{equation}
-and
-\begin{equation}	x'_{AB} =\frac{2 x_{A}x_{B} }{\bar{w} }  \Big(1-x_{B}(d-1/2)(1-F) \Big)	\end{equation}
-
-its change in frequency across generations $\Delta f_B$ is
-\begin{equation}
-\frac{f_A f_B^2}{\bar{w}} \left( [1-F][ d(1-2 f_A s) - f_B s-1/2] - s  F \right)
+f_A f_B^2 \left( [1-F][ d(1-2 f_A s) - f_B s-1/2] - s  F \right)/\bar{w}
 \label{deltadriver}
 \end{equation}
-where $F$ is the deviation from Hardy-Weinberg expectations (HWE), and $f_A$ is the frequency of the non-driving allele. 
-For analytical solutions presented above, we assume that $F$ equals zero.
-
-We used this HWE approximation to investifgate the bistability of the
-invasion criteria. Under this approximation we find that a common, recessive, self-promoting driver fixes if
- $s<(2d-1)/2$. Furthermore, when $(2 d - 1)/(4 d)<s<(2d-1)/(2)$, a
+where $F$ is the deviation from Hardy-Weinberg expectations (HWE). 
+%For analytical solutions presented above, we assume that $F$ equals zero. 
+Assuming HWE a common, recessive, self-promoting driver invades if $(2 d - 1)/(4 d)$, and fixes if
+ $s<(2d-1)/2$. 
+ Therefore, when \begin{equation}(2 d - 1)/(4 d)<s<(2d-1)/2\label{ineqA}\end{equation} a
  	recessive, self-promoting driver will deterministically fix if drift,
@@
 \begin{equation} 
-(1-2d+4ds)/(2s(2d-1)) \label{bistab_homozy_approx}
+f^*_{B\text{ recessive}} = (1-2d+4ds)/(2s(2d-1)) \label{bistab_homozy_approx}
 \end{equation}
- but it will be lost if it is introduced below this frequency
- (i.e. this is an unstable equilibrium point). In Supplementary Figure \ref{Bistab_homozyg_cost_fig} we show
- this approximation condition for bistability andd how it compared to
- that obtained using the recursion. This equation
- \eqref{bistab_homozy_approx} offers a reasonable, but imperfect
- approximation 
-and seems to serve as a lower bound to the
-true unstable equilibrium frequency.
+ but it will be lost if it is introduced below this frequency. 
+ %(i.e. this is an unstable equilibrium point). 
+% In Supplementary Figure \ref{Bistab_homozyg_cost_fig} we show
+% this approximation condition for bistability andd how it compared to
+% that obtained using the recursion. 
+Compared to exact results (Supplementary Figure \ref{Bistab_homozyg_cost_fig}), Equation
+ \eqref{bistab_homozy_approx} offers a reasonable, but imperfect approximation. 
+%s_h and seems to serve as a lower bound to the true unstable equilibrium frequency.
 
-\paragraph{cost of driver in heterozygotes}
-We explored a more complex model where the self-promoting driver has a cost, $s_h$ in heterozygotes (so $w_{AB} = 1-h_s$) in addition to the cost $s$, in homozygotes.
-Model construction is similar to that above.
-So long as there is any heterozygous fitness cost (i.e. $s_h > 0$) the self-promoting driver cannot invade when rare;
-	however, as is the case without a heterozygous cost, this system is bistable.
-Assuming no deviation from Hardy-Weinberg a self-promoting driver with a heterozygous fitness cost deterministically fixes if it's frequency exceeds some threshold, $f_B^*$.
+\paragraph{Cost of driver in heterozygotes:} %We explored a more complex model where the self-promoting driver has a cost, $s_h$ in heterozygotes (so $w_{AB} = 1-s_h$) in addition to the cost $s$, in homozygotes.
+When the fitness of drive heterozygotes is compromised ($s_h>0$), a rare, self promoting driver cannot invade when rare.
+This results from the fact that
+when the allele is rare ($f_B \ll 1$) sperm carrying it rarely
+encounter heterozygous eggs ($\sim f_B^2$) but the allele still pays a
+cost in heterozygous individuals. 
+%Model construction is similar to that above.
+%So long as there is any heterozygous fitness cost (i.e. $s_h > 0$) the self-promoting driver cannot invade when rare;
+However, this system too, is bistable -- as the driver becomes more frequency it is more often fertilized by a driving sperm and therefore drives more effectively.
+%Assuming no deviation from Hardy-Weinberg a self-promoting driver with a heterozygous fitness cost 
+This allele deterministically fixes if its frequency exceeds some threshold, $f_B^*$.
+Assuming HWE, this threshold equals 
 \begin{equation}
@@
 where, $a_1=1-2 ds_h+4 d s-2 d-3 s_h$ and $a_2=2(s-s_h)(2d-1)$.
-
- range of other costs to the drive alleles other than a
-simple recessive cost, for example an additive cost. We modified the recursions given in eqn
-\ref{eqn.A1}-\ref{eqn.A3} to feature an additive cost $s$ for each
-copy of the $B$ allele. A none recessive cost to the drive allele can
-cause bistability of the drive allele. The B allele is unable to enter
-the population when rare, this results from the fact that
-when the allele is rare ($x_B \ll 1$) sperm carrying it rarely
-encounter heterozygous eggs ($\sim x^2$) but the allele still pays a
-cost in heterozygous individuals.  By contrast, when the B allele is common ($x_B
-\gg 0$) encounters between heterozygous eggs and $B$ allele sperm
-become more common allowing drive to occur this can allow the $B$
-allele to spread when common. In Supplementary Figure
-\ref{bistable_additive} we show the unstable equilibrium frequency
-above which the $B$ allele can invade. This equilibrium frequency was found by a simple
-parameter search.  \gc{wonder if a bit of work could find this}
+\yb{We compare this analytical approximation to exact results obtained by a simple parameter search in Supplementary Figure \ref{bistable_additive}. }
+ %range of other costs to the drive alleles other than a
+%simple recessive cost, for example an additive cost. We modified the recursions given in eqn
+%\ref{eqn.A1}-\ref{eqn.A3} to feature an additive cost $s$ for each
+%copy of the $B$ allele. A none recessive cost to the drive allele can
+%cause bistability of the drive allele. The B allele is unable to enter
+%the population when rare, this results from the fact that
+%when the allele is rare ($x_B \ll 1$) sperm carrying it rarely
+%encounter heterozygous eggs ($\sim x^2$) but the allele still pays a
+%cost in heterozygous individuals.  By contrast, when the B allele is common ($x_B
+%\gg 0$) encounters between heterozygous eggs and $B$ allele sperm
+%become more common allowing drive to occur this can allow the $B$
+%allele to spread when common. In Supplementary Figure
+%\ref{bistable_additive} we show the unstable equilibrium frequency
+%above which the $B$ allele can invade. This equilibrium frequency was found by a simple
+%parameter search.  \gc{wonder if a bit of work could find this}
 

### 2014-05-13 writeup_and_notes/male_dependent_drive.tex
Updated / overhauled the in-text appendix. Added writeup_and_notes/TransmissionRules.xlsx

@@
 All models include a balletic locus with non-driving and driving alleles in frequencies $f_A$ and $f_B$, respectively, while more complex models include more loci and/or alleles.  
+Transmission rules describing the outcomes of all matings in each models are presented in a \yb{Supplementary Worksheet}.  
 Because many of our analyses assume Hardy Weinberg Equilibrium (HWE), we note which analytical results are approximations. 
@@
 
-\subsection*{Model 1. Traditional driver}
+\subsection*{Models 1-3. Single-locus drive}
+\subsubsection*{Model 1. Traditional driver}
 In the standard drive model meiosis in males is fair such that $A/B$ heterozygotes contribute $A$ and $B$ alleles with equal probabilities; however, $A/B$ females transmit the $B$ allele with probability $d>1/2$.
@@
 	where $w_i$ is genotypic fitness, and mean fitness, $\bar{w} = \Sigma f_i'w_i$. 
-	Mathematical approximations assuming HWE show that a rare driver will spread when $s_h <(2d-1)/(1+2d)$, and will fix when $s< d -1/2 + 3 s_h /2- d s_h$.
-This later inequality reduces to $s<(2d-1)/2$ when the cost of drive is fully recessive.
+	Assuming HWE, a rare driver will spread when $s_h \lesssim(2d-1)/(1+2d)$, and will fix when $s\lesssim d -1/2 + 3 s_h /2- d s_h$.
+This later inequality reduces to $s\lesssim(2d-1)/2$ when the cost of drive is fully recessive.
 
-\subsection*{Model 2. Single locus, sperm-dependent drive.}
+\subsubsection*{Model 2. Single locus, sperm-dependent drive.}
 Our single-locus model of sperm dependent drive resembles the traditional driver with the caveat that the $B$ allele drives in heterozygous females only when fertilized by $B$ bearing sperm. 
@@
 %\begin{equation}	x'_{AB} =\frac{2 x_{A}x_{B} }{\bar{w} }  \Big(1-x_{B}(d-1/2)(1-F) \Big)	%\end{equation}
-When fully recessive, the change in frequency of the self-promoting driver across generations $\Delta f_B$ is exactly 
+When fully recessive, the change in frequency of the self-promoting driver across generations equals 
 \begin{equation}
-f_A f_B^2 \left( [1-F][ d(1-2 f_A s) - f_B s-1/2] - s  F \right)/\bar{w}
+\Delta f_B=f_A f_B^2 \left( [1-F][ d(1-2 f_A s) - f_B s-1/2] - s  F \right)/\bar{w}
 \label{deltadriver}
 \end{equation}
-where $F$ is the deviation from Hardy-Weinberg expectations (HWE). 
+where $F$ is the deviation from HWE. 
 %For analytical solutions presented above, we assume that $F$ equals zero. 
-Assuming HWE a common, recessive, self-promoting driver invades if $(2 d - 1)/(4 d)$, and fixes if
- $s<(2d-1)/2$. 
- Therefore, when \begin{equation}(2 d - 1)/(4 d)<s<(2d-1)/2\label{ineqA}\end{equation} a
+Assuming HWE a common, recessive, self-promoting driver invades if $s\lesssim(2 d - 1)/(4 d)$, and fixes if
+ $s\lesssim(2d-1)/2$. 
+ Therefore, when \begin{equation}(2 d - 1)/(4 d)  \lesssim  s  \lesssim  (2d-1)/2\label{ineqA}\end{equation} a
  	recessive, self-promoting driver will deterministically fix if drift,
- 	mutation, or migration pressure bring its frequency above approximately
+ 	mutation, or migration pressure bring its frequency above
 \begin{equation} 
-f^*_{B\text{ recessive}} = (1-2d+4ds)/(2s(2d-1)) \label{bistab_homozy_approx}
+f^*_{B\text{ recessive}} \approx (1-2d+4ds)/(2s(2d-1)) \label{bistab_homozy_approx}
 \end{equation}
@@
 \paragraph{Cost of driver in heterozygotes:} %We explored a more complex model where the self-promoting driver has a cost, $s_h$ in heterozygotes (so $w_{AB} = 1-s_h$) in addition to the cost $s$, in homozygotes.
-When the fitness of drive heterozygotes is compromised ($s_h>0$), a rare, self promoting driver cannot invade when rare.
+When the fitness of drive heterozygotes is compromised ($s_h>0$), a self promoting driver cannot invade when rare.
 This results from the fact that
-when the allele is rare ($f_B \ll 1$) sperm carrying it rarely
+when rare %($f_B \ll 1$) 
+$B$-bearing sperm rarely
 encounter heterozygous eggs ($\sim f_B^2$) but the allele still pays a
-cost in heterozygous individuals. 
+cost in heterozygous individuals ($\sim f_B$). 
 %Model construction is similar to that above.
@@
 \begin{equation}
-	f_B^* = \frac{a_1+\sqrt{a_1^2+8 s_h a_2}}{2 a_2}
+	f_B^* \approx \frac{a_1+\sqrt{a_1^2+8 s_h a_2}}{2 a_2}
 	\label{thresh2}
@@
 where, $a_1=1-2 ds_h+4 d s-2 d-3 s_h$ and $a_2=2(s-s_h)(2d-1)$.
-\yb{We compare this analytical approximation to exact results obtained by a simple parameter search in Supplementary Figure \ref{bistable_additive}. }
+\yb{Comparison of Equation \ref{thresh2} to exact results obtained by a simple parameter search (Supplementary Figure \ref{bistable_additive}) show that this approximation is qualitatively reasonable, but quantitatively imperfect.
+}
  %range of other costs to the drive alleles other than a
@@
 
-\paragraph{Dominance} we also explored a model where
-the genotype of the male who produced the fertilizing sperm affects female drive,
-instead of just the allelic type of the sperm.
+\subsubsection*{Model 3. Single locus, paternal genotype dependent drive:}
+When female meiotic drive depends on paternal genotype
+%\paragraph{Dominance} we also explored a model where
+%the genotype of the male who produced the fertilizing sperm affects female drive,
+%instead of just the allelic type of the sperm.
  We assume a heterozygous female will transmit the $B$ allele 
-  with probabilities  $\frac{1}{2}$,  $\frac{1}{2} + h(d-\frac{1}{2}) $ or $d$, 
- if she mated with a n $AA$, $AB$, or $BB$ male,  respectively.
- Above, $d > 1/2$ and $h$ ($0 \geq h \geq 1$) acts to specify the dominance of the $B$
+  with probabilities  $\frac{1}{2}$,  $d_{het}\geq 1/2 $ or $d_{hom}\geq d_{het}$, 
+%YB RE-PARAMETERIZING  %  $\frac{1}{2}$,  $\frac{1}{2} + h(d-\frac{1}{2}) $ or $d$,  YB RE-PARAMETERIZING 
+ when mated with to $AA$, $AB$, or $BB$ males,  respectively. 
+  In this model, genotype frequencies after drive, mating and selection are 
+  	$f_{AA}'' =   w_{AA}\left( f_{AA} [2 f_A + f_{AB} ] + f_{AB}^2 [1 - d_{het}] \right)/(2\bar{w})$,
+	$f_{AB}'' =   w_{AB}(f_{AA} [f_B + f_{AB}/2] + f_{AB} f_B + f_{BB} [ f_A - f_{AB} d_{hom}])/\bar{w}$, 
+	and 	$f_{AB}'' =   w_{AB}(f_{AB} (f_{AB} d_{het}/2 + f_{BB} d_{hom}) + f_{BB} f_B)/\bar{w}$. 
+% Above, $d > 1/2$ and $h$ ($0 \geq h \geq 1$) acts to specify the dominance of the $B$
 %allele's effect on female meiotic drive through the males genotype.
@@
 %allele's effect on female meiotic drive through the males genotype. 
-We set up a set of recursions using a recessive sex independent cost to
-the driver, similar to those in eqn.
-\ref{eqn.A1}-\ref{eqn.A1}, but now featuring these male genotype
-transmission probabilites. In Supplementary Figure \ref{Effect_of_dominance} we show the invasion
-criteria for both the A and B alleles when they are rare. 
-We find a narrow parameter range where B 
-can invade but not fix.
-This corresponds to B being less than fully recessive (i.e. $h>0$) and to a modest cost in homozygotes. 
-This combination of parameters reflects the fact that when the allele is rare, 
-it can increase in frequency due to fertilization of heterozygous eggs
-by sperm from heterozygote males, where the sperm do not themselves
-carry the allele. But as the B allele becomes common it will frequently be
-present in the homozygous state in males, and so will be unable to
-avoid the production of unfit homozygote zygotes leaving it trapped in
-the population.
+%We set up a set of recursions using a recessive sex independent cost to
+%the driver, similar to those in eqn.
+%\ref{eqn.A1}-\ref{eqn.A1}, but now featuring these male genotype
+%transmission probabilities.
 
-\paragraph{three allele system}
+If the cost of drive is fully recessive (i.e. $s_h=0$), assuming HWE, 
+	a rare paternal genotype dependent driver invades when 
+	$s\lesssim (d_{het}-1/2)/(d_{het})$, and a common paternal genotype dependent driver fixes when 
+	$s\lesssim d_{hom}-1/2$, approximations well supported by exact results (Supplementary Figure \ref{Effect_of_dominance}).
+Specifically, when drive in heterozygotes is large relative to that in homozygotes, 
+	($d_{het} \gtrsim 1/(3-2d_{hom})$), fixation criteria are more stringent than invasion criteria, 
+	and therefore some values of $s$ can maintain a stable polymorphism. 
+Under these parameter values, a rare paternal genotype dependent driver 
+	can increase in frequency as drivers gain a transmission advantage and suffer 
+	no fitness cost when heterozygous eggs are fertilized by 
+	$A$-bearing sperm of heterozygous males. 
+As the frequency of the $B$ allele increases, 
+	occur more often as a homozygote, and so will be unable to 
+	avoid producing unfit homozygote zygotes, leaving it trapped in
+	the population at frequency $f_{B\text{ pat}}^*$. 
+Assuming HWE, a recessive fitness cost ($h_s=0$), and dominance of driver ($d_{het}=d_{hom}$) this equilibrium frequency is
+\begin{equation}f_{B\text{ pat}}^*\approx2 (1 + 2d_{hom}[s - 1])/([2 s - 1][2d_{hom}- 1]) \label{eqmale} \end{equation}
+By contrast, when $d_{het} \lesssim 1/(3-2d_{hom})$ the case is reversed, and the model is bistable.
+%In Supplementary Figure \ref{Effect_of_dominance} we show the invasion
+%criteria for both the A and B alleles when they are rare. 
+%We find a narrow parameter range where B 
+%can invade but not fix.{eqmale}
+%This corresponds to B being less than fully recessive (i.e. $h>0$) and to a modest cost in homozygotes. 
+%This combination of parameters reflects the fact that when the allele is rare, 
+%it can increase in frequency due to fertilization of heterozygous eggs
+%by sperm from heterozygote males, where the sperm do not themselves
+%carry the allele. But as the B allele becomes common it will frequently be
+%present in the homozygous state in males, and so will be unable to
+%avoid the production of unfit homozygote zygotes leaving it trapped in
+%the population.
+
+\subsection*{Models 4-6. Two-locus, sperm-dependent drive.} 
+Assuming that a traditional driver segregates at its equilibrium frequency (Model 1), we investigate the evolution of tightly linked (Models 4 and 5), and unlinked sperm-dependent modifiers of drive. 
+In models 4 and 5, we treat this two loci system as if it consists of three alleles: $A$, $B$ and $C$, which is appropriate if these lowly are tightly linked, such that recombination is unexpected. 
+In all cases, fitness is independent of the drive modifier, so, $w_{AA}=1$, $w_{AB}=1-s_h$, and $w_{BB}$ = 1-s. 
+We assume the traditional driver is transmitted to $d_0$ of gametes from female heterozygotes  when fertilized by wild-type sperm, and $d_1=d_0+\epsilon$ when fertilized by a sperm acting drive modifier.
+
+\subsubsection*{Model 4. Drive-modifier in coupling phase:}
+When the $C$ allele is tightly linked to and on the same genetic background as the driver, 
+	genotypic fitnesses equal $w_{AC}=w_{AB}=1-s_h$, and $w_{BC}=w_{CC}=w_{BB}=1-s$. 
+%We address two questions concerning the $C$ allele - can it invade when rare  and can it fix when common?
+Although analytical approximation do not reveal simple inequalities like those above,
+	graphical exploration of analytical results show that this mutant is lost when it enhances drive ($\epsilon>0$ - the $B^+$ allele, above),
+	and spreads from rarity to fixation when it suppresses drive ($\epsilon<0$ - the $B^-$ allele, above).	
+This result is supported by exact numerical iterations, as well \yb{FIG?}.
+
+\subsubsection*{Model 5. Drive-modifier in repulsion phase:}
+When the $C$ allele is tightly linked to and on a different background than the driver, 
+	genotypic fitnesses equal $w_{AC}=w_{CC}=w_{AA}=1$, and $w_{BC}=1-s_h$. 
+%We address two questions concerning the $C$ allele - can it invade when rare  and can it fix when common?
+Again, while simple inequalities are lacking, 
+	graphical exploration of our analytical results, and exact numerical iteration \yb{FIG?} shows that this modifier only spreads from rarity to fixation when it suppresses drive ($\epsilon<0$ - the $A^-$ allele, above).
+Somewhat counterintuitively, the spread of the $A^-$ allele is substantially slower than that of the $B^-$ allele (it is on the order of $C^2$ rather than $C$). 
+This is because the $A^-$ allele will never occur in drive-homozygotes and therefore only spreads by preventing itself from the cheating of the driver.
+
+\subsubsection*{Model 6. Unlinked drive-modifier:}
+For the unlinked, model we introduce another locus where drive is modified in A/B females fertilized by $F$ allele, 
+	while the wild-type $E$ does not influence drive. 
+Assuming HWE and linkage equilibrium, an $F$ allele increases in frequency when 
+	$-\epsilon f_A f_B (f_B s + s_h (fA - fB) ) >0$, corresponding to any level of drive suppression ($\epsilon<0$) 
+	when the the fitness cost of the driver does not display over- or under-dominance. 
+Notably, this result holds while assuming complete linkage equilibrium, and therefore sperm suppression of female meiotic drive is likely more powerful than female enforced systems which usually spread due to the indirect benefit of avoiding drive haplotypes \citep[e.g.
+][]{Brandvain2012}.
+The spread of this driver does not, however, depend on assumption of HWE, and linkage equilibrium - exact results resemble analytical approximations. 
+%\paragraph{three allele system}
 %$x_{AA}$, $x_{AB}$,$x_{AB'}$, $w_{BB'} x_{BB'}$, $w_{B'B'} x_{BB}$, $w_{B'B'} x_{B'B'}$
-We denote the frequency of the three genotypes in the population at birth are
-$x_{AA}$, $x_{AB}$, $x_{AB'}$, $x_{BB'}$, $x_{BB}$, $x_{B'B'}$. 
-The mean relative fitness of the population is 
-\begin{equation}
-x_{AA}+x_{AB}+ x_{AB'}+ w_{BB'} +x_{BB'}+w_{B'B'}x_{BB} + w_{B'B'} x_{B'B'} 
-\end{equation}
-The frequency of the genotypes in meiotic-cells of female adults is 
-\begin{equation}
-\frac{1}{\bar{w}}\Big( x_{AA}, ~x_{AB},~ x_{AB'},~ w_{BB'} x_{BB'},~ w_{B'B'}x_{BB},
-~w_{B'B'} x_{B'B'} \Big)
-\end{equation}
-The frequency of fertilization events can be found by combining these
-female frequencies with the frequency of the three alleles in sperm is
-\begin{eqnarray}
-x_A = & x_{AA} + \frac{1}{2} x_{AB}+ \frac{1}{2} x_{AB'} \\
-x_B = & w_{BB}  x_{BB} +   w_{BB'} \frac{1}{2} x_{BB'}+ \frac{1}{2} x_{AB} \\
-x_{B'} = & w_{B'B'}  x_{B'B'} +   w_{BB'} \frac{1}{2} + x_{BB'}+ \frac{1}{2} x_{AB'}  
-\end{eqnarray}
-in combination with the transmission probabilities given in Figure
-\ref{Eggsperm_3_allele_cartoon}A.
+%We denote the frequency of the three genotypes in the population at birth are
+%$x_{AA}$, $x_{AB}$, $x_{AB'}$, $x_{BB'}$, $x_{BB}$, $x_{B'B'}$. 
+%The mean relative fitness of the population is 
+%\begin{equation}
+%x_{AA}+x_{AB}+ x_{AB'}+ w_{BB'} +x_{BB'}+w_{B'B'}x_{BB} + w_{B'B'} x_{B'B'} 
+%\end{equation}
+%The frequency of the genotypes in meiotic-cells of female adults is 
+%\begin{equation}
+%\frac{1}{\bar{w}}\Big( x_{AA}, ~x_{AB},~ x_{AB'},~ w_{BB'} x_{BB'},~ w_{B'B'}x_{BB},
+%~w_{B'B'} x_{B'B'} \Big)
+%\end{equation}
+%The frequency of fertilization events can be found by combining these
+%female frequencies with the frequency of the three alleles in sperm is
+%\begin{eqnarray}
+%x_A = & x_{AA} + \frac{1}{2} x_{AB}+ \frac{1}{2} x_{AB'} \\
+%x_B = & w_{BB}  x_{BB} +   w_{BB'} \frac{1}{2} x_{BB'}+ \frac{1}{2} x_{AB} \\
+%x_{B'} = & w_{B'B'}  x_{B'B'} +   w_{BB'} \frac{1}{2} + x_{BB'}+ \frac{1}{2} x_{AB'}  
+%\end{eqnarray}
+%in combination with the transmission probabilities given in Figure
+%\ref{Eggsperm_3_allele_cartoon}A.
 

### 2014-05-14 writeup_and_notes/male_dependent_drive.tex
pulled out Equaltion 6 [theshold value of drive mod in het males] and pointed to it in text

@@
 	when the cost of the drive is recessive
-	(see Figure S\ref{Effect_of_dominance}, \yb{and Equations XXX in the Appendix}).  
+	(see Figure S\ref{Effect_of_dominance}, \yb{and Equation \ref{polymale}}).  
 %Presumably this zone of polymorphism comes about because positive frequency 
@@
 Specifically, when drive in heterozygotes is large relative to that in homozygotes, 
-	($d_{het} \gtrsim 1/(3-2d_{hom})$), fixation criteria are more stringent than invasion criteria, 
+	\begin{equation} d_{het} \gtrsim 1/(3-2d_{hom}) \label{polymale} \end{equation}
+	fixation criteria are more stringent than invasion criteria, 
 	and therefore some values of $s$ can maintain a stable polymorphism. 

### 2014-05-14 writeup_and_notes/male_dependent_drive.tex
added invasion crit for A- allele

@@
 	or is unlinked to the drive locus, it readily invades a population segregating
-	for the drive system ($AB$, see Figure and Equations XXX). 
+	for the drive system ($AB$, see Figure and \yb{Equations \ref{A-}).} %still coming together 
 This allele lowers the frequency of the original drive system (perhaps to zero),
@@
 When the $C$ allele is tightly linked to and on a different background than the driver, 
-	genotypic fitnesses equal $w_{AC}=w_{CC}=w_{AA}=1$, and $w_{BC}=1-s_h$. 
+	genotypic fitnesses equal $w_{CC}=w_{AC}=w_{AA}=1$, and $w_{BC}=1-s_h$. 
+Assuming HWE, a recessive fitness cost to drive, and assuming that the $A/B$ locus is at its equilibrium frequency, the change in frequency of a rare drive modifier is
+\begin{equation}
+	\Delta C^- \approx -\epsilon f_{C^-}^2 \frac{ \left(2 d_0 s-\sqrt{2} \sqrt{s (2 d_0 (d_0 (s-2)+2)-1)}\right)}{2\bar{w}s (2 d_0 -1) }. \label{A-}
+\end{equation}
+For all values of interest ($0<s<1$, $0.5<d_0<1$), the change in frequency a rare $C$ is positive when it decreases drive (i.e.
+$\epsilon <0$, corresponding to the $A^-$ model, above), a result which hold qualitatively for a common $C$ allele, as well (Mathematical Supplement).
 %We address two questions concerning the $C$ allele - can it invade when rare  and can it fix when common?
-Again, while simple inequalities are lacking, 
-	graphical exploration of our analytical results, and exact numerical iteration \yb{FIG?} shows that this modifier only spreads from rarity to fixation when it suppresses drive ($\epsilon<0$ - the $A^-$ allele, above).
-Somewhat counterintuitively, the spread of the $A^-$ allele is substantially slower than that of the $B^-$ allele (it is on the order of $C^2$ rather than $C$). 
+%Again, while simple inequalities are lacking, 
+%	graphical exploration of our analytical results, and exact numerical iteration \yb{FIG?} shows that this modifier only spreads from rarity to fixation when it suppresses drive ($\epsilon<0$ - the $A^-$ allele, above).
+Somewhat counterintuitively, the spread of the $A^-$ allele is substantially slower than that of the $B^-$ allele (it is on the order of $C^{2}$ rather than $C$). 
 This is because the $A^-$ allele will never occur in drive-homozygotes and therefore only spreads by preventing itself from the cheating of the driver.

### 2014-05-14 writeup_and_notes/male_dependent_drive.tex
made invasion of unlinked mod an equation and pointed to it in main text

@@
 	or is unlinked to the drive locus, it readily invades a population segregating
-	for the drive system ($AB$, see Figure and \yb{Equations \ref{A-}).} %still coming together 
+	for the drive system ($AB$, see Figure and Equations \ref{A-} and \ref{unlinked}).  
 This allele lowers the frequency of the original drive system (perhaps to zero),
@@
 	while the wild-type $E$ does not influence drive. 
-Assuming HWE and linkage equilibrium, an $F$ allele increases in frequency when 
-	$-\epsilon f_A f_B (f_B s + s_h (fA - fB) ) >0$, corresponding to any level of drive suppression ($\epsilon<0$) 
-	when the the fitness cost of the driver does not display over- or under-dominance. 
+Assuming HWE and linkage equilibrium, the change in frequency of a rare unlinked, sperm-acting drive modifier is 
+\begin{equation} \Delta F =-\epsilon f_A f_B (f_B s + s_h (fA - fB) ) \frac{f_F}{\bar{w}}  \label{unlinked} \end{equation}
+% an $F$ allele increases in frequency when 
+%	$-\epsilon f_A f_B (f_B s + s_h (fA - fB) ) >0$, corresponding to any level of drive suppression ($\epsilon<0$) 
+Thus, a rare drive suppressor ($\epsilon<0$) will spread so long as  the fitness cost of the driver does not display over- or under-dominance. 
 Notably, this result holds while assuming complete linkage equilibrium, and therefore sperm suppression of female meiotic drive is likely more powerful than female enforced systems which usually spread due to the indirect benefit of avoiding drive haplotypes \citep[e.g.

### 2014-05-14 writeup_and_notes/male_dependent_drive.tex
edited through appendix

--- 2014-05-14 writeup_and_notes/male_dependent_drive.tex
@@
 Below, we highlight some of our derivations and analytical results. 
-The fitness of genotype, $i$ is $w_g$ and its frequency before drive, after drive, and after drive and selection are $f_g$, $f_g'$ and $f_g''$, respectively, and population mean fitness,  $\bar{w}$ equals  $\Sigma f_g'w_g$.
-Genotypic fitnesses are $w_{AA} = 1$, $w_{AB} = 1-s_h$, and $w_{BB}=1-s$, and are sex-independent.
-All models include a balletic locus with non-driving and driving alleles in frequencies $f_A$ and $f_B$, respectively, while more complex models include more loci and/or alleles.  
-Transmission rules describing the outcomes of all matings in each models are presented in a \yb{Supplementary Worksheet}.  
+The fitness of genotype, $i$ is $w_g$ independent of its sex. 
+We denote its frequency in adults in the current generation is $f_g$. In the next generation of
+zygotes (accounting for drive) but before selection
+the frequency is $f_g'$, and in the adults after selection the
+frequency is $f_g''$ and population mean fitness,  $\bar{w}=\Sigma f_g'w_g$. Genotypic fitnesses are $w_{AA} = 1$, $w_{AB} = 1-s_h$, and $w_{BB}=1-s$, and are sex-independent. 
+All models include a biallelic locus with non-driving and driving alleles in frequencies $f_A$ and $f_B$, respectively, while more complex models include more loci and/or alleles.  
+Transmission rules describing the outcomes of all matings in each
+models are presented in a \yb{Supplementary Worksheet}, and  our
+\gc{Mathematical supplement} for our analytical approximations work.  
 Because many of our analyses assume Hardy Weinberg Equilibrium (HWE), we note which analytical results are approximations. 
@@
 \subsubsection*{Model 1. Traditional driver}
-In the standard drive model meiosis in males is fair such that $A/B$ heterozygotes contribute $A$ and $B$ alleles with equal probabilities; however, $A/B$ females transmit the $B$ allele with probability $d>1/2$.
+In the standard female drive model meiosis in males is fair such that $A/B$ heterozygotes contribute $A$ and $B$ alleles with equal probabilities; however, $A/B$ females transmit the $B$ allele with probability $d>1/2$.
 After drive and random mating genotype frequencies are 
@@
 	where $w_i$ is genotypic fitness, and mean fitness, $\bar{w} = \Sigma f_i'w_i$. 
-	Assuming HWE, a rare driver will spread when $s_h \lesssim(2d-1)/(1+2d)$, and will fix when $s\lesssim d -1/2 + 3 s_h /2- d s_h$.
-This later inequality reduces to $s\lesssim(2d-1)/2$ when the cost of drive is fully recessive.
+	Assuming HWE, a rare driver will spread when $s_h
+        \lesssim(2d-1)/(1+2d)$, and will fix when $s\lesssim d -1/2 +
+        3 s_h /2- d s_h$. This later inequality reduces to
+        $s\lesssim(2d-1)/2$ when the cost of drive is fully
+        recessive. \gc{isn't this last expression exact?}
 
@@
 We iterate this recursion over generations, to produce the frequency trajectories shown in the inset
-Figure \ref{Invasion_space} by plotting $x_{BB}+ \frac{1}{2}x_{AB}$
-over time, and to assess the analytical approximations, below.
+Figure \ref{Invasion_space} by plotting \gc{$f_B=f_{BB}+ \frac{1}{2}g_{AB}$}
+over time, and to assess the analytical approximations, below. \gc{To
+assess invasion or fixation criteria we iterate this system and
+testing whether $f_B$ increases over a grid of parameters. }
 
@@
 \end{equation}
-where $F$ is the deviation from HWE. 
+where $F$ is the deviation from \gc{Hardy Weinberg Expectations} (HWE). 
 %For analytical solutions presented above, we assume that $F$ equals zero. 
-Assuming HWE a common, recessive, self-promoting driver invades if $s\lesssim(2 d - 1)/(4 d)$, and fixes if
+Assuming HWE \gc{($F=0$)} a common, recessive, self-promoting driver invades if $s\lesssim(2 d - 1)/(4 d)$, and fixes if
  $s\lesssim(2d-1)/2$. 
- Therefore, when \begin{equation}(2 d - 1)/(4 d)  \lesssim  s  \lesssim  (2d-1)/2\label{ineqA}\end{equation} a
- 	recessive, self-promoting driver will deterministically fix if drift,
+ Therefore, when 
+\begin{equation}
+(2 d - 1)/(4 d)  \lesssim  s  \lesssim  (2d-1)/2\label{ineqA}
+\end{equation} 
+a recessive, self-promoting driver will deterministically fix if drift,
  	mutation, or migration pressure bring its frequency above
@@
 % that obtained using the recursion. 
-Compared to exact results (Supplementary Figure \ref{Bistab_homozyg_cost_fig}), Equation
- \eqref{bistab_homozy_approx} offers a reasonable, but imperfect approximation. 
+Compared to exact results (Supplementary Figure \ref{Bistab_homozyg_cost_fig}), Equations
+ \gc{\eqref{ineqA} and} \eqref{bistab_homozy_approx} offers a reasonable, but imperfect approximation. 
 %s_h and seems to serve as a lower bound to the true unstable equilibrium frequency.
@@
 where, $a_1=1-2 ds_h+4 d s-2 d-3 s_h$ and $a_2=2(s-s_h)(2d-1)$.
-\yb{Comparison of Equation \ref{thresh2} to exact results obtained by a simple parameter search (Supplementary Figure \ref{bistable_additive}) show that this approximation is qualitatively reasonable, but quantitatively imperfect.
-}
+Comparison of Equation \ref{thresh2} to exact results obtained by a
+simple parameter search (Supplementary Figure \ref{bistable_additive})
+show that this approximation is reasonablely correct \gc{for small
+parameter values. }
  %range of other costs to the drive alleles other than a
@@
 \subsubsection*{Model 3. Single locus, paternal genotype dependent drive:}
-When female meiotic drive depends on paternal genotype
+\gc{We now turn to the case when} female meiotic drive depends on paternal genotype.
 %\paragraph{Dominance} we also explored a model where
@@
 Assuming that a traditional driver segregates at its equilibrium frequency (Model 1), we investigate the evolution of tightly linked (Models 4 and 5), and unlinked sperm-dependent modifiers of drive. 
-In models 4 and 5, we treat this two loci system as if it consists of three alleles: $A$, $B$ and $C$, which is appropriate if these lowly are tightly linked, such that recombination is unexpected. 
+In models 4 and 5, we treat this two loci system as if it consists of
+three alleles: $A$, $B$ and $C$, corresponding to the case \gc{when the
+sperm-modifier arises once, tightly linked to one or other of the
+alleles (such that recombination is unexpected). }
 In all cases, fitness is independent of the drive modifier, so, $w_{AA}=1$, $w_{AB}=1-s_h$, and $w_{BB}$ = 1-s. 
@@
 %We address two questions concerning the $C$ allele - can it invade when rare  and can it fix when common?
-Although analytical approximation do not reveal simple inequalities like those above,
+Although our analytical approximations (see our \gc{Mathematical supplement}) do not reveal simple inequalities like those above,
 	graphical exploration of analytical results show that this mutant is lost when it enhances drive ($\epsilon>0$ - the $B^+$ allele, above),

### 2014-05-14 writeup_and_notes/male_dependent_drive.tex

@@
 	or is unlinked to the drive locus, it readily invades a population segregating
-	for the drive system ($AB$, see Figure and Equations XXX). 
+	for the drive system ($AB$, see Figure and Equations \ref{A-} and \ref{unlinked}).  
 This allele lowers the frequency of the original drive system (perhaps to zero),

### 2014-05-14 writeup_and_notes/male_dependent_drive.tex
edited through appendix model 5

@@
 \section*{Appendix}
+\gc{Our appendix covers standard female meiotic drive model (1); a sperm
+dependent female meiotic drive model (2); a male genotype dependent
+female meiotic drive model (3); and more complex two locus sperm
+dependent drive models (4-6). }
+
 Below, we highlight some of our derivations and analytical results. 
@@
 alleles (such that recombination is unexpected). }
-In all cases, fitness is independent of the drive modifier, so, $w_{AA}=1$, $w_{AB}=1-s_h$, and $w_{BB}$ = 1-s. 
+In all cases, fitness is independent of the drive modifier, so, $w_{AA}=1$, $w_{AB}=1-s_h$, and $w_{BB} = 1-s$. 
 We assume the traditional driver is transmitted to $d_0$ of gametes from female heterozygotes  when fertilized by wild-type sperm, and $d_1=d_0+\epsilon$ when fertilized by a sperm acting drive modifier.
@@
 \subsubsection*{Model 4. Drive-modifier in coupling phase:}
-When the $C$ allele is tightly linked to and on the same genetic background as the driver, 
+When the $C$ allele is tightly linked to the driver allele, 
 	genotypic fitnesses equal $w_{AC}=w_{AB}=1-s_h$, and $w_{BC}=w_{CC}=w_{BB}=1-s$. 
@@
 	and spreads from rarity to fixation when it suppresses drive ($\epsilon<0$ - the $B^-$ allele, above).	
-This result is supported by exact numerical iterations, as well \yb{FIG?}.
+This result is supported by exact numerical iterations, as well
+\yb{FIG?}. \gc{Lets not bother}
 
 \subsubsection*{Model 5. Drive-modifier in repulsion phase:}
-When the $C$ allele is tightly linked to and on a different background than the driver, 
+When the $C$ allele is tightly linked to the non-driver, 
 	genotypic fitnesses equal $w_{CC}=w_{AC}=w_{AA}=1$, and $w_{BC}=1-s_h$. 
@@
 \subsubsection*{Model 6. Unlinked drive-modifier:}
+\gc{Can we use something other than F, as that was HWE deviation}
 For the unlinked, model we introduce another locus where drive is modified in A/B females fertilized by $F$ allele, 

### 2014-05-15 writeup_and_notes/male_dependent_drive.tex
YB edited supp throuh all single locus models

@@
 \section*{Appendix}
-\gc{Our appendix covers standard female meiotic drive model (1); a sperm
-dependent female meiotic drive model (2); a male genotype dependent
-female meiotic drive model (3); and more complex two locus sperm
-dependent drive models (4-6). }
+Our appendix highlights some of our derivations and analytical results from: 
+	the standard female meiotic drive (Model 1); a sperm
+	dependent female meiotic drive (Model 2); a male genotype dependent
+female meiotic drive (Model 3); and more complex two locus sperm
+dependent drive (Models 4-6). %changed this to minimize confusion with refs, figs, equations, etc.
 
-Below, we highlight some of our derivations and analytical results. 
-The fitness of genotype, $i$ is $w_g$ independent of its sex. 
-We denote its frequency in adults in the current generation is $f_g$. In the next generation of
-zygotes (accounting for drive) but before selection
-the frequency is $f_g'$, and in the adults after selection the
-frequency is $f_g''$ and population mean fitness,  $\bar{w}=\Sigma f_g'w_g$. Genotypic fitnesses are $w_{AA} = 1$, $w_{AB} = 1-s_h$, and $w_{BB}=1-s$, and are sex-independent. 
-All models include a biallelic locus with non-driving and driving alleles in frequencies $f_A$ and $f_B$, respectively, while more complex models include more loci and/or alleles.  
+%Below, we highlight some of our derivations and analytical results. 
+%The fitness of genotype, $g$ is $w_g$ independent of its sex, and its 
+All models include a biallelic locus with non-driving and driving alleles in frequencies $f_A$ and $f_B$, 
+	respectively, while more complex models include more loci and/or alleles.  
 Transmission rules describing the outcomes of all matings in each
-models are presented in a \yb{Supplementary Worksheet}, and  our
-\gc{Mathematical supplement} for our analytical approximations work.  
+	models are presented in a \yb{Supplementary Worksheet}. 
+The fitness of genotype, $g$ is sex-independent and equals $1$, $1-h_s$, and $1-s$ for genotypes $AA$, $AB$, and $BB$, respectively. %  and  $w_{AA} = 1$, $w_{AB} = 1-s_h$, and $w_{BB}=1-s$. 
+Genotypic frequencies  equal  $f_g$ for adults in the current generation, 
+	$f_g'$ in the next generation of zygotes (i.e. after recombination, random mating, drive, and syngamy, but before selection)
+	and $f_g''$ in adults after selection. 
+After a complete generation $f_g'' = f_g' w_g/\bar{w}$, where $\bar{w}$ is the population mean fitness and equals $\Sigma w_g f_g'$. 
+%All models include a biallelic locus with non-driving and driving alleles in frequencies $f_A$ and $f_B$, respectively, while more complex models include more loci and/or alleles.  
+%Transmission rules describing the outcomes of all matings in each
+%models are presented in a \yb{Supplementary Worksheet}, and complete derivations of our analytical approximations are in the
+%\gc{Mathematical Supplement}.  
 Because many of our analyses assume Hardy Weinberg Equilibrium (HWE), we note which analytical results are approximations. 
-We verify these approximations with exact numerical iterations in supplementary figures and scripts, and detail the derivation and details of these results in our mathematical supplement.
+We verify these approximations with exact numerical iterations in Supplementary Figures and Supplementary Scripts, and detail the derivation of these approximations in our Mathematical Supplement.
 
@@
 \subsubsection*{Model 1. Traditional driver}
-In the standard female drive model meiosis in males is fair such that $A/B$ heterozygotes contribute $A$ and $B$ alleles with equal probabilities; however, $A/B$ females transmit the $B$ allele with probability $d>1/2$.
-After drive and random mating genotype frequencies are 
+In the standard female drive model, meiosis in males is fair such that $A/B$ heterozygotes contribute $A$ and $B$ alleles with equal probabilities; however, $A/B$ females transmit the $B$ allele with probability $d>1/2$.
+After drive and random mating, genotype frequencies are 
 	$f_{AA}'=f_A (f_{AA }+ f_{AB} (1 - d))$, 
@@
 %Assuming fitnesses $w_{AA}=1$, $w_{AB}-1 s_h$ and $w_{BB}=1-s$, 
-Frequencies of genotype, $i$, after drive and selection are 
-	\begin{equation} f_i''= f_i'w_i/\bar{w}\end{equation}
-	where $w_i$ is genotypic fitness, and mean fitness, $\bar{w} = \Sigma f_i'w_i$. 
+%Frequencies of genotype, $i$, after drive and selection are 
+%	\begin{equation} f_i''= f_i'w_i/\bar{w}\end{equation}
+%	where $w_i$ is genotypic fitness, and mean fitness, $\bar{w} = \Sigma f_i'w_i$. 
+As detailed above, exact frequencies after drive, random mating and selection are $f_g''= f_g'w_g/\bar{w}$. 
 Assuming HWE, a rare driver will spread when $s_h
-        \lesssim(2d-1)/(1+2d)$, and will fix when $s\lesssim d -1/2 +
+        \lessapprox(2d-1)/(1+2d)$, and will fix when $s\lessapprox d -1/2 +
         3 s_h /2- d s_h$. This later inequality reduces to
-        $s\lesssim(2d-1)/2$ when the cost of drive is fully
+        $s\lessapprox(2d-1)/2$ when the cost of drive is fully
         recessive. \gc{isn't this last expression exact?}
@@
 	and $f_{BB}'= f_B (f_{AB} d + f_{BB} )$. 
-As above, exact frequencies after drive, random mating and selection are $f_i''= f_i'w_i/\bar{w}$. 
-We iterate this recursion over generations, to produce the frequency trajectories shown in the inset
-Figure \ref{Invasion_space} by plotting \gc{$f_B=f_{BB}+ \frac{1}{2}g_{AB}$}
-over time, and to assess the analytical approximations, below. \gc{To
-assess invasion or fixation criteria we iterate this system and
+%As above, exact frequencies after drive, random mating and selection are $f_i''= f_i'w_i/\bar{w}$. 
+We iterate exact genotype frequency recursions ($f_g''= f_g'w_g/\bar{w}$) over generations to produce the frequency trajectories shown in the inset
+Figure \ref{Invasion_space} by plotting \gc{$f_B=f_{BB}+ \frac{1}{2}f_{AB}$}
+over time. %, and to assess the analytical approximations, below.
+ \gc{To assess invasion or fixation criteria, \yb{as well as bistability points,} we iterate this system and
 testing whether $f_B$ increases over a grid of parameters. }
@@
 \end{equation}
-where $F$ is the deviation from \gc{Hardy Weinberg Expectations} (HWE). 
+where $F$ is the deviation from genotypic frequencies expected under Hardy-Weinberg. % Trying to clarify some.   \gc{Hardy Weinberg Expectations} (HWE). 
 %For analytical solutions presented above, we assume that $F$ equals zero. 
-Assuming HWE \gc{($F=0$)} a common, recessive, self-promoting driver invades if $s\lesssim(2 d - 1)/(4 d)$, and fixes if
- $s\lesssim(2d-1)/2$. 
+Assuming HWE  ($F=0$) a common, recessive, self-promoting driver invades if $s\lessapprox(2 d - 1)/(4 d)$, and fixes if
+ $s\lessapprox(2d-1)/2$. 
  Therefore, when 
 \begin{equation}
-(2 d - 1)/(4 d)  \lesssim  s  \lesssim  (2d-1)/2\label{ineqA}
+(2 d - 1)/(4 d)  \lessapprox  s  \lessapprox  (2d-1)/2\label{ineqA}
 \end{equation} 
@@
 Compared to exact results (Supplementary Figure \ref{Bistab_homozyg_cost_fig}), Equations
- \gc{\eqref{ineqA} and} \eqref{bistab_homozy_approx} offers a reasonable, but imperfect approximation. 
+ \gc{\eqref{ineqA} and} \eqref{bistab_homozy_approx} offer reasonable, but imperfect approximations. 
 %s_h and seems to serve as a lower bound to the true unstable equilibrium frequency.
@@
 simple parameter search (Supplementary Figure \ref{bistable_additive})
-show that this approximation is reasonablely correct \gc{for small
-parameter values. }
+show that this approximation is reasonably correct for small
+parameter values; \yb{however, this approximation underestimates $f_B^*$ for large parameter values, presumably because they result in strong departures from HWE.}
  %range of other costs to the drive alleles other than a
@@
 	a rare paternal genotype dependent driver invades when 
-	$s\lesssim (d_{het}-1/2)/(d_{het})$, and a common paternal genotype dependent driver fixes when 
-	$s\lesssim d_{hom}-1/2$, approximations well supported by exact results (Supplementary Figure \ref{Effect_of_dominance}).
+	$s\lessapprox (d_{het}-1/2)/(d_{het})$, and a common paternal genotype dependent driver fixes when 
+	$s\lessapprox d_{hom}-1/2$, approximations well supported by exact results (Supplementary Figure \ref{Effect_of_dominance}).
 Specifically, when drive in heterozygotes is large relative to that in homozygotes, 
-	\begin{equation} d_{het} \gtrsim 1/(3-2d_{hom}) \label{polymale} \end{equation}
+	\begin{equation} d_{het}  \gtrapprox 1/(3-2d_{hom}) \label{polymale} \end{equation}
 	fixation criteria are more stringent than invasion criteria, 
@@
 \begin{equation}f_{B\text{ pat}}^*\approx2 (1 + 2d_{hom}[s - 1])/([2 s - 1][2d_{hom}- 1]) \label{eqmale} \end{equation}
-By contrast, when $d_{het} \lesssim 1/(3-2d_{hom})$ the case is reversed, and the model is bistable.
+By contrast, when $d_{het} \lessapprox 1/(3-2d_{hom})$ the case is reversed, and the model is bistable.
 %In Supplementary Figure \ref{Effect_of_dominance} we show the invasion

### 2014-05-15 writeup_and_notes/male_dependent_drive.tex
added delta C for coupling phase model instead of saying we got nothing

@@
 \subsubsection*{Model 3. Single locus, paternal genotype dependent drive:}
-\gc{We now turn to the case when} female meiotic drive depends on paternal genotype.
+We now turn to the case when female meiotic drive depends on paternal genotype.
 %\paragraph{Dominance} we also explored a model where
@@
 	genotypic fitnesses equal $w_{AC}=w_{AB}=1-s_h$, and $w_{BC}=w_{CC}=w_{BB}=1-s$. 
+Assuming HWE, a recessive fitness cost to drive, and assuming that the $A/B$ locus is at its equilibrium frequency, the change in frequency of a rare drive modifier is
+\begin{equation}
+	\Delta C\approx -\epsilon f_C \frac{ 1-2 d_0 (-2 d_0+s+2)+\sqrt{2} \sqrt{s (2 d_0 (d_0(s-2)+2)-1)}}{2 \bar{w}(1-2 d_0)^2}
+\end{equation}
+%YB Considered this ugly math more impressive.
+For all parameters sustaining a polymorphism at the drive locus ($s>d_0-1/2$), this corresponds to a decrease in frequency of the $C$ allele when it enhances drive ($\epsilon >0$ -- the $B^+$ model, above),
+	and a decrease in frequency of the $C$ allele when it suppresses drive ($\epsilon <0$ -- the $B^-$ model, above). 
+More generally even when the cost of drive is not fully recessive, the $B^-$ allele will invade and fix under all parameters sustaining a previous polymorphism at the drive locus (see Mathematical Supplement).
+%When the C allele acts to suppress drive for all parameters sustaining a polymorphism at the drive locus ($s>d_0-1/2$)
 %We address two questions concerning the $C$ allele - can it invade when rare  and can it fix when common?
-Although our analytical approximations (see our \gc{Mathematical supplement}) do not reveal simple inequalities like those above,
-	graphical exploration of analytical results show that this mutant is lost when it enhances drive ($\epsilon>0$ - the $B^+$ allele, above),
-	and spreads from rarity to fixation when it suppresses drive ($\epsilon<0$ - the $B^-$ allele, above).	
-This result is supported by exact numerical iterations, as well
-\yb{FIG?}. \gc{Lets not bother}
+%Although our analytical approximations (see our \gc{Mathematical supplement}) do not reveal simple inequalities like those above,
+%	graphical exploration of analytical results show that this mutant is lost when it enhances drive ($\epsilon>0$ - the $B^+$ allele, above),
+%	and spreads from rarity to fixation when it suppresses drive ($\epsilon<0$ - the $B^-$ allele, above).	
+%This result is supported by exact numerical iterations, as well
+%\yb{FIG?}. \gc{Lets not bother}
 
@@
 \begin{equation}
-	\Delta C^- \approx -\epsilon f_{C^-}^2 \frac{ \left(2 d_0 s-\sqrt{2} \sqrt{s (2 d_0 (d_0 (s-2)+2)-1)}\right)}{2\bar{w}s (2 d_0 -1) }. \label{A-}
+	\Delta C \approx -\epsilon f_{C^-}^2 \frac{ \left(2 d_0 s-\sqrt{2} \sqrt{s (2 d_0 (d_0 (s-2)+2)-1)}\right)}{2\bar{w}s (2 d_0 -1) }. \label{A-}
 \end{equation}
@@
 \subsubsection*{Model 6. Unlinked drive-modifier:}
-\gc{Can we use something other than F, as that was HWE deviation}
-For the unlinked, model we introduce another locus where drive is modified in A/B females fertilized by $F$ allele, 
-	while the wild-type $E$ does not influence drive. 
+%\gc{Can we use something other than F, as that was HWE deviation}
+For the unlinked, model we introduce another locus where drive is modified in A/B females fertilized by $M$ allele, 
+	while the wild-type $L$ does not influence drive. 
 Assuming HWE and linkage equilibrium, the change in frequency of a rare unlinked, sperm-acting drive modifier is 
-\begin{equation} \Delta F =-\epsilon f_A f_B (f_B s + s_h (fA - fB) ) \frac{f_F}{\bar{w}}  \label{unlinked} \end{equation}
+\begin{equation} \Delta M =-\epsilon f_A f_B (f_B s + s_h (fA - fB) ) \frac{f_M}{\bar{w}}  \label{unlinked} \end{equation}
 % an $F$ allele increases in frequency when 

### 2014-05-15 writeup_and_notes/male_dependent_drive.tex
fixed minor typos in abstract

@@
 Because only one of four products of female meiosis is transmitted to offspring, oogenesis is particularly
-	 susceptible to genomic conflicts arising when alleles gains in frequency at a cost to organismal fitness. 
+	 susceptible to genomic conflicts arising when an allele gains in frequency at a cost to organismal fitness. 
 In fact, many facets of oogenesis have been interpreted as mechanisms of protection against genomic outlaws. 
@@
 	sperm nearly always evolve to increase the fairness of female meiosis when possible. 
-These results are constant with current knowledge of sperm-dependent meiotic drivers, 
+These results are consistent with current knowledge of sperm-dependent meiotic drivers, 
 	and suggest that the requirement of fertilization for the completion of meiosis may represent 

### 2014-05-15 writeup_and_notes/male_dependent_drive.tex
pushing ybs minor edits

@@
 {\bf Abstract:}
-Because only one of four products of female meiosis is transmitted to offspring, oogenesis is particularly
+Because only one of four products of female meiosis is transmitted to offspring, o\"{o}genesis is particularly
 	 susceptible to genomic conflicts arising when an allele gains in frequency at a cost to organismal fitness. 
@@
  While ruling out the alternative hypothesis of early selection on zygotes \citep[pages
-52-54][]{Burt2006} in these cases is challenging, it appears that the extent to which \emph{In} and \emph{Om} distort the second female meiotic division partially depends on the genotype of the fertilizing sperm \citep{Agulnik1993,Wu2005}.
+52-54 in][]{Burt2006} in these cases is challenging, it appears that the extent to which \emph{In} and \emph{Om} distort the second female meiotic division partially depends on the genotype of the fertilizing sperm \citep{Agulnik1993,Wu2005}.
 % Both of these female drive systems operate by distorting the second meiotic division, 
@@
 the driving allele is transmitted to the egg  in heterozygotes with
-probability  $d > 1/2$, regardless of sperm genotype. To depict a case
+probability  $d > 1/2$, regardless of sperm genotype (Model 1). To depict a case
 of a self-promoting meiotic driver,  we modify this standard model such 
-	that the driver is only effective when fertilized by a sperm carrying that allele (see Figure \ref{Eggsperm_2_allele_cartoon})
+	that the driver is only effective when fertilized by a sperm carrying that allele (see Figure \ref{Eggsperm_2_allele_cartoon} and Model 2 in the Supplementary Worksheet)
 \footnote{the timing of fertilization relative to female meiosis places another constraint on $d$, for example, if fertilization (and therefore, sperm dependent drive) takes place at MII (as in mammals),
@@
 
-For comparative purposes, we present the standard drive model as Model 1 in the Appendix, 
+For comparative purposes, we briefly present the standard drive model %as Model 1 in the Appendix, 
 %equations \ref{eqn.A1}-\ref{eqn.A3}, 
-	and refer readers to the previous literature on this model \citep[e.g. ][]{Prout1973,Ubeda2004} for additional results. 
-In this case, assuming the driving allele is deleterious in both sexes but fully recessive 
+	and refer readers to the previous literature on this model 
+	\citep[see e.g. ][ for additional results]{Prout1973,Ubeda2004}. 
+Assuming the driving allele is deleterious in both sexes, but fully recessive 
 	(i.e. the fitness of drive homozygotes equals $w_{BB}=1-s$ and other genotypic fitnesses equal $w_{AA}=w_{AB}=1$), 
 	it always invades because,  when rare, it occurs predominantly in heterozygotes and therefore drives without a fitness cost. 
-However, when $s$ is large ($s>(2d-1)/(2)$) a driver cannot fix, and
-will be maintained as a protected polymorphism \citep{Prout1973}. The
-        parameter space where the allele can invade but not fix is shown in white
-        in Figure \ref{Invasion_space}A, and where it can fix in the green
-        and orange areasin Figure \ref{Invasion_space}B. When the allele is
-        maintained as a polymorphism, it provides an opportunity for the evolution of
+However, when $s$ is large ($s>(2d-1)/(2)$), a driver cannot fix, and
+will be maintained as a protected polymorphism \citep{Prout1973}. 
+The parameter space where the allele can invade but not fix is shown in white
+        in Figure \yb{\ref{Invasion_space}A}, and it can fix in the colored regions. 
+When the allele is maintained as a polymorphism, it provides an opportunity for the evolution of
 	drive suppressors, corresponding well to empirical examples of

### 2014-05-16 writeup_and_notes/male_dependent_drive.tex
modifierd absrtact and intro, new title suggestion

@@
 %Or 
-Sperm do not evolve to collaborate in female meiotic drive. 
+Sperm make female meiosis fair.
+%Sperm do not evolve to collaborate in female meiotic drive. 
 %Or Why does female meiosis allow the opportunity for exploitation by selfish sperm? 
@@
 {\bf Abstract:}
-Because only one of four products of female meiosis is transmitted to offspring, o\"{o}genesis is particularly
-	 susceptible to genomic conflicts arising when an allele gains in frequency at a cost to organismal fitness. 
-In fact, many facets of oogenesis have been interpreted as mechanisms of protection against genomic outlaws. 
+Genomic conflicts arise when an allele gains an evolutionary advantage at a cost to organismal fitness. 
+Because only one of four products of female meiosis is transmitted to offspring, o\"{o}genesis is inherantly susceptible to  such conflicts. 
+In fact, many facets of o\"{o}genesis have been interpreted as mechanisms of protection against genomic outlaws. 
 Why then is female meiosis often left uncompleted until after fertilization - 
-	potentially providing an opportunity for sperm to meddle with its outcome and facilitate meiotic drive? 
+	potentially providing an opportunity for sperm to meddle with its outcome? %and facilitate meiotic drive? 
 The population genetic theory presented herein suggests that
@@
 \section*{Introduction}
-Despite the apparent unity of the organism, alleles can
+Despite the apparent unity of the organism, `selfish' alleles can
 gain an evolutionary advantage at a cost to  individual fitness
@@
 creating ample opportunity for interaction between the sperm and female meiotic machinery. 
-Therefore, a `green-bearded' \citep{Gardner2010} sperm-acting allele that recognizes and facilitates the meiotic drive of a genetically equivalent allele in female meiosis %(a  heteromorphic dyad), 
+Therefore, a `green-bearded' \citep{Gardner2010} sperm-acting allele that recognizes and facilitates the meiotic drive of a genetically equivalent allele in heterozygous females %female meiosis %(a  heteromorphic dyad),
      could presumably rapidly spread through a population. 
@@
 The stage at which female meiosis requires fertilization varies widely across taxa, and in a few animals, 
-	female meiosis is complete upon ovulation (that is, fertilization is not required for female meiosis \citep[see table 1 of ref 
+	female meiosis is complete upon ovulation (that is, fertilization is not required for female meiosis \citep[e.g. Table 1 in 
 ][]{Masui_book}).
@@
 
-We develop population genetic models exploring the evolution of sperm influence on female meiotic drive, focussing on models in which `self-promoting' alleles in sperm facilitate drive of like genotypes during gametogenesis in heterozygous females.
+We develop population genetic models exploring the evolution of sperm influence on female meiotic drive.
+%, focussing on models in which `self-promoting' alleles in sperm facilitate drive of like genotypes during gametogenesis in heterozygous females.
+\yb{We first focus on models in which `self-promoting' alleles in sperm facilitate drive of like genotypes during gametogenesis in heterozygous females. }
 %In this article we explore through simple population genetic models the consequences of alleles that influence the outcome of female meiosis.
@@
 	female meiosis will have little incentive to evolve resistance to them.
-In fact, the interests of sperm and egg  genomes' are often aligned, as both are invested in the fate of the resultant zygote (as was speculated for the \emph{In} locus \citep{Pomiankowski1993}).
-Therefore, there is little selective benefit to females in preventing sperm to influence female meioses,
-	indeed features of meiosis may evolve that facilitate the interaction of sperm with female meiosis.
+\yb{We then examine models in which an allele in sperm acts to modify the efficacy of a polymorphic meiotic driver - 
+	which universally reveal that sperm acting drive modifiers are favored only if the suppress drive. }
+The results highlight the fact that the interests of sperm and egg  genomes' are often aligned, as both are invested in the fate of the resultant zygote (as was speculated for the \emph{In} locus \citep{Pomiankowski1993}).
+Therefore, there is little selective benefit to females in preventing sperm to influence female meioses.
+\yb{Rather, features of o\"{o}genesis  may evolve to facilitate an interaction with sperm %with female meiosis
+	as yet another mechanism to enforce an equitable transfer of genetic material.}
 

### 2014-05-16 writeup_and_notes/male_dependent_drive.tex
changes to abstract

--- 2014-05-15 writeup_and_notes/male_dependent_drive.tex
@@
 Because only one of four products of female meiosis is transmitted to offspring, oogenesis is particularly
-	 susceptible to genomic conflicts arising when an allele gains in frequency at a cost to organismal fitness. 
+	 susceptible to genomic conflicts arising when an allele
+         spreads at a cost to organismal fitness. 
 In fact, many facets of oogenesis have been interpreted as mechanisms of protection against genomic outlaws. 

### 2014-05-16 writeup_and_notes/male_dependent_drive.tex

--- 2014-05-16 writeup_and_notes/male_dependent_drive.tex
@@
 %Or 
-Sperm make female meiosis fair.
+Sperm should evolve to make female meiosis fair.
 %Sperm do not evolve to collaborate in female meiotic drive. 
@@
 Genomic conflicts arise when an allele gains an evolutionary advantage at a cost to organismal fitness. 
-Because only one of four products of female meiosis is transmitted to offspring, o\"{o}genesis is inherantly susceptible to  such conflicts. 
-In fact, many facets of o\"{o}genesis have been interpreted as mechanisms of protection against genomic outlaws. 
+Because only one of four products of female meiosis is transmitted to
+offspring, o\"{o}genesis is inherantly susceptible to such conflicts,
+as alleles can compete to be transmitted into the egg
+(meiotic drive). 
+Genomics conflicts drive the rapid evolution of surpression mechanisms, and
+many facets of gametogenesis have been interpreted as mechanisms of protection against genomic outlaws. 
 Why then is female meiosis often left uncompleted until after fertilization - 
-	potentially providing an opportunity for sperm to meddle with its outcome? %and facilitate meiotic drive? 
+	potentially providing an opportunity for sperm alleles to
+        meddle with its outcome in their own favor? %and facilitate meiotic drive? 
 The population genetic theory presented herein suggests that
-	sperm nearly always evolve to increase the fairness of female meiosis when possible. 
+	sperm nearly always evolve to increase the fairness of female
+        meiosis in the face of genomic conflicts. %%felt this
+                                %%might be an important addition as
+                                %%they will spread if no cost to organism
 These results are consistent with current knowledge of sperm-dependent meiotic drivers, 
-	and suggest that the requirement of fertilization for the completion of meiosis may represent 
-	yet another mechanism of female meiosis designed to prevent meiotic drive.
+	and suggest that the requirement of fertilization for the
+        completion of female meiosis, and variation in stage of meiosis that
+        fertilization occurs at 
+ may represent a mechanism to prevent meiotic drive.
 \newpage

### 2014-05-16 writeup_and_notes/male_dependent_drive.tex
YB modified and edited results

--- 2014-05-16 writeup_and_notes/male_dependent_drive.tex
@@
 
-\subsection*{ Invasion of the population by a driving allele that promotes
-itself.}
+\subsection*{ Invasion of the population by a driving allele that promotes itself.}
 %%%%%%%%%%%%%%%%%%%%%%%%%
@@
 	and evaluate whether a driver of this form could generate a sustained conflict favoring the evolution of suppressors. 
+\yb{We concluded with an analysis of a related mode (Model 3) - in which drivers influence their transmission in females via paternal genotype, rather than sperm haplotype.}
 
@@
 %equations \ref{eqn.A1}-\ref{eqn.A3}, 
-	and refer readers to the previous literature on this model 
+	%and refer readers to the previous literature on this model 
 	\citep[see e.g. ][ for additional results]{Prout1973,Ubeda2004}. 
@@
 The parameter space where the allele can invade but not fix is shown in white
-        in Figure \yb{\ref{Invasion_space}A}, and it can fix in the colored regions. 
+        in Figure \yb{\ref{Invasion_space}A}. %, and it can fix in the colored regions. 
 When the allele is maintained as a polymorphism, it provides an opportunity for the evolution of
@@
 	no fitness cost when rare, 
-	a self-promoting driver specifically creates homozygote
-        zygotes when it drives. %in heterozygotes. 
+\yb{	a self-promoting driver specifically creates low fitness drive homozygotes 
+	by uniting driving female gametes with sperm enabling that drive.} 
+        % zygotes when it drives. in heterozygotes. 
 %, and therefore there is a drive-associated fitness cost even when rare. 
@@
 	positive frequency dependent behavior (Inset in Figure \ref{Invasion_space}).  
-This lack of a balanced polymorphism %(Figure \ref{Invasion_space}) 
-	precludes the evolution of an
-	allele that suppresses this form of meiotic drive in females.
+%This lack of a balanced polymorphism %(Figure \ref{Invasion_space}) 
+%	precludes the evolution of an
+%	allele that suppresses this form of meiotic drive in females.
+The positive frequency dependence of a self-promoting driver can also
+	induce bistability in its spread -- some values of $s$ 
+	allow the fixation of this driver when common, but 
+	preclude its invasion when rare (Equation \ref{ineqA}). 
+In this case, the driver will be fixed if it reaches a frequency greater than some threshold, 
+	(approximated in Equation \ref{bistab_homozy_approx} and  presented exactly in 
+	Figure \ref{Bistab_homozyg_cost_fig}), and lost otherwise. 
+For most parameters, this threshold is likely too high to be reached by drift, 
+	and therefore  the fate of self-promoting drivers is determined
+	by the more restrictive invasion criteria rather than the fixation criteria.
 
 %I SUBSTANTIALLY REWROTE THIS
-The positive frequency dependence of a self-promoting driver can
-induce bistability in its spread.
+%The positive frequency dependence of a self-promoting driver can
+%induce bistability in its spread.
 %OLD GRAHAM WORDS
@@
 %approximation to this bistability. 
-When $s$ is too large to allow a rare driver to increase in frequency, but small enough to allow a common driver to fix (the inequality in Equation \ref{ineqA}),
-there is  an unstable equilibrium frequency, $f^*_{B\text{ recessive}}$  
-	above which it can spread quickly to fixation and below which it moves towards loss 
-	($f^*_{B\text{ recessive}}$ is approximated in Equation \ref{bistab_homozy_approx}.  
-	We compare this approximation to the exact result in Figure \ref{Bistab_homozyg_cost_fig}).	
+%When $s$ is too large to allow a rare driver to increase in frequency, but small enough to allow a common driver to fix (the inequality in Equation \ref{ineqA}),
+%there is  an unstable equilibrium frequency, $f^*_{B\text{ recessive}}$  
+%	above which it can spread quickly to fixation and below which it moves towards loss 
+%	($f^*_{B\text{ recessive}}$ is approximated in Equation \ref{bistab_homozy_approx}.  
+%	We compare this approximation to the exact result in Figure \ref{Bistab_homozyg_cost_fig}).	
 %OLD GRAHAM WORDS
@@
 %\ref{Bistab_homozyg_cost_fig}).
-This unstable equilibrium frequency is high across much of
-	parameter space, likely too high to be reached by drift, 
-	and therefore  the fate of self-promoting drivers  will  likely be determined
-	by the more restrictive invasion criteria rather than the fixation criteria.
+%This unstable equilibrium frequency is high across much of
+%	parameter space, likely too high to be reached by drift, 
+%	and therefore  the fate of self-promoting drivers  will  likely be determined
+%	by the more restrictive invasion criteria rather than the fixation criteria.
 
@@
 In fact, with any heterozygous fitness cost, a rare self-promoting
-driver cannot spread deterministically. 
+	driver is always selected against. 
 However, this case also displays bistability -- 
-	mathematical approximation assuming Hardy-Weinberg reveal that 
-	if $s < d - 1/2+s_h(3-2d)/2$, it will fix deterministically if its 
-	frequency is greater than a threshold, $f_B^*$ [Equation \eqref{thresh2}. 
-	See \yb{FIGURES SXX} for exact results].
+	when $s$ is sufficiently small (Equation \ref{sForHetFix}) this allele fixes deterministically if its 
+	frequency exceeds some threshold,  (Equation \ref{thresh2}, exact results in 
+	\yb{FIGURES SXX}).
+%	$f_B^*$
 %	because it pays the  heterozygous fitness cost but drives
@@
 This bistability prevents self-promoting drivers from invading 	
-	reasonably sized populations, and assures that, if they do invade, they will rapidly fix.
+	reasonably sized populations, and assures that if they do invade, they will rapidly fix.
 Our model therefore predicts that self-promoting drivers will not be
 	observed as stable polymorphisms in natural populations. 
+This lack of a balanced polymorphism %(Figure \ref{Invasion_space}) 
+	precludes the evolution of an
+	allele that suppresses this form of meiotic drive in females.
 
@@
 	diploid genotype (perhaps due to proteins and RNAs packaged into the sperm), rather than sperm haplotype.
-This male-genotype dependent model requires one additional parameter, as we exchange $d$ in the sperm dependent case for $d_{het}$ and $d_{hom}$ which describe the transmission of the drive allele in a heterozygous female mating with males heterozygous and homozygous for the self-promoting drive allele, respectively.
+This paternal-genotype dependent model (Model 3) requires one additional parameter, as we exchange $d$ in the sperm dependent case for $d_{het}$ and $d_{hom}$ which describe the transmission of the drive allele in a heterozygous female mating with males heterozygous and homozygous for the self-promoting drive allele, respectively.
 Here, a rare driver invades when $s$ is less than $(d_{het}-1/2)/d_{het}$, and usually fixes when it invades.
@@
 	when the cost of the drive is recessive
-	(see Figure S\ref{Effect_of_dominance}, \yb{and Equation \ref{polymale}}).  
+	(see Figure S\ref{Effect_of_dominance}, and Equation \ref{polymale}).  
 %Presumably this zone of polymorphism comes about because positive frequency 
@@
 	male genotype, while the cost would be recessive. 
-Therefore, although this model can sustain a polymorphism, 
+Therefore, although Model 3 can sustain a polymorphism, 
 	the lack of biological reality underlying the small portion of parameter values 
@@
 
-\subsection*{ sperm-egg haplotype balanced drive systems}
+\subsection*{Two locus models of sperm-dependent female drive}
+
+%\subsection*{ sperm-egg haplotype balanced drive systems}
 
@@
 
-Above, we explored a model of an allele drove in females when signaled by a complementary allele in sperm.  
-We complement this single locus model %of an allele with pleiotropic effects in both sexes, 
-	with  an alternative model of two linked loci - one a female driver, 
-	and the other a sperm allele which modifies the effect of drive upon fertilization. 
+Models 2 and 3, above, explored the dynamics of an allele drove in females when signaled by a complementary signal in sperm.   
+We complement this single-locus approach %of an allele with pleiotropic effects in both sexes, 
+	with alternative models of two linked loci - one a female driver, 
+	and the other, a sperm allele which modifies the effect of drive upon fertilization. 
 In this model, a female meiotic driver with no sperm-dependency 
@@
         linked to this balanced locus (effectively creating a third
-        allele/haplotype at this locus). Tight linkage offers the best
+        allele/haplotype at this locus). 
+Tight linkage offers the best
         chance for a collaboration to evolve between a driver and
        an enhancer of the drive, as recombination breaks up drive haplotypes \citep{Thomson1974,Charlesworth1978,Haig1991}. 
-In addition tight linkage between female driver and sperm modifier is consistent with the nature of well characterized drive systems which are often maintained as polymorphic inversions with numerous linked modifiers (CITE).
+Additionally, tight linkage between female driver and sperm modifier is consistent with the nature of well characterized drive systems which are often maintained as polymorphic inversions with numerous linked modifiers (CITE).
 
@@
 %%%sperm-acting facilitator
-We first consider a sperm-acting enhancer of
-	female meiotic drive tightly linked to the $B$ background, 
-	generating a third `$B^{+}$' allele (i.e. the non-recombining haplotype of a
-        driver allele and sperm-acting drive modifier). 
-This new allele acts in sperm to increase the effectiveness of both
-	the $B$ and  $B^{+}$ drive in $AB$ and $AB^{+}$ heterozygotes (see Figure \ref{Eggsperm_3_allele_cartoon})
+We first consider a sperm-acting modifier of
+	female meiotic drive tightly linked to the $B$ background (Model 4).
+We treat this non-recombining drive/modifier haplotype as a third allele, and we label this allele as $B^+$ when it enhances drive. 
+%(i.e. the non-recombining haplotype of a
+ %       driver allele and sperm-acting drive modifier)
+%The $B^+$ allele 
+%	the $B$ and  $B^{+}$ drive in $AB$ and $AB^{+}$ heterozygotes (see Figure \ref{Eggsperm_3_allele_cartoon})
+%		generating a third `$B^{+}$' allele (i.e. the non-recombining haplotype of a
+   %     driver allele and sperm-acting drive modifier). 
+The $B^+$ allele acts in sperm to increase the effectiveness of drive for both
+	the $B$ and  $B^{+}$ alleles in $AB$ and $AB^{+}$ heterozygotes (see Figure \ref{Eggsperm_3_allele_cartoon}). 
 %	for a simple schematic of this drive model.  
-Naively, the $B^{+}$ may spread by capitalizing on the additional drive it causes; however,  
+Naively, $B^{+}$ may spread by capitalizing on the additional drive it causes; however,  
 	this is not the case for a few simple reasons. 
-First, the sperm-acting drive facilitator (the novel $B^{+}$ haplotype) 
-	arises when the ancestral driver ($B$) is at drive-selection balance, 
+First, %the sperm-acting drive facilitator (
+	the novel $B^{+}$ haplotype
+	%) 
+	arises when the ancestral driver 
+	%($B$) 
+	is at drive-selection balance, 
 	and therefore immediately suffers a homozygous fitness cost.  
@@
 	the $B$  allele drive ($B^+$ sperm meeting $AB$ eggs), because $B$ is initially more common than $B^{+}$. 
-As $B^{+}$ facilitates $B$ these interactions frequently form 
-	$BB^{+}$ heterozygotes that suffer the same fitness costs as $BB$. 
-Therefore, sperm-acting drive facilitator alleles experience a profound disadvantage 
+%As $B^{+}$ facilitates $B$ these interactions frequently form 
+%	$BB^{+}$ heterozygotes that suffer the same fitness costs as $BB$. 
+Therefore, a sperm-acting drive facilitator alleles experience a profound disadvantage 
 	in this scenario, even more so than under the previous two allele model. 
@@
 	three allele system that allows the sperm-acting drive facilitator $B^{+}$ to
-	invade the population.
+	invade the population \yb{(Mathematical Supplement and Figure XXX)}.
 
 %%%sperm-acting drive suppressors
-While  sperm enhancement of a female drive cannot displace a polymorphic female driver, sperm based drive suppression can. 
+While  sperm enhancement of a female drive cannot displace a polymorphic female driver, sperm based drive suppressors can. 
 %	that is sperm-acting alleles that act to restore  fairness of female meiosis, can. %If these sperm-acting drive suppressors arise on
-If a sperm-acting allele that restores fairness to female meiosis arises on
-	the non-driving $A$ background (creating a third allele $A^{-}$), 
-	or is unlinked to the drive locus, it readily invades a population segregating
-	for the drive system ($AB$, see Figure and Equations \ref{A-} and \ref{unlinked}).  
+\yb{If a sperm-acting allele that restores fairness to female meiosis arises on
+	the drive background (creating a third allele $B^{-}$, Model 4), it displaces the ancestral driver (Equation \ref{coupling}), 
+	and often returns to a lower equilibrium frequency than the the $B$ allele, further decreasing the extent of drive in the population. } 
+If this sperm-acting drive suppressor arises on 
+	the non-driving $A$ background (creating a third allele $A^{-}$, Model 5), 
+	or is unlinked to the drive locus (Model 6), it readily invades a population segregating
+	for the drive system (see \yb{Figures XXX} and Equations \ref{A-} and \ref{unlinked}).  
 This allele lowers the frequency of the original drive system (perhaps to zero),
@@
 This result is consistent with previous work showing that drive suppressors unlinked to, or in repulsion phase with drivers usually invade polymorphic drive systems \citep[e.g. ][]{Brandvain2012}.  
+\yb{Therefore, all two-locus models of sperm influence on female drive suggest that
+	sperm will evolve to oppose female meiotic drive.}
 
-More surprisingly, sperm-acting suppressors of female drive can spread
-	even when tightly linked to the original drive system (Figures \ref{Eggsperm_3_allele_cartoon} and \ref{Trajectories_of_supressors}B). 
-The $B^{-}$ haplotype (representing the tightly linked drive/drive suppressor complex) displaces the original driver 
-	by driving when not costly, but avoiding low fitness homozygotes by disrupting drive when transmitted to sperm. %  ($BB^-$ and $B^-B^-$). 
-Despite displacing the ancestral driver (the $B$ haplotype), the equilibrium frequency of the $B^-$ haplotype is often lower than that of $B$ \yb{(point to supp?)}.
+%More surprisingly, sperm-acting suppressors of female drive can spread
+%	even when tightly linked to the original drive system (Figures \ref{Eggsperm_3_allele_cartoon} and \ref{Trajectories_of_supressors}B). 
+%The $B^{-}$ haplotype (representing the tightly linked drive/drive suppressor complex) displaces the original driver 
+%	by driving when not costly, but avoiding low fitness homozygotes by disrupting drive when transmitted to sperm. %  ($BB^-$ and $B^-B^-$). 
+%Despite displacing the ancestral driver (the $B$ haplotype), the equilibrium frequency of the $B^-$ haplotype is often lower than that of $B$ \yb{(point to supp?)}.
 
@@
 %So long as there is any heterozygous fitness cost (i.e. $s_h > 0$) the self-promoting driver cannot invade when rare;
-However, this system too, is bistable -- as the driver becomes more frequency it is more often fertilized by a driving sperm and therefore drives more effectively.
-%Assuming no deviation from Hardy-Weinberg a self-promoting driver with a heterozygous fitness cost 
-This allele deterministically fixes if its frequency exceeds some threshold, $f_B^*$.
-Assuming HWE, this threshold equals 
+However, this system too, is bistable -- as the driver increases in frequency it is more often fertilized by a driving sperm and therefore drives more effectively. 
+Therefore, assuming HWE, if 
+\begin{equation}
+	s < d -1/2 +s_h(3-2d)/2
+	\label{sForHetFix}
+\end{equation}
+this self-promoting driver deterministically fixes when its frequency is greater than
 \begin{equation}
@@
 	\Delta C\approx -\epsilon f_C \frac{ 1-2 d_0 (-2 d_0+s+2)+\sqrt{2} \sqrt{s (2 d_0 (d_0(s-2)+2)-1)}}{2 \bar{w}(1-2 d_0)^2}
+	\label{coupling}
 \end{equation}

### 2014-05-16 writeup_and_notes/male_dependent_drive.tex
small changes to intro

--- 2014-05-16 writeup_and_notes/male_dependent_drive.tex
@@
 %, focussing on models in which `self-promoting' alleles in sperm facilitate drive of like genotypes during gametogenesis in heterozygous females.
-\yb{We first focus on models in which `self-promoting' alleles in sperm facilitate drive of like genotypes during gametogenesis in heterozygous females. }
+We first focus on models in which `self-promoting' alleles in
+  sperm facilitate drive of like \gc{alleles} during gametogenesis in heterozygous females. 
 %In this article we explore through simple population genetic models the consequences of alleles that influence the outcome of female meiosis.
 These models show that sperm modifiers of female meiotic drive are
-	unlikely to create a sustained conflict, and therefore 
+	unlikely to create a sustained genomic conflict, and therefore 
 	female meiosis will have little incentive to evolve resistance to them.
-\yb{We then examine models in which an allele in sperm acts to modify the efficacy of a polymorphic meiotic driver - 
-	which universally reveal that sperm acting drive modifiers are favored only if the suppress drive. }
+We then examine models in which an allele in sperm acts to modify the efficacy of a polymorphic meiotic driver - 
+	which universally reveal that sperm acting drive modifiers are favored only if the suppress drive. 
 The results highlight the fact that the interests of sperm and egg  genomes' are often aligned, as both are invested in the fate of the resultant zygote (as was speculated for the \emph{In} locus \citep{Pomiankowski1993}).
 Therefore, there is little selective benefit to females in preventing sperm to influence female meioses.
-\yb{Rather, features of o\"{o}genesis  may evolve to facilitate an interaction with sperm %with female meiosis
-	as yet another mechanism to enforce an equitable transfer of genetic material.}
+Rather, features of o\"{o}genesis  may evolve to facilitate an interaction with sperm %with female meiosis
+	as yet another mechanism to enforce an equitable transfer of genetic material.
 

### 2014-05-16 writeup_and_notes/male_dependent_drive.tex
YB modified and edited results

--- 2014-05-16 writeup_and_notes/male_dependent_drive.tex
@@
 Because only one of four products of female meiosis is transmitted to
-offspring, o\"{o}genesis is inherantly susceptible to such conflicts,
+offspring, o\"{o}genesis is inherently susceptible to such conflicts,
 as alleles can compete to be transmitted into the egg
 (meiotic drive). 
-Genomics conflicts drive the rapid evolution of surpression mechanisms, and
+Genomics conflicts drive the rapid evolution of suppression mechanisms, and
 many facets of gametogenesis have been interpreted as mechanisms of protection against genomic outlaws. 
@@
 	and evaluate whether a driver of this form could generate a sustained conflict favoring the evolution of suppressors. 
-\yb{We concluded with an analysis of a related mode (Model 3) - in which drivers influence their transmission in females via paternal genotype, rather than sperm haplotype.}
+\yb{We concluded with an analysis of a related model (Model 3) - in which drivers influence their transmission in females via paternal genotype, rather than sperm haplotype.}
 
@@
 In this case, the driver will be fixed if it reaches a frequency greater than some threshold, 
-	(approximated in Equation \ref{bistab_homozy_approx} and  presented exactly in 
-	Figure \ref{Bistab_homozyg_cost_fig}), and lost otherwise. 
+	approximated in Equation \ref{bistab_homozy_approx} and  presented exactly in 
+	Figure \ref{Bistab_homozyg_cost_fig}, and lost otherwise. 
 For most parameters, this threshold is likely too high to be reached by drift, 
@@
 \begin{equation}
-	s < d -1/2 +s_h(3-2d)/2
+	s \lessapprox  d -1/2 +s_h(3-2d)/2
 	\label{sForHetFix}
@@
 %\gc{Can we use something other than F, as that was HWE deviation}
-For the unlinked, model we introduce another locus where drive is modified in A/B females fertilized by $M$ allele, 
+For the unlinked model we introduce another locus where drive is modified in A/B females fertilized by $M$ allele, 
 	while the wild-type $L$ does not influence drive. 

### 2014-05-16 writeup_and_notes/male_dependent_drive.tex
small changes to intro

--- 2014-05-16 writeup_and_notes/male_dependent_drive.tex
@@
 the driving allele is transmitted to the egg in heterozygotes with
-probability  $d > 1/2$, regardless of sperm genotype (Model 1). To depict a case
+probability  $d > 1/2$, regardless of sperm genotype (Model 1) \gc{REF}. To depict a case
 of a self-promoting meiotic driver,  we modify this standard model such 

### 2014-05-17 writeup_and_notes/male_dependent_drive.tex
small edits throughout

@@
 
-\author{Yaniv Brandvain \\ email: ybrandvain@gmail.com  \and Graham Coop \\ email: gmcoop@ucdavis.edu }
+\author{Yaniv Brandvain$^1$ \\ email: ybrandvain@gmail.com  \and Graham Coop$^2$ \\ email: gmcoop@ucdavis.edu }
 
@@
 \begin{center}
-Center for Population Biology \& Department of Evolution and Ecology \\ University of California - Davis \\ Davis, CA, 95616
+$^1$Department of Plant Biology, \\ University of Minnesota – Twin
+Cities, MN, 55108 
+$^2$Center for Population Biology \& Department of Evolution and Ecology \\ University of California - Davis \\ Davis, CA, 95616
 \end{center}
@@
 the driving allele is transmitted to the egg in heterozygotes with
-probability  $d > 1/2$, regardless of sperm genotype (Model 1) \gc{REF}. To depict a case
+probability  $d > 1/2$, regardless of sperm genotype (Model 1) \gc{REF}.
+The timing of fertilization relative to female meiosis places another constraint on $d$, for example, if fertilization (and therefore, sperm dependent drive) takes place at MII (as in mammals),
+	female drive requires an uneven number of crossovers between the centromere and the drive locus, 
+	so $d$ is bounded to be $<0.83$ \citep[see ][ for
+        discussion]{Buckler1999}.  To depict a case
 of a self-promoting meiotic driver,  we modify this standard model such 
 	that the driver is only effective when fertilized by a sperm carrying that allele (see Figure \ref{Eggsperm_2_allele_cartoon} and Model 2 in the Supplementary Worksheet)
-\footnote{the timing of fertilization relative to female meiosis places another constraint on $d$, for example, if fertilization (and therefore, sperm dependent drive) takes place at MII (as in mammals),
-	female drive requires an uneven number of crossovers between the centromere and the drive locus, 
-	so $d$ is bounded to be $<0.75$? \citep[see ][ for
-        discussion]{Buckler1999}. }. %PUT THIS BACK IN
 We then identify the conditions allowing for the spread of this self-promoting driver, 
 	and evaluate whether a driver of this form could generate a sustained conflict favoring the evolution of suppressors. 
-\yb{We concluded with an analysis of a related model (Model 3) - in which drivers influence their transmission in females via paternal genotype, rather than sperm haplotype.}
+We concluded with an analysis of a related model (Model 3) - in which
+drivers influence their transmission in females via paternal genotype,
+rather than sperm haplotype.
 
@@
 The parameter space where the allele can invade but not fix is shown in white
-        in Figure \yb{\ref{Invasion_space}A}. %, and it can fix in the colored regions. 
+        in Figure \ref{Invasion_space}A. %, and it can fix in the colored regions. 
 When the allele is maintained as a polymorphism, it provides an opportunity for the evolution of
@@
 	no fitness cost when rare, 
-\yb{	a self-promoting driver specifically creates low fitness drive homozygotes 
-	by uniting driving female gametes with sperm enabling that drive.} 
+a self-promoting driver specifically creates low fitness drive homozygotes 
+	by uniting driving female gametes with sperm enabling that drive.
         % zygotes when it drives. in heterozygotes. 
@@
 	frequency exceeds some threshold,  (Equation \ref{thresh2}, exact results in 
-	\yb{FIGURES SXX}).
+	Figure \ref{bistable_additive}).
 %	$f_B^*$
@@
 	when the cost of the drive is recessive
-	(see Figure S\ref{Effect_of_dominance}, and Equation \ref{polymale}).  
+	(see Figure \ref{Effect_of_dominance}, and Equation \ref{polymale}).  
 %Presumably this zone of polymorphism comes about because positive frequency 
@@
 \caption{transmission probabilities for alleles through female
-  meiosis depend on sperm genotype. 2 allele model}  
+  meiosis depend on sperm genotype. 2 allele model. \gc{merge with 2,
+    rewrite both captions}}  
 \label{Eggsperm_2_allele_cartoon}
@@
        an enhancer of the drive, as recombination breaks up drive haplotypes \citep{Thomson1974,Charlesworth1978,Haig1991}. 
-Additionally, tight linkage between female driver and sperm modifier is consistent with the nature of well characterized drive systems which are often maintained as polymorphic inversions with numerous linked modifiers (CITE).
+Additionally, tight linkage between female driver and sperm modifier is consistent with the nature of well characterized drive systems which are often maintained as polymorphic inversions with numerous linked modifiers \gc{CITE}.
 
@@
 	three allele system that allows the sperm-acting drive facilitator $B^{+}$ to
-	invade the population \yb{(Mathematical Supplement and Figure XXX)}.
+	invade the population (eqn. \eqref{coupling}, Appendix Model 4, Mathematical Supplement).
 
@@
 %	that is sperm-acting alleles that act to restore  fairness of female meiosis, can. %If these sperm-acting drive suppressors arise on
-\yb{If a sperm-acting allele that restores fairness to female meiosis arises on
+If a sperm-acting allele that restores fairness to female meiosis arises on
 	the drive background (creating a third allele $B^{-}$, Model 4), it displaces the ancestral driver (Equation \ref{coupling}), 
@@
 	or is unlinked to the drive locus (Model 6), it readily invades a population segregating
-	for the drive system (see \yb{Figures XXX} and Equations \ref{A-} and \ref{unlinked}).  
+	for the drive system (Figure \ref{Eggsperm_3_allele_cartoon} and eqns \ref{A-} and \ref{unlinked}).  
 This allele lowers the frequency of the original drive system (perhaps to zero),
@@
 This result is consistent with previous work showing that drive suppressors unlinked to, or in repulsion phase with drivers usually invade polymorphic drive systems \citep[e.g. ][]{Brandvain2012}.  
-\yb{Therefore, all two-locus models of sperm influence on female drive suggest that
-	sperm will evolve to oppose female meiotic drive.}
+Therefore, all two-locus models of sperm influence on female drive suggest that
+	sperm will evolve to oppose female meiotic drive.
 
@@
 show that this approximation is reasonably correct for small
-parameter values; \yb{however, this approximation underestimates $f_B^*$ for large parameter values, presumably because they result in strong departures from HWE.}
+parameter values; however, this approximation underestimates $f_B^*$ for large parameter values, presumably because they result in strong departures from HWE.
  %range of other costs to the drive alleles other than a
@@
 
+\newpage
+
+  \setcounter{figure}{0}\global\def\thefigure{S\arabic{figure}}
+
 \section*{Supplementary Material}
@@
 
-\includegraphics[width = 0.5
-          \textwidth]{../Scripts/homozyg_bistability.eps}
+\includegraphics[width = \textwidth]{../Scripts/homozyg_bistability.eps}
 \caption{Invasion analysis for a self-promoting female meiotic drive allele with

### 2014-05-17 writeup_and_notes/male_dependent_drive.tex
small edits throughout appendix

@@
 	can succeed evolutionarily, even if they lower the 
-        fitness of the population, 
-	the mother whose offspring they will inhabit, 
-	or the organism in which they reside. 
+        fitness or the organism in which they reside. 
 This generates numerous conflicts  \cite{Burt2006} including sexual conflicts between mates \cite{Arnqvist2005}, 
@@
 of forming low fitness homozygotes. 
-In fact, an allele in sperm which causes the alternate allele to drive can be favored (FIG), a counterintuitive result 
+In fact, an allele in sperm which causes the alternate allele to drive can be favored, a counterintuitive result 
 	consistent with the tenuous evidence that the non-driving
@@
 In light of that it is interesting to consider whether the phylogenetic variation in the stage of oogenesis when fertilization takes place reflects not
-	only species' different reproductive ecologies, but also different selection pressures shaping the interaction of sperm alleles with female meioses. 
+	only species' different reproductive ecologies, but also different selection pressures shaping the interaction of sperm alleles with female meiosis. 
 %\gc{Add short paragraph about sex chromosomes here? XY systems don't care. ZW systems (males ZZ, females ZW). If there is
@@
 %The fitness of genotype, $g$ is $w_g$ independent of its sex, and its 
-All models include a biallelic locus with non-driving and driving alleles in frequencies $f_A$ and $f_B$, 
+All models include a biallelic locus (A/B) with non-driving and driving alleles in frequencies $f_A$ and $f_B$, 
 	respectively, while more complex models include more loci and/or alleles.  
@@
 over time. %, and to assess the analytical approximations, below.
- \gc{To assess invasion or fixation criteria, \yb{as well as bistability points,} we iterate this system and
-testing whether $f_B$ increases over a grid of parameters. }
+To assess invasion or fixation criteria, as well as bistability points, we iterate this system and
+testing whether $f_B$ increases over a grid of parameters.
 
@@
 In models 4 and 5, we treat this two loci system as if it consists of
-three alleles: $A$, $B$ and $C$, corresponding to the case \gc{when the
+three alleles: $A$, $B$ and $C$, corresponding to the case when the
 sperm-modifier arises once, tightly linked to one or other of the
-alleles (such that recombination is unexpected). }
+alleles (such that recombination is unexpected). 
 In all cases, fitness is independent of the drive modifier, so, $w_{AA}=1$, $w_{AB}=1-s_h$, and $w_{BB} = 1-s$. 
@@
 Assuming HWE and linkage equilibrium, the change in frequency of a rare unlinked, sperm-acting drive modifier is 
-\begin{equation} \Delta M =-\epsilon f_A f_B (f_B s + s_h (fA - fB) ) \frac{f_M}{\bar{w}}  \label{unlinked} \end{equation}
+\begin{equation} \Delta M =-\epsilon f_A f_B (f_B s + s_h (f_A - f_B) ) \frac{f_M}{\bar{w}}  \label{unlinked} \end{equation}
 % an $F$ allele increases in frequency when 

### 2014-05-17 writeup_and_notes/male_dependent_drive.tex
yb made a trivial change to see if the commit works
yb modified abstract

@@
 offspring, o\"{o}genesis is inherently susceptible to such conflicts,
-as alleles can compete to be transmitted into the egg
-(meiotic drive). 
+as alleles can compete to be transmitted into the egg. 
+%(meiotic drive).  
 Genomics conflicts drive the rapid evolution of suppression mechanisms, and
-many facets of gametogenesis have been interpreted as mechanisms of protection against genomic outlaws. 
+	many facets of  o\"{o}genesis and gametogenesis have been 
+	interpreted as mechanisms of protection against genomic outlaws. 
 Why then is female meiosis often left uncompleted until after fertilization - 
@@
                                 %%they will spread if no cost to organism
-These results are consistent with current knowledge of sperm-dependent meiotic drivers, 
+These results are consistent with current knowledge of sperm-dependent meiotic drivers 
+	\yb{(loci whose distortion of female meiosis depends on sperm genotype),} 
 	and suggest that the requirement of fertilization for the
         completion of female meiosis, and variation in stage of meiosis that
-        fertilization occurs at 
- may represent a mechanism to prevent meiotic drive.
+        fertilization occurs at may represent a mechanism employed by females 
+	 to ensure a fair meiosis. %prevent meiotic drive.
 \newpage
@@
 	the drive background (creating a third allele $B^{-}$, Model 4), it displaces the ancestral driver (Equation \ref{coupling}), 
-	and often returns to a lower equilibrium frequency than the the $B$ allele, further decreasing the extent of drive in the population. } 
+	and often returns to a lower equilibrium frequency than the the $B$ allele, further decreasing the extent of drive in the population.  
 If this sperm-acting drive suppressor arises on 

### 2014-05-17 writeup_and_notes/male_dependent_drive.tex
yb modified intro [near complete gc should look at the minor highlighted changes]

@@
 An allele that biases female meiosis in its favor (i.e.
-a meiotic driver), may increase in frequency even if this driver entails a pleiotropic fitness cost \citep{Prout1973}, generating a genetic conflict between the success of the driver and organismal fitness.
+a meiotic driver), may increase in frequency even if it entails a pleiotropic fitness cost \citep{Prout1973}, generating a genetic conflict between the success of the driver and organismal fitness.
 Meiotic drivers observed in both plants
@@
 \citep{Agulnik1990,Wu2005,Pardo-ManuelDeVillena2001c} highlight this
-conflict -- the selfish benefits of drive and the associated
-pleiotropic fitness costs sustain a balanced polymorphism
+conflict -- the selfish benefits and the associated
+pleiotropic fitness costs of drive sustain a balanced polymorphism
 \citep{Prout1973}, 
@@
 meiosis I \citep{Haig1991}, arrested female meiosis
-\citep{Mira1998}, centromere machinery \citep{Malik2002a,Malik2009,Malik2009}, and sex differences in the recombination rate \citep{Haig2010,Brandvain2012} 
+\citep{Mira1998}, the structure of centromere machinery \citep{Malik2002a,Malik2009,Malik2009}, and sex differences in the recombination rate \citep{Haig2010,Brandvain2012} 
 have perhaps evolved to enforce fairness by disrupting meiotic drive \citep{Rice2013}.
@@
 It is certainly not the case that animals are mechanistically incapable of circumventing this requirement.
-The stage at which female meiosis requires fertilization varies widely across taxa, and in a few animals, 
+The stage at which female meiosis requires fertilization varies widely across taxa, and in a few animals  
 	female meiosis is complete upon ovulation (that is, fertilization is not required for female meiosis \citep[e.g. Table 1 in 
@@
  transmission of alternate alleles in heterozygous females can depend on sperm haplotype.  
- While ruling out the alternative hypothesis of early selection on zygotes \citep[pages
-52-54 in][]{Burt2006} in these cases is challenging, it appears that the extent to which \emph{In} and \emph{Om} distort the second female meiotic division partially depends on the genotype of the fertilizing sperm \citep{Agulnik1993,Wu2005}.
+ While ruling out the alternative hypothesis of early selection on zygotes in these cases is challenging  \citep[see pages
+52-54 in][]{Burt2006}, it appears that the extent to which \emph{In} and \emph{Om} distort the second female meiotic division partially depends on the genotype of the fertilizing sperm \citep{Agulnik1993,Wu2005}.
 % Both of these female drive systems operate by distorting the second meiotic division, 
@@
 We first focus on models in which `self-promoting' alleles in
-  sperm facilitate drive of like \gc{alleles} during gametogenesis in heterozygous females. 
+  sperm facilitate drive of like alleles during gametogenesis in heterozygous females. 
 %In this article we explore through simple population genetic models the consequences of alleles that influence the outcome of female meiosis.
-These models show that sperm modifiers of female meiotic drive are
-	unlikely to create a sustained genomic conflict, and therefore 
-	female meiosis will have little incentive to evolve resistance to them.
-We then examine models in which an allele in sperm acts to modify the efficacy of a polymorphic meiotic driver - 
-	which universally reveal that sperm acting drive modifiers are favored only if the suppress drive. 
-The results highlight the fact that the interests of sperm and egg  genomes' are often aligned, as both are invested in the fate of the resultant zygote (as was speculated for the \emph{In} locus \citep{Pomiankowski1993}).
+These models show that such sperm-acting alleles 
+	\yb{have more difficulty invading a population than do traditional meiotic drivers, 
+	 and under most circumstances, cannot be maintained as a balanced polymorphism.
+Because self-promoting drivers are unlikely to create a sustained genomic conflict, 
+	female meiosis will have little incentive to evolve resistance to them.} 
+We then examine models in which a \yb{novel} sperm-acting allele modifies the efficacy of a polymorphic meiotic driver. 
+Such model universally reveal that sperm-acting drive modifiers are favored only if the suppress drive. 
+These results highlight the fact that the interests of sperm and \yb{maternal}  genomes' are often aligned, as both are invested in the fate of the resultant zygote (as was speculated for the \emph{In} locus \citep{Pomiankowski1993}).
 Therefore, there is little selective benefit to females in preventing sperm to influence female meioses.

### 2014-05-17 writeup_and_notes/male_dependent_drive.tex
yb read text through the end of the results

@@
 meiosis I \citep{Haig1991}, arrested female meiosis
-\citep{Mira1998}, the structure of centromere machinery \citep{Malik2002a,Malik2009,Malik2009}, and sex differences in the recombination rate \citep{Haig2010,Brandvain2012} 
+\citep{Mira1998}, the structure of centromere machinery \citep{Malik2002a,Malik2009}, and sex differences in the recombination rate \citep{Haig2010,Brandvain2012} 
 have perhaps evolved to enforce fairness by disrupting meiotic drive \citep{Rice2013}.
@@
 the driving allele is transmitted to the egg in heterozygotes with
-probability  $d > 1/2$, regardless of sperm genotype (Model 1) \gc{REF}.
+probability  $d > 1/2$, regardless of sperm genotype (Model 1) \cite[e.g.
+][]{Ubeda2004}.
 The timing of fertilization relative to female meiosis places another constraint on $d$, for example, if fertilization (and therefore, sperm dependent drive) takes place at MII (as in mammals),
@@
 of a self-promoting meiotic driver,  we modify this standard model such 
-	that the driver is only effective when fertilized by a sperm carrying that allele (see Figure \ref{Eggsperm_2_allele_cartoon} and Model 2 in the Supplementary Worksheet)
+	that the driver is only effective when fertilized by a sperm carrying that allele (see Figure \ref{Eggsperm_2_allele_cartoon} and Model 2 in the Supplementary Worksheet). 
 We then identify the conditions allowing for the spread of this self-promoting driver, 
@@
 	preclude its invasion when rare (Equation \ref{ineqA}). 
-In this case, the driver will be fixed if it reaches a frequency greater than some threshold, 
-	approximated in Equation \ref{bistab_homozy_approx} and  presented exactly in 
-	Figure \ref{Bistab_homozyg_cost_fig}, and lost otherwise. 
+In this case, the driver will be fixed if it reaches a frequency greater than some threshold 
+	(approximated in Equation \ref{bistab_homozy_approx} and  presented exactly in 
+	Figure \ref{Bistab_homozyg_cost_fig}) and lost otherwise. 
 For most parameters, this threshold is likely too high to be reached by drift, 
@@
 	when $s$ is sufficiently small (Equation \ref{sForHetFix}) this allele fixes deterministically if its 
-	frequency exceeds some threshold,  (Equation \ref{thresh2}, exact results in 
+	frequency exceeds some threshold  (Equation \ref{thresh2}, exact results in 
 	Figure \ref{bistable_additive}).
@@
 	limited gene expression in sperm \citep[e.g.][]{Joseph2004}
-	suggest a model where sperm influence female meiosis via expression of the fertilizing male's
+	suggests a model where sperm influence female meiosis via expression of the fertilizing male's
 	diploid genotype (perhaps due to proteins and RNAs packaged into the sperm), rather than sperm haplotype.
@@
 
-Models 2 and 3, above, explored the dynamics of an allele drove in females when signaled by a complementary signal in sperm.   
+Models 2 and 3, above, explored the dynamics of an allele that drove in females when signaled by a complementary signal in sperm.   
 We complement this single-locus approach %of an allele with pleiotropic effects in both sexes, 
-	with alternative models of two linked loci - one a female driver, 
+	with alternative models of two loci - one a female driver, 
 	and the other, a sperm allele which modifies the effect of drive upon fertilization. 
@@
 	$A$ and $B$ are the ancestral non-driver and driver alleles). 
-Subsequently, a sperm-acting modifier of female meiotic drive arises tightly
-        linked to this balanced locus (effectively creating a third
-        allele/haplotype at this locus). 
+Subsequently, a sperm-acting modifier of female meiotic drive arises at another lous.
+
+We first assume that the modifier is tightly
+        linked to the drive locus (effectively creating a third
+        allele/haplotype at this locus) and arises on the drive-background. 
+%We complement this single-locus approach %of an allele with pleiotropic effects in both sexes, 
+%	with alternative models of two linked loci - one a female driver, 
+%	and the other, a sperm allele which modifies the effect of drive upon fertilization. 
+%In this model, a female meiotic driver with no sperm-dependency 
+%	initially reaches drive-viability equilibrium (with two alleles
+%	$A$ and $B$ are the ancestral non-driver and driver alleles). 
+%Subsequently, a sperm-acting modifier of female meiotic drive arises tightly
+ %      linked to this balanced locus (effectively creating a third
+ %       allele/haplotype at this locus). 
 Tight linkage offers the best
@@
        an enhancer of the drive, as recombination breaks up drive haplotypes \citep{Thomson1974,Charlesworth1978,Haig1991}. 
-Additionally, tight linkage between female driver and sperm modifier is consistent with the nature of well characterized drive systems which are often maintained as polymorphic inversions with numerous linked modifiers \gc{CITE}.
+Additionally, tight linkage between female driver and sperm modifier is consistent with the nature of well characterized drive systems which are often maintained as polymorphic inversions with numerous linked modifiers \cite{Burt2006}.
+\yb{We conclude by analyzing models with alternative linkage relationship between driver and drive modifier - 
+	in Model 5 the modifier arises tightly linked to the non-driving allele, 
+	and in Model 6 it is unlinked to the driver.}
 
@@
 %%%sperm-acting facilitator
-We first consider a sperm-acting modifier of
-	female meiotic drive tightly linked to the $B$ background (Model 4).
-We treat this non-recombining drive/modifier haplotype as a third allele, and we label this allele as $B^+$ when it enhances drive. 
+When the modifier of drive is tightly linked to the driver and enhances drive, 
+	we label  this non-recombining drive/modifier haplotype as the $B^+$ allele.  
+%We first consider a sperm-acting modifier of
+%	female meiotic drive tightly linked to the $B$ background (Model 4).
+%We treat this non-recombining drive/modifier haplotype as a third allele, and we label this allele as $B^+$ when it enhances drive. 
 %(i.e. the non-recombining haplotype of a
@@
 The $B^+$ allele acts in sperm to increase the effectiveness of drive for both
-	the $B$ and  $B^{+}$ alleles in $AB$ and $AB^{+}$ heterozygotes (see Figure \ref{Eggsperm_3_allele_cartoon}). 
+	the $B$ and  $B^{+}$ alleles in $AB$ and $AB^{+}$ heterozygotes (see Figure \ref{Eggsperm_3_allele_cartoon}, and Model 4 in the Supplementary Worksheet). 
 %	for a simple schematic of this drive model.  
@@
 %	$BB^{+}$ heterozygotes that suffer the same fitness costs as $BB$. 
-Therefore, a sperm-acting drive facilitator alleles experience a profound disadvantage 
+Therefore, sperm-acting drive facilitator alleles experience a profound disadvantage 
 	in this scenario, even more so than under the previous two allele model. 
@@
 	three allele system that allows the sperm-acting drive facilitator $B^{+}$ to
-	invade the population (eqn. \eqref{coupling}, Appendix Model 4, Mathematical Supplement).
+	invade the population (Appendix Model 4, eqn. \eqref{coupling}, and Mathematical Supplement).
 
@@
 	for the drive system (Figure \ref{Eggsperm_3_allele_cartoon} and eqns \ref{A-} and \ref{unlinked}).  
-This allele lowers the frequency of the original drive system (perhaps to zero),
+This allele lowers the frequency of the original driver (perhaps to zero),
 	and spreads to fixation if it does not carry strong fitness costs

### 2014-05-17 writeup_and_notes/male_dependent_drive.tex
yb read text through the end of the discussion

@@
 of forming low fitness homozygotes. 
-In fact, an allele in sperm which causes the alternate allele to drive can be favored, a counterintuitive result 
-	consistent with the tenuous evidence that the non-driving
-        $ln^-$ allele is over transmitted through female meiosis in $ln^-/ln^+$ heterozygote when fertilized by $ln^+$
-        sperm \cite{agulnick, Pomiankowski1993}.   
+%In fact, an allele in sperm which causes the alternate allele to drive can be favored, a counterintuitive result 
+%	consistent with the tenuous evidence that the non-driving
+%        $ln^-$ allele is over transmitted through female meiosis in $ln^-/ln^+$ heterozygote when fertilized by $ln^+$
+%        sperm \cite{agulnick, Pomiankowski1993}.   
+% YB - WE DID NOT TALK ABOUT THIS IN THE RESULTS [ALTHOUGH IT'S TRUE... PERHAPS TOO DISTRACTING]
 More generally, alleles that act through sperm to

### 2014-05-17 writeup_and_notes/male_dependent_drive.tex
yb voiced concerns about en of discussion [in text.. we should resolve and fix]

@@
 	only species' different reproductive ecologies, but also different selection pressures shaping the interaction of sperm alleles with female meiosis. 
+\yb{this feels weak if we don't say anything more explicit. I also think maybe we simply state predictions:
+	(1) MII drivers should drive less often when fertilized by complimentary sperm. 
+	(2) Drive may be more severs in interspecific crosses in which sperm have not specifically evolved to resists  a driver. 
+	(3) ???
+Any more specific hypotheses -- maybe making this last sentence more explicit, e.g.
+self-fertilizing or otherwise frequently inbreeding animals will not require fertilization for the completion of meiosis...
+[note - this runs counter to our empirical examples of mice and C-elegens]}
 %\gc{Add short paragraph about sex chromosomes here? XY systems don't care. ZW systems (males ZZ, females ZW). If there is

### 2014-05-17 writeup_and_notes/male_dependent_drive.tex
yb editted through appendix

@@
 the driving allele is transmitted to the egg in heterozygotes with
-probability  $d > 1/2$, regardless of sperm genotype (Model 1) \cite[e.g.
-][]{Ubeda2004}.
-The timing of fertilization relative to female meiosis places another constraint on $d$, for example, if fertilization (and therefore, sperm dependent drive) takes place at MII (as in mammals),
-	female drive requires an uneven number of crossovers between the centromere and the drive locus, 
-	so $d$ is bounded to be $<0.83$ \citep[see ][ for
-        discussion]{Buckler1999}.  To depict a case
+probability  $d > 1/2$, regardless of sperm genotype \cite[e.g.
+][ and see Model 1 in the Appendix for more details]{Ubeda2004}.
+%The timing of fertilization relative to female meiosis places another constraint on $d$, for example, if fertilization (and therefore, sperm dependent drive) takes place at MII (as in mammals),
+%	female drive requires an uneven number of crossovers between the centromere and the drive locus, 
+%	so $d$ is bounded to be $<0.83$ \citep[see ][ for discussion]{Buckler1999}.
+	%YB - MOVED THIS TO THE APPENDIX
+          To depict a case
 of a self-promoting meiotic driver,  we modify this standard model such 
@@
 \section*{Appendix}
-Our appendix highlights some of our derivations and analytical results from: 
-	the standard female meiotic drive (Model 1); a sperm
-	dependent female meiotic drive (Model 2); a male genotype dependent
-female meiotic drive (Model 3); and more complex two locus sperm
-dependent drive (Models 4-6). %changed this to minimize confusion with refs, figs, equations, etc.
+We highlight simple analytical results from: 
+	the standard model of female meiotic drive (Model 1); a sperm-dependent 
+	model of female meiotic drive (Model 2); a male-genotype-dependent model of
+	female meiotic drive (Model 3); and more complex two-locus sperm-dependent drive (Models 4-6). %changed this to minimize confusion with refs, figs, equations, etc.
 
@@
 %The fitness of genotype, $g$ is $w_g$ independent of its sex, and its 
-All models include a biallelic locus (A/B) with non-driving and driving alleles in frequencies $f_A$ and $f_B$, 
+All models include a biallelic locus ($A/B$) with non-driving and driving alleles in frequencies $f_A$ and $f_B$, 
 	respectively, while more complex models include more loci and/or alleles.  
 Transmission rules describing the outcomes of all matings in each
-	models are presented in a \yb{Supplementary Worksheet}. 
+	models are presented in a Supplementary Worksheet. 
 The fitness of genotype, $g$ is sex-independent and equals $1$, $1-h_s$, and $1-s$ for genotypes $AA$, $AB$, and $BB$, respectively. %  and  $w_{AA} = 1$, $w_{AB} = 1-s_h$, and $w_{BB}=1-s$. 
@@
 %\gc{Mathematical Supplement}.  
-Because many of our analyses assume Hardy Weinberg Equilibrium (HWE), we note which analytical results are approximations. 
+Because many of our analyses assume genotype frequencies expected under Hardy Weinberg Equilibrium (HWE), we note which analytical results are approximations. 
 We verify these approximations with exact numerical iterations in Supplementary Figures and Supplementary Scripts, and detail the derivation of these approximations in our Mathematical Supplement.
@@
 In the standard female drive model, meiosis in males is fair such that $A/B$ heterozygotes contribute $A$ and $B$ alleles with equal probabilities; however, $A/B$ females transmit the $B$ allele with probability $d>1/2$.
+We note that the timing of fertilization relative to female meiosis places another constraint on $d$, for example, if fertilization (and therefore, sperm dependent drive) takes place at MII (as in mammals),
+	female drive requires an uneven number of crossovers between the centromere and the drive locus, 
+	so $d$ is bounded to be $<0.83$ \citep[see ][ for discussion]{Buckler1999}. 
 After drive and random mating, genotype frequencies are 
@@
         $s\lessapprox(2d-1)/2$ when the cost of drive is fully
-        recessive. \gc{isn't this last expression exact?}
+        recessive. 
+        %\gc{isn't this last expression exact?} YB: exact-ish.. all assume HWE.. but i think we don't expect a deviation from HWE here??  it's prob easiest to kept them all as approx?
 
 \subsubsection*{Model 2. Single locus, sperm-dependent drive.}
-Our single-locus model of sperm dependent drive resembles the traditional driver with the caveat that the $B$ allele drives in heterozygous females only when fertilized by $B$ bearing sperm. 
+Our single-locus model of sperm-dependent drive resembles the traditional driver, with the caveat that the $B$ allele drives in heterozygous females only when fertilized by $B$-bearing sperm. 
 Therefore, genotype frequencies after drive are 
@@
 We iterate exact genotype frequency recursions ($f_g''= f_g'w_g/\bar{w}$) over generations to produce the frequency trajectories shown in the inset
-Figure \ref{Invasion_space} by plotting \gc{$f_B=f_{BB}+ \frac{1}{2}f_{AB}$}
+Figure \ref{Invasion_space} by plotting $f_B=f_{BB}+ \frac{1}{2}f_{AB}$
 over time. %, and to assess the analytical approximations, below.
 To assess invasion or fixation criteria, as well as bistability points, we iterate this system and
-testing whether $f_B$ increases over a grid of parameters.
+test whether $f_B$ increases over a grid of parameters.
 
@@
 \end{equation}
- but it will be lost if it is introduced below this frequency. 
+ but it will be lost when introduced below this frequency. 
  %(i.e. this is an unstable equilibrium point). 
@@
 Compared to exact results (Supplementary Figure \ref{Bistab_homozyg_cost_fig}), Equations
- \gc{\eqref{ineqA} and} \eqref{bistab_homozy_approx} offer reasonable, but imperfect approximations. 
+ \eqref{ineqA} and \eqref{bistab_homozy_approx} offer reasonable, but imperfect approximations. 
 %s_h and seems to serve as a lower bound to the true unstable equilibrium frequency.
@@
 \paragraph{Cost of driver in heterozygotes:} %We explored a more complex model where the self-promoting driver has a cost, $s_h$ in heterozygotes (so $w_{AB} = 1-s_h$) in addition to the cost $s$, in homozygotes.
-When the fitness of drive heterozygotes is compromised ($s_h>0$), a self promoting driver cannot invade when rare.
-This results from the fact that
-when rare %($f_B \ll 1$) 
-$B$-bearing sperm rarely
-encounter heterozygous eggs ($\sim f_B^2$) but the allele still pays a
+When the fitness of drive heterozygotes is compromised ($s_h>0$), a self-promoting driver cannot invade when rare.
+This results from the fact that,
+when rare, %($f_B \ll 1$) 
+$B$-bearing sperm and heterozygous eggs will rarely encounter one another ($\sim f_B^2$) but the allele still pays a
 cost in heterozygous individuals ($\sim f_B$). 
@@
 \subsubsection*{Model 3. Single locus, paternal genotype dependent drive:}
-We now turn to the case when female meiotic drive depends on paternal genotype.
+In the case when female meiotic drive depends on paternal genotype
 %\paragraph{Dominance} we also explored a model where
@@
 %instead of just the allelic type of the sperm.
- We assume a heterozygous female will transmit the $B$ allele 
+ %We assume a 
+ heterozygous female will transmit the $B$ allele 
   with probabilities  $\frac{1}{2}$,  $d_{het}\geq 1/2 $ or $d_{hom}\geq d_{het}$, 
@@
  when mated with to $AA$, $AB$, or $BB$ males,  respectively. 
-  In this model, genotype frequencies after drive, mating and selection are 
+  In this model, genotype frequencies after drive, random mating, and selection are 
   	$f_{AA}'' =   w_{AA}\left( f_{AA} [2 f_A + f_{AB} ] + f_{AB}^2 [1 - d_{het}] \right)/(2\bar{w})$,
@@
 If the cost of drive is fully recessive (i.e. $s_h=0$), assuming HWE, 
-	a rare paternal genotype dependent driver invades when 
-	$s\lessapprox (d_{het}-1/2)/(d_{het})$, and a common paternal genotype dependent driver fixes when 
+	a rare paternal-genotype-dependent driver invades when 
+	$s\lessapprox (d_{het}-1/2)/(d_{het})$, and when common this driver fixes if 
 	$s\lessapprox d_{hom}-1/2$, approximations well supported by exact results (Supplementary Figure \ref{Effect_of_dominance}).
@@
 	and therefore some values of $s$ can maintain a stable polymorphism. 
-Under these parameter values, a rare paternal genotype dependent driver 
-	can increase in frequency as drivers gain a transmission advantage and suffer 
+Under these parameter values, a rare paternal-genotype-dependent driver 
+	can increase in frequency because it gains a transmission advantage and suffers 
 	no fitness cost when heterozygous eggs are fertilized by 
@@
 As the frequency of the $B$ allele increases, 
-	occur more often as a homozygote, and so will be unable to 
-	avoid producing unfit homozygote zygotes, leaving it trapped in
+	%it occurs more often as a homozygote, and so 
+	it will be unable to 
+	avoid producing unfit homozygous offspring, leaving it trapped in
 	the population at frequency $f_{B\text{ pat}}^*$. 
@@
 \subsection*{Models 4-6. Two-locus, sperm-dependent drive.} 
-Assuming that a traditional driver segregates at its equilibrium frequency (Model 1), we investigate the evolution of tightly linked (Models 4 and 5), and unlinked sperm-dependent modifiers of drive. 
-In models 4 and 5, we treat this two loci system as if it consists of
-three alleles: $A$, $B$ and $C$, corresponding to the case when the
+Assuming that a traditional driver segregates at its equilibrium frequency (Model 1), we investigate the evolution of tightly linked (Models 4 and 5), and unlinked (Model 6) sperm-dependent modifiers of drive.
+In Models 4 and 5, we treat this two-lous system as if it consists of
+a single locus with three alleles: $A$, $B$ and $C$, corresponding to the case when the
 sperm-modifier arises once, tightly linked to one or other of the
-alleles (such that recombination is unexpected). 
-In all cases, fitness is independent of the drive modifier, so, $w_{AA}=1$, $w_{AB}=1-s_h$, and $w_{BB} = 1-s$. 
-We assume the traditional driver is transmitted to $d_0$ of gametes from female heterozygotes  when fertilized by wild-type sperm, and $d_1=d_0+\epsilon$ when fertilized by a sperm acting drive modifier.
+alleles at the drive locus (such that recombination is unexpected). 
+In all cases, we assume that fitness is independent of the drive modifier. %, so, $w_{AA}=1$, $w_{AB}=1-s_h$, and $w_{BB} = 1-s$. 
+We assume the traditional driver is transmitted to $d_0$ of gametes from female heterozygotes  when fertilized by wild-type sperm, and $d_1=d_0+\epsilon$ when fertilized by a sperm-acting drive-modifier.
 
@@
 	and a decrease in frequency of the $C$ allele when it suppresses drive ($\epsilon <0$ -- the $B^-$ model, above). 
-More generally even when the cost of drive is not fully recessive, the $B^-$ allele will invade and fix under all parameters sustaining a previous polymorphism at the drive locus (see Mathematical Supplement).
+More generally, even when the cost of drive is not fully recessive, the $B^-$ allele will invade and fix under all parameters sustaining a previous polymorphism at the drive locus (see Mathematical Supplement).
 %When the C allele acts to suppress drive for all parameters sustaining a polymorphism at the drive locus ($s>d_0-1/2$)
@@
 When the $C$ allele is tightly linked to the non-driver, 
-	genotypic fitnesses equal $w_{CC}=w_{AC}=w_{AA}=1$, and $w_{BC}=1-s_h$. 
+	genotypic fitnesses equal $w_{CC}=w_{AC}=w_{AA}=1$, and $w_{BC}=w_{AB}=1-s_h$. 
 Assuming HWE, a recessive fitness cost to drive, and assuming that the $A/B$ locus is at its equilibrium frequency, the change in frequency of a rare drive modifier is
@@
 \end{equation}
-For all values of interest ($0<s<1$, $0.5<d_0<1$), the change in frequency a rare $C$ is positive when it decreases drive (i.e.
-$\epsilon <0$, corresponding to the $A^-$ model, above), a result which hold qualitatively for a common $C$ allele, as well (Mathematical Supplement).
+For all values of interest ($0<s<1$, $0.5<d_0<1$), the change in frequency a rare $C$ allele is positive when it decreases drive (i.e.
+$\epsilon <0$, corresponding to the $A^-$ model, above), a result which holds qualitatively for a common $C$ allele, as well (Mathematical Supplement).
 %We address two questions concerning the $C$ allele - can it invade when rare  and can it fix when common?
@@
 %\gc{Can we use something other than F, as that was HWE deviation}
-For the unlinked model we introduce another locus where drive is modified in A/B females fertilized by $M$ allele, 
-	while the wild-type $L$ does not influence drive. 
+For the unlinked model, we introduce another locus where drive is modified in $A/B$ females fertilized by $M$ allele, 
+	while the wild-type $L$ allele does not influence drive. 
 Assuming HWE and linkage equilibrium, the change in frequency of a rare unlinked, sperm-acting drive modifier is 
@@
 Thus, a rare drive suppressor ($\epsilon<0$) will spread so long as the fitness cost of the driver does not display over- or under-dominance. 
-Notably, this result holds while assuming complete linkage equilibrium, and therefore sperm suppression of female meiotic drive is likely more powerful than female enforced systems which usually spread due to the indirect benefit of avoiding drive haplotypes \citep[e.g.
+Notably, this result holds while assuming complete linkage equilibrium, and therefore sperm suppression of female meiotic drive is likely more powerful than more direct female-enforced drive suppressors  which usually spread due to the indirect benefit of avoiding drive haplotypes \citep[e.g.
 ][]{Brandvain2012}.

### 2014-05-18 writeup_and_notes/male_dependent_drive.tex
merged figures 1 and 2 to form selfprom.eps. modified legend and in text refs to this fig

@@
 of a self-promoting meiotic driver,  we modify this standard model such 
-	that the driver is only effective when fertilized by a sperm carrying that allele (see Figure \ref{Eggsperm_2_allele_cartoon} and Model 2 in the Supplementary Worksheet). 
+	that the driver is only effective when fertilized by a sperm carrying that allele (see Figure \ref{Eggsperm}A and Model 2 in the Supplementary Worksheet). 
 We then identify the conditions allowing for the spread of this self-promoting driver, 
@@
 The parameter space where the allele can invade but not fix is shown in white
-        in Figure \ref{Invasion_space}A. %, and it can fix in the colored regions. 
+        in Figure \ref{Eggsperm}B. %, and it can fix in the colored regions. 
 When the allele is maintained as a polymorphism, it provides an opportunity for the evolution of
@@
 This analytical approximation, derived from Equation \eqref{deltadriver} assuming Hardy-Weinberg, 
-	closely matches results obtained by exact numerical iteration (Figure \ref{Invasion_space}).
+	closely matches results obtained by exact numerical iteration (Figure \ref{Eggsperm}B).
 
@@
 However, once relatively common, it  rapidly achieves fixation due to its
-	positive frequency dependent behavior (Inset in Figure \ref{Invasion_space}).  
+	positive frequency dependent behavior (Figure \ref{Eggsperm}B.1).  
 %This lack of a balanced polymorphism %(Figure \ref{Invasion_space}) 
@@
 \begin{figure}
-	\rotatebox{90}{
-\includegraphics[width = 0.5
-          \textwidth]{Figures/Sperm_egg_cartoon_paper_Fig1.eps}}
-\caption{transmission probabilities for alleles through female
-  meiosis depend on sperm genotype. 2 allele model. \gc{merge with 2,
-    rewrite both captions}}  
-\label{Eggsperm_2_allele_cartoon}
+\includegraphics[width = 0.8 \textwidth]{Figures/selfprom.eps}
+\caption{{\bf{A.}} A visual depiction of our model of 'self-promoting' driver. 
+	Transmission probabilities for alleles through female
+ 	 meiosis depend on sperm genotype. 
+	 The non-driving $A$-allele, and self-promoting $B$-allele
+	 are represented by unfilled and filled circles, respectively. 
+	 {\bf{B.}} Evolution of a self-promoter and standard driver. 
+	 	Assuming that the fitnesses of drive homozygotes and 
+		heterozygotes are $1-s$ and $1$, respectively. 
+	 	\emph{Main figure:} Boundary conditions for the invasion 
+		and fixation of self-promoting and standard meiotic drivers, 
+		with drive coefficient, $d$. Colored regions depict exact results, 
+		while lines  represent analytical approximations. 
+	 \emph{B1:}  Trajectories of sperm-dependent female drive
+		 %alleles that can invade the population, 
+		 each allele has $s =0.1$ against the
+		 homozygotes.  The drive coefficient is denoted by color. 
+ }   
+\label{Eggsperm}
 \end{figure}
+
+%\begin{figure}
+%	\rotatebox{90}{
+%\includegraphics[width = 0.5
+   %       \textwidth]{Figures/Sperm_egg_cartoon_paper_Fig1.eps}}
+%\caption{transmission probabilities for alleles through female
+%  meiosis depend on sperm genotype. 2 allele model. \gc{merge with 2,
+%    rewrite both captions}}  
+%\label{Eggsperm_2_allele_cartoon}
+%\end{figure}
 
@@
 
-\begin{figure}
-\includegraphics[width = 0.8 \textwidth]{Figures/invasion_space_recessive_driver.eps}
-\caption{{\bf A.} Invasion analysis for a female meiotic drive allele with
-  recessive costs (selection coefficient $s$). First consider a simple female meiotic
-  drive system, where the allele is transmitted a fraction $d>1/2$ of
-  meioses in female heterozygotes, inspective
-  of the allele carried by the fertilizing sperm. 
-Such an allele can invade the population over all of the parameter space, and will be maintained as a
-  protected polymorphism in the white region, and can fix in the
-  population in all of the parameter space colored green or orange. 
-The solid back line marks the criteria for the allele to fix $s< (2d-1)/2$.
-Now consider an allele that is fairly transmitted through
-  female meiosis when fertilized by a sperm carrying a different
-  allele, but at a fraction $d>1/2$ when a heterozygote egg is fertilized
-  by a sperm carrying the allele (as depicted in Figure
-  \ref{Eggsperm_2_allele_cartoon}). Such an allele can only invade in
-  the region of parameter space colored orange, and it fixes over all
-  of this parameter region. The approximate criteria for this allele
-  to fix is given by the dashed line showing $s < (2d- 1)/(4d)$. 
- {\bf B.} In the inset we show trajectories of sperm-dependent female drive
- alleles that can invade the population, each allele has $s =0.1$ against the
- homozygotes and $d =0.6$, $0.7$, $0.8$ shown in yellow, blue, and red. 
- \gc{Likely make this Figure 1B.} } \label{Invasion_space}
-\end{figure}
+%\begin{figure}
+%\includegraphics[width = 0.8 \textwidth]{Figures/invasion_space_recessive_driver.eps}
+%\caption{{\bf A.} Invasion analysis for a female meiotic drive allele with
+%  recessive costs (selection coefficient $s$). First consider a simple female meiotic
+%  drive system, where the allele is transmitted a fraction $d>1/2$ of
+  %meioses in female heterozygotes, inspective
+%  of the allele carried by the fertilizing sperm. 
+%Such an allele can invade the population over all of the parameter space, and will be maintained as a
+%  protected polymorphism in the white region, and can fix in the
+%  population in all of the parameter space colored green or orange. 
+%The solid back line marks the criteria for the allele to fix $s< (2d-1)/2$.
+%Now consider an allele that is fairly transmitted through
+%  female meiosis when fertilized by a sperm carrying a different
+%  allele, but at a fraction $d>1/2$ when a heterozygote egg is fertilized
+%  by a sperm carrying the allele (as depicted in Figure
+%  \ref{Eggsperm_2_allele_cartoon}). Such an allele can only invade in
+%  the region of parameter space colored orange, and it fixes over all
+ % of this parameter region. The approximate criteria for this allele
+%  to fix is given by the dashed line showing $s < (2d- 1)/(4d)$. 
+% {\bf B.} In the inset we show trajectories of sperm-dependent female drive
+ %alleles that can invade the population, each allele has $s =0.1$ against the
+% homozygotes and $d =0.6$, $0.7$, $0.8$ shown in yellow, blue, and red. 
+% \gc{Likely make this Figure 1B.} } \label{Invasion_space}
+%\end{figure}
 

### 2014-05-18 writeup_and_notes/male_dependent_drive.tex
removing ybs

--- 2014-05-17 writeup_and_notes/male_dependent_drive.tex
@@
 These results are consistent with current knowledge of sperm-dependent meiotic drivers 
-	\yb{(loci whose distortion of female meiosis depends on sperm genotype),} 
+(loci whose distortion of female meiosis depends on sperm genotype),}
 	and suggest that the requirement of fertilization for the
@@
 These models show that such sperm-acting alleles 
-	\yb{have more difficulty invading a population than do traditional meiotic drivers, 
+	have more difficulty invading a population than do traditional meiotic drivers, 
 	 and under most circumstances, cannot be maintained as a balanced polymorphism.
 Because self-promoting drivers are unlikely to create a sustained genomic conflict, 
-	female meiosis will have little incentive to evolve resistance to them.} 
-We then examine models in which a \yb{novel} sperm-acting allele modifies the efficacy of a polymorphic meiotic driver. 
+	female meiosis will have little incentive to evolve resistance to them.
+We then examine models in which a novel sperm-acting allele modifies the efficacy of a polymorphic meiotic driver. 
 Such model universally reveal that sperm-acting drive modifiers are favored only if the suppress drive. 
-These results highlight the fact that the interests of sperm and \yb{maternal}  genomes' are often aligned, as both are invested in the fate of the resultant zygote (as was speculated for the \emph{In} locus \citep{Pomiankowski1993}).
+These results highlight the fact that the interests of sperm and maternal  genomes' are often aligned, as both are invested in the fate of the resultant zygote (as was speculated for the \emph{In} locus \citep{Pomiankowski1993}).
 Therefore, there is little selective benefit to females in preventing sperm to influence female meioses.
@@
 Additionally, tight linkage between female driver and sperm modifier is consistent with the nature of well characterized drive systems which are often maintained as polymorphic inversions with numerous linked modifiers \cite{Burt2006}.
-\yb{We conclude by analyzing models with alternative linkage relationship between driver and drive modifier - 
+We conclude by analyzing models with alternative linkage relationship between driver and drive modifier - 
 	in Model 5 the modifier arises tightly linked to the non-driving allele, 
-	and in Model 6 it is unlinked to the driver.}
+	and in Model 6 it is unlinked to the driver.
 
@@
 	only species' different reproductive ecologies, but also different selection pressures shaping the interaction of sperm alleles with female meiosis. 
+We predict that female drive should be
+
 \yb{this feels weak if we don't say anything more explicit. I also think maybe we simply state predictions:
@@
 [note - this runs counter to our empirical examples of mice and C-elegens]}
-%\gc{Add short paragraph about sex chromosomes here? XY systems don't care. ZW systems (males ZZ, females ZW). If there is
-%sex ratio distortion how could this play out?}
 
-%\yb{DOES THIS GO ANYWHERE? }
-
-% Males can potentially influence the outcome of meiosis even before fertilization. 
-% For example in Drosophila and other taxa, sperm accessory
-%         gland proteins (ACPs) increase the female recombination rate
-%         \cite{SOMEthING}, and there is some evidence that there is
-%         variance across male genotypes in there effect on female
-%         recombination rate \cite{Stevison2012}.
-% Presumably this is a pleiotropic consequence of the resolution of double strand breaks initiated by the stressful/toxic composition of ACPs. 
-% However, a pleiotropic effect of this is to increase the recombination rate which can decrease the effectiveness of female meiotic drivers
-% active during MI \citep{Brandvain}. 
-% We speculate that other effects of sperm on female meiosis may also be tolerated by females as a way of ensuring meiotic fairness -- 
-% 	that is females evolve to allow male influence on female meiosis in cases in which the male influence is  egalitarian.
-
-%Gametogenesis and meiosis are common times for genetic conflict. 
-%This is a consequence of the fact that as the set of alleles that consitute
-%	an individual are in the process of parting ways, 
-%	with only a subset of them being assured transmission.
-%In constrast, the perspective of the fertilizing sperm's genome is different, it has to live with the consequences of any distortion it causes 
-%to female meiosis through the entire next life cycle of the organism. 
-%It is therefore in the interest of the fertilizing sperm's genome to
-%make a high fitness zygote even if that means alleles in sperm
-%disrupting other copies of themselves from gaining a 
-%transmission advantage through female meiosis.
   \appendix

### 2014-05-18 scripts/invasion_analysis.R

@@ -410,7 +410,7 @@ for(i in c(1,3,4,5)){
 	lines(d.range,fixation.w.dom[i,],lty=2,col=my.cols[i])
 	if(i==3){ 
 		legend("bottomright",lty=c(1,2),legend=c("invasion line","fixation line")); 
-		mtext(side=3,line=0,"h=0.5")))
+		mtext(side=3,line=0,"h=0.5")
 
 		#text(0.6,0.45,expression(paste(d[h],"=0.5d")))
 
@@ -437,7 +437,7 @@ for(i in c(1,3,4,5)){
 	}
 }
 lines(d.range, (2*d.range-1)/(2*d.range),col=my.cols[i],lty=3)
-legend("bottomright",lty=3,legend=c("HWE fixation line"),col=my.cols[i]); 
+legend("bottomright",lty=3,legend=c("HWE invasion line"),col=my.cols[i]); 
 dev.copy2eps(file=paste(directory,"effect_of_dominance_on_invasion_space_four_graph.eps",sep=""))
 
 

### 2014-05-18 writeup_and_notes/male_dependent_drive.tex
added hypotheses to discussion

@@
  reduce the opportunity of female meiotic drive regardless of linkage or phase. 
+We predict that female drive, at a stage influencable by meiotic
+drive,  should be reduced when fertilized by
+complimentary sperm. For example, MII female drive should be more severe in
+interspecific crosses as sperm will not specifically evolved to
+resists a female driver.
 
@@
 In light of that it is interesting to consider whether the phylogenetic variation in the stage of oogenesis when fertilization takes place reflects not
-	only species' different reproductive ecologies, but also different selection pressures shaping the interaction of sperm alleles with female meiosis. 
-We predict that female drive should be
+	only species' different reproductive ecologies, but also
+        different selection pressures shaping the interaction of sperm
+        alleles with female meiosis. 
+For example it would be interesting to know whether changes that are predicted to lower the opportunity for female meiotic drive,
+        e.g. an evolutionary transition to inbreeding or a high
+        selfing rate in animals, are often accompanied by a relaxation of the requirement of fertilization for the completion of meiosis.
 
-\yb{this feels weak if we don't say anything more explicit. I also think maybe we simply state predictions:
-	(1) MII drivers should drive less often when fertilized by complimentary sperm. 
-	(2) Drive may be more severs in interspecific crosses in which sperm have not specifically evolved to resists a driver. 
-	(3) ???
+% \yb{this feels weak if we don't say anything more explicit. I also think maybe we simply state predictions:
+% 	(1) MII drivers should drive less often when fertilized by complimentary sperm. 
+% 	(2) Drive may be more severs in interspecific crosses in which sperm have not specifically evolved to resists a driver. 
+% 	(3) ???
 Any more specific hypotheses -- maybe making this last sentence more explicit, e.g.

### 2014-05-19 writeup_and_notes/male_dependent_drive.tex
added hypotheses to discussion

@@
 %%%sperm-acting facilitator
-When the modifier of drive is tightly linked to the driver and enhances drive, 
+When the modifier of drive is tightly linked to the driver and
+enhances drive (i.e. in coupling phase), 
 	we label  this non-recombining drive/modifier haplotype as the $B^+$ allele.  
@@
 If a sperm-acting allele that restores fairness to female meiosis arises on
-	the drive background (creating a third allele $B^{-}$, Model 4), it displaces the ancestral driver (Equation \ref{coupling}), 
+	the drive background (i.e. in replusion phase, creating a third allele $B^{-}$, Model 4), it displaces the ancestral driver (Equation \ref{coupling}), 
 	and often returns to a lower equilibrium frequency than the the $B$ allele, further decreasing the extent of drive in the population.  

### 2014-05-19 writeup_and_notes/male_dependent_drive.tex
new version of trajectories_of_sperm_based_supressors.eps, and added to description of repulsion phase in text

@@
 %	that is sperm-acting alleles that act to restore  fairness of female meiosis, can. %If these sperm-acting drive suppressors arise on
-If a sperm-acting allele that restores fairness to female meiosis arises on
-	the drive background (i.e. in replusion phase, creating a third allele $B^{-}$, Model 4), it displaces the ancestral driver (Equation \ref{coupling}), 
+\gc{Imagine a sperm-acting allele that restores fairness to female meiosis arises on
+	the drive background, i.e. in repulsion phase, creating a
+        third allele $B^{-}$, Model 4. This new allele still
+        experiences femaledrive, but does so at a lower rate (or not
+        at all) when it
+        is fertilized by itself.} This $B^{-}$ displaces the ancestral driver (Equation \ref{coupling}), 
 	and often returns to a lower equilibrium frequency than the the $B$ allele, further decreasing the extent of drive in the population.  

### 2014-05-19 writeup_and_notes/male_dependent_drive.tex
added more to description of repulsion phase in text

@@
 
-% \begin{figure}
-% \includegraphics[width = 0.8 \textwidth]{Figures/trajectories_of_sperm_based_supressors.eps} 
-% \caption{Trajectories of two sperm-based supressors of drive. Possibly
-% merge this figure the 3 allele cartoon}  \label{Trajectories_of_supressors}
-% \end{figure}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%
@@
         allele/haplotype at this locus) and arises on the drive-background. 
-%We complement this single-locus approach %of an allele with pleiotropic effects in both sexes, 
-%	with alternative models of two linked loci - one a female driver, 
-%	and the other, a sperm allele which modifies the effect of drive upon fertilization. 
-%In this model, a female meiotic driver with no sperm-dependency 
-%	initially reaches drive-viability equilibrium (with two alleles
-%	$A$ and $B$ are the ancestral non-driver and driver alleles). 
-%Subsequently, a sperm-acting modifier of female meiotic drive arises tightly
- %      linked to this balanced locus (effectively creating a third
- %       allele/haplotype at this locus). 
 Tight linkage offers the best
@@
 
-%\yb{??In the SUPP we explore the dynamics of a sperm modifier of female drive unlinked to the female driver??}
-
 %%%sperm-acting facilitator
@@
 	we label  this non-recombining drive/modifier haplotype as the $B^+$ allele.  
-%We first consider a sperm-acting modifier of
-%	female meiotic drive tightly linked to the $B$ background (Model 4).
-%We treat this non-recombining drive/modifier haplotype as a third allele, and we label this allele as $B^+$ when it enhances drive. 
-%(i.e. the non-recombining haplotype of a
- %       driver allele and sperm-acting drive modifier)
-%The $B^+$ allele 
-%	the $B$ and  $B^{+}$ drive in $AB$ and $AB^{+}$ heterozygotes (see Figure \ref{Eggsperm_3_allele_cartoon})
-%		generating a third `$B^{+}$' allele (i.e. the non-recombining haplotype of a
-   %     driver allele and sperm-acting drive modifier). 
 The $B^+$ allele acts in sperm to increase the effectiveness of drive for both
 	the $B$ and  $B^{+}$ alleles in $AB$ and $AB^{+}$ heterozygotes (see Figure \ref{Eggsperm_3_allele_cartoon}, and Model 4 in the Supplementary Worksheet). 
-%	for a simple schematic of this drive model.  
 Naively, $B^{+}$ may spread by capitalizing on the additional drive it causes; however,  
 	this is not the case for a few simple reasons. 
-First, %the sperm-acting drive facilitator (
+First, 
 	the novel $B^{+}$ haplotype
-	%) 
 	arises when the ancestral driver 
-	%($B$) 
 	is at drive-selection balance, 
@@
 	the $B$  allele drive ($B^+$ sperm meeting $AB$ eggs), because $B$ is initially more common than $B^{+}$. 
-%As $B^{+}$ facilitates $B$ these interactions frequently form 
-%	$BB^{+}$ heterozygotes that suffer the same fitness costs as $BB$. 
 Therefore, sperm-acting drive facilitator alleles experience a profound disadvantage 
@@
 	the drive background, i.e. in repulsion phase, creating a
-        third allele $B^{-}$, Model 4. This new allele still
-        experiences femaledrive, but does so at a lower rate (or not
-        at all) when it
-        is fertilized by itself.} This $B^{-}$ displaces the ancestral driver (Equation \ref{coupling}), 
+        third allele $B^{-}$, Model 4. \gc{This new allele still
+        experiences female drive when fertilized by A or B sperm, but 
+       is does not drive when fertilized by another $B^{-}$ so it
+       avoid the excess formation of low fitness homozygotes.} This
+     allows the $B^{-}$ to displace the ancestral driver (Figure
+     \ref{Eggsperm_3_allele_cartoon}, Equation \ref{coupling}), 
 	and often returns to a lower equilibrium frequency than the the $B$ allele, further decreasing the extent of drive in the population.  

### 2014-05-19 writeup_and_notes/male_dependent_drive.tex
made the tec compile

@@
 These results are consistent with current knowledge of sperm-dependent meiotic drivers 
-(loci whose distortion of female meiosis depends on sperm genotype),}
+(loci whose distortion of female meiosis depends on sperm genotype),
 	and suggest that the requirement of fertilization for the
@@
 %	that is sperm-acting alleles that act to restore  fairness of female meiosis, can. %If these sperm-acting drive suppressors arise on
-\gc{Imagine a sperm-acting allele that restores fairness to female meiosis arises on
+Imagine a sperm-acting allele that restores fairness to female meiosis arises on
 	the drive background, i.e. in repulsion phase, creating a

### 2014-05-19 writeup_and_notes/male_dependent_drive.tex

@@
 \begin{center}
-$^1$Department of Plant Biology, \\ University of Minnesota – Twin
-Cities, MN, 55108 
-$^2$Center for Population Biology \& Department of Evolution and Ecology \\ University of California - Davis \\ Davis, CA, 95616
+$^1$Department of Plant Biology, \\ University of Minnesota --€“ Twin
+Cities. St. Paul MN, 55108 \\
+$^2$Center for Population Biology \& Department of Evolution and Ecology \\ University of California -- Davis. Davis, CA, 95616
 \end{center}
@@
 Genomic conflicts arise when an allele gains an evolutionary advantage at a cost to organismal fitness. 
-Because only one of four products of female meiosis is transmitted to
-offspring, o\"{o}genesis is inherently susceptible to such conflicts,
-as alleles can compete to be transmitted into the egg. 
+%Because only one of four products of female meiosis is transmitted to offspring, 
+O\"{o}genesis is inherently susceptible to such conflicts
+because alleles compete to be the  product of female meiosis transmitted to the egg. 
 %(meiotic drive).  
-Genomics conflicts drive the rapid evolution of suppression mechanisms, and
-	many facets of  o\"{o}genesis and gametogenesis have been 
+Alleles that distort meiosis in their favor (i.e. meiotic drivers) often decrease organismal fitness,
+	and therefore indirectly favor the evolution of mechanisms to suppress meiotic drive. % drive the rapid evolution of suppression mechanisms, and
+In this light, many facets of  o\"{o}genesis and gametogenesis have been 
 	interpreted as mechanisms of protection against genomic outlaws. 
@@
 	potentially providing an opportunity for sperm alleles to
-        meddle with its outcome in their own favor? %and facilitate meiotic drive? 
+        meddle with its outcome and help like-alleles drive in heterozygous females? %and facilitate meiotic drive? 
 The population genetic theory presented herein suggests that
@@
 	and suggest that the requirement of fertilization for the
-        completion of female meiosis, and variation in stage of meiosis that
-        fertilization occurs at may represent a mechanism employed by females 
+        completion of female meiosis
+        %, and variation in stage of meiosis that
+        %fertilization occurs at 
+        potentially represents a mechanism employed by females 
 	 to ensure a fair meiosis. %prevent meiotic drive.

### 2014-05-19 writeup_and_notes/male_dependent_drive.tex
yb minor modifications to self-promoting driver

@@
 	and evaluate whether a driver of this form could generate a sustained conflict favoring the evolution of suppressors. 
-We concluded with an analysis of a related model (Model 3) - in which
+We conclude with an analysis of a related model (Model 3) - in which
 drivers influence their transmission in females via paternal genotype,
@@
 	allow the fixation of this driver when common, but 
-	preclude its invasion when rare (Equation \ref{ineqA}). 
+	preclude its invasion when rare (Equation \ref{ineqA} and Figure \ref{Eggsperm}B). 
 In this case, the driver will be fixed if it reaches a frequency greater than some threshold 

### 2014-05-19 writeup_and_notes/male_dependent_drive.tex
yb minor modifications to two-locus models

@@
 	with alternative models of two loci - one a female driver, 
-	and the other, a sperm allele which modifies the effect of drive upon fertilization. 
+	and the other, a sperm-acting allele which modifies the effect of drive upon fertilization. 
 In this model, a female meiotic driver with no sperm-dependency 
@@
 	$A$ and $B$ are the ancestral non-driver and driver alleles). 
-Subsequently, a sperm-acting modifier of female meiotic drive arises at another lous.
+Subsequently, a sperm-acting modifier of female meiotic drive arises at another locus.
 
@@
         chance for a collaboration to evolve between a driver and
-       an enhancer of the drive, as recombination breaks up drive haplotypes \citep{Thomson1974,Charlesworth1978,Haig1991}. 
+       a sperm-acting drive  enhancer, as recombination breaks up drive haplotypes \citep{Thomson1974,Charlesworth1978,Haig1991}. 
 Additionally, tight linkage between female driver and sperm modifier is consistent with the nature of well characterized drive systems which are often maintained as polymorphic inversions with numerous linked modifiers \cite{Burt2006}.
@@
 	the drive background, i.e. in repulsion phase, creating a
-        third allele $B^{-}$, Model 4. \gc{This new allele still
+        third allele $B^{-}$, Model 4. This new allele still
         experiences female drive when fertilized by A or B sperm, but 
        is does not drive when fertilized by another $B^{-}$ so it
-       avoid the excess formation of low fitness homozygotes.} This
+       avoid the excess formation of low fitness homozygotes. This
      allows the $B^{-}$ to displace the ancestral driver (Figure
@@
 	or is unlinked to the drive locus (Model 6), it readily invades a population segregating
-	for the drive system (Figure \ref{Eggsperm_3_allele_cartoon} and eqns \ref{A-} and \ref{unlinked}).  
+	for the drive system (Figure \ref{Eggsperm_3_allele_cartoon} and Equations \ref{A-} and \ref{unlinked}).  
 This allele lowers the frequency of the original driver (perhaps to zero),

### 2014-05-19 writeup_and_notes/male_dependent_drive.tex
small edits with Yaniv to abstract, intro, discussion. I also deleted some blocks of commented out text from methods

@@
 	interpreted as mechanisms of protection against genomic outlaws. 
-Why then is female meiosis often left uncompleted until after fertilization - 
+Why then is female meiosis often left uncompleted until after
+fertilization \gc{in many animals}- 
 	potentially providing an opportunity for sperm alleles to
@@
 That is, in many animal species, female meiosis is completed only after fertilization \citep{Masui_book}, 
-creating ample opportunity for interaction between the sperm and female meiotic machinery. 
+creating ample opportunity for interaction between the sperm and
+female meiotic machinery \gc{(note that the alternation of generations in
+plants precludes this)}. 
 Therefore, a `green-bearded' \citep{Gardner2010} sperm-acting allele that recognizes and facilitates the meiotic drive of a genetically equivalent allele in heterozygous females %female meiosis %(a  heteromorphic dyad),
@@
         alleles with female meiosis. 
-For example it would be interesting to know whether changes that are predicted to lower the opportunity for female meiotic drive,
+\gc{For example, we would predict that a lower the opportunity for female meiotic drive,
         e.g. an evolutionary transition to inbreeding or a high
-        selfing rate in animals, are often accompanied by a relaxation of the requirement of fertilization for the completion of meiosis.
-
-% \yb{this feels weak if we don't say anything more explicit. I also think maybe we simply state predictions:
-% 	(1) MII drivers should drive less often when fertilized by complimentary sperm. 
-% 	(2) Drive may be more severs in interspecific crosses in which sperm have not specifically evolved to resists a driver. 
-% 	(3) ???
-Any more specific hypotheses -- maybe making this last sentence more explicit, e.g.
-self-fertilizing or otherwise frequently inbreeding animals will not require fertilization for the completion of meiosis...
-[note - this runs counter to our empirical examples of mice and C-elegens]}
+        selfing rate in animals, would often accompanied by a
+        relaxation of the requirement of fertilization for the
+        completion of female meiosis.}
 
@@
 	model of female meiotic drive (Model 2); a male-genotype-dependent model of
-	female meiotic drive (Model 3); and more complex two-locus sperm-dependent drive (Models 4-6). %changed this to minimize confusion with refs, figs, equations, etc.
-
-%Below, we highlight some of our derivations and analytical results. 
-%The fitness of genotype, $g$ is $w_g$ independent of its sex, and its 
+	female meiotic drive (Model 3); and more complex two-locus sperm-dependent drive (Models 4-6). 
 All models include a biallelic locus ($A/B$) with non-driving and driving alleles in frequencies $f_A$ and $f_B$, 
@@
 After a complete generation $f_g'' = f_g' w_g/\bar{w}$, where $\bar{w}$ is the population mean fitness and equals $\Sigma w_g f_g'$. 
-%All models include a biallelic locus with non-driving and driving alleles in frequencies $f_A$ and $f_B$, respectively, while more complex models include more loci and/or alleles.  
-%Transmission rules describing the outcomes of all matings in each
-%models are presented in a \yb{Supplementary Worksheet}, and complete derivations of our analytical approximations are in the
-%\gc{Mathematical Supplement}.  
 Because many of our analyses assume genotype frequencies expected under Hardy Weinberg Equilibrium (HWE), we note which analytical results are approximations. 

### 2014-05-19 writeup_and_notes/signif_statement.tex
added signif_statement.tex

There are a number of evolutionary mystries concerning meiosis, however,little attention has been directed at the odd fact that fertilization by sperm is required to complete female meiosis in many species of animal. 
At first sight this seems strange as it seems to presents a dangerous opportunity for selfish alleles in sperm to distort the outcome of female meiosis in their own favor creating the opportunity for conflicts between male and female genomes. In this paper, through mathematical modeling, we show that in fact quite the opposite is true, sperm are usually selected to ensure the fairness of meiosis.
Intuitively this comes about because selfish sperm have to immediately suffer any ill consequences of their actions in the resulting zygote, and so it is strongly in sperm's interests to keep female meiosis fair.
We hypthosize that the wide-spread requirement of fertilization to complete female meiosis may represent a long-term 
collaboration between male and female gametes to ensure fairness and the production of high quality gametes.

### 2014-05-19 writeup_and_notes/signif_statement.tex
updated the signif statement

@@
-There are a number of evolutionary mystries concerning meiosis, however,little attention has been directed at the odd fact that fertilization by sperm is required to complete female meiosis in many species of animal.
+Evolutionary biologists are fascinated by the oppotunities for benefits, costs, and opportunities for conflict during meiosis, however,little attention has been directed at the odd fact that fertilization by sperm is required to complete female meiosis in many species of animal.
 At first sight this seems strange as it seems to presents a dangerous opportunity for selfish alleles in sperm to distort the outcome of female meiosis in their own favor creating the opportunity for conflicts between male and female genomes.
 In this paper, through mathematical modeling, we show that in fact quite the opposite is true, sperm are usually selected to ensure the fairness of meiosis.
-Intuitively this comes about because selfish sperm have to immediately suffer any ill consequences of their actions in the resulting zygote, and so it is strongly in sperm's interests to keep female meiosis fair.
-We hypthosize that the wide-spread requirement of fertilization to complete female meiosis may represent a long-term 
-collaboration between male and female gametes to ensure fairness and the production of high quality gametes.
+Intuitively this comes about because selfish sperm have to immediately suffer any ill consequences of their actions in the resulting zygote, and so it is strongly in sperm's interests to keep genomes from gaming female meiosis.
+We hypthosize that the wide-spread requirement of fertilization to complete female meiosis may represent a long-term collaboration between male and female gametes to ensure fairness and the production of high quality gametes.

### 2014-05-19 writeup_and_notes/signif_statement.tex
i accidently lost the sigstatement. but recovered it and pushed it back
yb modified significance statement

@@
-Evolutionary biologists are fascinated by the oppotunities for benefits, costs, and opportunities for conflict during meiosis, however,little attention has been directed at the odd fact that fertilization by sperm is required to complete female meiosis in many species of animal.
+\documentclass{article}
+\usepackage{graphicx}
+
+\begin{document}
+Female meiosis is the time during development where an allele can act to directly modify its representation in the next generation. 
+The potential benefits to allele what succeed in this competition and collateral damage to organismal fitness are thought to have shaped basic features of female reproduction. 
+However, little attention has been directed at the odd fact that fertilization by sperm is required to complete female meiosis in many species of animal. 
 At first sight this seems strange as it seems to presents a dangerous opportunity for selfish alleles in sperm to distort the outcome of female meiosis in their own favor creating the opportunity for conflicts between male and female genomes.
@@
 Intuitively this comes about because selfish sperm have to immediately suffer any ill consequences of their actions in the resulting zygote, and so it is strongly in sperm's interests to keep genomes from gaming female meiosis.
-We hypthosize that the wide-spread requirement of fertilization to complete female meiosis may represent a long-term collaboration between male and female gametes to ensure fairness and the production of high quality gametes.
+We hypothesize that the wide-spread requirement of fertilization to complete female meiosis may represent a long-term collaboration between male and female gametes to ensure fairness and the production of high quality gametes.
+\end{document}

### 2014-05-19 writeup_and_notes/male_dependent_drive.tex

@@
 Why then is female meiosis often left uncompleted until after
-fertilization \gc{in many animals}- 
+fertilization in many animals -- 
 	potentially providing an opportunity for sperm alleles to
@@
 	creating ample opportunity for interaction between the sperm and
-female meiotic machinery \gc{(note that the alternation of generations in
-plants precludes this)}. 
+	female meiotic machinery (note that the alternation of generations in
+	likely precludes this interaction in plants). 
 Therefore, a `green-bearded' \citep{Gardner2010} sperm-acting allele that recognizes and facilitates the meiotic drive of a genetically equivalent allele in heterozygous females %female meiosis %(a  heteromorphic dyad),
@@
 As a consequence of this intense competition, alleles that gain a transmission advantage during reproduction 
-	can succeed evolutionarily, even if they lower the 
-        fitness or the organism in which they reside. 
+	can succeed evolutionarily even if they lower organismal fitness. 
 This generates numerous conflicts  \cite{Burt2006} including sexual conflicts between mates \cite{Arnqvist2005}, 
@@
 Additionally, when this self-promoting driver can spread, it can only rarely 
-	be maintained as a protected polymorphism, and due to its positive frequency dependence 
+	be maintained as a protected polymorphism, and due to its positive frequency dependence,  
 	it spends very little time at intermediate frequency.
@@
 	frequency until the fitness cost experienced when homozygous balances the transmission advantage.
-By contrast a sperm promotor of female drive is only effective when matched with a heterozygote female -- meaning that when rare this allele rarely enhances female drive. 
+By contrast, a sperm promoter of female drive is only effective when matched with a heterozygote female -- meaning that, when rare, this allele rarely enhances female drive. 
 Even worse, when it does so it will preferentially find itself in a low fitness drive homozygote. 
 Not only are drive-promoting sperm alleles unable to create a sustained genetic conflict, 
-	but alleles in sperm with the opposite effect - that is those that prevent their own drive through female meiosis do maintain a polymorphisms and 
+	but alleles in sperm with the opposite effect - that is those that prevent their own drive through female meiosis do maintain a polymorphism and 
 	provide evolution  with time and opportunity to further minimize drive.
@@
  reduce the opportunity of female meiotic drive regardless of linkage or phase.
-We predict that female drive, at a stage influencable by meiotic
-drive,  should be reduced when fertilized by
-complimentary sperm. For example, MII female drive should be more severe in
-interspecific crosses as sperm will not specifically evolved to
-resists a female driver.
+
+Our theory offers a novel interpretation of a baffling idiosyncrasy of female meiosis -- 
+	we argue that females do not complete meiosis until after fertilization as a mechanism to encourage meiotic fairness. 
+In addition to providing a new interpretation of a basic biological process, our theory provides numerous testable predictions. 
+We predict that female drive, at a stage influenceable by sperm,  
+	should be reduced when fertilized by
+	complimentary sperm. 
+Therefore, For example, MII female drive should be exceptionally severe in
+	interspecific crosses between animals which complete the second meiotic division after fertilization,  
+	as heterospecific sperm will not specifically evolve to resists a female driver in another species. 
+Our theory also encourages phylogenetic hypotheses concerning the relationship between the opportunity 
+	for female meiotic drive the requirement of fertilization for the completion of female meiosis. 
+For example, we predict that a lower the opportunity for female meiotic drive,
+        (e.g. an animal lineage with a long history of high inbreeding or selfing) 
+        would often accompanied by a relaxation of the requirement of fertilization for the
+        completion of female meiosis.
 
@@
 
-The interests of fertilizing sperm and female are quite well aligned during syngamy -- producing a viable and potentially fit offspring. 
+Our results highlight potentially counterintuitive results of complex genetic conflicts. 
+Despite much opportunity for conflict between sperm and females over fertilization \citep{Partridge1998},
+	the interests of fertilizing sperm and female are quite well aligned during syngamy.
+While conflict between mother and her alternative chromosomes ensue, 
+	fertilizing sperm decidedly side with mom, as both have a shared interest in producing a viable and potentially fit offspring. 
 %As noted above there are even opportunities for coevolution between sperm-based suppressors of female meiotic drive and the female genome.
-Thus it seems plausible that aspects of oogenesis may actively evolve to facilitate the influence of sperm on ensuring a fair  outcome of female meiosis which generates fit offspring.
-In light of that it is interesting to consider whether the phylogenetic variation in the stage of oogenesis when fertilization takes place reflects not
-	only species' different reproductive ecologies, but also
-        different selection pressures shaping the interaction of sperm
-        alleles with female meiosis. 
-\gc{For example, we would predict that a lower the opportunity for female meiotic drive,
-        e.g. an evolutionary transition to inbreeding or a high
-        selfing rate in animals, would often accompanied by a
-        relaxation of the requirement of fertilization for the
-        completion of female meiosis.}
+%Thus it seems plausible that aspects of oogenesis may actively evolve to facilitate the influence of sperm on ensuring a fair  outcome of female meiosis which generates fit offspring.
+%In light of that it is interesting to consider whether the phylogenetic variation in the stage of oogenesis when fertilization takes place reflects not
+%	only species' different reproductive ecologies, but also
+%        different selection pressures shaping the interaction of sperm
+%        alleles with female meiosis. 
+%\gc{For example, we would predict that a lower the opportunity for female meiotic drive,
+%        e.g. an evolutionary transition to inbreeding or a high
+%        selfing rate in animals, would often accompanied by a
+%        relaxation of the requirement of fertilization for the
+%        completion of female meiosis.}
 
@@
  The four panels correspond to different dominance relationships.
-In the parameter space below the invasion (solid) line the self-promotor
+In the parameter space below the invasion (solid) line the self-promoter
  driver can invade. In the parameter space below the fixation (long
- dash) line the self-promotor can fix in the population. In the last
+ dash) line the self-promoter can fix in the population. In the last
  two panels the invasion line is above the fixation line and so the

### 2014-05-19 writeup_and_notes/male_dependent_drive.tex

@@
 	model of female meiotic drive (Model 2); a male-genotype-dependent model of
-	female meiotic drive (Model 3); and more complex two-locus sperm-dependent drive (Models 4-6). 
+	female meiotic drive (Model 3); and more complex two-locus sperm-dependent drive models (Models 4-6). 
 All models include a biallelic locus ($A/B$) with non-driving and driving alleles in frequencies $f_A$ and $f_B$, 
@@
 Transmission rules describing the outcomes of all matings in each
-	models are presented in a Supplementary Worksheet. 
+	model are presented in a Supplementary Worksheet. 
 The fitness of genotype, $g$ is sex-independent and equals $1$, $1-h_s$, and $1-s$ for genotypes $AA$, $AB$, and $BB$, respectively. %  and  $w_{AA} = 1$, $w_{AB} = 1-s_h$, and $w_{BB}=1-s$. 
@@
 We iterate exact genotype frequency recursions ($f_g''= f_g'w_g/\bar{w}$) over generations to produce the frequency trajectories shown in the inset
-Figure \ref{Invasion_space} by plotting $f_B=f_{BB}+ \frac{1}{2}f_{AB}$
+Figure \ref{IEggsperm}B by plotting $f_B=f_{BB}+ \frac{1}{2}f_{AB}$
 over time. %, and to assess the analytical approximations, below.
@@
 show that this approximation is reasonably correct for small
-parameter values; however, this approximation underestimates $f_B^*$ for large parameter values, presumably because they result in strong departures from HWE.
+parameter values; however, it underestimates $f_B^*$ for large parameter values, presumably because they result in strong departures from HWE.
  %range of other costs to the drive alleles other than a
@@
 	a rare paternal-genotype-dependent driver invades when 
-	$s\lessapprox (d_{het}-1/2)/(d_{het})$, and when common this driver fixes if 
+	$s\lessapprox (d_{het}-1/2)/(d_{het})$, and when common, this driver fixes if 
 	$s\lessapprox d_{hom}-1/2$, approximations well supported by exact results (Supplementary Figure \ref{Effect_of_dominance}).
@@
 Assuming that a traditional driver segregates at its equilibrium frequency (Model 1), we investigate the evolution of tightly linked (Models 4 and 5), and unlinked (Model 6) sperm-dependent modifiers of drive.
-In Models 4 and 5, we treat this two-lous system as if it consists of
+In Models 4 and 5, we treat this two-locus system as if it consists of
 a single locus with three alleles: $A$, $B$ and $C$, corresponding to the case when the
@@
 For all parameters sustaining a polymorphism at the drive locus ($s>d_0-1/2$), this corresponds to a decrease in frequency of the $C$ allele when it enhances drive ($\epsilon >0$ -- the $B^+$ model, above),
-	and a decrease in frequency of the $C$ allele when it suppresses drive ($\epsilon <0$ -- the $B^-$ model, above). 
+	and an increase in frequency of the $C$ allele when it suppresses drive ($\epsilon <0$ -- the $B^-$ model, above). 
 More generally, even when the cost of drive is not fully recessive, the $B^-$ allele will invade and fix under all parameters sustaining a previous polymorphism at the drive locus (see Mathematical Supplement).
@@
 ][]{Brandvain2012}.
-The spread of this driver does not, however, depend on assumption of HWE, and linkage equilibrium - exact results resemble analytical approximations. 
+%The spread of this driver does not, however, depend on assumption of HWE, and linkage equilibrium - exact results resemble analytical approximations. 
 %\paragraph{three allele system}

### 2014-05-19 writeup_and_notes/male_dependent_drive.tex
yb made minor improvements to the supp fig legends

@@
   that these alleles reach fixation conditional on invading). In the
-  white area the allele can not invade ($x \geq1$), in the solid red
+  white area the allele cannot invade ($x \geq1$), in the solid red
   area the allele can invade and fix when rare. In the left panel we
@@
   function of the drive parameter. The solid line shows results
-  obtained using the recursion, the dots our approximation given by \eqref{thresh2}/}  \label{bistable_additive}
+  obtained using the recursion, the dots our approximation given by Equation \eqref{thresh2}}  \label{bistable_additive}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width = 0.8 \textwidth]{../Scripts/effect_of_dominance_on_invasion_space_four_graph.eps} 
-\caption{Invasion analysis of an allele whose effect
- on female meoisis is mediated by the genotype of the fertilizing
- male.  As above wee assume a heterozygous female will transmit the $B$ allele 
-  with probabilities  $\frac{1}{2}$,  $\frac{1}{2} + h(d-\frac{1}{2}) $ or $d$, 
- if she mated with a n $AA$, $AB$, or $BB$ male,  respectively.  The allele suffers a recessive fitness cost $s$.  
+\includegraphics[width = 0.7 \textwidth]{../Scripts/effect_of_dominance_on_invasion_space_four_graph.eps} 
+\caption{Exact results of invasion analysis of an allele whose effect
+ on female meiosis is mediated by the genotype of the fertilizing
+ male.  
+ A heterozygous female transmits the $B$ allele 
+  with probabilities  $\frac{1}{2}$,  $d_{het}=\frac{1}{2} + h(d_{hom}-\frac{1}{2}) $ or $d_{hom}$, 
+ if she mated with a n $AA$, $AB$, or $BB$ male,  respectively.  
+ The allele suffers a recessive fitness cost $s$.  
  The four panels correspond to different dominance relationships.
@@
  driver can invade. In the parameter space below the fixation (long
- dash) line the self-promoter can fix in the population. In the last
+ dash) line the self-promoter can fix. In the last
  two panels the invasion line is above the fixation line and so the
  allele can be maintained as a polymorphism in that thin slice of
- parameter space between the two lines. These results were 
-  obtained using the recursion. In the final panel we show the
+ parameter space between the two lines. 
+ In the final panel we show the
   fixation line (small dashes) as predicted by our HWE approximation
-  ($(d_{het}-1/2)/d_{het}$), see the `dominance' section of the
-  Appendix for more details.
+  ($(d_{het}-1/2)/d_{het}$) see the appendix for more details.
 }  \label{Effect_of_dominance}

### 2014-05-19 writeup_and_notes/male_dependent_drive.tex
fixed small typo

--- 2014-05-19 writeup_and_notes/male_dependent_drive.tex
@@
 	for female meiotic drive the requirement of fertilization for the completion of female meiosis. 
-For example, we predict that a lower the opportunity for female meiotic drive,
-        (e.g. an animal lineage with a long history of high inbreeding or selfing) 
+For example, we predict that a lower opportunity for female meiotic drive,
+        e.g. an animal lineage with a long history of high inbreeding
+        or selfing, 
         would often accompanied by a relaxation of the requirement of fertilization for the

### 2014-05-20 writeup_and_notes/male_dependent_drive.tex
fixed minor tyos in appendix

@@
 	$f_{AA}'=f_A (f_{AA }+ f_{AB} (1 - d))$, 
-	$f_{AB}'=f_A (f_{BB} + f_{ABd}) + f_B (f_{AA} + f_{AB} (1 - d))$,
-	and $f_{BB}= f_B (f_{AB} d + f_{BB})$. 
+	$f_{AB}'=f_A (f_{BB} + f_{AB}d) + f_B (f_{AA} + f_{AB} (1 - d))$,
+	and $f_{BB}'= f_B (f_{AB} d + f_{BB})$. 
 %Assuming fitnesses $w_{AA}=1$, $w_{AB}-1 s_h$ and $w_{BB}=1-s$, 
@@
 %As above, exact frequencies after drive, random mating and selection are $f_i''= f_i'w_i/\bar{w}$. 
-We iterate exact genotype frequency recursions ($f_g''= f_g'w_g/\bar{w}$) over generations to produce the frequency trajectories shown in the inset
-Figure \ref{IEggsperm}B by plotting $f_B=f_{BB}+ \frac{1}{2}f_{AB}$
+We iterate exact genotype frequency recursions ($f_g''= f_g'w_g/\bar{w}$) over generations to produce the frequency trajectories shown in the inset of 
+Figure \ref{Eggsperm}B by plotting $f_B=f_{BB}+ \frac{1}{2}f_{AB}$
 over time. %, and to assess the analytical approximations, below.
@@
 	$f_{AB}'' =   w_{AB}(f_{AA} [f_B + f_{AB}/2] + f_{AB} f_B + f_{BB} [ f_A - f_{AB} d_{hom}])/\bar{w}$, 
-	and 	$f_{AB}'' =   w_{AB}(f_{AB} (f_{AB} d_{het}/2 + f_{BB} d_{hom}) + f_{BB} f_B)/\bar{w}$. 
+	and 	$f_{BB}'' =   w_{BB}(f_{AB} (f_{AB} d_{het}/2 + f_{BB} d_{hom}) + f_{BB} f_B)/\bar{w}$. 
 % Above, $d > 1/2$ and $h$ ($0 \geq h \geq 1$) acts to specify the dominance of the $B$

### 2014-05-20 writeup_and_notes/pnasSubmission/PNASSubmission.TEX
fixed minor tyos in significance statement and push PNAS folder

--- 2014-05-19 writeup_and_notes/male_dependent_drive.tex
@@
-%%!TEX TS-program = latex
-\documentclass[12pt,letterpaper]{article}
-\usepackage{epsfig}                 
-\usepackage[authoryear]{natbib}
-\usepackage{graphicx}
+%% PNAStwoS.tex
+%% Sample file to use for PNAS articles prepared in LaTeX
+%% For two column PNAS articles
+%% Version1: Apr 15, 2008
+%% Version2: Oct 04, 2013
+\bibliographystyle{pnas}
+%% BASIC CLASS FILE
+\documentclass{pnastwo}
+
+%% ADDITIONAL OPTIONAL STYLE FILES Font specification
+
+%\usepackage{pnastwoF}
+
 \usepackage{amsmath}
-\usepackage{psfrag}
-\usepackage{mathabx}
-\renewcommand{\baselinestretch}{1.6}
-\large
-\pagenumbering{arabic}
-\usepackage[usenames,dvipsnames]{color}
-\usepackage{fullpage}
-\bibliographystyle{genetics}
-\usepackage{multirow}
-\newcommand{\ye}{\hat{y}}
-\newcommand{\xe}{\hat{x}}
-\usepackage{color}
-\usepackage[normalem]{ulem}  
-	\newcommand{\gc}[1]{{ \color{red} #1}}
-	\newcommand{\yb}[1]{{ \color{blue} #1}}
+%% OPTIONAL MACRO DEFINITIONS
+\def\s{\sigma}
+%%%%%%%%%%%%
+%% For PNAS Only:
+\url{www.pnas.org/...}
+\copyrightyear{20XX}
+\issuedate{Issue Date}
+\volume{Volume}
+\issuenumber{Issue Number}
+%\setcounter{page}{2687} %Set page number here if desired
+%%%%%%%%%%%%
 
-%Why do animals allow male control of female meiosis?
-%Why does female meiosis allow the opportunity for exploitation by selfish sperm?
-%sperm and egg genomes both favour the peaceful resolution of female meiotic drive
-%strange bed fellows
+\begin{document}
 
-%\title{Why does female meiosis allow the opportunity for exploitation by selfish sperm? BETTER TITLE?}
-\title{
-%Co-operation between the sexes trumps selfish control of female  meiotic drive. 
-%Or 
-Sperm should evolve to make female meiosis fair.
-%Sperm do not evolve to collaborate in female meiotic drive. 
-%Or Why does female meiosis allow the opportunity for exploitation by selfish sperm? 
+\title{Sperm should evolve to make female meiosis fair}
+
+\author{Yaniv Brandvain\affil{1}{Department of Plant Biology. University of Minnesota -- Twin Cities, St. Paul, MN, USA},
+\and Graham Cooop\affil{2}{Department of Ecology and Evolution. University of California -- Davis. Davis, CA, USA}
 }
 
-\author{Yaniv Brandvain$^1$ \\ email: ybrandvain@gmail.com  \and Graham Coop$^2$ \\ email: gmcoop@ucdavis.edu }
+\contributor{Submitted to Proceedings of the National Academy of Sciences
+of the United States of America}
 
-\usepackage{fancyhdr}
-\pagestyle{fancy}
-\lhead{}
-\rhead{}
-\renewcommand{\headrulewidth}{0.0pt}
-\rfoot{}
-\cfoot{\thepage}
-\date{}
-\bibliographystyle{plain}
-\begin{document}
+%%%Newly updated.
+%%% If significance statement need, then can use the below command otherwise just delete it.
+\significancetext{Female meiosis is the time during development where an allele can act to directly modify its representation in the next generation. 
+The potential benefits to an allele that succeeds in this competition and collateral damage to organismal fitness are thought to have shaped basic features of female reproduction. 
+However, little attention has been directed at the odd fact that fertilization by sperm is required to complete female meiosis in many species of animal. 
+At first sight this seems strange as it seems to presents a dangerous opportunity for selfish alleles in sperm to distort the outcome of female meiosis in their own favor creating the opportunity for conflicts between male and female genomes.
+In this paper, through mathematical modeling, we show that in fact quite the opposite is true, sperm are usually selected to ensure the fairness of meiosis. 
+Intuitively this comes about because selfish sperm have to immediately suffer any ill consequences of their actions in the resulting zygote, and so it is strongly in sperm's interests to keep genomes from gaming female meiosis.
+We hypothesize that the wide-spread requirement of fertilization to complete female meiosis may represent a long-term collaboration between male and female gametes to ensure fairness and the production of high quality gametes.}
+
 \maketitle
-\begin{center}
-$^1$Department of Plant Biology, \\ University of Minnesota --€“ Twin
-Cities. St. Paul MN, 55108 \\
-$^2$Center for Population Biology \& Department of Evolution and Ecology \\ University of California -- Davis. Davis, CA, 95616
-\end{center}
 
-%FIgures 
-{\bf Abstract:}
-Genomic conflicts arise when an allele gains an evolutionary advantage at a cost to organismal fitness. 
+\begin{article}
+\begin{abstract}
+{Genomic conflicts arise when an allele gains an evolutionary advantage at a cost to organismal fitness. 
 %Because only one of four products of female meiosis is transmitted to offspring, 
@@
 because alleles compete to be the  product of female meiosis transmitted to the egg. 
-%(meiotic drive).  
 Alleles that distort meiosis in their favor (i.e. meiotic drivers) often decrease organismal fitness,
-	and therefore indirectly favor the evolution of mechanisms to suppress meiotic drive. % drive the rapid evolution of suppression mechanisms, and
-In this light, many facets of  o\"{o}genesis and gametogenesis have been 
+	and therefore indirectly favor the evolution of mechanisms to suppress meiotic drive. In this light, many facets of  o\"{o}genesis and gametogenesis have been 
 	interpreted as mechanisms of protection against genomic outlaws. 
@@
 	potentially providing an opportunity for sperm alleles to
-        meddle with its outcome and help like-alleles drive in heterozygous females? %and facilitate meiotic drive? 
+        meddle with its outcome and help like-alleles drive in heterozygous females? 
 The population genetic theory presented herein suggests that
 	sperm nearly always evolve to increase the fairness of female
-        meiosis in the face of genomic conflicts. %%felt this
-                                %%might be an important addition as
-                                %%they will spread if no cost to organism
+        meiosis in the face of genomic conflicts.
 These results are consistent with current knowledge of sperm-dependent meiotic drivers 
@@
         completion of female meiosis
-        %, and variation in stage of meiosis that
-        %fertilization occurs at 
         potentially represents a mechanism employed by females 
-	 to ensure a fair meiosis. %prevent meiotic drive.
-\newpage
+	 to ensure a fair meiosis.}
+\end{abstract}
 
-\section*{Introduction}
-Despite the apparent unity of the organism, `selfish' alleles can
+\keywords{meiosis | genomic conflict | meiotic drive | fertilization}
+
+%\abbreviations{SAM, self-assembled monolayer; OTS, octadecyltrichlorosilane}
+
+\dropcap{D}espite the apparent unity of the organism, `selfish' alleles can
 gain an evolutionary advantage at a cost to  individual fitness
-\citep{Burt2006}, often by exploiting meiosis and gametogenesis.
-Because only one of the four products of female meiosis is transmitted to the egg, female meiosis is particularly vulnerable to such exploitation \citep{Sandler1957,Pardo-ManuelDeVillena2001a}. 
+\cite{Burt2006}, often by exploiting meiosis and gametogenesis.
+Because only one of the four products of female meiosis is transmitted to the egg, female meiosis is particularly vulnerable to such exploitation \cite{Sandler1957,Pardo-ManuelDeVillena2001a}. 
 An allele that biases female meiosis in its favor (i.e.
-a meiotic driver), may increase in frequency even if it entails a pleiotropic fitness cost \citep{Prout1973}, generating a genetic conflict between the success of the driver and organismal fitness.
+a meiotic driver), may increase in frequency even if it entails a pleiotropic fitness cost \cite{Prout1973}, generating a genetic conflict between the success of the driver and organismal fitness.
 Meiotic drivers observed in both plants
-\citep{Buckler1999,Fishman2005,Fishman2008}, and animals
-\citep{Agulnik1990,Wu2005,Pardo-ManuelDeVillena2001c} highlight this
+\cite{Buckler1999,Fishman2005,Fishman2008}, and animals
+\cite{Agulnik1990,Wu2005,Pardo-ManuelDeVillena2001c} highlight this
 conflict -- the selfish benefits and the associated
 pleiotropic fitness costs of drive sustain a balanced polymorphism
-\citep{Prout1973}, 
-and often generate ongoing evolutionary escalations of drive suppressors and enhancers \citep{Dawe1996,Fishman2008}. 
+\cite{Prout1973}, 
+and often generate ongoing evolutionary escalations of drive suppressors and enhancers \cite{Dawe1996,Fishman2008}. 
 The threat of meiotic drive to organismal fitness is potentially so
@@
 meiosis and o\"{o}genesis, including the initial genome doubling in
-meiosis I \citep{Haig1991}, arrested female meiosis
-\citep{Mira1998}, the structure of centromere machinery \citep{Malik2002a,Malik2009}, and sex differences in the recombination rate \citep{Haig2010,Brandvain2012} 
-have perhaps evolved to enforce fairness by disrupting meiotic drive \citep{Rice2013}.
+meiosis I \cite{Haig1991}, arrested female meiosis
+\cite{Mira1998}, the structure of centromere machinery \cite{Malik2002a,Malik2009}, and sex differences in the recombination rate \cite{Haig2010,Brandvain2012} 
+have perhaps evolved to enforce fairness by disrupting meiotic drive \cite{Rice2013}.
 
@@
 it is potentially open to sabotage by a virtual stranger -- a haploid sperm genome.
-That is, in many animal species, female meiosis is completed only after fertilization \citep{Masui_book}, 
+That is, in many animal species, female meiosis is completed only after fertilization \cite{Masui_book}, 
 	creating ample opportunity for interaction between the sperm and
@@
 	likely precludes this interaction in plants). 
-Therefore, a `green-bearded' \citep{Gardner2010} sperm-acting allele that recognizes and facilitates the meiotic drive of a genetically equivalent allele in heterozygous females %female meiosis %(a  heteromorphic dyad),
+Therefore, a `green-bearded' \cite{Gardner2010} sperm-acting allele that recognizes and facilitates the meiotic drive of a genetically equivalent allele in heterozygous females %female meiosis %(a  heteromorphic dyad),
      could presumably rapidly spread through a population. 
@@
 The stage at which female meiosis requires fertilization varies widely across taxa, and in a few animals  
-	female meiosis is complete upon ovulation (that is, fertilization is not required for female meiosis \citep[e.g. Table 1 in 
-][]{Masui_book}).
+	female meiosis is complete upon ovulation (that is, fertilization is not required for female meiosis (e.g. Table 1 in \cite {Masui_book}).
 
@@
 machinery provided by the sperm, disrupts MII meiotic segregation
-in the egg, leading to a triploid zygote \citep{McNally2012}. 
+in the egg, leading to a triploid zygote \cite{McNally2012}. 
 Thus, there is direct mechanistic evidence that sperm can influence female meiosis.
@@
  transmission of alternate alleles in heterozygous females can depend on sperm haplotype.  
- While ruling out the alternative hypothesis of early selection on zygotes in these cases is challenging  \citep[see pages
-52-54 in][]{Burt2006}, it appears that the extent to which \emph{In} and \emph{Om} distort the second female meiotic division partially depends on the genotype of the fertilizing sperm \citep{Agulnik1993,Wu2005}.
-% Both of these female drive systems operate by distorting the second meiotic division, 
-% and in both systems the outcome of female meiosis depends the genotype of the fertilizing sperm \citep{Agulnik1993,Wu2005}.  
-%While such genetic evidence is suggestive, we note the difficulty in
-%ruling out the alternative hypothesis of early selection on zygotes \citep[pages
-%52-54][]{Burt2006}.
+ While ruling out the alternative hypothesis of early selection on zygotes in these cases is challenging (see pages 52-54 in \cite{Burt2006}), it appears that the extent to which \emph{In} and \emph{Om} distort the second female meiotic division partially depends on the genotype of the fertilizing sperm \cite{Agulnik1993,Wu2005}.
 
 We develop population genetic models exploring the evolution of sperm influence on female meiotic drive. 
-%, focussing on models in which `self-promoting' alleles in sperm facilitate drive of like genotypes during gametogenesis in heterozygous females.
 We first focus on models in which `self-promoting' alleles in
   sperm facilitate drive of like alleles during gametogenesis in heterozygous females. 
-%In this article we explore through simple population genetic models the consequences of alleles that influence the outcome of female meiosis.
 These models show that such sperm-acting alleles 
@@
 Such model universally reveal that sperm-acting drive modifiers are favored only if the suppress drive. 
-These results highlight the fact that the interests of sperm and maternal  genomes' are often aligned, as both are invested in the fate of the resultant zygote (as was speculated for the \emph{In} locus \citep{Pomiankowski1993}).
+These results highlight the fact that the interests of sperm and maternal  genomes' are often aligned, as both are invested in the fate of the resultant zygote (as was speculated for the \emph{In} locus \cite{Pomiankowski1993}).
 Therefore, there is little selective benefit to females in preventing sperm to influence female meioses.
-Rather, features of o\"{o}genesis  may evolve to facilitate an interaction with sperm %with female meiosis
+Rather, features of o\"{o}genesis  may evolve to facilitate an interaction with sperm
 	as yet another mechanism to enforce an equitable transfer of genetic material.
 
-\section*{Results}
-
-\subsection*{ Invasion of the population by a driving allele that promotes itself.}
-%%%%%%%%%%%%%%%%%%%%%%%%%
-% PLEIOTROPY MODEL [self-promoting]
-%%%%%%%%%%%%%%%%%%%%%%%%%
-
+\section{Results}
+\subsection{Invasion of an allele that promotes its own drive}
 In the standard single-locus, biallelic model of female meiotic drive,
 the driving allele is transmitted to the egg in heterozygotes with
-probability  $d > 1/2$, regardless of sperm genotype \cite[e.g.
-][ and see Model 1 in the Appendix for more details]{Ubeda2004}.
-%The timing of fertilization relative to female meiosis places another constraint on $d$, for example, if fertilization (and therefore, sperm dependent drive) takes place at MII (as in mammals),
-%	female drive requires an uneven number of crossovers between the centromere and the drive locus, 
-%	so $d$ is bounded to be $<0.83$ \citep[see ][ for discussion]{Buckler1999}.
-	%YB - MOVED THIS TO THE APPENDIX
-          To depict a case
-of a self-promoting meiotic driver,  we modify this standard model such 
-	that the driver is only effective when fertilized by a sperm carrying that allele (see Figure \ref{Eggsperm}A and Model 2 in the Supplementary Worksheet). 
+probability  $d > 1/2$, regardless of sperm genotype (e.g. \cite{Ubeda2004} and see Model 1 in the Appendix for more details). 
+To depict a case of a self-promoting meiotic driver,  we modify this standard model such 
+	that the driver is only effective when fertilized by a sperm carrying that allele (see Figure 1A and Model 2 	in the Supplementary Worksheet). 
 We then identify the conditions allowing for the spread of this self-promoting driver, 
@@
 
-For comparative purposes, we briefly present the standard drive model %as Model 1 in the Appendix, 
-%equations \ref{eqn.A1}-\ref{eqn.A3}, 
-	%and refer readers to the previous literature on this model 
-	\citep[see e.g. ][ for additional results]{Prout1973,Ubeda2004}. 
+For comparative purposes, we briefly present the standard drive model 
+	(see e.g. \cite{Prout1973,Ubeda2004} for additional results). 
 Assuming the driving allele is deleterious in both sexes, but fully recessive 
@@
 However, when $s$ is large ($s>(2d-1)/(2)$), a driver cannot fix, and
-will be maintained as a protected polymorphism \citep{Prout1973}. 
+will be maintained as a protected polymorphism \cite{Prout1973}. 
 The parameter space where the allele can invade but not fix is shown in white
-        in Figure \ref{Eggsperm}B. %, and it can fix in the colored regions. 
+        in Figure 1B. . 
 When the allele is maintained as a polymorphism, it provides an opportunity for the evolution of
 	drive suppressors, corresponding well to empirical examples of
-        female meiotic drive \citep[reviewed in ][]{Burt2006}.
+        female meiotic drive (reviewed in \cite{Burt2006}).
 
@@
 	by uniting driving female gametes with sperm enabling that drive.
-        % zygotes when it drives. in heterozygotes. 
-%, and therefore there is a drive-associated fitness cost even when rare. 
 It must therefore overcome a drive-associated homozygous fitness cost simply to spread when rare. 
@@
 When rare, a fully recessive, self-promoting driver can only invade when $s$ 
-	%, the selection against the drive homozygote, 
 	is less than $(2 d - 1)/(4 d)$. 
 This analytical approximation, derived from Equation \eqref{deltadriver} assuming Hardy-Weinberg, 
-	closely matches results obtained by exact numerical iteration (Figure \ref{Eggsperm}B).
+	closely matches results obtained by exact numerical iteration (Figure 1B).
 
@@
 However, once relatively common, it  rapidly achieves fixation due to its
-	positive frequency dependent behavior (Figure \ref{Eggsperm}B.1).  
-%This lack of a balanced polymorphism %(Figure \ref{Invasion_space}) 
-%	precludes the evolution of an
-%	allele that suppresses this form of meiotic drive in females.
+	positive frequency dependent behavior (Figure 1B.1).  
 The positive frequency dependence of a self-promoting driver can also
@@
 	allow the fixation of this driver when common, but 
-	preclude its invasion when rare (Equation \ref{ineqA} and Figure \ref{Eggsperm}B). 
+	preclude its invasion when rare (Equation \ref{ineqA} and Figure 1B). 
 In this case, the driver will be fixed if it reaches a frequency greater than some threshold 
 	(approximated in Equation \ref{bistab_homozy_approx} and  presented exactly in 
-	Figure \ref{Bistab_homozyg_cost_fig}) and lost otherwise. 
+	Figure S1) and lost otherwise. 
 For most parameters, this threshold is likely too high to be reached by drift, 
@@
 
-%I SUBSTANTIALLY REWROTE THIS
-%The positive frequency dependence of a self-promoting driver can
-%induce bistability in its spread.
-%OLD GRAHAM WORDS
- %, i.e. there is a critical frequency
-%above which it can spread quickly to fixation and below which it moves towards loss (an
-%unstable equilibrium frequency). In 
-%the appendix, we develop a simple
-%approximation to this bistability. 
-%When $s$ is too large to allow a rare driver to increase in frequency, but small enough to allow a common driver to fix (the inequality in Equation \ref{ineqA}),
-%there is  an unstable equilibrium frequency, $f^*_{B\text{ recessive}}$  
-%	above which it can spread quickly to fixation and below which it moves towards loss 
-%	($f^*_{B\text{ recessive}}$ is approximated in Equation \ref{bistab_homozy_approx}.  
-%	We compare this approximation to the exact result in Figure \ref{Bistab_homozyg_cost_fig}).	
-%OLD GRAHAM WORDS
-%This bistability occurs in the region of parameter space in Figure
-%\ref{Invasion_space}. In Supplementary Figure
-%\ref{Bistab_homozyg_cost_fig} we further explore this bistability and
-%show that the unstable equilibrium frequency is high across much of
-%this parameter space, and so in practice the self promoting allele
-%only invade in the orange space. (In the appendix we develop a simple
-%approximation to the unstable equilibrium frequency
-%\eqref{bistab_homozy_approx} that is also shown in  Supplementary Figure
-%\ref{Bistab_homozyg_cost_fig}).
-%This unstable equilibrium frequency is high across much of
-%	parameter space, likely too high to be reached by drift, 
-%	and therefore  the fate of self-promoting drivers  will  likely be determined
-%	by the more restrictive invasion criteria rather than the fixation criteria.
-
-%OLD YANIV WORDS
-% Compared to the invasion criteria, the parameter space
-% allowing the fixation of the allele if it is above a cutoff frequency is much broader -- 
-% 	a common, recessive, self-promoting driver fixes if
-%         $s<(2d-1)/2$ approximately. 
-% Therefore, when $(2 d - 1)/(4 d)<s<(2d-1)/(2)$, 
-% 	a recessive, self-promoting driver will deterministically fix if drift, mutation, 
-% 	or migration pressure bring its frequency above $(1-2d+4ds)/(2s(2d-1))$, 
-% 	but it will be lost otherwise.
 Inclusion of a heterozygous fitness cost (i.e. $w_{AB}=1-s_h$)
@@
 	frequency exceeds some threshold  (Equation \ref{thresh2}, exact results in 
-	Figure \ref{bistable_additive}).
-%	$f_B^*$
-%	because it pays the  heterozygous fitness cost but drives
- %      ineffectively (see Figure S\ref{bistable_additive})
-%While a standard female driver invades from low frequency when the transmission advantage outweighs the fitness cost (XXX),
-%	a rare self-promoter pays a fitness cost, but drives only weakly because complementary sperm are uncommon. 
-%This frequency-independent fitness cost, and positive frequency dependent drive generates a bistable system. 
-%Below a cutoff frequency the self-promoter driver deterministically decreases in frequency 
-%	because it pays the  heterozygous fitness cost but drives
- %      ineffectively (see Figure S\ref{bistable_additive}). 
-%Above this frequency, the allele increases deterministically because the commonness of complementary sperm 
-%	make drive effective enough to overcome the associated fitness cost.
+	Figure S2).
+
 This bistability prevents self-promoting drivers from invading 	
@@
 	observed as stable polymorphisms in natural populations. 
-This lack of a balanced polymorphism %(Figure \ref{Invasion_space}) 
+This lack of a balanced polymorphism
 	precludes the evolution of an
@@
 Although the allelic identity of sperm could plausibly influence the outcome of female meiosis, 
-	limited gene expression in sperm \citep[e.g.][]{Joseph2004}
+	limited gene expression in sperm (e.g. \cite{Joseph2004}) 
 	suggests a model where sperm influence female meiosis via expression of the fertilizing male's
@@
 Here, a rare driver invades when $s$ is less than $(d_{het}-1/2)/d_{het}$, and usually fixes when it invades.
-	% fixes when $s$ is less than $d_{hom} - 1/2$.
 However, when the distorting effect of genotype displays strong dominance in its
@@
 	when the cost of the drive is recessive
-	(see Figure \ref{Effect_of_dominance}, and Equation \ref{polymale}).  
-%Presumably this zone of polymorphism comes about because positive frequency 
-%	dependence diminishes with the dominance of the male effect driver. 
-%As an analytical example, consider the case of male-genotype dependent deleterious recessive female driver (assuming, as above, $w_{BB}=1-s$, and $w_{AA}=w_{AB}=1$). 
-%Here, a rare driver invades when $s$ is less than $(d_{het}-1/2)/d_{het}$, and fixes when $s$ is less than $d_{hom} - 1/2$.
-%Intermediate values of $s$, which represent a small proportion of parameter space can maintain a polymorphism. 
+	(see Figure S3, and Equation \ref{polymale}).  
 While mathematically interesting, it does not seem particularly realistic to think that the
@@
 	required for this polymorphism make us doubt it general applicability.
-%Thus we regard this case as interesting but not very general.
 
@@
 
-\begin{figure}
-\includegraphics[width = 0.8 \textwidth]{Figures/selfprom.eps}
-\caption{{\bf{A.}} A visual depiction of our model of 'self-promoting' driver. 
-	Transmission probabilities for alleles through female
- 	 meiosis depend on sperm genotype. 
-	 The non-driving $A$-allele, and self-promoting $B$-allele
-	 are represented by unfilled and filled circles, respectively. 
-	 {\bf{B.}} Evolution of a self-promoter and standard driver. 
-	 	Assuming that the fitnesses of drive homozygotes and 
-		heterozygotes are $1-s$ and $1$, respectively. 
-	 	\emph{Main figure:} Boundary conditions for the invasion 
-		and fixation of self-promoting and standard meiotic drivers, 
-		with drive coefficient, $d$. Colored regions depict exact results, 
-		while lines  represent analytical approximations. 
-	 \emph{B1:}  Trajectories of sperm-dependent female drive
-		 %alleles that can invade the population, 
-		 each allele has $s =0.1$ against the
-		 homozygotes.  The drive coefficient is denoted by color. 
- }   
-\label{Eggsperm}
-\end{figure}
-
-%\begin{figure}
-%	\rotatebox{90}{
-%\includegraphics[width = 0.5
-   %       \textwidth]{Figures/Sperm_egg_cartoon_paper_Fig1.eps}}
-%\caption{transmission probabilities for alleles through female
-%  meiosis depend on sperm genotype. 2 allele model. \gc{merge with 2,
-%    rewrite both captions}}  
-%\label{Eggsperm_2_allele_cartoon}
-%\end{figure}
-
-% (we note that the timing of fertilization relative to female meiosis places another constraint on $d$, for example, when drive takes place druing the second meitoic division
-%	it requires an uneven number of crossovers between the centromere and the drive locus \citep[see ][ for discussion]{Buckler1999}).
-
-%\begin{figure}
-%\includegraphics[width = 0.8 \textwidth]{Figures/invasion_space_recessive_driver.eps}
-%\caption{{\bf A.} Invasion analysis for a female meiotic drive allele with
-%  recessive costs (selection coefficient $s$). First consider a simple female meiotic
-%  drive system, where the allele is transmitted a fraction $d>1/2$ of
-  %meioses in female heterozygotes, inspective
-%  of the allele carried by the fertilizing sperm. 
-%Such an allele can invade the population over all of the parameter space, and will be maintained as a
-%  protected polymorphism in the white region, and can fix in the
-%  population in all of the parameter space colored green or orange. 
-%The solid back line marks the criteria for the allele to fix $s< (2d-1)/2$.
-%Now consider an allele that is fairly transmitted through
-%  female meiosis when fertilized by a sperm carrying a different
-%  allele, but at a fraction $d>1/2$ when a heterozygote egg is fertilized
-%  by a sperm carrying the allele (as depicted in Figure
-%  \ref{Eggsperm_2_allele_cartoon}). Such an allele can only invade in
-%  the region of parameter space colored orange, and it fixes over all
- % of this parameter region. The approximate criteria for this allele
-%  to fix is given by the dashed line showing $s < (2d- 1)/(4d)$. 
-% {\bf B.} In the inset we show trajectories of sperm-dependent female drive
- %alleles that can invade the population, each allele has $s =0.1$ against the
-% homozygotes and $d =0.6$, $0.7$, $0.8$ shown in yellow, blue, and red. 
-% \gc{Likely make this Figure 1B.} } \label{Invasion_space}
-%\end{figure}
-
-\subsection*{Two locus models of sperm-dependent female drive}
-
-%\subsection*{ sperm-egg haplotype balanced drive systems}
-
-\begin{figure}
-	\rotatebox{90}{\includegraphics[width = 0.8
-          \textwidth]{Figures/Sperm_egg_cartoon_paper_Fig2.eps}}
-\caption{Dynamics in 3 allele system
-{\bf (A) } The sperm-dependent transmission probabilities of alleles through
-female meiosis (female germ cells depicted as circles, sperm indicated by wiggly
-tails). In each case the top and bottom entry in each cell give the
-transmission probability of the left and right allele in the female
-being transmitted to the egg. In the left panel we show the transmission probabilities for
-  A, B, and B'. Where B' is a driver allele which is either promotes
-  itself through through female meiosis when sperm (a B$^+$ allele,
-  $d'>d$) or self-avoiding through the interaction of sperm with female meiosis (a B$^-$ allele,
-$d'<d$). In the right panel we depict the transmission probabilities for
-  A, B, and A$^-$ allele a non-driven allele that acts to reduce the
-  effects of female drive through sperm.
-{\bf (B)} Frequency trajectories in the three allele system. }
- \label{Eggsperm_3_allele_cartoon}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%
-% TWO LOCUS MODEL [modify (general) / facilitate (helps ) / prevent (stops)] 
-%%%%%%%%%%%%%%%%%%%%%%%%%
-
+\subsection{Two locus models of sperm-dependent female drive}
 Models 2 and 3, above, explored the dynamics of an allele that drove in females when signaled by a complementary signal in sperm.   
@@
         chance for a collaboration to evolve between a driver and
-       a sperm-acting drive  enhancer, as recombination breaks up drive haplotypes \citep{Thomson1974,Charlesworth1978,Haig1991}. 
+       a sperm-acting drive  enhancer, as recombination breaks up drive haplotypes \cite{Thomson1974,Charlesworth1978,Haig1991}. 
 Additionally, tight linkage between female driver and sperm modifier is consistent with the nature of well characterized drive systems which are often maintained as polymorphic inversions with numerous linked modifiers \cite{Burt2006}.
@@
 The $B^+$ allele acts in sperm to increase the effectiveness of drive for both
-	the $B$ and  $B^{+}$ alleles in $AB$ and $AB^{+}$ heterozygotes (see Figure \ref{Eggsperm_3_allele_cartoon}, and Model 4 in the Supplementary Worksheet). 
+	the $B$ and  $B^{+}$ alleles in $AB$ and $AB^{+}$ heterozygotes (see Figure 2, and Model 4 in the Supplementary Worksheet). 
 Naively, $B^{+}$ may spread by capitalizing on the additional drive it causes; however,  
 	this is not the case for a few simple reasons. 
-First, 
-	the novel $B^{+}$ haplotype
+First, the novel $B^{+}$ haplotype
 	arises when the ancestral driver 
@@
 	three allele system that allows the sperm-acting drive facilitator $B^{+}$ to
-	invade the population (Appendix Model 4, eqn. \eqref{coupling}, and Mathematical Supplement).
+	invade the population (Appendix Model 4, Equation \eqref{coupling}, and Mathematical Supplement).
 
-%%%sperm-acting drive suppressors
 While  sperm enhancement of a female drive cannot displace a polymorphic female driver, sperm based drive suppressors can. 
-%	that is sperm-acting alleles that act to restore  fairness of female meiosis, can. %If these sperm-acting drive suppressors arise on
 Imagine a sperm-acting allele that restores fairness to female meiosis arises on
@@
      allows the $B^{-}$ to displace the ancestral driver (Figure
-     \ref{Eggsperm_3_allele_cartoon}, Equation \ref{coupling}), 
+     2, Equation \ref{coupling}), 
 	and often returns to a lower equilibrium frequency than the the $B$ allele, further decreasing the extent of drive in the population.  
@@
 	or is unlinked to the drive locus (Model 6), it readily invades a population segregating
-	for the drive system (Figure \ref{Eggsperm_3_allele_cartoon} and Equations \ref{A-} and \ref{unlinked}).  
+	for the drive system (Figure 2 and Equations \ref{A-} and \ref{unlinked}).  
 This allele lowers the frequency of the original driver (perhaps to zero),
 	and spreads to fixation if it does not carry strong fitness costs
-	(Figure \ref{Eggsperm_3_allele_cartoon} and \ref{Trajectories_of_supressors}A). 
-This result is consistent with previous work showing that drive suppressors unlinked to, or in repulsion phase with drivers usually invade polymorphic drive systems \citep[e.g. ][]{Brandvain2012}.  
+	(Figure 2). 
+This result is consistent with previous work showing that drive suppressors unlinked to, or in repulsion phase with drivers usually invade polymorphic drive systems (e.g. \cite{Brandvain2012}).  
 Therefore, all two-locus models of sperm influence on female drive suggest that
 	sperm will evolve to oppose female meiotic drive.
-
-%More surprisingly, sperm-acting suppressors of female drive can spread
-%	even when tightly linked to the original drive system (Figures \ref{Eggsperm_3_allele_cartoon} and \ref{Trajectories_of_supressors}B). 
-%The $B^{-}$ haplotype (representing the tightly linked drive/drive suppressor complex) displaces the original driver 
-%	by driving when not costly, but avoiding low fitness homozygotes by disrupting drive when transmitted to sperm. %  ($BB^-$ and $B^-B^-$). 
-%Despite displacing the ancestral driver (the $B$ haplotype), the equilibrium frequency of the $B^-$ haplotype is often lower than that of $B$ \yb{(point to supp?)}.
-
-%\gc{Do we mention the In and Om locus here or in the discussion as a
-%	 potential e.g. Do we mention the fact that the frequency of these
-%	 newly invade sperm dodging alleles can be lower than the original system?}\\
 
@@
 	and conflicts between alleles that are over-transmitted in meiosis and the organisms they injure while doing so. 
-Such conflicts and their resolution likely play a major role in the structure and evolution of many basic biological processes \citep{Rice2013}.
+Such conflicts and their resolution likely play a major role in the structure and evolution of many basic biological processes \cite{Rice2013}.
 
@@
 However, this is not the case.
-We find that an allele which acts through sperm to distort female meiosis in its favor %YB [too awkward / unclear?] 
+We find that an allele which acts through sperm to distort female meiosis in its favor 
 	can rarely spread through a population if it bears any cost. 
@@
         drivers and sperm alleles.
-%Male modification of recombination rates. \yb{trying to think through this \dots{}}
 
@@
 of forming low fitness homozygotes. 
-%In fact, an allele in sperm which causes the alternate allele to drive can be favored, a counterintuitive result 
-%	consistent with the tenuous evidence that the non-driving
-%        $ln^-$ allele is over transmitted through female meiosis in $ln^-/ln^+$ heterozygote when fertilized by $ln^+$
-%        sperm \cite{agulnick, Pomiankowski1993}.   
-% YB - WE DID NOT TALK ABOUT THIS IN THE RESULTS [ALTHOUGH IT'S TRUE... PERHAPS TOO DISTRACTING]
 More generally, alleles that act through sperm to
@@
 
-%More generally both linked and unlinked alleles may often spread that act through sperm to
-%reduce the opportunity of female meiotic drive in order to disable
-%segregating drive systems, by for example reducing the mechanistic
-%asymmetries that female drivers exploit.
-
 Our results highlight potentially counterintuitive results of complex genetic conflicts. 
-Despite much opportunity for conflict between sperm and females over fertilization \citep{Partridge1998},
+Despite much opportunity for conflict between sperm and females over fertilization \cite{Partridge1998},
 	the interests of fertilizing sperm and female are quite well aligned during syngamy.
@@
 	fertilizing sperm decidedly side with mom, as both have a shared interest in producing a viable and potentially fit offspring.
-%As noted above there are even opportunities for coevolution between sperm-based suppressors of female meiotic drive and the female genome.
-%Thus it seems plausible that aspects of oogenesis may actively evolve to facilitate the influence of sperm on ensuring a fair  outcome of female meiosis which generates fit offspring.
-%In light of that it is interesting to consider whether the phylogenetic variation in the stage of oogenesis when fertilization takes place reflects not
-%	only species' different reproductive ecologies, but also
-%        different selection pressures shaping the interaction of sperm
-%        alleles with female meiosis. 
-%\gc{For example, we would predict that a lower the opportunity for female meiotic drive,
-%        e.g. an evolutionary transition to inbreeding or a high
-%        selfing rate in animals, would often accompanied by a
-%        relaxation of the requirement of fertilization for the
-%        completion of female meiosis.}
 
-  \appendix
-
-\section*{Appendix}
+\section{Appendix}
 We highlight simple analytical results from: 
@@
 	model are presented in a Supplementary Worksheet. 
-The fitness of genotype, $g$ is sex-independent and equals $1$, $1-h_s$, and $1-s$ for genotypes $AA$, $AB$, and $BB$, respectively. %  and  $w_{AA} = 1$, $w_{AB} = 1-s_h$, and $w_{BB}=1-s$. 
+The fitness of genotype, $g$ is sex-independent and equals $1$, $1-h_s$, and $1-s$ for genotypes $AA$, $AB$, and $BB$, respectively. 
 Genotypic frequencies  equal  $f_g$ for adults in the current generation, 
@@
 
-\subsection*{Models 1-3. Single-locus drive}
-\subsubsection*{Model 1. Traditional driver}
+\subsection{Models 1-3. Single-locus drive}
+\subsubsection{Model 1. Traditional driver:}
 In the standard female drive model, meiosis in males is fair such that $A/B$ heterozygotes contribute $A$ and $B$ alleles with equal probabilities; however, $A/B$ females transmit the $B$ allele with probability $d>1/2$.
@@
 	female drive requires an uneven number of crossovers between the centromere and the drive locus, 
-	so $d$ is bounded to be $<0.83$ \citep[see ][ for discussion]{Buckler1999}. 
+	so $d$ is bounded to be $<0.83$ (see \cite{Buckler1999} for discussion). 
 After drive and random mating, genotype frequencies are 
-	$f_{AA}'=f_A (f_{AA }+ f_{AB} (1 - d))$, 
-	$f_{AB}'=f_A (f_{BB} + f_{ABd}) + f_B (f_{AA} + f_{AB} (1 - d))$,
-	and $f_{BB}= f_B (f_{AB} d + f_{BB})$. 
-%Assuming fitnesses $w_{AA}=1$, $w_{AB}-1 s_h$ and $w_{BB}=1-s$, 
-%Frequencies of genotype, $i$, after drive and selection are 
-%	\begin{equation} f_i''= f_i'w_i/\bar{w}\end{equation}
-%	where $w_i$ is genotypic fitness, and mean fitness, $\bar{w} = \Sigma f_i'w_i$. 
+\begin{eqnarray*}
+f_{AA}'&=&f_A (f_{AA }+ f_{AB} (1 - d)),\\
+f_{AB}'&=&f_A (f_{BB} + f_{ABd}) + f_B (f_{AA} + f_{AB} (1 - d)),\\
+f_{BB}'&=& f_B (f_{AB} d + f_{BB})
+\end{eqnarray*}
 As detailed above, exact frequencies after drive, random mating and selection are $f_g''= f_g'w_g/\bar{w}$. 
@@
         recessive.
-        %\gc{isn't this last expression exact?} YB: exact-ish.. all assume HWE.. but i think we don't expect a deviation from HWE here??  it's prob easiest to kept them all as approx?
 
-\subsubsection*{Model 2. Single locus, sperm-dependent drive.}
+\subsubsection{Model 2. Single locus, sperm-dependent drive:}
 Our single-locus model of sperm-dependent drive resembles the traditional driver, with the caveat that the $B$ allele drives in heterozygous females only when fertilized by $B$-bearing sperm. 
 Therefore, genotype frequencies after drive are 
-	$f_{AA}'=f_A^2$, 
-	$f_{AB}'=f_A f_B+f_B (f_{AA} + f_{AB}(1 - d)) $,
-	and $f_{BB}'= f_B (f_{AB} d + f_{BB} )$. 
-%As above, exact frequencies after drive, random mating and selection are $f_i''= f_i'w_i/\bar{w}$. 
-We iterate exact genotype frequency recursions ($f_g''= f_g'w_g/\bar{w}$) over generations to produce the frequency trajectories shown in the inset
-Figure \ref{IEggsperm}B by plotting $f_B=f_{BB}+ \frac{1}{2}f_{AB}$
-over time. %, and to assess the analytical approximations, below.
+\begin{eqnarray*}
+f_{AA}'&=&f_A^2,\\
+f_{AB}'&=&f_A f_B+f_B (f_{AA} + f_{AB}(1 - d)) ,\\
+f_{BB}'&=&f_B (f_{AB} d + f_{BB} )
+\end{eqnarray*}
+
+We iterate exact genotype frequency recursions ($f_g''= f_g'w_g/\bar{w}$) over generations to produce the frequency trajectories shown in the inset of
+Figure 1B by plotting $f_B=f_{BB}+ \frac{1}{2}f_{AB}$
+over time. 
 To assess invasion or fixation criteria, as well as bistability points, we iterate this system and
@@
 
-% the $B$ allele In our first scenerio allele $A$ is the ancestral allele and allele
-%$B$ is our sperm dependent drive allele.
-%The sperm dependence of the drive is a form of assortative
-%transmission (mating), and so we have to keep track of the genotype frequencies.
-%Our genotype frequencies at birth are $x_{AA},~x_{AB},$ and
-%$x_{BB}$. These three genotypes have fitnesses $w_{AA},~w_{AB},$ and
-%$w_{BB}$, which are identical between males and females. Because 
-%fitnesses are equivalent in both sexes, genotypic frequencies
-%are identical in males and females, but in the following
-%equations they are marked with $\Mars$ and $\Venus$ respectively to
-%allow the reader to more easily follow the logic. 
-%The frequency of the three genotypes at birth in the next generation is:
-%\begin{equation}
-%x'_{BB} =\frac{1}{\bar{w} } \Big( w_{BB} x_{BB}^{\Venus} (w_{BB} x_{BB}^{\Mars} +\frac{1}{2}
-%  w_{AB} x_{AB}^{\Mars} ) + w_{AB} x_{AB}^{\Venus} (d w_{BB}
-%  x_{BB}^{\Mars} +\frac{1}{2} d
-%w_{AB} x_{AB}^{\Mars} ) \Big),  \label{eqn.A1}
-%\end{equation}
-%\begin{equation}
-%x'_{AA} =\frac{1}{\bar{w} } \Big(  w_{AA} x_{AA}^{\Venus} (w_{AA} x_{AA}^{\Mars} + \frac{1}{2}
-%   w_{AB} x_{AB}^{\Mars}) + w_{AB} x_{AB}^{\Venus}(\frac{1}{2} w_{AA} x_{AA}^{\Mars}+ \frac{1}{2} 
-%  w_{AB} x_{AB}^{\Mars}) \Big),  \label{eqn.A2}
-%\end{equation}
-%and
-%\begin{eqnarray}
-%x'_{AB} = & \frac{1}{\bar{w} } \Big( w_{AA} x_{AA}^{\Venus}  ( w_{BB}  x_{BB}^{\Mars} + \frac{1}{2}
-%  w_{AB}  x_{AB}^{\Mars} ) +w_{BB} x_{BB}^{\Venus} ( w_{AA}  x_{AA}^{\Mars} + \frac{1}{2}
-%  w_{AB}  x_{AB}^{\Mars} )  \nonumber \\
-%& + x_{AB}^{\Venus}(\frac{1}{2}  w_{AA}  x_{AA}^{\Mars}+\frac{1}{2} (1-d)
-% w_{AB}  x_{AB}^{\Mars} + \frac{1}{4}  w_{AB}
- %x_{AB}^{\Mars} + (1-d)  w_{BB}  x_{BB}^{\Mars} ) \Big) \label{eqn.A3}
-%\end{eqnarray}
-%where the mean fitness of the population is
-%\begin{equation}
-%\bar{w} = w_{AA} x_{AA}+w_{AB}  x_{AB} +w_{BB}  x_{BB}.
-%\end{equation}
-%We can then iterate this recursion over generations, for
-%example to produce the frequency trajectories shown in the inset
-%Figure \ref{Invasion_space} by plotting $x_{BB}+ \frac{1}{2}x_{AB}$
-%over time. We can also use this to assess the
-%invasion criteria for the allele by starting allele B at very low
-%frequency, and seeing if it increases across the generations.
-
-\paragraph{Recessive fitness cost of self-promoting driver:} %When the self-promoting driver has a recessive fitness cost, $s$ genotype frequencies after selection and drive equal
-%\begin{equation}	x'_{BB} =\frac{x_{B}^2(1-s)}{\bar{w} }  \Big( F+(1-F)(x_{B}+2x_{A} d) \Big) 	\end{equation}
-%\begin{equation}	x'_{AA} =\frac{1}{\bar{w} }  \Big(x_{A}^2\Big) 	\end{equation}
-%and
-%\begin{equation}	x'_{AB} =\frac{2 x_{A}x_{B} }{\bar{w} }  \Big(1-x_{B}(d-1/2)(1-F) \Big)	%\end{equation}
+\paragraph{Recessive fitness cost of self-promoting driver -- }
 When fully recessive, the change in frequency of the self-promoting driver across generations equals 
 \begin{equation}
-\Delta f_B=f_A f_B^2 \left( [1-F][ d(1-2 f_A s) - f_B s-1/2] - s  F \right)/\bar{w}
+\Delta f_B=\frac{f_A f_B^2}{\bar{w}} \left( \left(1-F\right)\left( d\left(1-2 f_A s\right) - f_B s-\frac{1}{2}\right) - s  F \right)
 \label{deltadriver}
 \end{equation}
-where $F$ is the deviation from genotypic frequencies expected under Hardy-Weinberg. % Trying to clarify some.   \gc{Hardy Weinberg Expectations} (HWE). 
-%For analytical solutions presented above, we assume that $F$ equals zero. 
+where $F$ is the deviation from genotypic frequencies expected under Hardy-Weinberg. 
 Assuming HWE  ($F=0$) a common, recessive, self-promoting driver invades if $s\lessapprox(2 d - 1)/(4 d)$, and fixes if
@@
  but it will be lost when introduced below this frequency. 
- %(i.e. this is an unstable equilibrium point). 
-% In Supplementary Figure \ref{Bistab_homozyg_cost_fig} we show
-% this approximation condition for bistability andd how it compared to
-% that obtained using the recursion. 
-Compared to exact results (Supplementary Figure \ref{Bistab_homozyg_cost_fig}), Equations
+ Compared to exact results (Supplementary Figure S1), Equations
  \eqref{ineqA} and \eqref{bistab_homozy_approx} offer reasonable, but imperfect approximations.
-%s_h and seems to serve as a lower bound to the true unstable equilibrium frequency.
 
-\paragraph{Cost of driver in heterozygotes:} %We explored a more complex model where the self-promoting driver has a cost, $s_h$ in heterozygotes (so $w_{AB} = 1-s_h$) in addition to the cost $s$, in homozygotes.
+ \paragraph{Cost of driver in heterozygotes -- } 
 When the fitness of drive heterozygotes is compromised ($s_h>0$), a self-promoting driver cannot invade when rare.
 This results from the fact that,
-when rare, %($f_B \ll 1$) 
+when rare, 
 $B$-bearing sperm and heterozygous eggs will rarely encounter one another ($\sim f_B^2$) but the allele still pays a
 cost in heterozygous individuals ($\sim f_B$). 
-%Model construction is similar to that above.
-%So long as there is any heterozygous fitness cost (i.e. $s_h > 0$) the self-promoting driver cannot invade when rare;
 However, this system too, is bistable -- as the driver increases in frequency it is more often fertilized by a driving sperm and therefore drives more effectively. 
@@
 Comparison of Equation \ref{thresh2} to exact results obtained by a
-simple parameter search (Supplementary Figure \ref{bistable_additive})
+simple parameter search (Supplementary Figure S2)
 show that this approximation is reasonably correct for small
 parameter values; however, it underestimates $f_B^*$ for large parameter values, presumably because they result in strong departures from HWE.
- %range of other costs to the drive alleles other than a
-%simple recessive cost, for example an additive cost. We modified the recursions given in eqn
-%\ref{eqn.A1}-\ref{eqn.A3} to feature an additive cost $s$ for each
-%copy of the $B$ allele. A none recessive cost to the drive allele can
-%cause bistability of the drive allele. The B allele is unable to enter
-%the population when rare, this results from the fact that
-%when the allele is rare ($x_B \ll 1$) sperm carrying it rarely
-%encounter heterozygous eggs ($\sim x^2$) but the allele still pays a
-%cost in heterozygous individuals.  By contrast, when the B allele is common ($x_B
-%\gg 0$) encounters between heterozygous eggs and $B$ allele sperm
-%become more common allowing drive to occur this can allow the $B$
-%allele to spread when common. In Supplementary Figure
-%\ref{bistable_additive} we show the unstable equilibrium frequency
-%above which the $B$ allele can invade. This equilibrium frequency was found by a simple
-%parameter search.  \gc{wonder if a bit of work could find this}
 
-\subsubsection*{Model 3. Single locus, paternal genotype dependent drive:}
+\subsubsection{Model 3. Single locus, paternal genotype dependent drive:}
 In the case when female meiotic drive depends on paternal genotype
-%\paragraph{Dominance} we also explored a model where
-%the genotype of the male who produced the fertilizing sperm affects female drive,
-%instead of just the allelic type of the sperm.
- %We assume a 
  heterozygous female will transmit the $B$ allele 
   with probabilities  $\frac{1}{2}$,  $d_{het}\geq 1/2 $ or $d_{hom}\geq d_{het}$, 
-%YB RE-PARAMETERIZING  %  $\frac{1}{2}$,  $\frac{1}{2} + h(d-\frac{1}{2}) $ or $d$,  YB RE-PARAMETERIZING 
  when mated with to $AA$, $AB$, or $BB$ males,  respectively. 
   In this model, genotype frequencies after drive, random mating, and selection are
-  	$f_{AA}'' =   w_{AA}\left( f_{AA} [2 f_A + f_{AB} ] + f_{AB}^2 [1 - d_{het}] \right)/(2\bar{w})$,
-	$f_{AB}'' =   w_{AB}(f_{AA} [f_B + f_{AB}/2] + f_{AB} f_B + f_{BB} [ f_A - f_{AB} d_{hom}])/\bar{w}$, 
-	and 	$f_{AB}'' =   w_{AB}(f_{AB} (f_{AB} d_{het}/2 + f_{BB} d_{hom}) + f_{BB} f_B)/\bar{w}$. 
-% Above, $d > 1/2$ and $h$ ($0 \geq h \geq 1$) acts to specify the dominance of the $B$
-%allele's effect on female meiotic drive through the males genotype.
-%YB TRY_1 We assume that in a mating between a heterozygous female and an $AA$, $AB$, or $BB$ male, the female will transmit the $B$ allele 
-% with probabilities  $\frac{1}{2}$,  $\frac{1}{2} + h(d-\frac{1}{2}) $, and $d$, respectively.
-  %Here $d > 1/2$ and $h$ ($0 \geq h \geq 1$) acts to specify the dominance of the $B$
-%allele's effect on female meiotic drive through the males genotype.
-% GC ORIGINAL  We assume that if a
-%heterozygous egg ($AB$) is fertilized by a sperm with the following
-%male ($\Mars$) genotypes the $B$ allele will be transmitted through
-%female meiosis with the following probabilities \\
-%\begin{tabular}{cccc}
-%$\Mars$ geno. & AA & AB  & BB\\
-%P(B). & $\frac{1}{2}$ &  $\frac{1}{2} + h(d-\frac{1}{2}) $ & $d$ 
-%\end{tabular}\\
-%where $d > 1/2$ and $h$ ($0 \geq h \geq 1$) acts to specify the dominance of the $B$
-%allele's effect on female meiotic drive through the males genotype. 
-%We set up a set of recursions using a recessive sex independent cost to
-%the driver, similar to those in eqn.
-%\ref{eqn.A1}-\ref{eqn.A1}, but now featuring these male genotype
-%transmission probabilities.
+
+\begin{eqnarray*}
+f_{AA}''&=&\frac{w_{AA}}{2\bar{w}}\left( f_{AA} (2 f_A + f_{AB} ) + f_{AB}^2 (1 - d_{het}) \right),\\
+f_{AB}''&=&\frac{w_{AB}}{\bar{w}}\big(f_{AA} (f_B + f_{AB}/2) +  f_{AB} f_B \\&& + f_{BB}(f_A - f_{AB} d_{hom})\big),\\
+f_{BB}''&=&\frac{w_{BB}}{\bar{w}}\left((f_{AB} (f_{AB} d_{het}/2 + f_{BB} d_{hom}) + f_{BB} f_B\right)
+\end{eqnarray*}
 
@@
 	$s\lessapprox (d_{het}-1/2)/(d_{het})$, and when common, this driver fixes if 
-	$s\lessapprox d_{hom}-1/2$, approximations well supported by exact results (Supplementary Figure \ref{Effect_of_dominance}).
+	$s\lessapprox d_{hom}-1/2$, approximations well supported by exact results (Supplementary Figure S3).
 Specifically, when drive in heterozygotes is large relative to that in homozygotes, 
@@
 As the frequency of the $B$ allele increases, 
-	%it occurs more often as a homozygote, and so 
 	it will be unable to 
@@
 By contrast, when $d_{het} \lessapprox 1/(3-2d_{hom})$ the case is reversed, and the model is bistable.
-%In Supplementary Figure \ref{Effect_of_dominance} we show the invasion
-%criteria for both the A and B alleles when they are rare. 
-%We find a narrow parameter range where B 
-%can invade but not fix.{eqmale}
-%This corresponds to B being less than fully recessive (i.e. $h>0$) and to a modest cost in homozygotes. 
-%This combination of parameters reflects the fact that when the allele is rare, 
-%it can increase in frequency due to fertilization of heterozygous eggs
-%by sperm from heterozygote males, where the sperm do not themselves
-%carry the allele. But as the B allele becomes common it will frequently be
-%present in the homozygous state in males, and so will be unable to
-%avoid the production of unfit homozygote zygotes leaving it trapped in
-%the population.
 
-\subsection*{Models 4-6. Two-locus, sperm-dependent drive.} 
+\subsection{Models 4-6. Two-locus, sperm-dependent drive} 
 Assuming that a traditional driver segregates at its equilibrium frequency (Model 1), we investigate the evolution of tightly linked (Models 4 and 5), and unlinked (Model 6) sperm-dependent modifiers of drive.
@@
 alleles at the drive locus (such that recombination is unexpected). 
-In all cases, we assume that fitness is independent of the drive modifier. %, so, $w_{AA}=1$, $w_{AB}=1-s_h$, and $w_{BB} = 1-s$. 
+In all cases, we assume that fitness is independent of the drive modifier.  
 We assume the traditional driver is transmitted to $d_0$ of gametes from female heterozygotes  when fertilized by wild-type sperm, and $d_1=d_0+\epsilon$ when fertilized by a sperm-acting drive-modifier.
@@
 Assuming HWE, a recessive fitness cost to drive, and assuming that the $A/B$ locus is at its equilibrium frequency, the change in frequency of a rare drive modifier is
-\begin{equation}
-	\Delta C\approx -\epsilon f_C \frac{ 1-2 d_0 (-2 d_0+s+2)+\sqrt{2} \sqrt{s (2 d_0 (d_0(s-2)+2)-1)}}{2 \bar{w}(1-2 d_0)^2}
+
+\begin{eqnarray}\nonumber
+\Delta C\approx -\epsilon \frac{f_C}{2 \bar{w}(1-2 d_0)^2} \bigg( 1-2 d_0 (-2 d_0+s+2) \\
+ +\sqrt{2} \sqrt{s (2 d_0 (d_0(s-2)+2)-1)} \bigg)
 	\label{coupling}
-\end{equation}
-%YB Considered this ugly math more impressive.
+\end{eqnarray}
+
 For all parameters sustaining a polymorphism at the drive locus ($s>d_0-1/2$), this corresponds to a decrease in frequency of the $C$ allele when it enhances drive ($\epsilon >0$ -- the $B^+$ model, above),
@@
 More generally, even when the cost of drive is not fully recessive, the $B^-$ allele will invade and fix under all parameters sustaining a previous polymorphism at the drive locus (see Mathematical Supplement).
-%When the C allele acts to suppress drive for all parameters sustaining a polymorphism at the drive locus ($s>d_0-1/2$)
-%We address two questions concerning the $C$ allele - can it invade when rare  and can it fix when common?
-%Although our analytical approximations (see our \gc{Mathematical supplement}) do not reveal simple inequalities like those above,
-%	graphical exploration of analytical results show that this mutant is lost when it enhances drive ($\epsilon>0$ - the $B^+$ allele, above),
-%	and spreads from rarity to fixation when it suppresses drive ($\epsilon<0$ - the $B^-$ allele, above).	
-%This result is supported by exact numerical iterations, as well
-%\yb{FIG?}. \gc{Lets not bother}
 
-\subsubsection*{Model 5. Drive-modifier in repulsion phase:}
+ \subsubsection{Model 5. Drive-modifier in repulsion phase:}
 When the $C$ allele is tightly linked to the non-driver, 
@@
 \begin{equation}
-	\Delta C \approx -\epsilon f_{C^-}^2 \frac{ \left(2 d_0 s-\sqrt{2} \sqrt{s (2 d_0 (d_0 (s-2)+2)-1)}\right)}{2\bar{w}s (2 d_0 -1) }. \label{A-}
+	\Delta C \approx -\epsilon f_{C^-}^2 \frac{ \left(2 d_0 s-\sqrt{2} \\ \sqrt{s (2 d_0 (d_0 (s-2)+2)-1)}\right)}{2\bar{w}s (2 d_0 -1) } \label{A-}
 \end{equation}
@@
 $\epsilon <0$, corresponding to the $A^-$ model, above), a result which holds qualitatively for a common $C$ allele, as well (Mathematical Supplement).
-%We address two questions concerning the $C$ allele - can it invade when rare  and can it fix when common?
-%Again, while simple inequalities are lacking, 
-%	graphical exploration of our analytical results, and exact numerical iteration \yb{FIG?} shows that this modifier only spreads from rarity to fixation when it suppresses drive ($\epsilon<0$ - the $A^-$ allele, above).
 Somewhat counterintuitively, the spread of the $A^-$ allele is substantially slower than that of the $B^-$ allele (it is on the order of $C^{2}$ rather than $C$). 
@@
 \subsubsection*{Model 6. Unlinked drive-modifier:}
-%\gc{Can we use something other than F, as that was HWE deviation}
 For the unlinked model, we introduce another locus where drive is modified in $A/B$ females fertilized by $M$ allele, 
@@
 \begin{equation} \Delta M =-\epsilon f_A f_B (f_B s + s_h (f_A - f_B) ) \frac{f_M}{\bar{w}}  \label{unlinked} \end{equation}
-% an $F$ allele increases in frequency when 
-%	$-\epsilon f_A f_B (f_B s + s_h (fA - fB) ) >0$, corresponding to any level of drive suppression ($\epsilon<0$) 
 Thus, a rare drive suppressor ($\epsilon<0$) will spread so long as the fitness cost of the driver does not display over- or under-dominance. 
-Notably, this result holds while assuming complete linkage equilibrium, and therefore sperm suppression of female meiotic drive is likely more powerful than more direct female-enforced drive suppressors  which usually spread due to the indirect benefit of avoiding drive haplotypes \citep[e.g.
-][]{Brandvain2012}.
-%The spread of this driver does not, however, depend on assumption of HWE, and linkage equilibrium - exact results resemble analytical approximations. 
-%\paragraph{three allele system}
-%$x_{AA}$, $x_{AB}$,$x_{AB'}$, $w_{BB'} x_{BB'}$, $w_{B'B'} x_{BB}$, $w_{B'B'} x_{B'B'}$
-%We denote the frequency of the three genotypes in the population at birth are
-%$x_{AA}$, $x_{AB}$, $x_{AB'}$, $x_{BB'}$, $x_{BB}$, $x_{B'B'}$. 
-%The mean relative fitness of the population is 
-%\begin{equation}
-%x_{AA}+x_{AB}+ x_{AB'}+ w_{BB'} +x_{BB'}+w_{B'B'}x_{BB} + w_{B'B'} x_{B'B'} 
-%\end{equation}
-%The frequency of the genotypes in meiotic-cells of female adults is 
-%\begin{equation}
-%\frac{1}{\bar{w}}\Big( x_{AA}, ~x_{AB},~ x_{AB'},~ w_{BB'} x_{BB'},~ w_{B'B'}x_{BB},
-%~w_{B'B'} x_{B'B'} \Big)
-%\end{equation}
-%The frequency of fertilization events can be found by combining these
-%female frequencies with the frequency of the three alleles in sperm is
-%\begin{eqnarray}
-%x_A = & x_{AA} + \frac{1}{2} x_{AB}+ \frac{1}{2} x_{AB'} \\
-%x_B = & w_{BB}  x_{BB} +   w_{BB'} \frac{1}{2} x_{BB'}+ \frac{1}{2} x_{AB} \\
-%x_{B'} = & w_{B'B'}  x_{B'B'} +   w_{BB'} \frac{1}{2} + x_{BB'}+ \frac{1}{2} x_{AB'}  
-%\end{eqnarray}
-%in combination with the transmission probabilities given in Figure
-%\ref{Eggsperm_3_allele_cartoon}A.
+Notably, this result holds while assuming complete linkage equilibrium, and therefore sperm suppression of female meiotic drive is likely more powerful than more direct female-enforced drive suppressors  which usually spread due to the indirect benefit of avoiding drive haplotypes (e.g.
+\cite{Brandvain2012}).
 
-% \gc{Alternatively we could give the table below:}
-
-% \begin{table}
-% \begin{tabular}{lllll}
-% \hline 
-%   &  \multicolumn{4}{c}{Sperm genotype}   \\
-%   &  & A & B & B'  \\
-% $\Venus$ geno. & Freq.  & $x_{AA} + \frac{1}{2} x_{AB}+ \frac{1}{2} x_{AB'} $ & $ w_{BB}  x_{BB} +   w_{BB'} \frac{1}{2} x_{BB'}+
-% \frac{1}{2} x_{AB}$ & $ w_{B'B'}  x_{B'B'} +   w_{BB'} \frac{1}{2}
-% x_{BB'}+ \frac{1}{2} x_{AB'}  $\\
-% \hline\\
-% AA & $x_{AA}$ &  & & \\
-% AB & $x_{AB}$ &  & & \\
-% AB' & $x_{AB'}$ &  & & \\
-% BB' & $w_{BB'} x_{BB'}$ &  & & \\
-% BB & $w_{B'B'} x_{BB}$ &  & & \\
-% B'B' & $w_{B'B'} x_{B'B'}$ &  & & \\
-% \end{tabular}
-% \caption{ALTERNATIVELY I COULD FILL IN THIS MATRIX  with genotype }
-% \end{table}
+\begin{acknowledgments} \end{acknowledgments}
 \bibliography{refs}
-
-\newpage
-
-  \setcounter{figure}{0}\global\def\thefigure{S\arabic{figure}}
-
-\section*{Supplementary Material}
-
-\begin{figure}
-
-\includegraphics[width = \textwidth]{../Scripts/homozyg_bistability.eps}
-\caption{Invasion analysis for a self-promoting female meiotic drive allele with
-  recessive costs (selection coefficient $s$), showing the region of
-  bistability. The colors, and the thin dashed contours, indicate the
-  frequency the allele must be above to invade the population (note
-  that these alleles reach fixation conditional on invading). In the
-  white area the allele cannot invade ($x \geq1$), in the solid red
-  area the allele can invade and fix when rare. In the left panel we
-  show the results obtained by a grid search using the recursion, on
-  the right we show the approximation obtained assuming that HWE
-  holds. }  
-\label{Bistab_homozyg_cost_fig}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width = 0.8 \textwidth]{Figures/bistable_x_vs_d_additive_s.eps} 
-\caption{The unstable equilibrium frequency for a self-promoting
-  female meiotic drive allele with an additive cost ($s_h=s/2$) as a
-  function of the drive parameter. The solid line shows results
-  obtained using the recursion, the dots our approximation given by Equation \eqref{thresh2}}  \label{bistable_additive}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width = 0.7 \textwidth]{../Scripts/effect_of_dominance_on_invasion_space_four_graph.eps} 
-\caption{Exact results of invasion analysis of an allele whose effect
- on female meiosis is mediated by the genotype of the fertilizing
- male.  
- A heterozygous female transmits the $B$ allele 
-  with probabilities  $\frac{1}{2}$,  $d_{het}=\frac{1}{2} + h(d_{hom}-\frac{1}{2}) $ or $d_{hom}$, 
- if she mated with a n $AA$, $AB$, or $BB$ male,  respectively.  
- The allele suffers a recessive fitness cost $s$.  
- The four panels correspond to different dominance relationships.
-In the parameter space below the invasion (solid) line the self-promoter
- driver can invade. In the parameter space below the fixation (long
- dash) line the self-promoter can fix. In the last
- two panels the invasion line is above the fixation line and so the
- allele can be maintained as a polymorphism in that thin slice of
- parameter space between the two lines. 
- In the final panel we show the
-  fixation line (small dashes) as predicted by our HWE approximation
-  ($(d_{het}-1/2)/d_{het}$) see the appendix for more details.
-}  \label{Effect_of_dominance}
-\end{figure}
+\end{article}
 
 \end{document}
+
+Figure 1 {Eggsperm}
+Figure 2 {Eggsperm_3_allele_cartoon}
+Figure S1 {Bistab_homozyg_cost_fig}
+Figure S2 {bistable_additive}
+Figure S3 {Effect_of_dominance}

### 2014-05-20 writeup_and_notes/signif_statement.tex
fixed minor tyos in significance statement and push PNAS folder

@@
 Female meiosis is the time during development where an allele can act to directly modify its representation in the next generation. 
-The potential benefits to allele what succeed in this competition and collateral damage to organismal fitness are thought to have shaped basic features of female reproduction. 
+The potential benefits to an allele that succeeds in this competition and collateral damage to organismal fitness are thought to have shaped basic features of female reproduction. 
 However, little attention has been directed at the odd fact that fertilization by sperm is required to complete female meiosis in many species of animal. 

### 2014-05-20 writeup_and_notes/male_dependent_drive.tex
small changes throughout 1st 1/2 of methods, all marked with gc

@@
 of a self-promoting meiotic driver,  we modify this standard model such 
-	that the driver is only effective when fertilized by a sperm carrying that allele (see Figure \ref{Eggsperm}A and Model 2 in the Supplementary Worksheet). 
+	that the driver is only effective when fertilized by a sperm
+        carrying that allele (see Figure \ref{Eggsperm}A and Model 2
+        in the \gc{Appendix and} Supplementary Worksheet). 
 We then identify the conditions allowing for the spread of this self-promoting driver, 
 	and evaluate whether a driver of this form could generate a sustained conflict favoring the evolution of suppressors. 
-We conclude with an analysis of a related model (Model 3) - in which
+We conclude \gc{our single locus results} with an analysis of a related model (Model 3) - in which
 drivers influence their transmission in females via paternal genotype,
@@
 
-For comparative purposes, we briefly present the standard drive model %as Model 1 in the Appendix, 
+For comparative purposes, we \gc{first} briefly present the standard drive model %as Model 1 in the Appendix, 
 %equations \ref{eqn.A1}-\ref{eqn.A3}, 
@@
 	it always invades because,  when rare, it occurs predominantly in heterozygotes and therefore drives without a fitness cost. 
-However, when $s$ is large ($s>(2d-1)/(2)$), a driver cannot fix, and
+However, when $s$ is large ($s>(2d-1)/(2)$, \gc{solid black line in Figure \ref{Eggsperm}B}), a driver cannot fix, and
 will be maintained as a protected polymorphism \citep{Prout1973}. 
@@
 	%, the selection against the drive homozygote, 
-	is less than $(2 d - 1)/(4 d)$. 
+	is \gc{approximately} less than $(2 d - 1)/(4 d)$ ( \gc{dashed black line in Figure \ref{Eggsperm}B}). 
 This analytical approximation, derived from Equation \eqref{deltadriver} assuming Hardy-Weinberg, 
@@
 	diploid genotype (perhaps due to proteins and RNAs packaged into the sperm), rather than sperm haplotype.
-This paternal-genotype dependent model (Model 3) requires one additional parameter, as we exchange $d$ in the sperm dependent case for $d_{het}$ and $d_{hom}$ which describe the transmission of the drive allele in a heterozygous female mating with males heterozygous and homozygous for the self-promoting drive allele, respectively.
+This paternal-genotype dependent model (Model 3 \gc{in the appendix}) requires one additional parameter, as we exchange $d$ in the sperm dependent case for $d_{het}$ and $d_{hom}$ which describe the transmission of the drive allele in a heterozygous female mating with males heterozygous and homozygous for the self-promoting drive allele, respectively.
 Here, a rare driver invades when $s$ is less than $(d_{het}-1/2)/d_{het}$, and usually fixes when it invades.

### 2014-05-20 writeup_and_notes/male_dependent_drive.tex
small changes throughout 2nd 1/2 of result, all marked with gc

@@
 	initially reaches drive-viability equilibrium (with two alleles
-	$A$ and $B$ are the ancestral non-driver and driver alleles). 
+	$A$ and $B$ are the ancestral non-driver and driver alleles\gc{, Figure \ref{Eggsperm_3_allele_cartoon}A1}). 
 Subsequently, a sperm-acting modifier of female meiotic drive arises at another locus.
@@
 The $B^+$ allele acts in sperm to increase the effectiveness of drive for both
-	the $B$ and  $B^{+}$ alleles in $AB$ and $AB^{+}$ heterozygotes (see Figure \ref{Eggsperm_3_allele_cartoon}, and Model 4 in the Supplementary Worksheet). 
+	the $B$ and  $B^{+}$ alleles in $AB$ and $AB^{+}$
+        heterozygotes (see Figure \ref{Eggsperm_3_allele_cartoon}A2, and
+        Model 4 in the \gc{Appendix and} Supplementary Worksheet). 
 Naively, $B^{+}$ may spread by capitalizing on the additional drive it causes; however,  
@@
 	the drive background, i.e. in repulsion phase, creating a
-        third allele $B^{-}$, Model 4. This new allele still
+        third allele $B^{-}$ \gc{(Figure
+          \ref{Eggsperm_3_allele_cartoon}A3, Model 4 in Appendix)}. This new allele still
         experiences female drive when fertilized by A or B sperm, but 
@@
      allows the $B^{-}$ to displace the ancestral driver (Figure
-     \ref{Eggsperm_3_allele_cartoon}, Equation \ref{coupling}), 
+     \ref{Eggsperm_3_allele_cartoon}\gc{B1}, Equation \ref{coupling}), 
 	and often returns to a lower equilibrium frequency than the the $B$ allele, further decreasing the extent of drive in the population.  
 If this sperm-acting drive suppressor arises on 
-	the non-driving $A$ background (creating a third allele $A^{-}$, Model 5), 
+	the non-driving $A$ background (creating a third allele $A^{-}$, Figure \ref{Eggsperm_3_allele_cartoon}A4, Model 5), 
 	or is unlinked to the drive locus (Model 6), it readily invades a population segregating
-	for the drive system (Figure \ref{Eggsperm_3_allele_cartoon} and Equations \ref{A-} and \ref{unlinked}).  
+	for the drive system (\gc{Equations} \ref{A-} and
+        \ref{unlinked}).  %%DELETED FIG REF HERE.
 This allele lowers the frequency of the original driver (perhaps to zero),
 	and spreads to fixation if it does not carry strong fitness costs
-	(Figure \ref{Eggsperm_3_allele_cartoon} and \ref{Trajectories_of_supressors}A). 
+	\gc{(Figure \ref{Eggsperm_3_allele_cartoon}B2)}. 
 This result is consistent with previous work showing that drive suppressors unlinked to, or in repulsion phase with drivers usually invade polymorphic drive systems \citep[e.g. ][]{Brandvain2012}.  

### 2014-05-20 writeup_and_notes/male_dependent_drive.tex
incorporated coops comments

--- 2014-05-20 writeup_and_notes/male_dependent_drive.tex
@@
 of a self-promoting meiotic driver,  we modify this standard model such 
-	that the driver is only effective when fertilized by a sperm carrying that allele (see Figure \ref{Eggsperm}A and Model 2 in the Supplementary Worksheet). 
+	that the driver is only effective when fertilized by a sperm
+        carrying that allele (see Figure \ref{Eggsperm}A and Model 2
+        in the Appendix and Supplementary Worksheet). 
 We then identify the conditions allowing for the spread of this self-promoting driver, 
 	and evaluate whether a driver of this form could generate a sustained conflict favoring the evolution of suppressors. 
-We conclude with an analysis of a related model (Model 3) - in which
+We conclude our single locus results with an analysis of a related model (Model 3) - in which
 drivers influence their transmission in females via paternal genotype,
@@
 
-For comparative purposes, we briefly present the standard drive model %as Model 1 in the Appendix, 
+For comparative purposes, we first briefly present the standard drive model %as Model 1 in the Appendix, 
 %equations \ref{eqn.A1}-\ref{eqn.A3}, 
@@
 	it always invades because,  when rare, it occurs predominantly in heterozygotes and therefore drives without a fitness cost. 
-However, when $s$ is large ($s>(2d-1)/(2)$), a driver cannot fix, and
+However, when $s$ is large ($s>(2d-1)/(2)$, solid black line in Figure \ref{Eggsperm}B), a driver cannot fix, and
 will be maintained as a protected polymorphism \citep{Prout1973}. 
@@
 	%, the selection against the drive homozygote, 
-	is less than $(2 d - 1)/(4 d)$. 
+	is approximately less than $(2 d - 1)/(4 d)$ -- see dashed black line in Figure \ref{Eggsperm}B). 
 This analytical approximation, derived from Equation \eqref{deltadriver} assuming Hardy-Weinberg, 
@@
 	diploid genotype (perhaps due to proteins and RNAs packaged into the sperm), rather than sperm haplotype.
-This paternal-genotype dependent model (Model 3) requires one additional parameter, as we exchange $d$ in the sperm dependent case for $d_{het}$ and $d_{hom}$ which describe the transmission of the drive allele in a heterozygous female mating with males heterozygous and homozygous for the self-promoting drive allele, respectively.
+This paternal-genotype dependent model (Model 3 in the Appendix) requires one additional parameter, as we exchange $d$ in the sperm dependent case for $d_{het}$ and $d_{hom}$ which describe the transmission of the drive allele in a heterozygous female mating with males heterozygous and homozygous for the self-promoting drive allele, respectively.
 Here, a rare driver invades when $s$ is less than $(d_{het}-1/2)/d_{het}$, and usually fixes when it invades.
@@
 	initially reaches drive-viability equilibrium (with two alleles
-	$A$ and $B$ are the ancestral non-driver and driver alleles). 
+	$A$ and $B$ are the ancestral non-driver and driver alleles, Figure \ref{Eggsperm_3_allele_cartoon}A1). 
 Subsequently, a sperm-acting modifier of female meiotic drive arises at another locus.
@@
 The $B^+$ allele acts in sperm to increase the effectiveness of drive for both
-	the $B$ and  $B^{+}$ alleles in $AB$ and $AB^{+}$ heterozygotes (see Figure \ref{Eggsperm_3_allele_cartoon}, and Model 4 in the Supplementary Worksheet). 
+	the $B$ and  $B^{+}$ alleles in $AB$ and $AB^{+}$
+        heterozygotes (see Figure \ref{Eggsperm_3_allele_cartoon}A2, and
+        Model 4 in the Appendix and Supplementary Worksheet). 
 Naively, $B^{+}$ may spread by capitalizing on the additional drive it causes; however,  
@@
 	the drive background, i.e. in repulsion phase, creating a
-        third allele $B^{-}$, Model 4. This new allele still
+        third allele $B^{-}$ (Figure
+          \ref{Eggsperm_3_allele_cartoon}A3, Model 4 in Appendix). This new allele still
         experiences female drive when fertilized by A or B sperm, but 
@@
      allows the $B^{-}$ to displace the ancestral driver (Figure
-     \ref{Eggsperm_3_allele_cartoon}, Equation \ref{coupling}), 
+     \ref{Eggsperm_3_allele_cartoon}B1, Equation \ref{coupling}), 
 	and often returns to a lower equilibrium frequency than the the $B$ allele, further decreasing the extent of drive in the population.  
 If this sperm-acting drive suppressor arises on 
-	the non-driving $A$ background (creating a third allele $A^{-}$, Model 5), 
+	the non-driving $A$ background (creating a third allele $A^{-}$, Figure \ref{Eggsperm_3_allele_cartoon}A4, Model 5), 
 	or is unlinked to the drive locus (Model 6), it readily invades a population segregating
-	for the drive system (Figure \ref{Eggsperm_3_allele_cartoon} and Equations \ref{A-} and \ref{unlinked}).  
+	for the drive system (Equations \ref{A-} and
+        \ref{unlinked}).  %%DELETED FIG REF HERE.
 This allele lowers the frequency of the original driver (perhaps to zero),
 	and spreads to fixation if it does not carry strong fitness costs
-	(Figure \ref{Eggsperm_3_allele_cartoon} and \ref{Trajectories_of_supressors}A). 
+	(Figure \ref{Eggsperm_3_allele_cartoon}B2). 
 This result is consistent with previous work showing that drive suppressors unlinked to, or in repulsion phase with drivers usually invade polymorphic drive systems \citep[e.g. ][]{Brandvain2012}.  

### 2014-05-20 writeup_and_notes/pnasSubmission/PNASSubmission.TEX
incorporated coops comments

@@
 To depict a case of a self-promoting meiotic driver,  we modify this standard model such 
-	that the driver is only effective when fertilized by a sperm carrying that allele (see Figure 1A and Model 2 	in the Supplementary Worksheet). 
+	that the driver is only effective when fertilized by a sperm carrying that allele (see Figure 1A and Model 2 	in the Appendix and Supplementary Worksheet). 
 We then identify the conditions allowing for the spread of this self-promoting driver, 
 	and evaluate whether a driver of this form could generate a sustained conflict favoring the evolution of suppressors. 
-We conclude with an analysis of a related model (Model 3) - in which
+We conclude our single locus results with an analysis of a related model (Model 3) - in which
 drivers influence their transmission in females via paternal genotype,
@@
 
-For comparative purposes, we briefly present the standard drive model 
+For comparative purposes, we first briefly present the standard drive model 
 	(see e.g. \cite{Prout1973,Ubeda2004} for additional results). 
@@
 	it always invades because, when rare, it occurs predominantly in heterozygotes and therefore drives without a fitness cost. 
-However, when $s$ is large ($s>(2d-1)/(2)$), a driver cannot fix, and
+However, when $s$ is large ($s>(2d-1)/(2)$, solid black line in Figure 1B), a driver cannot fix, and
 will be maintained as a protected polymorphism \cite{Prout1973}. 
@@
 When rare, a fully recessive, self-promoting driver can only invade when $s$ 
-	is less than $(2 d - 1)/(4 d)$. 
+	is approximately less than $(2 d - 1)/(4 d)$ -- see dashed black line in Figure 1B. 
 This analytical approximation, derived from Equation \eqref{deltadriver} assuming Hardy-Weinberg, 
@@
 	diploid genotype (perhaps due to proteins and RNAs packaged into the sperm), rather than sperm haplotype.
-This paternal-genotype dependent model (Model 3) requires one additional parameter, as we exchange $d$ in the sperm dependent case for $d_{het}$ and $d_{hom}$ which describe the transmission of the drive allele in a heterozygous female mating with males heterozygous and homozygous for the self-promoting drive allele, respectively.
+This paternal-genotype dependent model (Model 3  in the Appendix) requires one additional parameter, as we exchange $d$ in the sperm dependent case for $d_{het}$ and $d_{hom}$ which describe the transmission of the drive allele in a heterozygous female mating with males heterozygous and homozygous for the self-promoting drive allele, respectively.
 Here, a rare driver invades when $s$ is less than $(d_{het}-1/2)/d_{het}$, and usually fixes when it invades.
@@
 	initially reaches drive-viability equilibrium (with two alleles
-	$A$ and $B$ are the ancestral non-driver and driver alleles). 
+	$A$ and $B$ are the ancestral non-driver and driver alleles, Figure 2A1). 
 Subsequently, a sperm-acting modifier of female meiotic drive arises at another locus.
@@
 The $B^+$ allele acts in sperm to increase the effectiveness of drive for both
-	the $B$ and  $B^{+}$ alleles in $AB$ and $AB^{+}$ heterozygotes (see Figure 2, and Model 4 in the Supplementary Worksheet). 
+	the $B$ and  $B^{+}$ alleles in $AB$ and $AB^{+}$ heterozygotes (see Figure 2A2, and Model 4 in the Appendix and Supplementary Worksheet). 
 Naively, $B^{+}$ may spread by capitalizing on the additional drive it causes; however,  
@@
 	the drive background, i.e. in repulsion phase, creating a
-        third allele $B^{-}$, Model 4. This new allele still
+        third allele $B^{-}$, (Figure 2A3, Model 4 in Appendix). 
+ This new allele still
         experiences female drive when fertilized by A or B sperm, but 
@@
 	or is unlinked to the drive locus (Model 6), it readily invades a population segregating
-	for the drive system (Figure 2 and Equations \ref{A-} and \ref{unlinked}).  
+	for the drive system (Equations \ref{A-} and \ref{unlinked}).  
 This allele lowers the frequency of the original driver (perhaps to zero),
 	and spreads to fixation if it does not carry strong fitness costs
-	(Figure 2). 
+	(Figure 2B2). 
 This result is consistent with previous work showing that drive suppressors unlinked to, or in repulsion phase with drivers usually invade polymorphic drive systems (e.g. \cite{Brandvain2012}).  

### 2014-05-20 writeup_and_notes/male_dependent_drive.tex
legend for fig 2 other minor rearangement

@@
 
+\begin{figure}
+\includegraphics[width = 1 \textwidth]{Figures/selfprom.eps}
+\caption{{\bf{A.}} A visual depiction of our model of `self-promoting' driver. 
+	Transmission probabilities for alleles through female
+ 	 meiosis depend on sperm genotype. 
+	 The non-driving $A$-allele, and self-promoting $B$-allele
+	 are represented by unfilled and filled circles, respectively. 
+	 {\bf{B.}} Evolution of a self-promoter and standard driver. 
+	 	Assuming that the fitnesses of drive homozygotes and 
+		heterozygotes are $1-s$ and $1$, respectively. 
+	 	\emph{Main figure:} Boundary conditions for the invasion 
+		and fixation of self-promoting and standard meiotic drivers, 
+		with drive coefficient, $d$. Colored regions depict exact results, 
+		while lines  represent analytical approximations. 
+	 \emph{B1:}  Trajectories of sperm-dependent female drive
+		 %alleles that can invade the population, 
+		 each allele has $s =0.1$ against the
+		 homozygotes.  The drive coefficient is denoted by color. 
+ }   
+\label{Eggsperm}
+\end{figure}
+
 When a self-promoting driver does spread, 
@@
 
-\begin{figure}
-\includegraphics[width = 0.8 \textwidth]{Figures/selfprom.eps}
-\caption{{\bf{A.}} A visual depiction of our model of 'self-promoting' driver. 
-	Transmission probabilities for alleles through female
- 	 meiosis depend on sperm genotype. 
-	 The non-driving $A$-allele, and self-promoting $B$-allele
-	 are represented by unfilled and filled circles, respectively. 
-	 {\bf{B.}} Evolution of a self-promoter and standard driver. 
-	 	Assuming that the fitnesses of drive homozygotes and 
-		heterozygotes are $1-s$ and $1$, respectively. 
-	 	\emph{Main figure:} Boundary conditions for the invasion 
-		and fixation of self-promoting and standard meiotic drivers, 
-		with drive coefficient, $d$. Colored regions depict exact results, 
-		while lines  represent analytical approximations. 
-	 \emph{B1:}  Trajectories of sperm-dependent female drive
-		 %alleles that can invade the population, 
-		 each allele has $s =0.1$ against the
-		 homozygotes.  The drive coefficient is denoted by color. 
- }   
-\label{Eggsperm}
-\end{figure}
-
 %\begin{figure}
@@
 %\subsection*{ sperm-egg haplotype balanced drive systems}
-
-\begin{figure}
-	\rotatebox{90}{\includegraphics[width = 0.8
-          \textwidth]{Figures/Sperm_egg_cartoon_paper_Fig2.eps}}
-\caption{Dynamics in 3 allele system
-{\bf (A) } The sperm-dependent transmission probabilities of alleles through
-female meiosis (female germ cells depicted as circles, sperm indicated by wiggly
-tails). In each case the top and bottom entry in each cell give the
-transmission probability of the left and right allele in the female
-being transmitted to the egg. In the left panel we show the transmission probabilities for
-  A, B, and B'. Where B' is a driver allele which is either promotes
-  itself through through female meiosis when sperm (a B$^+$ allele,
-  $d'>d$) or self-avoiding through the interaction of sperm with female meiosis (a B$^-$ allele,
-$d'<d$). In the right panel we depict the transmission probabilities for
-  A, B, and A$^-$ allele a non-driven allele that acts to reduce the
-  effects of female drive through sperm.
-{\bf (B)} Frequency trajectories in the three allele system. }
- \label{Eggsperm_3_allele_cartoon}
-\end{figure}
 
@@
 	invade the population (Appendix Model 4, eqn. \eqref{coupling}, and Mathematical Supplement).
+
+\begin{figure}
+	\includegraphics[width = 0.7
+          \textwidth]{Figures/twolocus.eps}
+\caption{Models of a sperm-acting drive modifier tightly linked to a meiotic driver. 
+{\bf (A) } Loci exist in tightly linked haplotypes (top of A). 
+	Genotype of sperm at the modifier locus (ancestral and sperm acting modifier empty and filled squares, respectively) alters transmission  
+	at the drive locus (circles) during female meiosis. 
+	{\bf{A1)}} In the standard model of drive, 
+	the driver (filled black circle) and non-driver (empty circle) are transmitted with probabilities $d_0$ 
+	and $1-d_0$, respectively. 
+	{\bf{A2)}} The $B^+$ haplotype increases the transmission of the drive allele ($d_1 > d_0$), and due to their shared genetic background, also increases its drive. 
+	{\bf{A3 \& A4)}} The sperm-acting modifier acts to decrease drive  ($d_1=1/2$ in A3 \& A4. More generally, $d_1 < d_0$). 
+		In {\bf{A3}} the drive suppressor exists on the drive background, forming the $B^-$ haplotype. 
+		In {\bf{A4}} the drive suppressor exists on the non-drive background, forming the $A^-$ haplotype. 		
+{\bf{(B)}} Invasion of a sperm-acting drive suppressor linked to a driver. 
+	After the driver ($B$ haplotype) reaches drive selection equilibrium, we introduce a sperm acting drive modifier. 
+	We assume full drive ($D_0 =1$), a recessive lethal fitness cost to drive ($h_s=0$, $s=1$) and that the sperm-acting modifier results in a fair meiosis. 
+{\bf {B1)}} The $B^-$ allele replaces the ancestral drive haplotype, but segregates at a lower equilibrium frequency.
+{\bf {B2)}} The $A^-$ allele replaces the ancestral non-driving haplotype, and in this case, removes the driver from the population. }
+ \label{Eggsperm_3_allele_cartoon}
+\end{figure}
 

### 2014-05-20 writeup_and_notes/male_dependent_drive.tex
rewrote caption 23, and centered figures

@@
 \begin{figure}
-\includegraphics[width = 1 \textwidth]{Figures/selfprom.eps}
+\begin{center}
+\includegraphics[width = 1 \textwidth]{Figures/selfprom.eps} \end{center}
 \caption{{\bf{A.}} A visual depiction of our model of `self-promoting' driver. 
@@
 \begin{figure}
+\begin{center}
 	\includegraphics[width = 0.7
-          \textwidth]{Figures/twolocus.eps}
+          \textwidth]{Figures/twolocus.eps} \end{center}
 \caption{Models of a sperm-acting drive modifier tightly linked to a meiotic driver. 
-{\bf (A) } Loci exist in tightly linked haplotypes (top of A). 
-	Genotype of sperm at the modifier locus (ancestral and sperm acting modifier empty and filled squares, respectively) alters transmission  
-	at the drive locus (circles) during female meiosis. 
-	{\bf{A1)}} In the standard model of drive, 
-	the driver (filled black circle) and non-driver (empty circle) are transmitted with probabilities $d_0$ 
-	and $1-d_0$, respectively. 
-	{\bf{A2)}} The $B^+$ haplotype increases the transmission of the drive allele ($d_1 > d_0$), and due to their shared genetic background, also increases its drive. 
-	{\bf{A3 \& A4)}} The sperm-acting modifier acts to decrease drive  ($d_1=1/2$ in A3 \& A4. More generally, $d_1 < d_0$). 
-		In {\bf{A3}} the drive suppressor exists on the drive background, forming the $B^-$ haplotype. 
-		In {\bf{A4}} the drive suppressor exists on the non-drive background, forming the $A^-$ haplotype. 		
+{\bf (A) } 
+Sperm carrying the derived allele at the modifier locus (filled squares) alters transmission  
+	at the driving allele (filled circles) during female meiosis.
+        Alleles at these two tightly linked loci form three haplotypes (top of A). 	
+	{\bf{A1)}} In the standard model of drive there is no
+        variation at the modifier, 
+	and the driver is transmitted to the egg with probability $d_0$ 
+	{\bf{A2)}} The modifier allele increases the transmission of the drive allele ($d_1 > d_0$), and due to their shared genetic background, also increases its drive. 
+	{\bf{A3 \& A4)}} The sperm-acting modifier acts to decrease
+        drive  ($d_1=1/2$ or more generally, $d_1 < d_0$) and arises on the same or opposite
+        background from the driver (A3 \& A4 respectively). 
+%		In {\bf{A3}} the drive suppressor exists on the drive background, forming the $B^-$ haplotype. 
+%		In {\bf{A4}} the drive suppressor exists on the non-drive background, forming the $A^-$ haplotype. 		
 {\bf{(B)}} Invasion of a sperm-acting drive suppressor linked to a driver. 

### 2014-05-20 writeup_and_notes/male_dependent_drive.tex
yb minor mods to legend for fig 2

@@
         variation at the modifier, 
-	and the driver is transmitted to the egg with probability $d_0$ 
+	and the driver is transmitted to the egg with probability $d_0$. 
 	{\bf{A2)}} The modifier allele increases the transmission of the drive allele ($d_1 > d_0$), and due to their shared genetic background, also increases its drive. 
 	{\bf{A3 \& A4)}} The sperm-acting modifier acts to decrease
-        drive  ($d_1=1/2$ or more generally, $d_1 < d_0$) and arises on the same or opposite
+        drive  ($d_1=1/2$ in A3 \& A4, or more generally, $d_1 < d_0$) and arises on the same or opposite
         background from the driver (A3 \& A4 respectively). 

### 2014-05-21 writeup_and_notes/pnasSubmission/PNASSubmission.TEX
changed PNAS files to [hopefully] comply with instructions

@@
 
-%\abbreviations{SAM, self-assembled monolayer; OTS, octadecyltrichlorosilane}
-
 \dropcap{D}espite the apparent unity of the organism, `selfish' alleles can
@@
 	likely precludes this interaction in plants). 
-Therefore, a `green-bearded' \cite{Gardner2010} sperm-acting allele that recognizes and facilitates the meiotic drive of a genetically equivalent allele in heterozygous females %female meiosis %(a  heteromorphic dyad),
+Therefore, a `green-bearded' \cite{Gardner2010} sperm-acting allele that recognizes and facilitates the meiotic drive of a genetically equivalent allele in heterozygous females 
      could presumably rapidly spread through a population. 
@@
 To depict a case of a self-promoting meiotic driver,  we modify this standard model such 
-	that the driver is only effective when fertilized by a sperm carrying that allele (see Figure 1A and Model 2 	in the Appendix and Supplementary Worksheet). 
+	that the driver is only effective when fertilized by a sperm carrying that allele (see Figure 1A and Model 2 	in the Appendix and Table S1). 
 We then identify the conditions allowing for the spread of this self-promoting driver, 
@@
 	(approximated in Equation \ref{bistab_homozy_approx} and  presented exactly in 
-	Figure S1) and lost otherwise. 
+	Figure S1), and lost otherwise. 
 For most parameters, this threshold is likely too high to be reached by drift, 
@@
 Models 2 and 3, above, explored the dynamics of an allele that drove in females when signaled by a complementary signal in sperm.   
-We complement this single-locus approach %of an allele with pleiotropic effects in both sexes, 
+We complement this single-locus approach 
 	with alternative models of two loci - one a female driver, 
@@
 The $B^+$ allele acts in sperm to increase the effectiveness of drive for both
-	the $B$ and  $B^{+}$ alleles in $AB$ and $AB^{+}$ heterozygotes (see Figure 2A2, and Model 4 in the Appendix and Supplementary Worksheet). 
+	the $B$ and  $B^{+}$ alleles in $AB$ and $AB^{+}$ heterozygotes (see Figure 2A2, and Model 4 in the Appendix and Table S1). 
 Naively, $B^{+}$ may spread by capitalizing on the additional drive it causes; however,  
@@
 	three allele system that allows the sperm-acting drive facilitator $B^{+}$ to
-	invade the population (Appendix Model 4, Equation \eqref{coupling}, and Mathematical Supplement).
+	invade the population (Appendix Model 4, Equation \eqref{coupling}, and Appendix S1).
 
@@
 Transmission rules describing the outcomes of all matings in each
-	model are presented in a Supplementary Worksheet. 
+	model are presented in Table S1. 
 The fitness of genotype, $g$ is sex-independent and equals $1$, $1-h_s$, and $1-s$ for genotypes $AA$, $AB$, and $BB$, respectively. 
@@
 Because many of our analyses assume genotype frequencies expected under Hardy Weinberg Equilibrium (HWE), we note which analytical results are approximations. 
-We verify these approximations with exact numerical iterations in Supplementary Figures and Supplementary Scripts, and detail the derivation of these approximations in our Mathematical Supplement.
+We verify these approximations with exact numerical iterations in Figures S1-S3 , and detail the derivation of these approximations in Appendix S1.
 
@@
  but it will be lost when introduced below this frequency. 
- Compared to exact results (Supplementary Figure S1), Equations
+ Compared to exact results (Figure S1), Equations
  \eqref{ineqA} and \eqref{bistab_homozy_approx} offer reasonable, but imperfect approximations.
@@
 Comparison of Equation \ref{thresh2} to exact results obtained by a
-simple parameter search (Supplementary Figure S2)
+simple parameter search (Figure S2)
 show that this approximation is reasonably correct for small
@@
 	$s\lessapprox (d_{het}-1/2)/(d_{het})$, and when common, this driver fixes if 
-	$s\lessapprox d_{hom}-1/2$, approximations well supported by exact results (Supplementary Figure S3).
+	$s\lessapprox d_{hom}-1/2$, approximations well supported by exact results (Figure S3).
 Specifically, when drive in heterozygotes is large relative to that in homozygotes, 
@@
 	and an increase in frequency of the $C$ allele when it suppresses drive ($\epsilon <0$ -- the $B^-$ model, above). 
-More generally, even when the cost of drive is not fully recessive, the $B^-$ allele will invade and fix under all parameters sustaining a previous polymorphism at the drive locus (see Mathematical Supplement).
+More generally, even when the cost of drive is not fully recessive, the $B^-$ allele will invade and fix under all parameters sustaining a previous polymorphism at the drive locus (see Appendix S1).
 
@@
 For all values of interest ($0<s<1$, $0.5<d_0<1$), the change in frequency a rare $C$ allele is positive when it decreases drive (i.e.
-$\epsilon <0$, corresponding to the $A^-$ model, above), a result which holds qualitatively for a common $C$ allele, as well (Mathematical Supplement).
+$\epsilon <0$, corresponding to the $A^-$ model, above), a result which holds qualitatively for a common $C$ allele, as well (Appendix S1).
 Somewhat counterintuitively, the spread of the $A^-$ allele is substantially slower than that of the $B^-$ allele (it is on the order of $C^{2}$ rather than $C$). 
@@
 \begin{acknowledgments} \end{acknowledgments}
-\bibliography{refs}
+%\bibliography{refs}
+\input{PNASSubmission.bbl}
 \end{article}
+
+\begin{figure*}
+\includegraphics{selfprom.eps}
+\caption{{\bf{A.}} A visual depiction of our model of `self-promoting' driver. 
+	Transmission probabilities for alleles through female
+ 	 meiosis depend on sperm genotype. 
+	 The non-driving $A$-allele, and self-promoting $B$-allele
+	 are represented by unfilled and filled circles, respectively. 
+	 {\bf{B.}} Evolution of a self-promoter and standard driver. 
+	 	Assuming that the fitnesses of drive homozygotes and 
+		heterozygotes are $1-s$ and $1$, respectively. 
+	 	\emph{Main figure:} Boundary conditions for the invasion 
+		and fixation of self-promoting and standard meiotic drivers, 
+		with drive coefficient, $d$. Colored regions depict exact results, 
+		while lines  represent analytical approximations. 
+	 \emph{B1:}  Trajectories of sperm-dependent female drive
+		 %alleles that can invade the population, 
+		 each allele has $s =0.1$ against the
+		 homozygotes.  The drive coefficient is denoted by color. 
+ }   
+\label{Eggsperm}
+\end{figure*}
+
+\begin{figure*}
+	\includegraphics[width = .95
+          \textwidth]{twolocus.pdf} 
+\caption{Models of a sperm-acting drive modifier tightly linked to a meiotic driver. 
+{\bf (A) } 
+Sperm carrying the derived allele at the modifier locus (filled squares) alters transmission  
+	at the driving allele (filled circles) during female meiosis.
+        Alleles at these two tightly linked loci form three haplotypes (top of A). 	
+	{\bf{A1)}} In the standard model of drive there is no
+        variation at the modifier, 
+	and the driver is transmitted to the egg with probability $d_0$. 
+	{\bf{A2)}} The modifier allele increases the transmission of the drive allele ($d_1 > d_0$), and due to their shared genetic background, also increases its drive. 
+	{\bf{A3 \& A4)}} The sperm-acting modifier acts to decrease
+        drive  ($d_1=1/2$ in A3 \& A4, or more generally, $d_1 < d_0$) and arises on the same or opposite
+        background from the driver (A3 \& A4 respectively). 
+{\bf{(B)}} Invasion of a sperm-acting drive suppressor linked to a driver. 
+	After the driver ($B$ haplotype) reaches drive selection equilibrium, we introduce a sperm acting drive modifier. 
+	We assume full drive ($D_0 =1$), a recessive lethal fitness cost to drive ($h_s=0$, $s=1$) and that the sperm-acting modifier results in a fair meiosis. 
+{\bf {B1)}} The $B^-$ allele replaces the ancestral drive haplotype, but segregates at a lower equilibrium frequency.
+{\bf {B2)}} The $A^-$ allele replaces the ancestral non-driving haplotype, and in this case, removes the driver from the population. }
+ \label{Eggsperm_3_allele_cartoon}
+\end{figure*}
 

### 2014-05-21 writeup_and_notes/pnasSubmission/Text_S1.tex
added text S1.tex [a description of other supp files for PNAS submission]

--- 2014-05-20 writeup_and_notes/male_dependent_drive.tex
@@
-%%!TEX TS-program = latex
 \documentclass[12pt,letterpaper]{article}
@@
 \usepackage{fullpage}
-\bibliographystyle{genetics}
-\usepackage{multirow}
-\newcommand{\ye}{\hat{y}}
-\newcommand{\xe}{\hat{x}}
-\usepackage{color}
-\usepackage[normalem]{ulem}  
-	\newcommand{\gc}[1]{{ \color{red} #1}}
-	\newcommand{\yb}[1]{{ \color{blue} #1}}
 
-%Why do animals allow male control of female meiosis?
-%Why does female meiosis allow the opportunity for exploitation by selfish sperm?
-%sperm and egg genomes both favour the peaceful resolution of female meiotic drive
-%strange bed fellows
-
-%\title{Why does female meiosis allow the opportunity for exploitation by selfish sperm? BETTER TITLE?}
-\title{
-%Co-operation between the sexes trumps selfish control of female  meiotic drive. 
-%Or 
-Sperm should evolve to make female meiosis fair.
-%Sperm do not evolve to collaborate in female meiotic drive. 
-%Or Why does female meiosis allow the opportunity for exploitation by selfish sperm? 
-}
-
-\author{Yaniv Brandvain$^1$ \\ email: ybrandvain@gmail.com  \and Graham Coop$^2$ \\ email: gmcoop@ucdavis.edu }
-
-\usepackage{fancyhdr}
-\pagestyle{fancy}
-\lhead{}
-\rhead{}
-\renewcommand{\headrulewidth}{0.0pt}
-\rfoot{}
-\cfoot{\thepage}
-\date{}
-\bibliographystyle{plain}
 \begin{document} 
-\maketitle
 \begin{center}
-$^1$Department of Plant Biology, \\ University of Minnesota --€“ Twin
-Cities. St. Paul MN, 55108 \\
-$^2$Center for Population Biology \& Department of Evolution and Ecology \\ University of California -- Davis. Davis, CA, 95616
+\Large{{\bf{Text S1 of Brandvain \& Coop. \\ ``Sperm should evolve to make female meiosis fair.''}}}
 \end{center}
 
-%FIgures 
-{\bf Abstract:}
-Genomic conflicts arise when an allele gains an evolutionary advantage at a cost to organismal fitness. 
-%Because only one of four products of female meiosis is transmitted to offspring, 
-O\"{o}genesis is inherently susceptible to such conflicts
-because alleles compete to be the  product of female meiosis transmitted to the egg. 
-%(meiotic drive).  
-Alleles that distort meiosis in their favor (i.e. meiotic drivers) often decrease organismal fitness,
-	and therefore indirectly favor the evolution of mechanisms to suppress meiotic drive. % drive the rapid evolution of suppression mechanisms, and
-In this light, many facets of  o\"{o}genesis and gametogenesis have been 
-	interpreted as mechanisms of protection against genomic outlaws. 
-Why then is female meiosis often left uncompleted until after
-fertilization in many animals -- 
-	potentially providing an opportunity for sperm alleles to
-        meddle with its outcome and help like-alleles drive in heterozygous females? %and facilitate meiotic drive? 
-The population genetic theory presented herein suggests that
-	sperm nearly always evolve to increase the fairness of female
-        meiosis in the face of genomic conflicts. %%felt this
-                                %%might be an important addition as
-                                %%they will spread if no cost to organism
-These results are consistent with current knowledge of sperm-dependent meiotic drivers 
-(loci whose distortion of female meiosis depends on sperm genotype),
-	and suggest that the requirement of fertilization for the
-        completion of female meiosis
-        %, and variation in stage of meiosis that
-        %fertilization occurs at 
-        potentially represents a mechanism employed by females 
-	 to ensure a fair meiosis. %prevent meiotic drive.
-\newpage
-
-\section*{Introduction}
-Despite the apparent unity of the organism, `selfish' alleles can
-gain an evolutionary advantage at a cost to  individual fitness
-\citep{Burt2006}, often by exploiting meiosis and gametogenesis.
-Because only one of the four products of female meiosis is transmitted to the egg, female meiosis is particularly vulnerable to such exploitation \citep{Sandler1957,Pardo-ManuelDeVillena2001a}. 
-An allele that biases female meiosis in its favor (i.e.
-a meiotic driver), may increase in frequency even if it entails a pleiotropic fitness cost \citep{Prout1973}, generating a genetic conflict between the success of the driver and organismal fitness.
-Meiotic drivers observed in both plants
-\citep{Buckler1999,Fishman2005,Fishman2008}, and animals
-\citep{Agulnik1990,Wu2005,Pardo-ManuelDeVillena2001c} highlight this
-conflict -- the selfish benefits and the associated
-pleiotropic fitness costs of drive sustain a balanced polymorphism
-\citep{Prout1973}, 
-and often generate ongoing evolutionary escalations of drive suppressors and enhancers \citep{Dawe1996,Fishman2008}. 
-The threat of meiotic drive to organismal fitness is potentially so
-severe that many basic properties of
-meiosis and o\"{o}genesis, including the initial genome doubling in
-meiosis I \citep{Haig1991}, arrested female meiosis
-\citep{Mira1998}, the structure of centromere machinery \citep{Malik2002a,Malik2009}, and sex differences in the recombination rate \citep{Haig2010,Brandvain2012} 
-have perhaps evolved to enforce fairness by disrupting meiotic drive \citep{Rice2013}.
-
-It is therefore somewhat surprising that despite the intense evolutionary pressure on female meiosis to prevent meiotic drive, 
-it is potentially open to sabotage by a virtual stranger -- a haploid sperm genome.
-That is, in many animal species, female meiosis is completed only after fertilization \citep{Masui_book}, 
-	creating ample opportunity for interaction between the sperm and
-	female meiotic machinery (note that the alternation of generations in
-	likely precludes this interaction in plants). 
-Therefore, a `green-bearded' \citep{Gardner2010} sperm-acting allele that recognizes and facilitates the meiotic drive of a genetically equivalent allele in heterozygous females %female meiosis %(a  heteromorphic dyad),
-     could presumably rapidly spread through a population. 
-At first sight, female meiosis appears primed for exploitation by such selfish systems.  
-Why then is the requirement of fertilization to complete female meiosis so ubiquitous?
-
-It is certainly not the case that animals are mechanistically incapable of circumventing this requirement.
-The stage at which female meiosis requires fertilization varies widely across taxa, and in a few animals  
-	female meiosis is complete upon ovulation (that is, fertilization is not required for female meiosis \citep[e.g. Table 1 in 
-][]{Masui_book}).
-
-It is also not the case that sperm is mechanistically incapable of influencing female meiosis.
-In \emph{C. elegans}, premature deployment of the aster, a vital component of mitotic
-machinery provided by the sperm, disrupts MII meiotic segregation
-in the egg, leading to a triploid zygote \citep{McNally2012}. 
-Thus, there is direct mechanistic evidence that sperm can influence female meiosis.
- Additionally, genetic evidence from the \emph{In} and \emph{Om}  drive systems in mice suggests that the 
- transmission of alternate alleles in heterozygous females can depend on sperm haplotype.  
- While ruling out the alternative hypothesis of early selection on zygotes in these cases is challenging  \citep[see pages
-52-54 in][]{Burt2006}, it appears that the extent to which \emph{In} and \emph{Om} distort the second female meiotic division partially depends on the genotype of the fertilizing sperm \citep{Agulnik1993,Wu2005}.
-% Both of these female drive systems operate by distorting the second meiotic division, 
-% and in both systems the outcome of female meiosis depends the genotype of the fertilizing sperm \citep{Agulnik1993,Wu2005}.  
-%While such genetic evidence is suggestive, we note the difficulty in
-%ruling out the alternative hypothesis of early selection on zygotes \citep[pages
-%52-54][]{Burt2006}.
-
-We develop population genetic models exploring the evolution of sperm influence on female meiotic drive.
-%, focussing on models in which `self-promoting' alleles in sperm facilitate drive of like genotypes during gametogenesis in heterozygous females.
-We first focus on models in which `self-promoting' alleles in
-  sperm facilitate drive of like alleles during gametogenesis in heterozygous females. 
-%In this article we explore through simple population genetic models the consequences of alleles that influence the outcome of female meiosis.
-These models show that such sperm-acting alleles 
-	have more difficulty invading a population than do traditional meiotic drivers, 
-	 and under most circumstances, cannot be maintained as a balanced polymorphism.
-Because self-promoting drivers are unlikely to create a sustained genomic conflict, 
-	female meiosis will have little incentive to evolve resistance to them.
-We then examine models in which a novel sperm-acting allele modifies the efficacy of a polymorphic meiotic driver. 
-Such model universally reveal that sperm-acting drive modifiers are favored only if the suppress drive. 
-These results highlight the fact that the interests of sperm and maternal  genomes' are often aligned, as both are invested in the fate of the resultant zygote (as was speculated for the \emph{In} locus \citep{Pomiankowski1993}).
-Therefore, there is little selective benefit to females in preventing sperm to influence female meioses.
-Rather, features of o\"{o}genesis  may evolve to facilitate an interaction with sperm %with female meiosis
-	as yet another mechanism to enforce an equitable transfer of genetic material.
-
-\section*{Results}
-
-\subsection*{ Invasion of the population by a driving allele that promotes itself.}
-%%%%%%%%%%%%%%%%%%%%%%%%%
-% PLEIOTROPY MODEL [self-promoting]
-%%%%%%%%%%%%%%%%%%%%%%%%%
-
-In the standard single-locus, biallelic model of female meiotic drive,
-the driving allele is transmitted to the egg in heterozygotes with
-probability  $d > 1/2$, regardless of sperm genotype \cite[e.g.
-][ and see Model 1 in the Appendix for more details]{Ubeda2004}.
-%The timing of fertilization relative to female meiosis places another constraint on $d$, for example, if fertilization (and therefore, sperm dependent drive) takes place at MII (as in mammals),
-%	female drive requires an uneven number of crossovers between the centromere and the drive locus, 
-%	so $d$ is bounded to be $<0.83$ \citep[see ][ for discussion]{Buckler1999}.
-	%YB - MOVED THIS TO THE APPENDIX
-          To depict a case
-of a self-promoting meiotic driver,  we modify this standard model such 
-	that the driver is only effective when fertilized by a sperm
-        carrying that allele (see Figure \ref{Eggsperm}A and Model 2
-        in the Appendix and Supplementary Worksheet). 
-We then identify the conditions allowing for the spread of this self-promoting driver, 
-	and evaluate whether a driver of this form could generate a sustained conflict favoring the evolution of suppressors. 
-We conclude our single locus results with an analysis of a related model (Model 3) - in which
-drivers influence their transmission in females via paternal genotype,
-rather than sperm haplotype.
-
-For comparative purposes, we first briefly present the standard drive model %as Model 1 in the Appendix, 
-%equations \ref{eqn.A1}-\ref{eqn.A3}, 
-	%and refer readers to the previous literature on this model 
-	\citep[see e.g. ][ for additional results]{Prout1973,Ubeda2004}. 
-Assuming the driving allele is deleterious in both sexes, but fully recessive 
-	(i.e. the fitness of drive homozygotes equals $w_{BB}=1-s$ and other genotypic fitnesses equal $w_{AA}=w_{AB}=1$), 
-	it always invades because,  when rare, it occurs predominantly in heterozygotes and therefore drives without a fitness cost. 
-However, when $s$ is large ($s>(2d-1)/(2)$, solid black line in Figure \ref{Eggsperm}B), a driver cannot fix, and
-will be maintained as a protected polymorphism \citep{Prout1973}. 
-The parameter space where the allele can invade but not fix is shown in white
-        in Figure \ref{Eggsperm}B. %, and it can fix in the colored regions. 
-When the allele is maintained as a polymorphism, it provides an opportunity for the evolution of
-	drive suppressors, corresponding well to empirical examples of
-        female meiotic drive \citep[reviewed in ][]{Burt2006}.
-
-In contrast to a traditional driver, which drives but pays effectively
-	no fitness cost when rare, 
-a self-promoting driver specifically creates low fitness drive homozygotes 
-	by uniting driving female gametes with sperm enabling that drive.
-        % zygotes when it drives. in heterozygotes. 
-%, and therefore there is a drive-associated fitness cost even when rare. 
-It must therefore overcome a drive-associated homozygous fitness cost simply to spread when rare. 
-The conditions allowing the invasion of a self-promoting driver
- 	are consequently far more restrictive than those for a standard meiotic driver.
-When rare, a fully recessive, self-promoting driver can only invade when $s$ 
-	%, the selection against the drive homozygote, 
-	is approximately less than $(2 d - 1)/(4 d)$ -- see dashed black line in Figure \ref{Eggsperm}B). 
-This analytical approximation, derived from Equation \eqref{deltadriver} assuming Hardy-Weinberg, 
-	closely matches results obtained by exact numerical iteration (Figure \ref{Eggsperm}B).
-
-\begin{figure}
-\includegraphics[width = 1 \textwidth]{Figures/selfprom.eps}
-\caption{{\bf{A.}} A visual depiction of our model of `self-promoting' driver. 
-	Transmission probabilities for alleles through female
- 	 meiosis depend on sperm genotype. 
-	 The non-driving $A$-allele, and self-promoting $B$-allele
-	 are represented by unfilled and filled circles, respectively. 
-	 {\bf{B.}} Evolution of a self-promoter and standard driver. 
-	 	Assuming that the fitnesses of drive homozygotes and 
-		heterozygotes are $1-s$ and $1$, respectively. 
-	 	\emph{Main figure:} Boundary conditions for the invasion 
-		and fixation of self-promoting and standard meiotic drivers, 
-		with drive coefficient, $d$. Colored regions depict exact results, 
-		while lines  represent analytical approximations. 
-	 \emph{B1:}  Trajectories of sperm-dependent female drive
-		 %alleles that can invade the population, 
-		 each allele has $s =0.1$ against the
-		 homozygotes.  The drive coefficient is denoted by color. 
- }   
-\label{Eggsperm}
-\end{figure}
-
-When a self-promoting driver does spread, 
-	it spends much time at low frequency, 
-	because the paucity of complimentary sperm compromises its ability to drive. 
-However, once relatively common, it  rapidly achieves fixation due to its
-	positive frequency dependent behavior (Figure \ref{Eggsperm}B.1).  
-%This lack of a balanced polymorphism %(Figure \ref{Invasion_space}) 
-%	precludes the evolution of an
-%	allele that suppresses this form of meiotic drive in females.
-The positive frequency dependence of a self-promoting driver can also
-	induce bistability in its spread -- some values of $s$ 
-	allow the fixation of this driver when common, but 
-	preclude its invasion when rare (Equation \ref{ineqA} and Figure \ref{Eggsperm}B). 
-In this case, the driver will be fixed if it reaches a frequency greater than some threshold 
-	(approximated in Equation \ref{bistab_homozy_approx} and  presented exactly in 
-	Figure \ref{Bistab_homozyg_cost_fig}) and lost otherwise. 
-For most parameters, this threshold is likely too high to be reached by drift, 
-	and therefore  the fate of self-promoting drivers is determined
-	by the more restrictive invasion criteria rather than the fixation criteria.
-
-%I SUBSTANTIALLY REWROTE THIS
-%The positive frequency dependence of a self-promoting driver can
-%induce bistability in its spread.
-%OLD GRAHAM WORDS
- %, i.e. there is a critical frequency
-%above which it can spread quickly to fixation and below which it moves towards loss (an
-%unstable equilibrium frequency). In 
-%the appendix, we develop a simple
-%approximation to this bistability. 
-%When $s$ is too large to allow a rare driver to increase in frequency, but small enough to allow a common driver to fix (the inequality in Equation \ref{ineqA}),
-%there is  an unstable equilibrium frequency, $f^*_{B\text{ recessive}}$  
-%	above which it can spread quickly to fixation and below which it moves towards loss 
-%	($f^*_{B\text{ recessive}}$ is approximated in Equation \ref{bistab_homozy_approx}.  
-%	We compare this approximation to the exact result in Figure \ref{Bistab_homozyg_cost_fig}).	
-%OLD GRAHAM WORDS
-%This bistability occurs in the region of parameter space in Figure
-%\ref{Invasion_space}. In Supplementary Figure
-%\ref{Bistab_homozyg_cost_fig} we further explore this bistability and
-%show that the unstable equilibrium frequency is high across much of
-%this parameter space, and so in practice the self promoting allele
-%only invade in the orange space. (In the appendix we develop a simple
-%approximation to the unstable equilibrium frequency
-%\eqref{bistab_homozy_approx} that is also shown in  Supplementary Figure
-%\ref{Bistab_homozyg_cost_fig}).
-%This unstable equilibrium frequency is high across much of
-%	parameter space, likely too high to be reached by drift, 
-%	and therefore  the fate of self-promoting drivers  will  likely be determined
-%	by the more restrictive invasion criteria rather than the fixation criteria.
-
-%OLD YANIV WORDS
-% Compared to the invasion criteria, the parameter space
-% allowing the fixation of the allele if it is above a cutoff frequency is much broader -- 
-% 	a common, recessive, self-promoting driver fixes if
-%         $s<(2d-1)/2$ approximately. 
-% Therefore, when $(2 d - 1)/(4 d)<s<(2d-1)/(2)$, 
-% 	a recessive, self-promoting driver will deterministically fix if drift, mutation, 
-% 	or migration pressure bring its frequency above $(1-2d+4ds)/(2s(2d-1))$, 
-% 	but it will be lost otherwise.
-Inclusion of a heterozygous fitness cost (i.e. $w_{AB}=1-s_h$)
-	further constrains the evolution of a self-promoting driver. 
-In fact, with any heterozygous fitness cost, a rare self-promoting
-	driver is always selected against. 
-However, this case also displays bistability -- 
-	when $s$ is sufficiently small (Equation \ref{sForHetFix}) this allele fixes deterministically if its 
-	frequency exceeds some threshold  (Equation \ref{thresh2}, exact results in 
-	Figure \ref{bistable_additive}).
-%	$f_B^*$
-%	because it pays the  heterozygous fitness cost but drives
- %      ineffectively (see Figure S\ref{bistable_additive})
-%While a standard female driver invades from low frequency when the transmission advantage outweighs the fitness cost (XXX),
-%	a rare self-promoter pays a fitness cost, but drives only weakly because complementary sperm are uncommon. 
-%This frequency-independent fitness cost, and positive frequency dependent drive generates a bistable system. 
-%Below a cutoff frequency the self-promoter driver deterministically decreases in frequency 
-%	because it pays the  heterozygous fitness cost but drives
- %      ineffectively (see Figure S\ref{bistable_additive}). 
-%Above this frequency, the allele increases deterministically because the commonness of complementary sperm 
-%	make drive effective enough to overcome the associated fitness cost.
-This bistability prevents self-promoting drivers from invading 	
-	reasonably sized populations, and assures that if they do invade, they will rapidly fix.
-Our model therefore predicts that self-promoting drivers will not be
-	observed as stable polymorphisms in natural populations. 
-This lack of a balanced polymorphism %(Figure \ref{Invasion_space}) 
-	precludes the evolution of an
-	allele that suppresses this form of meiotic drive in females.
-
-Although the allelic identity of sperm could plausibly influence the outcome of female meiosis, 
-	limited gene expression in sperm \citep[e.g.][]{Joseph2004}
-	suggests a model where sperm influence female meiosis via expression of the fertilizing male's
-	diploid genotype (perhaps due to proteins and RNAs packaged into the sperm), rather than sperm haplotype.
-This paternal-genotype dependent model (Model 3 in the Appendix) requires one additional parameter, as we exchange $d$ in the sperm dependent case for $d_{het}$ and $d_{hom}$ which describe the transmission of the drive allele in a heterozygous female mating with males heterozygous and homozygous for the self-promoting drive allele, respectively.
-Here, a rare driver invades when $s$ is less than $(d_{het}-1/2)/d_{het}$, and usually fixes when it invades.
-	% fixes when $s$ is less than $d_{hom} - 1/2$.
-However, when the distorting effect of genotype displays strong dominance in its
-	effect on female meiosis ($d_{het}$ is close to $d_{hom}$), 
-	there is a a narrow sliver of parameter space sustains a polymorphism
-	when the cost of the drive is recessive
-	(see Figure \ref{Effect_of_dominance}, and Equation \ref{polymale}).  
-%Presumably this zone of polymorphism comes about because positive frequency 
-%	dependence diminishes with the dominance of the male effect driver. 
-%As an analytical example, consider the case of male-genotype dependent deleterious recessive female driver (assuming, as above, $w_{BB}=1-s$, and $w_{AA}=w_{AB}=1$). 
-%Here, a rare driver invades when $s$ is less than $(d_{het}-1/2)/d_{het}$, and fixes when $s$ is less than $d_{hom} - 1/2$.
-%Intermediate values of $s$, which represent a small proportion of parameter space can maintain a polymorphism. 
-While mathematically interesting, it does not seem particularly realistic to think that the
-	effect of the drive allele would be dominant in its action through the
-	male genotype, while the cost would be recessive. 
-Therefore, although Model 3 can sustain a polymorphism, 
-	the lack of biological reality underlying the small portion of parameter values 
-	required for this polymorphism make us doubt it general applicability. 
-%Thus we regard this case as interesting but not very general.
-
-Given the difficulty that self-promoting meiotic drivers have entering the population, the speed at
-which they fix if they do, and the narrow parameter range permitting balanced polymorphisms at such loci,  
-it seems very unlikely that such alleles could drive the evolution of female suppressors of sperm-enabled
-female meiotic drive.
-
-%\begin{figure}
-%	\rotatebox{90}{
-%\includegraphics[width = 0.5
-   %       \textwidth]{Figures/Sperm_egg_cartoon_paper_Fig1.eps}}
-%\caption{transmission probabilities for alleles through female
-%  meiosis depend on sperm genotype. 2 allele model. \gc{merge with 2,
-%    rewrite both captions}}  
-%\label{Eggsperm_2_allele_cartoon}
-%\end{figure}
-
-% (we note that the timing of fertilization relative to female meiosis places another constraint on $d$, for example, when drive takes place druing the second meitoic division
-%	it requires an uneven number of crossovers between the centromere and the drive locus \citep[see ][ for discussion]{Buckler1999}).
-
-%\begin{figure}
-%\includegraphics[width = 0.8 \textwidth]{Figures/invasion_space_recessive_driver.eps}
-%\caption{{\bf A.} Invasion analysis for a female meiotic drive allele with
-%  recessive costs (selection coefficient $s$). First consider a simple female meiotic
-%  drive system, where the allele is transmitted a fraction $d>1/2$ of
-  %meioses in female heterozygotes, inspective
-%  of the allele carried by the fertilizing sperm. 
-%Such an allele can invade the population over all of the parameter space, and will be maintained as a
-%  protected polymorphism in the white region, and can fix in the
-%  population in all of the parameter space colored green or orange. 
-%The solid back line marks the criteria for the allele to fix $s< (2d-1)/2$.
-%Now consider an allele that is fairly transmitted through
-%  female meiosis when fertilized by a sperm carrying a different
-%  allele, but at a fraction $d>1/2$ when a heterozygote egg is fertilized
-%  by a sperm carrying the allele (as depicted in Figure
-%  \ref{Eggsperm_2_allele_cartoon}). Such an allele can only invade in
-%  the region of parameter space colored orange, and it fixes over all
- % of this parameter region. The approximate criteria for this allele
-%  to fix is given by the dashed line showing $s < (2d- 1)/(4d)$. 
-% {\bf B.} In the inset we show trajectories of sperm-dependent female drive
- %alleles that can invade the population, each allele has $s =0.1$ against the
-% homozygotes and $d =0.6$, $0.7$, $0.8$ shown in yellow, blue, and red. 
-% \gc{Likely make this Figure 1B.} } \label{Invasion_space}
-%\end{figure}
-
-\subsection*{Two locus models of sperm-dependent female drive}
-
-%\subsection*{ sperm-egg haplotype balanced drive systems}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%
-% TWO LOCUS MODEL [modify (general) / facilitate (helps ) / prevent (stops)] 
-%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Models 2 and 3, above, explored the dynamics of an allele that drove in females when signaled by a complementary signal in sperm.   
-We complement this single-locus approach %of an allele with pleiotropic effects in both sexes, 
-	with alternative models of two loci - one a female driver, 
-	and the other, a sperm-acting allele which modifies the effect of drive upon fertilization. 
-In this model, a female meiotic driver with no sperm-dependency 
-	initially reaches drive-viability equilibrium (with two alleles
-	$A$ and $B$ are the ancestral non-driver and driver alleles, Figure \ref{Eggsperm_3_allele_cartoon}A1). 
-Subsequently, a sperm-acting modifier of female meiotic drive arises at another locus.
-
-We first assume that the modifier is tightly
-        linked to the drive locus (effectively creating a third
-        allele/haplotype at this locus) and arises on the drive-background. 
-Tight linkage offers the best
-        chance for a collaboration to evolve between a driver and
-       a sperm-acting drive  enhancer, as recombination breaks up drive haplotypes \citep{Thomson1974,Charlesworth1978,Haig1991}. 
-Additionally, tight linkage between female driver and sperm modifier is consistent with the nature of well characterized drive systems which are often maintained as polymorphic inversions with numerous linked modifiers \cite{Burt2006}.
-We conclude by analyzing models with alternative linkage relationship between driver and drive modifier - 
-	in Model 5 the modifier arises tightly linked to the non-driving allele, 
-	and in Model 6 it is unlinked to the driver.
-
-%%%sperm-acting facilitator
-When the modifier of drive is tightly linked to the driver and
-enhances drive (i.e. in coupling phase), 
-	we label  this non-recombining drive/modifier haplotype as the $B^+$ allele.  
-The $B^+$ allele acts in sperm to increase the effectiveness of drive for both
-	the $B$ and  $B^{+}$ alleles in $AB$ and $AB^{+}$
-        heterozygotes (see Figure \ref{Eggsperm_3_allele_cartoon}A2, and
-        Model 4 in the Appendix and Supplementary Worksheet). 
-Naively, $B^{+}$ may spread by capitalizing on the additional drive it causes; however,  
-	this is not the case for a few simple reasons. 
-First, 
-	the novel $B^{+}$ haplotype
-	arises when the ancestral driver 
-	is at drive-selection balance, 
-	and therefore immediately suffers a homozygous fitness cost.  
-Worse yet, a novel $B^{+}$ haplotype most often helps 
-	the $B$  allele drive ($B^+$ sperm meeting $AB$ eggs), because $B$ is initially more common than $B^{+}$. 
-Therefore, sperm-acting drive facilitator alleles experience a profound disadvantage 
-	in this scenario, even more so than under the previous two allele model. 
-We have found no parameter range of this
-	three allele system that allows the sperm-acting drive facilitator $B^{+}$ to
-	invade the population (Appendix Model 4, eqn. \eqref{coupling}, and Mathematical Supplement).
-
-\begin{figure}
-	\includegraphics[width = 0.7
-          \textwidth]{Figures/twolocus.eps}
-\caption{Models of a sperm-acting drive modifier tightly linked to a meiotic driver. 
-{\bf (A) } Loci exist in tightly linked haplotypes (top of A). 
-	Genotype of sperm at the modifier locus (ancestral and sperm acting modifier empty and filled squares, respectively) alters transmission  
-	at the drive locus (circles) during female meiosis. 
-	{\bf{A1)}} In the standard model of drive, 
-	the driver (filled black circle) and non-driver (empty circle) are transmitted with probabilities $d_0$ 
-	and $1-d_0$, respectively. 
-	{\bf{A2)}} The $B^+$ haplotype increases the transmission of the drive allele ($d_1 > d_0$), and due to their shared genetic background, also increases its drive. 
-	{\bf{A3 \& A4)}} The sperm-acting modifier acts to decrease drive  ($d_1=1/2$ in A3 \& A4. More generally, $d_1 < d_0$). 
-		In {\bf{A3}} the drive suppressor exists on the drive background, forming the $B^-$ haplotype. 
-		In {\bf{A4}} the drive suppressor exists on the non-drive background, forming the $A^-$ haplotype. 		
-{\bf{(B)}} Invasion of a sperm-acting drive suppressor linked to a driver. 
-	After the driver ($B$ haplotype) reaches drive selection equilibrium, we introduce a sperm acting drive modifier. 
-	We assume full drive ($D_0 =1$), a recessive lethal fitness cost to drive ($h_s=0$, $s=1$) and that the sperm-acting modifier results in a fair meiosis. 
-{\bf {B1)}} The $B^-$ allele replaces the ancestral drive haplotype, but segregates at a lower equilibrium frequency.
-{\bf {B2)}} The $A^-$ allele replaces the ancestral non-driving haplotype, and in this case, removes the driver from the population. }
- \label{Eggsperm_3_allele_cartoon}
-\end{figure}
-
-%%%sperm-acting drive suppressors
-While  sperm enhancement of a female drive cannot displace a polymorphic female driver, sperm based drive suppressors can. 
-%	that is sperm-acting alleles that act to restore  fairness of female meiosis, can. %If these sperm-acting drive suppressors arise on
-Imagine a sperm-acting allele that restores fairness to female meiosis arises on
-	the drive background, i.e. in repulsion phase, creating a
-        third allele $B^{-}$ (Figure
-          \ref{Eggsperm_3_allele_cartoon}A3, Model 4 in Appendix). This new allele still
-        experiences female drive when fertilized by A or B sperm, but 
-       is does not drive when fertilized by another $B^{-}$ so it
-       avoid the excess formation of low fitness homozygotes. This
-     allows the $B^{-}$ to displace the ancestral driver (Figure
-     \ref{Eggsperm_3_allele_cartoon}B1, Equation \ref{coupling}), 
-	and often returns to a lower equilibrium frequency than the the $B$ allele, further decreasing the extent of drive in the population.  
-If this sperm-acting drive suppressor arises on 
-	the non-driving $A$ background (creating a third allele $A^{-}$, Figure \ref{Eggsperm_3_allele_cartoon}A4, Model 5), 
-	or is unlinked to the drive locus (Model 6), it readily invades a population segregating
-	for the drive system (Equations \ref{A-} and
-        \ref{unlinked}).  %%DELETED FIG REF HERE.
-This allele lowers the frequency of the original driver (perhaps to zero),
-	and spreads to fixation if it does not carry strong fitness costs
-	(Figure \ref{Eggsperm_3_allele_cartoon}B2). 
-This result is consistent with previous work showing that drive suppressors unlinked to, or in repulsion phase with drivers usually invade polymorphic drive systems \citep[e.g. ][]{Brandvain2012}.  
-Therefore, all two-locus models of sperm influence on female drive suggest that
-	sperm will evolve to oppose female meiotic drive.
-
-%More surprisingly, sperm-acting suppressors of female drive can spread
-%	even when tightly linked to the original drive system (Figures \ref{Eggsperm_3_allele_cartoon} and \ref{Trajectories_of_supressors}B). 
-%The $B^{-}$ haplotype (representing the tightly linked drive/drive suppressor complex) displaces the original driver 
-%	by driving when not costly, but avoiding low fitness homozygotes by disrupting drive when transmitted to sperm. %  ($BB^-$ and $B^-B^-$). 
-%Despite displacing the ancestral driver (the $B$ haplotype), the equilibrium frequency of the $B^-$ haplotype is often lower than that of $B$ \yb{(point to supp?)}.
-
-%\gc{Do we mention the In and Om locus here or in the discussion as a
-%	 potential e.g. Do we mention the fact that the frequency of these
-%	 newly invade sperm dodging alleles can be lower than the original system?}\\
-
-\section*{Discussion}
-
-Sexual reproduction is a high-stakes event that determines what gets transmitted to the next generation.  
-As a consequence of this intense competition, alleles that gain a transmission advantage during reproduction 
-	can succeed evolutionarily even if they lower organismal fitness. 
-This generates numerous conflicts  \cite{Burt2006} including sexual conflicts between mates \cite{Arnqvist2005}, 
-	and conflicts between alleles that are over-transmitted in meiosis and the organisms they injure while doing so. 
-Such conflicts and their resolution likely play a major role in the structure and evolution of many basic biological processes \citep{Rice2013}.
-
-It seems that allowing sperm to influence the outcome of female meiosis would generate a confluence of these potential conflicts -- 
-	sperm could actually assist an allele that distorts female meiosis.
-However, this is not the case.
-We find that an allele which acts through sperm to distort female meiosis in its favor %YB [too awkward / unclear?] 
-	can rarely spread through a population if it bears any cost. 
-Additionally, when this self-promoting driver can spread, it can only rarely 
-	be maintained as a protected polymorphism, and due to its positive frequency dependence,  
-	it spends very little time at intermediate frequency.
-As such, this type of exploitation cannot generate a sustained genetic conflict.
-It is therefore unlikely
-	that female oogenesis and meiosis will evolve to prevent their effect.  
-This likely explains why females can delay the completion of meiosis until after fertilization 
-	without risking exploitation by collaborations by female
-        drivers and sperm alleles.
-%Male modification of recombination rates. \yb{trying to think through this \dots{}}
-
-Why is it that an allele that biases female meiosis in its favor can generate a genetic conflict, but an allele in sperm that assists this female driver cannot? 
-So long as the transmission advantage of female meiotic drive outweigh the organismal fitness cost to heterozygotes the female driver can spread when rare, and it increases in 		
-	frequency until the fitness cost experienced when homozygous balances the transmission advantage.
-By contrast, a sperm promoter of female drive is only effective when matched with a heterozygote female -- meaning that, when rare, this allele rarely enhances female drive. 
-Even worse, when it does so it will preferentially find itself in a low fitness drive homozygote. 
-Not only are drive-promoting sperm alleles unable to create a sustained genetic conflict, 
-	but alleles in sperm with the opposite effect - that is those that prevent their own drive through female meiosis do maintain a polymorphism and 
-	provide evolution  with time and opportunity to further minimize drive.
-This is because such drive suppressing alleles reduced their chances
-of forming low fitness homozygotes. 
-%In fact, an allele in sperm which causes the alternate allele to drive can be favored, a counterintuitive result 
-%	consistent with the tenuous evidence that the non-driving
-%        $ln^-$ allele is over transmitted through female meiosis in $ln^-/ln^+$ heterozygote when fertilized by $ln^+$
-%        sperm \cite{agulnick, Pomiankowski1993}.   
-% YB - WE DID NOT TALK ABOUT THIS IN THE RESULTS [ALTHOUGH IT'S TRUE... PERHAPS TOO DISTRACTING]
-More generally, alleles that act through sperm to
- reduce the opportunity of female meiotic drive regardless of linkage or phase.
-
-Our theory offers a novel interpretation of a baffling idiosyncrasy of female meiosis -- 
-	we argue that females do not complete meiosis until after fertilization as a mechanism to encourage meiotic fairness. 
-In addition to providing a new interpretation of a basic biological process, our theory provides numerous testable predictions. 
-We predict that female drive, at a stage influenceable by sperm,  
-	should be reduced when fertilized by
-	complimentary sperm. 
-Therefore, For example, MII female drive should be exceptionally severe in
-	interspecific crosses between animals which complete the second meiotic division after fertilization,  
-	as heterospecific sperm will not specifically evolve to resists a female driver in another species. 
-Our theory also encourages phylogenetic hypotheses concerning the relationship between the opportunity 
-	for female meiotic drive the requirement of fertilization for the completion of female meiosis. 
-For example, we predict that a lower opportunity for female meiotic drive,
-        e.g. an animal lineage with a long history of high inbreeding
-        or selfing, 
-        would often accompanied by a relaxation of the requirement of fertilization for the
-        completion of female meiosis.
-
-%More generally both linked and unlinked alleles may often spread that act through sperm to
-%reduce the opportunity of female meiotic drive in order to disable
-%segregating drive systems, by for example reducing the mechanistic
-%asymmetries that female drivers exploit.
-
-Our results highlight potentially counterintuitive results of complex genetic conflicts. 
-Despite much opportunity for conflict between sperm and females over fertilization \citep{Partridge1998},
-	the interests of fertilizing sperm and female are quite well aligned during syngamy.
-While conflict between mother and her alternative chromosomes ensue, 
-	fertilizing sperm decidedly side with mom, as both have a shared interest in producing a viable and potentially fit offspring. 
-%As noted above there are even opportunities for coevolution between sperm-based suppressors of female meiotic drive and the female genome.
-%Thus it seems plausible that aspects of oogenesis may actively evolve to facilitate the influence of sperm on ensuring a fair  outcome of female meiosis which generates fit offspring.
-%In light of that it is interesting to consider whether the phylogenetic variation in the stage of oogenesis when fertilization takes place reflects not
-%	only species' different reproductive ecologies, but also
-%        different selection pressures shaping the interaction of sperm
-%        alleles with female meiosis. 
-%\gc{For example, we would predict that a lower the opportunity for female meiotic drive,
-%        e.g. an evolutionary transition to inbreeding or a high
-%        selfing rate in animals, would often accompanied by a
-%        relaxation of the requirement of fertilization for the
-%        completion of female meiosis.}
-
-  \appendix
-
-\section*{Appendix}
-We highlight simple analytical results from: 
-	the standard model of female meiotic drive (Model 1); a sperm-dependent 
-	model of female meiotic drive (Model 2); a male-genotype-dependent model of
-	female meiotic drive (Model 3); and more complex two-locus sperm-dependent drive models (Models 4-6). 
-All models include a biallelic locus ($A/B$) with non-driving and driving alleles in frequencies $f_A$ and $f_B$, 
-	respectively, while more complex models include more loci and/or alleles.  
-Transmission rules describing the outcomes of all matings in each
-	model are presented in a Supplementary Worksheet. 
-The fitness of genotype, $g$ is sex-independent and equals $1$, $1-h_s$, and $1-s$ for genotypes $AA$, $AB$, and $BB$, respectively. %  and  $w_{AA} = 1$, $w_{AB} = 1-s_h$, and $w_{BB}=1-s$. 
-Genotypic frequencies  equal  $f_g$ for adults in the current generation, 
-	$f_g'$ in the next generation of zygotes (i.e. after recombination, random mating, drive, and syngamy, but before selection)
-	and $f_g''$ in adults after selection. 
-After a complete generation $f_g'' = f_g' w_g/\bar{w}$, where $\bar{w}$ is the population mean fitness and equals $\Sigma w_g f_g'$. 
-Because many of our analyses assume genotype frequencies expected under Hardy Weinberg Equilibrium (HWE), we note which analytical results are approximations. 
-We verify these approximations with exact numerical iterations in Supplementary Figures and Supplementary Scripts, and detail the derivation of these approximations in our Mathematical Supplement.
-
-\subsection*{Models 1-3. Single-locus drive}
-\subsubsection*{Model 1. Traditional driver}
-In the standard female drive model, meiosis in males is fair such that $A/B$ heterozygotes contribute $A$ and $B$ alleles with equal probabilities; however, $A/B$ females transmit the $B$ allele with probability $d>1/2$.
-We note that the timing of fertilization relative to female meiosis places another constraint on $d$, for example, if fertilization (and therefore, sperm dependent drive) takes place at MII (as in mammals),
-	female drive requires an uneven number of crossovers between the centromere and the drive locus, 
-	so $d$ is bounded to be $<0.83$ \citep[see ][ for discussion]{Buckler1999}. 
-After drive and random mating, genotype frequencies are 
-	$f_{AA}'=f_A (f_{AA }+ f_{AB} (1 - d))$, 
-	$f_{AB}'=f_A (f_{BB} + f_{AB}d) + f_B (f_{AA} + f_{AB} (1 - d))$,
-	and $f_{BB}'= f_B (f_{AB} d + f_{BB})$. 
-%Assuming fitnesses $w_{AA}=1$, $w_{AB}-1 s_h$ and $w_{BB}=1-s$, 
-%Frequencies of genotype, $i$, after drive and selection are 
-%	\begin{equation} f_i''= f_i'w_i/\bar{w}\end{equation}
-%	where $w_i$ is genotypic fitness, and mean fitness, $\bar{w} = \Sigma f_i'w_i$. 
-As detailed above, exact frequencies after drive, random mating and selection are $f_g''= f_g'w_g/\bar{w}$. 
-Assuming HWE, a rare driver will spread when $s_h
-        \lessapprox(2d-1)/(1+2d)$, and will fix when $s\lessapprox d -1/2 +
-        3 s_h /2- d s_h$. This later inequality reduces to
-        $s\lessapprox(2d-1)/2$ when the cost of drive is fully
-        recessive. 
-        %\gc{isn't this last expression exact?} YB: exact-ish.. all assume HWE.. but i think we don't expect a deviation from HWE here??  it's prob easiest to kept them all as approx?
-
-\subsubsection*{Model 2. Single locus, sperm-dependent drive.}
-Our single-locus model of sperm-dependent drive resembles the traditional driver, with the caveat that the $B$ allele drives in heterozygous females only when fertilized by $B$-bearing sperm. 
-Therefore, genotype frequencies after drive are 
-	$f_{AA}'=f_A^2$, 
-	$f_{AB}'=f_A f_B+f_B (f_{AA} + f_{AB}(1 - d)) $,
-	and $f_{BB}'= f_B (f_{AB} d + f_{BB} )$. 
-%As above, exact frequencies after drive, random mating and selection are $f_i''= f_i'w_i/\bar{w}$. 
-We iterate exact genotype frequency recursions ($f_g''= f_g'w_g/\bar{w}$) over generations to produce the frequency trajectories shown in the inset of 
-Figure \ref{Eggsperm}B by plotting $f_B=f_{BB}+ \frac{1}{2}f_{AB}$
-over time. %, and to assess the analytical approximations, below.
-To assess invasion or fixation criteria, as well as bistability points, we iterate this system and
-test whether $f_B$ increases over a grid of parameters.
-
-% the $B$ allele In our first scenerio allele $A$ is the ancestral allele and allele
-%$B$ is our sperm dependent drive allele.
-%The sperm dependence of the drive is a form of assortative
-%transmission (mating), and so we have to keep track of the genotype frequencies.
-%Our genotype frequencies at birth are $x_{AA},~x_{AB},$ and
-%$x_{BB}$. These three genotypes have fitnesses $w_{AA},~w_{AB},$ and
-%$w_{BB}$, which are identical between males and females. Because 
-%fitnesses are equivalent in both sexes, genotypic frequencies
-%are identical in males and females, but in the following
-%equations they are marked with $\Mars$ and $\Venus$ respectively to
-%allow the reader to more easily follow the logic. 
-%The frequency of the three genotypes at birth in the next generation is:
-%\begin{equation}
-%x'_{BB} =\frac{1}{\bar{w} } \Big( w_{BB} x_{BB}^{\Venus} (w_{BB} x_{BB}^{\Mars} +\frac{1}{2}
-%  w_{AB} x_{AB}^{\Mars} ) + w_{AB} x_{AB}^{\Venus} (d w_{BB}
-%  x_{BB}^{\Mars} +\frac{1}{2} d
-%w_{AB} x_{AB}^{\Mars} ) \Big),  \label{eqn.A1}
-%\end{equation}
-%\begin{equation}
-%x'_{AA} =\frac{1}{\bar{w} } \Big(  w_{AA} x_{AA}^{\Venus} (w_{AA} x_{AA}^{\Mars} + \frac{1}{2}
-%   w_{AB} x_{AB}^{\Mars}) + w_{AB} x_{AB}^{\Venus}(\frac{1}{2} w_{AA} x_{AA}^{\Mars}+ \frac{1}{2} 
-%  w_{AB} x_{AB}^{\Mars}) \Big),  \label{eqn.A2}
-%\end{equation}
-%and
-%\begin{eqnarray}
-%x'_{AB} = & \frac{1}{\bar{w} } \Big( w_{AA} x_{AA}^{\Venus}  ( w_{BB}  x_{BB}^{\Mars} + \frac{1}{2}
-%  w_{AB}  x_{AB}^{\Mars} ) +w_{BB} x_{BB}^{\Venus} ( w_{AA}  x_{AA}^{\Mars} + \frac{1}{2}
-%  w_{AB}  x_{AB}^{\Mars} )  \nonumber \\
-%& + x_{AB}^{\Venus}(\frac{1}{2}  w_{AA}  x_{AA}^{\Mars}+\frac{1}{2} (1-d)
-% w_{AB}  x_{AB}^{\Mars} + \frac{1}{4}  w_{AB}
- %x_{AB}^{\Mars} + (1-d)  w_{BB}  x_{BB}^{\Mars} ) \Big) \label{eqn.A3}
-%\end{eqnarray}
-%where the mean fitness of the population is
-%\begin{equation}
-%\bar{w} = w_{AA} x_{AA}+w_{AB}  x_{AB} +w_{BB}  x_{BB}.
-%\end{equation}
-%We can then iterate this recursion over generations, for
-%example to produce the frequency trajectories shown in the inset
-%Figure \ref{Invasion_space} by plotting $x_{BB}+ \frac{1}{2}x_{AB}$
-%over time. We can also use this to assess the
-%invasion criteria for the allele by starting allele B at very low
-%frequency, and seeing if it increases across the generations.
-
-\paragraph{Recessive fitness cost of self-promoting driver:} %When the self-promoting driver has a recessive fitness cost, $s$ genotype frequencies after selection and drive equal
-%\begin{equation}	x'_{BB} =\frac{x_{B}^2(1-s)}{\bar{w} }  \Big( F+(1-F)(x_{B}+2x_{A} d) \Big) 	\end{equation}
-%\begin{equation}	x'_{AA} =\frac{1}{\bar{w} }  \Big(x_{A}^2\Big) 	\end{equation}
-%and
-%\begin{equation}	x'_{AB} =\frac{2 x_{A}x_{B} }{\bar{w} }  \Big(1-x_{B}(d-1/2)(1-F) \Big)	%\end{equation}
-When fully recessive, the change in frequency of the self-promoting driver across generations equals 
-\begin{equation}
-\Delta f_B=f_A f_B^2 \left( [1-F][ d(1-2 f_A s) - f_B s-1/2] - s  F \right)/\bar{w}
-\label{deltadriver}
-\end{equation}
-where $F$ is the deviation from genotypic frequencies expected under Hardy-Weinberg. % Trying to clarify some.   \gc{Hardy Weinberg Expectations} (HWE). 
-%For analytical solutions presented above, we assume that $F$ equals zero. 
-Assuming HWE  ($F=0$) a common, recessive, self-promoting driver invades if $s\lessapprox(2 d - 1)/(4 d)$, and fixes if
- $s\lessapprox(2d-1)/2$. 
- Therefore, when 
-\begin{equation}
-(2 d - 1)/(4 d)  \lessapprox  s  \lessapprox  (2d-1)/2\label{ineqA}
-\end{equation} 
-a recessive, self-promoting driver will deterministically fix if drift,
- 	mutation, or migration pressure bring its frequency above
-\begin{equation} 
-f^*_{B\text{ recessive}} \approx (1-2d+4ds)/(2s(2d-1)) \label{bistab_homozy_approx}
-\end{equation}
- but it will be lost when introduced below this frequency. 
- %(i.e. this is an unstable equilibrium point). 
-% In Supplementary Figure \ref{Bistab_homozyg_cost_fig} we show
-% this approximation condition for bistability andd how it compared to
-% that obtained using the recursion. 
-Compared to exact results (Supplementary Figure \ref{Bistab_homozyg_cost_fig}), Equations
- \eqref{ineqA} and \eqref{bistab_homozy_approx} offer reasonable, but imperfect approximations. 
-%s_h and seems to serve as a lower bound to the true unstable equilibrium frequency.
-
-\paragraph{Cost of driver in heterozygotes:} %We explored a more complex model where the self-promoting driver has a cost, $s_h$ in heterozygotes (so $w_{AB} = 1-s_h$) in addition to the cost $s$, in homozygotes.
-When the fitness of drive heterozygotes is compromised ($s_h>0$), a self-promoting driver cannot invade when rare.
-This results from the fact that,
-when rare, %($f_B \ll 1$) 
-$B$-bearing sperm and heterozygous eggs will rarely encounter one another ($\sim f_B^2$) but the allele still pays a
-cost in heterozygous individuals ($\sim f_B$). 
-%Model construction is similar to that above.
-%So long as there is any heterozygous fitness cost (i.e. $s_h > 0$) the self-promoting driver cannot invade when rare;
-However, this system too, is bistable -- as the driver increases in frequency it is more often fertilized by a driving sperm and therefore drives more effectively. 
-Therefore, assuming HWE, if 
-\begin{equation}
-	s \lessapprox  d -1/2 +s_h(3-2d)/2
-	\label{sForHetFix}
-\end{equation}
-this self-promoting driver deterministically fixes when its frequency is greater than
-\begin{equation}
-	f_B^* \approx \frac{a_1+\sqrt{a_1^2+8 s_h a_2}}{2 a_2}
-	\label{thresh2}
-\end{equation}
-where, $a_1=1-2 ds_h+4 d s-2 d-3 s_h$ and $a_2=2(s-s_h)(2d-1)$.
-Comparison of Equation \ref{thresh2} to exact results obtained by a
-simple parameter search (Supplementary Figure \ref{bistable_additive})
-show that this approximation is reasonably correct for small
-parameter values; however, it underestimates $f_B^*$ for large parameter values, presumably because they result in strong departures from HWE.
- %range of other costs to the drive alleles other than a
-%simple recessive cost, for example an additive cost. We modified the recursions given in eqn
-%\ref{eqn.A1}-\ref{eqn.A3} to feature an additive cost $s$ for each
-%copy of the $B$ allele. A none recessive cost to the drive allele can
-%cause bistability of the drive allele. The B allele is unable to enter
-%the population when rare, this results from the fact that
-%when the allele is rare ($x_B \ll 1$) sperm carrying it rarely
-%encounter heterozygous eggs ($\sim x^2$) but the allele still pays a
-%cost in heterozygous individuals.  By contrast, when the B allele is common ($x_B
-%\gg 0$) encounters between heterozygous eggs and $B$ allele sperm
-%become more common allowing drive to occur this can allow the $B$
-%allele to spread when common. In Supplementary Figure
-%\ref{bistable_additive} we show the unstable equilibrium frequency
-%above which the $B$ allele can invade. This equilibrium frequency was found by a simple
-%parameter search.  \gc{wonder if a bit of work could find this}
-
-\subsubsection*{Model 3. Single locus, paternal genotype dependent drive:}
-In the case when female meiotic drive depends on paternal genotype
-%\paragraph{Dominance} we also explored a model where
-%the genotype of the male who produced the fertilizing sperm affects female drive,
-%instead of just the allelic type of the sperm.
- %We assume a 
- heterozygous female will transmit the $B$ allele 
-  with probabilities  $\frac{1}{2}$,  $d_{het}\geq 1/2 $ or $d_{hom}\geq d_{het}$, 
-%YB RE-PARAMETERIZING  %  $\frac{1}{2}$,  $\frac{1}{2} + h(d-\frac{1}{2}) $ or $d$,  YB RE-PARAMETERIZING 
- when mated with to $AA$, $AB$, or $BB$ males,  respectively. 
-  In this model, genotype frequencies after drive, random mating, and selection are 
-  	$f_{AA}'' =   w_{AA}\left( f_{AA} [2 f_A + f_{AB} ] + f_{AB}^2 [1 - d_{het}] \right)/(2\bar{w})$,
-	$f_{AB}'' =   w_{AB}(f_{AA} [f_B + f_{AB}/2] + f_{AB} f_B + f_{BB} [ f_A - f_{AB} d_{hom}])/\bar{w}$, 
-	and 	$f_{BB}'' =   w_{BB}(f_{AB} (f_{AB} d_{het}/2 + f_{BB} d_{hom}) + f_{BB} f_B)/\bar{w}$. 
-% Above, $d > 1/2$ and $h$ ($0 \geq h \geq 1$) acts to specify the dominance of the $B$
-%allele's effect on female meiotic drive through the males genotype.
-%YB TRY_1 We assume that in a mating between a heterozygous female and an $AA$, $AB$, or $BB$ male, the female will transmit the $B$ allele 
-% with probabilities  $\frac{1}{2}$,  $\frac{1}{2} + h(d-\frac{1}{2}) $, and $d$, respectively.
-  %Here $d > 1/2$ and $h$ ($0 \geq h \geq 1$) acts to specify the dominance of the $B$
-%allele's effect on female meiotic drive through the males genotype.
-% GC ORIGINAL  We assume that if a
-%heterozygous egg ($AB$) is fertilized by a sperm with the following
-%male ($\Mars$) genotypes the $B$ allele will be transmitted through
-%female meiosis with the following probabilities \\
-%\begin{tabular}{cccc}
-%$\Mars$ geno. & AA & AB  & BB\\
-%P(B). & $\frac{1}{2}$ &  $\frac{1}{2} + h(d-\frac{1}{2}) $ & $d$ 
-%\end{tabular}\\
-%where $d > 1/2$ and $h$ ($0 \geq h \geq 1$) acts to specify the dominance of the $B$
-%allele's effect on female meiotic drive through the males genotype. 
-%We set up a set of recursions using a recessive sex independent cost to
-%the driver, similar to those in eqn.
-%\ref{eqn.A1}-\ref{eqn.A1}, but now featuring these male genotype
-%transmission probabilities.
-
-If the cost of drive is fully recessive (i.e. $s_h=0$), assuming HWE, 
-	a rare paternal-genotype-dependent driver invades when 
-	$s\lessapprox (d_{het}-1/2)/(d_{het})$, and when common, this driver fixes if 
-	$s\lessapprox d_{hom}-1/2$, approximations well supported by exact results (Supplementary Figure \ref{Effect_of_dominance}).
-Specifically, when drive in heterozygotes is large relative to that in homozygotes, 
-	\begin{equation} d_{het}  \gtrapprox 1/(3-2d_{hom}) \label{polymale} \end{equation}
-	fixation criteria are more stringent than invasion criteria, 
-	and therefore some values of $s$ can maintain a stable polymorphism. 
-Under these parameter values, a rare paternal-genotype-dependent driver 
-	can increase in frequency because it gains a transmission advantage and suffers 
-	no fitness cost when heterozygous eggs are fertilized by 
-	$A$-bearing sperm of heterozygous males. 
-As the frequency of the $B$ allele increases, 
-	%it occurs more often as a homozygote, and so 
-	it will be unable to 
-	avoid producing unfit homozygous offspring, leaving it trapped in
-	the population at frequency $f_{B\text{ pat}}^*$. 
-Assuming HWE, a recessive fitness cost ($h_s=0$), and dominance of driver ($d_{het}=d_{hom}$) this equilibrium frequency is
-\begin{equation}f_{B\text{ pat}}^*\approx2 (1 + 2d_{hom}[s - 1])/([2 s - 1][2d_{hom}- 1]) \label{eqmale} \end{equation}
-By contrast, when $d_{het} \lessapprox 1/(3-2d_{hom})$ the case is reversed, and the model is bistable.
-%In Supplementary Figure \ref{Effect_of_dominance} we show the invasion
-%criteria for both the A and B alleles when they are rare. 
-%We find a narrow parameter range where B 
-%can invade but not fix.{eqmale}
-%This corresponds to B being less than fully recessive (i.e. $h>0$) and to a modest cost in homozygotes. 
-%This combination of parameters reflects the fact that when the allele is rare, 
-%it can increase in frequency due to fertilization of heterozygous eggs
-%by sperm from heterozygote males, where the sperm do not themselves
-%carry the allele. But as the B allele becomes common it will frequently be
-%present in the homozygous state in males, and so will be unable to
-%avoid the production of unfit homozygote zygotes leaving it trapped in
-%the population.
-
-\subsection*{Models 4-6. Two-locus, sperm-dependent drive.} 
-Assuming that a traditional driver segregates at its equilibrium frequency (Model 1), we investigate the evolution of tightly linked (Models 4 and 5), and unlinked (Model 6) sperm-dependent modifiers of drive.
-In Models 4 and 5, we treat this two-locus system as if it consists of
-a single locus with three alleles: $A$, $B$ and $C$, corresponding to the case when the
-sperm-modifier arises once, tightly linked to one or other of the
-alleles at the drive locus (such that recombination is unexpected). 
-In all cases, we assume that fitness is independent of the drive modifier. %, so, $w_{AA}=1$, $w_{AB}=1-s_h$, and $w_{BB} = 1-s$. 
-We assume the traditional driver is transmitted to $d_0$ of gametes from female heterozygotes  when fertilized by wild-type sperm, and $d_1=d_0+\epsilon$ when fertilized by a sperm-acting drive-modifier.
-
-\subsubsection*{Model 4. Drive-modifier in coupling phase:}
-When the $C$ allele is tightly linked to the driver allele, 
-	genotypic fitnesses equal $w_{AC}=w_{AB}=1-s_h$, and $w_{BC}=w_{CC}=w_{BB}=1-s$. 
-Assuming HWE, a recessive fitness cost to drive, and assuming that the $A/B$ locus is at its equilibrium frequency, the change in frequency of a rare drive modifier is
-\begin{equation}
-	\Delta C\approx -\epsilon f_C \frac{ 1-2 d_0 (-2 d_0+s+2)+\sqrt{2} \sqrt{s (2 d_0 (d_0(s-2)+2)-1)}}{2 \bar{w}(1-2 d_0)^2}
-	\label{coupling}
-\end{equation}
-%YB Considered this ugly math more impressive.
-For all parameters sustaining a polymorphism at the drive locus ($s>d_0-1/2$), this corresponds to a decrease in frequency of the $C$ allele when it enhances drive ($\epsilon >0$ -- the $B^+$ model, above),
-	and an increase in frequency of the $C$ allele when it suppresses drive ($\epsilon <0$ -- the $B^-$ model, above). 
-More generally, even when the cost of drive is not fully recessive, the $B^-$ allele will invade and fix under all parameters sustaining a previous polymorphism at the drive locus (see Mathematical Supplement).
-%When the C allele acts to suppress drive for all parameters sustaining a polymorphism at the drive locus ($s>d_0-1/2$)
-%We address two questions concerning the $C$ allele - can it invade when rare  and can it fix when common?
-%Although our analytical approximations (see our \gc{Mathematical supplement}) do not reveal simple inequalities like those above,
-%	graphical exploration of analytical results show that this mutant is lost when it enhances drive ($\epsilon>0$ - the $B^+$ allele, above),
-%	and spreads from rarity to fixation when it suppresses drive ($\epsilon<0$ - the $B^-$ allele, above).	
-%This result is supported by exact numerical iterations, as well
-%\yb{FIG?}. \gc{Lets not bother}
-
-\subsubsection*{Model 5. Drive-modifier in repulsion phase:}
-When the $C$ allele is tightly linked to the non-driver, 
-	genotypic fitnesses equal $w_{CC}=w_{AC}=w_{AA}=1$, and $w_{BC}=w_{AB}=1-s_h$. 
-Assuming HWE, a recessive fitness cost to drive, and assuming that the $A/B$ locus is at its equilibrium frequency, the change in frequency of a rare drive modifier is
-\begin{equation}
-	\Delta C \approx -\epsilon f_{C^-}^2 \frac{ \left(2 d_0 s-\sqrt{2} \sqrt{s (2 d_0 (d_0 (s-2)+2)-1)}\right)}{2\bar{w}s (2 d_0 -1) }. \label{A-}
-\end{equation}
-For all values of interest ($0<s<1$, $0.5<d_0<1$), the change in frequency a rare $C$ allele is positive when it decreases drive (i.e.
-$\epsilon <0$, corresponding to the $A^-$ model, above), a result which holds qualitatively for a common $C$ allele, as well (Mathematical Supplement).
-%We address two questions concerning the $C$ allele - can it invade when rare  and can it fix when common?
-%Again, while simple inequalities are lacking, 
-%	graphical exploration of our analytical results, and exact numerical iteration \yb{FIG?} shows that this modifier only spreads from rarity to fixation when it suppresses drive ($\epsilon<0$ - the $A^-$ allele, above).
-Somewhat counterintuitively, the spread of the $A^-$ allele is substantially slower than that of the $B^-$ allele (it is on the order of $C^{2}$ rather than $C$). 
-This is because the $A^-$ allele will never occur in drive-homozygotes and therefore only spreads by preventing itself from the cheating of the driver.
-
-\subsubsection*{Model 6. Unlinked drive-modifier:}
-%\gc{Can we use something other than F, as that was HWE deviation}
-For the unlinked model, we introduce another locus where drive is modified in $A/B$ females fertilized by $M$ allele, 
-	while the wild-type $L$ allele does not influence drive. 
-Assuming HWE and linkage equilibrium, the change in frequency of a rare unlinked, sperm-acting drive modifier is 
-\begin{equation} \Delta M =-\epsilon f_A f_B (f_B s + s_h (f_A - f_B) ) \frac{f_M}{\bar{w}}  \label{unlinked} \end{equation}
-% an $F$ allele increases in frequency when 
-%	$-\epsilon f_A f_B (f_B s + s_h (fA - fB) ) >0$, corresponding to any level of drive suppression ($\epsilon<0$) 
-Thus, a rare drive suppressor ($\epsilon<0$) will spread so long as the fitness cost of the driver does not display over- or under-dominance. 
-Notably, this result holds while assuming complete linkage equilibrium, and therefore sperm suppression of female meiotic drive is likely more powerful than more direct female-enforced drive suppressors  which usually spread due to the indirect benefit of avoiding drive haplotypes \citep[e.g.
-][]{Brandvain2012}.
-%The spread of this driver does not, however, depend on assumption of HWE, and linkage equilibrium - exact results resemble analytical approximations. 
-%\paragraph{three allele system}
-%$x_{AA}$, $x_{AB}$,$x_{AB'}$, $w_{BB'} x_{BB'}$, $w_{B'B'} x_{BB}$, $w_{B'B'} x_{B'B'}$
-%We denote the frequency of the three genotypes in the population at birth are
-%$x_{AA}$, $x_{AB}$, $x_{AB'}$, $x_{BB'}$, $x_{BB}$, $x_{B'B'}$. 
-%The mean relative fitness of the population is 
-%\begin{equation}
-%x_{AA}+x_{AB}+ x_{AB'}+ w_{BB'} +x_{BB'}+w_{B'B'}x_{BB} + w_{B'B'} x_{B'B'} 
-%\end{equation}
-%The frequency of the genotypes in meiotic-cells of female adults is 
-%\begin{equation}
-%\frac{1}{\bar{w}}\Big( x_{AA}, ~x_{AB},~ x_{AB'},~ w_{BB'} x_{BB'},~ w_{B'B'}x_{BB},
-%~w_{B'B'} x_{B'B'} \Big)
-%\end{equation}
-%The frequency of fertilization events can be found by combining these
-%female frequencies with the frequency of the three alleles in sperm is
-%\begin{eqnarray}
-%x_A = & x_{AA} + \frac{1}{2} x_{AB}+ \frac{1}{2} x_{AB'} \\
-%x_B = & w_{BB}  x_{BB} +   w_{BB'} \frac{1}{2} x_{BB'}+ \frac{1}{2} x_{AB} \\
-%x_{B'} = & w_{B'B'}  x_{B'B'} +   w_{BB'} \frac{1}{2} + x_{BB'}+ \frac{1}{2} x_{AB'}  
-%\end{eqnarray}
-%in combination with the transmission probabilities given in Figure
-%\ref{Eggsperm_3_allele_cartoon}A.
-
-% \gc{Alternatively we could give the table below:}
-
-% \begin{table}
-% \begin{tabular}{lllll}
-% \hline 
-%   &  \multicolumn{4}{c}{Sperm genotype}   \\
-%   &  & A & B & B'  \\
-% $\Venus$ geno. & Freq.  & $x_{AA} + \frac{1}{2} x_{AB}+ \frac{1}{2} x_{AB'} $ & $ w_{BB}  x_{BB} +   w_{BB'} \frac{1}{2} x_{BB'}+
-% \frac{1}{2} x_{AB}$ & $ w_{B'B'}  x_{B'B'} +   w_{BB'} \frac{1}{2}
-% x_{BB'}+ \frac{1}{2} x_{AB'}  $\\
-% \hline\\
-% AA & $x_{AA}$ &  & & \\
-% AB & $x_{AB}$ &  & & \\
-% AB' & $x_{AB'}$ &  & & \\
-% BB' & $w_{BB'} x_{BB'}$ &  & & \\
-% BB & $w_{B'B'} x_{BB}$ &  & & \\
-% B'B' & $w_{B'B'} x_{B'B'}$ &  & & \\
-% \end{tabular}
-% \caption{ALTERNATIVELY I COULD FILL IN THIS MATRIX  with genotype }
-% \end{table}
-\bibliography{refs}
-
-\newpage
-
-  \setcounter{figure}{0}\global\def\thefigure{S\arabic{figure}}
-
-\section*{Supplementary Material}
-
-\begin{figure}
-
-\includegraphics[width = \textwidth]{../Scripts/homozyg_bistability.eps}
-\caption{Invasion analysis for a self-promoting female meiotic drive allele with
+\section*{SI Figure legends}
+\subsection*{Figure S1}
+Invasion analysis for a self-promoting female meiotic drive allele with
   recessive costs (selection coefficient $s$), showing the region of
@@
   the right we show the approximation obtained assuming that HWE
-  holds. }  
-\label{Bistab_homozyg_cost_fig}
-\end{figure}
+  holds.
 
-\begin{figure}
-\includegraphics[width = 0.8 \textwidth]{Figures/bistable_x_vs_d_additive_s.eps} 
-\caption{The unstable equilibrium frequency for a self-promoting
+\subsection*{Figure S2}
+The unstable equilibrium frequency for a self-promoting
   female meiotic drive allele with an additive cost ($s_h=s/2$) as a
   function of the drive parameter. The solid line shows results
-  obtained using the recursion, the dots our approximation given by Equation \eqref{thresh2}}  \label{bistable_additive}
-\end{figure}
+  obtained using the recursion, the dots our approximation given by Equation 5.
 
-\begin{figure}
-\includegraphics[width = 0.7 \textwidth]{../Scripts/effect_of_dominance_on_invasion_space_four_graph.eps} 
-\caption{Exact results of invasion analysis of an allele whose effect
+\subsection*{Figure S3}
+Exact results of invasion analysis of an allele whose effect
  on female meiosis is mediated by the genotype of the fertilizing
@@
   ($(d_{het}-1/2)/d_{het}$) see the appendix for more details.
-}  \label{Effect_of_dominance}
-\end{figure}
+
+\section*{SI Table legends}
+  \subsection*{Table S1}
+  We detail the expected frequencies of each zygote from a given mating in Table S1. 
+  Table S1 is an Excel Spreadsheet with six tabs - each one corresponding to Models 1-6. 
+  Table S1 can be used to generate genotype frequencies after drive and selection, 
+  by summing the  frequency of zygotes produced by a given mating weighted by the frequency of this mating. 
+  Assuming random mating this weighting is simply the product of maternal and paternal genotype frequencies after selection in the previous generation.
+
+\section*{SI Appendix legends}
+  \subsection*{Appendix S1}
+A PDF of the Mathematica file in which we derived our analytical results and approximations.
 

### 2014-05-21 writeup_and_notes/signif_statement.tex
yb got sig statement down to 149 words, still need to cut 30

--- 2014-05-19 writeup_and_notes/signif_statement.tex
@@
-Evolutionary biologists are fascinated by the oppotunities for benefits, costs, and opportunities for conflict during meiosis, however,little attention has been directed at the odd fact that fertilization by sperm is required to complete female meiosis in many species of animal.
-At first sight this seems strange as it seems to presents a dangerous opportunity for selfish alleles in sperm to distort the outcome of female meiosis in their own favor creating the opportunity for conflicts between male and female genomes.
-In this paper, through mathematical modeling, we show that in fact quite the opposite is true, sperm are usually selected to ensure the fairness of meiosis.
-Intuitively this comes about because selfish sperm have to immediately suffer any ill consequences of their actions in the resulting zygote, and so it is strongly in sperm's interests to keep genomes from gaming female meiosis.
-We hypthosize that the wide-spread requirement of fertilization to complete female meiosis may represent a long-term collaboration between male and female gametes to ensure fairness and the production of high quality gametes.
+\documentclass{article}
+\usepackage{graphicx}
+
+\begin{document}
+Female meiosis is the time when alleles can directly modify their transmission.
+Benefits to alleles that succeeds in this competition and collateral damage to organismal fitness likely shaped basic features of female reproduction.
+However, the observation that fertilization is often required for the completion of female meiosis appears incompatible with this view.
+In fact, this fertilization requirement seems to presents an opportunity for sperm-acting alleles to distort the in their own favor, creating an additional arena for conflict.
+We show that, counterintuitively, the opposite is true, sperm are usually selected to ensure that female meiosis is fair.
+Intuitively this comes about because sperm immediately suffer the company they keep in the zygote, and it is in sperm's interests to enforce fairness in female meiosis.
+The wide-spread requirement of fertilization to complete female meiosis may represent a long-term collaboration between sperm and mothers to ensure the production of fit offspring.
+\end{document}

### 2014-05-21 writeup_and_notes/signif_statement.tex
edited signif down to 119

@@
 \begin{document}
-Female meiosis is the time when alleles can directly modify their transmission.
-Benefits to alleles that succeeds in this competition and collateral damage to organismal fitness likely shaped basic features of female reproduction.
-However, the observation that fertilization is often required for the completion of female meiosis appears incompatible with this view.
-In fact, this fertilization requirement seems to presents an opportunity for sperm-acting alleles to distort the in their own favor, creating an additional arena for conflict.
-We show that, counterintuitively, the opposite is true, sperm are usually selected to ensure that female meiosis is fair.
-Intuitively this comes about because sperm immediately suffer the company they keep in the zygote, and it is in sperm's interests to enforce fairness in female meiosis.
-The wide-spread requirement of fertilization to complete female meiosis may represent a long-term collaboration between sperm and mothers to ensure the production of fit offspring.
+Meiosis is the time when alleles compete to ensure their transmission at the expense of organismal fitness, and many features of meiosis may have evolved to reduce competition.
+The observation that fertilization is often required for the completion of female meiosis appears incompatible with this view; sperm-acting alleles could distort female meiosis in favor, creating an additional conflict.
+Counterintuitively, we show that sperm are selected to ensure a fair female meiosis.
+This comes about because sperm immediately suffer the result of cheating, and so it is in sperm's interests to enforce fairness in female meiosis.
+The requirement of fertilization to complete female meiosis may represent a long-term collaboration between sperm and mothers to ensure the production of fit offspring.
 \end{document}

### 2014-05-21 writeup_and_notes/signif_statement.tex
yb [hopefully] improved  gcs sig statement while keeping it at 119 words

@@
 \begin{document}
-Meiosis is the time when alleles compete to ensure their transmission at the expense of organismal fitness, and many features of meiosis may have evolved to reduce competition.
-The observation that fertilization is often required for the completion of female meiosis appears incompatible with this view; sperm-acting alleles could distort female meiosis in favor, creating an additional conflict.
-Counterintuitively, we show that sperm are selected to ensure a fair female meiosis.
-This comes about because sperm immediately suffer the result of cheating, and so it is in sperm's interests to enforce fairness in female meiosis.
+During female meiosis, alleles can take charge of their transmission, and those that bias meiosis in their favor can spread at a cost to organismal fitness.
+It is therefore believed that basic features of oogenesis likely evolved to control such renegade genes.
+The observation that female meiosis requires fertilization for its completion appears incompatible with this view; sperm-acting alleles could help like alleles distort  female meiosis in their favor, creating an additional conflict.
+Counterintuitively, we show that sperm succeed by enforcing fairness in females meiosis because this allows them to avoid pairing with low-fitness cheaters.
 The requirement of fertilization to complete female meiosis may represent a long-term collaboration between sperm and mothers to ensure the production of fit offspring.

### 2014-05-21 writeup_and_notes/pnasSubmission/PNASSubmission.TEX
yb [hopefully] included modified sig statement in pnas submission

@@
 %%% If significance statement need, then can use the below command otherwise just delete it.
-\significancetext{Female meiosis is the time during development where an allele can act to directly modify its representation in the next generation. 
-The potential benefits to an allele that succeeds in this competition and collateral damage to organismal fitness are thought to have shaped basic features of female reproduction. 
-However, little attention has been directed at the odd fact that fertilization by sperm is required to complete female meiosis in many species of animal. 
-At first sight this seems strange as it seems to presents a dangerous opportunity for selfish alleles in sperm to distort the outcome of female meiosis in their own favor creating the opportunity for conflicts between male and female genomes.
-In this paper, through mathematical modeling, we show that in fact quite the opposite is true, sperm are usually selected to ensure the fairness of meiosis. 
-Intuitively this comes about because selfish sperm have to immediately suffer any ill consequences of their actions in the resulting zygote, and so it is strongly in sperm's interests to keep genomes from gaming female meiosis.
-We hypothesize that the wide-spread requirement of fertilization to complete female meiosis may represent a long-term collaboration between male and female gametes to ensure fairness and the production of high quality gametes.}
+\significancetext{During female meiosis, alleles can take charge of their transmission, and those that bias meiosis in their favor can spread at a cost to organismal fitness.
+It is therefore believed that basic features of o\"{o}genesis likely evolved to control such renegade genes.
+The observation that female meiosis requires fertilization for its completion appears incompatible with this view; sperm-acting alleles could help like alleles distort  female meiosis in their favor, creating an additional conflict.
+Counterintuitively, we show that sperm succeed by enforcing fairness in females meiosis because this allows them to avoid pairing with low-fitness cheaters.
+The requirement of fertilization to complete female meiosis may represent a long-term collaboration between sperm and mothers to ensure the production of fit offspring.}
 

### 2014-05-21 writeup_and_notes/pnasSubmission/PNASSubmission.TEX
yb changed replaced refs to Table S1 with refs to Dataset S1 because datasets and not tables can be Excel format

@@
 To depict a case of a self-promoting meiotic driver,  we modify this standard model such 
-	that the driver is only effective when fertilized by a sperm carrying that allele (see Figure 1A and Model 2 	in the Appendix and Table S1). 
+	that the driver is only effective when fertilized by a sperm carrying that allele (see Figure 1A and Model 2 	in the Appendix and Dataset S1). 
 We then identify the conditions allowing for the spread of this self-promoting driver, 
@@
 The $B^+$ allele acts in sperm to increase the effectiveness of drive for both
-	the $B$ and  $B^{+}$ alleles in $AB$ and $AB^{+}$ heterozygotes (see Figure 2A2, and Model 4 in the Appendix and Table S1). 
+	the $B$ and  $B^{+}$ alleles in $AB$ and $AB^{+}$ heterozygotes (see Figure 2A2, and Model 4 in the Appendix and Dataset S1). 
 Naively, $B^{+}$ may spread by capitalizing on the additional drive it causes; however,  
@@
 Transmission rules describing the outcomes of all matings in each
-	model are presented in Table S1. 
+	model are presented in Dataset S1. 
 The fitness of genotype, $g$ is sex-independent and equals $1$, $1-h_s$, and $1-s$ for genotypes $AA$, $AB$, and $BB$, respectively. 

### 2014-05-21 writeup_and_notes/pnasSubmission/Text_S1.tex
yb changed replaced refs to Table S1 with refs to Dataset S1 because datasets and not tables can be Excel format

@@
 \section*{SI Table legends}
-  \subsection*{Table S1}
-  We detail the expected frequencies of each zygote from a given mating in Table S1. 
-  Table S1 is an Excel Spreadsheet with six tabs - each one corresponding to Models 1-6. 
-  Table S1 can be used to generate genotype frequencies after drive and selection, 
+  \subsection*{Dataset S1}
+  We detail the expected frequencies of each zygote from a given mating in Dataset S1. 
+ Dataset S1 is an Excel Spreadsheet with six tabs - each one corresponding to Models 1-6. 
+ Dataset S1 can be used to generate genotype frequencies after drive and selection, 
   by summing the  frequency of zygotes produced by a given mating weighted by the frequency of this mating. 

### 2014-05-22 writeup_and_notes/male_dependent_drive.tex
yb fixed numerous typos [in both versions]... perhaps too late

@@
 It is also not the case that sperm is mechanistically incapable of influencing female meiosis.
-In \emph{C. elegans}, premature deployment of the aster, a vital component of mitotic
+In \emph{C. elegans}, premature deployment of the aster -- a vital component of mitotic
 machinery provided by the sperm, disrupts MII meiotic segregation
@@
 We then examine models in which a novel sperm-acting allele modifies the efficacy of a polymorphic meiotic driver. 
-Such model universally reveal that sperm-acting drive modifiers are favored only if the suppress drive. 
+Such models universally reveal that sperm-acting drive modifiers are favored only if they suppress drive. 
 These results highlight the fact that the interests of sperm and maternal  genomes' are often aligned, as both are invested in the fate of the resultant zygote (as was speculated for the \emph{In} locus \citep{Pomiankowski1993}).
@@
 	\citep[see e.g. ][ for additional results]{Prout1973,Ubeda2004}. 
-Assuming the driving allele is deleterious in both sexes, but fully recessive 
+Assuming that the driving allele is deleterious in both sexes, but fully recessive 
 	(i.e. the fitness of drive homozygotes equals $w_{BB}=1-s$ and other genotypic fitnesses equal $w_{AA}=w_{AB}=1$), 
-	it always invades because,  when rare, it occurs predominantly in heterozygotes and therefore drives without a fitness cost. 
-However, when $s$ is large ($s>(2d-1)/(2)$, solid black line in Figure \ref{Eggsperm}B), a driver cannot fix, and
+	it always invades because,  when rare it occurs predominantly in heterozygotes and therefore drives without a fitness cost. 
+However, when $s$ is large ($s>(2d-1)/(2)$, solid black line in Figure \ref{Eggsperm}B) a driver cannot fix and
 will be maintained as a protected polymorphism \citep{Prout1973}. 
@@
 	%, the selection against the drive homozygote, 
-	is approximately less than $(2 d - 1)/(4 d)$ -- see dashed black line in Figure \ref{Eggsperm}B). 
+	is less than approximately $(2 d - 1)/(4 d)$ -- see dashed black line in Figure \ref{Eggsperm}B. 
 This analytical approximation, derived from Equation \eqref{deltadriver} assuming Hardy-Weinberg, 
@@
 
-When a self-promoting driver does spread, 
-	it spends much time at low frequency, 
+When a self-promoting driver does spread 
+	it spends much of its time at low frequency, 
 	because the paucity of complimentary sperm compromises its ability to drive. 
@@
 %	allele that suppresses this form of meiotic drive in females.
-The positive frequency dependence of a self-promoting driver can also
+This positive frequency dependence can 
 	induce bistability in its spread -- some values of $s$ 
@@
 	preclude its invasion when rare (Equation \ref{ineqA} and Figure \ref{Eggsperm}B). 
-In this case, the driver will be fixed if it reaches a frequency greater than some threshold 
+In this case, the driver will be fixed if its frequency exceeds some threshold 
 	(approximated in Equation \ref{bistab_homozy_approx} and  presented exactly in 
@@
 For most parameters, this threshold is likely too high to be reached by drift, 
-	and therefore  the fate of self-promoting drivers is determined
+	and therefore  the fate of a self-promoting driver is determined
 	by the more restrictive invasion criteria rather than the fixation criteria.
@@
 	effect on female meiosis ($d_{het}$ is close to $d_{hom}$), 
-	there is a a narrow sliver of parameter space sustains a polymorphism
+	a narrow sliver of parameter space sustains a polymorphism
 	when the cost of the drive is recessive
@@
 	the lack of biological reality underlying the small portion of parameter values 
-	required for this polymorphism make us doubt it general applicability. 
+	required for this polymorphism make us doubt its general applicability. 
 %Thus we regard this case as interesting but not very general.
@@
 %%%sperm-acting facilitator
-When the modifier of drive is tightly linked to the driver and
-enhances drive (i.e. in coupling phase), 
+When the modifier of drive arises on the drive background (i.e. in coupling phase),
+	is tightly linked to the driver, and enhances drive 
 	we label  this non-recombining drive/modifier haplotype as the $B^+$ allele.  
@@
 Imagine a sperm-acting allele that restores fairness to female meiosis arises on
-	the drive background, i.e. in repulsion phase, creating a
+	the drive background, creating a
         third allele $B^{-}$ (Figure
@@
 If this sperm-acting drive suppressor arises on 
-	the non-driving $A$ background (creating a third allele $A^{-}$, Figure \ref{Eggsperm_3_allele_cartoon}A4, Model 5), 
+	the non-driving $A$ background ( i.e. in repulsion phase, 
+	creating a third allele $A^{-}$, Figure \ref{Eggsperm_3_allele_cartoon}A4, Model 5), 
 	or is unlinked to the drive locus (Model 6), it readily invades a population segregating
@@
 This likely explains why females can delay the completion of meiosis until after fertilization 
-	without risking exploitation by collaborations by female
+	without risking exploitation by collaborations between female
         drivers and sperm alleles.
@@
 Why is it that an allele that biases female meiosis in its favor can generate a genetic conflict, but an allele in sperm that assists this female driver cannot? 
-So long as the transmission advantage of female meiotic drive outweigh the organismal fitness cost to heterozygotes the female driver can spread when rare, and it increases in 		
-	frequency until the fitness cost experienced when homozygous balances the transmission advantage.
+So long as the transmission advantage of female meiotic drive outweigh the organismal fitness cost to heterozygotes, the female driver can spread when rare, and it increases in 		
+	frequency until the fitness cost to homozygotes balances the transmission advantage.
 By contrast, a sperm promoter of female drive is only effective when matched with a heterozygote female -- meaning that, when rare, this allele rarely enhances female drive. 
@@
 % YB - WE DID NOT TALK ABOUT THIS IN THE RESULTS [ALTHOUGH IT'S TRUE... PERHAPS TOO DISTRACTING]
-More generally, alleles that act through sperm to
+More generally, natural selection favors alleles that act through sperm to
  reduce the opportunity of female meiotic drive regardless of linkage or phase.
 
-Our theory offers a novel interpretation of a baffling idiosyncrasy of female meiosis -- 
-	we argue that females do not complete meiosis until after fertilization as a mechanism to encourage meiotic fairness. 
+Our theory offers a novel interpretation of a baffling idiosyncrasy of female meiosis in many animals -- the requirement of fertilization before fertilization is complete. 
+We argue that females do not complete meiosis until after fertilization as a mechanism to encourage meiotic fairness.
 In addition to providing a new interpretation of a basic biological process, our theory provides numerous testable predictions.
@@
 	complimentary sperm. 
-Therefore, For example, MII female drive should be exceptionally severe in
+Therefore, for example, MII female drive should be exceptionally severe in
 	interspecific crosses between animals which complete the second meiotic division after fertilization,  
@@
 Our theory also encourages phylogenetic hypotheses concerning the relationship between the opportunity 
-	for female meiotic drive the requirement of fertilization for the completion of female meiosis. 
+	for female meiotic drive and the requirement of fertilization for the completion of female meiosis. 
 For example, we predict that a lower opportunity for female meiotic drive,
@@
         or selfing, 
-        would often accompanied by a relaxation of the requirement of fertilization for the
+        would often be accompanied by a relaxation of the requirement of fertilization for the
         completion of female meiosis.
@@
 \subsubsection*{Model 3. Single locus, paternal genotype dependent drive:}
-In the case when female meiotic drive depends on paternal genotype
+In the case when female meiotic drive depends on paternal genotype,
 %\paragraph{Dominance} we also explored a model where
@@
  %We assume a 
- heterozygous female will transmit the $B$ allele 
+ a heterozygous female will transmit the $B$ allele 
   with probabilities  $\frac{1}{2}$,  $d_{het}\geq 1/2 $ or $d_{hom}\geq d_{het}$, 

### 2014-05-22 writeup_and_notes/pnasSubmission/PNASSubmission.TEX
yb fixed numerous typos [in both versions]... perhaps too late

@@
 We then examine models in which a novel sperm-acting allele modifies the efficacy of a polymorphic meiotic driver. 
-Such model universally reveal that sperm-acting drive modifiers are favored only if the suppress drive. 
+Such models universally reveal that sperm-acting drive modifiers are favored only if they suppress drive. 
 These results highlight the fact that the interests of sperm and maternal  genomes' are often aligned, as both are invested in the fate of the resultant zygote (as was speculated for the \emph{In} locus \cite{Pomiankowski1993}).
@@
 	(see e.g. \cite{Prout1973,Ubeda2004} for additional results). 
-Assuming the driving allele is deleterious in both sexes, but fully recessive 
+Assuming that the driving allele is deleterious in both sexes, but fully recessive 
 	(i.e. the fitness of drive homozygotes equals $w_{BB}=1-s$ and other genotypic fitnesses equal $w_{AA}=w_{AB}=1$), 
-	it always invades because, when rare, it occurs predominantly in heterozygotes and therefore drives without a fitness cost. 
-However, when $s$ is large ($s>(2d-1)/(2)$, solid black line in Figure 1B), a driver cannot fix, and
+	it always invades because, when rare it occurs predominantly in heterozygotes and therefore drives without a fitness cost. 
+However, when $s$ is large ($s>(2d-1)/(2)$, solid black line in Figure 1B) a driver cannot fix and
 will be maintained as a protected polymorphism \cite{Prout1973}. 
@@
 When rare, a fully recessive, self-promoting driver can only invade when $s$ 
-	is approximately less than $(2 d - 1)/(4 d)$ -- see dashed black line in Figure 1B. 
+	is less than approximately $(2 d - 1)/(4 d)$ -- see dashed black line in Figure 1B. 
 This analytical approximation, derived from Equation \eqref{deltadriver} assuming Hardy-Weinberg, 
@@
 	positive frequency dependent behavior (Figure 1B.1).  
-The positive frequency dependence of a self-promoting driver can also
+This positive frequency dependence can 
 	induce bistability in its spread -- some values of $s$ 
@@
 	preclude its invasion when rare (Equation \ref{ineqA} and Figure 1B). 
-In this case, the driver will be fixed if it reaches a frequency greater than some threshold 
+In this case, the driver will be fixed if its frequency exceeds some threshold 
 	(approximated in Equation \ref{bistab_homozy_approx} and  presented exactly in 
@@
 	the lack of biological reality underlying the small portion of parameter values 
-	required for this polymorphism make us doubt it general applicability.
+	required for this polymorphism make us doubt its general applicability.
 
@@
 Imagine a sperm-acting allele that restores fairness to female meiosis arises on
-	the drive background, i.e. in repulsion phase, creating a
+	the drive background, creating a
         third allele $B^{-}$, (Figure 2A3, Model 4 in Appendix). 
@@
 If this sperm-acting drive suppressor arises on 
-	the non-driving $A$ background (creating a third allele $A^{-}$, Model 5), 
+	the non-driving $A$ background (i.e. in repulsion phase, 
+	creating a third allele $A^{-}$, Figure \ref{Eggsperm_3_allele_cartoon}A4, Model 5), 
 	or is unlinked to the drive locus (Model 6), it readily invades a population segregating
@@
 of forming low fitness homozygotes. 
-More generally, alleles that act through sperm to
+More generally, natural selection favors alleles that act through sperm to
  reduce the opportunity of female meiotic drive regardless of linkage or phase.
 
-Our theory offers a novel interpretation of a baffling idiosyncrasy of female meiosis -- 
-	we argue that females do not complete meiosis until after fertilization as a mechanism to encourage meiotic fairness. 
+Our theory offers a novel interpretation of a baffling idiosyncrasy of female meiosis in many animals -- the requirement of fertilization before fertilization is complete. 
+We argue that females do not complete meiosis until after fertilization as a mechanism to encourage meiotic fairness. 
 In addition to providing a new interpretation of a basic biological process, our theory provides numerous testable predictions. 
@@
 \subsubsection{Model 3. Single locus, paternal genotype dependent drive:}
-In the case when female meiotic drive depends on paternal genotype
- heterozygous female will transmit the $B$ allele 
+In the case when female meiotic drive depends on paternal genotype,
+ a heterozygous female will transmit the $B$ allele 
   with probabilities  $\frac{1}{2}$,  $d_{het}\geq 1/2 $ or $d_{hom}\geq d_{het}$, 

### 2014-05-26 writeup_and_notes/male_dependent_drive.tex
fixing typos found by Jeremy

@@
 	creating ample opportunity for interaction between the sperm and
-	female meiotic machinery (note that the alternation of generations in
+	female meiotic machinery (note that the alternation of generations
 	likely precludes this interaction in plants). 
@@
 It is therefore unlikely
-	that female oogenesis and meiosis will evolve to prevent their effect.  
+	that female o\"{o}genesis and meiosis will evolve to prevent their effect.  
 This likely explains why females can delay the completion of meiosis until after fertilization 
@@
 
-Our theory offers a novel interpretation of a baffling idiosyncrasy of female meiosis in many animals -- the requirement of fertilization before fertilization is complete. 
+Our theory offers a novel interpretation of a baffling idiosyncrasy of female meiosis in many animals -- the requirement of fertilization before meiosis is complete. 
 We argue that females do not complete meiosis until after fertilization as a mechanism to encourage meiotic fairness.

### 2014-05-31 writeup_and_notes/male_dependent_drive.tex
Fixed additional typos

@@
 	it spends much of its time at low frequency, 
-	because the paucity of complimentary sperm compromises its ability to drive. 
+	because the paucity of complementary sperm compromises its ability to drive. 
 However, once relatively common, it  rapidly achieves fixation due to its
 	positive frequency dependent behavior (Figure \ref{Eggsperm}B.1).  
-%This lack of a balanced polymorphism %(Figure \ref{Invasion_space}) 
-%	precludes the evolution of an
-%	allele that suppresses this form of meiotic drive in females.
 This positive frequency dependence can 
@@
 	Figure \ref{bistable_additive}).
-%	$f_B^*$
-%	because it pays the  heterozygous fitness cost but drives
- %      ineffectively (see Figure S\ref{bistable_additive})
-%While a standard female driver invades from low frequency when the transmission advantage outweighs the fitness cost (XXX),
-%	a rare self-promoter pays a fitness cost, but drives only weakly because complementary sperm are uncommon. 
-%This frequency-independent fitness cost, and positive frequency dependent drive generates a bistable system. 
-%Below a cutoff frequency the self-promoter driver deterministically decreases in frequency 
-%	because it pays the  heterozygous fitness cost but drives
- %      ineffectively (see Figure S\ref{bistable_additive}). 
-%Above this frequency, the allele increases deterministically because the commonness of complementary sperm 
-%	make drive effective enough to overcome the associated fitness cost.
 This bistability prevents self-promoting drivers from invading 	
@@
         experiences female drive when fertilized by A or B sperm, but 
-       is does not drive when fertilized by another $B^{-}$ so it
+       it does not drive when fertilized by another $B^{-}$ so it
        avoid the excess formation of low fitness homozygotes. This
@@
 	should be reduced when fertilized by
-	complimentary sperm. 
-Therefore, for example, MII female drive should be exceptionally severe in
+	complementary sperm. 
+Therefore, MII female drive should be exceptionally severe in
 	interspecific crosses between animals which complete the second meiotic division after fertilization,  
-	as heterospecific sperm will not specifically evolve to resists a female driver in another species. 
+	as heterospecific sperm will not specifically evolve to resist a female driver in another species. 
 Our theory also encourages phylogenetic hypotheses concerning the relationship between the opportunity 
@@
 
-%More generally both linked and unlinked alleles may often spread that act through sperm to
-%reduce the opportunity of female meiotic drive in order to disable
-%segregating drive systems, by for example reducing the mechanistic
-%asymmetries that female drivers exploit.
-
 Our results highlight potentially counterintuitive results of complex genetic conflicts. 
@@
 	the interests of fertilizing sperm and female are quite well aligned during syngamy.
-While conflict between mother and her alternative chromosomes ensue, 
+While conflict between mother and her alternative chromosomes ensues, 
 	fertilizing sperm decidedly side with mom, as both have a shared interest in producing a viable and potentially fit offspring.
-%As noted above there are even opportunities for coevolution between sperm-based suppressors of female meiotic drive and the female genome.
-%Thus it seems plausible that aspects of oogenesis may actively evolve to facilitate the influence of sperm on ensuring a fair  outcome of female meiosis which generates fit offspring.
-%In light of that it is interesting to consider whether the phylogenetic variation in the stage of oogenesis when fertilization takes place reflects not
-%	only species' different reproductive ecologies, but also
-%        different selection pressures shaping the interaction of sperm
-%        alleles with female meiosis. 
-%\gc{For example, we would predict that a lower the opportunity for female meiotic drive,
-%        e.g. an evolutionary transition to inbreeding or a high
-%        selfing rate in animals, would often accompanied by a
-%        relaxation of the requirement of fertilization for the
-%        completion of female meiosis.}
 

### 2014-05-31 writeup_and_notes/PLoSBioSubmission/CoverLetter.tex
added PLoS Bio folder, cover letter and ms

%%!TEX TS-program = latex
% Template for PLoS
% Version 1.0 January 2009
%
% To compile to pdf, run:
% latex plos.template
% bibtex plos.template
% latex plos.template
% latex plos.template
% dvipdf plos.template

\documentclass[10pt]{article}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

\usepackage{color}

% Use doublespacing - comment out for single spacing
%\usepackage{setspace} 
%\doublespacing

% Text layout
\topmargin 0.0cm
\oddsidemargin 0.5cm
\evensidemargin 0.5cm
\textwidth 16cm 
\textheight 21cm

% Bold the 'Figure #' in the caption and separate it with a period
% Captions will be left justified
\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}

% Use the PLoS provided bibtex style
\bibliographystyle{plos2009}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}

\pagestyle{myheadings}
%% ** EDIT HERE **

%% ** EDIT HERE **
%% PLEASE INCLUDE ALL MACROS BELOW

%% END MACROS SECTION

\begin{document}

% Title must be 150 characters or less
\begin{flushleft}
\textbf{Cover Letter for: Sperm should evolve to make female meiosis fair.}
% Insert Author names, affiliations and corresponding author email.
\newline by Yaniv Brandvain$^{\ast}$ and  Graham Coop \\

\end{flushleft}

We thank you for consideration of our manuscript \emph{Sperm should evolve to make female meiosis fair} for publication in \emph{PLoS Biology}. 
\\

The requirement of fertilization for the completion of female meiosis in many animal species presents an opportunity for sperm to interfere with the outcome of female meiosis. We employ population genetic theory to show that sperm that enforce fairness in female meiosis are nearly universally favored. As such, we offer a novel interpretation of the requirement of fertilization for the completion of female meiosis ? this requirement potentially represent a collaboration between mother and sperm in the control of fitness reducing meiotic drivers. The application of theory to a basic and unexplained feature of organismal biology, and the interplay between conflict and co-option by numerous parties involved in fertilization make this paper of broad interest to a wide-range of biologists.
\\

The broad appeal of this paper is evidenced by the reception of our bioR$\chi$iv -- preprint (link: http://biorxiv.org/content/early/2014/05/22/005363) which has been downloaded viewed more than 450 times since its submission 10 days ago, and whose Altmetric score places it above the 90$^{th}$ percentile for similar aged articles (link http://www.altmetric.com/details.php?domain=biorxiv.org\&citation\_id=2368591).
\\

We note that while we submitted this as a `Research article,' we are open to reformatting it as an `Unsolved mystery' if the editorial board feels that is more appropriate. 
\\
\\ 
\begin{flushright}
Sincerely,\\
Yaniv Brandvain \\ \& Graham Coop\\
May 31$^{st}$ 2014
\end{flushright}
\end{document}

### 2014-05-31 writeup_and_notes/PLoSBioSubmission/PLoSBio_male_dependent_drive.tex
added PLoS Bio folder, cover letter and ms

--- 2014-05-31 writeup_and_notes/male_dependent_drive.tex
@@
 %%!TEX TS-program = latex
-\documentclass[12pt,letterpaper]{article}
-\usepackage{epsfig}                 
-\usepackage[authoryear]{natbib}
+% Template for PLoS
+% Version 1.0 January 2009
+%
+% To compile to pdf, run:
+% latex plos.template
+% bibtex plos.template
+% latex plos.template
+% latex plos.template
+% dvipdf plos.template
+
+\documentclass[10pt]{article}
+
+% amsmath package, useful for mathematical formulas
+\usepackage{amsmath}
+% amssymb package, useful for mathematical symbols
+\usepackage{amssymb}
+
+% graphicx package, useful for including eps and pdf graphics
+% include graphics with the command \includegraphics
 \usepackage{graphicx}
-\usepackage{amsmath}
-\usepackage{psfrag}
-\usepackage{mathabx}
-\renewcommand{\baselinestretch}{1.6}
-\large
-\pagenumbering{arabic}
-\usepackage[usenames,dvipsnames]{color}
-\usepackage{fullpage}
-\bibliographystyle{genetics}
-\usepackage{multirow}
-\newcommand{\ye}{\hat{y}}
-\newcommand{\xe}{\hat{x}}
+
+% cite package, to clean up citations in the main text. Do not remove.
+\usepackage{cite}
+
 \usepackage{color}
-\usepackage[normalem]{ulem}  
-	\newcommand{\gc}[1]{{ \color{red} #1}}
-	\newcommand{\yb}[1]{{ \color{blue} #1}}
 
-%Why do animals allow male control of female meiosis?
-%Why does female meiosis allow the opportunity for exploitation by selfish sperm?
-%sperm and egg genomes both favour the peaceful resolution of female meiotic drive
-%strange bed fellows
+% Use doublespacing - comment out for single spacing
+%\usepackage{setspace} 
+%\doublespacing
 
-%\title{Why does female meiosis allow the opportunity for exploitation by selfish sperm? BETTER TITLE?}
-\title{
-%Co-operation between the sexes trumps selfish control of female  meiotic drive. 
-%Or 
-Sperm should evolve to make female meiosis fair.
-%Sperm do not evolve to collaborate in female meiotic drive. 
-%Or Why does female meiosis allow the opportunity for exploitation by selfish sperm? 
+% Text layout
+\topmargin 0.0cm
+\oddsidemargin 0.5cm
+\evensidemargin 0.5cm
+\textwidth 16cm 
+\textheight 21cm
+
+% Bold the 'Figure #' in the caption and separate it with a period
+% Captions will be left justified
+\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}
+
+% Use the PLoS provided bibtex style
+\bibliographystyle{plos2009}
+
+% Remove brackets from numbering in List of References
+\makeatletter
+\renewcommand{\@biblabel}[1]{\quad#1.}
+\makeatother
+
+% Leave date blank
+\date{}
+
+\pagestyle{myheadings}
+%% ** EDIT HERE **
+
+%% ** EDIT HERE **
+%% PLEASE INCLUDE ALL MACROS BELOW
+
+%% END MACROS SECTION
+
+\begin{document}
+
+% Title must be 150 characters or less
+\begin{flushleft}
+{\Large
+\textbf{Sperm should evolve to make female meiosis fair.}
 }
+% Insert Author names, affiliations and corresponding author email.
+\\
+Yaniv Brandvain$^{1,\ast}$, 
+Graham Coop$^{2}$, 
+\\
+\bf{1} Yaniv Brandvain Department of Plant Biology, University of Minnesota --€“ Twin Cities, St, Paul, MN, USA
+\\
+\bf{2} Graham Coop Center for Population Biology \& Department of Evolution and Ecology, University of California -- Davis. Davis, CA, USA
+\\
+$\ast$ E-mail: ybrandva@umn.edu
+\end{flushleft}
 
-\author{Yaniv Brandvain$^1$ \\ email: ybrandvain@gmail.com  \and Graham Coop$^2$ \\ email: gmcoop@ucdavis.edu }
-
-\usepackage{fancyhdr}
-\pagestyle{fancy}
-\lhead{}
-\rhead{}
-\renewcommand{\headrulewidth}{0.0pt}
-\rfoot{}
-\cfoot{\thepage}
-\date{}
-\bibliographystyle{plain}
-\begin{document}
-\maketitle
-\begin{center}
-$^1$Department of Plant Biology, \\ University of Minnesota --€“ Twin
-Cities. St. Paul MN, 55108 \\
-$^2$Center for Population Biology \& Department of Evolution and Ecology \\ University of California -- Davis. Davis, CA, 95616
-\end{center}
-
-%FIgures 
-{\bf Abstract:}
+% Please keep the abstract between 250 and 300 words
+\section*{Abstract}
 Genomic conflicts arise when an allele gains an evolutionary advantage at a cost to organismal fitness. 
-%Because only one of four products of female meiosis is transmitted to offspring, 
 O\"{o}genesis is inherently susceptible to such conflicts
 because alleles compete to be the  product of female meiosis transmitted to the egg.   
-%(meiotic drive).  
 Alleles that distort meiosis in their favor (i.e. meiotic drivers) often decrease organismal fitness,
-	and therefore indirectly favor the evolution of mechanisms to suppress meiotic drive. % drive the rapid evolution of suppression mechanisms, and
+	and therefore indirectly favor the evolution of mechanisms to suppress meiotic drive. 
 In this light, many facets of  o\"{o}genesis and gametogenesis have been 
@@
 	potentially providing an opportunity for sperm alleles to
-        meddle with its outcome and help like-alleles drive in heterozygous females? %and facilitate meiotic drive? 
+        meddle with its outcome and help like-alleles drive in heterozygous females? 
 The population genetic theory presented herein suggests that
 	sperm nearly always evolve to increase the fairness of female
-        meiosis in the face of genomic conflicts. %%felt this
-                                %%might be an important addition as
-                                %%they will spread if no cost to organism
+        meiosis in the face of genomic conflicts. 
 These results are consistent with current knowledge of sperm-dependent meiotic drivers 
@@
         completion of female meiosis
-        %, and variation in stage of meiosis that
-        %fertilization occurs at 
         potentially represents a mechanism employed by females 
-	 to ensure a fair meiosis. %prevent meiotic drive.
+	 to ensure a fair meiosis.
+
+% Please keep the Author Summary between 150 and 200 words
+% Use first person. PLoS ONE authors please skip this step. 
+% Author Summary not valid for PLoS ONE submissions.   
+\section*{Author Summary}
+During female meiosis, alleles can take charge of their transmission, and those that bias meiosis in their favor can spread at a cost to organismal fitness.
+It is therefore believed that basic features of o\"{o}genesis likely evolved to control such renegade genes.
+The observation that female meiosis requires fertilization for its completion appears incompatible with this view; sperm-acting alleles could help like alleles distort female meiosis in their favor, creating an additional conflict.
+Counterintuitively, we show that sperm succeed by enforcing fairness in females meiosis because this allows them to avoid pairing with low-fitness cheaters.
+The requirement of fertilization to complete female meiosis may represent a long-term collaboration between sperm and mothers to ensure the production of fit offspring.
+
 \newpage
-
 \section*{Introduction}
@@
 gain an evolutionary advantage at a cost to  individual fitness
-\citep{Burt2006}, often by exploiting meiosis and gametogenesis.
-Because only one of the four products of female meiosis is transmitted to the egg, female meiosis is particularly vulnerable to such exploitation \citep{Sandler1957,Pardo-ManuelDeVillena2001a}. 
+\cite{Burt2006}, often by exploiting meiosis and gametogenesis.
+Because only one of the four products of female meiosis is transmitted to the egg, female meiosis is particularly vulnerable to such exploitation \cite{Sandler1957,Pardo-ManuelDeVillena2001a}. 
 An allele that biases female meiosis in its favor (i.e.
-a meiotic driver), may increase in frequency even if it entails a pleiotropic fitness cost \citep{Prout1973}, generating a genetic conflict between the success of the driver and organismal fitness.
+a meiotic driver), may increase in frequency even if it entails a pleiotropic fitness cost \cite{Prout1973}, generating a genetic conflict between the success of the driver and organismal fitness.
 Meiotic drivers observed in both plants
-\citep{Buckler1999,Fishman2005,Fishman2008}, and animals
-\citep{Agulnik1990,Wu2005,Pardo-ManuelDeVillena2001c} highlight this
+\cite{Buckler1999,Fishman2005,Fishman2008}, and animals
+\cite{Agulnik1990,Wu2005,Pardo-ManuelDeVillena2001c} highlight this
 conflict -- the selfish benefits and the associated
 pleiotropic fitness costs of drive sustain a balanced polymorphism
-\citep{Prout1973}, 
-and often generate ongoing evolutionary escalations of drive suppressors and enhancers \citep{Dawe1996,Fishman2008}. 
+\cite{Prout1973}, 
+and often generate ongoing evolutionary escalations of drive suppressors and enhancers \cite{Dawe1996,Fishman2008}. 
 The threat of meiotic drive to organismal fitness is potentially so
@@
 The stage at which female meiosis requires fertilization varies widely across taxa, and in a few animals  
-	female meiosis is complete upon ovulation (that is, fertilization is not required for female meiosis \citep[e.g. Table 1 in 
-][]{Masui_book}).
+	female meiosis is complete upon ovulation (that is, fertilization is not required for female meiosis (e.g. Table 1 in \cite{Masui_book}).
 
@@
  transmission of alternate alleles in heterozygous females can depend on sperm haplotype.  
- While ruling out the alternative hypothesis of early selection on zygotes in these cases is challenging  \citep[see pages
-52-54 in][]{Burt2006}, it appears that the extent to which \emph{In} and \emph{Om} distort the second female meiotic division partially depends on the genotype of the fertilizing sperm \citep{Agulnik1993,Wu2005}.
-% Both of these female drive systems operate by distorting the second meiotic division, 
-% and in both systems the outcome of female meiosis depends the genotype of the fertilizing sperm \citep{Agulnik1993,Wu2005}.  
-%While such genetic evidence is suggestive, we note the difficulty in
-%ruling out the alternative hypothesis of early selection on zygotes \citep[pages
-%52-54][]{Burt2006}.
+ While ruling out the alternative hypothesis of early selection on zygotes in these cases is challenging  (see pages 52-54 in \cite{Burt2006}), it appears that the extent to which \emph{In} and \emph{Om} distort the second female meiotic division partially depends on the genotype of the fertilizing sperm \cite{Agulnik1993,Wu2005}.
 
 We develop population genetic models exploring the evolution of sperm influence on female meiotic drive. 
-%, focussing on models in which `self-promoting' alleles in sperm facilitate drive of like genotypes during gametogenesis in heterozygous females.
 We first focus on models in which `self-promoting' alleles in
   sperm facilitate drive of like alleles during gametogenesis in heterozygous females. 
-%In this article we explore through simple population genetic models the consequences of alleles that influence the outcome of female meiosis.
 These models show that such sperm-acting alleles 
@@
 Such models universally reveal that sperm-acting drive modifiers are favored only if they suppress drive. 
-These results highlight the fact that the interests of sperm and maternal  genomes' are often aligned, as both are invested in the fate of the resultant zygote (as was speculated for the \emph{In} locus \citep{Pomiankowski1993}).
+These results highlight the fact that the interests of sperm and maternal  genomes' are often aligned, as both are invested in the fate of the resultant zygote (as was speculated for the \emph{In} locus \cite{Pomiankowski1993}).
 Therefore, there is little selective benefit to females in preventing sperm to influence female meioses.
@@
 
+% Results and Discussion can be combined.
 \section*{Results}
@@
 the driving allele is transmitted to the egg in heterozygotes with
-probability  $d > 1/2$, regardless of sperm genotype \cite[e.g.
-][ and see Model 1 in the Appendix for more details]{Ubeda2004}.
-%The timing of fertilization relative to female meiosis places another constraint on $d$, for example, if fertilization (and therefore, sperm dependent drive) takes place at MII (as in mammals),
-%	female drive requires an uneven number of crossovers between the centromere and the drive locus, 
-%	so $d$ is bounded to be $<0.83$ \citep[see ][ for discussion]{Buckler1999}.
-	%YB - MOVED THIS TO THE APPENDIX
-          To depict a case
-of a self-promoting meiotic driver,  we modify this standard model such 
+probability  $d > 1/2$, regardless of sperm genotype (e.g. \cite{Ubeda2004}, and see Model 1 in the Appendix for more details). 
+To depict a case of a self-promoting meiotic driver,  we modify this standard model such 
 	that the driver is only effective when fertilized by a sperm
@@
 
-For comparative purposes, we first briefly present the standard drive model %as Model 1 in the Appendix, 
-%equations \ref{eqn.A1}-\ref{eqn.A3}, 
-	%and refer readers to the previous literature on this model 
-	\citep[see e.g. ][ for additional results]{Prout1973,Ubeda2004}. 
+For comparative purposes, we first briefly present the standard drive model 
+	(see e.g. \cite{Prout1973,Ubeda2004} for additional results). 
 Assuming that the driving allele is deleterious in both sexes, but fully recessive 
@@
 However, when $s$ is large ($s>(2d-1)/(2)$, solid black line in Figure \ref{Eggsperm}B) a driver cannot fix and
-will be maintained as a protected polymorphism \citep{Prout1973}. 
+will be maintained as a protected polymorphism \cite{Prout1973}. 
 The parameter space where the allele can invade but not fix is shown in white
@@
 	drive suppressors, corresponding well to empirical examples of
-        female meiotic drive \citep[reviewed in ][]{Burt2006}.
+        female meiotic drive (reviewed in \cite{Burt2006}).
 
@@
 When rare, a fully recessive, self-promoting driver can only invade when $s$ 
-	%, the selection against the drive homozygote, 
 	is less than approximately $(2 d - 1)/(4 d)$ -- see dashed black line in Figure \ref{Eggsperm}B. 
@@
 
-\begin{figure}
-\begin{center}
-\includegraphics[width = 1 \textwidth]{Figures/selfprom.eps} \end{center}
-\caption{{\bf{A.}} A visual depiction of our model of `self-promoting' driver. 
-	Transmission probabilities for alleles through female
- 	 meiosis depend on sperm genotype. 
-	 The non-driving $A$-allele, and self-promoting $B$-allele
-	 are represented by unfilled and filled circles, respectively. 
-	 {\bf{B.}} Evolution of a self-promoter and standard driver. 
-	 	Assuming that the fitnesses of drive homozygotes and 
-		heterozygotes are $1-s$ and $1$, respectively. 
-	 	\emph{Main figure:} Boundary conditions for the invasion 
-		and fixation of self-promoting and standard meiotic drivers, 
-		with drive coefficient, $d$. Colored regions depict exact results, 
-		while lines  represent analytical approximations. 
-	 \emph{B1:}  Trajectories of sperm-dependent female drive
-		 %alleles that can invade the population, 
-		 each allele has $s =0.1$ against the
-		 homozygotes.  The drive coefficient is denoted by color. 
- }   
-\label{Eggsperm}
-\end{figure}
+%%%%%%%%%%%
+%Figure1 was here%
+%%%%%%%%%%%
 
@@
 Although the allelic identity of sperm could plausibly influence the outcome of female meiosis, 
-	limited gene expression in sperm \citep[e.g.][]{Joseph2004}
+	limited gene expression in sperm (e.g. \cite{Joseph2004})
 	suggests a model where sperm influence female meiosis via expression of the fertilizing male's
@@
 	(see Figure \ref{Effect_of_dominance}, and Equation \ref{polymale}).  
-%Presumably this zone of polymorphism comes about because positive frequency 
-%	dependence diminishes with the dominance of the male effect driver. 
-%As an analytical example, consider the case of male-genotype dependent deleterious recessive female driver (assuming, as above, $w_{BB}=1-s$, and $w_{AA}=w_{AB}=1$). 
-%Here, a rare driver invades when $s$ is less than $(d_{het}-1/2)/d_{het}$, and fixes when $s$ is less than $d_{hom} - 1/2$.
-%Intermediate values of $s$, which represent a small proportion of parameter space can maintain a polymorphism. 
 While mathematically interesting, it does not seem particularly realistic to think that the
@@
 	required for this polymorphism make us doubt its general applicability.
-%Thus we regard this case as interesting but not very general.
 
@@
 female meiotic drive.
-
-%\begin{figure}
-%	\rotatebox{90}{
-%\includegraphics[width = 0.5
-   %       \textwidth]{Figures/Sperm_egg_cartoon_paper_Fig1.eps}}
-%\caption{transmission probabilities for alleles through female
-%  meiosis depend on sperm genotype. 2 allele model. \gc{merge with 2,
-%    rewrite both captions}}  
-%\label{Eggsperm_2_allele_cartoon}
-%\end{figure}
-
-% (we note that the timing of fertilization relative to female meiosis places another constraint on $d$, for example, when drive takes place druing the second meitoic division
-%	it requires an uneven number of crossovers between the centromere and the drive locus \citep[see ][ for discussion]{Buckler1999}).
-
-%\begin{figure}
-%\includegraphics[width = 0.8 \textwidth]{Figures/invasion_space_recessive_driver.eps}
-%\caption{{\bf A.} Invasion analysis for a female meiotic drive allele with
-%  recessive costs (selection coefficient $s$). First consider a simple female meiotic
-%  drive system, where the allele is transmitted a fraction $d>1/2$ of
-  %meioses in female heterozygotes, inspective
-%  of the allele carried by the fertilizing sperm. 
-%Such an allele can invade the population over all of the parameter space, and will be maintained as a
-%  protected polymorphism in the white region, and can fix in the
-%  population in all of the parameter space colored green or orange. 
-%The solid back line marks the criteria for the allele to fix $s< (2d-1)/2$.
-%Now consider an allele that is fairly transmitted through
-%  female meiosis when fertilized by a sperm carrying a different
-%  allele, but at a fraction $d>1/2$ when a heterozygote egg is fertilized
-%  by a sperm carrying the allele (as depicted in Figure
-%  \ref{Eggsperm_2_allele_cartoon}). Such an allele can only invade in
-%  the region of parameter space colored orange, and it fixes over all
- % of this parameter region. The approximate criteria for this allele
-%  to fix is given by the dashed line showing $s < (2d- 1)/(4d)$. 
-% {\bf B.} In the inset we show trajectories of sperm-dependent female drive
- %alleles that can invade the population, each allele has $s =0.1$ against the
-% homozygotes and $d =0.6$, $0.7$, $0.8$ shown in yellow, blue, and red. 
-% \gc{Likely make this Figure 1B.} } \label{Invasion_space}
-%\end{figure}
 
@@
 
-\begin{figure}
-\begin{center}
-	\includegraphics[width = 0.7
-          \textwidth]{Figures/twolocus.eps} \end{center}
-\caption{Models of a sperm-acting drive modifier tightly linked to a meiotic driver. 
-{\bf (A) } 
-Sperm carrying the derived allele at the modifier locus (filled squares) alters transmission  
-	at the driving allele (filled circles) during female meiosis.
-        Alleles at these two tightly linked loci form three haplotypes (top of A). 	
-	{\bf{A1)}} In the standard model of drive there is no
-        variation at the modifier, 
-	and the driver is transmitted to the egg with probability $d_0$. 
-	{\bf{A2)}} The modifier allele increases the transmission of the drive allele ($d_1 > d_0$), and due to their shared genetic background, also increases its drive. 
-	{\bf{A3 \& A4)}} The sperm-acting modifier acts to decrease
-        drive  ($d_1=1/2$ in A3 \& A4, or more generally, $d_1 < d_0$) and arises on the same or opposite
-        background from the driver (A3 \& A4 respectively). 
-%		In {\bf{A3}} the drive suppressor exists on the drive background, forming the $B^-$ haplotype. 
-%		In {\bf{A4}} the drive suppressor exists on the non-drive background, forming the $A^-$ haplotype. 		
-{\bf{(B)}} Invasion of a sperm-acting drive suppressor linked to a driver. 
-	After the driver ($B$ haplotype) reaches drive selection equilibrium, we introduce a sperm acting drive modifier. 
-	We assume full drive ($D_0 =1$), a recessive lethal fitness cost to drive ($h_s=0$, $s=1$) and that the sperm-acting modifier results in a fair meiosis. 
-{\bf {B1)}} The $B^-$ allele replaces the ancestral drive haplotype, but segregates at a lower equilibrium frequency.
-{\bf {B2)}} The $A^-$ allele replaces the ancestral non-driving haplotype, and in this case, removes the driver from the population. }
- \label{Eggsperm_3_allele_cartoon}
-\end{figure}
+%%%%%%%%%%%
+%Figure2 was here%
+%%%%%%%%%%%
 
@@
 	(Figure \ref{Eggsperm_3_allele_cartoon}B2). 
-This result is consistent with previous work showing that drive suppressors unlinked to, or in repulsion phase with drivers usually invade polymorphic drive systems \citep[e.g. ][]{Brandvain2012}.  
+This result is consistent with previous work showing that drive suppressors unlinked to, or in repulsion phase with drivers usually invade polymorphic drive systems (e.g. \cite{Brandvain2012}).  
 Therefore, all two-locus models of sperm influence on female drive suggest that
 	sperm will evolve to oppose female meiotic drive.
-
-%More surprisingly, sperm-acting suppressors of female drive can spread
-%	even when tightly linked to the original drive system (Figures \ref{Eggsperm_3_allele_cartoon} and \ref{Trajectories_of_supressors}B). 
-%The $B^{-}$ haplotype (representing the tightly linked drive/drive suppressor complex) displaces the original driver 
-%	by driving when not costly, but avoiding low fitness homozygotes by disrupting drive when transmitted to sperm. %  ($BB^-$ and $B^-B^-$). 
-%Despite displacing the ancestral driver (the $B$ haplotype), the equilibrium frequency of the $B^-$ haplotype is often lower than that of $B$ \yb{(point to supp?)}.
-
-%\gc{Do we mention the In and Om locus here or in the discussion as a
-%	 potential e.g. Do we mention the fact that the frequency of these
-%	 newly invade sperm dodging alleles can be lower than the original system?}\\
 
@@
 of forming low fitness homozygotes. 
-%In fact, an allele in sperm which causes the alternate allele to drive can be favored, a counterintuitive result 
-%	consistent with the tenuous evidence that the non-driving
-%        $ln^-$ allele is over transmitted through female meiosis in $ln^-/ln^+$ heterozygote when fertilized by $ln^+$
-%        sperm \cite{agulnick, Pomiankowski1993}.   
-% YB - WE DID NOT TALK ABOUT THIS IN THE RESULTS [ALTHOUGH IT'S TRUE... PERHAPS TOO DISTRACTING]
 More generally, natural selection favors alleles that act through sperm to
@@
 Our results highlight potentially counterintuitive results of complex genetic conflicts. 
-Despite much opportunity for conflict between sperm and females over fertilization \citep{Partridge1998},
+Despite much opportunity for conflict between sperm and females over fertilization \cite{Partridge1998},
 	the interests of fertilizing sperm and female are quite well aligned during syngamy.
@@
 
-  \appendix
-
-\section*{Appendix}
+% You may title this section "Methods" or "Models". 
+% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
+% authors may use "Analysis" 
+\section*{Models}
 We highlight simple analytical results from: 
@@
 	female drive requires an uneven number of crossovers between the centromere and the drive locus, 
-	so $d$ is bounded to be $<0.83$ \citep[see ][ for discussion]{Buckler1999}. 
+	so $d$ is bounded to be $<0.83$ (see \cite{Buckler1999} for discussion). 
 After drive and random mating, genotype frequencies are 
-	$f_{AA}'=f_A (f_{AA }+ f_{AB} (1 - d))$, 
-	$f_{AB}'=f_A (f_{BB} + f_{AB}d) + f_B (f_{AA} + f_{AB} (1 - d))$,
-	and $f_{BB}'= f_B (f_{AB} d + f_{BB})$. 
-%Assuming fitnesses $w_{AA}=1$, $w_{AB}-1 s_h$ and $w_{BB}=1-s$, 
-%Frequencies of genotype, $i$, after drive and selection are 
-%	\begin{equation} f_i''= f_i'w_i/\bar{w}\end{equation}
-%	where $w_i$ is genotypic fitness, and mean fitness, $\bar{w} = \Sigma f_i'w_i$. 
+	\begin{equation} 	f_{AA}'=f_A (f_{AA }+ f_{AB} (1 - d)), \end{equation} 
+	\begin{equation} 	f_{AB}'=f_A (f_{BB} + f_{AB}d) + f_B (f_{AA} + f_{AB} (1 - d)), \text{and }\end{equation}
+	\begin{equation} 	f_{BB}'= f_B (f_{AB} d + f_{BB}). \end{equation}
 As detailed above, exact frequencies after drive, random mating and selection are $f_g''= f_g'w_g/\bar{w}$. 
@@
         recessive.
-        %\gc{isn't this last expression exact?} YB: exact-ish.. all assume HWE.. but i think we don't expect a deviation from HWE here??  it's prob easiest to kept them all as approx?
 
@@
 Therefore, genotype frequencies after drive are 
-	$f_{AA}'=f_A^2$, 
-	$f_{AB}'=f_A f_B+f_B (f_{AA} + f_{AB}(1 - d)) $,
-	and $f_{BB}'= f_B (f_{AB} d + f_{BB} )$. 
-%As above, exact frequencies after drive, random mating and selection are $f_i''= f_i'w_i/\bar{w}$. 
+	\begin{equation} 	f_{AA}'=f_A^2, \end{equation} 
+	\begin{equation} 	f_{AB}'=f_A f_B+f_B (f_{AA} + f_{AB}(1 - d)), \text{and } \end{equation}
+	\begin{equation} 	f_{BB}'= f_B (f_{AB} d + f_{BB} ).\end{equation}
 We iterate exact genotype frequency recursions ($f_g''= f_g'w_g/\bar{w}$) over generations to produce the frequency trajectories shown in the inset of 
 Figure \ref{Eggsperm}B by plotting $f_B=f_{BB}+ \frac{1}{2}f_{AB}$
-over time. %, and to assess the analytical approximations, below.
+over time. 
 To assess invasion or fixation criteria, as well as bistability points, we iterate this system and
@@
 
-% the $B$ allele In our first scenerio allele $A$ is the ancestral allele and allele
-%$B$ is our sperm dependent drive allele.
-%The sperm dependence of the drive is a form of assortative
-%transmission (mating), and so we have to keep track of the genotype frequencies.
-%Our genotype frequencies at birth are $x_{AA},~x_{AB},$ and
-%$x_{BB}$. These three genotypes have fitnesses $w_{AA},~w_{AB},$ and
-%$w_{BB}$, which are identical between males and females. Because 
-%fitnesses are equivalent in both sexes, genotypic frequencies
-%are identical in males and females, but in the following
-%equations they are marked with $\Mars$ and $\Venus$ respectively to
-%allow the reader to more easily follow the logic. 
-%The frequency of the three genotypes at birth in the next generation is:
-%\begin{equation}
-%x'_{BB} =\frac{1}{\bar{w} } \Big( w_{BB} x_{BB}^{\Venus} (w_{BB} x_{BB}^{\Mars} +\frac{1}{2}
-%  w_{AB} x_{AB}^{\Mars} ) + w_{AB} x_{AB}^{\Venus} (d w_{BB}
-%  x_{BB}^{\Mars} +\frac{1}{2} d
-%w_{AB} x_{AB}^{\Mars} ) \Big),  \label{eqn.A1}
-%\end{equation}
-%\begin{equation}
-%x'_{AA} =\frac{1}{\bar{w} } \Big(  w_{AA} x_{AA}^{\Venus} (w_{AA} x_{AA}^{\Mars} + \frac{1}{2}
-%   w_{AB} x_{AB}^{\Mars}) + w_{AB} x_{AB}^{\Venus}(\frac{1}{2} w_{AA} x_{AA}^{\Mars}+ \frac{1}{2} 
-%  w_{AB} x_{AB}^{\Mars}) \Big),  \label{eqn.A2}
-%\end{equation}
-%and
-%\begin{eqnarray}
-%x'_{AB} = & \frac{1}{\bar{w} } \Big( w_{AA} x_{AA}^{\Venus}  ( w_{BB}  x_{BB}^{\Mars} + \frac{1}{2}
-%  w_{AB}  x_{AB}^{\Mars} ) +w_{BB} x_{BB}^{\Venus} ( w_{AA}  x_{AA}^{\Mars} + \frac{1}{2}
-%  w_{AB}  x_{AB}^{\Mars} )  \nonumber \\
-%& + x_{AB}^{\Venus}(\frac{1}{2}  w_{AA}  x_{AA}^{\Mars}+\frac{1}{2} (1-d)
-% w_{AB}  x_{AB}^{\Mars} + \frac{1}{4}  w_{AB}
- %x_{AB}^{\Mars} + (1-d)  w_{BB}  x_{BB}^{\Mars} ) \Big) \label{eqn.A3}
-%\end{eqnarray}
-%where the mean fitness of the population is
-%\begin{equation}
-%\bar{w} = w_{AA} x_{AA}+w_{AB}  x_{AB} +w_{BB}  x_{BB}.
-%\end{equation}
-%We can then iterate this recursion over generations, for
-%example to produce the frequency trajectories shown in the inset
-%Figure \ref{Invasion_space} by plotting $x_{BB}+ \frac{1}{2}x_{AB}$
-%over time. We can also use this to assess the
-%invasion criteria for the allele by starting allele B at very low
-%frequency, and seeing if it increases across the generations.
-
-\paragraph{Recessive fitness cost of self-promoting driver:} %When the self-promoting driver has a recessive fitness cost, $s$ genotype frequencies after selection and drive equal
-%\begin{equation}	x'_{BB} =\frac{x_{B}^2(1-s)}{\bar{w} }  \Big( F+(1-F)(x_{B}+2x_{A} d) \Big) 	\end{equation}
-%\begin{equation}	x'_{AA} =\frac{1}{\bar{w} }  \Big(x_{A}^2\Big) 	\end{equation}
-%and
-%\begin{equation}	x'_{AB} =\frac{2 x_{A}x_{B} }{\bar{w} }  \Big(1-x_{B}(d-1/2)(1-F) \Big)	%\end{equation}
+\paragraph{Recessive fitness cost of self-promoting driver:}
 When fully recessive, the change in frequency of the self-promoting driver across generations equals 
@@
 \end{equation}
-where $F$ is the deviation from genotypic frequencies expected under Hardy-Weinberg. % Trying to clarify some.   \gc{Hardy Weinberg Expectations} (HWE). 
-%For analytical solutions presented above, we assume that $F$ equals zero. 
+where $F$ is the deviation from genotypic frequencies expected under Hardy-Weinberg.
 Assuming HWE  ($F=0$) a common, recessive, self-promoting driver invades if $s\lessapprox(2 d - 1)/(4 d)$, and fixes if
@@
  but it will be lost when introduced below this frequency. 
- %(i.e. this is an unstable equilibrium point). 
-% In Supplementary Figure \ref{Bistab_homozyg_cost_fig} we show
-% this approximation condition for bistability andd how it compared to
-% that obtained using the recursion. 
 Compared to exact results (Supplementary Figure \ref{Bistab_homozyg_cost_fig}), Equations
  \eqref{ineqA} and \eqref{bistab_homozy_approx} offer reasonable, but imperfect approximations.
-%s_h and seems to serve as a lower bound to the true unstable equilibrium frequency.
 
-\paragraph{Cost of driver in heterozygotes:} %We explored a more complex model where the self-promoting driver has a cost, $s_h$ in heterozygotes (so $w_{AB} = 1-s_h$) in addition to the cost $s$, in homozygotes.
+\paragraph{Cost of driver in heterozygotes:} 
 When the fitness of drive heterozygotes is compromised ($s_h>0$), a self-promoting driver cannot invade when rare.
@@
 cost in heterozygous individuals ($\sim f_B$). 
-%Model construction is similar to that above.
-%So long as there is any heterozygous fitness cost (i.e. $s_h > 0$) the self-promoting driver cannot invade when rare;
 However, this system too, is bistable -- as the driver increases in frequency it is more often fertilized by a driving sperm and therefore drives more effectively. 
@@
 parameter values; however, it underestimates $f_B^*$ for large parameter values, presumably because they result in strong departures from HWE.
- %range of other costs to the drive alleles other than a
-%simple recessive cost, for example an additive cost. We modified the recursions given in eqn
-%\ref{eqn.A1}-\ref{eqn.A3} to feature an additive cost $s$ for each
-%copy of the $B$ allele. A none recessive cost to the drive allele can
-%cause bistability of the drive allele. The B allele is unable to enter
-%the population when rare, this results from the fact that
-%when the allele is rare ($x_B \ll 1$) sperm carrying it rarely
-%encounter heterozygous eggs ($\sim x^2$) but the allele still pays a
-%cost in heterozygous individuals.  By contrast, when the B allele is common ($x_B
-%\gg 0$) encounters between heterozygous eggs and $B$ allele sperm
-%become more common allowing drive to occur this can allow the $B$
-%allele to spread when common. In Supplementary Figure
-%\ref{bistable_additive} we show the unstable equilibrium frequency
-%above which the $B$ allele can invade. This equilibrium frequency was found by a simple
-%parameter search.  \gc{wonder if a bit of work could find this}
 
@@
 In the case when female meiotic drive depends on paternal genotype,
-%\paragraph{Dominance} we also explored a model where
-%the genotype of the male who produced the fertilizing sperm affects female drive,
-%instead of just the allelic type of the sperm.
- %We assume a 
  a heterozygous female will transmit the $B$ allele 
   with probabilities  $\frac{1}{2}$,  $d_{het}\geq 1/2 $ or $d_{hom}\geq d_{het}$, 
-%YB RE-PARAMETERIZING  %  $\frac{1}{2}$,  $\frac{1}{2} + h(d-\frac{1}{2}) $ or $d$,  YB RE-PARAMETERIZING 
  when mated with to $AA$, $AB$, or $BB$ males,  respectively. 
   In this model, genotype frequencies after drive, random mating, and selection are 
-  	$f_{AA}'' =   w_{AA}\left( f_{AA} [2 f_A + f_{AB} ] + f_{AB}^2 [1 - d_{het}] \right)/(2\bar{w})$,
-	$f_{AB}'' =   w_{AB}(f_{AA} [f_B + f_{AB}/2] + f_{AB} f_B + f_{BB} [ f_A - f_{AB} d_{hom}])/\bar{w}$, 
-	and 	$f_{BB}'' =   w_{BB}(f_{AB} (f_{AB} d_{het}/2 + f_{BB} d_{hom}) + f_{BB} f_B)/\bar{w}$. 
-% Above, $d > 1/2$ and $h$ ($0 \geq h \geq 1$) acts to specify the dominance of the $B$
-%allele's effect on female meiotic drive through the males genotype.
-%YB TRY_1 We assume that in a mating between a heterozygous female and an $AA$, $AB$, or $BB$ male, the female will transmit the $B$ allele 
-% with probabilities  $\frac{1}{2}$,  $\frac{1}{2} + h(d-\frac{1}{2}) $, and $d$, respectively.
-  %Here $d > 1/2$ and $h$ ($0 \geq h \geq 1$) acts to specify the dominance of the $B$
-%allele's effect on female meiotic drive through the males genotype.
-% GC ORIGINAL  We assume that if a
-%heterozygous egg ($AB$) is fertilized by a sperm with the following
-%male ($\Mars$) genotypes the $B$ allele will be transmitted through
-%female meiosis with the following probabilities \\
-%\begin{tabular}{cccc}
-%$\Mars$ geno. & AA & AB  & BB\\
-%P(B). & $\frac{1}{2}$ &  $\frac{1}{2} + h(d-\frac{1}{2}) $ & $d$ 
-%\end{tabular}\\
-%where $d > 1/2$ and $h$ ($0 \geq h \geq 1$) acts to specify the dominance of the $B$
-%allele's effect on female meiotic drive through the males genotype. 
-%We set up a set of recursions using a recessive sex independent cost to
-%the driver, similar to those in eqn.
-%\ref{eqn.A1}-\ref{eqn.A1}, but now featuring these male genotype
-%transmission probabilities.
+  	\begin{equation} f_{AA}'' =   w_{AA}\left( f_{AA} [2 f_A + f_{AB} ] + f_{AB}^2 [1 - d_{het}] \right)/(2\bar{w}) , \end{equation}
+	\begin{equation} f_{AB}'' =   w_{AB}(f_{AA} [f_B + f_{AB}/2] + f_{AB} f_B + f_{BB} [ f_A - f_{AB} d_{hom}])/\bar{w}  , \text{and } \end{equation} 
+	\begin{equation} f_{BB}'' =   w_{BB}(f_{AB} (f_{AB} d_{het}/2 + f_{BB} d_{hom}) + f_{BB} f_B)/\bar{w} . \end{equation}
 
@@
 By contrast, when $d_{het} \lessapprox 1/(3-2d_{hom})$ the case is reversed, and the model is bistable.
-%In Supplementary Figure \ref{Effect_of_dominance} we show the invasion
-%criteria for both the A and B alleles when they are rare. 
-%We find a narrow parameter range where B 
-%can invade but not fix.{eqmale}
-%This corresponds to B being less than fully recessive (i.e. $h>0$) and to a modest cost in homozygotes. 
-%This combination of parameters reflects the fact that when the allele is rare, 
-%it can increase in frequency due to fertilization of heterozygous eggs
-%by sperm from heterozygote males, where the sperm do not themselves
-%carry the allele. But as the B allele becomes common it will frequently be
-%present in the homozygous state in males, and so will be unable to
-%avoid the production of unfit homozygote zygotes leaving it trapped in
-%the population.
 
@@
 alleles at the drive locus (such that recombination is unexpected). 
-In all cases, we assume that fitness is independent of the drive modifier. %, so, $w_{AA}=1$, $w_{AB}=1-s_h$, and $w_{BB} = 1-s$. 
+In all cases, we assume that fitness is independent of the drive modifier. 
 We assume the traditional driver is transmitted to $d_0$ of gametes from female heterozygotes  when fertilized by wild-type sperm, and $d_1=d_0+\epsilon$ when fertilized by a sperm-acting drive-modifier.
@@
 More generally, even when the cost of drive is not fully recessive, the $B^-$ allele will invade and fix under all parameters sustaining a previous polymorphism at the drive locus (see Mathematical Supplement).
-%When the C allele acts to suppress drive for all parameters sustaining a polymorphism at the drive locus ($s>d_0-1/2$)
-%We address two questions concerning the $C$ allele - can it invade when rare  and can it fix when common?
-%Although our analytical approximations (see our \gc{Mathematical supplement}) do not reveal simple inequalities like those above,
-%	graphical exploration of analytical results show that this mutant is lost when it enhances drive ($\epsilon>0$ - the $B^+$ allele, above),
-%	and spreads from rarity to fixation when it suppresses drive ($\epsilon<0$ - the $B^-$ allele, above).	
-%This result is supported by exact numerical iterations, as well
-%\yb{FIG?}. \gc{Lets not bother}
 
@@
 \begin{equation} \Delta M =-\epsilon f_A f_B (f_B s + s_h (f_A - f_B) ) \frac{f_M}{\bar{w}}  \label{unlinked} \end{equation}
-% an $F$ allele increases in frequency when 
-%	$-\epsilon f_A f_B (f_B s + s_h (fA - fB) ) >0$, corresponding to any level of drive suppression ($\epsilon<0$) 
 Thus, a rare drive suppressor ($\epsilon<0$) will spread so long as the fitness cost of the driver does not display over- or under-dominance. 
-Notably, this result holds while assuming complete linkage equilibrium, and therefore sperm suppression of female meiotic drive is likely more powerful than more direct female-enforced drive suppressors  which usually spread due to the indirect benefit of avoiding drive haplotypes \citep[e.g.
-][]{Brandvain2012}.
-%The spread of this driver does not, however, depend on assumption of HWE, and linkage equilibrium - exact results resemble analytical approximations. 
-%\paragraph{three allele system}
-%$x_{AA}$, $x_{AB}$,$x_{AB'}$, $w_{BB'} x_{BB'}$, $w_{B'B'} x_{BB}$, $w_{B'B'} x_{B'B'}$
-%We denote the frequency of the three genotypes in the population at birth are
-%$x_{AA}$, $x_{AB}$, $x_{AB'}$, $x_{BB'}$, $x_{BB}$, $x_{B'B'}$. 
-%The mean relative fitness of the population is 
-%\begin{equation}
-%x_{AA}+x_{AB}+ x_{AB'}+ w_{BB'} +x_{BB'}+w_{B'B'}x_{BB} + w_{B'B'} x_{B'B'} 
-%\end{equation}
-%The frequency of the genotypes in meiotic-cells of female adults is 
-%\begin{equation}
-%\frac{1}{\bar{w}}\Big( x_{AA}, ~x_{AB},~ x_{AB'},~ w_{BB'} x_{BB'},~ w_{B'B'}x_{BB},
-%~w_{B'B'} x_{B'B'} \Big)
-%\end{equation}
-%The frequency of fertilization events can be found by combining these
-%female frequencies with the frequency of the three alleles in sperm is
-%\begin{eqnarray}
-%x_A = & x_{AA} + \frac{1}{2} x_{AB}+ \frac{1}{2} x_{AB'} \\
-%x_B = & w_{BB}  x_{BB} +   w_{BB'} \frac{1}{2} x_{BB'}+ \frac{1}{2} x_{AB} \\
-%x_{B'} = & w_{B'B'}  x_{B'B'} +   w_{BB'} \frac{1}{2} + x_{BB'}+ \frac{1}{2} x_{AB'}  
-%\end{eqnarray}
-%in combination with the transmission probabilities given in Figure
-%\ref{Eggsperm_3_allele_cartoon}A.
+Notably, this result holds while assuming complete linkage equilibrium, and therefore sperm suppression of female meiotic drive is likely more powerful than more direct female-enforced drive suppressors  which usually spread due to the indirect benefit of avoiding drive haplotypes (e.g.
+\cite{Brandvain2012}).
 
-% \gc{Alternatively we could give the table below:}
+% Do NOT remove this, even if you are not including acknowledgments
+\section*{Acknowledgments}
 
-% \begin{table}
-% \begin{tabular}{lllll}
-% \hline 
-%   &  \multicolumn{4}{c}{Sperm genotype}   \\
-%   &  & A & B & B'  \\
-% $\Venus$ geno. & Freq.  & $x_{AA} + \frac{1}{2} x_{AB}+ \frac{1}{2} x_{AB'} $ & $ w_{BB}  x_{BB} +   w_{BB'} \frac{1}{2} x_{BB'}+
-% \frac{1}{2} x_{AB}$ & $ w_{B'B'}  x_{B'B'} +   w_{BB'} \frac{1}{2}
-% x_{BB'}+ \frac{1}{2} x_{AB'}  $\\
-% \hline\\
-% AA & $x_{AA}$ &  & & \\
-% AB & $x_{AB}$ &  & & \\
-% AB' & $x_{AB'}$ &  & & \\
-% BB' & $w_{BB'} x_{BB'}$ &  & & \\
-% BB & $w_{B'B'} x_{BB}$ &  & & \\
-% B'B' & $w_{B'B'} x_{B'B'}$ &  & & \\
-% \end{tabular}
-% \caption{ALTERNATIVELY I COULD FILL IN THIS MATRIX  with genotype }
-% \end{table}
+%\section*{References}
+% The bibtex filename
 \bibliography{refs}
 
-\newpage
+\section*{Figure Legends}
+%\begin{figure}[!ht]
+%\begin{center}
+%%\includegraphics[width=4in]{figure_name.2.eps}
+%\end{center}
+%\caption{
+%{\bf Bold the first sentence.}  Rest of figure 2  caption.  Caption 
+%should be left justified, as specified by the options to the caption 
+%package.
+%}
+%\label{Figure_label}
+%\end{figure}
+
+\begin{figure}
+\begin{center} \includegraphics[width = 1 \textwidth]{Figures/selfprom.eps} \end{center}
+\caption{ {\bf{A model of single locus, male-dependent meiotic drive. }} {\bf{A.}} A visual depiction of our model of `self-promoting' driver. 
+	Transmission probabilities for alleles through female
+ 	 meiosis depend on sperm genotype. 
+	 The non-driving $A$-allele, and self-promoting $B$-allele
+	 are represented by unfilled and filled circles, respectively. 
+	 {\bf{B.}} Evolution of a self-promoter and standard driver. 
+	 	Assuming that the fitnesses of drive homozygotes and 
+		heterozygotes are $1-s$ and $1$, respectively. 
+	 	\emph{Main figure:} Boundary conditions for the invasion 
+		and fixation of self-promoting and standard meiotic drivers, 
+		with drive coefficient, $d$. Colored regions depict exact results, 
+		while lines  represent analytical approximations. 
+	 \emph{B1:}  Trajectories of sperm-dependent female drive
+		 %alleles that can invade the population, 
+		 each allele has $s =0.1$ against the
+		 homozygotes.  The drive coefficient is denoted by color. 
+ }   
+\label{Eggsperm}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+	\includegraphics[width = 0.7
+          \textwidth]{Figures/twolocus.eps} \end{center}
+\caption{ {\bf{Models of a sperm-acting drive modifier tightly linked to a meiotic driver. }}
+{\bf A.} 
+Sperm carrying the derived allele at the modifier locus (filled squares) alters transmission  
+	at the driving allele (filled circles) during female meiosis.
+        Alleles at these two tightly linked loci form three haplotypes (top of A). 	
+	\emph{A1:} In the standard model of drive there is no
+        variation at the modifier, 
+	and the driver is transmitted to the egg with probability $d_0$. 
+	\emph{A2:} The modifier allele increases the transmission of the drive allele ($d_1 > d_0$), and due to their shared genetic background, also increases its drive. 
+	\emph{A3 and A4:} The sperm-acting modifier acts to decrease
+        drive  ($d_1=1/2$ in A3 \& A4, or more generally, $d_1 < d_0$) and arises on the same or opposite
+        background from the driver (A3 \& A4 respectively). 
+%		In {\bf{A3}} the drive suppressor exists on the drive background, forming the $B^-$ haplotype. 
+%		In {\bf{A4}} the drive suppressor exists on the non-drive background, forming the $A^-$ haplotype. 		
+{\bf{ B.}} Invasion of a sperm-acting drive suppressor linked to a driver. 
+	After the driver ($B$ haplotype) reaches drive selection equilibrium, we introduce a sperm acting drive modifier. 
+	We assume full drive ($D_0 =1$), a recessive lethal fitness cost to drive ($h_s=0$, $s=1$) and that the sperm-acting modifier results in a fair meiosis. 
+\emph{B1:} The $B^-$ allele replaces the ancestral drive haplotype, but segregates at a lower equilibrium frequency.
+\emph{B2:} The $A^-$ allele replaces the ancestral non-driving haplotype, and in this case, removes the driver from the population. }
+ \label{Eggsperm_3_allele_cartoon}
+\end{figure}
 
@@
 
-\section*{Supplementary Material}
-
 \begin{figure}
 
-\includegraphics[width = \textwidth]{../Scripts/homozyg_bistability.eps}
-\caption{Invasion analysis for a self-promoting female meiotic drive allele with
+\includegraphics[width = \textwidth]{../../Scripts/homozyg_bistability.eps}
+\caption{  {\bf{Invasion analysis for a self-promoting female meiotic drive allele with
   recessive costs (selection coefficient $s$), showing the region of
-  bistability. The colors, and the thin dashed contours, indicate the
+  bistability.}} The colors, and the thin dashed contours, indicate the
   frequency the allele must be above to invade the population (note
@@
 \includegraphics[width = 0.8 \textwidth]{Figures/bistable_x_vs_d_additive_s.eps} 
-\caption{The unstable equilibrium frequency for a self-promoting
+\caption{{\bf{The unstable equilibrium frequency for a self-promoting
   female meiotic drive allele with an additive cost ($s_h=s/2$) as a
-  function of the drive parameter. The solid line shows results
+  function of the drive parameter.}}  The solid line shows results
   obtained using the recursion, the dots our approximation given by Equation \eqref{thresh2}}  \label{bistable_additive}
@@
 \begin{figure}
-\includegraphics[width = 0.7 \textwidth]{../Scripts/effect_of_dominance_on_invasion_space_four_graph.eps} 
-\caption{Exact results of invasion analysis of an allele whose effect
+\includegraphics[width = 0.7 \textwidth]{../../Scripts/effect_of_dominance_on_invasion_space_four_graph.eps} 
+\caption{{\bf{Comparing exact results to analytical approximations for the model of paternal-dependent female drive.}} Exact results of invasion analysis of an allele whose effect
  on female meiosis is mediated by the genotype of the fertilizing
@@
 \end{document}
+
+ \bibliography{refs}
+
+\newpage
+
+  \setcounter{figure}{0}\global\def\thefigure{S\arabic{figure}}
+
+\section*{Supplementary Material}

### 2014-05-31 writeup_and_notes/PLoSBioSubmission/PLoSBio_male_dependent_drive.tex
made things more PLoS-sy

@@
 the driving allele is transmitted to the egg in heterozygotes with
-probability  $d > 1/2$, regardless of sperm genotype (e.g. \cite{Ubeda2004}, and see Model 1 in the Appendix for more details). 
+probability  $d > 1/2$, regardless of sperm genotype (e.g. \cite{Ubeda2004}, and see Model 1 in the \emph{Models} for more details). 
 To depict a case of a self-promoting meiotic driver,  we modify this standard model such 
@@
         carrying that allele (see Figure \ref{Eggsperm}A and Model 2
-        in the Appendix and Supplementary Worksheet). 
+        in \emph{Models} and Appendix S1). 
 We then identify the conditions allowing for the spread of this self-promoting driver, 
@@
 	diploid genotype (perhaps due to proteins and RNAs packaged into the sperm), rather than sperm haplotype.
-This paternal-genotype dependent model (Model 3 in the Appendix) requires one additional parameter, as we exchange $d$ in the sperm dependent case for $d_{het}$ and $d_{hom}$ which describe the transmission of the drive allele in a heterozygous female mating with males heterozygous and homozygous for the self-promoting drive allele, respectively.
+This paternal-genotype dependent model (Model 3 in \emph{Models}) requires one additional parameter, as we exchange $d$ in the sperm dependent case for $d_{het}$ and $d_{hom}$ which describe the transmission of the drive allele in a heterozygous female mating with males heterozygous and homozygous for the self-promoting drive allele, respectively.
 Here, a rare driver invades when $s$ is less than $(d_{het}-1/2)/d_{het}$, and usually fixes when it invades.
@@
         heterozygotes (see Figure \ref{Eggsperm_3_allele_cartoon}A2, and
-        Model 4 in the Appendix and Supplementary Worksheet). 
+        Model 4 in \emph{Models} and Appendix S1). 
 Naively, $B^{+}$ may spread by capitalizing on the additional drive it causes; however,  
@@
 	three allele system that allows the sperm-acting drive facilitator $B^{+}$ to
-	invade the population (Appendix Model 4, eqn. \eqref{coupling}, and Mathematical Supplement).
+	invade the population (Model 4 in \emph{Models}, eqn. \eqref{coupling}, and Appendix S2).
 
@@
         third allele $B^{-}$ (Figure
-          \ref{Eggsperm_3_allele_cartoon}A3, Model 4 in Appendix). This new allele still
+          \ref{Eggsperm_3_allele_cartoon}A3, Model 4 in \emph{Models}). This new allele still
         experiences female drive when fertilized by A or B sperm, but 
@@
 Transmission rules describing the outcomes of all matings in each
-	model are presented in a Supplementary Worksheet. 
+	model are presented in a Appendix S1. 
 The fitness of genotype, $g$ is sex-independent and equals $1$, $1-h_s$, and $1-s$ for genotypes $AA$, $AB$, and $BB$, respectively. %  and  $w_{AA} = 1$, $w_{AB} = 1-s_h$, and $w_{BB}=1-s$. 
@@
 Because many of our analyses assume genotype frequencies expected under Hardy Weinberg Equilibrium (HWE), we note which analytical results are approximations. 
-We verify these approximations with exact numerical iterations in Supplementary Figures and Supplementary Scripts, and detail the derivation of these approximations in our Mathematical Supplement.
+We verify these approximations with exact numerical iterations in Figures S1-S3 and Appendix S3, and detail the derivation of these approximations in our Appendix S2.
 
@@
 	and an increase in frequency of the $C$ allele when it suppresses drive ($\epsilon <0$ -- the $B^-$ model, above). 
-More generally, even when the cost of drive is not fully recessive, the $B^-$ allele will invade and fix under all parameters sustaining a previous polymorphism at the drive locus (see Mathematical Supplement).
+More generally, even when the cost of drive is not fully recessive, the $B^-$ allele will invade and fix under all parameters sustaining a previous polymorphism at the drive locus (see Appendix S2).
 
@@
 For all values of interest ($0<s<1$, $0.5<d_0<1$), the change in frequency a rare $C$ allele is positive when it decreases drive (i.e.
-$\epsilon <0$, corresponding to the $A^-$ model, above), a result which holds qualitatively for a common $C$ allele, as well (Mathematical Supplement).
+$\epsilon <0$, corresponding to the $A^-$ model, above), a result which holds qualitatively for a common $C$ allele, as well (Appendix S2).
 Somewhat counterintuitively, the spread of the $A^-$ allele is substantially slower than that of the $B^-$ allele (it is on the order of $C^{2}$ rather than $C$). 
@@
 
-\section*{Figure Legends}
+\newpage
+\section*{Figures and Figure Legends}
 %\begin{figure}[!ht]
@@
 %\end{figure}
-
+\clearpage 
+\newpage
 \begin{figure}
@@
 
+\newpage
+\clearpage 
+\section*{Supplementary Information}
+
+\subsection*{Legends for Supplementary Appendices}
+
+\paragraph{Appendix S1: Transmission rules.} We detail the frequency of offspring genotypes produced by each possible mating for Models 1-6 in the tabs of this Excel spreadsheet.
+
+\paragraph{Appendix S2: Analytical results.} We attach a PDF of our Mathematica spreadsheet in which we derive analytical results for models 1-6.
+
+\paragraph{Appendix S3: Exact approach.} We attach an R Script used for exact recursions.
+
+\subsection*{Supplementary Figures and Supplementary  Figure Legends}
+
+\clearpage
   \setcounter{figure}{0}\global\def\thefigure{S\arabic{figure}}
@@
   fixation line (small dashes) as predicted by our HWE approximation
-  ($(d_{het}-1/2)/d_{het}$) see the appendix for more details.
+  ($(d_{het}-1/2)/d_{het}$) see \emph{Models} for more details.
 }  \label{Effect_of_dominance}

### 2014-05-31 writeup_and_notes/PLoSBioSubmission/CoverLetter.tex
pointed out that figs where in ms
replaced question mark with dash

@@
 We employ population genetic theory to show that sperm that enforce fairness in female meiosis are nearly universally favored.
-As such, we offer a novel interpretation of the requirement of fertilization for the completion of female meiosis ?
-this requirement potentially represent a collaboration between mother and sperm in the control of fitness reducing meiotic drivers.
+As such, we offer a novel interpretation of the requirement of fertilization for the completion of female meiosis -- this requirement potentially represent a collaboration between mother and sperm in the control of fitness reducing meiotic drivers.
 The application of theory to a basic and unexplained feature of organismal biology, and the interplay between conflict and co-option by numerous parties involved in fertilization make this paper of broad interest to a wide-range of biologists.
@@
 We note that while we submitted this as a `Research article,' we are open to reformatting it as an `Unsolved mystery' if the editorial board feels that is more appropriate.
+We also note that we have included Figures 1 and 2, as wells as Figures S1, S2 and S3 directly in our uploaded PDF.
 \\

### 2014-07-11 writeup_and_notes/GeneticsResub/male_dependent_drive.tex
begin revisions

--- 2014-05-31 writeup_and_notes/male_dependent_drive.tex
@@
 \documentclass[12pt,letterpaper]{article}
+\bibliographystyle{genetics}
 \usepackage{epsfig}                 
@@
 \usepackage{fullpage}
-\bibliographystyle{genetics}
 \usepackage{multirow}
@@
 
-%Why do animals allow male control of female meiosis?
-%Why does female meiosis allow the opportunity for exploitation by selfish sperm?
-%sperm and egg genomes both favour the peaceful resolution of female meiotic drive
-%strange bed fellows
+\title{  Sperm should evolve to make female meiosis fair.  }
 
-%\title{Why does female meiosis allow the opportunity for exploitation by selfish sperm? BETTER TITLE?}
-\title{
-%Co-operation between the sexes trumps selfish control of female  meiotic drive. 
-%Or 
-Sperm should evolve to make female meiosis fair.
-%Sperm do not evolve to collaborate in female meiotic drive. 
-%Or Why does female meiosis allow the opportunity for exploitation by selfish sperm? 
-}
-
-\author{Yaniv Brandvain$^1$ \\ email: ybrandvain@gmail.com  \and Graham Coop$^2$ \\ email: gmcoop@ucdavis.edu }
+\author{Yaniv Brandvain$^1$ \and Graham Coop$^2$ }
 
@@
 \date{}
-\bibliographystyle{plain}
 \begin{document}
@@
 
-%FIgures 
+\newpage 
+\flushleft{{\bf{Running title:}} Sperm prevent female meiotic drive} \\
+{\bf{Keywords:}} Meiosis; Meiotic drive; Fertilization; Genomic conflict \\
+{\bf{Corresponding author:}} Yaniv Brandvain. \emph{Address:} Department of Plant Biology, University of Minnesota --€“ Twin
+Cities. St. Paul MN, 55108. \emph{Phone:} 612-624-4375 \emph{Email:} ybrandvain@gmail.com
+
+\newpage
+
 {\bf Abstract:}
 Genomic conflicts arise when an allele gains an evolutionary advantage at a cost to organismal fitness. 
-%Because only one of four products of female meiosis is transmitted to offspring, 
 O\"{o}genesis is inherently susceptible to such conflicts
 because alleles compete to be the  product of female meiosis transmitted to the egg. 
-%(meiotic drive).  
 Alleles that distort meiosis in their favor (i.e. meiotic drivers) often decrease organismal fitness,
-	and therefore indirectly favor the evolution of mechanisms to suppress meiotic drive. % drive the rapid evolution of suppression mechanisms, and
+	and therefore indirectly favor the evolution of mechanisms to suppress meiotic drive. 
 In this light, many facets of  o\"{o}genesis and gametogenesis have been 
@@
         completion of female meiosis
-        %, and variation in stage of meiosis that
-        %fertilization occurs at 
         potentially represents a mechanism employed by females 
-	 to ensure a fair meiosis. %prevent meiotic drive.
+	 to ensure a fair meiosis.
 \newpage
@@
 At first sight, female meiosis appears primed for exploitation by such selfish systems.  
-Why then is the requirement of fertilization to complete female meiosis so ubiquitous?
+%Why then is the requirement of fertilization to complete female meiosis so ubiquitous?
 
-It is certainly not the case that animals are mechanistically incapable of circumventing this requirement.
-The stage at which female meiosis requires fertilization varies widely across taxa, and in a few animals  
-	female meiosis is complete upon ovulation (that is, fertilization is not required for female meiosis \citep[e.g. Table 1 in 
-][]{Masui_book}).
+%It is certainly not the case that animals are mechanistically incapable of circumventing this requirement.
+%The stage at which female meiosis requires fertilization varies widely across taxa, and in a few animals  
+%	female meiosis is complete upon ovulation (that is, fertilization is not required for female meiosis \citep[e.g. Table 1 in 
+%][]{Masui_book}).
 
-It is also not the case that sperm is mechanistically incapable of influencing female meiosis.
-In \emph{C. elegans}, premature deployment of the aster -- a vital component of mitotic
-machinery provided by the sperm, disrupts MII meiotic segregation
-in the egg, leading to a triploid zygote \citep{McNally2012}. 
-Thus, there is direct mechanistic evidence that sperm can influence female meiosis.
- Additionally, genetic evidence from the \emph{In} and \emph{Om}  drive systems in mice suggests that the 
- transmission of alternate alleles in heterozygous females can depend on sperm haplotype.  
- While ruling out the alternative hypothesis of early selection on zygotes in these cases is challenging  \citep[see pages
-52-54 in][]{Burt2006}, it appears that the extent to which \emph{In} and \emph{Om} distort the second female meiotic division partially depends on the genotype of the fertilizing sperm \citep{Agulnik1993,Wu2005}.
-% Both of these female drive systems operate by distorting the second meiotic division, 
-% and in both systems the outcome of female meiosis depends the genotype of the fertilizing sperm \citep{Agulnik1993,Wu2005}.  
-%While such genetic evidence is suggestive, we note the difficulty in
-%ruling out the alternative hypothesis of early selection on zygotes \citep[pages
-%52-54][]{Burt2006}.
+%It is also not the case that sperm is mechanistically incapable of influencing female meiosis.
+%In \emph{C. elegans}, premature deployment of the aster -- a vital component of mitotic
+%machinery provided by the sperm, disrupts MII meiotic segregation
+%in the egg, leading to a triploid zygote \citep{McNally2012}. 
+%Thus, there is direct mechanistic evidence that sperm can influence female meiosis.
+% Additionally, genetic evidence from the \emph{In} and \emph{Om}  drive systems in mice suggests that the 
+% transmission of alternate alleles in heterozygous females can depend on sperm haplotype.  
+% While ruling out the alternative hypothesis of early selection on zygotes in these cases is challenging  \citep[see pages
+%52-54 in][]{Burt2006}, it appears that the extent to which \emph{In} and \emph{Om} distort the second female meiotic division partially depends on the genotype of the fertilizing sperm \citep{Agulnik1993,Wu2005}.
 
 We develop population genetic models exploring the evolution of sperm influence on female meiotic drive. 
-%, focussing on models in which `self-promoting' alleles in sperm facilitate drive of like genotypes during gametogenesis in heterozygous females.
 We first focus on models in which `self-promoting' alleles in
   sperm facilitate drive of like alleles during gametogenesis in heterozygous females. 
-%In this article we explore through simple population genetic models the consequences of alleles that influence the outcome of female meiosis.
 These models show that such sperm-acting alleles 
@@
 These results highlight the fact that the interests of sperm and maternal  genomes' are often aligned, as both are invested in the fate of the resultant zygote (as was speculated for the \emph{In} locus \citep{Pomiankowski1993}).
-Therefore, there is little selective benefit to females in preventing sperm to influence female meioses.
-Rather, features of o\"{o}genesis  may evolve to facilitate an interaction with sperm %with female meiosis
-	as yet another mechanism to enforce an equitable transfer of genetic material.
+Thus, there is little selective benefit to females in preventing sperm to influence female meioses. 
+\yb{While, we do not claim that female meiotic arrest evolved specifically to recruit helpful sperm to defend against meiotic drive, the delayed completion of female meiotic drive will not in fact encourage sperm to assist renegade drivers.}
+%Rather, features of o\"{o}genesis  may evolve to facilitate an interaction with sperm %with female meiosis
+%	as yet another mechanism to enforce an equitable transfer of genetic material.
 
@@
 the driving allele is transmitted to the egg in heterozygotes with
-probability  $d > 1/2$, regardless of sperm genotype \cite[e.g.
-][ and see Model 1 in the Appendix for more details]{Ubeda2004}.
-%The timing of fertilization relative to female meiosis places another constraint on $d$, for example, if fertilization (and therefore, sperm dependent drive) takes place at MII (as in mammals),
-%	female drive requires an uneven number of crossovers between the centromere and the drive locus, 
-%	so $d$ is bounded to be $<0.83$ \citep[see ][ for discussion]{Buckler1999}.
-	%YB - MOVED THIS TO THE APPENDIX
-          To depict a case
-of a self-promoting meiotic driver,  we modify this standard model such 
+probability  $d > 1/2$, regardless of sperm genotype \cite[e.g. ][ and see Model 1 in the Appendix for more details]{Ubeda2004}. 
+ To depict a case of a self-promoting meiotic driver,  we modify this standard model such 
 	that the driver is only effective when fertilized by a sperm
         carrying that allele (see Figure \ref{Eggsperm}A and Model 2
-        in the Appendix and Supplementary Worksheet). 
+        in the Appendix and File S1 ). 
 We then identify the conditions allowing for the spread of this self-promoting driver, 
@@
 
-For comparative purposes, we first briefly present the standard drive model %as Model 1 in the Appendix, 
-%equations \ref{eqn.A1}-\ref{eqn.A3}, 
-	%and refer readers to the previous literature on this model 
+For comparative purposes, we first briefly present the standard drive model 
 	\citep[see e.g. ][ for additional results]{Prout1973,Ubeda2004}. 
@@
 \begin{center}
-\includegraphics[width = 1 \textwidth]{Figures/selfprom.eps} \end{center}
+\includegraphics[width = 1 \textwidth]{../Figures/selfprom.eps} \end{center}
 \caption{{\bf{A.}} A visual depiction of our model of `self-promoting' driver. 
@@
 
-%\begin{figure}
-%	\rotatebox{90}{
-%\includegraphics[width = 0.5
-   %       \textwidth]{Figures/Sperm_egg_cartoon_paper_Fig1.eps}}
-%\caption{transmission probabilities for alleles through female
-%  meiosis depend on sperm genotype. 2 allele model. \gc{merge with 2,
-%    rewrite both captions}}  
-%\label{Eggsperm_2_allele_cartoon}
-%\end{figure}
-
-% (we note that the timing of fertilization relative to female meiosis places another constraint on $d$, for example, when drive takes place druing the second meitoic division
-%	it requires an uneven number of crossovers between the centromere and the drive locus \citep[see ][ for discussion]{Buckler1999}).
-
-%\begin{figure}
-%\includegraphics[width = 0.8 \textwidth]{Figures/invasion_space_recessive_driver.eps}
-%\caption{{\bf A.} Invasion analysis for a female meiotic drive allele with
-%  recessive costs (selection coefficient $s$). First consider a simple female meiotic
-%  drive system, where the allele is transmitted a fraction $d>1/2$ of
-  %meioses in female heterozygotes, inspective
-%  of the allele carried by the fertilizing sperm. 
-%Such an allele can invade the population over all of the parameter space, and will be maintained as a
-%  protected polymorphism in the white region, and can fix in the
-%  population in all of the parameter space colored green or orange. 
-%The solid back line marks the criteria for the allele to fix $s< (2d-1)/2$.
-%Now consider an allele that is fairly transmitted through
-%  female meiosis when fertilized by a sperm carrying a different
-%  allele, but at a fraction $d>1/2$ when a heterozygote egg is fertilized
-%  by a sperm carrying the allele (as depicted in Figure
-%  \ref{Eggsperm_2_allele_cartoon}). Such an allele can only invade in
-%  the region of parameter space colored orange, and it fixes over all
- % of this parameter region. The approximate criteria for this allele
-%  to fix is given by the dashed line showing $s < (2d- 1)/(4d)$. 
-% {\bf B.} In the inset we show trajectories of sperm-dependent female drive
- %alleles that can invade the population, each allele has $s =0.1$ against the
-% homozygotes and $d =0.6$, $0.7$, $0.8$ shown in yellow, blue, and red. 
-% \gc{Likely make this Figure 1B.} } \label{Invasion_space}
-%\end{figure}
-
 \subsection*{Two locus models of sperm-dependent female drive}
-
-%\subsection*{ sperm-egg haplotype balanced drive systems}
 
@@
         heterozygotes (see Figure \ref{Eggsperm_3_allele_cartoon}A2, and
-        Model 4 in the Appendix and Supplementary Worksheet). 
+        Model 4 in the Appendix and File S1 ). 
 Naively, $B^{+}$ may spread by capitalizing on the additional drive it causes; however,  
@@
 	three allele system that allows the sperm-acting drive facilitator $B^{+}$ to
-	invade the population (Appendix Model 4, eqn. \eqref{coupling}, and Mathematical Supplement).
+	invade the population (Appendix Model 4, eqn. \eqref{coupling}, and File S2  ).
 
@@
 \begin{center}
-	\includegraphics[width = 0.7
-          \textwidth]{Figures/twolocus.eps} \end{center}
+	\includegraphics[width = 0.7 \textwidth]{../Figures/twolocus.eps} 
+          \end{center}
 \caption{Models of a sperm-acting drive modifier tightly linked to a meiotic driver. 
@@
         background from the driver (A3 \& A4 respectively). 
-%		In {\bf{A3}} the drive suppressor exists on the drive background, forming the $B^-$ haplotype. 
-%		In {\bf{A4}} the drive suppressor exists on the non-drive background, forming the $A^-$ haplotype. 		
 {\bf{(B)}} Invasion of a sperm-acting drive suppressor linked to a driver. 
@@
 While  sperm enhancement of a female drive cannot displace a polymorphic female driver, sperm based drive suppressors can. 
-%	that is sperm-acting alleles that act to restore  fairness of female meiosis, can. %If these sperm-acting drive suppressors arise on
 Imagine a sperm-acting allele that restores fairness to female meiosis arises on
@@
 	sperm will evolve to oppose female meiotic drive.
-
-%More surprisingly, sperm-acting suppressors of female drive can spread
-%	even when tightly linked to the original drive system (Figures \ref{Eggsperm_3_allele_cartoon} and \ref{Trajectories_of_supressors}B). 
-%The $B^{-}$ haplotype (representing the tightly linked drive/drive suppressor complex) displaces the original driver 
-%	by driving when not costly, but avoiding low fitness homozygotes by disrupting drive when transmitted to sperm. %  ($BB^-$ and $B^-B^-$). 
-%Despite displacing the ancestral driver (the $B$ haplotype), the equilibrium frequency of the $B^-$ haplotype is often lower than that of $B$ \yb{(point to supp?)}.
-
-%\gc{Do we mention the In and Om locus here or in the discussion as a
-%	 potential e.g. Do we mention the fact that the frequency of these
-%	 newly invade sperm dodging alleles can be lower than the original system?}\\
 
@@
 Transmission rules describing the outcomes of all matings in each
-	model are presented in a Supplementary Worksheet. 
+	model are presented in a File S1 . 
 The fitness of genotype, $g$ is sex-independent and equals $1$, $1-h_s$, and $1-s$ for genotypes $AA$, $AB$, and $BB$, respectively. %  and  $w_{AA} = 1$, $w_{AB} = 1-s_h$, and $w_{BB}=1-s$. 
@@
 Because many of our analyses assume genotype frequencies expected under Hardy Weinberg Equilibrium (HWE), we note which analytical results are approximations. 
-We verify these approximations with exact numerical iterations in Supplementary Figures and Supplementary Scripts, and detail the derivation of these approximations in our Mathematical Supplement.
+We verify these approximations with exact numerical iterations in Figures S1-S3 and File S3 , and detail the derivation of these approximations in our File S2  .
 
@@
 After drive and random mating, genotype frequencies are 
-	$f_{AA}'=f_A (f_{AA }+ f_{AB} (1 - d))$, 
-	$f_{AB}'=f_A (f_{BB} + f_{AB}d) + f_B (f_{AA} + f_{AB} (1 - d))$,
-	and $f_{BB}'= f_B (f_{AB} d + f_{BB})$. 
-%Assuming fitnesses $w_{AA}=1$, $w_{AB}-1 s_h$ and $w_{BB}=1-s$, 
-%Frequencies of genotype, $i$, after drive and selection are 
-%	\begin{equation} f_i''= f_i'w_i/\bar{w}\end{equation}
-%	where $w_i$ is genotypic fitness, and mean fitness, $\bar{w} = \Sigma f_i'w_i$. 
+	$$f_{AA}'=f_A (f_{AA }+ f_{AB} (1 - d)),$$ 
+	$$f_{AB}'=f_A (f_{BB} + f_{AB}d) + f_B (f_{AA} + f_{AB} (1 - d)),\text{ and}$$
+	$$f_{BB}'= f_B (f_{AB} d + f_{BB}).$$ 
 As detailed above, exact frequencies after drive, random mating and selection are $f_g''= f_g'w_g/\bar{w}$. 
@@
 Therefore, genotype frequencies after drive are 
-	$f_{AA}'=f_A^2$, 
-	$f_{AB}'=f_A f_B+f_B (f_{AA} + f_{AB}(1 - d)) $,
-	and $f_{BB}'= f_B (f_{AB} d + f_{BB} )$. 
+	$$f_{AA}'=f_A^2 , $$
+	$$f_{AB}'=f_A f_B+f_B (f_{AA} + f_{AB}(1 - d)), \text{and} $$
+	$$f_{BB}'= f_B (f_{AB} d + f_{BB} ).$$
 %As above, exact frequencies after drive, random mating and selection are $f_i''= f_i'w_i/\bar{w}$. 
@@
 
-% the $B$ allele In our first scenerio allele $A$ is the ancestral allele and allele
-%$B$ is our sperm dependent drive allele.
-%The sperm dependence of the drive is a form of assortative
-%transmission (mating), and so we have to keep track of the genotype frequencies.
-%Our genotype frequencies at birth are $x_{AA},~x_{AB},$ and
-%$x_{BB}$. These three genotypes have fitnesses $w_{AA},~w_{AB},$ and
-%$w_{BB}$, which are identical between males and females. Because 
-%fitnesses are equivalent in both sexes, genotypic frequencies
-%are identical in males and females, but in the following
-%equations they are marked with $\Mars$ and $\Venus$ respectively to
-%allow the reader to more easily follow the logic. 
-%The frequency of the three genotypes at birth in the next generation is:
-%\begin{equation}
-%x'_{BB} =\frac{1}{\bar{w} } \Big( w_{BB} x_{BB}^{\Venus} (w_{BB} x_{BB}^{\Mars} +\frac{1}{2}
-%  w_{AB} x_{AB}^{\Mars} ) + w_{AB} x_{AB}^{\Venus} (d w_{BB}
-%  x_{BB}^{\Mars} +\frac{1}{2} d
-%w_{AB} x_{AB}^{\Mars} ) \Big),  \label{eqn.A1}
-%\end{equation}
-%\begin{equation}
-%x'_{AA} =\frac{1}{\bar{w} } \Big(  w_{AA} x_{AA}^{\Venus} (w_{AA} x_{AA}^{\Mars} + \frac{1}{2}
-%   w_{AB} x_{AB}^{\Mars}) + w_{AB} x_{AB}^{\Venus}(\frac{1}{2} w_{AA} x_{AA}^{\Mars}+ \frac{1}{2} 
-%  w_{AB} x_{AB}^{\Mars}) \Big),  \label{eqn.A2}
-%\end{equation}
-%and
-%\begin{eqnarray}
-%x'_{AB} = & \frac{1}{\bar{w} } \Big( w_{AA} x_{AA}^{\Venus}  ( w_{BB}  x_{BB}^{\Mars} + \frac{1}{2}
-%  w_{AB}  x_{AB}^{\Mars} ) +w_{BB} x_{BB}^{\Venus} ( w_{AA}  x_{AA}^{\Mars} + \frac{1}{2}
-%  w_{AB}  x_{AB}^{\Mars} )  \nonumber \\
-%& + x_{AB}^{\Venus}(\frac{1}{2}  w_{AA}  x_{AA}^{\Mars}+\frac{1}{2} (1-d)
-% w_{AB}  x_{AB}^{\Mars} + \frac{1}{4}  w_{AB}
- %x_{AB}^{\Mars} + (1-d)  w_{BB}  x_{BB}^{\Mars} ) \Big) \label{eqn.A3}
-%\end{eqnarray}
-%where the mean fitness of the population is
-%\begin{equation}
-%\bar{w} = w_{AA} x_{AA}+w_{AB}  x_{AB} +w_{BB}  x_{BB}.
-%\end{equation}
-%We can then iterate this recursion over generations, for
-%example to produce the frequency trajectories shown in the inset
-%Figure \ref{Invasion_space} by plotting $x_{BB}+ \frac{1}{2}x_{AB}$
-%over time. We can also use this to assess the
-%invasion criteria for the allele by starting allele B at very low
-%frequency, and seeing if it increases across the generations.
-
-\paragraph{Recessive fitness cost of self-promoting driver:} %When the self-promoting driver has a recessive fitness cost, $s$ genotype frequencies after selection and drive equal
-%\begin{equation}	x'_{BB} =\frac{x_{B}^2(1-s)}{\bar{w} }  \Big( F+(1-F)(x_{B}+2x_{A} d) \Big) 	\end{equation}
-%\begin{equation}	x'_{AA} =\frac{1}{\bar{w} }  \Big(x_{A}^2\Big) 	\end{equation}
-%and
-%\begin{equation}	x'_{AB} =\frac{2 x_{A}x_{B} }{\bar{w} }  \Big(1-x_{B}(d-1/2)(1-F) \Big)	%\end{equation}
+\paragraph{Recessive fitness cost of self-promoting driver:} 
 When fully recessive, the change in frequency of the self-promoting driver across generations equals 
@@
 \end{equation}
-where $F$ is the deviation from genotypic frequencies expected under Hardy-Weinberg. % Trying to clarify some.   \gc{Hardy Weinberg Expectations} (HWE). 
-%For analytical solutions presented above, we assume that $F$ equals zero. 
+where $F$ is the deviation from genotypic frequencies expected under Hardy-Weinberg. 
 Assuming HWE  ($F=0$) a common, recessive, self-promoting driver invades if $s\lessapprox(2 d - 1)/(4 d)$, and fixes if
@@
  but it will be lost when introduced below this frequency. 
- %(i.e. this is an unstable equilibrium point). 
-% In Supplementary Figure \ref{Bistab_homozyg_cost_fig} we show
-% this approximation condition for bistability andd how it compared to
-% that obtained using the recursion. 
-Compared to exact results (Supplementary Figure \ref{Bistab_homozyg_cost_fig}), Equations
+Compared to exact results (Figure S\ref{Bistab_homozyg_cost_fig}), Equations
  \eqref{ineqA} and \eqref{bistab_homozy_approx} offer reasonable, but imperfect approximations.
-%s_h and seems to serve as a lower bound to the true unstable equilibrium frequency.
 
-\paragraph{Cost of driver in heterozygotes:} %We explored a more complex model where the self-promoting driver has a cost, $s_h$ in heterozygotes (so $w_{AB} = 1-s_h$) in addition to the cost $s$, in homozygotes.
+\paragraph{Cost of driver in heterozygotes:} 
 When the fitness of drive heterozygotes is compromised ($s_h>0$), a self-promoting driver cannot invade when rare.
@@
 Comparison of Equation \ref{thresh2} to exact results obtained by a
-simple parameter search (Supplementary Figure \ref{bistable_additive})
+simple parameter search (Figure S\ref{bistable_additive})
 show that this approximation is reasonably correct for small
 parameter values; however, it underestimates $f_B^*$ for large parameter values, presumably because they result in strong departures from HWE.
- %range of other costs to the drive alleles other than a
-%simple recessive cost, for example an additive cost. We modified the recursions given in eqn
-%\ref{eqn.A1}-\ref{eqn.A3} to feature an additive cost $s$ for each
-%copy of the $B$ allele. A none recessive cost to the drive allele can
-%cause bistability of the drive allele. The B allele is unable to enter
-%the population when rare, this results from the fact that
-%when the allele is rare ($x_B \ll 1$) sperm carrying it rarely
-%encounter heterozygous eggs ($\sim x^2$) but the allele still pays a
-%cost in heterozygous individuals.  By contrast, when the B allele is common ($x_B
-%\gg 0$) encounters between heterozygous eggs and $B$ allele sperm
-%become more common allowing drive to occur this can allow the $B$
-%allele to spread when common. In Supplementary Figure
-%\ref{bistable_additive} we show the unstable equilibrium frequency
-%above which the $B$ allele can invade. This equilibrium frequency was found by a simple
-%parameter search.  \gc{wonder if a bit of work could find this}
 
@@
 In the case when female meiotic drive depends on paternal genotype,
-%\paragraph{Dominance} we also explored a model where
-%the genotype of the male who produced the fertilizing sperm affects female drive,
-%instead of just the allelic type of the sperm.
- %We assume a 
  a heterozygous female will transmit the $B$ allele 
   with probabilities  $\frac{1}{2}$,  $d_{het}\geq 1/2 $ or $d_{hom}\geq d_{het}$, 
-%YB RE-PARAMETERIZING  %  $\frac{1}{2}$,  $\frac{1}{2} + h(d-\frac{1}{2}) $ or $d$,  YB RE-PARAMETERIZING 
  when mated with to $AA$, $AB$, or $BB$ males,  respectively. 
   In this model, genotype frequencies after drive, random mating, and selection are 
-  	$f_{AA}'' =   w_{AA}\left( f_{AA} [2 f_A + f_{AB} ] + f_{AB}^2 [1 - d_{het}] \right)/(2\bar{w})$,
-	$f_{AB}'' =   w_{AB}(f_{AA} [f_B + f_{AB}/2] + f_{AB} f_B + f_{BB} [ f_A - f_{AB} d_{hom}])/\bar{w}$, 
-	and 	$f_{BB}'' =   w_{BB}(f_{AB} (f_{AB} d_{het}/2 + f_{BB} d_{hom}) + f_{BB} f_B)/\bar{w}$. 
-% Above, $d > 1/2$ and $h$ ($0 \geq h \geq 1$) acts to specify the dominance of the $B$
-%allele's effect on female meiotic drive through the males genotype.
-%YB TRY_1 We assume that in a mating between a heterozygous female and an $AA$, $AB$, or $BB$ male, the female will transmit the $B$ allele 
-% with probabilities  $\frac{1}{2}$,  $\frac{1}{2} + h(d-\frac{1}{2}) $, and $d$, respectively.
-  %Here $d > 1/2$ and $h$ ($0 \geq h \geq 1$) acts to specify the dominance of the $B$
-%allele's effect on female meiotic drive through the males genotype.
-% GC ORIGINAL  We assume that if a
-%heterozygous egg ($AB$) is fertilized by a sperm with the following
-%male ($\Mars$) genotypes the $B$ allele will be transmitted through
-%female meiosis with the following probabilities \\
-%\begin{tabular}{cccc}
-%$\Mars$ geno. & AA & AB  & BB\\
-%P(B). & $\frac{1}{2}$ &  $\frac{1}{2} + h(d-\frac{1}{2}) $ & $d$ 
-%\end{tabular}\\
-%where $d > 1/2$ and $h$ ($0 \geq h \geq 1$) acts to specify the dominance of the $B$
-%allele's effect on female meiotic drive through the males genotype. 
-%We set up a set of recursions using a recessive sex independent cost to
-%the driver, similar to those in eqn.
-%\ref{eqn.A1}-\ref{eqn.A1}, but now featuring these male genotype
-%transmission probabilities.
+  	$$f_{AA}'' =   w_{AA}\left( f_{AA} [2 f_A + f_{AB} ] + f_{AB}^2 [1 - d_{het}] \right)/(2\bar{w}), $$
+	$$f_{AB}'' =   w_{AB}(f_{AA} [f_B + f_{AB}/2] + f_{AB} f_B + f_{BB} [ f_A - f_{AB} d_{hom}])/\bar{w}, \text{ and}$$ 
+	$$f_{BB}'' =   w_{BB}(f_{AB} (f_{AB} d_{het}/2 + f_{BB} d_{hom}) + f_{BB} f_B)/\bar{w}.$$
 
@@
 	$s\lessapprox (d_{het}-1/2)/(d_{het})$, and when common, this driver fixes if 
-	$s\lessapprox d_{hom}-1/2$, approximations well supported by exact results (Supplementary Figure \ref{Effect_of_dominance}).
+	$s\lessapprox d_{hom}-1/2$, approximations well supported by exact results (Figure S\ref{Effect_of_dominance}).
 Specifically, when drive in heterozygotes is large relative to that in homozygotes, 
@@
 	and an increase in frequency of the $C$ allele when it suppresses drive ($\epsilon <0$ -- the $B^-$ model, above). 
-More generally, even when the cost of drive is not fully recessive, the $B^-$ allele will invade and fix under all parameters sustaining a previous polymorphism at the drive locus (see Mathematical Supplement).
-%When the C allele acts to suppress drive for all parameters sustaining a polymorphism at the drive locus ($s>d_0-1/2$)
-%We address two questions concerning the $C$ allele - can it invade when rare  and can it fix when common?
-%Although our analytical approximations (see our \gc{Mathematical supplement}) do not reveal simple inequalities like those above,
-%	graphical exploration of analytical results show that this mutant is lost when it enhances drive ($\epsilon>0$ - the $B^+$ allele, above),
-%	and spreads from rarity to fixation when it suppresses drive ($\epsilon<0$ - the $B^-$ allele, above).	
-%This result is supported by exact numerical iterations, as well
-%\yb{FIG?}. \gc{Lets not bother}
+More generally, even when the cost of drive is not fully recessive, the $B^-$ allele will invade and fix under all parameters sustaining a previous polymorphism at the drive locus (see File S2  ).
 
@@
 For all values of interest ($0<s<1$, $0.5<d_0<1$), the change in frequency a rare $C$ allele is positive when it decreases drive (i.e.
-$\epsilon <0$, corresponding to the $A^-$ model, above), a result which holds qualitatively for a common $C$ allele, as well (Mathematical Supplement).
-%We address two questions concerning the $C$ allele - can it invade when rare  and can it fix when common?
-%Again, while simple inequalities are lacking, 
-%	graphical exploration of our analytical results, and exact numerical iteration \yb{FIG?} shows that this modifier only spreads from rarity to fixation when it suppresses drive ($\epsilon<0$ - the $A^-$ allele, above).
+$\epsilon <0$, corresponding to the $A^-$ model, above), a result which holds qualitatively for a common $C$ allele, as well (File S2  ).
 Somewhat counterintuitively, the spread of the $A^-$ allele is substantially slower than that of the $B^-$ allele (it is on the order of $C^{2}$ rather than $C$). 
@@
 \begin{equation} \Delta M =-\epsilon f_A f_B (f_B s + s_h (f_A - f_B) ) \frac{f_M}{\bar{w}}  \label{unlinked} \end{equation}
-% an $F$ allele increases in frequency when 
-%	$-\epsilon f_A f_B (f_B s + s_h (fA - fB) ) >0$, corresponding to any level of drive suppression ($\epsilon<0$) 
 Thus, a rare drive suppressor ($\epsilon<0$) will spread so long as the fitness cost of the driver does not display over- or under-dominance. 
@@
 ][]{Brandvain2012}.
-%The spread of this driver does not, however, depend on assumption of HWE, and linkage equilibrium - exact results resemble analytical approximations. 
-%\paragraph{three allele system}
-%$x_{AA}$, $x_{AB}$,$x_{AB'}$, $w_{BB'} x_{BB'}$, $w_{B'B'} x_{BB}$, $w_{B'B'} x_{B'B'}$
-%We denote the frequency of the three genotypes in the population at birth are
-%$x_{AA}$, $x_{AB}$, $x_{AB'}$, $x_{BB'}$, $x_{BB}$, $x_{B'B'}$. 
-%The mean relative fitness of the population is 
-%\begin{equation}
-%x_{AA}+x_{AB}+ x_{AB'}+ w_{BB'} +x_{BB'}+w_{B'B'}x_{BB} + w_{B'B'} x_{B'B'} 
-%\end{equation}
-%The frequency of the genotypes in meiotic-cells of female adults is 
-%\begin{equation}
-%\frac{1}{\bar{w}}\Big( x_{AA}, ~x_{AB},~ x_{AB'},~ w_{BB'} x_{BB'},~ w_{B'B'}x_{BB},
-%~w_{B'B'} x_{B'B'} \Big)
-%\end{equation}
-%The frequency of fertilization events can be found by combining these
-%female frequencies with the frequency of the three alleles in sperm is
-%\begin{eqnarray}
-%x_A = & x_{AA} + \frac{1}{2} x_{AB}+ \frac{1}{2} x_{AB'} \\
-%x_B = & w_{BB}  x_{BB} +   w_{BB'} \frac{1}{2} x_{BB'}+ \frac{1}{2} x_{AB} \\
-%x_{B'} = & w_{B'B'}  x_{B'B'} +   w_{BB'} \frac{1}{2} + x_{BB'}+ \frac{1}{2} x_{AB'}  
-%\end{eqnarray}
-%in combination with the transmission probabilities given in Figure
-%\ref{Eggsperm_3_allele_cartoon}A.
 
-% \gc{Alternatively we could give the table below:}
-
-% \begin{table}
-% \begin{tabular}{lllll}
-% \hline 
-%   &  \multicolumn{4}{c}{Sperm genotype}   \\
-%   &  & A & B & B'  \\
-% $\Venus$ geno. & Freq.  & $x_{AA} + \frac{1}{2} x_{AB}+ \frac{1}{2} x_{AB'} $ & $ w_{BB}  x_{BB} +   w_{BB'} \frac{1}{2} x_{BB'}+
-% \frac{1}{2} x_{AB}$ & $ w_{B'B'}  x_{B'B'} +   w_{BB'} \frac{1}{2}
-% x_{BB'}+ \frac{1}{2} x_{AB'}  $\\
-% \hline\\
-% AA & $x_{AA}$ &  & & \\
-% AB & $x_{AB}$ &  & & \\
-% AB' & $x_{AB'}$ &  & & \\
-% BB' & $w_{BB'} x_{BB'}$ &  & & \\
-% BB & $w_{B'B'} x_{BB}$ &  & & \\
-% B'B' & $w_{B'B'} x_{B'B'}$ &  & & \\
-% \end{tabular}
-% \caption{ALTERNATIVELY I COULD FILL IN THIS MATRIX  with genotype }
-% \end{table}
 \bibliography{refs}
@@
 
+{\bf{File S1:}} Transmission rules. We detail the frequency of offspring genotypes produced by each possible mating for Models 1-6 in the tabs of this Excel spreadsheet.
+
+{\bf{Files S2A \& S2B:}} A Mathematica file (FileS2A) and a PDF of this file (File S2B) in which we derive analytical results for models 1-6.
+
+{\bf{File S3:}} Exact approach. The R Script used for exact recursions.
+\clearpage
 \begin{figure}
 
-\includegraphics[width = \textwidth]{../Scripts/homozyg_bistability.eps}
+\includegraphics[width = \textwidth]{../../Scripts/homozyg_bistability.eps}
 \caption{Invasion analysis for a self-promoting female meiotic drive allele with
@@
 \begin{figure}
-\includegraphics[width = 0.8 \textwidth]{Figures/bistable_x_vs_d_additive_s.eps} 
+\includegraphics[width = 0.8 \textwidth]{../Figures/bistable_x_vs_d_additive_s.eps} 
 \caption{The unstable equilibrium frequency for a self-promoting
@@
 \begin{figure}
-\includegraphics[width = 0.7 \textwidth]{../Scripts/effect_of_dominance_on_invasion_space_four_graph.eps} 
+\includegraphics[width = 0.7 \textwidth]{../../Scripts/effect_of_dominance_on_invasion_space_four_graph.eps} 
 \caption{Exact results of invasion analysis of an allele whose effect

### 2014-07-14 writeup_and_notes/GeneticsResub/male_dependent_drive.tex
first go at revisions by YB next do some math and sims for female vs male suppression of female drive

@@
 	interpreted as mechanisms of protection against genomic outlaws. 
-Why then is female meiosis often left uncompleted until after
-fertilization in many animals -- 
-	potentially providing an opportunity for sperm alleles to
-        meddle with its outcome and help like-alleles drive in heterozygous females? 
+\yb{The fact that female meiosis often left uncompleted until after
+      fertilization in many animals  appears to run counter to this interpretation, 
+      as this delay provides an opportunity for sperm alleles to
+     meddle with its outcome and help like-alleles drive in heterozygous females. } 
+%Why then is female meiosis often left uncompleted until after
+%fertilization in many animals -- 
+%	potentially providing an opportunity for sperm alleles to
+%     meddle with its outcome and help like-alleles drive in heterozygous females? 
 The population genetic theory presented herein suggests that
@@
 These results are consistent with current knowledge of sperm-dependent meiotic drivers 
-(loci whose distortion of female meiosis depends on sperm genotype),
-	and suggest that the requirement of fertilization for the
-        completion of female meiosis
-        potentially represents a mechanism employed by females 
-	 to ensure a fair meiosis.
+(loci whose distortion of female meiosis depends on sperm genotype). 
+\yb{Rather than providing an opportunity for sperm collaboration in female meiotic drive,
+        the requirement of fertilization for the completion of female meiosis
+        likely indirectly provides protection to females against meiotic drivers as it provides an opportunity for sperm to suppress female meiotic drive.}
+%	and suggest that the requirement of fertilization for the
+%%        completion of female meiosis
+%        potentially represents a mechanism employed by females 
+%	 to ensure a fair meiosis.
 \newpage
@@
 \citep{Mira1998}, the structure of centromere machinery \citep{Malik2002a,Malik2009}, and sex differences in the recombination rate \citep{Haig2010,Brandvain2012} 
-have perhaps evolved to enforce fairness by disrupting meiotic drive \citep{Rice2013}.
+have perhaps evolved to enforce fairness by disrupting meiotic drive \citep{Rice2013}. \newline
 
@@
      could presumably rapidly spread through a population. 
-At first sight, female meiosis appears primed for exploitation by such selfish systems.  
+At first sight, female meiosis appears primed for exploitation by such selfish systems, \yb{and here we ask if spam do indeed evolve to collaborate with female drivers to exploit this apparent weakness in the defense against meiotic drive.} \newline
 %Why then is the requirement of fertilization to complete female meiosis so ubiquitous?
 
-%It is certainly not the case that animals are mechanistically incapable of circumventing this requirement.
+\yb{The threat to meiotic fairness posed by the fertilization requirement for completion of female meiosis seems both real and avoidable. 
+We say this threat is avoidable because the stage at which female meiosis requires fertilization 
+	varies widely across taxa, and in a few animals female meiosis is complete upon ovulation \
+	(that is, fertilization is not required for female meiosis \citep[e.g. Table 1 in ][]{Masui_book}). 
+We say this threat is real because there is substantial evidence that female meiotic drive is an 
+	evolutionary force and considerable opportunity for sperm to influence female drive.  
+While research to date has identified a few of what are likely the most extreme female meiotic 
+	drivers \citep{Agulnik1990, Fishman2008, Hiatt2003,Novitski1951, Pardo-ManuelDeVillena2001c}, 
+	rapid evolution of centromeres, telomeres and other basic components of the meiotic machinery, 
+	suggest consistent evolution of meiotic drivers and suppressors of meiotic drive
+	 \citep[e.g. ][]{Anderson2008, Anderson2009, Axelsson2010, Malik2009a}, and the inexpense of 
+	 genotyping-by-sequencing platforms for non-model systems will likely 
+	identify subtle meiotic drivers in numerous non-model systems.  
+Both mechanistic and genetic evidence suggests that sperm can influence the outcome of female meiosis.
+An example of mechanistic evidence for sperm's influence on female meiosis comes from \emph{C.
+elegans}, where premature deployment of the aster -- a vital 	component of mitotic machinery provided by the sperm,
+	disrupts MII meiotic segregation in the egg, leading to a triploid zygote \citep{McNally2012}.
+{\bf{GRAHAM DO MORE????} } 
+Genetic evidence from the \emph{In} and \emph{Om}  drive systems in mice suggests that the 
+	 transmission of alternate alleles in heterozygous females can depend on sperm haplotype.  
+ While ruling out the alternative hypothesis of early selection on zygotes in these cases is challenging  \citep[see pages
+	52-54 in][]{Burt2006}, it appears that the extent to which \emph{In} and \emph{Om} distort the second female meiotic division partially depends on the 	
+	genotype of the fertilizing sperm \citep{Agulnik1993,Wu2005}. 
+}\newline
+
 %The stage at which female meiosis requires fertilization varies widely across taxa, and in a few animals  
@@
 Thus, there is little selective benefit to females in preventing sperm to influence female meioses. 
-\yb{While, we do not claim that female meiotic arrest evolved specifically to recruit helpful sperm to defend against meiotic drive, the delayed completion of female meiotic drive will not in fact encourage sperm to assist renegade drivers.}
+\yb{While we do not claim that female meiotic arrest evolved specifically to recruit helpful sperm to defend against meiotic drive, the delayed completion of female meiotic drive will not in fact encourage sperm to assist renegade drivers.}
 %Rather, features of o\"{o}genesis  may evolve to facilitate an interaction with sperm %with female meiosis
@@
 
+\yb{
+\section*{Methods}
+We present deterministic one- and two- locus population genetic models of sperm influence on female meiotic drive to evaluate wether sperm are likely to collaborate with female meiotic drivers or to stop them.
+\newline
+
+We present six related models, the first three of which represent the system with a single locus.
+{\bf{Model 1}} describes a traditional, single-locus, female meiotic driver, 
+{\bf{Model 2}} describes a single-locus sperm-dependent female driver -- that is, an allele whose transmission in female meiosis depends on sperm haplotype. 
+{\bf{Model 3}} describes a single-locus paternal-dependent female driver -- an allele whose transmission in female meiosis depends on paternal genotype. 
+Assuming that a traditional driver segregates at its equilibrium frequency ({\bf{Model 1}}), we investigate the evolution of tightly linked ({\bf{Models 4 and 5}}), and unlinked ({\bf{Model 6}}) sperm-dependent modifiers of drive.
+In {\bf{Models 4 and 5}}, we treat this two-locus system as if it consists of
+a single locus with three alleles: $A$, $B$ and $C$, corresponding to the case when the
+sperm-modifier arises once, tightly linked to one or other of the
+alleles at the drive locus (such that recombination is unexpected). 
+In all cases, we assume that fitness is independent of the drive modifier. 
+ \newline
+
+All models include a biallelic locus ($A/B$) with non-driving and driving alleles in frequencies $f_A$ and $f_B$, 
+	respectively, while more complex models include more loci and/or alleles.  
+Transmission rules describing the outcomes of all matings in each
+	model are presented in a File S1 . 
+The fitness of genotype, $g$, is sex-independent and equals $1$, $1-h_s$, and $1-s$ for genotypes $AA$, $AB$, and $BB$, respectively. %  and  $w_{AA} = 1$, $w_{AB} = 1-s_h$, and $w_{BB}=1-s$. 
+Genotypic frequencies  equal  $f_g$ for adults in the current generation, 
+	$f_g'$ in the next generation of zygotes (i.e. after recombination, random mating, drive, and syngamy, but before selection)
+	and $f_g''$ in adults after selection. 
+After a complete generation $f_g'' = f_g' w_g/\bar{w}$, where $\bar{w}$ is the population mean fitness and equals $\Sigma w_g f_g'$.  \newline
+
+We verbally describe our main results below.  
+Readers interested in more quantitative results should turn to the Appendix for a mathematical treatment. 
+Here we present critical analytical results in Equations 1-10, and describe our analyses and results in more detail.  
+Because many of our analyses assume genotype frequencies expected under Hardy Weinberg Equilibrium (HWE), we note which analytical results are approximations. 
+We verify these approximations with exact numerical iterations in Figures S1-S3 and File S3 , and detail the derivation of these approximations in our File S2.
+}
+
 \section*{Results}
 
-\subsection*{ Invasion of the population by a driving allele that promotes itself.}
+\subsection*{Invasion of the population by a driving allele that promotes itself.}
 %%%%%%%%%%%%%%%%%%%%%%%%%
@@
         carrying that allele (see Figure \ref{Eggsperm}A and Model 2
-        in the Appendix and File S1 ). 
+        in the Appendix and File S1). 
 We then identify the conditions allowing for the spread of this self-promoting driver, 
@@
 drivers influence their transmission in females via paternal genotype,
-rather than sperm haplotype.
+rather than sperm haplotype. \newline
 
@@
 	drive suppressors, corresponding well to empirical examples of
-        female meiotic drive \citep[reviewed in ][]{Burt2006}.
+        female meiotic drive \citep[reviewed in ][]{Burt2006}. \newline
 
@@
 This analytical approximation, derived from Equation \eqref{deltadriver} assuming Hardy-Weinberg, 
-	closely matches results obtained by exact numerical iteration (Figure \ref{Eggsperm}B).
+	closely matches results obtained by exact numerical iteration (Figure \ref{Eggsperm}B.
+\yb{We note that Equation 1 and all equations discussed in the main text are presented in the appendix and derived in File S2}).
+\newline
 
@@
 	and therefore  the fate of a self-promoting driver is determined
-	by the more restrictive invasion criteria rather than the fixation criteria.
+	by the more restrictive invasion criteria rather than the fixation criteria. \newline
 
@@
 	precludes the evolution of an
-	allele that suppresses this form of meiotic drive in females.
+	allele that suppresses this form of meiotic drive in females. \newline
 
@@
 	the lack of biological reality underlying the small portion of parameter values 
-	required for this polymorphism make us doubt its general applicability.
+	required for this polymorphism make us doubt its general applicability. \newline
 
@@
 it seems very unlikely that such alleles could drive the evolution of female suppressors of sperm-enabled
-female meiotic drive.
+female meiotic drive. \newline
 
@@
 Subsequently, a sperm-acting modifier of female meiotic drive arises at another locus.  
+In these two-locus models, the driver is transmitted to $d_0$ of gametes from female heterozygotes when fertilized by wild-type sperm, and $d_1=d_0+\epsilon$ when fertilized by a sperm-acting drive-modifier.
+\newline
 
@@
 	in Model 5 the modifier arises tightly linked to the non-driving allele, 
-	and in Model 6 it is unlinked to the driver.
+	and in Model 6 it is unlinked to the driver. \newline
 
@@
 	three allele system that allows the sperm-acting drive facilitator $B^{+}$ to
-	invade the population (Appendix Model 4, eqn. \eqref{coupling}, and File S2  ).
+	invade the population (Appendix Model 4, eqn. \eqref{coupling}, and File S2  ). \newline
 
@@
 Therefore, all two-locus models of sperm influence on female drive suggest that
-	sperm will evolve to oppose female meiotic drive.
+	sperm will evolve to oppose female meiotic drive. \newline
 
@@
 	and conflicts between alleles that are over-transmitted in meiosis and the organisms they injure while doing so. 
-Such conflicts and their resolution likely play a major role in the structure and evolution of many basic biological processes \citep{Rice2013}.
+Such conflicts and their resolution likely play a major role in the structure and evolution of many basic biological processes \citep{Rice2013}. \newline
 
@@
 	without risking exploitation by collaborations between female
-        drivers and sperm alleles.
+        drivers and sperm alleles. \newline
 
@@
 More generally, natural selection favors alleles that act through sperm to
- reduce the opportunity of female meiotic drive regardless of linkage or phase.
+ reduce the opportunity of female meiotic drive regardless of linkage or phase. \newline
 
-Our theory offers a novel interpretation of a baffling idiosyncrasy of female meiosis in many animals -- the requirement of fertilization before meiosis is complete. 
-We argue that females do not complete meiosis until after fertilization as a mechanism to encourage meiotic fairness.
-In addition to providing a new interpretation of a basic biological process, our theory provides numerous testable predictions.
-We predict that female drive, at a stage influenceable by sperm,  
-	should be reduced when fertilized by
-	complementary sperm. 
-Therefore, MII female drive should be exceptionally severe in
-	interspecific crosses between animals which complete the second meiotic division after fertilization,  
-	as heterospecific sperm will not specifically evolve to resist a female driver in another species. 
-Our theory also encourages phylogenetic hypotheses concerning the relationship between the opportunity 
-	for female meiotic drive and the requirement of fertilization for the completion of female meiosis. 
-For example, we predict that a lower opportunity for female meiotic drive,
-        e.g. an animal lineage with a long history of high inbreeding
-        or selfing, 
-        would often be accompanied by a relaxation of the requirement of fertilization for the
-        completion of female meiosis.
+%Our theory offers a novel interpretation of a baffling idiosyncrasy of female meiosis in many animals -- the requirement of fertilization before meiosis is complete. 
+%We argue that females do not complete meiosis until after fertilization as a mechanism to encourage meiotic fairness.  
+%In addition to providing a new interpretation of a basic biological process, our theory provides numerous testable predictions.
+
+\yb{The theory presented herein provides a strong prediction -- in animals and other organisms in which female meiosis is not completed until after fertilization sperm will act to suppress female drive at the stage at which they can influence meiosis.
+%Therefore, for example, when fertilized by drying sperm, we expect meiosis to be fairer when  females are fertilized by 
+contrary to expectations of a ``green-beard'' model, we are unsurprised by the observation that female meiosis is fairer when fertilized by sperm bearing the drive allele in two of the best described cases of female meiotic drive in animals (the \emph{Om} and \emph{In} loci in mice \citep{Agulnik1993,Wu2005}).
+Given the immense benefit to sperm of hampering female drive, drive-suppressing sperm are often likely to be fixed within a species, making the hypothesis of sperm acting drive-suppression difficult to test from intra-population crosses.
+However, crosses between populations or species are likely to provide critical tests of our theory -- specifically we predict that female meiosis will be less fair when a species (or population) is fertilized by heterospecific sperm because such sperm have not evolved to counter novel females meiotic drivers.
+We further predict that segregation in F1 females backcrossed to parental species will likely be biased with a deficit of transmission of the paternal species allele from the F1  female.   
+We note that while these predictions follow straightforwardly from the theory presented above, tests of meiotic drive, and especially sperm dependent meiotic drive require a high standard of evidence to exclude plausible alternative hypotheses such as genotypic inviability including epistatic maternal X zygotic lethality.}  \newline
+
+%We predict that female drive, at a stage influenceable by sperm,  
+%	should be reduced when fertilized by
+%	complementary sperm. 
+%Therefore, MII female drive should be exceptionally severe in
+%	interspecific crosses between animals which complete the second meiotic division after fertilization,  
+%	as heterospecific sperm will not specifically evolve to resist a female driver in another species. 
+%Our theory also encourages phylogenetic hypotheses concerning the relationship between the opportunity 
+%	for female meiotic drive and the requirement of fertilization for the completion of female meiosis. 
+%For example, we predict that a lower opportunity for female meiotic drive,
+%        e.g. an animal lineage with a long history of high inbreeding
+%        or selfing, 
+%        would often be accompanied by a relaxation of the requirement of fertilization for the
+%        completion of female meiosis. \newline
 
@@
 	fertilizing sperm decidedly side with mom, as both have a shared interest in producing a viable and potentially fit offspring. 
+\yb{Our model does not directly speak to the evolutionary origin of female meiotic arrest (for a review and evaluation of such hypotheses see \citep{Mira1998}); however, our work does imply that the fertilization requirement does not leave female meiosis open to sabotage by sperm collaborators, but rather that  an evolutionary loss of female meiotic arrest will result in the  loss of sperm defense against female meiotic drivers.
+}
 
@@
 
-\section*{Appendix}
-We highlight simple analytical results from: 
-	the standard model of female meiotic drive (Model 1); a sperm-dependent 
-	model of female meiotic drive (Model 2); a male-genotype-dependent model of
-	female meiotic drive (Model 3); and more complex two-locus sperm-dependent drive models (Models 4-6). 
-All models include a biallelic locus ($A/B$) with non-driving and driving alleles in frequencies $f_A$ and $f_B$, 
-	respectively, while more complex models include more loci and/or alleles.  
-Transmission rules describing the outcomes of all matings in each
-	model are presented in a File S1 . 
-The fitness of genotype, $g$ is sex-independent and equals $1$, $1-h_s$, and $1-s$ for genotypes $AA$, $AB$, and $BB$, respectively. %  and  $w_{AA} = 1$, $w_{AB} = 1-s_h$, and $w_{BB}=1-s$. 
-Genotypic frequencies  equal  $f_g$ for adults in the current generation, 
-	$f_g'$ in the next generation of zygotes (i.e. after recombination, random mating, drive, and syngamy, but before selection)
-	and $f_g''$ in adults after selection. 
-After a complete generation $f_g'' = f_g' w_g/\bar{w}$, where $\bar{w}$ is the population mean fitness and equals $\Sigma w_g f_g'$. 
-Because many of our analyses assume genotype frequencies expected under Hardy Weinberg Equilibrium (HWE), we note which analytical results are approximations. 
-We verify these approximations with exact numerical iterations in Figures S1-S3 and File S3 , and detail the derivation of these approximations in our File S2  .
+%\section*{Appendix}
+%We highlight simple analytical results from: 
+%	the standard model of female meiotic drive (Model 1); a sperm-dependent 
+%	model of female meiotic drive (Model 2); a male-genotype-dependent model of
+%	female meiotic drive (Model 3); and more complex two-locus sperm-dependent drive models (Models 4-6). 
+%All models include a biallelic locus ($A/B$) with non-driving and driving alleles in frequencies $f_A$ and $f_B$, 
+%	respectively, while more complex models include more loci and/or alleles.  
+%Transmission rules describing the outcomes of all matings in each
+%	model are presented in a File S1 . 
+%The fitness of genotype, $g$ ,is sex-independent and equals $1$, $1-h_s$, and $1-s$ for genotypes $AA$, $AB$, and $BB$, respectively. %  and  $w_{AA} = 1$, $w_{AB} = 1-s_h$, and $w_{BB}=1-s$. 
+%Genotypic frequencies  equal  $f_g$ for adults in the current generation, 
+%	$f_g'$ in the next generation of zygotes (i.e. after recombination, random mating, drive, and syngamy, but before selection)
+%	and $f_g''$ in adults after selection. 
+%After a complete generation $f_g'' = f_g' w_g/\bar{w}$, where $\bar{w}$ is the population mean fitness and equals $\Sigma w_g f_g'$. 
+%Because many of our analyses assume genotype frequencies expected under Hardy Weinberg Equilibrium (HWE), we note which analytical results are approximations. 
+%We verify these approximations with exact numerical iterations in Figures S1-S3 and File S3 , and detail the derivation of these approximations in our File S2.
 
@@
 As detailed above, exact frequencies after drive, random mating and selection are $f_g''= f_g'w_g/\bar{w}$. 
-Assuming HWE, a rare driver will spread when $s_h
-        \lessapprox(2d-1)/(1+2d)$, and will fix when $s\lessapprox d -1/2 +
-        3 s_h /2- d s_h$. This later inequality reduces to
-        $s\lessapprox(2d-1)/2$ when the cost of drive is fully
+Assuming HWE, a rare driver will spread when $\left(s_h
+        \lessapprox \left(2d-1 \right)/\left(1+2d\right)\right)$, and will fix when $\left( s\lessapprox d -1/2 +
+        3 s_h /2- d s_h \right) $. This later inequality reduces to
+        $\left( s\lessapprox\left(2d-1\right/2 \right)$ when the cost of drive is fully
         recessive.
@@
 where $F$ is the deviation from genotypic frequencies expected under Hardy-Weinberg. 
-Assuming HWE  ($F=0$) a common, recessive, self-promoting driver invades if $s\lessapprox(2 d - 1)/(4 d)$, and fixes if
+Assuming HWE  ($F=0$) a common, recessive, self-promoting driver invades if $\left(s\lessapprox\left(2 d - 1\right)/\left(4 d\right)\right)$, and fixes if
  $s\lessapprox(2d-1)/2$. 
@@
  but it will be lost when introduced below this frequency. 
-Compared to exact results (Figure S\ref{Bistab_homozyg_cost_fig}), Equations
+Compared to exact results (Figure \ref{Bistab_homozyg_cost_fig}), Equations
  \eqref{ineqA} and \eqref{bistab_homozy_approx} offer reasonable, but imperfect approximations.
@@
 \end{equation}
-where, $a_1=1-2 ds_h+4 d s-2 d-3 s_h$ and $a_2=2(s-s_h)(2d-1)$.
+where, $a_1=\left(1-2 ds_h+4 d s-2 d-3 s_h\right)$ and $a_2=\left(2\left(s-s_h\right)\left(2d-1\right)\right)$.
 Comparison of Equation \ref{thresh2} to exact results obtained by a
-simple parameter search (Figure S\ref{bistable_additive})
+simple parameter search (Figure \ref{bistable_additive})
 show that this approximation is reasonably correct for small
@@
 	a rare paternal-genotype-dependent driver invades when 
-	$s\lessapprox (d_{het}-1/2)/(d_{het})$, and when common, this driver fixes if 
-	$s\lessapprox d_{hom}-1/2$, approximations well supported by exact results (Figure S\ref{Effect_of_dominance}).
+	$\left(s\lessapprox \left(d_{het}-1/2\right)/d_{het}\right)$, and when common, this driver fixes if 
+	$\left(s\lessapprox d_{hom}-1/2\right)$, approximations well supported by exact results (Figure \ref{Effect_of_dominance}).
 Specifically, when drive in heterozygotes is large relative to that in homozygotes, 
@@
 \begin{equation}f_{B\text{ pat}}^*\approx2 (1 + 2d_{hom}[s - 1])/([2 s - 1][2d_{hom}- 1]) \label{eqmale} \end{equation}
-By contrast, when $d_{het} \lessapprox 1/(3-2d_{hom})$ the case is reversed, and the model is bistable.
+By contrast, when $\left(d_{het} \lessapprox 1/\left(3-2d_{hom}\right)\right)$ the case is reversed, and the model is bistable.
 
 \subsection*{Models 4-6. Two-locus, sperm-dependent drive.}
-Assuming that a traditional driver segregates at its equilibrium frequency (Model 1), we investigate the evolution of tightly linked (Models 4 and 5), and unlinked (Model 6) sperm-dependent modifiers of drive.
-In Models 4 and 5, we treat this two-locus system as if it consists of
-a single locus with three alleles: $A$, $B$ and $C$, corresponding to the case when the
-sperm-modifier arises once, tightly linked to one or other of the
-alleles at the drive locus (such that recombination is unexpected). 
-In all cases, we assume that fitness is independent of the drive modifier. 
-We assume the traditional driver is transmitted to $d_0$ of gametes from female heterozygotes  when fertilized by wild-type sperm, and $d_1=d_0+\epsilon$ when fertilized by a sperm-acting drive-modifier.
 

### 2014-07-14 writeup_and_notes/GeneticsResub/reviews.tex
first go at revisions by YB next do some math and sims for female vs male suppression of female drive

\documentclass[12pt,letterpaper]{article}
\usepackage{epsfig}                 
\usepackage[authoryear]{natbib}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{psfrag}
\usepackage{mathabx}
\renewcommand{\baselinestretch}{1.6}
\large
\pagenumbering{arabic}
\usepackage[usenames,dvipsnames]{color}
\usepackage{fullpage}
\bibliographystyle{genetics}
\usepackage{multirow}
\newcommand{\ye}{\hat{y}}
\newcommand{\xe}{\hat{x}}
\usepackage{color}
\usepackage[normalem]{ulem}  
	\newcommand{\gc}[1]{{ \color{red} #1}}
	\newcommand{\yb}[1]{{ \color{blue} #1}}

\begin{document}
MS ID\#: GENETICS/2014/167064

Dear Yaniv:

Please visit http://submit.genetics.org to view the comments of the reviewers of your submission, "Sperm should evolve to make female meiosis fair." \\

Both reviewers found the manuscript inventive and well-written.  However, both have raised substantive issues that indicate that the interesting ideas presented need more development before consideration for publication.

The paper as written establishes that factors expressed in sperm should generally evolve to suppress drivers of meiotic drive in female meiosis.  However, as Reviewer 1 points out, this is different from showing that factors expressed in females should evolve to delay completion of meiosis until after fertilization.  There is general agreement that the argument needs to be completed so that it supports all conclusions and that the justification for study of the question be improved.  Reviewer 2 has provided many valuable suggestions along these lines.

The format of this submission departs from the style of GENETICS (model description and analysis, literature citations, etc.).  As a result, it is unclear whether GENETICS is the appropriate or even the intended forum for this work.  I am recommending rejection of the present submission.  A manuscript that addressed all concerns raised would essentially constitute a new work, and could be reviewed as a new submission. \\
Best wishes,
Marcy
\\
I regret to have to bring this unwelcome news.  I hope you will continue to submit your best work for publication in the journal, and that I can bring more favorable news when you do.
\\ \\
Sincerely,\\
Marcy Uyenoyama\\
Associate Editor\\
Senior Editor
\newpage

\section*{Reviewer 1 -- Reviewer's Comments to Author \dots}

This manuscript uses simple population genetic models to evaluate the potential
for sperm haplotype (or paternal genotype) to influence female meiotic drive. 
The models are quite straightforward and the results are very clearly explained.
 In that sense, the manuscript was a pleasure to read.  This manuscript should
be published somewhere but I am not sure it is enough of a sufficient advance to
warrant publication in Genetics.
\begin{enumerate}
\item
A major part of the motivation for the model seems to be to explain why, in
animals, meiosis is typically suspended until fertilization occurs (providing
the opportunity for males to influence female drive).  This seems a poor
motivation for the project because it would require a bizarre form of logic by
which the evolution of this delay in meiosis evolved with the ``foresight'' that
there would then be subsequent evolution of male influence on female drive. 
More directly, if the authors' interest is in the evolution of arrest in female
meiosis, then it would be better to do a model that explicitly addressed that
issue.

Obviously, the authors can avoid this problem by simply saying, given that
meiotic arrest occurs and provides the opportunity for males to influence drive,
we want to see what will happen.  However, it is misleading to present this
model as though it addresses the issue of the evolution of meiotic arrest.

\yb{We thank the reviewer for this comment, and have reworked our manuscript to focus 
	around our major result -- that sperm are expected to evolve to prevent female meiotic drive.
	We view this as an important point, and as the significant change in this new manuscript. 
	As suggested by R1, we focus our discussion of the fertilization requirement to pointing out that it provides sperm an opportunity to influence female meiosis, 
	and we only briefly point out that given the potential reasons for this delay, the threat of sperm collaboration in meiotic drive does not represent a selective force favoring the completion of meiosis before fertilization.}

\item
I did not feel that the authors made a convincing case that this is an
important phenomenon in its own right.  P. 5 has one short paragraph pointing to
what seems like some rather indirect evidence of male influence on drive.
Consequently, I find myself wondering whether this manuscript addresses a topic
of sufficient importance for Genetics.  The authors do make an interesting
prediction regarding drive effects in hybrids but in the absence of any current
evidence it is hard to know whether this is just an interesting idea that never
really happens. 
\end{enumerate}

\newpage

\section*{Reviewer 2 -- Reviewer's Comments to Author \dots}

\subsection*{R2 Summary}
The focus of the manuscript entitled  ``Sperm should evolve to make female meiosis
fair'' by Yaniv Brandvain and Graham Coop is to provide a theoretical foundation
for two seemingly unrelated enigmatic observations: i) that it is common for
female meiosis to arrested at an early stage and then complete the process only
after the egg is fertilized, and ii) that there is evidence (i.e., a couple of
established cases) that the strength of female meiotic drive can be influenced
by the genotype of the sperm that fertilizes an egg.  These two seemingly
unrelated empirical patterns are functionally coupled by the authors by solving
for the conditions under which different forms of sperm-mediated modifiers of
female meiotic drive will be selectively favored.  The authors show that
``Greenbeard-effect'' modifiers that bias female meiotic drive in favor of the
allele found in the sperm evolve under a very restricted part of the total
parameter space (and rapidly fix under these conditions) while those that
prevent or reduce meiotic drive evolve under a much expanded parameter space
(and are more likely to remain polymorphic).  This asymmetry would be expected
to lead to the evolution of sperm that suppress female meiotic drive, but only
if most of the steps of female meiosis are delayed until after fertilization. So
arrested meiotic progression (until fertilization) may be an adaptation that
expands the ability of females to reduce the deleterious effects of meiotic
drive by creating the opportunity for sperm to also suppress the trait.

	 Overall I found the paper to be well written and organized, and because of
this, the theory that the authors describe will be accessible to a broad
spectrum of the readership of Genetics -- not just other theoreticians.  Although
I have not re-derived all of the equations used in the author's models, the
special cases that I have worked through, and the logic used by the authors to
motivate the models, suggest to me that they are sound. I would like to see the
authors consider (but not necessarily follow) a few suggestions that I describe
below.

\subsection*{R2 Primary suggestions}
\begin{enumerate}
\item	The importance of the wok hinges critically on the the degree to which:  {\bf{\emph{i)}}}
female meiotic drive operates in nature, and {\bf{\emph{ii)}}} sperm influence the level of
female meiotic drive.  In species like maize, where every chromosome has two or
more established meiotic drivers (``chromosomal knobs''), the importance of
meiotic drive is obvious, but in most species there is little or no evidence for
the current operation of meiotic drive -- although low levels of transmission
ratio distortion (TRD) are becoming increasingly well established from
genome-wide SNP studies.  However, the rapid divergence of telomeric sequences
among closely related species is certainly consistent with the widespread
operation of transient episodes of meiotic drive via centromere drive.  In the
present manuscript, the prevalence and importance of female meiotic drive is
quickly summarized by the terse statements ``Meiotic drivers observed in both
plants [5, 10, 9], and animals [1, 29, 21] highlight this conflict -- the selfish
benefits and the associated pleiotropic fitness costs of drive sustain a
balanced polymorphism [24], and often generate ongoing evolutionary escalations
of drive suppressors and enhancers [8, 9].''  I think the paper would be far more
convincing and appreciated if the authors put a tad more effort into convincing
the reader that female meiotic drive is sufficiently common to be a widespread
selective force.

\item	 To my knowledge, the evidence for sperm influencing the outcome of meiotic
drive is restricted to the In and Om drive systems in mice and aster transfer by
the sperm in C. elegans.  These are clearly proof of process examples but I
think that many readers will take issue with the assumed general ability of
sperm to substantially influence the outcome of female meiotic drive. This
skepticism is to be expected  because the egg typically completes meiosis
immediately after fertilization, at a time when the sperm pronucleus is
decondensing and presumably not yet transcriptionally active. The more evidence
that the authors can provide for sperm being capable of influencing female
meiotic drive, the better this paper will be received. The authors could bolster
their case by briefly reviewing the recent evidence for extensive paternal
effects via RNA and protein packaged in the sperm, delivered to the egg, and
influencing subsequent early embryonic development.

\item	Finally, female meiotic drive potentially can be suppressed by alleles
expressed in the female (traditional models) or by alleles originating in the
sperm and acting in the egg immediately after fertilization (the authors new
model).  It would be useful to compare the constraints for the evolution of both
types of suppressors to evaluate if sperm-mediated suppression of female meiosis
has some evolutionary advantage that make it easier to evolve or non-redundant
to female-coded suppressors.
\end{enumerate}

\subsection*{Minor Points:}
\begin{enumerate}
\item   On the top of page 7 the statement ``This analytical approximation, derived
from Equation (1) assuming Hardy-Weinberg'', -- precedes the introduction of
Equation (1) in the Appendix.  I suggest pointing out here that all equations
described in the text are located in the Appendix (so the reader doesn't stop
reading and start looking for them, thinking they somehow overlooked them in the
body of the manuscript).
\yb{We can relate to the confusion and frustration of struggling to find the relevant equation. We have therefore added the parenthetical comment --   ``\emph{We note that Equation 1 and all equations discussed in the main text are presented in the appendix and derived in File S2}'' -- to the end of this paragraph.}

\item There are 6 models described in the text but these are not described in any
detail until the Appendix.  The authors might consider placing a table in the
main body of the text that tersely describes the 6 models.?
\yb{We have added a Methods section, which introduces these six models, and makes our manuscript more in line with the GENETICS format.}

\item On the bottom of page 14 the authors write ``we predict that a lower
opportunity for female meiotic drive, e.g. an animal lineage with a long history
of high inbreeding or selfing, would often be accompanied by a relaxation of the
requirement of fertilization for the completion of female meiosis.''  As I
recall, species with a history if inbreeding or selfing are only at the tips of
most phylogenetic trees (i.e., they do not persist over geological time and do
not have a ``LONG history of high inbreeding or selfing'') so this prediction may
be limited.

\item On the bottom of page 14, do you also predict that observed drive in species
hybrids will depend on the direction (which male) is used in the backcross?

\item I think that equations like $(s < d -1/2 + 3sh/2 - dsh)$ need parentheses to be
clear. 
\yb{To add clarity, we have surrounded longer in-text equations like the example above, with parentheses.}

\end{enumerate}
\end{document}

### 2014-07-14 writeup_and_notes/male_dependent_drive.tex
first go at revisions by YB next do some math and sims for female vs male suppression of female drive

@@
 52-54 in][]{Burt2006}, it appears that the extent to which \emph{In} and \emph{Om} distort the second female meiotic division partially depends on the genotype of the fertilizing sperm \citep{Agulnik1993,Wu2005}.
-% Both of these female drive systems operate by distorting the second meiotic division, 
-% and in both systems the outcome of female meiosis depends the genotype of the fertilizing sperm \citep{Agulnik1993,Wu2005}.  
-%While such genetic evidence is suggestive, we note the difficulty in
-%ruling out the alternative hypothesis of early selection on zygotes \citep[pages
-%52-54][]{Burt2006}.
 
 We develop population genetic models exploring the evolution of sperm influence on female meiotic drive. 
-%, focussing on models in which `self-promoting' alleles in sperm facilitate drive of like genotypes during gametogenesis in heterozygous females.
 We first focus on models in which `self-promoting' alleles in
   sperm facilitate drive of like alleles during gametogenesis in heterozygous females. 
-%In this article we explore through simple population genetic models the consequences of alleles that influence the outcome of female meiosis.
 These models show that such sperm-acting alleles 
@@
 
+{\bf{File S1:}} Transmission rules. We detail the frequency of offspring genotypes produced by each possible mating for Models 1-6 in the tabs of this Excel spreadsheet.
+
+{\bf{Files S2A \& S2B:}} A Mathematica file (FileS2A) and a PDF of this file (File S2B) in which we derive analytical results for models 1-6.
+
+{\bf{File S3:}} Exact approach. The R Script used for exact recursions.
+\clearpage
 \begin{figure}

### 2014-07-16 writeup_and_notes/GeneticsResub/male_dependent_drive.tex
graham's changes to intro

@@
 	likely precludes this interaction in plants). 
+\gc{It is becoming
+        increasingly clear that sperm bring a wide variety of RNA and proteins into the
+        egg REF. Some of these have known functions, for example, in most
+        animal species sperm-- not eggs-- are
+        responsible for the transmission of the centrosome, a vitial
+        component of the mitotic machinary for the zygote. However,
+        the function of many of th products the sperm beings into the
+        egg is completely unknown and 
+      these products vary widely over species. 
+The existance of paternal effect mutations in model systems highlights
+ the fact that some of these components have a wide-range of functions in
+ the completion of synagamy and the development of the zygote. 
+Thus sperm could potentially have functional effects on
+various aspects of female meiosis, e.g. in some systems 
+sperm actually carry the factor responsibly for the resumption of
+female meiosis, rather than this being a simple consequence of egg
+activation. } 
 Therefore, a `green-bearded' \citep{Gardner2010} sperm-acting allele that recognizes and facilitates the meiotic drive of a genetically equivalent allele in heterozygous females %female meiosis %(a  heteromorphic dyad),
      could presumably rapidly spread through a population. 
-At first sight, female meiosis appears primed for exploitation by such selfish systems, \yb{and here we ask if spam do indeed evolve to collaborate with female drivers to exploit this apparent weakness in the defense against meiotic drive.} \newline
+At first sight, female meiosis appears primed for exploitation by such selfish systems.
+Here we ask if sperm do indeed evolve to collaborate with female drivers to exploit this apparent weakness in the defense against meiotic drive.
+\newline
 %Why then is the requirement of fertilization to complete female meiosis so ubiquitous?
 
-\yb{The threat to meiotic fairness posed by the fertilization requirement for completion of female meiosis seems both real and avoidable. 
-We say this threat is avoidable because the stage at which female meiosis requires fertilization 
-	varies widely across taxa, and in a few animals female meiosis is complete upon ovulation \
-	(that is, fertilization is not required for female meiosis \citep[e.g. Table 1 in ][]{Masui_book}). 
-We say this threat is real because there is substantial evidence that female meiotic drive is an 
-	evolutionary force and considerable opportunity for sperm to influence female drive.  
-While research to date has identified a few of what are likely the most extreme female meiotic 
-	drivers \citep{Agulnik1990, Fishman2008, Hiatt2003,Novitski1951, Pardo-ManuelDeVillena2001c}, 
-	rapid evolution of centromeres, telomeres and other basic components of the meiotic machinery, 
-	suggest consistent evolution of meiotic drivers and suppressors of meiotic drive
-	 \citep[e.g. ][]{Anderson2008, Anderson2009, Axelsson2010, Malik2009a}, and the inexpense of 
-	 genotyping-by-sequencing platforms for non-model systems will likely 
-	identify subtle meiotic drivers in numerous non-model systems.  
+It is not the case that sperm is mechanistically incapable of influencing female meiosis.
 Both mechanistic and genetic evidence suggests that sperm can influence the outcome of female meiosis.
@@
 	disrupts MII meiotic segregation in the egg, leading to a triploid zygote \citep{McNally2012}.
-{\bf{GRAHAM DO MORE????} } 
-Genetic evidence from the \emph{In} and \emph{Om}  drive systems in mice suggests that the 
+There are current only two polymorphic,characterized females meiotic drive systems in
+animals-- and both of them, the \emph{In} and \emph{Om}  drive systems in mice-- suggests that the 
 	 transmission of alternate alleles in heterozygous females can depend on sperm haplotype.  
  While ruling out the alternative hypothesis of early selection on zygotes in these cases is challenging  \citep[see pages
-	52-54 in][]{Burt2006}, it appears that the extent to which \emph{In} and \emph{Om} distort the second female meiotic division partially depends on the 	
-	genotype of the fertilizing sperm \citep{Agulnik1993,Wu2005}. 
-}\newline
+	52-54 in][ for comment]{Burt2006}, it appears that the extent to which \emph{In} and \emph{Om} distort the second female meiotic division partially depends on the 	
+	genotype of the fertilizing sperm
+        \citep{Agulnik1993,Wu2005}. 
+The fact that both of polymorphic, putative female meiotic drive systems identified
+to date in animals (\emph{In} and \emph{Om} in mice) seem to show an
+effect of sperm genotype suggests that if female meiotic drive is
+common the role of sperm will be important as well.
+
+There is good evidence that female meiotic drive is an 
+	evolutionary force and offering considerable opportunities for sperm to influence female drive.  
+While research to date has identified a few extreme female meiotic 
+	drivers in the small number model systems systematically
+        studied to date \citep{Agulnik1990, Fishman2008, Hiatt2003,Novitski1951, Pardo-ManuelDeVillena2001c}, 
+	rapid evolution of centromeres, telomeres and other basic components of the meiotic machinery, 
+	suggest consistent evolution of female meiotic drivers and suppressors of meiotic drive
+	 \citep[e.g. ][]{Anderson2008, Anderson2009, Axelsson2010,
+           Malik2009a}. We expect that over the next decade the spread of sequencing to a
+         range of systems will reveal many more female meiotic drive
+         systems, however, carefully characterizing them will still
+         remain a challenging task.
+\newline
 
@@
 Such models universally reveal that sperm-acting drive modifiers are favored only if they suppress drive. 
-These results highlight the fact that the interests of sperm and maternal  genomes' are often aligned, as both are invested in the fate of the resultant zygote (as was speculated for the \emph{In} locus \citep{Pomiankowski1993}).
+These results highlight the fact that the interests of sperm and
+maternal  genomes' are often aligned, 
+as both are invested in the fate of the resultant zygote (as was speculated for the \emph{In} locus \citep{Pomiankowski1993}).
 Thus, there is little selective benefit to females in preventing sperm to influence female meioses. 
-\yb{While we do not claim that female meiotic arrest evolved specifically to recruit helpful sperm to defend against meiotic drive, the delayed completion of female meiotic drive will not in fact encourage sperm to assist renegade drivers.}
-%Rather, features of o\"{o}genesis  may evolve to facilitate an interaction with sperm %with female meiosis
-%	as yet another mechanism to enforce an equitable transfer of genetic material.
+\gc{We do not claim that female meiotic arrest evolved
+  specifically to recruit helpful sperm to defend against meiotic
+  drive, the delayed completion of female meiotic drive will not in
+  fact encourage sperm to assist renegade drivers. Rather, we think
+  that features of
+  o\"{o}genesis  may evolve to facilitate an interaction with sperm 
+with female meiosis
+as yet another mechanism to enforce an equitable transfer of genetic material.}
 

### 2014-07-16 writeup_and_notes/GeneticsResub/male_dependent_drive.tex
graham's changes to intro

@@
         component of the mitotic machinary for the zygote. However,
-        the function of many of th products the sperm beings into the
+        the function of many of the products the sperm brings into the
         egg is completely unknown and 
@@
 %Why then is the requirement of fertilization to complete female meiosis so ubiquitous? 
+\gc{THERE may be too much bio. detail in previous paragraph}
 

### 2014-07-17 writeup_and_notes/GeneticsResub/male_dependent_drive.tex
small edits to intro and methods

@@
         responsible for the transmission of the centrosome, a vitial
-        component of the mitotic machinary for the zygote. However,
+        component of the mitotic machinary for the zygote. 
+Detailed functional studies and paternal effect mutations in model
+systems further highlight
+ that sperm-transmitted products have a wide-range of functions in
+ the completion of syngamy, zygotic development, and the resumption of
+ female meiosis \cite{Miller:01}. 
+However,
         the function of many of the products the sperm brings into the
@@
       these products vary widely over species. 
-The existance of paternal effect mutations in model systems highlights
- the fact that some of these components have a wide-range of functions in
- the completion of synagamy and the development of the zygote. 
-Thus sperm could potentially have functional effects on
-various aspects of female meiosis, e.g. in some systems 
-sperm actually carry the factor responsibly for the resumption of
-female meiosis, rather than this being a simple consequence of egg
-activation. } 
+It seems quite plausible that sperm based products and
+henece alleles could potentially have functional effects on
+various aspects of female meiosis that occur after fertilization. } 
 Therefore, a `green-bearded' \citep{Gardner2010} sperm-acting allele that recognizes and facilitates the meiotic drive of a genetically equivalent allele in heterozygous females %female meiosis %(a  heteromorphic dyad),
      could presumably rapidly spread through a population. 
-At first sight, female meiosis appears primed for exploitation by such selfish systems.
-Here we ask if sperm do indeed evolve to collaborate with female drivers to exploit this apparent weakness in the defense against meiotic drive.
-\newline
+At first sight, female meiosis appears primed for \gc{conflict caused by
+the such selfish systems.} Here we ask if sperm do indeed evolve to collaborate with female drivers to exploit this apparent weakness in the defense against meiotic drive. \newline 
 %Why then is the requirement of fertilization to complete female meiosis so ubiquitous? 
@@
 We verbally describe our main results below.  
-Readers interested in more quantitative results should turn to the Appendix for a mathematical treatment. 
-Here we present critical analytical results in Equations 1-10, and describe our analyses and results in more detail.  
-Because many of our analyses assume genotype frequencies expected under Hardy Weinberg Equilibrium (HWE), we note which analytical results are approximations. 
+Readers interested in the details of these results should turn to the Appendix for a mathematical treatment. 
+There we present critical analytical results in Equations 1-10, and describe our analyses and results in more detail.  
+Because a number of our analyses are approximations based on assuming
+that genotype frequencies follow Hardy Weinberg Equilibrium (HWE), we note which analytical results are approximations. 
 We verify these approximations with exact numerical iterations in Figures S1-S3 and File S3 , and detail the derivation of these approximations in our File S2.

### 2014-07-18 writeup_and_notes/GeneticsResub/male_dependent_drive.tex
added model 6prime to text

@@
 {\bf{Model 3}} describes a single-locus paternal-dependent female driver -- an allele whose transmission in female meiosis depends on paternal genotype. 
-Assuming that a traditional driver segregates at its equilibrium frequency ({\bf{Model 1}}), we investigate the evolution of tightly linked ({\bf{Models 4 and 5}}), and unlinked ({\bf{Model 6}}) sperm-dependent modifiers of drive.
+Assuming that a traditional driver segregates at its equilibrium frequency ({\bf{Model 1}}), we investigate the evolution of tightly linked ({\bf{Models 4 and 5}}), and unlinked ({\bf{Model 6}}) sperm-dependent modifiers of drive (We also briefly compare Model 6, to Model 6' - the suppression of female drive by an unlinked female acting suppressor).
 In {\bf{Models 4 and 5}}, we treat this two-locus system as if it consists of
@@
 Readers interested in the details of these results should turn to the Appendix for a mathematical treatment. 
-There we present critical analytical results in Equations 1-10, and describe our analyses and results in more detail.  
+There we present critical analytical results in Equations 1-11, and describe our analyses and results in more detail.  
 Because a number of our analyses are approximations based on assuming
@@
         \ref{unlinked}).  %%DELETED FIG REF HERE. 
-This allele lowers the frequency of the original driver (perhaps to zero),
+        \newline
+
+The sperm acting drive suppressing allele lowers the frequency of the original driver (perhaps to zero),
 	and spreads to fixation if it does not carry strong fitness costs
@@
 This result is consistent with previous work showing that drive suppressors unlinked to, or in repulsion phase with drivers usually invade polymorphic drive systems \citep[e.g. ][]{Brandvain2012}.  
+\yb{We note that the evolution of sperm-acting drive  suppressors unlinked to a driver (Model 6) is both qualitatively and quantitatively similar to the evolution  of a female-acting drive suppressor (compare Equations \ref{unlinked} and \ref{femunlinked}).}
 Therefore, all two-locus models of sperm influence on female drive suggest that
-	sperm will evolve to oppose female meiotic drive. \newline
+	sperm will evolve to oppose female meiotic drive. 
+\newline
 
@@
 $\epsilon <0$, corresponding to the $A^-$ model, above), a result which holds qualitatively for a common $C$ allele, as well (File S2  ).
-Somewhat counterintuitively, the spread of the $A^-$ allele is substantially slower than that of the $B^-$ allele (it is on the order of $C^{2}$ rather than $C$). 
-This is because the $A^-$ allele will never occur in drive-homozygotes and therefore only spreads by preventing itself from the cheating of the driver.
+%Somewhat counterintuitively, the spread of the $A^-$ allele is somewhat slower than that of the $B^-$ allele (it is on the order of $C^{2}$ rather than $C$). 
+%This is because the $A^-$ allele will never occur in drive-homozygotes and therefore only spreads by preventing itself from the cheating of the driver.
+% YB COMMENT {I removed bc this was actually quite a minor influence}
 
@@
 Thus, a rare drive suppressor ($\epsilon<0$) will spread so long as the fitness cost of the driver does not display over- or under-dominance.
-Notably, this result holds while assuming complete linkage equilibrium, and therefore sperm suppression of female meiotic drive is likely more powerful than more direct female-enforced drive suppressors  which usually spread due to the indirect benefit of avoiding drive haplotypes \citep[e.g.
-][]{Brandvain2012}.
+
+\subsubsection*{Model $6^\prime$. Unlinked female acting drive-modifier:}
+\yb{The dynamics of a female-acting drive modifier are comparable to those describing a female acting drive suppressor.  
+When rare and fully dominant and in linkage and in HWE, the change in frequency of a rare unlinked, female-acting drive modifier is 
+\begin{equation} \Delta M =-(d_h-d_0) f_A f_B (f_B s + s_h (f_A - f_B) ) \frac{f_M}{\bar{w}}  \label{femunlinked} \end{equation}
+When  when drive-modification is dominant ($d_h = d_1 = d_0 + \epsilon$), Equation \ref{femunlinked} is equal to Equation \ref{unlinked}.
+ However, if female drive suppression is less than fully dominant, sperm-acting drive suppressors are more efficacious when rare than are female-acting suppressors, and are therefore more likely to spread.
+}
 

### 2014-07-18 writeup_and_notes/GeneticsResub/male_dependent_drive.tex
Changes to discussion

--- 2014-07-17 writeup_and_notes/GeneticsResub/male_dependent_drive.tex
@@
  the completion of syngamy, zygotic development, and the resumption of
- female meiosis \cite{Miller:01}. 
+ female meiosis \cite{Miller:01, McNally:05,REF}. 
 However,
@@
 	that female o\"{o}genesis and meiosis will evolve to prevent their effect.  
-This likely explains why females can delay the completion of meiosis until after fertilization 
+\gc{Thus females can} delay the completion of meiosis until after fertilization 
 	without risking exploitation by collaborations between female
         drivers and sperm alleles. \newline 
+%%G: felt like this made it sound less why they do it.
 
-Why is it that an allele that biases female meiosis in its favor can generate a genetic conflict, but an allele in sperm that assists this female driver cannot? 
-So long as the transmission advantage of female meiotic drive outweigh the organismal fitness cost to heterozygotes, the female driver can spread when rare, and it increases in 		
+Why is it that an allele that biases female meiosis in its favor can
+generate a genetic conflict, 
+but an allele in sperm that assists this female driver cannot? 
+So long as the transmission advantage of female meiotic drive outweigh
+the organismal fitness cost to heterozygotes, 
+the female driver can spread when rare, and it increases in 		
 	frequency until the fitness cost to homozygotes balances the transmission advantage.
@@
 
-\yb{The theory presented herein provides a strong prediction -- in animals and other organisms in which female meiosis is not completed until after fertilization sperm will act to suppress female drive at the stage at which they can influence meiosis.
+\yb{The theory presented herein provides a strong prediction -- 
+organisms in which female 
+meiosis is not completed until after fertilization sperm will act to suppress female drive at the stage at which they can influence meiosis. 
 %Therefore, for example, when fertilized by drying sperm, we expect meiosis to be fairer when  females are fertilized by 
-contrary to expectations of a ``green-beard'' model, we are unsurprised by the observation that female meiosis is fairer when fertilized by sperm bearing the drive allele in two of the best described cases of female meiotic drive in animals (the \emph{Om} and \emph{In} loci in mice \citep{Agulnik1993,Wu2005}).
-Given the immense benefit to sperm of hampering female drive, drive-suppressing sperm are often likely to be fixed within a species, making the hypothesis of sperm acting drive-suppression difficult to test from intra-population crosses.
+Contrary to expectations of a \gc{naive verbal} ``green-beard'' model, 
+we are unsurprised by the observation that female meiosis is fairer when fertilized by sperm bearing the drive allele in two of the best described cases of female meiotic drive in animals (the \emph{Om} and \emph{In} loci in mice \citep{Agulnik1993,Wu2005}).
+Given the immense benefit to sperm of hampering female drive, 
+drive-suppressing sperm are often likely to be fixed within a species, 
+making the hypothesis of sperm acting drive-suppression difficult to test from intra-population crosses. 
 However, crosses between populations or species are likely to provide critical tests of our theory -- specifically we predict that female meiosis will be less fair when a species (or population) is fertilized by heterospecific sperm because such sperm have not evolved to counter novel females meiotic drivers.
 We further predict that segregation in F1 females backcrossed to parental species will likely be biased with a deficit of transmission of the paternal species allele from the F1  female.   
-We note that while these predictions follow straightforwardly from the theory presented above, tests of meiotic drive, and especially sperm dependent meiotic drive require a high standard of evidence to exclude plausible alternative hypotheses such as genotypic inviability including epistatic maternal X zygotic lethality.}  \newline
+We note that while these predictions follow straightforwardly from the
+theory presented above, tests of meiotic drive, 
+and especially sperm dependent meiotic drive require a high standard
+of evidence 
+to exclude plausible alternative hypotheses such as genotypic inviability including epistatic maternal X zygotic lethality.}  \newline
+
+\gc{Our theory also encourages phylogenetic hypotheses concerning the relationship between the opportunity 
+	for female meiotic drive and the requirement of fertilization for the completion of female meiosis. 
+For example, we predict that a lower opportunity for female meiotic drive,
+        e.g. an animal lineage with a history of high inbreeding
+        or selfing, 
+        may be accompanied by a relaxation of the requirement of fertilization for the
+        completion of female meiosis (although this effect may be
+        weaks as such
+        lineages may only persist for a short time). Another
+        hypothesis that our work suggests is we might
+        expect that phylogenetic variation in the stage of female meiosis when
+        fertilization occurs to play a role in the prevalance of
+        female meiotic drive.
+For example, in taxa where meiosis I is completed
+by the egg pre-fertilization, including humans, centromere drive may
+be more common as sperm based modifiers can not intercede than in
+species where fertilization occurs during MI. 
+Such hypotheses could be tested if the speed of centromere
+turnover serves as a proxy for the efficacy of MI drive. }
 
@@
 	fertilizing sperm decidedly side with mom, as both have a shared interest in producing a viable and potentially fit offspring. 
-\yb{Our model does not directly speak to the evolutionary origin of female meiotic arrest (for a review and evaluation of such hypotheses see \citep{Mira1998}); however, our work does imply that the fertilization requirement does not leave female meiosis open to sabotage by sperm collaborators, but rather that  an evolutionary loss of female meiotic arrest will result in the  loss of sperm defense against female meiotic drivers.
-}
+\yb{Our model does not directly speak to the evolutionary origin of
+  female meiotic arrest \citep[(for a review and evaluation of such
+  hypotheses see][]{Mira1998}.
+However, given that arrest does occur its timing relative to
+fertilisation, various features of the interaction between sperm and
+egg and divergence of these features across species  \citep{Karrbook:09} could be subject
+to selection to conflict between female meiotic drivers and alleles
+elsewhere in the genomes of sperm and eggs.
+the mechanism by which it occur
+mechanistic features, and evolutionary changes among taxa
+
+various features of the
+mechanism of fertilization 
+however, our work does imply that the fertilization requirement does
+not 
+leave female meiosis open to sabotage by sperm collaborators, but
+rather that  an 
+evolutionary loss of female meiotic arrest will result in the loss of sperm defense against female meiotic drivers. }
 

### 2014-07-18 writeup_and_notes/GeneticsResub/male_dependent_drive.tex
Changes to intro

@@
 	interpreted as mechanisms of protection against genomic outlaws. 
-\yb{The fact that female meiosis often left uncompleted until after
+The fact that female meiosis often left uncompleted until after
       fertilization in many animals appears to run counter to this interpretation, 
       as this delay provides an opportunity for sperm alleles to
-     meddle with its outcome and help like-alleles drive in heterozygous females. } 
+     meddle with its outcome and help like-alleles drive in heterozygous females. 
 %Why then is female meiosis often left uncompleted until after
@@
 The population genetic theory presented herein suggests that
-	sperm nearly always evolve to increase the fairness of female
+	\gc{in fact} sperm nearly always evolve to increase the fairness of female
         meiosis in the face of genomic conflicts. 
-These results are consistent with current knowledge of sperm-dependent meiotic drivers 
-(loci whose distortion of female meiosis depends on sperm genotype). 
-\yb{Rather than providing an opportunity for sperm collaboration in female meiotic drive,
+These results are consistent with current knowledge of sperm-dependence
+at the \gc{for the well characterized female meiotic drivers in animals to date.} 
+Rather than providing an opportunity for sperm collaboration in female meiotic drive,
         the requirement of fertilization for the completion of female meiosis
-        likely indirectly provides protection to females against meiotic drivers as it provides an opportunity for sperm to suppress female meiotic drive.}
+        likely indirectly provides protection to females against
+        meiotic drivers as it provides an opportunity for sperm to
+        suppress female meiotic drive. }
 %	and suggest that the requirement of fertilization for the
@@
         animal species sperm-- not eggs-- are
-        responsible for the transmission of the centrosome, a vitial
+        responsible for the transmission of the centriole, a vitial
         component of the mitotic machinary for the zygote. 
@@
         egg is completely unknown and 
-      these products vary widely over species. 
-It seems quite plausible that sperm based products and
-henece alleles could potentially have functional effects on
+      these products vary widely over species \citep{Karrbook:09}. 
+It seems quite plausible that sperm-based products and
+hence alleles could potentially have functional effects on
 various aspects of female meiosis that occur after fertilization. } 
@@
 %Why then is the requirement of fertilization to complete female meiosis so ubiquitous? 
-\gc{THERE may be too much bio. detail in previous paragraph}
+\gc{THERE may be too much bio. detail in previous paragraph, I can
+  add more refs. once we've decided on content}
 
-It is not the case that sperm is mechanistically incapable of influencing female meiosis.
+It is certainly not the case that sperm is mechanistically incapable of influencing female meiosis.
 Both mechanistic and genetic evidence suggests that sperm can influence the outcome of female meiosis.

### 2014-07-21 writeup_and_notes/GeneticsResub/male_dependent_drive.tex
Changes to bib
yb modified abstract and intro

@@
 O\"{o}genesis is inherently susceptible to such conflicts
-because alleles compete to be the  product of female meiosis transmitted to the egg. 
+because alleles compete \yb{for inclusion in} the  product of female meiosis transmitted to the egg. 
 Alleles that distort meiosis in their favor (i.e. meiotic drivers) often decrease organismal fitness,
@@
 	interpreted as mechanisms of protection against genomic outlaws. 
-The fact that female meiosis often left uncompleted until after
+The fact that female meiosis \yb{is} often left uncompleted until after
       fertilization in many animals appears to run counter to this interpretation, 
-      as this delay provides an opportunity for sperm alleles to
+      as this delay provides an opportunity for sperm\yb{-acting} alleles to
      meddle with its outcome and help like-alleles drive in heterozygous females. 
@@
         meiosis in the face of genomic conflicts. 
-These results are consistent with current knowledge of sperm-dependence
-at the \gc{for the well characterized female meiotic drivers in animals to date.} 
-Rather than providing an opportunity for sperm collaboration in female meiotic drive,
-        the requirement of fertilization for the completion of female meiosis
-        likely indirectly provides protection to females against
-        meiotic drivers as it provides an opportunity for sperm to
+These results are consistent with \yb{the apparent sperm- dependence}  \gc{of the best characterized female meiotic drivers in animals to date.} 
+ \yb{Rather than providing an opportunity for sperm collaboration in female meiotic drive,
+        the `fertilization requirement'  indirectly protects females from
+        meiotic drivers by providing sperm an opportunity to
         suppress female meiotic drive. }
@@
 An allele that biases female meiosis in its favor (i.e.
-a meiotic driver), may increase in frequency even if it entails a pleiotropic fitness cost \citep{Prout1973}, generating a genetic conflict between the success of the driver and organismal fitness.
+a meiotic driver), may increase in frequency even if it entails a pleiotropic fitness cost \citep{Prout1973}, generating a genetic conflict between the success of the driver and the organism.
 Meiotic drivers observed in both plants
@@
 	likely precludes this interaction in plants). 
+Therefore, a `green-bearded' \citep{Gardner2010} sperm-acting allele that recognizes and facilitates the meiotic drive of a genetically equivalent allele in heterozygous females %female meiosis %(a  heteromorphic dyad),
+     could presumably rapidly spread through a population. 
+At first sight, female meiosis appears primed for \gc{conflict caused by
+the such selfish systems.} Here we ask if sperm do indeed evolve to collaborate with female drivers to exploit this apparent weakness in the defense against meiotic drive. \newline
+
+\yb{Before doing so, it is worth considering whether there is any evidence that sperm could (or do), influence female meiosis. }
 \gc{It is becoming
         increasingly clear that sperm bring a wide variety of RNA and proteins into the
-        egg REF. Some of these have known functions, for example, in most
+        egg REF. 
+ Some of these have known functions, for example, in most
         animal species sperm-- not eggs-- are
-        responsible for the transmission of the centriole, a vitial
-        component of the mitotic machinary for the zygote. 
+        responsible for the transmission of the centriole, a vital
+        component of the mitotic machinery for the zygote. 
 Detailed functional studies and paternal effect mutations in model
@@
  female meiosis \cite{Miller:01, McNally:05,REF}. 
+ For example, in \emph{C. elegans}, premature deployment of the aster 
+ 	-- a vital component of mitotic machinery provided by the sperm, 
+	disrupts MII meiotic segregation in the egg, leading to a triploid zygote \citep{McNally2012}.
 However,
@@
 various aspects of female meiosis that occur after fertilization. } 
-Therefore, a `green-bearded' \citep{Gardner2010} sperm-acting allele that recognizes and facilitates the meiotic drive of a genetically equivalent allele in heterozygous females %female meiosis %(a  heteromorphic dyad),
-     could presumably rapidly spread through a population. 
-At first sight, female meiosis appears primed for \gc{conflict caused by
-the such selfish systems.} Here we ask if sperm do indeed evolve to collaborate with female drivers to exploit this apparent weakness in the defense against meiotic drive. \newline 
 %Why then is the requirement of fertilization to complete female meiosis so ubiquitous? 
 \gc{THERE may be too much bio. detail in previous paragraph, I can
-  add more refs. once we've decided on content}
+  add more refs. once we've decided on content
+centriole : http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3144413/?report=classic}
+\newline
 
-It is certainly not the case that sperm is mechanistically incapable of influencing female meiosis.
-Both mechanistic and genetic evidence suggests that sperm can influence the outcome of female meiosis.
-An example of mechanistic evidence for sperm's influence on female meiosis comes from \emph{C.
+%It is certainly not the case that sperm is mechanistically incapable of influencing female meiosis.
+%Both mechanistic and genetic evidence suggests that sperm can influence the outcome of female meiosis.
+%An example of mechanistic evidence for sperm's influence on female meiosis comes from \emph{C.
 elegans}, where premature deployment of the aster -- a vital component of mitotic machinery provided by the sperm,
-	disrupts MII meiotic segregation in the egg, leading to a triploid zygote \citep{McNally2012}.
-There are current only two polymorphic,characterized females meiotic drive systems in
-animals-- and both of them, the \emph{In} and \emph{Om}  drive systems in mice-- suggests that the 
+%	disrupts MII meiotic segregation in the egg, leading to a triploid zygote \citep{McNally2012}.
+\yb{Current evidence suggests that sperm influence on female meiotic drive is not only possible, but likely. }
+Evidence from both the \emph{In} and \emph{Om}  -- suggests that the 
 	 transmission of alternate alleles in heterozygous females can depend on sperm haplotype.  
- While ruling out the alternative hypothesis of early selection on zygotes in these cases is challenging  \citep[see pages
-	52-54 in][ for comment]{Burt2006}, it appears that the extent to which \emph{In} and \emph{Om} distort the second female meiotic division partially depends on the 	
-	genotype of the fertilizing sperm
-        \citep{Agulnik1993,Wu2005}. 
-The fact that both of polymorphic, putative female meiotic drive systems identified
-to date in animals (\emph{In} and \emph{Om} in mice) seem to show an
-effect of sperm genotype suggests that if female meiotic drive is
-common the role of sperm will be important as well.
+%There are current only two polymorphic, well-characterized female meiotic drive systems in
+%animals-- and both of them, the \emph{In} and \emph{Om}  drive systems in mice-- suggests that the 
+	 transmission of alternate alleles in heterozygous females can depend on sperm haplotype.  
+ While ruling out the alternative hypothesis of early selection on zygotes in these cases is challenging  \citep[see pages 52-54 in][ for comment]{Burt2006}, it appears that the extent to which \emph{In} and \emph{Om} distort the second female meiotic division partially depends on the
+	genotype of the fertilizing sperm \citep{Agulnik1993,Wu2005}. 
+\yb{The fact that the two best characterized, polymorphic systems of putative female 
+	meiotic drive systems identified to date in animals (\emph{In} and \emph{Om}) 
+	show this effect suggests that if female meiotic drive is
+	common the role of sperm in modifying female drive will be important as well. } \newline
 
-There is good evidence that female meiotic drive is an 
-	evolutionary force and offering considerable opportunities for sperm to influence female drive.  
-While research to date has identified a few extreme female meiotic 
-	drivers in the small number model systems systematically
-        studied to date \citep{Agulnik1990, Fishman2008, Hiatt2003,Novitski1951, Pardo-ManuelDeVillena2001c}, 
+\yb{Numerous lines of evidence suggest that female meiotic drive is an 
+	evolutionary force, and therefore the opportunity for sperm to influence female drive likely had applications to many animals}.   
+While, research to date has identified a few extreme cases of female meiotic 
+	drive in the small number model systems systematically
+        studied \citep{Agulnik1990, Fishman2008, Hiatt2003,Novitski1951, Pardo-ManuelDeVillena2001c},
 	rapid evolution of centromeres, telomeres and other basic components of the meiotic machinery, 
@@
 	 \citep[e.g. ][]{Anderson2008, Anderson2009, Axelsson2010,
-           Malik2009a}. We expect that over the next decade the spread of sequencing to a
+           Malik2009a}. 
+ We expect that over the next decade the spread of sequencing to a
          range of systems will reveal many more female meiotic drive
@@
 %][]{Masui_book}). 
-
 %It is also not the case that sperm is mechanistically incapable of influencing female meiosis.
@@
 
-We develop population genetic models exploring the evolution of sperm influence on female meiotic drive. 
+\yb{Because female meiotic drive is likely common and consequential, 
+	and because sperm have ample opportunity to influence female drive,  
+	we develop population genetic models exploring the evolution of sperm influence on female meiotic drive -- 
+	that is, what is the expected influence of sperm on female drive? } 
 We first focus on models in which `self-promoting' alleles in
@@
 Thus, there is little selective benefit to females in preventing sperm to influence female meioses. 
-\gc{We do not claim that female meiotic arrest evolved
-  specifically to recruit helpful sperm to defend against meiotic
-  drive, the delayed completion of female meiotic drive will not in
-  fact encourage sperm to assist renegade drivers. Rather, we think
-  that features of
-  o\"{o}genesis  may evolve to facilitate an interaction with sperm 
-with female meiosis
-as yet another mechanism to enforce an equitable transfer of genetic material.}
+ \yb{In fact, a mutant female which completes meiosis before 
+ 	fertilization is giving up the protection from meiotic drivers potentially offered by sperm, 
+	and is therefore more likely to produce low fitness offspring. }
+%\gc{We do not claim that female meiotic arrest evolved
+%  specifically to recruit helpful sperm to defend against meiotic
+%  drive, the delayed completion of female meiotic drive will not in
+%  fact encourage sperm to assist renegade drivers. Rather, we think
+%  that features of
+%  o\"{o}genesis  may evolve to facilitate an interaction with sperm 
+%with female meiosis
+%as yet another mechanism to enforce an equitable transfer of genetic material.}
 

### 2014-07-21 writeup_and_notes/GeneticsResub/male_dependent_drive.tex
yb modified discussion and fixed some refs

@@
         may be accompanied by a relaxation of the requirement of fertilization for the
-        completion of female meiosis (although this effect may be
-        weaks as such
-        lineages may only persist for a short time). Another
-        hypothesis that our work suggests is we might
-        expect that phylogenetic variation in the stage of female meiosis when
-        fertilization occurs to play a role in the prevalance of
+        completion of female meiosis (although 
+   \yb{     opportunities to test this hypothesis may be limited because }
+        %this effect may be weaks as such
+        lineages may only persist for a short time). 
+Our results also suggests that 
+%Another   hypothesis that our work suggests is we might expect 
+        that phylogenetic variation in the stage of female meiosis when
+        fertilization occurs to influence the prevalence of
         female meiotic drive.
@@
 turnover serves as a proxy for the efficacy of MI drive. }
+\newline
 
@@
 \yb{Our model does not directly speak to the evolutionary origin of
-  female meiotic arrest \citep[(for a review and evaluation of such
+  female meiotic arrest \citep[for a review and evaluation of such
   hypotheses see][]{Mira1998}.
-However, given that arrest does occur its timing relative to
-fertilisation, various features of the interaction between sperm and
-egg and divergence of these features across species  \citep{Karrbook:09} could be subject
-to selection to conflict between female meiotic drivers and alleles
-elsewhere in the genomes of sperm and eggs.
-the mechanism by which it occur
-mechanistic features, and evolutionary changes among taxa
-
-various features of the
-mechanism of fertilization 
-however, our work does imply that the fertilization requirement does
-not 
-leave female meiosis open to sabotage by sperm collaborators, but
-rather that  an 
-evolutionary loss of female meiotic arrest will result in the loss of sperm defense against female meiotic drivers. }
+However, given the existence of female meiotic arrest, 
+	and that its timing relative to
+	fertilization as well as the interactions between sperm and egg are variable across species  \citep{Karrbook:09} 
+	the nature of the meiotic arrest and interactions between sperm and egg may be molded by selection to reduce the opportunity for female meiotic drive, 
+	while selfish drivers may evolve in concert, to overcome these adaptations. 
+%leave female meiosis open to sabotage by sperm collaborators, but
+%rather that  an 
+%evolutionary loss of female meiotic arrest will result in the loss of sperm defense against female meiotic drivers. } 
+  }
+%However, given that arrest does occur its timing relative to
+%fertilization, various features of the interaction between sperm and
+%egg and divergence of these features across species  \citep{Karrbook:09} could be subject
+%to selection to conflict between female meiotic drivers and alleles
+%elsewhere in the genomes of sperm and eggs.
+%the mechanism by which it occur
+%mechanistic features, and evolutionary changes among taxa
+%various features of the
+%mechanism of fertilization 
+%however, our work does imply that the fertilization requirement does
+%not 
+%leave female meiosis open to sabotage by sperm collaborators, but
+%rather that  an 
+%evolutionary loss of female meiotic arrest will result in the loss of sperm defense against female meiotic drivers. }
 

### 2014-07-22 writeup_and_notes/GeneticsResub/reviews.tex
responded to reviewer 1, and staatted letter... now to r2

@@
 \pagenumbering{arabic}
+\usepackage{setspace}
+\singlespacing
+\onehalfspacing
+\doublespacing
 \usepackage[usenames,dvipsnames]{color}
@@
 \begin{document}
-MS ID\#: GENETICS/2014/167064
+Please find attached a new manuscript for your evaluation for publication in GENETICS, titled  `Sperm should evolve to make female meiosis fair.' 
+The positive and constructive reviews of our previous submission with the same title (\emph{MS ID\#: GENETICS/2014/167064})  where extremely useful in our crafting of this revision. 
+Below we discuss how this paper was improved by addressing, in detail, previous concerns \emph{point-by-point}, but first we address the major  concerns with our previous submission.
+\begin{itemize}
+\begin{singlespace}
+\item \emph{{\bf{Major concern 1 -- }} The paper as written establishes that factors expressed in sperm should generally evolve to suppress drivers of meiotic drive in female meiosis.
+However, as Reviewer 1 points out, this is different from showing that factors expressed in females should evolve to delay completion of meiosis until after fertilization.
+There is general agreement that the argument needs to be completed so that it supports all conclusions and that the justification for study of the question be improved.
+Reviewer 2 has provided many valuable suggestions along these lines.}  \newline \\
+{\bf{Response -- }} This revision makes it clear that our primary argument is that, given a chance to influence female meiosis, sperm will evolve to increase its fairness (as reflected in our title), {\bf{AND NOT}} that females should delay the completion of meiosis so as to recruit help from sperm in making meiosis fair (as could have been interpreted from our previous draft).
+\end{singlespace}
 
-Dear Yaniv:
+\begin{singlespace}
+\item \emph{{\bf{Major concern 2 -- }} The format of this submission departs from the style of GENETICS (model description and analysis, literature citations, etc.).
+As a result, it is unclear whether GENETICS is the appropriate or even the intended forum for this work.} \newline \\
+{\bf{Response -- }}  We believe our manuscript presents 
+(1) a rigorous and theoretical model, and 
+(2) research questions that can be addressed by molecular genetics, transmission genetics, studies of genome structure, phylogenetics and population genetics.
+As such, GENETICS is a particularly appropriate venue for our paper.
+Our view is echoed by comments from Reviewer 2 who stated  `Overall I found the paper to be well written and organized, and because of this, the theory that the authors describe will be accessible to a broad spectrum of the readership of Genetics ?
+not just other theoreticians.'
+The suitability of the topic of our manuscript is clear from our reference list -- a strong plurality (over 25\%) of our works cited appeared in GENETICS, placing our work squarely within the tradition of GENETICS papers.
+We have therefor corrected issues with our formatting, and included a METHODS section to introduce our models, as is the tradition in GENETICS. \\
+\end{singlespace} 
+\end{itemize}
+We thank you for your previous comments and for consideration of this revised manuscript. 
+To aid in your consideration, we provide  previous reviews and a description of how we have addressed individual comments, below. 
+\\
+\\
+Sincerely,  \\
+Yaniv Brandvain and Graham Coop.
 
-Please visit http://submit.genetics.org to view the comments of the reviewers of your submission, "Sperm should evolve to make female meiosis fair." \\
+\newpage
+\section*{Decision Letter}
+
+Dear Yaniv:  \newline
+%Please visit http://submit.genetics.org to view the comments of the reviewers of your submission, "Sperm should evolve to make female meiosis fair." \\
 
@@
 However, both have raised substantive issues that indicate that the interesting ideas presented need more development before consideration for publication.
+\newline
 
@@
 Reviewer 2 has provided many valuable suggestions along these lines.
+\newline
 
@@
 A manuscript that addressed all concerns raised would essentially constitute a new work, and could be reviewed as a new submission.
-\\
-Best wishes,
-Marcy
-\\
+\\ \\
+Best wishes, \\
+Marcy \newline 
 I regret to have to bring this unwelcome news.  I hope you will continue to submit your best work for publication in the journal, and that I can bring more favorable news when you do.
-\\ \\
-Sincerely,\\
-Marcy Uyenoyama\\
-Associate Editor\\
-Senior Editor
+%\\ 
+%Sincerely,\\
+%Marcy Uyenoyama\\
+%Associate Editor\\
+%Senior Editor
 \newpage
@@
 
-This manuscript uses simple population genetic models to evaluate the potential
+\emph{{\bf{Overview -- }} This manuscript uses simple population genetic models to evaluate the potential
 for sperm haplotype (or paternal genotype) to influence female meiotic drive. 
@@
 be published somewhere but I am not sure it is enough of a sufficient advance to
-warrant publication in Genetics.
+warrant publication in Genetics.}
+
+{\bf{Response:}} We appreciate that the Reviewer found the paper to be `a pleasure 
+to read' and that the models  where `straightforward and clearly explained.'
+We do however insist that this is a sufficient advance for publication in GENETICS --
+ we illustrate the expected outcomes of a previously under-appreciated feature of sperm-
+ egg interactions and present a series of novel and testable hypotheses. 
 \begin{enumerate}
 \item
-A major part of the motivation for the model seems to be to explain why, in
+\begin{singlespace}
+\emph{A major part of the motivation for the model seems to be to explain why, in
 animals, meiosis is typically suspended until fertilization occurs (providing
@@
 meiosis, then it would be better to do a model that explicitly addressed that
-issue.
+issue.  }
 
-Obviously, the authors can avoid this problem by simply saying, given that
+\emph{Obviously, the authors can avoid this problem by simply saying, given that
 meiotic arrest occurs and provides the opportunity for males to influence drive,
 we want to see what will happen.  However, it is misleading to present this
-model as though it addresses the issue of the evolution of meiotic arrest.
+model as though it addresses the issue of the evolution of meiotic arrest.}
+\end{singlespace}
 
-\yb{We thank the reviewer for this comment, and have reworked our manuscript to focus 
+\begin{singlespace}
+{\bf{Response -- }} We thank the reviewer for this comment, and have reworked our manuscript to focus 
 	around our major result -- that sperm are expected to evolve to prevent female meiotic drive. 
+	We completely agree that this delay connote be favored by ``evolutionary foresight'' and have 
+	modified our manuscript to ensure that readers do not interpret our argument in this way 
+	(in fact, we specifically clarify that this is not our argument in the revised introduction and discussion). 
 	We view this as an important point, and as the significant change in this new manuscript. 
-	As suggested by R1, we focus our discussion of the fertilization requirement to pointing out that it provides sperm an opportunity to influence female meiosis, 
-	and we only briefly point out that given the potential reasons for this delay, the threat of sperm collaboration in meiotic drive does not represent a selective force favoring the completion of meiosis before fertilization.}
+	As suggested by R1, we focus our discussion of the fertilization requirement to pointing out that it 
+	provides sperm an opportunity to influence female meiosis.
+
+	We clarify this point by referring readers to a list of hypotheses proposed to explain female meiotic arrest. 
+	We point pout that, whatever favors this arrest,	 the threat of sperm collaboration in meiotic drive does not 
+	represent a selective force favoring the  completion of meiosis before fertilization.
+	We do argue, however, that whatever the evolutionary reason for the origin of this delay, the cost o losing 
+	sperm protection against meiotic drive could play a role in the evolutionary maintenance  of the female meiotic arrest. \\
+\end{singlespace}
 
 \item
+\begin{singlespace}
 I did not feel that the authors made a convincing case that this is an
@@
 evidence it is hard to know whether this is just an interesting idea that never
-really happens. 
+really happens. \\
+\end{singlespace}
+
+\begin{singlespace}
+	{\bf{Response -- }}  We believe that this is an important phenomenon. 
+	We have modified the manuscript to clarify that it is  quite impressive that we find evidence
+	for male influence on female drive in BOTH cases where this phenomenon could possibly be observed.
+	We note that although direct evidence for female meiotic drive is modest, 
+	most experiments to dates have lacked the power to detect most modest drivers. 
+	We believe that given our theoretical predictions and the affordability of large scale GBS experiments,
+	data weighing in on this paper is on its way. 
+	We appreciate the opportunity to get ahead of this data by offering strong predictions that could guide analyses.
+\end{singlespace}
 \end{enumerate}
@@
 
-\subsection*{R2 Summary}
+\emph{{\bf{Overview -- }} 
 The focus of the manuscript entitled  ``Sperm should evolve to make female meiosis
@@
 drive by creating the opportunity for sperm to also suppress the trait. 
-
 	 Overall I found the paper to be well written and organized, and because of
@@
 below.  
+}
 
-\subsection*{R2 Primary suggestions}
+{\bf{Response:}} We appreciate the reviewers comments and encouragement. We address the specific comments below.
+
 \begin{enumerate}

### 2014-07-25 writeup_and_notes/GeneticsResub/male_dependent_drive.tex
added refs and small changes to intro

@@
 
-\yb{Before doing so, it is worth considering whether there is any evidence that sperm could (or do), influence female meiosis. }
-\gc{It is becoming
+Before doing so, it is worth considering whether there is any evidence that sperm could (or do), influence female meiosis. 
+It is becoming
         increasingly clear that sperm bring a wide variety of RNA and proteins into the
-        egg REF. 
+        egg \citep{Miller:05}. 
  Some of these have known functions, for example, in most
@@
         responsible for the transmission of the centriole, a vital
-        component of the mitotic machinery for the zygote. 
+        component of the mitotic machinery for the zygote \citep{Schatten:94}. 
 Detailed functional studies and paternal effect mutations in model
@@
  that sperm-transmitted products have a wide-range of functions in
- the completion of syngamy, zygotic development, and the resumption of
- female meiosis \cite{Miller:01, McNally:05,REF}. 
- For example, in \emph{C. elegans}, premature deployment of the aster 
- 	-- a vital component of mitotic machinery provided by the sperm, 
-	disrupts MII meiotic segregation in the egg, leading to a triploid zygote \citep{McNally2012}.
+ the \gc{egg activation, completion of syngamy, zygotic development, and the resumption
+ and successful completion of
+ female meiosis} \cite[e.g.][]{Yasuda:95,Loppin:05,Miller:01, McNally:05,Churchill:03}. 
+ For example, in \emph{C. elegans}, \gc{premature deployment of the sperm aster 
+	disrupts MII meiotic segregation in the egg, leading to a triploid zygote} \citep{McNally2012}.
 However,
@@
 hence alleles could potentially have functional effects on
-various aspects of female meiosis that occur after fertilization. } 
+various aspects of female meiosis that occur after fertilization. 
 %Why then is the requirement of fertilization to complete female meiosis so ubiquitous? 
-\gc{THERE may be too much bio. detail in previous paragraph, I can
-  add more refs. once we've decided on content
-centriole : http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3144413/?report=classic}
 \newline
@@
 %	disrupts MII meiotic segregation in the egg, leading to a triploid zygote \citep{McNally2012}.
-\yb{Current evidence suggests that sperm influence on female meiotic drive is not only possible, but likely. }
+Current evidence suggests that sperm influence on female meiotic drive is not only possible, but likely. 
 Evidence from both the \emph{In} and \emph{Om}  -- suggests that the 
@@
 Thus, there is little selective benefit to females in preventing sperm to influence female meioses. 
- \yb{In fact, a mutant female which completes meiosis before 
- 	fertilization is giving up the protection from meiotic drivers potentially offered by sperm, 
-	and is therefore more likely to produce low fitness offspring. }
-%\gc{We do not claim that female meiotic arrest evolved
+\gc{Given the wide-spread requirement of fertilization for the completion
+of  female meiosis it seems possible that various feature of the
+interaction between sperm and egg have evolved as another mechanism to enforce an equitable transfer of genetic material.}
+\gc{HOW ABOUT THE ABOVE.}
+
+%%don't thi k this helped: 
+%\yb{In fact, a mutant female which completes meiosis before 
+ %	fertilization is giving up the protection from meiotic drivers potentially offered by sperm, 
+%	and is therefore more likely to produce low fitness offspring. }
+%%\gc{We do not claim that female meiotic arrest evolved
 %  specifically to recruit helpful sperm to defend against meiotic
@@
         that phylogenetic variation in the stage of female meiosis when
-        fertilization occurs to influence the prevalence of
+        fertilization occurs may influence the prevalence of
         female meiotic drive.

### 2014-07-25 writeup_and_notes/GeneticsResub/reviews.tex

@@
 	We believe that given our theoretical predictions and the affordability of large scale GBS experiments,
-	data weighing in on this paper is on its way. 
+	data weighing in on this paper are on its way. 
 	We appreciate the opportunity to get ahead of this data by offering strong predictions that could guide analyses.
@@
 \begin{enumerate}
-\item	The importance of the wok hinges critically on the the degree to which:  {\bf{\emph{i)}}}
+\begin{singlespace}
+\item	\emph{The importance of the wok hinges critically on the the degree to which:  {\bf{\emph{i)}}}
 female meiotic drive operates in nature, and {\bf{\emph{ii)}}} sperm influence the level of
@@
 the reader that female meiotic drive is sufficiently common to be a widespread
-selective force.
+selective force.}
+\end{singlespace}
+\begin{singlespace}
+	{\bf{Response -- }}  We agree and have added more support of the likely importance and prevalence of female meiotic drive. 
+	We hope that new theoretical work (including ours) coupled with the falling expense of GBS, will encourage the discovery of additional drive systems in non-model organisms in the coming years.
+\end{singlespace}
 
-\item	 To my knowledge, the evidence for sperm influencing the outcome of meiotic
+\begin{singlespace}
+\item	 \emph{To my knowledge, the evidence for sperm influencing the outcome of meiotic
 drive is restricted to the In and Om drive systems in mice and aster transfer by
-the sperm in C. elegans.  These are clearly proof of process examples but I
+the sperm in \emph{C. elegans}.  These are clearly proof of process examples but I
 think that many readers will take issue with the assumed general ability of
@@
 effects via RNA and protein packaged in the sperm, delivered to the egg, and
-influencing subsequent early embryonic development.
+influencing subsequent early embryonic development. }
+\end{singlespace}
+\begin{singlespace}
+	{\bf{Response -- }}  
+	We have added a discussion of these points to our introduction. 
+	Specifically, we now discuss the influence of sperm on basic processes in the egg, 
+	and highlight that although data are preliminary $In$ and $Om$ are the best cases of polymorphic female drivers in animals, 
+	and therefore current support for our theory is actually quite high. 
+\end{singlespace}
 
-\item	Finally, female meiotic drive potentially can be suppressed by alleles
+\begin{singlespace}
+\item	 \emph{Finally, female meiotic drive potentially can be suppressed by alleles
 expressed in the female (traditional models) or by alleles originating in the
@@
 has some evolutionary advantage that make it easier to evolve or non-redundant
-to female-coded suppressors.
+to female-coded suppressors. }
+\end{singlespace}
+\begin{singlespace}
+	{\bf{Response -- }}  For comparison, we have added a female meiotic drive suppressor unlinked to a driver as model 
+	$6^\prime$, and compared its evolution to a sperm acting driver.
+	In general, the spread of an unlinked female drive suppressor is qualitatively similar to that of an unlinked sperm-acting drive suppressor; 
+	however, the initial spread of the sperm suppressor is equal to (when the female suppressor is dominant), or more rapid than (otherwise) the female suppressor. 
+	Given the difference in mechanism of female drive suppression in sperm as compared to females, sperm suppression of female drive is likely not functionally redundant with female drive suppression and provides a qualitatively different mechanism of suppression.
+\end{singlespace}
+
 \end{enumerate}
@@
 \begin{enumerate}
-\item   On the top of page 7 the statement ``This analytical approximation, derived
+\begin{singlespace}
+\item \emph{  On the top of page 7 the statement ``This analytical approximation, derived
 from Equation (1) assuming Hardy-Weinberg'', -- precedes the introduction of
@@
 reading and start looking for them, thinking they somehow overlooked them in the
-body of the manuscript).
-\yb{We can relate to the confusion and frustration of struggling to find the relevant equation.
-We have therefore added the parenthetical comment --   ``\emph{We note that Equation 1 and all equations discussed in the main text are presented in the appendix and derived in File S2}'' -- to the end of this paragraph.}
+body of the manuscript). \\ \\}
+{\bf{Response -- }} We can relate to the confusion and frustration of struggling to find the relevant equation.
+We have therefore added the parenthetical comment --   ``\emph{We note that Equation 1 and all equations discussed in the main text are presented in the appendix and derived in File S2}'' -- to the end of this paragraph.
+We also believe that the METHODS section, which we present before the results in this new manuscript, will help prepare readers for this. 
+\end{singlespace}
 
-\item There are 6 models described in the text but these are not described in any
+\begin{singlespace}
+\item \emph{There are 6 models described in the text but these are not described in any
 detail until the Appendix.  The authors might consider placing a table in the
-main body of the text that tersely describes the 6 models.?
-\yb{We have added a Methods section, which introduces these six models, and makes our manuscript more in line with the GENETICS format.}
+main body of the text that tersely describes the 6 models? \\ \\}
+{\bf{Response -- }} We have added a Methods section, which introduces these six models, and makes our manuscript more in line with the GENETICS format.
+\end{singlespace}
 
-\item On the bottom of page 14 the authors write ``we predict that a lower
+\begin{singlespace}
+\item \emph{On the bottom of page 14 the authors write ``we predict that a lower
 opportunity for female meiotic drive, e.g. an animal lineage with a long history
@@
 not have a ``LONG history of high inbreeding or selfing'') so this prediction may
-be limited.
+be limited. \\ \\ }
+{\bf{Response -- }} We agree that biology (e.g. extinction) might conspire to make this a challenging 
+prediction to test, and we point this out in our new manuscript. 
+However, we believe that this hypothesis follows from our results and therefore still propose it.
+\end{singlespace}
 
-\item On the bottom of page 14, do you also predict that observed drive in species
-hybrids will depend on the direction (which male) is used in the backcross?
+\begin{singlespace}
+\item \emph{On the bottom of page 14, do you also predict that observed drive in species
+hybrids will depend on the direction (which male) is used in the backcross?\\ \\}
+{\bf{Response -- }}  We have expanded and clarified our prediction regarding interspecific crosses in our new manuscript. 
+\end{singlespace}
 
-\item I think that equations like $(s < d -1/2 + 3sh/2 - dsh)$ need parentheses to be
-clear. 
-\yb{To add clarity, we have surrounded longer in-text equations like the example above, with parentheses.}
+\begin{singlespace}
+\item 
+\emph{I think that equations like $(s < d -1/2 + 3s_h/2 - ds_h)$ need parentheses to be clear. \\ \\ }
+	{\bf{Response -- }}  To add clarity, we have surrounded longer in-text equations like the example above, with parentheses.
+\end{singlespace}
+%\yb{To add clarity, we have surrounded longer in-text equations like the example above, with parentheses.}
 

### 2014-07-25 writeup_and_notes/GeneticsResub/reviews.tex
changes to reviews

@@
 Reviewer 2 has provided many valuable suggestions along these lines.}  \newline \\
-{\bf{Response -- }} This revision makes it clear that our primary argument is that, given a chance to influence female meiosis, sperm will evolve to increase its fairness (as reflected in our title), {\bf{AND NOT}} that females should delay the completion of meiosis so as to recruit help from sperm in making meiosis fair (as could have been interpreted from our previous draft).
+{\bf{Response -- }} \gc{We thank the reviewers and editor for highlighting
+this point. In this revision we make it clear} that our primary
+argument is that, given a chance to influence female meiosis, sperm
+will evolve to increase its fairness (as reflected in our title), {\it
+  and not} that females should delay the completion of meiosis so as to recruit help from sperm in making meiosis fair (as could have been interpreted from our previous draft). 
 \end{singlespace}
@@
 (1) a rigorous and theoretical model, and 
-(2) research questions that can be addressed by molecular genetics, transmission genetics, studies of genome structure, phylogenetics and population genetics.
-As such, GENETICS is a particularly appropriate venue for our paper.
-Our view is echoed by comments from Reviewer 2 who stated  `Overall I found the paper to be well written and organized, and because of this, the theory that the authors describe will be accessible to a broad spectrum of the readership of Genetics ?
-not just other theoreticians.'
+(2) research questions that can be addressed by molecular genetics,
+transmission genetics, studies of genome structure, phylogenetics and
+population genetics. As such, GENETICS is a particularly appropriate
+venue for our paper.
+Our view is echoed by comments from Reviewer 2 who stated  `Overall I found the paper to be well written and organized, and because of this, the theory that the authors describe will be accessible to a broad spectrum of the readership of Genetics not just other theoreticians.'
 The suitability of the topic of our manuscript is clear from our reference list -- a strong plurality (over 25\%) of our works cited appeared in GENETICS, placing our work squarely within the tradition of GENETICS papers.
-We have therefor corrected issues with our formatting, and included a METHODS section to introduce our models, as is the tradition in GENETICS. \\
+\gc{We apologize for the formatting issues. We have corrected these
+  issues, and included a short METHODS section to introduce our models.} \\
 \end{singlespace} 
@@
 We thank you for your previous comments and for consideration of this revised manuscript. 
-To aid in your consideration, we provide  previous reviews and a description of how we have addressed individual comments, below. 
+To aid in your consideration, we provide \gc{the previous reviews and a
+detailed response below. }
 \\
@@
 
-{\bf{Response:}} We appreciate that the Reviewer found the paper to be `a pleasure 
-to read' and that the models  where `straightforward and clearly explained.'
+%%I'd not use quotes right below the quoted text, seems trite.
+{\bf{Response:}} \gc{We appreciate that the Reviewer enjoyed the paper.}
 We do however insist that this is a sufficient advance for publication in GENETICS --
@@
 	around our major result -- that sperm are expected to evolve to prevent female meiotic drive. 
-	We completely agree that this delay connote be favored by ``evolutionary foresight'' and have 
-	modified our manuscript to ensure that readers do not interpret our argument in this way 
-	(in fact, we specifically clarify that this is not our argument in the revised introduction and discussion). 
-	We view this as an important point, and as the significant change in this new manuscript. 
-	As suggested by R1, we focus our discussion of the fertilization requirement to pointing out that it 
-	provides sperm an opportunity to influence female meiosis.
+	We completely agree that this delay \gc{cannot} be favored by ``evolutionary foresight'' and have 
+	modified our manuscript to ensure that readers do not
+        interpret our argument in this way. \gc{We have now added
+          significantly to our  revised introduction and discussion to
+          clarify this point. 	
+We also now more explicitly refer readers to a discussion of
+        hypotheses proposed to explain female meiotic arrest (Mira
+        '98), so that our discussion can be seen as being distinct
+        from that. }
 
-	We clarify this point by referring readers to a list of hypotheses proposed to explain female meiotic arrest. 
-	We point pout that, whatever favors this arrest,	 the threat of sperm collaboration in meiotic drive does not 
-	represent a selective force favoring the  completion of meiosis before fertilization.
-	We do argue, however, that whatever the evolutionary reason for the origin of this delay, the cost o losing 
-	sperm protection against meiotic drive could play a role in the evolutionary maintenance  of the female meiotic arrest. \\
+%	As suggested by R1 \gc{???}, we focus our discussion of the fertilization requirement to pointing out that it 
+%	provides sperm an opportunity to influence female meiosis.  
+%	
+%	We point out that, whatever favors this arrest, the threat of sperm collaboration in meiotic drive does not 
+%	represent a selective force favoring the completion of meiosis before fertilization.
+%	We do argue, however, that whatever the evolutionary reason for the origin of this delay, the cost o losing 
+%	sperm protection against meiotic drive could play a role in the evolutionary maintenance of the female meiotic arrest. \\
 \end{singlespace}
@@
 \begin{singlespace}
-	{\bf{Response -- }}  We believe that this is an important phenomenon. 
-	We have modified the manuscript to clarify that it is  quite impressive that we find evidence
-	for male influence on female drive in BOTH cases where this phenomenon could possibly be observed.
-	We note that although direct evidence for female meiotic drive is modest, 
-	most experiments to dates have lacked the power to detect most modest drivers. 
-	We believe that given our theoretical predictions and the affordability of large scale GBS experiments,
-	data weighing in on this paper are on its way. 
-	We appreciate the opportunity to get ahead of this data by offering strong predictions that could guide analyses.
+	{\bf{Response -- }}  \gc{We believe that this is an important
+        phenomenon. We have strengthened our discussion of this point
+        throughout the introduction,
+        we have also highlighted the points raised by R2 about 
+        the evidence of pervassive female drive and the opportunities for
+        sperm-based products
+        to interfer with female meiosis.
+	We have modified the manuscript to clarify that it is quite impressive that there is evidence
+	for male influence on female drive at both of the two described
+        female meiotic drive elements in animals. 
+	We note that although direct evidence for polymorphic female meiotic drive
+        in animals is modest (i.e. these two loci in mice), 
+	this likely reflects the fact that it is very hard to study
+        outside of model organisms. 
+Given the increasing tractibility of genetics in many traditionally
+        non-model systems, and the growing awareness of the importance
+        of genetic conflicts in shaping molecular systems, we think
+        that it is timely to offer a framework to understand one
+        aspect of sperm-egg 
+        interactions and provide
+strong predictions that could guide analyses.}
 \end{singlespace}
@@
 
-{\bf{Response:}} We appreciate the reviewers comments and encouragement. We address the specific comments below.
+{\bf{Response:}} We appreciate the reviewers comments and
+encouragement. 
+We address the specific comments below.
 
@@
 	{\bf{Response -- }}  We agree and have added more support of the likely importance and prevalence of female meiotic drive. 
-	We hope that new theoretical work (including ours) coupled with the falling expense of GBS, will encourage the discovery of additional drive systems in non-model organisms in the coming years.
+	We hope that new theoretical work (including ours) coupled
+        with the increasing easy of genetics in many system, will encourage the discovery of additional drive systems in non-model organisms in the coming years.
 \end{singlespace}

### 2014-07-26 writeup_and_notes/GeneticsResub/male_dependent_drive.tex
small final changes to main text

@@
 \gc{Given the wide-spread requirement of fertilization for the completion
-of  female meiosis it seems possible that various feature of the
-interaction between sperm and egg have evolved as another mechanism to enforce an equitable transfer of genetic material.}
-\gc{HOW ABOUT THE ABOVE.}
+of  female meiosis it seems possible that various features of the
+interaction between sperm and egg have evolved as a mechanism to enforce an equitable transfer of genetic material.}
 
@@
         completion of female meiosis (although 
-   \yb{     opportunities to test this hypothesis may be limited because }
+        opportunities to test this hypothesis may be limited because 
         %this effect may be weaks as such
@@
         female meiotic drive.
-For example, in taxa where meiosis I is completed
-by the egg pre-fertilization, including humans, centromere drive may
-be more common as sperm based modifiers can not intercede than in
-species where fertilization occurs during MI. 
+\gc{For example, in taxa where meiosis I is completed
+by the egg pre-fertilization centromere drive may
+be more common as sperm-based modifiers can not intercede, as compared to
+species where fertilization occurs during MI. }
 Such hypotheses could be tested if the speed of centromere
-turnover serves as a proxy for the efficacy of MI drive. }
+turnover serves as a proxy for the efficacy of MI drive. 
 \newline
@@
 	fertilizing sperm decidedly side with mom, as both have a shared interest in producing a viable and potentially fit offspring. 
-\yb{Our model does not directly speak to the evolutionary origin of
+Our model does not directly speak to the evolutionary origin of
   female meiotic arrest \citep[for a review and evaluation of such
@@
 However, given the existence of female meiotic arrest, 
-	and that its timing relative to
-	fertilization as well as the interactions between sperm and egg are variable across species  \citep{Karrbook:09} 
-	the nature of the meiotic arrest and interactions between sperm and egg may be molded by selection to reduce the opportunity for female meiotic drive, 
-	while selfish drivers may evolve in concert, to overcome these adaptations. 
+	\gc{and that its timing and mechanistic details are variable across species \citep{Masui_book,Karrbook:09} 
+	the nature of the meiotic arrest and interactions between
+        sperm and egg may  be molded by selection to reduce the opportunity for female meiotic drive, 
+	and selfish drivers evolving in concert to overcome these adaptations. }
 %leave female meiosis open to sabotage by sperm collaborators, but
@@
 %evolutionary loss of female meiotic arrest will result in the loss of sperm defense against female meiotic drivers. }
-  }
+
 %However, given that arrest does occur its timing relative to

### 2014-07-27 writeup_and_notes/GeneticsResub/male_dependent_drive.tex
yb (nearly) finalized changes for resub

@@
 O\"{o}genesis is inherently susceptible to such conflicts
-because alleles compete \yb{for inclusion in} the  product of female meiosis transmitted to the egg. 
+because alleles compete for inclusion in the  product of female meiosis that will be transmitted to the egg. 
 Alleles that distort meiosis in their favor (i.e. meiotic drivers) often decrease organismal fitness,
@@
 	interpreted as mechanisms of protection against genomic outlaws. 
-The fact that female meiosis \yb{is} often left uncompleted until after
-      fertilization in many animals appears to run counter to this interpretation, 
-      as this delay provides an opportunity for sperm\yb{-acting} alleles to
-     meddle with its outcome and help like-alleles drive in heterozygous females. 
-%Why then is female meiosis often left uncompleted until after
-%fertilization in many animals -- 
-%	potentially providing an opportunity for sperm alleles to
-%     meddle with its outcome and help like-alleles drive in heterozygous females? 
-The population genetic theory presented herein suggests that
-	\gc{in fact} sperm nearly always evolve to increase the fairness of female
+That females  of many animal species do not complete meiosis until after
+     fertilization appears to run counter to this interpretation
+     because this delay provides an opportunity for sperm-acting alleles to
+     meddle with the outcome of female meiosis and help like-alleles drive in heterozygous females. 
+Contrary to this perceived danger, the population genetic theory presented herein suggests that, 
+	in fact, sperm nearly always evolve to increase the fairness of female
         meiosis in the face of genomic conflicts. 
-These results are consistent with \yb{the apparent sperm- dependence}  \gc{of the best characterized female meiotic drivers in animals to date.} 
- \yb{Rather than providing an opportunity for sperm collaboration in female meiotic drive,
+These results are consistent with the apparent sperm-dependence of the best characterized female meiotic drivers in animals to date.  
+Rather than providing an opportunity for sperm collaboration in female meiotic drive,
         the `fertilization requirement'  indirectly protects females from
         meiotic drivers by providing sperm an opportunity to
-        suppress female meiotic drive. }
+        suppress drive. 
 %	and suggest that the requirement of fertilization for the
@@
      could presumably rapidly spread through a population. 
-At first sight, female meiosis appears primed for \gc{conflict caused by
-the such selfish systems.} Here we ask if sperm do indeed evolve to collaborate with female drivers to exploit this apparent weakness in the defense against meiotic drive. \newline
+At first sight, female meiosis appears primed for conflict caused by such selfish systems. 
+Here we ask if sperm do indeed evolve to collaborate with female drivers to exploit this 
+	apparent weakness in the defense against meiotic drive. \newline
 
@@
  Some of these have known functions, for example, in most
-        animal species sperm-- not eggs-- are
+        animal species sperm -- not eggs -- are
         responsible for the transmission of the centriole, a vital
         component of the mitotic machinery for the zygote \citep{Schatten:94}. 
-Detailed functional studies and paternal effect mutations in model
-systems further highlight
- that sperm-transmitted products have a wide-range of functions in
- the \gc{egg activation, completion of syngamy, zygotic development, and the resumption
+Detailed functional studies and analyses of paternal effect mutations in model
+systems further highlight that sperm-transmitted products have a wide-range of functions in
+ egg activation, completion of syngamy, zygotic development, and the resumption
  and successful completion of
- female meiosis} \cite[e.g.][]{Yasuda:95,Loppin:05,Miller:01, McNally:05,Churchill:03}. 
- For example, in \emph{C. elegans}, \gc{premature deployment of the sperm aster 
-	disrupts MII meiotic segregation in the egg, leading to a triploid zygote} \citep{McNally2012}.
-However,
-        the function of many of the products the sperm brings into the
+ female meiosis \cite[e.g.][]{Yasuda:95,Loppin:05,Miller:01, McNally:05,Churchill:03}. 
+ For example, in \emph{C. elegans}, premature deployment of the sperm aster 
+	disrupts MII meiotic segregation in the egg, leading to a triploid zygote \citep{McNally2012}.
+However, the function of many of the products the sperm brings into the
         egg is completely unknown and 
       these products vary widely over species \citep{Karrbook:09}. 
-It seems quite plausible that sperm-based products and
-hence alleles could potentially have functional effects on
+It seems quite plausible that sperm-based products, and
+	hence sperm haplotype or paternal genotype could influence 
 	various aspects of female meiosis that occur after fertilization. 
-%Why then is the requirement of fertilization to complete female meiosis so ubiquitous? 
 \newline
 
-%It is certainly not the case that sperm is mechanistically incapable of influencing female meiosis.
-%Both mechanistic and genetic evidence suggests that sperm can influence the outcome of female meiosis.
-%An example of mechanistic evidence for sperm's influence on female meiosis comes from \emph{C.
-elegans}, where premature deployment of the aster -- a vital component of mitotic machinery provided by the sperm,
-%	disrupts MII meiotic segregation in the egg, leading to a triploid zygote \citep{McNally2012}.
-Current evidence suggests that sperm influence on female meiotic drive is not only possible, but likely. 
-Evidence from both the \emph{In} and \emph{Om}  -- suggests that the 
-	 transmission of alternate alleles in heterozygous females can depend on sperm haplotype.  
+Current evidence from the best characterized systems of female meiotic drive in animals 
+	(the  \emph{In} and \emph{Om} loci in mice) 
+	suggests that sperm influence on female meiotic drive is not only possible, but likely. 
+%Evidence from both the \emph{In} and \emph{Om}  -- suggests that the 
+%	 transmission of alternate alleles in heterozygous females can depend on sperm haplotype.  
 %There are current only two polymorphic, well-characterized female meiotic drive systems in
 %animals-- and both of them, the \emph{In} and \emph{Om}  drive systems in mice-- suggests that the 
-	 transmission of alternate alleles in heterozygous females can depend on sperm haplotype.  
+%	 transmission of alternate alleles in heterozygous females can depend on sperm haplotype.  
  While ruling out the alternative hypothesis of early selection on zygotes in these cases is challenging  \citep[see pages 52-54 in][ for comment]{Burt2006}, it appears that the extent to which \emph{In} and \emph{Om} distort the second female meiotic division partially depends on the
 	genotype of the fertilizing sperm \citep{Agulnik1993,Wu2005}. 
-\yb{The fact that the two best characterized, polymorphic systems of putative female 
-	meiotic drive systems identified to date in animals (\emph{In} and \emph{Om}) 
-	show this effect suggests that if female meiotic drive is
-	common the role of sperm in modifying female drive will be important as well. } \newline
+The fact that the two best characterized, polymorphic systems of putative female 
+	meiotic drive systems in animals 
+	show this effect suggests that if female meiotic drive is common 
+	the role of sperm in modifying female drive will be important as well.  \newline
 
-\yb{Numerous lines of evidence suggest that female meiotic drive is an 
-	evolutionary force, and therefore the opportunity for sperm to influence female drive likely had applications to many animals}.   
-While, research to date has identified a few extreme cases of female meiotic 
+Numerous lines of evidence suggest that female meiotic drive is a common and important 
+	evolutionary force, and therefore the opportunity for sperm to influence female drive likely had applications to many animals.   
+While research to date has identified a few extreme cases of female meiotic 
 	drive in the small number model systems systematically
         studied \citep{Agulnik1990, Fishman2008, Hiatt2003,Novitski1951, Pardo-ManuelDeVillena2001c},
-	rapid evolution of centromeres, telomeres and other basic components of the meiotic machinery, 
-	suggest consistent evolution of female meiotic drivers and suppressors of meiotic drive
-	 \citep[e.g. ][]{Anderson2008, Anderson2009, Axelsson2010,
-           Malik2009a}. 
+	rapid evolution of the basic components of the meiotic machinery (e.g. centromeres, telomeres, etc \dots) 
+	suggest consistent selection on female meiotic drivers and suppressors of meiotic drive in many animal species 
+	 \citep[e.g. ][]{Anderson2008, Anderson2009, Axelsson2010, Malik2009a}. 
  We expect that over the next decade the spread of sequencing to a
@@
 
-\yb{Because female meiotic drive is likely common and consequential, 
+Because female meiotic drive is likely a common force with predictably negative effects on organismal fitness,  
 	and because sperm have ample opportunity to influence female drive,  
-	we develop population genetic models exploring the evolution of sperm influence on female meiotic drive -- 
-	that is, what is the expected influence of sperm on female drive? } 
+	we develop population genetic models to address the the expected influence of sperm on female drive.  
 We first focus on models in which `self-promoting' alleles in
@@
 as both are invested in the fate of the resultant zygote (as was speculated for the \emph{In} locus \citep{Pomiankowski1993}).
-Thus, there is little selective benefit to females in preventing sperm to influence female meioses. 
-\gc{Given the wide-spread requirement of fertilization for the completion
-of  female meiosis it seems possible that various features of the
-interaction between sperm and egg have evolved as a mechanism to enforce an equitable transfer of genetic material.}
+Thus, there is little selective benefit to females in preventing sperm to influence female meioses, 
+	\yb{and in fact, females eschewing this delay would potentially lose sperm assistance in the suppression of meiotic drivers.}  
+Given the wide-spread requirement of fertilization for the completion
+	of  female meiosis, various features of the interaction between sperm 
+	and egg likely evolved as a mechanism 
+	to enforce an equitable transfer of genetic material.
 
@@
 
-\yb{
 \section*{Methods}
@@
 
-We present six related models, the first three of which represent the system with a single locus.
-{\bf{Model 1}} describes a traditional, single-locus, female meiotic driver, 
+We present six related models -- three single-locus `pleiotropy' models and three two-locus `drive-modifier' models.
+	{\bf{Model 1}} describes a single-locus female meiotic driver.  
 	{\bf{Model 2}} describes a single-locus sperm-dependent female driver -- that is, an allele whose transmission in female meiosis depends on sperm haplotype. 
 	{\bf{Model 3}} describes a single-locus paternal-dependent female driver -- an allele whose transmission in female meiosis depends on paternal genotype. 
-Assuming that a traditional driver segregates at its equilibrium frequency ({\bf{Model 1}}), we investigate the evolution of tightly linked ({\bf{Models 4 and 5}}), and unlinked ({\bf{Model 6}}) sperm-dependent modifiers of drive (We also briefly compare Model 6, to Model 6' - the suppression of female drive by an unlinked female acting suppressor).
+Assuming that a traditional driver segregates at its equilibrium frequency (identified in Model 1), 
+	we investigate the evolution of tightly linked ({\bf{Models 4 and 5}}), and unlinked ({\bf{Model 6}}) 
+	sperm-dependent modifiers of drive. 
 In {\bf{Models 4 and 5}}, we treat this two-locus system as if it consists of
 	a single locus with three alleles: $A$, $B$ and $C$, corresponding to the case when the
-sperm-modifier arises once, tightly linked to one or other of the
-alleles at the drive locus (such that recombination is unexpected). 
+	sperm-modifier is very tightly linked to the driving ({\bf{Model 4}}) or non-driving allele ({\bf{Model 5}})
+ 	at the drive locus such that recombination is unexpected. 
+To evaluate the feasibility of sperm-modification of female meiotic drive, as compared to female suppression of drive, 
+	we conclude with a model of female drive suppression by an unlinked female acting suppressor ({\bf{Model $6^\prime$}}). 
 In all cases, we assume that fitness is independent of the drive modifier. 
@@
 
-All models include a biallelic locus ($A/B$) with non-driving and driving alleles in frequencies $f_A$ and $f_B$, 
-	respectively, while more complex models include more loci and/or alleles.  
-Transmission rules describing the outcomes of all matings in each
-	model are presented in a File S1 . 
+All models include a biallelic locus ($A/B$) with non-driving and driving alleles in frequencies $f_A$ and $f_B = 1 - f_A$, 
+	respectively, while Models 4-6 include a drive-modifying locus.  
+Transmission rules describing the outcomes of all matings in each model are presented in a File S1. 
 The fitness of genotype, $g$, is sex-independent and equals $1$, $1-h_s$, and $1-s$ for genotypes $AA$, $AB$, and $BB$, respectively. %  and  $w_{AA} = 1$, $w_{AB} = 1-s_h$, and $w_{BB}=1-s$. 
@@
 	and $f_g''$ in adults after selection. 
-After a complete generation $f_g'' = f_g' w_g/\bar{w}$, where $\bar{w}$ is the population mean fitness and equals $\Sigma w_g f_g'$.  \newline
+After a complete generation genotype frequencies are $f_g'' = f_g' w_g/\bar{w}$, where $\bar{w}$ is the population mean fitness and equals $\Sigma w_g f_g'$.  \newline
 
 We verbally describe our main results below.  
-Readers interested in the details of these results should turn to the Appendix for a mathematical treatment. 
+Readers interested in the details of these results should turn to the Appendix for a mathematical treatment, 
+	and to our Mathematica worksheet  (File S2) for our complete derivations. 
 There we present critical analytical results in Equations 1-11, and describe our analyses and results in more detail.  
@@
 that genotype frequencies follow Hardy Weinberg Equilibrium (HWE), we note which analytical results are approximations. 
-We verify these approximations with exact numerical iterations in Figures S1-S3 and File S3 , and detail the derivation of these approximations in our File S2.
-}
+We verify these approximations with exact numerical iterations in Figures S1-S3 and File S3.
 
@@
 	closely matches results obtained by exact numerical iteration (Figure \ref{Eggsperm}B.
-\yb{We note that Equation 1 and all equations discussed in the main text are presented in the appendix and derived in File S2}).
+We remind readers that Equation 1 and all equations discussed in the main text are presented in the Appendix and derived in File S2).
 \newline
@@
         \ref{unlinked}).  %%DELETED FIG REF HERE. 
+ We note that the evolution of sperm-acting drive  suppressors unlinked to a driver (Model 6)
+ 	is both qualitatively and quantitatively similar to the evolution  of a female-acting drive 
+	suppressor (Model 6$^\prime$ -- compare Equations \ref{unlinked} and \ref{femunlinked}). 
         \newline
@@
 This result is consistent with previous work showing that drive suppressors unlinked to, or in repulsion phase with drivers usually invade polymorphic drive systems \citep[e.g. ][]{Brandvain2012}.  
-\yb{We note that the evolution of sperm-acting drive  suppressors unlinked to a driver (Model 6) is both qualitatively and quantitatively similar to the evolution  of a female-acting drive suppressor (compare Equations \ref{unlinked} and \ref{femunlinked}).}
 Therefore, all two-locus models of sperm influence on female drive suggest that
@@
 	that female o\"{o}genesis and meiosis will evolve to prevent their effect.  
-\gc{Thus females can} delay the completion of meiosis until after fertilization 
+Thus, females can delay the completion of meiosis until after fertilization 
 	without risking exploitation by collaborations between female
@@
 
-\yb{The theory presented herein provides a strong prediction -- 
-organisms in which female 
-meiosis is not completed until after fertilization sperm will act to suppress female drive at the stage at which they can influence meiosis. 
+The theory presented herein provides a strong prediction -- 
+	for organisms in which female  meiosis is not completed until after fertilization, 
+	sperm will act to suppress female drive at the stage at which they can influence meiosis. 
 %Therefore, for example, when fertilized by drying sperm, we expect meiosis to be fairer when  females are fertilized by 
-Contrary to expectations of a \gc{naive verbal} ``green-beard'' model, 
-we are unsurprised by the observation that female meiosis is fairer when fertilized by sperm bearing the drive allele in two of the best described cases of female meiotic drive in animals (the \emph{Om} and \emph{In} loci in mice \citep{Agulnik1993,Wu2005}).
-Given the immense benefit to sperm of hampering female drive, 
+Contrary to expectations of a naive verbal ``green-beard'' model, 
+	our results are consistent with the observation that female meiosis 
+	is more fair when fertilized by sperm bearing the drive allele in 
+	two of the best described cases of female meiotic drive in 
+	animals (the \emph{Om} and \emph{In} loci in mice \citep{Agulnik1993,Wu2005}).  
+Given the benefit to sperm of hampering female drive, 
 	drive-suppressing sperm are often likely to be fixed within a species, 
-making the hypothesis of sperm acting drive-suppression difficult to test from intra-population crosses. 
-However, crosses between populations or species are likely to provide critical tests of our theory -- specifically we predict that female meiosis will be less fair when a species (or population) is fertilized by heterospecific sperm because such sperm have not evolved to counter novel females meiotic drivers.
-We further predict that segregation in F1 females backcrossed to parental species will likely be biased with a deficit of transmission of the paternal species allele from the F1  female.   
+	making the hypothesis of sperm-acting drive-suppression difficult to test from intra-population crosses. 
+However, crosses between populations or species are likely to provide critical tests of our theory -- 
+	specifically we predict that female meiosis will be less fair when a species (or population) is 
+	fertilized by heterospecific sperm because such sperm have not evolved to counter novel females meiotic drivers. 
+We further predict that segregation in F1 females backcrossed to parental species will likely be biased with a deficit 
+	of transmission of the paternal species allele from the F1  female.   
 We note that while these predictions follow straightforwardly from the
 	theory presented above, tests of meiotic drive, 
-and especially sperm dependent meiotic drive require a high standard
-of evidence 
-to exclude plausible alternative hypotheses such as genotypic inviability including epistatic maternal X zygotic lethality.}  \newline
+	and especially sperm-dependent meiotic drive require a high standard
+	of evidence  to exclude plausible alternative hypotheses such as genotypic 
+	inviability including epistatic maternal X zygotic lethality \citep[e.g. ][]{Sawamura:1993aa}.  \newline
 
@@
         female meiotic drive.
-\gc{For example, in taxa where meiosis I is completed
-by the egg pre-fertilization centromere drive may
-be more common as sperm-based modifiers can not intercede, as compared to
-species where fertilization occurs during MI. }
-Such hypotheses could be tested if the speed of centromere
-turnover serves as a proxy for the efficacy of MI drive. 
+For example, in taxa where females complete meiosis
+	before fertilization, centromeric drive may
+	be more common than in species which sperm interact with eggs arrested  in MI, 
+	because in the prior case sperm-based modifiers can only intercede during the $2^\text{nd}$, but not $1^\text{st}$ meiotic division. }
+As a potential test of this hypothesis the speed of centromere
+	turnover could be compared in species in which sperm iterate with eggs paused at MI and MII 
+	(assuming the pace of centromere turnover serves as a proxy for the efficacy of MI drive). 
 \newline
@@
 
-\subsubsection*{Model $6^\prime$. Unlinked female acting drive-modifier:}
-\yb{The dynamics of a female-acting drive modifier are comparable to those describing a female acting drive suppressor.  
-When rare and fully dominant and in linkage and in HWE, the change in frequency of a rare unlinked, female-acting drive modifier is 
+\paragraph{Model $6^\prime$. Unlinked female acting drive-modifier:}
+The dynamics of a female-acting drive modifier are comparable to those describing a sperm-acting drive modifier.  
+Assuming Hardy-Weinberg and linkage equilibrium, the change in frequency of a rare, unlinked, female-acting drive modifier is 
 \begin{equation} \Delta M =-(d_h-d_0) f_A f_B (f_B s + s_h (f_A - f_B) ) \frac{f_M}{\bar{w}}  \label{femunlinked} \end{equation}
@@
  However, if female drive suppression is less than fully dominant, sperm-acting drive suppressors are more efficacious when rare than are female-acting suppressors, and are therefore more likely to spread.
-}
 

### 2014-07-27 writeup_and_notes/GeneticsResub/reviews.tex
yb (nearly) finalized changes for resub

@@
 Reviewer 2 has provided many valuable suggestions along these lines.}  \newline \\
-{\bf{Response -- }} \gc{We thank the reviewers and editor for highlighting
-this point. In this revision we make it clear} that our primary
+{\bf{Response -- }} We thank the reviewers and editor for highlighting
+this point. In this revision we make it clear that our primary
 argument is that, given a chance to influence female meiosis, sperm
@@
 The suitability of the topic of our manuscript is clear from our reference list -- a strong plurality (over 25\%) of our works cited appeared in GENETICS, placing our work squarely within the tradition of GENETICS papers.
-\gc{We apologize for the formatting issues. We have corrected these
-  issues, and included a short METHODS section to introduce our models.} \\
+We apologize for the formatting issues. We have corrected these
+  issues, and included a short METHODS section to introduce our models. \\
 \end{singlespace} 
@@
 We thank you for your previous comments and for consideration of this revised manuscript. 
-To aid in your consideration, we provide \gc{the previous reviews and a
-detailed response below. }
+To aid in your consideration, we provide the previous reviews and a
+detailed response below. 
 \\
@@
 	around our major result -- that sperm are expected to evolve to prevent female meiotic drive. 
-	We completely agree that this delay \gc{cannot} be favored by ``evolutionary foresight'' and have 
+	We completely agree that this delay cannot be favored by ``evolutionary foresight'' and have 
 	modified our manuscript to ensure that readers do not
-        interpret our argument in this way. \gc{We have now added
+        interpret our argument in this way. We have now added
           significantly to our  revised introduction and discussion to
@@
         '98), so that our discussion can be seen as being distinct
-        from that. }
+        from that.
 
@@
 \begin{singlespace}
-	{\bf{Response -- }}  \gc{We believe that this is an important
+	{\bf{Response -- }}  We believe that this is an important
         phenomenon. We have strengthened our discussion of this point
@@
         we have also highlighted the points raised by R2 about 
-        the evidence of pervassive female drive and the opportunities for
+        the evidence of pervasive female drive and the opportunities for
         sperm-based products
-        to interfer with female meiosis.
+        to interfere with female meiosis.
 	We have modified the manuscript to clarify that it is quite impressive that there is evidence
@@
         outside of model organisms. 
-Given the increasing tractibility of genetics in many traditionally
+Given the increasing tractability of genetics in many traditionally
         non-model systems, and the growing awareness of the importance
@@
         interactions and provide
-strong predictions that could guide analyses.}
+strong predictions that could guide analyses.
 \end{singlespace}

### 2014-07-28 writeup_and_notes/GeneticsResub/male_dependent_drive.tex
yb finalized changes for resub 7/27

@@
 	interpreted as mechanisms of protection against genomic outlaws. 
-That females  of many animal species do not complete meiosis until after
-     fertilization appears to run counter to this interpretation
+That females  of many animal species do not complete meiosis until after fertilization, 
+     appears to run counter to this interpretation,
      because this delay provides an opportunity for sperm-acting alleles to
-     meddle with the outcome of female meiosis and help like-alleles drive in heterozygous females. 
+     meddle with the outcome of female meiosis and help like alleles drive in heterozygous females. 
 Contrary to this perceived danger, the population genetic theory presented herein suggests that, 
@@
         meiosis in the face of genomic conflicts. 
-These results are consistent with the apparent sperm-dependence of the best characterized female meiotic drivers in animals to date.  
+These results are consistent with the apparent sperm-dependence of the best characterized female meiotic drivers in animals.   
 Rather than providing an opportunity for sperm collaboration in female meiotic drive,
@@
 An allele that biases female meiosis in its favor (i.e.
-a meiotic driver), may increase in frequency even if it entails a pleiotropic fitness cost \citep{Prout1973}, generating a genetic conflict between the success of the driver and the organism.
+a meiotic driver), can increase in frequency even if it entails a pleiotropic fitness cost \citep{Prout1973}, generating a genetic conflict between the success of the driver and the organism.
 Meiotic drivers observed in both plants
@@
 
-Before doing so, it is worth considering whether there is any evidence that sperm could (or do), influence female meiosis. 
-It is becoming
-        increasingly clear that sperm bring a wide variety of RNA and proteins into the
-        egg \citep{Miller:05}. 
+Before doing so, we highlight the evidence that sperm can (or do), influence female meiosis. 
+It is becoming increasingly clear that sperm bring a wide variety of RNA and proteins into the egg \citep{Miller:05}. 
  Some of these have known functions, for example, in most
-        animal species sperm -- not eggs -- are
+        animal species, sperm -- not eggs -- are
         responsible for the transmission of the centriole, a vital
@@
 Numerous lines of evidence suggest that female meiotic drive is a common and important 
-	evolutionary force, and therefore the opportunity for sperm to influence female drive likely had applications to many animals.   
+	evolutionary force, and therefore the opportunity for sperm to influence female drive is likely relevant to many animals.    
 While research to date has identified a few extreme cases of female meiotic 
@@
 Thus, there is little selective benefit to females in preventing sperm to influence female meioses, 
-	\yb{and in fact, females eschewing this delay would potentially lose sperm assistance in the suppression of meiotic drivers.}  
+	and in fact, females eschewing this delay would potentially lose sperm assistance in the suppression of meiotic drivers.  
 Given the wide-spread requirement of fertilization for the completion
 	of  female meiosis, various features of the interaction between sperm 
-	and egg likely evolved as a mechanism 
-	to enforce an equitable transfer of genetic material.
+	and egg may function to enforce an equitable transfer of genetic material.
 
@@
         heterozygotes (see Figure \ref{Eggsperm_3_allele_cartoon}A2, and
-        Model 4 in the Appendix and File S1 ). 
+        Model 4 in the Appendix and File S1). 
 Naively, $B^{+}$ may spread by capitalizing on the additional drive it causes; however,  
@@
 	three allele system that allows the sperm-acting drive facilitator $B^{+}$ to
-	invade the population (Appendix Model 4, eqn. \eqref{coupling}, and File S2  ). \newline
+	invade the population (Appendix Model 4, eqn. \eqref{coupling}, and File S2). \newline
 
@@
 If this sperm-acting drive suppressor arises on 
-	the non-driving $A$ background ( i.e. in repulsion phase, 
+	the non-driving $A$ background (i.e. in repulsion phase, 
 	creating a third allele $A^{-}$, Figure \ref{Eggsperm_3_allele_cartoon}A4, Model 5), 
@@
 Therefore, all two-locus models of sperm influence on female drive suggest that
-	sperm will evolve to oppose female meiotic drive. 
+	sperm will evolve to oppose female meiotic drive, and can do so as effectively (or more effectively) than female-acting drive modifiers. 
 \newline
@@
 	our results are consistent with the observation that female meiosis 
-	is more fair when fertilized by sperm bearing the drive allele in 
+	is fairer when fertilized by sperm bearing the drive allele in 
 	two of the best described cases of female meiotic drive in 
@@
 	fertilized by heterospecific sperm because such sperm have not evolved to counter novel females meiotic drivers. 
-We further predict that segregation in F1 females backcrossed to parental species will likely be biased with a deficit 
-	of transmission of the paternal species allele from the F1  female.   
-We note that while these predictions follow straightforwardly from the
-	theory presented above, tests of meiotic drive, 
+We further predict that segregation in F1 females backcrossed to parental species will likely be biased, 
+	with a deficit of transmission of the paternal species allele from the F1  female.   
+Tthese predictions follow straightforwardly from the
+	theory presented above; however, we caution that tests of meiotic drive, 
 	and especially sperm-dependent meiotic drive require a high standard
@@
 
-\gc{Our theory also encourages phylogenetic hypotheses concerning the relationship between the opportunity 
+Our theory also encourages phylogenetic hypotheses concerning the relationship between the opportunity 
 	for female meiotic drive and the requirement of fertilization for the completion of female meiosis. 
@@
         lineages may only persist for a short time). 
-Our results also suggests that 
-%Another   hypothesis that our work suggests is we might expect 
-        that phylogenetic variation in the stage of female meiosis when
-        fertilization occurs may influence the prevalence of
-        female meiotic drive.
-For example, in taxa where females complete meiosis
-	before fertilization, centromeric drive may
-	be more common than in species which sperm interact with eggs arrested  in MI, 
-	because in the prior case sperm-based modifiers can only intercede during the $2^\text{nd}$, but not $1^\text{st}$ meiotic division. }
-As a potential test of this hypothesis the speed of centromere
-	turnover could be compared in species in which sperm iterate with eggs paused at MI and MII 
-	(assuming the pace of centromere turnover serves as a proxy for the efficacy of MI drive). 
+Our results also suggests that phylogenetic variation in the stage of female meiosis when
+        fertilization occurs may influence the prevalence of  female meiotic drive.
+For example, centromeric drive may be more common  in taxa where females complete MI before fertilization, 
+	as compared to species in which sperm interact with eggs arrested  in MI, 
+	because in the prior case sperm-based modifiers can only intercede during the second, but not the first  meiotic division. 
+As a potential test of this hypothesis, the speed of centromere
+	turnover could be compared in species in which sperm iteract with eggs paused at MI and MII 
+	(assuming the pace of centromere turnover serves as a proxy for the frequency of MI drivers). 
 \newline
@@
   female meiotic arrest \citep[for a review and evaluation of such
-  hypotheses see][]{Mira1998}.
+  hypotheses see][]{Mira1998}, in fact, we presuppose its existence.
 However, given the existence of female meiotic arrest, 
-	\gc{and that its timing and mechanistic details are variable across species \citep{Masui_book,Karrbook:09} 
+	and that its timing and mechanistic details are variable across species \citep{Masui_book,Karrbook:09} ,
 	the nature of the meiotic arrest and interactions between
         sperm and egg may  be molded by selection to reduce the opportunity for female meiotic drive, 
-	and selfish drivers evolving in concert to overcome these adaptations. }
+	and counteracted by selfish drivers evolving to overcome these adaptations. 
 %leave female meiosis open to sabotage by sperm collaborators, but
@@
   bistability. The colors, and the thin dashed contours, indicate the
-  frequency the allele must be above to invade the population (note
+  frequency the allele must reach, $f^*$ in order to invade the population (note
   that these alleles reach fixation conditional on invading). In the
-  white area the allele cannot invade ($x \geq1$), in the solid red
+  white area, the allele cannot invade, in the solid red
   area the allele can invade and fix when rare. In the left panel we

### 2014-07-28 writeup_and_notes/GeneticsResub/reviews.tex
yb finalized changes for resub 7/27

@@
 The positive and constructive reviews of our previous submission with the same title (\emph{MS ID\#: GENETICS/2014/167064})  where extremely useful in our crafting of this revision. 
-Below we discuss how this paper was improved by addressing, in detail, previous concerns \emph{point-by-point}, but first we address the major  concerns with our previous submission.
+Below we discuss how this paper was improved by addressing, in detail, previous concerns \emph{point-by-point}, but first we address the main points raised by the AE in evaluation of our previous submission.
 \begin{itemize}
 \begin{singlespace}
-\item \emph{{\bf{Major concern 1 -- }} The paper as written establishes that factors expressed in sperm should generally evolve to suppress drivers of meiotic drive in female meiosis.
+\item \emph{{\bf{Point 1 -- }} The paper as written establishes that factors expressed in sperm should generally evolve to suppress drivers of meiotic drive in female meiosis.
 However, as Reviewer 1 points out, this is different from showing that factors expressed in females should evolve to delay completion of meiosis until after fertilization.
@@
 \begin{singlespace}
-\item \emph{{\bf{Major concern 2 -- }} The format of this submission departs from the style of GENETICS (model description and analysis, literature citations, etc.).
+\item \emph{{\bf{Point 2 -- }} The format of this submission departs from the style of GENETICS (model description and analysis, literature citations, etc.).
 As a result, it is unclear whether GENETICS is the appropriate or even the intended forum for this work.} \newline \\
@@
 %%I'd not use quotes right below the quoted text, seems trite.
-{\bf{Response:}} \gc{We appreciate that the Reviewer enjoyed the paper.}
+{\bf{Response:}} We appreciate that the Reviewer enjoyed the paper. 
 We do however insist that this is a sufficient advance for publication in GENETICS --
@@
         hypotheses proposed to explain female meiotic arrest (Mira
-        '98), so that our discussion can be seen as being distinct
-        from that.
+        '98), and emphasize that our model presupposes meiotic arrest, 
+        so that our discussion can be seen as being distinct from that.
 
@@
 	We hope that new theoretical work (including ours) coupled
-        with the increasing easy of genetics in many system, will encourage the discovery of additional drive systems in non-model organisms in the coming years.
+        with the increasing ease of genetics in many non-model systems, 
+        will encourage the discovery of additional drive systems in non-model organisms in the coming years.
 \end{singlespace}

### 2014-07-28 writeup_and_notes/GeneticsResub/male_dependent_drive.tex
yb minor changes for resub 7/27

@@
 maternal  genomes' are often aligned, 
-as both are invested in the fate of the resultant zygote (as was speculated for the \emph{In} locus \citep{Pomiankowski1993}).
+as both are invested in the fate of the resultant zygote \citep[as was speculated for the \emph{In} locus ][]{Pomiankowski1993}.
 Thus, there is little selective benefit to females in preventing sperm to influence female meioses, 
@@
 	can succeed evolutionarily even if they lower organismal fitness. 
-This generates numerous conflicts  \cite{Burt2006} including sexual conflicts between mates \cite{Arnqvist2005}, 
+This generates numerous conflicts  \citep{Burt2006} including sexual conflicts between mates \citep{Arnqvist2005}, 
 	and conflicts between alleles that are over-transmitted in meiosis and the organisms they injure while doing so. 
@@
 	two of the best described cases of female meiotic drive in 
-	animals (the \emph{Om} and \emph{In} loci in mice \citep{Agulnik1993,Wu2005}).  
+	animals \citep[the \emph{Om} and \emph{In} loci in mice][]{Agulnik1993,Wu2005}.  
 Given the benefit to sperm of hampering female drive, 
@@
 
-{\bf{Files S2A \& S2B:}} A Mathematica file (FileS2A) and a PDF of this file (File S2B) in which we derive analytical results for models 1-6.
+{\bf{Files S2A \& S2B:}} A Mathematica file (FileS2A) and a PDF of this file (File S2B) in which we derive analytical results for models 1-6 and $6^\prime$.
 

### 2014-07-28 writeup_and_notes/GeneticsResub/male_dependent_drive.tex
fixed couple of citations

@@
 maternal  genomes' are often aligned, 
-as both are invested in the fate of the resultant zygote \citep[as was speculated for the \emph{In} locus ][]{Pomiankowski1993}.
+as both are invested in the fate of the resultant zygote \citep[as was speculated for the \emph{In} locus, ][]{Pomiankowski1993}.
 Thus, there is little selective benefit to females in preventing sperm to influence female meioses, 
@@
 	can succeed evolutionarily even if they lower organismal fitness. 
-This generates numerous conflicts  \citep{Burt2006} including sexual conflicts between mates \citep{Arnqvist2005}, 
-	and conflicts between alleles that are over-transmitted in meiosis and the organisms they injure while doing so. 
+This generates numerous conflicts including sexual conflicts between mates \citep{Arnqvist2005}, 
+	and conflicts between alleles that are over-transmitted in
+        meiosis and the organisms they injure while doing so  \citep{Burt2006}. 
 Such conflicts and their resolution likely play a major role in the structure and evolution of many basic biological processes \citep{Rice2013}. \newline
@@
 	two of the best described cases of female meiotic drive in 
-	animals \citep[the \emph{Om} and \emph{In} loci in mice][]{Agulnik1993,Wu2005}.  
+	animals \citep[the \emph{Om} and \emph{In} loci in mice, ][]{Agulnik1993,Wu2005}.  
 Given the benefit to sperm of hampering female drive, 

### 2014-07-28 writeup_and_notes/GeneticsResub/male_dependent_drive.tex
yb minor changes for resub 7/28 (again)

@@
         opportunities to test this hypothesis may be limited because 
-        %this effect may be weaks as such
         lineages may only persist for a short time). 
@@
 	as compared to species in which sperm interact with eggs arrested  in MI, 
-	because in the prior case sperm-based modifiers can only intercede during the second, but not the first  meiotic division. 
+	because in the prior case, sperm-based modifiers can only intercede during the second, but not the first  meiotic division. 
 As a potential test of this hypothesis, the speed of centromere
-	turnover could be compared in species in which sperm iteract with eggs paused at MI and MII 
+	turnover could be compared in species in which sperm interact with eggs paused at MI and MII 
 	(assuming the pace of centromere turnover serves as a proxy for the frequency of MI drivers). 

### 2014-07-28 writeup_and_notes/GeneticsResub/reviews.tex
yb minor changes for response on  7/28 (again)

@@
 \begin{singlespace}
-I did not feel that the authors made a convincing case that this is an
+\emph{I did not feel that the authors made a convincing case that this is an
 important phenomenon in its own right.  P. 5 has one short paragraph pointing to
@@
 evidence it is hard to know whether this is just an interesting idea that never
-really happens. \\
+really happens.} \\
 \end{singlespace}
@@
 	In general, the spread of an unlinked female drive suppressor is qualitatively similar to that of an unlinked sperm-acting drive suppressor; 
-	however, the initial spread of the sperm suppressor is equal to (when the female suppressor is dominant), or more rapid than (otherwise) the female suppressor. 
+	however, the initial spread of the sperm suppressor is equal to (when the female suppressor is dominant), or more rapid (otherwise) than the female suppressor. 
 	Given the difference in mechanism of female drive suppression in sperm as compared to females, sperm suppression of female drive is likely not functionally redundant with female drive suppression and provides a qualitatively different mechanism of suppression.
@@
 main body of the text that tersely describes the 6 models? \\ \\}
-{\bf{Response -- }} We have added a Methods section, which introduces these six models, and makes our manuscript more in line with the GENETICS format.
+{\bf{Response -- }} We have added a Methods section, which introduces these six models (+ model 6$^\prime$), and makes our manuscript more in line with the GENETICS format.
 \end{singlespace}

### 2014-11-17 writeup_and_notes/evolution/response_drive_MS.tex
added a new folder [evolution] with a draft of our evolution resubmision and response to reviewers

%%!TEX TS-program = latex
\documentclass[12pt,letterpaper]{article}
\bibliographystyle{evolution}
\usepackage{epsfig}                 
\usepackage[authoryear]{natbib}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{psfrag}
\usepackage{mathabx}
\renewcommand{\baselinestretch}{1.6}
\large
\pagenumbering{arabic}
\usepackage[usenames,dvipsnames]{color}
\usepackage{fullpage}
\usepackage{multirow}
\newcommand{\ye}{\hat{y}}
\newcommand{\xe}{\hat{x}}
\usepackage{color}
\usepackage[normalem]{ulem}  
	\newcommand{\gc}[1]{{ \color{red} #1}}
	\newcommand{\yb}[1]{{ \color{blue} #1}}
\usepackage{lineno}
%\linenumbers 
\begin{document}
Dear EVOLUTION editorial team,
\\

Thank you for your consideration of our manuscript 14-0566 entitled "Sperm should evolve to make female meiosis fair." We appreciate the thorough comments by the reviewers and Associate Editor.
We are submitting a new version of this manuscript in which we have incorporated reviewer comments as advised by the AE. Our revision includes versions of Models 1 and 2 with some degree of inbreeding, and has been edited to ensure clarity, and minimize the possibility of misleading our readers. We believe that this has resulted in a stronger manuscript. Below we describe in greater detail, and a point-by-point basis how we responded to specific comments of the AE and the reviewers.
\\
\\
Sincerely \\
Yaniv Brandvain and Graham Coop
\newpage

{\bf{Directive from Editor}} \\ \\
Montpellier, Nov. 6th 2014,
\\ \\
Dear Mr. Brandvain: 
I am pleased to tell you that your manuscript 14-0566 entitled "Sperm should evolve to make female meiosis fair." will be suitable for publication in Evolution pending some additional, minor revision. We appreciate your submission as well as your willingness to respond to the constructive suggestions of the reviewers and Associate Editor, Dr. Jon Wilkins.

You will find the comments from the Associate Editor and the Reviewers below. I invite you to respond to these and revise your manuscript accordingly. If your responses require additions to the text of the manuscript, please take care not to exceed our manuscript length guidelines, as these apply to the final, accepted version of the manuscript. \dots{}

Once again, thank you for submitting your manuscript to Evolution.  I look forward to the published paper.
\\  \\
Sincerely, \\
Dr. Thomas Lenormand \\
Editor, Evolution \\
thomas.lenormand@cefe.cnrs.fr

\newpage

{\bf{Comments from the Associate Editor}} -- \emph{Authors respond in italics}  \\

Thanks for the interesting paper. As both reviewers note, it is an interesting and important topic. I also found the biological framing compelling, and I think it should penetrate deeper into the non-theory part of the Evolution audience than many theory papers do.

Let me apologize for the slowness of the review process here. I had a heck of a time getting reviewers to agree to this one. I don't know. Maybe they somehow saw that the combined PDF was 97 pages and got scared off. Or maybe when they read the abstract and saw the umlauts on oogenesis they assumed that this was actually a mis-addressed New Yorker submission.
\\ \\
\emph{We appreciate your feedback and efforts to co\"{o}rdinate reviewers}
\\

You've got some nice reviewer comments here, which should definitely be addressed. In particular, reviewer 2, aka Gardner, made two suggestions that I see as critical.
\\ \\
{\bf{AE comment 1:}} First is the suggestion to consider inbreeding. Personally, I'm not sure it would be worth the effort at this point to write out and analyze a full-on model of a structured population with migration, etc, etc. It's a fantastic idea, and definitely worth doing, but maybe it's the next paper.

However, it would be easy to tweak each model with the inclusion of one or two F values, to give you excess homozygosity (or excess probability of green-beard matching). For example, in model 3, you mention this narrow region of parameter space that exists only when the paternal modifier functions similarly in sperm from heterozygous and homozygous fathers. One thing that inbreeding would do is increase the likelihood of occupying that space. Would that then make a sustained polymorphism biologically plausible? If so, what would that predict for the evolution of female meiosis in highly inbred populations?

{\bf{Response to AE comment 1:}}  \emph{We note that both R1 and R2 noted an interest in the case of inbreeding. \yb{We therefore have explored the behavior of model 2 [as well as model one for reference to the case of traditional drive] under six values of inbreeding. We briefly discuss these results in the modified text and make the figures [analogous to Figure 1B] in the Supplement.}}
\\
\\
\\
{\bf{AE comment 2:}} Second is the taxonomic question. The phylogenetic distribution of post-fertilization completion of female meiosis is critical to the evolutionary implications of this work. Even if the data is spotty at the moment, it would be worth spelling out what is known. Plus, you could maybe make your phylogenetic hypotheses a bit more specific that way.

{\bf{Response to AE comment 2:}}  \emph{WE NEED TO DO SOMETHING HERE \dots{}}
\\
\\
\\
{\bf{AE comment 3:}} 
A couple of comments regarding the discussion section:

In general, the discussion section is a little bit fast and loose with the evolutionary reasoning, and it is not always entirely clear what is being claimed. I get that you're trying to go for a readable voice here, to spell out the take-home message for the reader who only skimmed the results. However, I think this would benefit from thinking carefully about what the four or five key points are, and dedicating a paragraph to each.

{\bf{Response to AE comment 3:}}  \emph{We have included clear `sign-posts' throughout the discussion, and have more fully articulated the logic underlying our predictions and conclusions.}
\\
\\
\\
{\bf{AE comment 4:}} 
You mention at the end that you're not modeling female meiotic arrest. However, earlier in the discussion (ca. lines 309 \& 349) and at the end of the introduction (line 109), you make a couple of statements that border on the teleological. The current text borders on suggesting that delayed female meiosis could be favored because it permits the sperm to participate in the suppression of female meiotic drivers.

Now, that's a cool idea, but it suffers from the same limitation as evolvability arguments: it poses an adaptive benefit in terms of future selective events. It can work, but you have to be careful about the conditions. For example, either drivers would have to crop up often enough that the transient fitness reductions create a substantial fitness advantage -- or that there is a persistent fitness cost after fixation. And maybe it requires (or would benefit from) some higher-level selection.

Plus, you can flip the just-so argument around. We might expect that in species without post-fertilization completion of female meiosis, this is a consequence of female drivers that have accelerated meiosis in order to prevent interference by sperm modifiers.

Basically, there's a bunch of stuff here that would need to be explicitly modeled. So, just make sure all the appropriate caveats are in place.

{\bf{Response to AE comment 4:}}  \emph{We agree and have further clarified this point. This includes pointing out the circularity in logic of a simple version of the claim that female meiotic arrest evolves to allow sperm to combat female drivers. We do however, maintain our 3rd prediction [line 249 in our first submission] that reducing opportunities for drive could allow organisms to skip the fertilization requirement. We point out that this is a question of evolutionary maintenance rather than of evolution origin.}
\\
\\
\\
{\bf{AE comment 5:}}
More compelling in the context of the present work is the line of reasoning followed from line 354 onward, that the form of meiosis will affect the prevalence of female drivers of different sorts. I would actually like to see this expanded on a bit more.

The Om and In bit is nice, and a key piece of evidence supporting the applicability of your analysis. I do have a question regarding the predictions on crosses though. It seems that the prediction only holds to the extent that there are drivers that are in the process of being suppressed, which depends on the frequency with which new drivers arise and the rate at which they get suppressed (and presumably eventually eliminated). However, it seems like you might also expect to see the effect if there is some divergence between versions of the same drivers in different populations. Intuitively, if there is sperm-mediated suppression going on, there might be opportunities for antagonistic coevolution, with drivers selected to evade suppression (depending on the molecular mechanisms of drive and suppression, of course). This would make a similar set of predictions, I think, but would not hinge on a new driver having arisen in one of the two.

If the coevolution argument holds water, you would also expect to see faster evolutionary divergence of drive elements in species with post-fertilization meiosis. However, it might be hard to establish apples-to-apples comparisons to test that sort of thing.

{\bf{Response to AE comment 5:}} \emph{We agree with this ida, and have incorporated it into our discussion.}
\\
\\
\\
{\bf{AE minor comments:}}
\\
A couple of minor things
\\
throughout: hyphenate compound adjectives, e.g. "sperm-acting modifier" and "drive-suppressing allele"
\\
line 315: "outweighs"
\\
line 330: you forgot the gratuitous umlaut on "naive"
\\
line 372: extra space after )

{\bf{Response to AE minor comments:}} \emph{We fixed these, thanks!}

\newpage
Reviewer(s)' Comments to Author:
\\
Reviewer: 1
\\
Comments to the Author
\\Sperm favor fair female meiosis.

I enjoyed reading this interesting paper. The question is clearly one where intuition fails because of the complexity of the interacting factors and the authors? models clarify these interactions.
\\
\\
\\
{\bf{R1 comment 1:}}
Models 1-3 cover the situation in which drive in female meiosis depends on sperm (padumnal) genotype or male (paternal) genotype. The models find that the invasion of such a system of drive is unlikely because in a random-mating population drive is very rare because both sexual partners need to carry the rare allele and because drive, when it does occur, creates driving homozygotes which are assumed to have low viability and which cannot themselves undergo drive. If there is no cost to homozygotes the ?driver? should go to fixation slowly at first. Forms of inbreeding would probably also facilitate initial invasion.

{\bf{Response to R1 comment 1:}} \emph{We have added inbreeding to our analyses of models 1-2, see response to AE comment 1}
\\
\\
\\
{\bf{R1 comment 2:}}
Models 4-5 cover the situation in which an allele drives in female meiosis and this is modified by sperm genotype at the same locus (by assumption of tight linkage). Model 6 considers an unlinked sperm modifier. The interesting case in Model 4 is the suppressor allele B-. This allele initially does not suppress its own drive (because rare) but suppresses the drive of B and thus ends up in A/B- heterozygotes rather than B/B- heterozygotes. The authors say ``avoid the excess formation of low fitness homozygotes'' but I think this is slightly misleading because B and B- are different ?alleles?. B- displaces B and often comes to a lower equilibrium frequency presumably because of the effects of suppressing its own drive
(B- causes the production of higher fitness A/B- heterozygotes rather than lower fitness B-/B- homozygotes. If the difference in fitness is not too great then B- ?loses? from suppressing its own drive).

{\bf{Response to R1 comment 2:}} \emph{We have changed this section to (e.g. changing ``avoid the excess formation of low fitness homozygotes'' to ``avoid the excess formation of low fitness genotypes'') in accordance with these suggestions.}
\\
\\
\\
{\bf{R2 minor comments:}}
\\ line 272 ``so it avoid[s]'' line 274 ``the the''
\\
The take home message is that sperm alleles usually evolve to suppress female meiotic drive, whether linked (in coupling or repulsion) or unlinked to the driving locus.
\\
line 339 ``novel females meiotic drivers'' rather ``novel female meiotic drivers'' or ``novel females' meiotic drivers''
\\
line 342 ``Tthese''

{\bf{Response to R2 minor comments:}} \emph{We fixed these, thanks!}

\newpage{}

Reviewer: 2

Comments to the Author

{\bf{R2 summary:}}
This study raises an interesting question: are sperm favoured to meddle in female meiosis and, if so, what are they favoured to do? This question is motivated by the very interesting observation that, in many animals, female meiosis takes place after fertilization, giving the sperm some possible role in deciding the fate of meiotic drivers. The authors develop some mathematical models, considering cases in which sperm promote drive of matching alleles only or promote or suppress drive in general. The overall message coming from the models is that sperm do not act to promote drive, and can be favoured to suppress it.

Although I found the motivating question and background biology really interesting, I am not really convinced by the analysis.

{\bf{Response to R2 summary:}} \emph{We appreciate that you find the question, the biology, and the results compelling, and are disappointed that you are unconvinced by the analysis. However, we can find no statement in your specific comments below that suggest that you are unconvinced. We therefore respond to the comments below, and hope that the responses to your specific comments are convincing.}
\\
\\
\\
Some comments:
\\
{\bf{R2 comment 1 :}}
(1) The results are pretty obvious a priori, and don?t really warrant the development and analysis of explicit models. As the authors note in the Introduction, the matching-alleles models are instances of greenbeard genes, and we already know that these are under positively frequency-dependent selection, and cannot invade from rarity in unstructured populations if there is any obligate cost involved, such as the cost of the machinery that does the allelic recognition. (This link also needs to be made in the Results and Discussion sections.) And in the scenarios where sperm are deciding the fate of drivers in an egg to which they have no genetic relatedness, it is obvious that their only concern will be to maximize the viability of the ensuing zygote, which means suppressing drive.

{\bf{Response to R2 comment 1:}} \emph{We believe that intuition without rigorous theoretical support can be dangerous. Therefore, while the results may be intuitive to R2, they were not to ourselves or to R1. We therefore believe that our results are not trivial and represent an important contribution. The non-triviality of our results are apparent in this comment -- in our model drive presents no cost when extremely rear, and there is some amount of cost for which a sperm acting drive promoter can invade and fix. The complexity of our analysis [and the bit that can choke simple intuition] is the competition between two forms of frequency dependence in the costs and benefits of ``green-bearded drive''.}
\\
\\
\\
{\bf{R2 comment 2:}} (2) Explicit models would be a lot more useful if the scope of the study were extended to inbred populations, where there are more selective pressures at play and where it is less obvious to see what will be favoured. Here, driving genes would be a lot more likely to encounter identical-by-descent copies in sperm, which could act to promote collaboration on the part of the sperm (analogous to the boost that greenbeards get in structured populations, owing to increased rate of encounters). But drivers would also suffer from the reduced frequency of heterozygotes under inbreeding, and this would tend to inhibit drive. And there may be additional pressures owing to inbreeding favouring bias sex ratios, which driving sex chromosomes or, indeed, whole genomes could be important in bringing about, under female heterogamety e.g. see Gardner \& Ross (2014, ?Mating ecology explains patterns of genome elimination?, Ecol Lett, doi: 10.1111/ele.12383). An explicit model would be doing really useful work here, showing which of these effects is more important, and why. The present results for the outbreeding case, that seem a bit too trivial to stand on their own, would usefully serve to prepare the reader for more substantial results for inbreeding scenarios.

{\bf{Response to R2 comment 2:}} \emph{We agree that the case of inbreeding could be interesting. In line with comments from the AE and R1, we implement a simple model of inbreeding [see response to AE comment 1]; however, as stated by R1, completely constructing a detailed inbreeding model is beyond the scope of this ms and is best left to future work. }

\emph{As stated above, we disagree with the notion that a model which confirms the intuition of R2 is trivial. We believe that clearly and rigorously presenting novel predictions and highlighting relevant biology is a substantial contribution. We are also somewhat concerned that chasing special scenarios in the hopes of unintuitive results could district from our major and clear finding, and therefore limited our analysis of the case of inbreeding.}
\\
\\
\\
{\bf{R2 comment 3:}}  (3) I really enjoyed the biology discussed at the beginning of the Introduction. I think this could be fleshed out a lot more. For example, when you say ?in many animal species? (line 46), it would be useful to explain this more fully, perhaps with the help of a figure giving a taxonomic overview. Is post-fertilization female meiosis monophyletic, have there been reversals, are there obvious ecological or other correlates, and so on?

{\bf{Response to R2 comment 3:}}  \emph{\yb{NEED TO WORK ON THIS}}
\\
\\
\\
\\
As is my policy, I waive anonymity
\\
Andy Gardner

\end{document}

### 2014-11-17 writeup_and_notes/evolution/RESUBMISSIONmale_dependent_drive.tex
added a new folder [evolution] with a draft of our evolution resubmision and response to reviewers

--- 2014-07-28 writeup_and_notes/GeneticsResub/male_dependent_drive.tex
@@
 \documentclass[12pt,letterpaper]{article}
-\bibliographystyle{genetics}
+\bibliographystyle{evolution}
 \usepackage{epsfig}                 
@@
 	\newcommand{\yb}[1]{{ \color{blue} #1}}
+\usepackage{lineno}
+\linenumbers
 
@@
 \newpage
-\flushleft{{\bf{Running title:}} Sperm prevent female meiotic drive} \\
-{\bf{Keywords:}} Meiosis; Meiotic drive; Fertilization; Genomic conflict \\
-{\bf{Corresponding author:}} Yaniv Brandvain. \emph{Address:} Department of Plant Biology, University of Minnesota --€“ Twin
-Cities. St. Paul MN, 55108. \emph{Phone:} 612-624-4375 \emph{Email:} ybrandvain@gmail.com
-
-\newpage
 
@@
 O\"{o}genesis is inherently susceptible to such conflicts
-because alleles compete for inclusion in the  product of female meiosis that will be transmitted to the egg. 
+because alleles compete for inclusion into the egg. 
 Alleles that distort meiosis in their favor (i.e. meiotic drivers) often decrease organismal fitness,
@@
         meiosis in the face of genomic conflicts. 
-These results are consistent with the apparent sperm-dependence of the best characterized female meiotic drivers in animals.   
+These results are consistent with the apparent sperm dependence of the best characterized female meiotic drivers in animals.   
 Rather than providing an opportunity for sperm collaboration in female meiotic drive,
@@
         suppress drive. 
-%	and suggest that the requirement of fertilization for the
-%%        completion of female meiosis
-%        potentially represents a mechanism employed by females 
-%	 to ensure a fair meiosis.
 \newpage
@@
 	likely precludes this interaction in plants). 
-Therefore, a `green-bearded' \citep{Gardner2010} sperm-acting allele that recognizes and facilitates the meiotic drive of a genetically equivalent allele in heterozygous females %female meiosis %(a  heteromorphic dyad),
+\yb{INSERT BIOLOGY HERE}
+Therefore, a `green-bearded' \citep{Gardner2010} sperm-acting allele that recognizes and facilitates the meiotic drive of a genetically equivalent allele in heterozygous females 
      could presumably rapidly spread through a population. 
@@
 	suggests that sperm influence on female meiotic drive is not only possible, but likely. 
-%Evidence from both the \emph{In} and \emph{Om}  -- suggests that the 
-%	 transmission of alternate alleles in heterozygous females can depend on sperm haplotype.  
-%There are current only two polymorphic, well-characterized female meiotic drive systems in
-%animals-- and both of them, the \emph{In} and \emph{Om}  drive systems in mice-- suggests that the 
-%	 transmission of alternate alleles in heterozygous females can depend on sperm haplotype.  
  While ruling out the alternative hypothesis of early selection on zygotes in these cases is challenging  \citep[see pages 52-54 in][ for comment]{Burt2006}, it appears that the extent to which \emph{In} and \emph{Om} distort the second female meiotic division partially depends on the
@@
          range of systems will reveal many more female meiotic drive
-         systems, however, carefully characterizing them will still
+         systems; however, carefully characterizing them will still
          remain a challenging task. 
 \newline
-
-%The stage at which female meiosis requires fertilization varies widely across taxa, and in a few animals  
-%	female meiosis is complete upon ovulation (that is, fertilization is not required for female meiosis \citep[e.g. Table 1 in 
-%][]{Masui_book}). 
-%It is also not the case that sperm is mechanistically incapable of influencing female meiosis.
-%In \emph{C. elegans}, premature deployment of the aster -- a vital component of mitotic
-%machinery provided by the sperm, disrupts MII meiotic segregation
-%in the egg, leading to a triploid zygote \citep{McNally2012}. 
-%Thus, there is direct mechanistic evidence that sperm can influence female meiosis.
-% Additionally, genetic evidence from the \emph{In} and \emph{Om}  drive systems in mice suggests that the 
-% transmission of alternate alleles in heterozygous females can depend on sperm haplotype.  
-% While ruling out the alternative hypothesis of early selection on zygotes in these cases is challenging  \citep[see pages
-%52-54 in][]{Burt2006}, it appears that the extent to which \emph{In} and \emph{Om} distort the second female meiotic division partially depends on the genotype of the fertilizing sperm \citep{Agulnik1993,Wu2005}.
 
@@
 	and in fact, females eschewing this delay would potentially lose sperm assistance in the suppression of meiotic drivers.  
-Given the wide-spread requirement of fertilization for the completion
+\yb{Given the wide-spread requirement of fertilization for the completion
 	of  female meiosis, various features of the interaction between sperm 
-	and egg may function to enforce an equitable transfer of genetic material.
-
-%%don't thi k this helped: 
-%\yb{In fact, a mutant female which completes meiosis before 
- %	fertilization is giving up the protection from meiotic drivers potentially offered by sperm, 
-%	and is therefore more likely to produce low fitness offspring. }
-%%\gc{We do not claim that female meiotic arrest evolved
-%  specifically to recruit helpful sperm to defend against meiotic
-%  drive, the delayed completion of female meiotic drive will not in
-%  fact encourage sperm to assist renegade drivers. Rather, we think
-%  that features of
-%  o\"{o}genesis  may evolve to facilitate an interaction with sperm 
-%with female meiosis
-%as yet another mechanism to enforce an equitable transfer of genetic material.}
+	and egg may result in an equitable transfer of genetic material -- 
+	wether this result is the ultimate evolutionary function of the fertilization requirement or a coincidental pleiotropic outcome is beyond the scope of this manuscript, 
+	but our intuition argues against the prior (see Discussion).}
 
@@
  	at the drive locus such that recombination is unexpected. 
-To evaluate the feasibility of sperm-modification of female meiotic drive, as compared to female suppression of drive, 
-	we conclude with a model of female drive suppression by an unlinked female acting suppressor ({\bf{Model $6^\prime$}}). 
+To evaluate the feasibility of sperm modification of female meiotic drive, as compared to female suppression of drive, 
+	we conclude with a model of female drive suppression by an unlinked female-acting suppressor ({\bf{Model $6^\prime$}}). 
 In all cases, we assume that fitness is independent of the drive modifier. 
@@
 	and to our Mathematica worksheet  (File S2) for our complete derivations. 
-There we present critical analytical results in Equations 1-11, and describe our analyses and results in more detail.  
+There, we present critical analytical results in Equations 1-11, and describe our analyses and results in more detail.  
 Because a number of our analyses are approximations based on assuming
-that genotype frequencies follow Hardy Weinberg Equilibrium (HWE), we note which analytical results are approximations. 
+that genotype frequencies follow Hardy Weinberg Equilibrium (HWE), we note which analytical results are approximate. 
 We verify these approximations with exact numerical iterations in Figures S1-S3 and File S3.
@@
 
-\begin{figure}
-\begin{center}
-\includegraphics[width = 1 \textwidth]{../Figures/selfprom.eps} \end{center}
-\caption{{\bf{A.}} A visual depiction of our model of `self-promoting' driver. 
-	Transmission probabilities for alleles through female
- 	 meiosis depend on sperm genotype. 
-	 The non-driving $A$-allele, and self-promoting $B$-allele
-	 are represented by unfilled and filled circles, respectively. 
-	 {\bf{B.}} Evolution of a self-promoter and standard driver. 
-	 	Assuming that the fitnesses of drive homozygotes and 
-		heterozygotes are $1-s$ and $1$, respectively. 
-	 	\emph{Main figure:} Boundary conditions for the invasion 
-		and fixation of self-promoting and standard meiotic drivers, 
-		with drive coefficient, $d$. Colored regions depict exact results, 
-		while lines  represent analytical approximations. 
-	 \emph{B1:}  Trajectories of sperm-dependent female drive
-		 %alleles that can invade the population, 
-		 each allele has $s =0.1$ against the
-		 homozygotes.  The drive coefficient is denoted by color. 
- }   
-\label{Eggsperm}
-\end{figure}
+%%%FIG!
 
@@
 Subsequently, a sperm-acting modifier of female meiotic drive arises at another locus.  
-In these two-locus models, the driver is transmitted to $d_0$ of gametes from female heterozygotes when fertilized by wild-type sperm, and $d_1=d_0+\epsilon$ when fertilized by a sperm-acting drive-modifier.
+In these two-locus models, the driver is transmitted to $d_0$ of gametes from female heterozygotes when fertilized by wild-type sperm, and $d_1=d_0+\epsilon$ when fertilized by a sperm-acting drive modifier.
 \newline
@@
         Model 4 in the Appendix and File S1). 
-Naively, $B^{+}$ may spread by capitalizing on the additional drive it causes; however,  
+Na\"{i}vely, $B^{+}$ may spread by capitalizing on the additional drive it causes; however,  
 	this is not the case for a few simple reasons. 
@@
 	is at drive-selection balance, 
-	and therefore immediately suffers a homozygous fitness cost.  
+	and therefore immediately suffers a genotypic fitness cost equivalent to the BB homozygote.  
 Worse yet, a novel $B^{+}$ haplotype most often helps 
@@
 
-\begin{figure}
-\begin{center}
-	\includegraphics[width = 0.7 \textwidth]{../Figures/twolocus.eps} 
-          \end{center}
-\caption{Models of a sperm-acting drive modifier tightly linked to a meiotic driver. 
-{\bf (A) } 
-Sperm carrying the derived allele at the modifier locus (filled squares) alters transmission  
-	at the driving allele (filled circles) during female meiosis.
-        Alleles at these two tightly linked loci form three haplotypes (top of A). 	
-	{\bf{A1)}} In the standard model of drive there is no
-        variation at the modifier, 
-	and the driver is transmitted to the egg with probability $d_0$. 
-	{\bf{A2)}} The modifier allele increases the transmission of the drive allele ($d_1 > d_0$), and due to their shared genetic background, also increases its drive. 
-	{\bf{A3 \& A4)}} The sperm-acting modifier acts to decrease
-        drive  ($d_1=1/2$ in A3 \& A4, or more generally, $d_1 < d_0$) and arises on the same or opposite
-        background from the driver (A3 \& A4 respectively). 
-{\bf{(B)}} Invasion of a sperm-acting drive suppressor linked to a driver. 
-	After the driver ($B$ haplotype) reaches drive selection equilibrium, we introduce a sperm acting drive modifier. 
-	We assume full drive ($D_0 =1$), a recessive lethal fitness cost to drive ($h_s=0$, $s=1$) and that the sperm-acting modifier results in a fair meiosis. 
-{\bf {B1)}} The $B^-$ allele replaces the ancestral drive haplotype, but segregates at a lower equilibrium frequency.
-{\bf {B2)}} The $A^-$ allele replaces the ancestral non-driving haplotype, and in this case, removes the driver from the population. }
- \label{Eggsperm_3_allele_cartoon}
-\end{figure}
+%%%%%%%FIG2
 
@@
        it does not drive when fertilized by another $B^{-}$ so it
-       avoid the excess formation of low fitness homozygotes. This
+       avoids the excess formation of low fitness genotypes. This
      allows the $B^{-}$ to displace the ancestral driver (Figure
      \ref{Eggsperm_3_allele_cartoon}B1, Equation \ref{coupling}), 
-	and often returns to a lower equilibrium frequency than the the $B$ allele, further decreasing the extent of drive in the population.  
+	and often returns to a lower equilibrium frequency than the $B$ allele (likely because it surprises its own drive), further decreasing the extent of drive in the population.  
 If this sperm-acting drive suppressor arises on 
@@
 
-The sperm acting drive suppressing allele lowers the frequency of the original driver (perhaps to zero),
+The sperm-acting drive suppressing allele lowers the frequency of the original driver (perhaps to zero),
 	and spreads to fixation if it does not carry strong fitness costs
@@
 
+\subsection*{Major result: Sperm evolve to enforce fairness in female meiosis.} 
 It seems that allowing sperm to influence the outcome of female meiosis would generate a confluence of these potential conflicts -- 
@@
 	without risking exploitation by collaborations between female
-        drivers and sperm alleles. \newline 
+        drivers and sperm alleles. 
+\yb{Although the fertilization requirement allows sperm an opportunity to enforce fairness in female meiosis, 
+	this is unlikely it evolutionary raison d'\^{e}tre. 
+In fact, to suggest so, presupposes that sperm have an evolved system, to prevent meiotic drive before they have a mechanism to do so.}
+\newline 
 %%G: felt like this made it sound less why they do it.
 
+\subsection*{Explaining why sperm evolve to enforce fairness in female meiosis.} 
 Why is it that an allele that biases female meiosis in its favor can
@@
 but an allele in sperm that assists this female driver cannot? 
-So long as the transmission advantage of female meiotic drive outweigh
+So long as the transmission advantage of female meiotic drive outweighs 
 the organismal fitness cost to heterozygotes, 
@@
 
-%Our theory offers a novel interpretation of a baffling idiosyncrasy of female meiosis in many animals -- the requirement of fertilization before meiosis is complete. 
-%We argue that females do not complete meiosis until after fertilization as a mechanism to encourage meiotic fairness.  
-%In addition to providing a new interpretation of a basic biological process, our theory provides numerous testable predictions.
+\subsection*{Predictions from theory.}
 
-The theory presented herein provides a strong prediction -- 
-	for organisms in which female  meiosis is not completed until after fertilization, 
+\yb{The theory developed above has one overarching conclusion -- that when possible, males evolve to make female meiosis fair. 
+This simple result provides numerous novel predictions, many of which are directly testable.
+
+Our most direct prediction is that for organisms in which female  meiosis is not completed until after fertilization, 
 	sperm will act to suppress female drive at the stage at which they can influence meiosis. 
-%Therefore, for example, when fertilized by drying sperm, we expect meiosis to be fairer when  females are fertilized by 
-Contrary to expectations of a naive verbal ``green-beard'' model, 
-	our results are consistent with the observation that female meiosis 
+This prediction, which holds when modifier and driver are the same gene (Model 2) or are in tight linkage (Model 4),  
+	is strongly supported by the observation that female meiosis 
 	is fairer when fertilized by sperm bearing the drive allele in 
@@
 	animals \citep[the \emph{Om} and \emph{In} loci in mice, ][]{Agulnik1993,Wu2005}.  
+Both this prediction, and the empirical support for it run contrary to expectations of a na\"{i}ve verbal ``green-beard'' model. 
+}
+
+\yb{Our model of a sperm-acting drive suppressor unlinked to a female driver (Model 6) also predicts that sperm should evolve to prevent meiotic drive; 
+	however, it contains no simple mechanism to maintain polymorphism for sperm-acting drive suppression.
 Given the benefit to sperm of hampering female drive, 
@@
 	specifically we predict that female meiosis will be less fair when a species (or population) is 
-	fertilized by heterospecific sperm because such sperm have not evolved to counter novel females meiotic drivers. 
-We further predict that segregation in F1 females backcrossed to parental species will likely be biased, 
+	fertilized by heterospecific sperm because either such sperm have not evolved to counter novel female meiotic drivers, 
+	or because antagonistic coevolution between a driver-suppressor pair has been independent since two populations have separated. 
+We can therefore predict that segregation in F1 females backcrossed to parental species will likely be biased, 
 	with a deficit of transmission of the paternal species allele from the F1  female.   
-Tthese predictions follow straightforwardly from the
+These predictions follow straightforwardly from the
 	theory presented above; however, we caution that tests of meiotic drive, 
@@
 	inviability including epistatic maternal X zygotic lethality \citep[e.g. ][]{Sawamura:1993aa}.  \newline
+}
 
@@
 	for female meiotic drive and the requirement of fertilization for the completion of female meiosis.
+
+\yb{
 For example, we predict that a lower opportunity for female meiotic drive,
-        e.g. an animal lineage with a history of high inbreeding
-        or selfing, 
+        e.g. an animal lineage with a history of high inbreeding or selfing, 
         may be accompanied by a relaxation of the requirement of fertilization for the
@@
         lineages may only persist for a short time). 
+This prediction follows from the logic that although the benefit of sperm protection from drivers did not necessarily favor the evolution of the fertilization requirement, 
+	mutants who forge this requirement will experience a higher level of meiotic drive than individuals who do not.  
+Therefore removing this requirement is safest in populations with little drive. 
+We caution that other constraints on the fertilization requirement could prevent species from conforming to this prediction.
+}
+
 Our results also suggests that phylogenetic variation in the stage of female meiosis when
@@
 	(assuming the pace of centromere turnover serves as a proxy for the frequency of MI drivers). 
+\yb{INSERT BIOLOGY HERE}
 \newline
 
-%We predict that female drive, at a stage influenceable by sperm,  
-%	should be reduced when fertilized by
-%	complementary sperm. 
-%Therefore, MII female drive should be exceptionally severe in
-%	interspecific crosses between animals which complete the second meiotic division after fertilization,  
-%	as heterospecific sperm will not specifically evolve to resist a female driver in another species. 
-%Our theory also encourages phylogenetic hypotheses concerning the relationship between the opportunity 
-%	for female meiotic drive and the requirement of fertilization for the completion of female meiosis. 
-%For example, we predict that a lower opportunity for female meiotic drive,
-%        e.g. an animal lineage with a long history of high inbreeding
-%        or selfing, 
-%        would often be accompanied by a relaxation of the requirement of fertilization for the
-%        completion of female meiosis. \newline
-
+\section*{Conclusion}
 Our results highlight potentially counterintuitive results of complex genetic conflicts. 
@@
 However, given the existence of female meiotic arrest, 
-	and that its timing and mechanistic details are variable across species \citep{Masui_book,Karrbook:09} ,
+	and that its timing and mechanistic details are variable across species \citep{Masui_book,Karrbook:09},
 	the nature of the meiotic arrest and interactions between
@@
 	and counteracted by selfish drivers evolving to overcome these adaptations.
-%leave female meiosis open to sabotage by sperm collaborators, but
-%rather that  an 
-%evolutionary loss of female meiotic arrest will result in the loss of sperm defense against female meiotic drivers. }
-
-%However, given that arrest does occur its timing relative to
-%fertilization, various features of the interaction between sperm and
-%egg and divergence of these features across species  \citep{Karrbook:09} could be subject
-%to selection to conflict between female meiotic drivers and alleles
-%elsewhere in the genomes of sperm and eggs.
-%the mechanism by which it occur
-%mechanistic features, and evolutionary changes among taxa
-%various features of the
-%mechanism of fertilization 
-%however, our work does imply that the fertilization requirement does
-%not 
-%leave female meiosis open to sabotage by sperm collaborators, but
-%rather that  an 
-%evolutionary loss of female meiotic arrest will result in the loss of sperm defense against female meiotic drivers. }
 
   \appendix
-
-%\section*{Appendix}
-%We highlight simple analytical results from: 
-%	the standard model of female meiotic drive (Model 1); a sperm-dependent 
-%	model of female meiotic drive (Model 2); a male-genotype-dependent model of
-%	female meiotic drive (Model 3); and more complex two-locus sperm-dependent drive models (Models 4-6). 
-%All models include a biallelic locus ($A/B$) with non-driving and driving alleles in frequencies $f_A$ and $f_B$, 
-%	respectively, while more complex models include more loci and/or alleles.  
-%Transmission rules describing the outcomes of all matings in each
-%	model are presented in a File S1 . 
-%The fitness of genotype, $g$ ,is sex-independent and equals $1$, $1-h_s$, and $1-s$ for genotypes $AA$, $AB$, and $BB$, respectively. %  and  $w_{AA} = 1$, $w_{AB} = 1-s_h$, and $w_{BB}=1-s$. 
-%Genotypic frequencies  equal  $f_g$ for adults in the current generation, 
-%	$f_g'$ in the next generation of zygotes (i.e. after recombination, random mating, drive, and syngamy, but before selection)
-%	and $f_g''$ in adults after selection. 
-%After a complete generation $f_g'' = f_g' w_g/\bar{w}$, where $\bar{w}$ is the population mean fitness and equals $\Sigma w_g f_g'$. 
-%Because many of our analyses assume genotype frequencies expected under Hardy Weinberg Equilibrium (HWE), we note which analytical results are approximations. 
-%We verify these approximations with exact numerical iterations in Figures S1-S3 and File S3 , and detail the derivation of these approximations in our File S2.
 
@@
 \newpage
+\section*{Figure Legends}
+\clearpage
+\begin{figure}
+%\begin{center} \includegraphics[width = 1 \textwidth]{../Figures/selfprom.eps} \end{center}
+\caption{{\bf{A.}} A visual depiction of our model of `self-promoting' driver. 
+	Transmission probabilities for alleles through female
+ 	 meiosis depend on sperm genotype. 
+	 The non-driving $A$-allele, and self-promoting $B$-allele
+	 are represented by unfilled and filled circles, respectively. 
+	 {\bf{B.}} Evolution of a self-promoter and standard driver. 
+	 	Assuming that the fitnesses of drive homozygotes and 
+		heterozygotes are $1-s$ and $1$, respectively. 
+	 	\emph{Main figure:} Boundary conditions for the invasion 
+		and fixation of self-promoting and standard meiotic drivers, 
+		with drive coefficient, $d$. Colored regions depict exact results, 
+		while lines  represent analytical approximations. 
+	 \emph{B1:}  Trajectories of sperm-dependent female drive
+		 %alleles that can invade the population, 
+		 each allele has $s =0.1$ against the
+		 homozygotes.  The drive coefficient is denoted by color. 
+ }   
+\label{Eggsperm}
+\end{figure}
 
+\clearpage
+\newpage
+
+\begin{figure}
+\begin{center}
+%	\includegraphics[width = 0.7 \textwidth]{../Figures/twolocus.eps} 
+          \end{center}
+\caption{Models of a sperm-acting drive modifier tightly linked to a meiotic driver. 
+{\bf (A) } 
+Sperm carrying the derived allele at the modifier locus (filled squares) alters transmission  
+	at the driving allele (filled circles) during female meiosis.
+        Alleles at these two tightly linked loci form three haplotypes (top of A). 	
+	{\bf{A1)}} In the standard model of drive there is no
+        variation at the modifier, 
+	and the driver is transmitted to the egg with probability $d_0$. 
+	{\bf{A2)}} The modifier allele increases the transmission of the drive allele ($d_1 > d_0$), and due to their shared genetic background, also increases its drive. 
+	{\bf{A3 \& A4)}} The sperm-acting modifier acts to decrease
+        drive  ($d_1=1/2$ in A3 \& A4, or more generally, $d_1 < d_0$) and arises on the same or opposite
+        background from the driver (A3 \& A4 respectively). 
+{\bf{(B)}} Invasion of a sperm-acting drive suppressor linked to a driver. 
+	After the driver ($B$ haplotype) reaches drive selection equilibrium, we introduce a sperm acting drive modifier. 
+	We assume full drive ($D_0 =1$), a recessive lethal fitness cost to drive ($h_s=0$, $s=1$) and that the sperm-acting modifier results in a fair meiosis. 
+{\bf {B1)}} The $B^-$ allele replaces the ancestral drive haplotype, but segregates at a lower equilibrium frequency.
+{\bf {B2)}} The $A^-$ allele replaces the ancestral non-driving haplotype, and in this case, removes the driver from the population. }
+ \label{Eggsperm_3_allele_cartoon}
+\end{figure}
+
+\clearpage
+\newpage
   \setcounter{figure}{0}\global\def\thefigure{S\arabic{figure}}
@@
 \clearpage
+
+\newpage
+\subsection*{Supplementary Figure Legends}
+\clearpage
 \begin{figure}
-
-\includegraphics[width = \textwidth]{../../Scripts/homozyg_bistability.eps}
+%\includegraphics[width = \textwidth]{../../Scripts/homozyg_bistability.eps}
 \caption{Invasion analysis for a self-promoting female meiotic drive allele with
@@
 \begin{figure}
-\includegraphics[width = 0.8 \textwidth]{../Figures/bistable_x_vs_d_additive_s.eps} 
+%\includegraphics[width = 0.8 \textwidth]{../Figures/bistable_x_vs_d_additive_s.eps} 
 \caption{The unstable equilibrium frequency for a self-promoting
@@
 \begin{figure}
-\includegraphics[width = 0.7 \textwidth]{../../Scripts/effect_of_dominance_on_invasion_space_four_graph.eps} 
+%\includegraphics[width = 0.7 \textwidth]{../../Scripts/effect_of_dominance_on_invasion_space_four_graph.eps} 
 \caption{Exact results of invasion analysis of an allele whose effect

### 2014-11-17 writeup_and_notes/evolution/first_sub/male_dependent_driveFIRSTSUB.tex
added a new folder [evolution] with a draft of our evolution resubmision and response to reviewers

--- 2014-07-28 writeup_and_notes/GeneticsResub/male_dependent_drive.tex
@@
 
-\newpage 
-\flushleft{{\bf{Running title:}} Sperm prevent female meiotic drive} \\
-{\bf{Keywords:}} Meiosis; Meiotic drive; Fertilization; Genomic conflict \\
-{\bf{Corresponding author:}} Yaniv Brandvain. \emph{Address:} Department of Plant Biology, University of Minnesota --€“ Twin
-Cities. St. Paul MN, 55108. \emph{Phone:} 612-624-4375 \emph{Email:} ybrandvain@gmail.com
+%\newpage 
+%\flushleft{{\bf{Running title:}} Sperm prevent female meiotic drive} \\
+%{\bf{Keywords:}} Meiosis; Meiotic drive; Fertilization; Genomic conflict \\
+%{\bf{Corresponding author:}} Yaniv Brandvain. \emph{Address:} Department of Plant Biology, University of Minnesota --€“ Twin
+%Cities. St. Paul MN, 55108. \emph{Phone:} 612-624-4375 \emph{Email:} ybrandvain@gmail.com
 

### 2014-11-17 writeup_and_notes/evolution/supp/FileS3.R

@@ -1,0 +1,596 @@
+
+
+library("RColorBrewer")   #color scheme used in many figures
+
+###BASIC FUNCTIONS TO SETUP AND ITERATE DRIVE MODEL
+
+iterate.1.locus.drive<-function(s.array,num.iterations,female.transmission.probs,my.geno.freqs,initialize.allele.freqs=c(0.999,0.001,0)){
+	geno.freqs<-rep(NA,6)
+	names(geno.freqs)<-geno.names
+
+	my.freqs<-numeric()
+	if(missing(my.geno.freqs)){
+       #     print("running 2 allele system to stablity")
+		allele.freqs<-initialize.allele.freqs	
+		names(allele.freqs)<-c("1","2","3")	
+
+	
+		geno.freqs["11"]<-allele.freqs["1"]^2
+		geno.freqs["22"]<-allele.freqs["2"]^2
+		geno.freqs["33"]<-allele.freqs["3"]^2
+
+		geno.freqs["12"]<-2*allele.freqs["1"]*allele.freqs["2"]
+		geno.freqs["13"]<-2*allele.freqs["1"]*allele.freqs["3"]
+		geno.freqs["23"]<-2*allele.freqs["3"]*allele.freqs["2"]
+		stopifnot(sum(geno.freqs)>0.9999999999 & sum(geno.freqs)< 1.0000000001)
+
+	}else{
+		print("initiating from my.geno.freqs")
+		geno.freqs[c("11","12","22")]<-my.geno.freqs[["geno.freqs"]][nrow(my.geno.freqs[["geno.freqs"]]),c("11","12","22")]*.99
+		geno.freqs[c("13","23")]<-my.geno.freqs[["allele.freqs"]][nrow(my.geno.freqs[["geno.freqs"]]),1:2]*.99*.01
+		geno.freqs["33"]<-.01^2
+		
+		}
+
+	my.freqs<-rbind(my.freqs,geno.freqs)
+
+	for(i in 1:num.iterations){
+	#selection
+		geno.freqs["11"]<-geno.freqs["11"] * (1-s.array["11"])
+		geno.freqs["12"]<-geno.freqs["12"] * (1-s.array["12"])
+		geno.freqs["13"]<-geno.freqs["13"] * (1-s.array["13"])
+
+		geno.freqs["22"]<-geno.freqs["22"] * (1-s.array["22"])
[…]

### 2014-11-17 writeup_and_notes/evolution/response_drive_MS.tex
minor changes to response letter. acknowledgements added to ms... graham add more

@@
 \\
-{\bf{R2 minor comments:}}
+{\bf{R1 minor comments:}}
 \\ line 272 ``so it avoid[s]'' line 274 ``the the''
@@
 
-{\bf{Response to R2 minor comments:}} \emph{We fixed these, thanks!}
+{\bf{Response to R1 minor comments:}} \emph{We fixed these, thanks!}
 
@@
 We therefore believe that our results are not trivial and represent an important contribution.
-The non-triviality of our results are apparent in this comment -- in our model drive presents no cost when extremely rear, and there is some amount of cost for which a sperm acting drive promoter can invade and fix.
-The complexity of our analysis [and the bit that can choke simple intuition] is the competition between two forms of frequency dependence in the costs and benefits of ``green-bearded drive''.}
+The complexity of our analysis [and the bit that can choke simple intuition] is that a model of drive that maintains polymorphism requires negative frequency dependnce [e.g.
+in the form or of recessivity], and that the benefit of the geenbeard is [as R2 points out] positively frequency depednent.
+This makes co-existance [i.e.
+a joint polymorphism] unlikely.
+We note that the inutition of R2 with regard to our analysis appears somewhat off-base -- under some non-zero costs to a self-promoting driver, it can sweep through a population [Figure 1], a simple result that R2 does not explain with his simple intuitive assertions, above.}
 \\

### 2014-11-17 writeup_and_notes/evolution/RESUBMISSIONmale_dependent_drive.tex
minor changes to response letter. acknowledgements added to ms... graham add more

@@
 \begin{center}
-$^1$Department of Plant Biology, \\ University of Minnesota --€“ Twin
+$^1$Department of Plant Biology, \\ University of Minnesota -- Twin
 Cities. St. Paul MN, 55108 \\
@@
 
+\newpage 
+\section*{Acknowledgements}
+We would like to thank members of the Coop lab for comments. This work was made possible by the NSF Postdoctoral Fellowship in Biology FY 2010 project number 1002942 awarded to Y.B..
+\newpage 
+\bibliography{refs}
+\newpage 
   \appendix
@@
 
-\bibliography{refs}
-
 \newpage

### 2014-12-15 writeup_and_notes/evolution/RESUBMISSIONmale_dependent_drive.tex
added bit about selfing and drive on lines 201-215

@@
 	precludes the evolution of an
-	allele that suppresses this form of meiotic drive in females. \newline
+	allele that suppresses this form of meiotic drive in females. 
+\yb{Allowing for arbitrary levels of inbreeding (in the form of self-fertilization in Figures SXX-SXY), as opposed to the panmictic model explored above -- more thoroughly aligns the interests of both parents and parental chromosomes, restricting further the possibility for invasion of both traditional female drivers and 'self-promoting' drivers (Figures SXX-SXY).
+Additionally, because inbreeding reduces the frequency of heterozygotes, the invasion and fixation criteria converge as both become more strict with increasing selfing rates.} \newline
 

### 2014-12-15 scripts/functions_for_sperm_dependent_drive_paper.R
setting up selfing

@@ -1,4 +1,4 @@
-
+directory<-"~/Dropbox/Ideas/Om/scripts/"
 
 library("RColorBrewer")   #color scheme used in many figures
 
@@ -13,7 +13,7 @@ iterate.1.locus.drive<-function(s.array,num.iterations,female.transmission.probs
        #     print("running 2 allele system to stablity")
 		allele.freqs<-initialize.allele.freqs	
 		names(allele.freqs)<-c("1","2","3")	
-
+ 
 	
 		geno.freqs["11"]<-allele.freqs["1"]^2
 		geno.freqs["22"]<-allele.freqs["2"]^2
@@ -48,12 +48,15 @@ iterate.1.locus.drive<-function(s.array,num.iterations,female.transmission.probs
 		geno.freqs<-geno.freqs/sum(geno.freqs)	
 
 		geno.freqs.array<-outer(rep(geno.freqs,each=2),rep(geno.freqs,each=2),"*")  ##one for each allele
-
-	geno.freqs.array<-(1-selfing.rate) * geno.freqs.array  ##selfing  (1-s) * pi *pj
-	diag(geno.freqs.array)<-diag(geno.freqs.array) +  selfing.rate * rep(geno.freqs,each=2)  #selfing  (1-s) * pi *pj + delta_{ij} s pi
+		geno.freqs.array<-(1-selfing.rate) * geno.freqs.array  ##selfing  (1-s) * pi *pj
+		for(geno in names(geno.freqs)){  #selfing  (1-s) * pi *pj + delta_{ij} s pi  
+			these.cols<-colnames(geno.freqs.array)==geno; these.rows<-rownames(geno.freqs.array)==geno
+			geno.freqs.array[these.rows,these.cols]<- geno.freqs.array[these.rows,these.cols] +  selfing.rate* geno.freqs[geno] 
+		}
+	
 
 	##transmission
-		tot.trans<-geno.freqs.array*female.transmission.probs*0.5
+		tot.trans<-geno.freqs.array*female.transmission.probs*0.5   ## 1/2 prob. of transmission of each male allele
 
 		new.geno.freqs<-rep(NA,6)	
 		names(new.geno.freqs)<-geno.names

### 2014-12-16 scripts/functions_for_sperm_dependent_drive_paper.R
added inbreeding supp. figure and code to generate it

@@ -422,6 +422,76 @@ figures.of.traj.for.figure.2<-function(){
 
 
 
+#####Affect of inbreeding in Figure 1.
+
+make.selfing.invasion.grid.supp.fig<-function(selfing.rate=0){
+	D<-matrix(1/2,nrow=2,ncol=3,dimnames=list(c("1 or 2","3"),c("12","13","23"))) ##entries are drive coeff of the 2nd allele listed against the 1st.
+	
+	d.range<-seq(0.5,1,length=200)
+	s3.range<-seq(0,1,length=200)
+	print("invasion of self prop. driver")
+	####ivasion of self prop. driver
+	invasion.grid<-sapply(s3.range,function(s3){
+		sapply(d.range,function(d){
+			test.invasion.self.prop(d=d,s.het=0,s.hom=s3,selfing.rate=selfing.rate)
+		})
+	})
+	invasion.lines.selfprop<-s3.range[apply(invasion.grid,1,function(x){max(which(x))})]
+
+	print("fixation of simpler driver")
+
+	###fixation of simpler driver
+	invasion.grid<-sapply(s3.range,function(s3){
+		sapply(d.range,function(d){
+			test.fixation.of.simple.driver(d=d,s.het=0,x=0.001,s.hom=s3,selfing.rate=selfing.rate)
+		})
+	})
+	invasion.lines.simple.driver<-s3.range[apply(invasion.grid,1,function(x){max(which(x))})]
+
+	print("fixation of self prop. driver")
+
+	####fixation of self prop. driver
+	fixation.grid<-sapply(s3.range,function(s3){
+		sapply(d.range,function(d){
+			test.invasion.self.prop(d=d,s.het=0,x=0.999,s.hom=s3,selfing.rate=selfing.rate)
+		})
+	})
+	fixation.lines.selfprop<-s3.range[apply(fixation.grid,1,function(x){max(which(x))})]
+
+	print("fixation of simpler driver")
+		
+	###fixation of simpler driver
+	fixation.grid<-sapply(s3.range,function(s3){
+		sapply(d.range,function(d){
+			test.fixation.of.simple.driver(d=d,s.het=0,s.hom=s3,selfing.rate=selfing.rate)
+		})
+	})
+	fixation.lines.simple.driver<-s3.range[apply(fixation.grid,1,function(x){max(which(x))})]
+	
[…]

### 2014-12-17 scripts/meiotic_pause_phylogeny.R
added scripts/meiotic_pause_phylogeny.R and scripts/redone_meiotic_table.csv

@@ -1,0 +1,93 @@
+#https://github.com/fmichonneau/rotl
+
+#library(devtools)
+#install_github("fmichonneau/rncl")
+#install_github("fmichonneau/rotl")
+
+
+library("rotl")
+library("rncl")
+library("ape")
+
+
+
+write.csv(file="~/Dropbox/Ideas/Om/scripts/redone_meiotic_table.csv",cbind(meiotic.table[reordered.meiotic.table,],resolved_names.no.NAs))
+
+new.meiotic.table<-read.csv(file="~/Dropbox/Ideas/Om/scripts/redone_meiotic_table.csv",as.is=TRUE)[,-1]
+
+new.meiotic.table[grep("Lymnea",new.meiotic.table$unique_name),colnames(resolved_names.no.NAs)]<-tnrs_match_names("Lymnaeidae")
+new.meiotic.table[grep("Pectinaria",new.meiotic.table$unique_name),colnames(resolved_names.no.NAs)]<-tnrs_match_names("Pectinariidae")
+
+tr <- tol_induced_subtree(ott_ids=new.meiotic.table$ott_id)
+
+old.names<- tr$tip.label
+ new.names<-sapply(tr$tip.label,function(x){strsplit(x,"_")[[1]][1]})
+ phylogeny.ott<-sapply(tr$tip.label,function(x){strsplit(x,"ott")[[1]][2]})
+ 
+order.meiotic.table.phy<- match(phylogeny.ott,new.meiotic.table$ott_id)
+ tr$tip.label<-paste(new.meiotic.table[order.meiotic.table.phy,1],new.names)
+
+
+sperm.entry.stage<-new.meiotic.table[order.meiotic.table.phy,]$V3
+sperm.entry.stage[sperm.entry.stage ==" GV (-MI)"]<-" GV-MI"
+sperm.entry.stage[sperm.entry.stage ==" Pron. "]<- " Pron."
+ sperm.entry.stage[sperm.entry.stage =="GV"]<- " GV"
+stage.cols<-rainbow(length(unique( sperm.entry.stage)))
+
+
+unfertilized.stage<-new.meiotic.table[order.meiotic.table.phy,]$V4
+unfertilized.stage[unfertilized.stage =="Pron. "]<-" Pron."
+unfertilized.stage[unfertilized.stage ==" ?"]<-"?"
+
+stage.cols<-c(rainbow(length(unique( sperm.entry.stage))),"grey")
+names(stage.cols)<- c(unique(sperm.entry.stage),"?")
[…]

### 2014-12-17 scripts/meiotic_pause_phylogeny.R
small changes to meiotic_pause_phylogeny.R

@@ -10,13 +10,12 @@ library("rncl")
 library("ape")
 
 
-
-write.csv(file="~/Dropbox/Ideas/Om/scripts/redone_meiotic_table.csv",cbind(meiotic.table[reordered.meiotic.table,],resolved_names.no.NAs))
-
 new.meiotic.table<-read.csv(file="~/Dropbox/Ideas/Om/scripts/redone_meiotic_table.csv",as.is=TRUE)[,-1]
 
 new.meiotic.table[grep("Lymnea",new.meiotic.table$unique_name),colnames(resolved_names.no.NAs)]<-tnrs_match_names("Lymnaeidae")
 new.meiotic.table[grep("Pectinaria",new.meiotic.table$unique_name),colnames(resolved_names.no.NAs)]<-tnrs_match_names("Pectinariidae")
+
+new.meiotic.table[grep("Patina",new.meiotic.table$unique_name),colnames(resolved_names.no.NAs)]<-tnrs_match_names("Patiria")
 
 tr <- tol_induced_subtree(ott_ids=new.meiotic.table$ott_id)
 
@@ -49,7 +48,7 @@ fertilization.site<-new.meiotic.table[order.meiotic.table.phy,]$V5
 names(fertilization.pch)<- c( "External"   ,  "Internal" ,    "Intraovarian", " ? " )
 # Abbreviations: GV germinal vesicle stage; Pron. pronucleus; MI metaphase I; Al anaphase I; MII metaphase II; AII anaphase II. 
 
-pdf(file=paste(directory,"../meiotic_phylogeny.pdf",sep=""))
+#pdf(file=paste(directory,"../meiotic_phylogeny.pdf",sep=""))
 
  par(mar=c(0,0,0,0))
  plot.phylo(tr,cex=0.75,label.offset=7) #,type="fan")

### 2014-12-18 scripts/meiotic_pause_phylogeny.R
added tree Robj

@@ -26,6 +26,8 @@ old.names<- tr$tip.label
 order.meiotic.table.phy<- match(phylogeny.ott,new.meiotic.table$ott_id)
  tr$tip.label<-paste(new.meiotic.table[order.meiotic.table.phy,1],new.names)
 
+save(file="~/Dropbox/Ideas/Om/scripts/meiotic_tree.Robj",tr,order.meiotic.table.phy,new.meiotic.table)
+load(file="~/Dropbox/Ideas/Om/scripts/meiotic_tree.Robj")
 
 sperm.entry.stage<-new.meiotic.table[order.meiotic.table.phy,]$V3
 sperm.entry.stage[sperm.entry.stage ==" GV (-MI)"]<-" GV-MI"

### 2014-12-22 writeup_and_notes/evolution/RESUBMISSIONmale_dependent_drive.tex
updated the main paper

--- 2014-11-17 writeup_and_notes/evolution/RESUBMISSIONmale_dependent_drive.tex
@@
 it is potentially open to sabotage by a virtual stranger -- a haploid sperm genome.
-That is, in many animal species, female meiosis is completed only after fertilization \citep{Masui_book}, 
-	creating ample opportunity for interaction between the sperm and
+That is, in many animal species, female meiosis is completed only
+after fertilization \citep{Masui_book}, creating ample opportunity for interaction between the sperm and
 	female meiotic machinery (note that the alternation of generations
 	likely precludes this interaction in plants). 
-\yb{INSERT BIOLOGY HERE}
+\gc{The timing of enty of sperm into the egg relative to the
+  progression of female meiosis is variable across animals, as is the
+  stage at which female meiosis
+  stalls without fertilization (we show the phylogenetic distribution
+  in supplementary Figure XXX).}
 Therefore, a `green-bearded' \citep{Gardner2010} sperm-acting allele that recognizes and facilitates the meiotic drive of a genetically equivalent allele in heterozygous females 
@@
 	precludes the evolution of an
-	allele that suppresses this form of meiotic drive in females. \newline
+	allele that suppresses this form of meiotic drive in females. 
+\gc{Allowing for arbitrary levels of inbreeding (in the form of
+  self-fertilization), as opposed to the panmictic model explored
+  above -- more thoroughly aligns the interests of both parents and
+  parental chromosomes, restricting further the possibility for
+  invasion of both traditional female drivers and 'self-promoting'
+  drivers (Figures SXX-SXY). Additionally, because inbreeding reduces
+  the frequency of heterozygotes, the invasion and fixation criteria
+  converge as both become more strict with selfing promoting rates.} \newline
 
@@
 	of evidence  to exclude plausible alternative hypotheses such as genotypic 
-	inviability including epistatic maternal X zygotic lethality \citep[e.g. ][]{Sawamura:1993aa}.  \newline
+	inviability including epistatic maternal by zygotic lethality \citep[e.g. ][]{Sawamura:1993aa}.  \newline
 }
@@
 Our results also suggests that phylogenetic variation in the stage of female meiosis when
-        fertilization occurs may influence the prevalence of  female meiotic drive.
+        fertilization occurs \gc{(see Supplementary Figure XXX)} may influence the prevalence of  female meiotic drive.
 For example, centromeric drive may be more common  in taxa where females complete MI before fertilization, 
@@
 	(assuming the pace of centromere turnover serves as a proxy for the frequency of MI drivers). 
-\yb{INSERT BIOLOGY HERE}
 \newline
@@
 However, given the existence of female meiotic arrest, 
-	and that its timing and mechanistic details are variable across species \citep{Masui_book,Karrbook:09},
+	and that its timing and mechanistic details are variable
+        across species \citep[Supplementary Figure XXX, and][ ]{Masui_book,Karrbook:09},
 	the nature of the meiotic arrest and interactions between

### 2014-12-22 scripts/meiotic_pause_phylogeny.R

@@ -63,6 +63,7 @@ legend("topleft",pch=c(rep(19,length(stage.cols)),fertilization.pch) ,legend=c(n
 dev.off()
 
 
+phyla.midpoint<-sapply(unique(phyla),function(phylum){mean((1:length(phyla))[phyla==phylum])})
 
 #	reading in OCR text
 

### 2014-12-22 writeup_and_notes/evolution/RESUBMISSIONmale_dependent_drive.tex
In blue, YB makes minor changes to GC improvements

@@
 after fertilization \citep{Masui_book}, creating ample opportunity for interaction between the sperm and
-	female meiotic machinery (note that the alternation of generations
-	likely precludes this interaction in plants). 
-\gc{The timing of enty of sperm into the egg relative to the
-  progression of female meiosis is variable across animals, as is the
-  stage at which female meiosis
-  stalls without fertilization (we show the phylogenetic distribution
-  in supplementary Figure XXX).}
-Therefore, a `green-bearded' \citep{Gardner2010} sperm-acting allele that recognizes and facilitates the meiotic drive of a genetically equivalent allele in heterozygous females 
-     could presumably rapidly spread through a population. 
+%	female meiotic machinery (note that the alternation of generations
+%	likely precludes this interaction in plants). 
+\yb{	female meiotic machinery (note that, across animals the variation in timing of sperm entry into the egg and the timing at which female meiosis stalls [supplementary Figure XXX] complicates this opportunity in some taxa, and that the alternation of generations likely precludes this interaction in plants).}
+%\gc{The timing of entry of sperm into the egg relative to the
+%  progression of female meiosis is variable across animals, as is the
+%  stage at which female meiosis
+%  stalls without fertilization (we show the phylogenetic distribution
+%  in supplementary Figure XXX).}
+Therefore, in many species a `green-bearded' \citep{Gardner2010} sperm-acting allele that recognizes and facilitates the meiotic drive of a genetically equivalent allele in heterozygous females could presumably rapidly spread through a population.
 At first sight, female meiosis appears primed for conflict caused by such selfish systems. 
@@
 	allele that suppresses this form of meiotic drive in females. 
-\gc{Allowing for arbitrary levels of inbreeding (in the form of
-  self-fertilization), as opposed to the panmictic model explored
-  above -- more thoroughly aligns the interests of both parents and
+\gc{ \yb{Relaxing our assumptions of panmixia by} allowing for 
+  arbitrary levels of inbreeding (in the form of self-fertilization),
+  %, as opposed to the panmictic model exploredabove -- 
+  more thoroughly aligns the interests of both parents and
   parental chromosomes, restricting further the possibility for
@@
   the frequency of heterozygotes, the invasion and fixation criteria
-  converge as both become more strict with selfing promoting rates.} \newline
+  converge, \yb{as both become stricter with increased inbreeding rates}.} \newline
 

### 2014-12-23 writeup_and_notes/evolution/response_drive_MS.tex
setting up selfing
added to response to reviewers

@@
 We are submitting a new version of this manuscript in which we have incorporated reviewer comments as advised by the AE.
-Our revision includes versions of Models 1 and 2 with some degree of inbreeding, and has been edited to ensure clarity, and minimize the possibility of misleading our readers.
+Our revision includes versions of Models 1 and 2 with  \gc{inbreeding}, and has been edited to ensure clarity, and minimize the possibility of misleading our readers.
 We believe that this has resulted in a stronger manuscript.
@@
 
-{\bf{Response to AE comment 2:}}  \emph{WE NEED TO DO SOMETHING HERE \dots{}}
+{\bf{Response to AE comment 2:}}  \emph{\gc{We have now included a phylogeny of animals showing the phylogenetic distribution of the stage of female meiosis, at which fertilization occurs.
+This has been left as a supplementary figure as we do not view this as key to our arguments.
+However, it is a nice addition to the paper, and we thank reviewers \& editor for this suggestion.
+}}
+
 \\
@@
 
-{\bf{Response to AE comment 5:}} \emph{We agree with this ida, and have incorporated it into our discussion.}
+{\bf{Response to AE comment 5:}} \emph{We agree with this idea, and have incorporated it into our discussion.}
 \\
@@
 
-{\bf{Response to R2 summary:}} \emph{We appreciate that you find the question, the biology, and the results compelling, and are disappointed that you are unconvinced by the analysis.
-However, we can find no statement in your specific comments below that suggest that you are unconvinced.
-We therefore respond to the comments below, and hope that the responses to your specific comments are convincing.}
+{\bf{Response to R2 summary:}} \emph{We appreciate that \gc{the reviewer} find the question, the biology, and the results compelling, and are disappointed that the \gc{reviewer is} unconvinced by the analysis.
+However, we can find no statement in \gc{the specific comments below that suggest why the reviewer is unconvinced}.
+We therefore respond to the comments \gc{presented below}.}
 \\
@@
 {\bf{R2 comment 1 :}}
-(1) The results are pretty obvious a priori, and don?t really warrant the development and analysis of explicit models.
+(1) The results are pretty obvious a priori, and don't really warrant the development and analysis of explicit models.
 As the authors note in the Introduction, the matching-alleles models are instances of greenbeard genes, and we already know that these are under positively frequency-dependent selection, and cannot invade from rarity in unstructured populations if there is any obligate cost involved, such as the cost of the machinery that does the allelic recognition.
@@
 
-{\bf{Response to R2 comment 3:}}  \emph{\yb{NEED TO WORK ON THIS}}
+{\bf{Response to R2 comment 3:}}  \emph{\gc{We have now included a phylogeny of animals showing the phylogenetic distribution of the stage of female meiosis, at which fertilization occurs}}
 \\
-\\
-\\
-\\
-As is my policy, I waive anonymity
-\\
-Andy Gardner
 

### 2014-12-24 writeup_and_notes/evolution/response_drive_MS.tex
yaniv accepted most of grahams changes to the response. But made changes to AE comment 2.

@@
 We are submitting a new version of this manuscript in which we have incorporated reviewer comments as advised by the AE.
-Our revision includes versions of Models 1 and 2 with  \gc{inbreeding}, and has been edited to ensure clarity, and minimize the possibility of misleading our readers.
+Our revision includes versions of Models 1 and 2 with inbreeding, and has been edited to ensure clarity, and minimize the possibility of misleading our readers.
 We believe that this has resulted in a stronger manuscript.
@@
 {\bf{Response to AE comment 1:}}  \emph{We note that both R1 and R2 noted an interest in the case of inbreeding.
-\yb{We therefore have explored the behavior of model 2 [as well as model one for reference to the case of traditional drive] under six values of inbreeding.
-We briefly discuss these results in the modified text and make the figures [analogous to Figure 1B] in the Supplement.}}
+We therefore have explored the behavior of model 2 [as well as model one for reference to the case of traditional drive] under six values of inbreeding.
+We briefly discuss these results in the modified text and make the figures [analogous to Figure 1B] in the Supplement.}
 \\
@@
 {\bf{Response to AE comment 2:}}  \emph{\gc{We have now included a phylogeny of animals showing the phylogenetic distribution of the stage of female meiosis, at which fertilization occurs.
-This has been left as a supplementary figure as we do not view this as key to our arguments.
-However, it is a nice addition to the paper, and we thank reviewers \& editor for this suggestion.
+This has been left as a supplementary figure as we do not view this as key to our arguments; however, \yb{we believe} it is a nice addition to the paper.
+\yb{We hope that this data and our pointing to it in the main text, highlights this under appreciated biology and encourages additional research}.
+We thank reviewers \& editor for this suggestion.
 }}
-
 \\
@@
 
-{\bf{Response to R2 summary:}} \emph{We appreciate that \gc{the reviewer} find the question, the biology, and the results compelling, and are disappointed that the \gc{reviewer is} unconvinced by the analysis.
-However, we can find no statement in \gc{the specific comments below that suggest why the reviewer is unconvinced}.
-We therefore respond to the comments \gc{presented below}.}
+{\bf{Response to R2 summary:}} \emph{We appreciate that the reviewer find the question, the biology, and the results compelling, and are disappointed that the reviewer is unconvinced by the analysis.
+However, we can find no statement in the specific comments below that suggest why the reviewer is unconvinced.
+We therefore respond to the comments presented below.}
 \\
@@
 We therefore believe that our results are not trivial and represent an important contribution.
-The complexity of our analysis [and the bit that can choke simple intuition] is that a model of drive that maintains polymorphism requires negative frequency dependnce [e.g.
-in the form or of recessivity], and that the benefit of the geenbeard is [as R2 points out] positively frequency depednent.
-This makes co-existance [i.e.
+The complexity of our analysis [and the bit that can choke simple intuition] is that a model of drive that maintains polymorphism requires negative frequency dependance [e.g.
+in the form or of recessivity], and that the benefit of the greenbeard is [as R2 points out] positively frequency dependent.
+This makes coexistence [i.e.
 a joint polymorphism] unlikely.
-We note that the inutition of R2 with regard to our analysis appears somewhat off-base -- under some non-zero costs to a self-promoting driver, it can sweep through a population [Figure 1], a simple result that R2 does not explain with his simple intuitive assertions, above.}
+We note that the intuition of R2 with regard to our analysis appears somewhat off-base -- under some non-zero costs to a self-promoting driver, it can sweep through a population [Figure 1], a simple result that R2 does not explain with his simple intuitive assertions, above.}
 \\
@@
 
-{\bf{Response to R2 comment 3:}}  \emph{\gc{We have now included a phylogeny of animals showing the phylogenetic distribution of the stage of female meiosis, at which fertilization occurs}}
+{\bf{Response to R2 comment 3:}}  \emph{We have now included a phylogeny of animals showing the phylogenetic distribution of the stage of female meiosis, at which fertilization occurs.}
 \\

### 2014-12-24 writeup_and_notes/evolution/RESUBMISSIONmale_dependent_drive.tex
added caption & phylo. figure

@@
 \clearpage
+
+\begin{figure}
+%\includegraphics[width = \textwidth]{supp/meiotic_phylogeny.pdf}
+\caption{The phylogenetic distribution of the female meiotic
+  arrest. The first two columns of symbols gives the stage of the ooycte when
+  the sperm penerates and the stage at which it arrests if
+  unfertilized. The third column of symbols shows the site for natural fertilization. 
+  These columns of phenotypic data were extracted from Table
+  1 of \citet{Masui_book}. The tree was extracted from the Open Tree of
+Life project. The raw data table and script to do this is included in
+the supplement. }  
+\label{Meiotic_fig}
+\end{figure}
+
 \begin{figure}

### 2014-12-24 scripts/meiotic_pause_phylogeny.R

@@ -16,6 +16,8 @@ new.meiotic.table[grep("Lymnea",new.meiotic.table$unique_name),colnames(resolved
 new.meiotic.table[grep("Pectinaria",new.meiotic.table$unique_name),colnames(resolved_names.no.NAs)]<-tnrs_match_names("Pectinariidae")
 
 new.meiotic.table[grep("Patina",new.meiotic.table$unique_name),colnames(resolved_names.no.NAs)]<-tnrs_match_names("Patiria")
+new.meiotic.table[grep("Helix",new.meiotic.table$unique_name),colnames(resolved_names.no.NAs)]<-tnrs_match_names("Helicidae")
+new.meiotic.table[grep("Pedicellina",new.meiotic.table$unique_name),colnames(resolved_names.no.NAs)]<-tnrs_match_names("Entoprocta")
 
 tr <- tol_induced_subtree(ott_ids=new.meiotic.table$ott_id)
 
@@ -25,6 +27,9 @@ old.names<- tr$tip.label
  
 order.meiotic.table.phy<- match(phylogeny.ott,new.meiotic.table$ott_id)
  tr$tip.label<-paste(new.meiotic.table[order.meiotic.table.phy,1],new.names)
+
+ tr$tip.label[ tr$tip.label=="Entoprocta Entoprocta"]<-"Entoprocta Pedicellina"
+
 
 save(file="~/Dropbox/Ideas/Om/scripts/meiotic_tree.Robj",tr,order.meiotic.table.phy,new.meiotic.table)
 load(file="~/Dropbox/Ideas/Om/scripts/meiotic_tree.Robj")
@@ -40,8 +45,10 @@ unfertilized.stage<-new.meiotic.table[order.meiotic.table.phy,]$V4
 unfertilized.stage[unfertilized.stage =="Pron. "]<-" Pron."
 unfertilized.stage[unfertilized.stage ==" ?"]<-"?"
 
-stage.cols<-c(rainbow(length(unique( sperm.entry.stage))),"grey")
-names(stage.cols)<- c(unique(sperm.entry.stage),"?")
+stage.cols<-   stage.cols<-    c(brewer.pal(7,name="Accent"),"light grey")    #c(rainbow(length(unique( sperm.entry.stage))),"grey")
+
+											##ordered
+names(stage.cols)<- c(c(" GV"," GV-MI"," MI"," AI"," MII"," AII", " Pron."),"?")
[…]

### 2014-12-24 writeup_and_notes/evolution/response_drive_MS.tex
yb for submission 12/24/14

@@
 Sincerely \\
-Yaniv Brandvain and Graham Coop
+Yaniv Brandvain and Graham Coop\\
+12/24/14
 \newpage
@@
 
-{\bf{Response to AE comment 2:}}  \emph{\gc{We have now included a phylogeny of animals showing the phylogenetic distribution of the stage of female meiosis, at which fertilization occurs.
-This has been left as a supplementary figure as we do not view this as key to our arguments; however, \yb{we believe} it is a nice addition to the paper.
-\yb{We hope that this data and our pointing to it in the main text, highlights this under appreciated biology and encourages additional research}.
+{\bf{Response to AE comment 2:}}  \emph{We have now included a phylogeny of animals showing the phylogenetic distribution of the stage of female meiosis, at which fertilization occurs.
+This has been left as a supplementary figure as we do not view this as key to our arguments; however, we believe it is a nice addition to the paper.
+We hope that this data and our pointing to it in the main text, highlights this under appreciated biology and encourages additional research}.
 We thank reviewers \& editor for this suggestion.
-}}
 \\

### 2014-12-24 writeup_and_notes/evolution/RESUBMISSIONmale_dependent_drive.tex
yb for submission 12/24/14

@@
 after fertilization \citep{Masui_book}, creating ample opportunity for interaction between the sperm and
-%	female meiotic machinery (note that the alternation of generations
-%	likely precludes this interaction in plants). 
-\yb{	female meiotic machinery (note that, across animals the variation in timing of sperm entry into the egg and the timing at which female meiosis stalls [supplementary Figure XXX] complicates this opportunity in some taxa, and that the alternation of generations likely precludes this interaction in plants).}
-%\gc{The timing of entry of sperm into the egg relative to the
-%  progression of female meiosis is variable across animals, as is the
-%  stage at which female meiosis
-%  stalls without fertilization (we show the phylogenetic distribution
-%  in supplementary Figure XXX).}
+female meiotic machinery (note that, across animals the variation in timing of sperm entry into the egg and the timing at which female meiosis stalls (Figure \ref{Meiotic_fig}) complicates this opportunity in some taxa, and that the alternation of generations likely precludes this interaction in plants).
 Therefore, in many species a `green-bearded' \citep{Gardner2010} sperm-acting allele that recognizes and facilitates the meiotic drive of a genetically equivalent allele in heterozygous females could presumably rapidly spread through a population.
@@
 	and in fact, females eschewing this delay would potentially lose sperm assistance in the suppression of meiotic drivers.  
-\yb{Given the wide-spread requirement of fertilization for the completion
+Given the wide-spread requirement of fertilization for the completion
 	of  female meiosis, various features of the interaction between sperm 
@@
 	wether this result is the ultimate evolutionary function of the fertilization requirement or a coincidental pleiotropic outcome is beyond the scope of this manuscript, 
-	but our intuition argues against the prior (see Discussion).}
+	but our intuition argues against the prior (see Discussion).
 
@@
 that genotype frequencies follow Hardy Weinberg Equilibrium (HWE), we note which analytical results are approximate. 
-We verify these approximations with exact numerical iterations in Figures S1-S3 and File S3.
+We verify these approximations with exact numerical iterations in Figures S2-S4 and File S3.
 
@@
 	allele that suppresses this form of meiotic drive in females. 
-\gc{ \yb{Relaxing our assumptions of panmixia by} allowing for 
-  arbitrary levels of inbreeding (in the form of self-fertilization),
+Relaxing our assumptions of panmixia by allowing for 
+  arbitrary levels of inbreeding (in the form of self-fertilization, implemented in File S3),
   %, as opposed to the panmictic model exploredabove -- 
@@
   invasion of both traditional female drivers and 'self-promoting'
-  drivers (Figures SXX-SXY). Additionally, because inbreeding reduces
+  drivers (Figure \ref{self}). Additionally, because inbreeding reduces
   the frequency of heterozygotes, the invasion and fixation criteria
-  converge, \yb{as both become stricter with increased inbreeding rates}.} \newline
+  converge, as both become stricter with increased inbreeding rates.\newline
 
@@
         drivers and sperm alleles. 
-\yb{Although the fertilization requirement allows sperm an opportunity to enforce fairness in female meiosis, 
+Although the fertilization requirement allows sperm an opportunity to enforce fairness in female meiosis, 
 	this is unlikely it evolutionary raison d'\^{e}tre. 
-In fact, to suggest so, presupposes that sperm have an evolved system, to prevent meiotic drive before they have a mechanism to do so.}
+In fact, to suggest so, presupposes that sperm have an evolved system, to prevent meiotic drive before they have a mechanism to do so.
 \newline 
@@
 
-\yb{The theory developed above has one overarching conclusion -- that when possible, males evolve to make female meiosis fair. 
+The theory developed above has one overarching conclusion -- that when possible, males evolve to make female meiosis fair. 
 This simple result provides numerous novel predictions, many of which are directly testable.
@@
 Both this prediction, and the empirical support for it run contrary to expectations of a na\"{i}ve verbal ``green-beard'' model.
-}
 
-\yb{Our model of a sperm-acting drive suppressor unlinked to a female driver (Model 6) also predicts that sperm should evolve to prevent meiotic drive; 
+Our model of a sperm-acting drive suppressor unlinked to a female driver (Model 6) also predicts that sperm should evolve to prevent meiotic drive; 
 	however, it contains no simple mechanism to maintain polymorphism for sperm-acting drive suppression.
@@
 	of evidence  to exclude plausible alternative hypotheses such as genotypic 
-	inviability including epistatic maternal by zygotic lethality \citep[e.g. ][]{Sawamura:1993aa}.  \newline
-}
+	inviability including epistatic maternal by zygotic lethality \citep[e.g. ][]{Sawamura:1993aa}.
 
@@
 
-\yb{
 For example, we predict that a lower opportunity for female meiotic drive,
@@
 We caution that other constraints on the fertilization requirement could prevent species from conforming to this prediction.
-}
 
 Our results also suggests that phylogenetic variation in the stage of female meiosis when
-        fertilization occurs \gc{(see Supplementary Figure XXX)} may influence the prevalence of  female meiotic drive.
+        fertilization occurs (see Figure \ref{Meiotic_fig}) may influence the prevalence of  female meiotic drive.
 For example, centromeric drive may be more common  in taxa where females complete MI before fertilization, 
@@
 	and that its timing and mechanistic details are variable
-        across species \citep[Supplementary Figure XXX, and][ ]{Masui_book,Karrbook:09},
+        across species \citep[Figure \ref{Meiotic_fig}, and] []{Masui_book,Karrbook:09}
 	the nature of the meiotic arrest and interactions between
@@
 	and an increase in frequency of the $C$ allele when it suppresses drive ($\epsilon <0$ -- the $B^-$ model, above). 
-More generally, even when the cost of drive is not fully recessive, the $B^-$ allele will invade and fix under all parameters sustaining a previous polymorphism at the drive locus (see File S2  ).
+More generally, even when the cost of drive is not fully recessive, the $B^-$ allele will invade and fix under all parameters sustaining a previous polymorphism at the drive locus (see File S2).
 
@@
 For all values of interest ($0<s<1$, $0.5<d_0<1$), the change in frequency a rare $C$ allele is positive when it decreases drive (i.e.
-$\epsilon <0$, corresponding to the $A^-$ model, above), a result which holds qualitatively for a common $C$ allele, as well (File S2  ).
+$\epsilon <0$, corresponding to the $A^-$ model, above), a result which holds qualitatively for a common $C$ allele, as well (File S2).
 %Somewhat counterintuitively, the spread of the $A^-$ allele is somewhat slower than that of the $B^-$ allele (it is on the order of $C^{2}$ rather than $C$). 
@@
 {\bf{File S1:}} Transmission rules. We detail the frequency of offspring genotypes produced by each possible mating for Models 1-6 in the tabs of this Excel spreadsheet. 
-
+\\ \\
 {\bf{Files S2A \& S2B:}} A Mathematica file (FileS2A) and a PDF of this file (File S2B) in which we derive analytical results for models 1-6 and $6^\prime$. 
-
-{\bf{File S3:}} Exact approach. The R Script used for exact recursions.
+\\ \\
+{\bf{File S3:}} Exact approach. The R Script used for exact recursions for all models, including cases with inbreeding.
+\\ \\
+{\bf{File S4:}} Variation in critical time-points during female meiosis across taxa (adapted from \citet{Masui_book}).
+\\ \\
+{\bf{File S5:}} An R object containing the phylogeny and raw data used to generate Figure \ref{Meiotic_fig}.
+\\ \\
+{\bf{File S6:}} The R Script used to generate Figure \ref{Meiotic_fig}. This requires that File S5 is loaded into the R environment.
 \clearpage
@@
   arrest. The first two columns of symbols gives the stage of the ooycte when
-  the sperm penerates and the stage at which it arrests if
+  the sperm penetrates and the stage at which it arrests if
   unfertilized. The third column of symbols shows the site for natural fertilization. 
@@
   1 of \citet{Masui_book}. The tree was extracted from the Open Tree of
-Life project. The raw data table and script to do this is included in
-the supplement. }  
+Life project. The raw data table, the phylogeny/supporting R objects, and the script to do this is are included in
+the supplement (Files S4-S6). }  
 \label{Meiotic_fig}
@@
 
+\begin{figure}
+%\includegraphics[width = 0.8 \textwidth]{../Figures/bistable_x_vs_d_additive_s.eps} 
+\caption{Evolution of a self-promoter and standard driver with variable levels of inbreeding (modifying the selfing rate from 0 to 0.9, in 0.1 increments). 
+	 	Assuming that the fitnesses of drive homozygotes and 
+		heterozygotes are $1-s$ and $1$, respectively. 
+	 	Boundary conditions for the invasion (solid lines) 
+		and fixation (dotted lines) of self-promoting (red) and standard (black) meiotic drivers, 
+		with drive coefficient, $d$. 
+		We derived these conditions from the simulation in File S3.}
+		\label{self}
+\end{figure}
+
 \end{document}

### 2014-12-24 writeup_and_notes/evolution/supp/FileS3.R

@@ -1,10 +1,10 @@
-
+directory<-"~/Dropbox/Ideas/Om/scripts/"
 
 library("RColorBrewer")   #color scheme used in many figures
 
 ###BASIC FUNCTIONS TO SETUP AND ITERATE DRIVE MODEL
 
-iterate.1.locus.drive<-function(s.array,num.iterations,female.transmission.probs,my.geno.freqs,initialize.allele.freqs=c(0.999,0.001,0)){
+iterate.1.locus.drive<-function(s.array,num.iterations,female.transmission.probs,my.geno.freqs,initialize.allele.freqs=c(0.999,0.001,0),selfing.rate=0){
 	geno.freqs<-rep(NA,6)
 	names(geno.freqs)<-geno.names
 
@@ -13,7 +13,7 @@ iterate.1.locus.drive<-function(s.array,num.iterations,female.transmission.probs
        #     print("running 2 allele system to stablity")
 		allele.freqs<-initialize.allele.freqs	
 		names(allele.freqs)<-c("1","2","3")	
-
+ 
 	
 		geno.freqs["11"]<-allele.freqs["1"]^2
 		geno.freqs["22"]<-allele.freqs["2"]^2
@@ -47,10 +47,16 @@ iterate.1.locus.drive<-function(s.array,num.iterations,female.transmission.probs
 	
 		geno.freqs<-geno.freqs/sum(geno.freqs)	
 
-		geno.freqs.array<-outer(rep(geno.freqs,each=2),rep(geno.freqs,each=2),"*")
+		geno.freqs.array<-outer(rep(geno.freqs,each=2),rep(geno.freqs,each=2),"*")  ##one for each allele
+		geno.freqs.array<-(1-selfing.rate) * geno.freqs.array  ##selfing  (1-s) * pi *pj
+		for(geno in names(geno.freqs)){  #selfing  (1-s) * pi *pj + delta_{ij} s pi  
+			these.cols<-colnames(geno.freqs.array)==geno; these.rows<-rownames(geno.freqs.array)==geno
+			geno.freqs.array[these.rows,these.cols]<- geno.freqs.array[these.rows,these.cols] +  selfing.rate* geno.freqs[geno] 
+		}
+	
 
 	##transmission
-		tot.trans<-geno.freqs.array*female.transmission.probs*0.5
+		tot.trans<-geno.freqs.array*female.transmission.probs*0.5   ## 1/2 prob. of transmission of each male allele
 
[…]

### 2014-12-24 writeup_and_notes/evolution/supp/FileS6.R

@@ -1,0 +1,106 @@
+#https://github.com/fmichonneau/rotl
+
+#library(devtools)
+#install_github("fmichonneau/rncl")
+#install_github("fmichonneau/rotl")
+
+
+library("rotl")
+library("rncl")
+library("ape")
+
+
+new.meiotic.table<-read.csv(file="FileS4.csv",as.is=TRUE)[,-1]
+
+new.meiotic.table[grep("Lymnea",new.meiotic.table$unique_name),colnames(resolved_names.no.NAs)]<-tnrs_match_names("Lymnaeidae")
+new.meiotic.table[grep("Pectinaria",new.meiotic.table$unique_name),colnames(resolved_names.no.NAs)]<-tnrs_match_names("Pectinariidae")
+
+new.meiotic.table[grep("Patina",new.meiotic.table$unique_name),colnames(resolved_names.no.NAs)]<-tnrs_match_names("Patiria")
+new.meiotic.table[grep("Helix",new.meiotic.table$unique_name),colnames(resolved_names.no.NAs)]<-tnrs_match_names("Helicidae")
+new.meiotic.table[grep("Pedicellina",new.meiotic.table$unique_name),colnames(resolved_names.no.NAs)]<-tnrs_match_names("Entoprocta")
+
+tr <- tol_induced_subtree(ott_ids=new.meiotic.table$ott_id)
+
+old.names<- tr$tip.label
+ new.names<-sapply(tr$tip.label,function(x){strsplit(x,"_")[[1]][1]})
+ phylogeny.ott<-sapply(tr$tip.label,function(x){strsplit(x,"ott")[[1]][2]})
+ 
+order.meiotic.table.phy<- match(phylogeny.ott,new.meiotic.table$ott_id)
+ tr$tip.label<-paste(new.meiotic.table[order.meiotic.table.phy,1],new.names)
+
+ tr$tip.label[ tr$tip.label=="Entoprocta Entoprocta"]<-"Entoprocta Pedicellina"
+
+
+save(file="~/FileS5.Robj",tr,order.meiotic.table.phy,new.meiotic.table)
+load(file="~/FileS5.Robj")
+
+sperm.entry.stage<-new.meiotic.table[order.meiotic.table.phy,]$V3
+sperm.entry.stage[sperm.entry.stage ==" GV (-MI)"]<-" GV-MI"
+sperm.entry.stage[sperm.entry.stage ==" Pron. "]<- " Pron."
[…]

### 2015-01-18 writeup_and_notes/evolution/RESUBMISSIONmale_dependent_drive.tex
added to acknowledgements

@@
 \usepackage{psfrag}
+\usepackage{hyperref}
 \usepackage{mathabx}
@@
 \section*{Acknowledgements}
-We would like to thank members of the Coop lab for comments. This work was made possible by the NSF Postdoctoral Fellowship in Biology FY 2010 project number 1002942 awarded to Y.B..
+We would like to thank members of the Coop lab for comments, and Jon
+Wilkins, Andrew Gardner, and an anonymous reviewer for helpful
+comments on an earlier version of the paper. This work
+was made possible by the NSF Postdoctoral Fellowship in Biology FY
+2010 project number 1002942 awarded to Y.B. and by grants from the
+National Science Foundation under Grant No. 1262645 and by the
+National Institute of General Medical Sciences of the National
+Institutes of Health under award numbers NIH RO1GM83098 and
+RO1GM107374 to GC. We thank Karen Cranston, Fran\c{c}ois Michonneau,
+and David Winter for assistance with the R open tree of life package
+used to produce Figure \ref{Meiotic_fig}. % ç
+
 \newpage 
@@
   1 of \citet{Masui_book}. The tree was extracted from the Open Tree of
-Life project. The raw data table, the phylogeny/supporting R objects, and the script to do this is are included in
+Life project (\url{https://github.com/fmichonneau/rotl}). The raw data table, the phylogeny/supporting R objects, and the script to do this is are included in
 the supplement (Files S4-S6). }  

### 2015-01-18 scripts/meiotic_pause_phylogeny.R

@@ -5,7 +5,8 @@
 #install_github("fmichonneau/rotl")
 
 
-library("rotl")
+library("
+")
 library("rncl")
 library("ape")
 
